### 2019-12-18 submission/manuscript.Rmd
Inital optimization of the process.

---
title: "**Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces**"
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: header.tex
fontsize: 12pt
geometry: margin=1.0in
csl: citation_style.csl #Get themes at https://github.com/citation-style-language/styles
bibliography: references.bib
---

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE}
opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)
opts_chunk$set("message" = FALSE)

inline_hook <- function(x){
	print(x)

	if(is.list(x)){
		x <- unlist(x)
	}

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=1, nsmall=1, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
knitr::knit_hooks$set(inline=inline_hook)
```

\vspace{80mm}

^1$\dagger$^

\vspace{40mm}

$\dagger$ To whom correspondence should be addressed: marino.korlevic@irb.hr

1\. Ruđer Bošković Institute, Center for Marine Research, G. Paliaga 5, Rovinj, Croatia

2\. University of Vienna, Department of Limnology and Bio-Oceanography, Althanstraße 14, Vienna, Austria
\newpage
\linenumbers
\sisetup{mode=text}
\setlength\parindent{24pt}

## Abstract

\newpage
## Introduction

\newpage
## Materials and Methods
### Sampling
Leaves of *Cymodocea nodosa* were sampled in a *Cymodocea nodosa* meadow in the Bay of Saline (45°7´5˝N, 13°37´20˝E) and in a *Cymodocea nodosa* meadow invaded by *Caulerpa cylindracea* in the proximity of the village of Funtana (45°10´39˝N, 13°35´42˝E). Thalli of *Caulerpa cylindracea* were sampled in the same *Cymodocea nodosa* invaded meadow in Funtana and on a locality of only *Caulerpa cylindracea* located in the proximity of the invaded meadow. Leaves and thalli were collected on the same day in two contrasting seasons, on 4 December 2017 and 18 June 2018. During spring 2018 the *Cymodocea nodosa* meadow in the Bay of Saline decayed to an extent that no leaves could be retrieved (Najdek *et al.*, unpublished data). Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with site seawater. Upon arrival to the laboratory, *Cymodocea nodosa* leaves were cut into sections of 1 -- 2 \si{\cm}, while *Caulerpa cylindracea* thalli were cut into 5 -- 8 \si{\cm} long sections. Leaves and thalli were washed three times with sterile artificial seawater (ASW) to remove loosely attached microbial cells.

### DNA Isolation
The DNA was isolated according to the protocol for isolation from filters described in @Massana1997. This protocol was modified and adapted for DNA isolation from microbial communities from macrophyte surfaces as described below. 5 \si{\ml} of lysis buffer (40 \si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar} sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or \si{\g} wet-weight of thalli. Lysozyme was added (final concentration 1 \si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius} for 30 \si{\minute}. Subsequently, proteinase K (final concentration 0.5 \si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were added and the samples were incubated at 55 \si{\degreeCelsius} for 2 \si{\hour}. Following the incubation, tubes were vortexed for 10 \si{\minute} and the mixture containing lyzed epiphytic cells was separated from host leaves or thalli by transferring the solution into a clean tube. The lysate was extracted twice with a mixture of phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with chloroform:isoamyl alcohol (24:1). After each organic solvent mixture addition tubes were slightly vortexed and centrifuged at 4,500 × g for 10 \si{\minute}. Following each centrifugation aqueous phases were retrieved. After the final extraction 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added. DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} °C and centrifuging at 16,000 × g and 4 \si{\degreeCelsius} for 20 \si{\minute}. The pellet was washed twice with 1 \si{\ml} of chilled isopropanol and centrifuged after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 10 \si{\minute}. After the first washing step duplicate pellets form the same sample were pooled and transferred to a clean 1.5 \si{\ml} tube. The dried pellet was resuspended in 100 \si{\ul} of deionized water.

### Illumina 16S rRNA Sequencing
An aliquot of isolated DNA was treated with RNase A (final concentration 200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA concentration was determined using the Quant-iT PicoGreen dsDNA Assay Kit (Invitrogen, USA) according to the manufacturer’s instructions and diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure. In the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used to amplify the target region [@Caporaso2012; @Apprill2015; @Parada2016]. These primers contained on their 5' end a tagged sequence. Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained: 1 × Q5 Reaction Buffer , 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template. Cycling conditions were: initial denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50 \si{\degreeCelsius} for 60 \si{\s} and elongation at 72 \si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at 72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions volumes were pooled and PCR products were purified using the GeneJET PCR Purification Kit (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and following the protocol that included isopropanol addition for better small DNA fragment yield. The column was eluted in 30 \si{\ul} of deionized water. Purified PCR products were sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories, Martinsried, Germany. Before sequencing, at IMGM the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR. In addition, these primers contained adapter and sample-specific index sequences. The second PCR was carried out for 8 cycles. Beside samples, a positive and negative control were sequenced. A negative control was comprised of four parallel PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used. The sequences obtained in this study have been submitted to the European Nucleotide Archive (ENA) under accession numbers **TO BE ADDED LATER!**.

### Sequence Analysis
```{r}
# Calculation of the error rate based on the ATCC MSA-1002 mock community
error <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.error.summary", na="") %>%
  filter(numparents==1) %>%
  mutate(mismatches_weight=weight*mismatches) %>%
  mutate(total_weight=weight*total)
error <- sum(error$mismatches_weight)/sum(error$total_weight)*100
# Calculation of the number of sequences in the negative control
nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary") %>%
  filter(taxon=="Root") %>%
  select(NC_1)
```
Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/). For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 132; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014]. Sequences classified as chloroplasts by mothur were exported, aligned using SILVA incremental aligner (SINA, version 1.6.0) [@Pruesse2012] against the same SILVA SSU Ref NR 99 database (release 132) and imported into ARB (version 6.0.6) [@Ludwig2004] for further phylogenetic analysis using the same database. Reference sequences close to imported ones were selected and used to calculate a phylogenetic tree using the Maximum Likelihood algorithm RAxML (version 7.0.3) with 1000 tree replicates [@Stamatakis2006]. Imported partial chloroplast sequences were added to the tree using the maximum parsimony criteria and not allowing changes to tree topology. Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019], package tidyverse (version 1.2.1) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Zhu2019; @Allaire2019; @Xie2018]. The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (**TO BE ADDED LATER!**). Based on the ATCC MSA-1002 mock community included in the analysis a sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016]. In addition, the negative control processed together with the samples yielded only `r nc` sequences after sequence quality curation.

### Protein Isolation
Proteins were isolated according to the protocol for isolation from soil described in @Chourey2010 and modified by @Hultman2015. These protocols were further modified and adapted for protein isolation from microbial communities form macrophyte surfaces as described below. 20 \si{\ml} of protein extraction buffer (4 \si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl [pH 8.0]) was added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of thalli. The mixture was incubated in boiling water for 5 \si{\minute}, vortexed for 10 \si{\minute} and incubated again in boiling water for 5 \si{\minute}. After a brief vortex the lysate was transferred to a clean tube separating the host leaves or thalli from the mixture containing lyzed epiphytic cells. Dithiothreitol (DTT; final concentration 24 \si{\milli\Molar}) was added and proteins were precipitated with chilled 100 \si{\percent} trichloroacetic acid (TCA; final concentration 20 \si{\percent}) overnight at \num{-20} °C. Precipitated proteins were centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40 \si{\minute}. The obtained protein pellet was washed three times with chilled acetone. During the first washing step the pellet was transferred to a clean 1.5 \si{\ml} tube. After each washing step samples were centrifuged at 20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets were stored at \num{-80} °C until further analysis.

### Metaproteomics
Isolated proteins were whole trypsin digested using the FASP (filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK) according to the manufacturer's instructions [@Wisniewski2009] with small modifications. Before loading the solution to the column, protein pellets were solubilized in a urea sample buffer included in the kit amended with DTT (final concentration 100 \si{\milli\Molar}) for 45 \si{\minute} at room temperature and centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to remove larger particles. The first washing step after protein solution loading was repeated twice. In addition, centrifugation steps were prolonged if the column was clogged. Trypsin digestion was performed on column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}. The final filtrate containing digested proteins was acidified with 1 \si{\percent} (final concentration) trifluoroacetic acid, freezed at \num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent to the Vienna Metabolomics Center (University of Vienna) for metaproteomic analysis. Peptides were resuspended in 1 \si{\percent} (final concentration) trifluoroacetic acid, desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA). Obtained MS/MS spectra were searched against a protein database from metagenomic assembly published in @Burke2011a using SEQUEST-HIT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA). The target-decoy approach was used to reduce the probability of false peptide identification. Results whose false discovery rate at the peptide level was \SI{< 1}{\percent} were kept. For protein identification a minimum of of two peptides and one unique peptide were required. For protein quantification, a chromatographic peak area-based free quantitative method was applied.

### Confocal Microscopy
Host leaves and thalli from DNA and protein isolation steps were washed seven times in deionized water and fixed with formaldehyde (final concentration ~ 3 \si{\percent}). In addition, nontreated leaves and thalli, washed three times in ASW to remove loosely attached microbial cells, were fixed in the same concentration of formaldehyde and used as a positive control. For long therm storage, fixed leaves and thalli were immersed in a mixture of phosphate-buffered saline (PBS) and ethanol (1:1) and stored at \num{-20} \si{\degreeCelsius}. Treated and untreated leaves and thalli segments were stained in a 2 × solution of SYBR Green I and examined under a Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).

## Results

## Discussion

## Acknowledgements

\newpage
## References
<div id="refs"></div>

\newpage 
\setlength\parindent{0pt}

## Figure Captions
**\autoref{community}.** \nameref{community}

\newpage
## Figures
\newpage
```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant bacterial sequences.\\label{community}", fig.pos="ht"}
include_graphics("../results/figures/community_barplot_phylum.jpg")
```

### 2019-12-18 submission/manuscript.tex
Inital optimization of the process.

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{\textbf{Dynamics of Epiphytic Microbial Communities on Marine Macrophyte
Surfaces}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{times} % Times New Roman font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}

\usepackage{indentfirst}

\usepackage[labelsep=period, labelfont=bf]{caption}
\renewcommand{\thefigure}{\arabic{figure}}
\captionsetup{justification=raggedright,singlelinecheck=false}

\usepackage{pdflscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

\usepackage{siunitx}
\DeclareSIUnit\molar{\mole\per\cubic\deci\metre}
\DeclareSIUnit\Molar{\textsc{m}}

\usepackage{caption}
\captionsetup{justification=justified}

\begin{document}
\maketitle

\vspace{80mm}

\textsuperscript{1\(\dagger\)}

\vspace{40mm}

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:marino.korlevic@irb.hr}{\nolinkurl{marino.korlevic@irb.hr}}

1. Ruđer Bošković Institute, Center for Marine Research, G. Paliaga 5,
Rovinj, Croatia

2. University of Vienna, Department of Limnology and Bio-Oceanography,
Althanstraße 14, Vienna, Austria \newpage \linenumbers
\sisetup{mode=text} \setlength\parindent{24pt}

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

\newpage

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

\newpage

\hypertarget{materials-and-methods}{%
\subsection{Materials and Methods}\label{materials-and-methods}}

\hypertarget{sampling}{%
\subsubsection{Sampling}\label{sampling}}

Leaves of \emph{Cymodocea nodosa} were sampled in a \emph{Cymodocea
nodosa} meadow in the Bay of Saline (45°7´5˝N, 13°37´20˝E) and in a
\emph{Cymodocea nodosa} meadow invaded by \emph{Caulerpa cylindracea} in
the proximity of the village of Funtana (45°10´39˝N, 13°35´42˝E). Thalli
of \emph{Caulerpa cylindracea} were sampled in the same \emph{Cymodocea
nodosa} invaded meadow in Funtana and on a locality of only
\emph{Caulerpa cylindracea} located in the proximity of the invaded
meadow. Leaves and thalli were collected on the same day in two
contrasting seasons, on 4 December 2017 and 18 June 2018. During spring
2018 the \emph{Cymodocea nodosa} meadow in the Bay of Saline decayed to
an extent that no leaves could be retrieved (Najdek \emph{et al.},
unpublished data). Leaves and thalli were collected by diving and
transported to the laboratory in containers placed on ice and filled
with site seawater. Upon arrival to the laboratory, \emph{Cymodocea
nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while
\emph{Caulerpa cylindracea} thalli were cut into 5 -- 8 \si{\cm} long
sections. Leaves and thalli were washed three times with sterile
artificial seawater (ASW) to remove loosely attached microbial cells.

\hypertarget{dna-isolation}{%
\subsubsection{DNA Isolation}\label{dna-isolation}}

The DNA was isolated according to the protocol for isolation from
filters described in Massana \emph{et al.} (1997). This protocol was
modified and adapted for DNA isolation from microbial communities from
macrophyte surfaces as described below. 5 \si{\ml} of lysis buffer (40
\si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar}
sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or \si{\g}
wet-weight of thalli. Lysozyme was added (final concentration 1
\si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius}
for 30 \si{\minute}. Subsequently, proteinase K (final concentration 0.5
\si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were
added and the samples were incubated at 55 \si{\degreeCelsius} for 2
\si{\hour}. Following the incubation, tubes were vortexed for 10
\si{\minute} and the mixture containing lyzed epiphytic cells was
separated from host leaves or thalli by transferring the solution into a
clean tube. The lysate was extracted twice with a mixture of
phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with
chloroform:isoamyl alcohol (24:1). After each organic solvent mixture
addition tubes were slightly vortexed and centrifuged at 4,500 × g for
10 \si{\minute}. Following each centrifugation aqueous phases were
retrieved. After the final extraction 1/10 of chilled 3 \si{\Molar}
sodium acetate (pH 5.2) was added. DNA was precipitated by adding 1
volume of chilled isopropanol, incubating the mixtures overnight at
\num{-20} °C and centrifuging at 16,000 × g and 4 \si{\degreeCelsius}
for 20 \si{\minute}. The pellet was washed twice with 1 \si{\ml} of
chilled isopropanol and centrifuged after each washing step at 20,000 ×
g and 4 \si{\degreeCelsius} for 10 \si{\minute}. After the first washing
step duplicate pellets form the same sample were pooled and transferred
to a clean 1.5 \si{\ml} tube. The dried pellet was resuspended in 100
\si{\ul} of deionized water.

\hypertarget{illumina-16s-rrna-sequencing}{%
\subsubsection{Illumina 16S rRNA
Sequencing}\label{illumina-16s-rrna-sequencing}}

An aliquot of isolated DNA was treated with RNase A (final concentration
200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA
concentration was determined using the Quant-iT PicoGreen dsDNA Assay
Kit (Invitrogen, USA) according to the manufacturer's instructions and
diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was
amplified using a two-step PCR procedure. In the first PCR the 515F
(5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3')
primers from the Earth Microbiome Project
(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
were used to amplify the target region (Caporaso \emph{et al.}, 2012;
Apprill \emph{et al.}, 2015; Parada \emph{et al.}, 2016). These primers
contained on their 5' end a tagged sequence. Each sample was amplified
in four parallel 25 \si{\ul} reactions of which each contained: 1 × Q5
Reaction Buffer , 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml}
BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse
primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs,
USA) and 5 \si{\ng} of DNA template. Cycling conditions were: initial
denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of
denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50
\si{\degreeCelsius} for 60 \si{\s} and elongation at 72
\si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at
72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions
volumes were pooled and PCR products were purified using the GeneJET PCR
Purification Kit (Thermo Fisher Scientific, USA) according to the
manufacturer's instructions and following the protocol that included
isopropanol addition for better small DNA fragment yield. The column was
eluted in 30 \si{\ul} of deionized water. Purified PCR products were
sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories,
Martinsried, Germany. Before sequencing, at IMGM the second PCR
amplification of the two-step PCR procedure was performed using primers
targeting the tagged region incorporated in the first PCR. In addition,
these primers contained adapter and sample-specific index sequences. The
second PCR was carried out for 8 cycles. Beside samples, a positive and
negative control were sequenced. A negative control was comprised of
four parallel PCR reactions without DNA template, while for a positive
control a mock community composed of evenly mixed DNA material
originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
used. The sequences obtained in this study have been submitted to the
European Nucleotide Archive (ENA) under accession numbers \textbf{TO BE
ADDED LATER!}.

\hypertarget{sequence-analysis}{%
\subsubsection{Sequence Analysis}\label{sequence-analysis}}

Obtained sequences were analyzed on the computer cluster Isabella
(University Computing Center, University of Zagreb) using mothur
(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
Standard Operating Procedure (MiSeq SOP;
\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
and recommendations given from the Riffomonas project to enhance data
reproducibility (\url{http://www.riffomonas.org/}). For alignment and
classification of sequences the SILVA SSU Ref NR 99 database (release
132; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
mothur were exported, aligned using SILVA incremental aligner (SINA,
version 1.6.0) (Pruesse \emph{et al.}, 2012) against the same SILVA SSU
Ref NR 99 database (release 132) and imported into ARB (version 6.0.6)
(Ludwig \emph{et al.}, 2004) for further phylogenetic analysis using the
same database. Reference sequences close to imported ones were selected
and used to calculate a phylogenetic tree using the Maximum Likelihood
algorithm RAxML (version 7.0.3) with 1000 tree replicates (Stamatakis,
2006). Imported partial chloroplast sequences were added to the tree
using the maximum parsimony criteria and not allowing changes to tree
topology. Pipeline data processing and visualization was done using R
(version 3.6.0) (R Core Team, 2019), package tidyverse (version 1.2.1)
(Wickham \emph{et al.}, 2019) and multiple other packages (Xie, 2014,
2015, 2019a, 2019b; Xie \emph{et al.}, 2018; Zhu, 2019; Allaire \emph{et
al.}, 2019). The detailed analysis procedure including the R Markdown
file for this paper are available as a GitHub repository (\textbf{TO BE
ADDED LATER!}). Based on the ATCC MSA-1002 mock community included in
the analysis a sequencing error rate of 0.01 \si{\percent} was
determined, which is in line with previously reported values for
next-generation sequencing data (Kozich \emph{et al.}, 2013; Schloss
\emph{et al.}, 2016). In addition, the negative control processed
together with the samples yielded only 2 sequences after sequence
quality curation.

\hypertarget{protein-isolation}{%
\subsubsection{Protein Isolation}\label{protein-isolation}}

Proteins were isolated according to the protocol for isolation from soil
described in Chourey \emph{et al.} (2010) and modified by Hultman
\emph{et al.} (2015). These protocols were further modified and adapted
for protein isolation from microbial communities form macrophyte
surfaces as described below. 20 \si{\ml} of protein extraction buffer (4
\si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl {[}pH 8.0{]}) was
added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of
thalli. The mixture was incubated in boiling water for 5 \si{\minute},
vortexed for 10 \si{\minute} and incubated again in boiling water for 5
\si{\minute}. After a brief vortex the lysate was transferred to a clean
tube separating the host leaves or thalli from the mixture containing
lyzed epiphytic cells. Dithiothreitol (DTT; final concentration 24
\si{\milli\Molar}) was added and proteins were precipitated with chilled
100 \si{\percent} trichloroacetic acid (TCA; final concentration 20
\si{\percent}) overnight at \num{-20} °C. Precipitated proteins were
centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40 \si{\minute}.
The obtained protein pellet was washed three times with chilled acetone.
During the first washing step the pellet was transferred to a clean 1.5
\si{\ml} tube. After each washing step samples were centrifuged at
20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets
were stored at \num{-80} °C until further analysis.

\hypertarget{metaproteomics}{%
\subsubsection{Metaproteomics}\label{metaproteomics}}

Isolated proteins were whole trypsin digested using the FASP
(filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK)
according to the manufacturer's instructions (Wiśniewski \emph{et al.},
2009) with small modifications. Before loading the solution to the
column, protein pellets were solubilized in a urea sample buffer
included in the kit amended with DTT (final concentration 100
\si{\milli\Molar}) for 45 \si{\minute} at room temperature and
centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to
remove larger particles. The first washing step after protein solution
loading was repeated twice. In addition, centrifugation steps were
prolonged if the column was clogged. Trypsin digestion was performed on
column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
The final filtrate containing digested proteins was acidified with 1
\si{\percent} (final concentration) trifluoroacetic acid, freezed at
\num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent
to the Vienna Metabolomics Center (University of Vienna) for
metaproteomic analysis. Peptides were resuspended in 1 \si{\percent}
(final concentration) trifluoroacetic acid, desalted using the Pierce
C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's
instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap
Mass Spectrometer (Thermo Fisher Scientific, USA). Obtained MS/MS
spectra were searched against a protein database from metagenomic
assembly published in Burke \emph{et al.} (2011) using SEQUEST-HIT
engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo
Fisher Scientific, USA). The target-decoy approach was used to reduce
the probability of false peptide identification. Results whose false
discovery rate at the peptide level was \SI{< 1}{\percent} were kept.
For protein identification a minimum of of two peptides and one unique
peptide were required. For protein quantification, a chromatographic
peak area-based free quantitative method was applied.

\hypertarget{confocal-microscopy}{%
\subsubsection{Confocal Microscopy}\label{confocal-microscopy}}

Host leaves and thalli from DNA and protein isolation steps were washed
seven times in deionized water and fixed with formaldehyde (final
concentration \textasciitilde{} 3 \si{\percent}). In addition,
nontreated leaves and thalli, washed three times in ASW to remove
loosely attached microbial cells, were fixed in the same concentration
of formaldehyde and used as a positive control. For long therm storage,
fixed leaves and thalli were immersed in a mixture of phosphate-buffered
saline (PBS) and ethanol (1:1) and stored at \num{-20}
\si{\degreeCelsius}. Treated and untreated leaves and thalli segments
were stained in a 2 × solution of SYBR Green I and examined under a
Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).

\hypertarget{results}{%
\subsection{Results}\label{results}}

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Allaire2019}{}%
Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
A., et al. (2019) rmarkdown: Dynamic Documents for R.

\leavevmode\hypertarget{ref-Apprill2015}{}%
Apprill, A., McNally, S., Parsons, R., and Weber, L. (2015) Minor
revision to V4 region SSU rRNA 806R gene primer greatly increases
detection of SAR11 bacterioplankton. \emph{Aquatic Microbial Ecology}
\textbf{75}: 129--137.

\leavevmode\hypertarget{ref-Burke2011a}{}%
Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
(2011) Bacterial community assembly based on functional genes rather
than species. \emph{Proceedings of the National Academy of Sciences of
the United States of America} \textbf{108}: 14288--14293.

\leavevmode\hypertarget{ref-Caporaso2012}{}%
Caporaso, J.G., Lauber, C.L., Walters, W.A., Berg-Lyons, D., Huntley,
J., Fierer, N., et al. (2012) Ultra-high-throughput microbial community
analysis on the Illumina HiSeq and MiSeq platforms. \emph{The ISME
Journal} \textbf{6}: 1621--1624.

\leavevmode\hypertarget{ref-Chourey2010}{}%
Chourey, K., Jansson, J., VerBerkmoes, N., Shah, M., Chavarria, K.L.,
Tom, L.M., et al. (2010) Direct Cellular Lysis/Protein Extraction
Protocol for Soil Metaproteomics. \emph{Journal of Proteome Research}
\textbf{9}: 6615--6622.

\leavevmode\hypertarget{ref-Hultman2015}{}%
Hultman, J., Waldrop, M.P., Mackelprang, R., David, M.M., McFarland, J.,
Blazewicz, S.J., et al. (2015) Multi-omics of permafrost, active layer
and thermokarst bog soil microbiomes. \emph{Nature} \textbf{521}:
208--212.

\leavevmode\hypertarget{ref-Kozich2013}{}%
Kozich, J.J., Westcott, S.L., Baxter, N.T., Highlander, S.K., and
Schloss, P.D. (2013) Development of a dual-index sequencing strategy and
curation pipeline for analyzing amplicon sequence data on the MiSeq
Illumina sequencing platform. \emph{Applied and environmental
microbiology} \textbf{79}: 5112--5120.

\leavevmode\hypertarget{ref-Ludwig2004}{}%
Ludwig, W., Strunk, O., Westram, R., Richter, L., Meier, H., Yadhukumar,
et al. (2004) ARB: a software environment for sequence data.
\emph{Nucleic Acids Research} \textbf{32}: 1363--1371.

\leavevmode\hypertarget{ref-Massana1997}{}%
Massana, R., Murray, A.E., Preston, C.M., and DeLong, E.F. (1997)
Vertical distribution and phylogenetic characterization of marine
planktonic Archaea in the Santa Barbara Channel. \emph{Applied and
environmental microbiology} \textbf{63}: 50--56.

\leavevmode\hypertarget{ref-Parada2016}{}%
Parada, A.E., Needham, D.M., and Fuhrman, J.A. (2016) Every base
matters: assessing small subunit rRNA primers for marine microbiomes
with mock communities, time series and global field samples.
\emph{Environmental Microbiology} \textbf{18}: 1403--1414.

\leavevmode\hypertarget{ref-Pruesse2012}{}%
Pruesse, E., Peplies, J., and Glöckner, F.O. (2012) SINA: accurate
high-throughput multiple sequence alignment of ribosomal RNA genes.
\emph{Bioinformatics (Oxford, England)} \textbf{28}: 1823--1829.

\leavevmode\hypertarget{ref-Quast2013}{}%
Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
et al. (2013) The SILVA ribosomal RNA gene database project: improved
data processing and web-based tools. \emph{Nucleic acids research}
\textbf{41}: D590--6.

\leavevmode\hypertarget{ref-RCoreTeam2019}{}%
R Core Team (2019) R: A Language and Environment for Statistical
Computing, Vienna, Austria: R Foundation for Statistical Computing.

\leavevmode\hypertarget{ref-Schloss2016}{}%
Schloss, P.D., Girard, R.A., Martin, T., Edwards, J., and Thrash, J.C.
(2016) Status of the Archaeal and Bacterial Census: an Update.
\emph{mBio} \textbf{7}: e00201--16.

\leavevmode\hypertarget{ref-Schloss2009}{}%
Schloss, P.D., Westcott, S.L., Ryabin, T., Hall, J.R., Hartmann, M.,
Hollister, E.B., et al. (2009) Introducing mothur: open-source,
platform-independent, community-supported software for describing and
comparing microbial communities. \emph{Applied and environmental
microbiology} \textbf{75}: 7537--7541.

\leavevmode\hypertarget{ref-Stamatakis2006}{}%
Stamatakis, A. (2006) RAxML-VI-HPC: maximum likelihood-based
phylogenetic analyses with thousands of taxa and mixed models.
\emph{Bioinformatics} \textbf{22}: 2688--2690.

\leavevmode\hypertarget{ref-Wickham2019}{}%
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.D., François,
R., et al. (2019) Welcome to the tidyverse. \emph{Journal of Open Source
Software} \textbf{4}: 1686.

\leavevmode\hypertarget{ref-Wisniewski2009}{}%
Wiśniewski, J.R., Zougman, A., Nagaraj, N., and Mann, M. (2009)
Universal sample preparation method for proteome analysis. \emph{Nature
Methods} \textbf{6}: 359--362.

\leavevmode\hypertarget{ref-Xie2015}{}%
Xie, Y. (2015) Dynamic Documents with \{R\} and knitr, 2nd ed. Boca
Raton, Florida: Chapman; Hall/CRC.

\leavevmode\hypertarget{ref-Xie2014}{}%
Xie, Y. (2014) knitr: A Comprehensive Tool for Reproducible Research in
\{R\}. In, Stodden,V., Leisch,F., and Peng,R.D. (eds),
\emph{Implementing reproducible computational research}. Chapman;
Hall/CRC.

\leavevmode\hypertarget{ref-Xie2019a}{}%
Xie, Y. (2019a) knitr: A General-Purpose Package for Dynamic Report
Generation in R.

\leavevmode\hypertarget{ref-Xie2019}{}%
Xie, Y. (2019b) tinytex: Helper Functions to Install and Maintain 'TeX
Live', and Compile 'LaTeX' Documents.

\leavevmode\hypertarget{ref-Xie2018}{}%
Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: The
Definitive Guide, Boca Raton, Florida: Chapman; Hall/CRC.

\leavevmode\hypertarget{ref-Yilmaz2014}{}%
Yilmaz, P., Parfrey, L.W., Yarza, P., Gerken, J., Pruesse, E., Quast,
C., et al. (2014) The SILVA and "All-species Living Tree Project (LTP)"
taxonomic frameworks. \emph{Nucleic acids research} \textbf{42}:
D643--8.

\leavevmode\hypertarget{ref-Zhu2019}{}%
Zhu, H. (2019) kableExtra: Construct Complex Table with 'kable' and Pipe
Syntax.

\newpage 
\setlength\parindent{0pt}

\hypertarget{figure-captions}{%
\subsection{Figure Captions}\label{figure-captions}}

\textbf{\autoref{community}.} \nameref{community}

\newpage

\hypertarget{figures}{%
\subsection{Figures}\label{figures}}

\newpage
\begin{figure}[ht]

{\centering \includegraphics[width=1\linewidth]{../results/figures/community_barplot_phylum}

}

\caption{Taxonomic classification and relative contribution of the most abundant bacterial sequences.\label{community}}\label{fig:unnamed-chunk-2}
\end{figure}

\end{document}

### 2019-12-18 code/format_taxonomy.R

@@ -1,0 +1,78 @@
+map.in <- read.table("data/references/tax_slv_ssu_132.txt", header=F,sep="\t", stringsAsFactors=F)
+map.in <- map.in[, c(1,3)]
+colnames(map.in) <- c("taxlabel", "taxlevel")
+taxlevels <- c("root", "domain", "major_clade", "superkingdom", "kingdom", "subkingdom", "infrakingdom", "superphylum", "phylum", "subphylum", "infraphylum", "superclass", "class", "subclass", "infraclass", "superorder", "order", "suborder", "superfamily", "family", "subfamily", "genus")
+taxabb <- c("ro", "do", "mc", "pk", "ki", "bk", "ik", "pp", "ph", "bp", "ip", "pc", "cl", "bc", "ic", "po", "or", "bo", "pf", "fa", "bf", "ge")
+tax.mat <- matrix(data="", nrow=nrow(map.in), ncol=length(taxlevels))
+tax.mat[,1] <- "root"
+colnames(tax.mat) <- taxlevels
+
+outlevels <- c("domain", "phylum", "class", "order", "family", "genus")
+
+for(i in 1:nrow(map.in)) {
+  taxname <- unlist(strsplit(as.character(map.in[i,1]), split=';'))
+  # Print(taxname);
+  
+  while ( length(taxname) > 0) {
+    # Regex to look for exact match
+    
+    tax.exp <- paste(paste(taxname,collapse=";"), ";", sep="")
+    tax.match <- match(tax.exp,map.in$taxlabel)
+    tax.mat[i, map.in[tax.match,2]] <- tail(taxname, 1)
+    taxname <- head(taxname, -1)
+  }
+}
+
+for(i in 1:nrow(tax.mat)) {
+  # This fills in the empty gaps by using the closest higher taxonomic level appended with an abbreviation for the current taxonomic level
+  # if you don't want this behavior, cut it out
+  for(j in 1:ncol(tax.mat)) {
+    if(tax.mat[i, j] < 0) { tax.mat[i, j] <- paste(tmptax, taxabb[j], sep="_")}
+    else { tmptax <- tax.mat[i, j]}
+  }
+  
+  # This maps the new name to the input taxonomic levels
+  map.in[i, "taxout"] <- paste(paste(tax.mat[i, outlevels], collapse=";"), ";", sep="")
+}
+
+# Replace spaces with underscores
[…]

### 2019-12-18 code/plot_calculators.R

@@ -1,0 +1,55 @@
+#################################################################################################################
+# plot_calculators.R
+# 
+# A script to plot richness and diversity calculators.
+# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.groups.ave-std.summary
+# Produces: results/figures/calculators.jpg
+#           
+#################################################################################################################
+
+library(tidyverse)
+library(gridExtra)
+
+alpha <- read_tsv(file="data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.groups.ave-std.summary",
+                 col_types=cols(group=col_character())) %>%
+  filter(method=='ave')
+  
+metadata <- read_tsv("data/raw/metadata.csv")
+
+meta_alpha <- inner_join(metadata, alpha, by=c("ID"="group"))
+
+plot <- meta_alpha %>%
+    filter(!is.na(station)) %>%
+    group_by(station) %>%
+    summarize(mean_sobs=mean(sobs),
+              mean_chao=mean(chao),
+              mean_ace=mean(ace),
+              mean_shannon=mean(shannon),
+              mean_invsimpson=mean(invsimpson))
+
+sobs <- ggplot(plot, aes(x=station, y=mean_sobs)) +
+  geom_boxplot(notch=TRUE) +
+  labs(x="Depth", y="Number of OTUs") +
+  theme_classic()
+chao <- ggplot(plot, aes(x=station, y=mean_chao)) +
+  geom_boxplot(notch=TRUE) +
+  labs(x="Depth", y="Number of OTUs") +
+  theme_classic()
+ace <- ggplot(plot, aes(x=station, y=mean_ace)) +
+  geom_boxplot(notch=TRUE) +
+  labs(x="Depth", y="Number of OTUs") +
+  theme_classic()
+shannon <- ggplot(plot, aes(x=station, y=mean_shannon)) +
+  geom_boxplot(notch=TRUE) +
+  labs(x="Depth", y="Shannon Diversity Index") +
+  theme_classic()
+invsimpson <- ggplot(plot, aes(x=station, y=mean_invsimpson)) +
+  geom_boxplot(notch=TRUE) +
+  labs(x="Depth", y="Inverse Simpson Diversity Index") +
+  theme_classic()
+grid.arrange(sobs, chao, ace, shannon, invsimpson, nrow=2)
[…]

### 2019-12-18 code/plot_community_bar_plot.R

@@ -1,0 +1,62 @@
+#################################################################################################################
+# plot_community_barplot_taxlevel.R
+# 
+# A script to plot the community structure of each sample at every taxonomic level (Domain-Genus).
+# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+# Produces: results/figures/community_barplot_domain.jpg
+#           results/figures/community_barplot_phylum.jpg
+#           results/figures/community_barplot_class.jpg
+#           results/figures/community_barplot_order.jpg
+#           results/figures/community_barplot_family.jpg
+#           results/figures/community_barplot_genus.jpg
+#################################################################################################################
+
+library(tidyverse)
+
+community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary") %>%
+  filter(!str_detect(taxon, "^Eukaryota")) %>%
+  filter(taxon!="Root") %>%
+  filter(!str_detect(taxon, "^Mitochondria")) %>%
+  group_by(taxlevel) %>%
+  mutate_at(5:ncol(.), funs(. / sum(.) * 100)) %>%
+  ungroup()
+
+metadata <- read_tsv("data/raw/metadata.csv") %>%
+  filter(ID %in% colnames(community[, 6:ncol(community)])) %>%
+  select(ID, label) %>%
+  deframe()
+
+taxlevel <- c("domain", "phylum", "class", "order", "family", "genus")
+
+for (i in seq(1:6)) {
+plot <- filter(community, taxlevel==i) %>%
+  filter_at(6:ncol(.), any_vars(. >= 2)) %>%
+  mutate(taxon=str_replace_all(taxon, c("unknown_unclassified"="No Relative", "unknown"="No Relative"))) %>%
+  filter_at(6:ncol(.), any_vars(. >= 2)) %>%
+  bind_rows(summarise_all(., funs(ifelse(is.numeric(.), 100-sum(.), paste("Other"))))) %>%
+  rename_at(names(metadata), ~unname(metadata)) %>%
+  arrange(-row_number())
+
+color <- read_tsv("data/raw/group_colors.csv") %>%
+  deframe()
+
+gather(plot, key="sample", value="abundance", 6:(ncol(plot))) %>%
+  mutate(taxon=factor(taxon, levels=unique(plot$taxon))) %>%
[…]

### 2019-12-18 code/plot_pcoa.R

@@ -1,0 +1,27 @@
+#################################################################################################################
+# plot_pcoa.R
+# 
+# A script to generate the PCoA figure.
+# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.braycurtis.0.03.lt.ave.pcoa.axes
+#               data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.braycurtis.0.03.lt.ave.pcoa.loadings
+#               data/raw/metadata.csv
+# Produces: results/figures/pcoa_figure.jpg
+#
+#################################################################################################################
+
+library(tidyverse)
+pcoa <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.braycurtis.0.03.lt.ave.pcoa.axes", col_types=cols(group = col_character()))
+metadata <- read_tsv("data/raw/metadata.csv")
+metadata_pcoa <- inner_join(metadata, pcoa, by=c('ID'='group'))
+
+pcoa_loadings <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.braycurtis.0.03.lt.ave.pcoa.loadings")
+
+ggplot(metadata_pcoa, aes(x=axis1, y=axis2)) +
+  geom_point(shape=19, size=2) +
+  geom_text(aes(label=label), nudge_x=0.07) +
+  coord_fixed() + 
+  labs(x=paste("PCoA 1 (", format(round(pcoa_loadings$loading[1], 2), nsmall=2), " %)"),
+       y=paste("PCoA 2 (", format(round(pcoa_loadings$loading[2], 2), nsmall=2), " %)")) +
+  theme_classic()
+
+ggsave("results/figures/pcoa_figure.jpg", width=297, height=210, units="mm")

### 2020-01-28 code/plot_calculators.R
Temporal indices.

@@ -49,7 +49,7 @@ estimators_indices <- inner_join(estimators, shannon, by="Group") %>%
 # Loading metadata 
 metadata <- read_tsv("data/raw/metadata.csv")
 
-# Joining metada with estimators and indices
+# Joining metadata with estimators and indices
 Sys.setlocale(locale="en_GB.utf8")
 estimators_indices_metadata <- inner_join(metadata, estimators_indices, by=c("ID"="Group")) %>%
   mutate(date=as.Date(date, "%d.%m.%Y"))
@@ -69,13 +69,13 @@ estimators_indices_metadata <- estimators_indices_metadata %>%
   gather("S.obs", "S.chao1", "S.ACE", "eshannon", "invsimpson", key="estimator_index", value="value")
 
 # Defining line types, dot shapes and fill dot colors
-lines_p1 <- c(c("S.obs"="dotted", "S.chao1"="solid", "S.ACE"="dotted"))
+lines_p1 <- c("S.obs"="dotted", "S.chao1"="solid", "S.ACE"="dotted")
 shapes_p1 <- c("S.obs"=21, "S.chao1"=23, "S.ACE"=25)
-fill_p1 <- c("S.obs"="white", "S.chao1"="black", "S.ACE"="white")
+fills_p1 <- c("S.obs"="white", "S.chao1"="black", "S.ACE"="white")
 
 lines_p2 <- c(c("eshannon"="solid", "invsimpson"="dotted"))
 shapes_p2 <- c("eshannon"=21, "invsimpson"=24)
-fill_p2 <- c("eshannon"="black", "invsimpson"="white")
+fills_p2 <- c("eshannon"="black", "invsimpson"="white")
 
 # Plots generation
 # Seawater samples
@@ -87,7 +87,7 @@ p1 <- filter(estimators_indices_metadata, station=="F") %>%
   geom_point(size=3) +
   scale_linetype_manual(values=lines_p1) +
   scale_shape_manual(values=shapes_p1) +
-  scale_fill_manual(values=fill_p1) +
+  scale_fill_manual(values=fills_p1) +
   scale_y_continuous(limits=c(300, 3000)) +
   labs(x="", y="") +
   ggtitle(parse(text="bold('Seawater')")) +
@@ -102,7 +102,7 @@ p2 <- filter(estimators_indices_metadata, station=="F") %>%
   geom_point(size=3) +
   scale_linetype_manual(values=lines_p2) +
   scale_shape_manual(values=shapes_p2) +
[…]

### 2020-01-28 code/plot_community_bar_plot.R

@@ -3,6 +3,7 @@
 # 
 # A script to plot the community structure of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+#               data/raw/metadata.csv
 # Produces: results/figures/community_bar_plot.jpg
 #
 #################################################################################################################

### 2020-01-28 code/plot_community_bar_plot_alphaproteobacteria.R

@@ -3,6 +3,7 @@
 # 
 # A script to plot sequence abundances of groups from the class Alphaproteobacteria of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+#               data/raw/metadata.csv
 # Produces: results/figures/alphaproteobacteria_bar_plot.jpg
 #
 #################################################################################################################

### 2020-01-28 code/plot_community_bar_plot_bacteroidetes.R

@@ -3,6 +3,7 @@
 # 
 # A script to plot sequence abundances of groups from the phylum Bacteoidetes of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+#               data/raw/metadata.csv
 # Produces: results/figures/bacteroidetes_bar_plot.jpg
 #
 #################################################################################################################

### 2020-01-28 code/plot_community_bar_plot_chloroplast.R

@@ -3,6 +3,7 @@
 # 
 # A script to plot chloroplast sequence abundances of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+#               data/raw/metadata.csv
 # Produces: results/figures/chloroplast_bar_plot.jpg
 #
 #################################################################################################################

### 2020-01-28 code/plot_community_bar_plot_cyanobacteria.R

@@ -3,6 +3,7 @@
 # 
 # A script to plot sequence abundances of groups from the phylum Cyanobacteria of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+#               data/raw/metadata.csv
 # Produces: results/figures/cyanobacteria_bar_plot.jpg
 #
 #################################################################################################################

### 2020-01-28 code/plot_community_bar_plot_deltaproteobacteria.R

@@ -3,6 +3,7 @@
 # 
 # A script to plot sequence abundances of groups from the class Gammaproteobacteria of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+#               data/raw/metadata.csv
 # Produces: results/figures/deltaproteobacteria_bar_plot.jpg
 #
 #################################################################################################################

### 2020-01-28 code/plot_community_bar_plot_gammaproteobacteria.R

@@ -3,6 +3,7 @@
 # 
 # A script to plot sequence abundances of groups from the class Gammaproteobacteria of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+#               data/raw/metadata.csv
 # Produces: results/figures/gammaproteobacteria_bar_plot.jpg
 #
 #################################################################################################################

### 2020-01-28 code/plot_seasonal_shared.R
Changes temporal to seasonal name

@@ -1,0 +1,262 @@
+#################################################################################################################
+# code/plot_seasonal_shared.R
+# 
+# A script to plot the Bray-Curtis and Jaccard's Similarity Coefficients of subsequent sampling dates.
+# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared
+#               data/raw/metadata.csv
+# Produces: results/figures/temporal_shared.jpg
+#
+#################################################################################################################
+
+# Loading OTU/sample data
+shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")
+
+# Generation of random rarified community data
+rarefied <- shared %>%
+  select(-label, -Group, -numOtus) %>%
+  rrarefy(., min(rowSums(.))) %>%
+  as.tibble() %>%
+  add_column("Group"=shared$Group, .before=TRUE) %>%
+  select_if(funs(!is.numeric(.) || sum(.)!=0))
+
+# Copying the sample labels to the rows (input for library vegan)
+row.names(rarefied) <- rarefied$Group
+
+# Removing column containing sample labels
+rarefied <- rarefied %>%
+  select(-Group)
+
+# Calculating dissimilarity indices
+jaccard <- vegdist(rarefied, method="jaccard", binary=T)
+jaccard <- as.tibble(data.frame(t(combn(rownames(rarefied),2)), as.numeric(jaccard))) %>%
+  rename(V1=X1, V2=X2, jaccard=as.numeric.jaccard.)
+
+bray <- vegdist(rarefied, method="bray", binary=F)
+bray <- as.tibble(data.frame(t(combn(rownames(rarefied),2)), as.numeric(bray))) %>%
+  rename(V1=X1, V2=X2, bray=as.numeric.bray.)
+
+distance <- inner_join(jaccard, bray, by=c("V1"="V1", "V2"="V2"))
+
+# Loading metadata 
+metadata <- read_tsv("data/raw/metadata.csv")
+
+# Joining metadata with dissimilarity indices data
+Sys.setlocale(locale="en_GB.utf8")
[…]

### 2020-01-31 submission/manuscript.Rmd
Switch to Silva database v138.

@@
 # Calculation of the number of sequences in the negative control
-nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary") %>%
+nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
   filter(taxon=="Root") %>%
@@
 ```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant bacterial sequences.\\label{community}", fig.pos="ht"}
-include_graphics("../results/figures/community_barplot_phylum.jpg")
+include_graphics("../results/figures/community_bar_plot.jpg")
 ```

### 2020-01-31 code/format_taxonomy.R

@@ -1,4 +1,4 @@
-map.in <- read.table("data/references/tax_slv_ssu_132.txt", header=F,sep="\t", stringsAsFactors=F)
+map.in <- read.table("data/references/tax_slv_ssu_138.txt", header=F,sep="\t", stringsAsFactors=F)
 map.in <- map.in[, c(1,3)]
 colnames(map.in) <- c("taxlabel", "taxlevel")
 taxlevels <- c("root", "domain", "major_clade", "superkingdom", "kingdom", "subkingdom", "infrakingdom", "superphylum", "phylum", "subphylum", "infraphylum", "superclass", "class", "subclass", "infraclass", "superorder", "order", "suborder", "superfamily", "family", "subfamily", "genus")
@@ -39,19 +39,21 @@ for(i in 1:nrow(tax.mat)) {
 map.in$taxout <- gsub(" ", "_", map.in$taxout)
 
 # Bring in the old taxonomic levels from SILVA and remap them using the new levels
-tax.in <- read.table("data/references/silva.nr_v132.full", header=F, stringsAsFactors=F, sep="\t")
+tax.in <- read.table("data/references/silva.nr_v138.full", header=F, stringsAsFactors=F, sep="\t")
 colnames(tax.in) <- c("taxid", "taxlabel")
 
-# Problem corrections observed in the SILVA ARB file (silva.nr_v132.full)
+# Problem corrections observed in the SILVA ARB file (silva.nr_v138.full)
 # Following line corrects the Bacteria;Bacteroidetes;Bacteroidia;Flavobacteriales;Flavobacteriaceae;Polaribacter;Polaribacter; problem
 tax.in$taxlabel <- gsub("Polaribacter;Polaribacter", "Polaribacter", tax.in$taxlabel)
 # Following line corrects the Bacteria;RsaHf231; problem
-tax.in$taxlabel <- gsub("RsaHf231;", "RsaHF231;", tax.in$taxlabel)
+# tax.in$taxlabel <- gsub("RsaHf231;", "RsaHF231;", tax.in$taxlabel)
 # Following line corrects the Bacteria;GBS-1;; problem
-tax.in$taxlabel <- gsub("Bacteria;GBS-1;;", "Bacteria;GBS-1;", tax.in$taxlabel)
[…]

### 2020-01-31 code/plot_calculators.R

@@ -17,7 +17,7 @@ rarefied <- shared %>%
   rrarefy(., min(rowSums(.))) %>%
   as.tibble() %>%
   add_column("Group"=shared$Group, .before=TRUE) %>%
-  select_if(funs(!is.numeric(.) || sum(.)!=0))
+  select_if(list(~!is.numeric(.) || sum(.)!=0))
 
 # Copying the sample labels to the rows (input for library vegan)
 row.names(rarefied) <- rarefied$Group

### 2020-01-31 code/plot_community_bar_plot.R

@@ -2,14 +2,14 @@
 # plot_community_bar_plot.R
 # 
 # A script to plot the community structure of each sample.
-# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary
 #               data/raw/metadata.csv
 # Produces: results/figures/community_bar_plot.jpg
 #
 #################################################################################################################
 
 # Loading input data containing sequence abundances and subsequent input data customization
-community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary") %>%
+community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
   filter(!str_detect(taxon, "^Eukaryota")) %>%
   filter(taxon!="Root")
 
@@ -17,13 +17,13 @@ community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.uniq
 # they belong
 chloroplast <- filter(community, str_detect(taxon, "^Chloroplast$"))$rankID
 mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
-community <- mutate_at(community, 5:ncol(community), funs(case_when(
+community <- mutate_at(community, 5:ncol(community), list(~case_when(
     rankID==str_extract(chloroplast, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Chloroplast"],
     rankID==str_extract(chloroplast, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Chloroplast"],
     rankID==str_extract(chloroplast, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Chloroplast"],
   TRUE ~ .))) %>%
   filter(!str_detect(taxon, "^Chloroplast")) %>%
-  mutate_at(5:ncol(.), funs(case_when(
+  mutate_at(5:ncol(.), list(~case_when(
[…]

### 2020-01-31 code/plot_community_bar_plot_alphaproteobacteria.R

@@ -2,14 +2,14 @@
 # plot_community_bar_plot_alphaproteobacteria.R
 # 
 # A script to plot sequence abundances of groups from the class Alphaproteobacteria of each sample.
-# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary
 #               data/raw/metadata.csv
 # Produces: results/figures/alphaproteobacteria_bar_plot.jpg
 #
 #################################################################################################################
 
 # Loading input data containing sequence abundances and subsequent input data customization
-community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary") %>%
+community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
   filter(!str_detect(taxon, "^Eukaryota")) %>%
   filter(taxon!="Root")
 
@@ -17,13 +17,13 @@ community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.uniq
 # they belong
 chloroplast <- filter(community, str_detect(taxon, "^Chloroplast$"))$rankID
 mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
-community <- mutate_at(community, 5:ncol(community), funs(case_when(
+community <- mutate_at(community, 5:ncol(community), list(~case_when(
   rankID==str_extract(chloroplast, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Chloroplast"],
   rankID==str_extract(chloroplast, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Chloroplast"],
   rankID==str_extract(chloroplast, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Chloroplast"],
   TRUE ~ .))) %>%
   filter(!str_detect(taxon, "^Chloroplast")) %>%
-  mutate_at(5:ncol(.), funs(case_when(
+  mutate_at(5:ncol(.), list(~case_when(
[…]

### 2020-01-31 code/plot_community_bar_plot_bacteroidota.R

@@ -1,15 +1,15 @@
 #################################################################################################################
-# plot_community_bar_plot_bacteroidetes.R
+# plot_community_bar_plot_bacteroidota.R
 # 
 # A script to plot sequence abundances of groups from the phylum Bacteoidetes of each sample.
-# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary
 #               data/raw/metadata.csv
-# Produces: results/figures/bacteroidetes_bar_plot.jpg
+# Produces: results/figures/bacteroidota_bar_plot.jpg
 #
 #################################################################################################################
 
 # Loading input data containing sequence abundances and subsequent input data customization
-community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary") %>%
+community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
   filter(!str_detect(taxon, "^Eukaryota")) %>%
   filter(taxon!="Root")
 
@@ -17,13 +17,13 @@ community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.uniq
 # they belong
 chloroplast <- filter(community, str_detect(taxon, "^Chloroplast$"))$rankID
 mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
-community <- mutate_at(community, 5:ncol(community), funs(case_when(
+community <- mutate_at(community, 5:ncol(community), list(~case_when(
   rankID==str_extract(chloroplast, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Chloroplast"],
   rankID==str_extract(chloroplast, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Chloroplast"],
   rankID==str_extract(chloroplast, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Chloroplast"],
   TRUE ~ .))) %>%
[…]

### 2020-01-31 code/plot_community_bar_plot_chloroplast.R

@@ -2,21 +2,21 @@
 # plot_community_bar_plot_chloroplast.R
 # 
 # A script to plot chloroplast sequence abundances of each sample.
-# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary
 #               data/raw/metadata.csv
 # Produces: results/figures/chloroplast_bar_plot.jpg
 #
 #################################################################################################################
 
 # Loading input data containing sequence abundances and subsequent input data customization
-community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary") %>%
+community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
   filter(!str_detect(taxon, "^Eukaryota")) %>%
   filter(taxon!="Root")
 
 # Remove and Mitochondria sequences and subtract their number from higher taxonomic levels to which
 # they belong
 mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
-community <- mutate_at(community, 5:ncol(community), funs(case_when(
+community <- mutate_at(community, 5:ncol(community), list(~case_when(
     rankID==str_extract(mitochondria, "(\\d+\\.){4}\\d+") ~ . - .[taxon=="Mitochondria"],
     rankID==str_extract(mitochondria, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Mitochondria"],
     rankID==str_extract(mitochondria, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Mitochondria"],
@@ -27,16 +27,17 @@ community <- mutate_at(community, 5:ncol(community), funs(case_when(
   mutate(`23`=`23_1`+`23_2`) %>%
   select(-`23_1`, -`23_2`) %>%
   group_by(taxlevel) %>%
-  mutate_at(5:ncol(.), funs(. / sum(.) * 100)) %>%
+  mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
   ungroup()
 
 # Selection of groups for plotting
 plot <- community %>%
[…]

### 2020-01-31 code/plot_community_bar_plot_cyanobacteria.R

@@ -2,14 +2,14 @@
 # plot_community_bar_plot_cyanobacteria.R
 # 
 # A script to plot sequence abundances of groups from the phylum Cyanobacteria of each sample.
-# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary
 #               data/raw/metadata.csv
 # Produces: results/figures/cyanobacteria_bar_plot.jpg
 #
 #################################################################################################################
 
 # Loading input data containing sequence abundances and subsequent input data customization
-community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary") %>%
+community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
   filter(!str_detect(taxon, "^Eukaryota")) %>%
   filter(taxon!="Root")
 
@@ -17,13 +17,13 @@ community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.uniq
 # they belong
 chloroplast <- filter(community, str_detect(taxon, "^Chloroplast$"))$rankID
 mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
-community <- mutate_at(community, 5:ncol(community), funs(case_when(
+community <- mutate_at(community, 5:ncol(community), list(~case_when(
   rankID==str_extract(chloroplast, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Chloroplast"],
   rankID==str_extract(chloroplast, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Chloroplast"],
   rankID==str_extract(chloroplast, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Chloroplast"],
   TRUE ~ .))) %>%
   filter(!str_detect(taxon, "^Chloroplast")) %>%
-  mutate_at(5:ncol(.), funs(case_when(
+  mutate_at(5:ncol(.), list(~case_when(
     rankID==str_extract(mitochondria, "(\\d+\\.){4}\\d+") ~ . - .[taxon=="Mitochondria"],
[…]

### 2020-01-31 code/plot_community_bar_plot_desulfobacterota.R

@@ -1,15 +1,15 @@
 #################################################################################################################
-# plot_community_bar_plot_deltaroteobacteria.R
+# plot_community_bar_plot_desulfobacterota.R
 # 
 # A script to plot sequence abundances of groups from the class Gammaproteobacteria of each sample.
-# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary
 #               data/raw/metadata.csv
-# Produces: results/figures/deltaproteobacteria_bar_plot.jpg
+# Produces: results/figures/desulfobacterota_bar_plot.jpg
 #
 #################################################################################################################
 
 # Loading input data containing sequence abundances and subsequent input data customization
-community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary") %>%
+community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
   filter(!str_detect(taxon, "^Eukaryota")) %>%
   filter(taxon!="Root")
 
@@ -17,13 +17,13 @@ community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.uniq
 # they belong
 chloroplast <- filter(community, str_detect(taxon, "^Chloroplast$"))$rankID
 mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
-community <- mutate_at(community, 5:ncol(community), funs(case_when(
+community <- mutate_at(community, 5:ncol(community), list(~case_when(
   rankID==str_extract(chloroplast, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Chloroplast"],
   rankID==str_extract(chloroplast, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Chloroplast"],
   rankID==str_extract(chloroplast, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Chloroplast"],
   TRUE ~ .))) %>%
[…]

### 2020-01-31 code/plot_community_bar_plot_gammaproteobacteria.R

@@ -2,14 +2,14 @@
 # plot_community_bar_plot_gammaproteobacteria.R
 # 
 # A script to plot sequence abundances of groups from the class Gammaproteobacteria of each sample.
-# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary
 #               data/raw/metadata.csv
 # Produces: results/figures/gammaproteobacteria_bar_plot.jpg
 #
 #################################################################################################################
 
 # Loading input data containing sequence abundances and subsequent input data customization
-community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary") %>%
+community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
   filter(!str_detect(taxon, "^Eukaryota")) %>%
   filter(taxon!="Root")
 
@@ -17,13 +17,13 @@ community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.uniq
 # they belong
 chloroplast <- filter(community, str_detect(taxon, "^Chloroplast$"))$rankID
 mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
-community <- mutate_at(community, 5:ncol(community), funs(case_when(
+community <- mutate_at(community, 5:ncol(community), list(~case_when(
   rankID==str_extract(chloroplast, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Chloroplast"],
   rankID==str_extract(chloroplast, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Chloroplast"],
   rankID==str_extract(chloroplast, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Chloroplast"],
   TRUE ~ .))) %>%
   filter(!str_detect(taxon, "^Chloroplast")) %>%
-  mutate_at(5:ncol(.), funs(case_when(
+  mutate_at(5:ncol(.), list(~case_when(
[…]

### 2020-01-31 code/plot_pcoa.R

@@ -17,7 +17,7 @@ rarefied <- shared %>%
   rrarefy(., min(rowSums(.))) %>%
   as.tibble() %>%
   add_column("Group"=shared$Group, .before=TRUE) %>%
-  select_if(funs(!is.numeric(.) || sum(.)!=0))
+  select_if(list(~!is.numeric(.) || sum(.)!=0))
 
 # Loading metadata 
 metadata <- read_tsv("data/raw/metadata.csv")
@@ -73,7 +73,7 @@ p1 <- ggplot() +
 # Seawater samples
 # Filter samples from the water column
 rarefied_metadata_select <- filter(rarefied_metadata, station=="F") %>%
-  select_if(funs(!is.numeric(.) || sum(.)!=0))
+  select_if(list(~!is.numeric(.) || sum(.)!=0))
 
 # Generation of PCoA data
 spe.bray <- vegdist(select(rarefied_metadata_select, starts_with("Otu")))
@@ -99,7 +99,7 @@ p2 <- ggplot() +
 # Cymodocea nodosa samples
 # Filter samples from the water column
 rarefied_metadata_select <- filter(rarefied_metadata, station=="FCyM") %>%
-  select_if(funs(!is.numeric(.) || sum(.)!=0))
+  select_if(list(~!is.numeric(.) || sum(.)!=0))
 
 # Generation of PCoA data
 spe.bray <- vegdist(select(rarefied_metadata_select, starts_with("Otu")))
@@ -125,7 +125,7 @@ p3 <- ggplot() +
 # Caulerpa cylindracea samples
 # Filter samples from the water column
 rarefied_metadata_select <- filter(rarefied_metadata, str_detect(station, "^FCa")) %>%
-  select_if(funs(!is.numeric(.) || sum(.)!=0))
+  select_if(list(~!is.numeric(.) || sum(.)!=0))
 
 # Generation of PCoA data
 spe.bray <- vegdist(select(rarefied_metadata_select, starts_with("Otu")))

### 2020-01-31 code/plot_rarefaction.R

@@ -111,7 +111,7 @@ p <- ggplot(metadata_rarefaction, aes(x=numsampled, y=sobs, group=ID,
   labs(x="Number of Sequences", y="") +
   theme +
   theme(legend.position="bottom", legend.title=element_blank(),
-        legend.text=element_text(size=10, margin=margin(r=0.2, unit="cm")),
+        legend.text=element_text(size=10, margin=margin(r=0.1, unit="cm")),
         legend.key.width=unit(1.4, "cm"), legend.key.height=unit(0.5, "cm"),
         legend.key=element_rect(fill="white"), legend.justification=c("top"),
         legend.text.align=0)

### 2020-01-31 code/plot_rarefaction.R
Resolved legend spacing problem.

@@ -111,10 +111,10 @@ p <- ggplot(metadata_rarefaction, aes(x=numsampled, y=sobs, group=ID,
   labs(x="Number of Sequences", y="") +
   theme +
   theme(legend.position="bottom", legend.title=element_blank(),
-        legend.text=element_text(size=10, margin=margin(r=0.1, unit="cm")),
+        legend.text=element_text(size=10, margin=margin(r=0.2, unit="cm")),
         legend.key.width=unit(1.4, "cm"), legend.key.height=unit(0.5, "cm"),
         legend.key=element_rect(fill="white"), legend.justification=c("top"),
-        legend.text.align=0)
+        legend.text.align=0, legend.spacing.x=unit(0, "cm"))
 
 # Common legend extraction
 legend <- cowplot::get_legend(p)

### 2020-01-31 code/plot_community_bar_plot.R
.gitignore changed

@@ -4,6 +4,7 @@
 # A script to plot the community structure of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary
 #               data/raw/metadata.csv
+#               data/raw/group_colors.csv
 # Produces: results/figures/community_bar_plot.jpg
 #
 #################################################################################################################

### 2020-01-31 code/plot_community_bar_plot_alphaproteobacteria.R

@@ -4,6 +4,7 @@
 # A script to plot sequence abundances of groups from the class Alphaproteobacteria of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary
 #               data/raw/metadata.csv
+#               data/raw/group_colors.csv
 # Produces: results/figures/alphaproteobacteria_bar_plot.jpg
 #
 #################################################################################################################

### 2020-01-31 code/plot_community_bar_plot_bacteroidota.R

@@ -4,6 +4,7 @@
 # A script to plot sequence abundances of groups from the phylum Bacteoidetes of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary
 #               data/raw/metadata.csv
+#               data/raw/group_colors.csv
 # Produces: results/figures/bacteroidota_bar_plot.jpg
 #
 #################################################################################################################

### 2020-01-31 code/plot_community_bar_plot_chloroplast.R

@@ -4,6 +4,7 @@
 # A script to plot chloroplast sequence abundances of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary
 #               data/raw/metadata.csv
+#               data/raw/group_colors.csv
 # Produces: results/figures/chloroplast_bar_plot.jpg
 #
 #################################################################################################################

### 2020-01-31 code/plot_community_bar_plot_cyanobacteria.R

@@ -4,6 +4,7 @@
 # A script to plot sequence abundances of groups from the phylum Cyanobacteria of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary
 #               data/raw/metadata.csv
+#               data/raw/group_colors.csv
 # Produces: results/figures/cyanobacteria_bar_plot.jpg
 #
 #################################################################################################################

### 2020-01-31 code/plot_community_bar_plot_desulfobacterota.R

@@ -4,6 +4,7 @@
 # A script to plot sequence abundances of groups from the class Gammaproteobacteria of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary
 #               data/raw/metadata.csv
+#               data/raw/group_colors.csv
 # Produces: results/figures/desulfobacterota_bar_plot.jpg
 #
 #################################################################################################################

### 2020-01-31 code/plot_community_bar_plot_gammaproteobacteria.R

@@ -4,6 +4,7 @@
 # A script to plot sequence abundances of groups from the class Gammaproteobacteria of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary
 #               data/raw/metadata.csv
+#               data/raw/group_colors.csv
 # Produces: results/figures/gammaproteobacteria_bar_plot.jpg
 #
 #################################################################################################################

### 2020-01-31 code/plot_community_bar_plot_alphaproteobacteria.R
Bar plots code updated.

@@ -68,11 +68,12 @@ color <- read_tsv("data/raw/group_colors.csv") %>%
   deframe()
 
 # Generation of italic names for taxonomic groups
-names <- parse(text=case_when(plot$taxon=="SAR116_clade_unclassified" ~ "plain('SAR116 Clade (NR)')",
-                              str_detect(plot$taxon, "uncultured") ~ paste0("plain('Uncultured')~italic('", str_remove(plot$taxon, "uncultured_"), "')"),
+names <- parse(text=case_when(str_detect(plot$taxon, "uncultured") ~ paste0("plain('Uncultured')~italic('", str_remove(plot$taxon, "uncultured_"), "')"),
                               str_detect(plot$taxon, "unclassified") ~ paste0("italic('", str_remove(plot$taxon, "_unclassified"), "')~plain('(NR)')"),
                               plot$taxon=="OCS116_clade_ge" ~ "plain('OCS116 Clade')",
+                              plot$taxon=="Candidatus_Puniceispirillum" ~ "italic('\"Candidatus')~plain('Puniceispirillum\"')",
                               plot$taxon=="SAR116_clade_ge" ~ "plain('SAR116 Clade')",
+                              plot$taxon=="Stappiaceae_ge" ~ "italic('Stappiaceae')",
                               plot$taxon=="HIMB11" ~ "plain('HIMB11')",
                               plot$taxon=="AEGEAN-169_marine_group_ge" ~ "plain('AEGEAN-169 Marine Group')",
                               plot$taxon=="Clade_Ia" ~ "plain('Clade Ia')",
@@ -114,7 +115,7 @@ theme <- theme(text=element_text(family="Times"), line=element_line(color="black
                axis.line.x=element_line(), axis.line.y=element_blank(),
                axis.ticks.x=element_line(), axis.ticks.y.left=element_line(),
                axis.text.y=element_text(size=12, color="black"),
-               axis.text.x=element_blank(), axis.title.y=element_text(size=14, color="black", vjust=-0.75),
+               axis.text.x=element_blank(), axis.title.y=element_text(size=14, color="black", vjust=-0.75, hjust=0.365),
                panel.background=element_blank(), plot.margin=unit(c(5.5, 5.5, 5.5, -11), "pt"), legend.position="none",
[…]

### 2020-01-31 code/plot_community_bar_plot_cyanobacteria.R

@@ -106,7 +106,7 @@ theme <- theme(text=element_text(family="Times"), line=element_line(color="black
                axis.line.x=element_line(), axis.line.y=element_blank(),
                axis.ticks.x=element_line(), axis.ticks.y.left=element_line(),
                axis.text.y=element_text(size=12, color="black"),
-               axis.text.x=element_blank(), axis.title.y=element_text(size=14, color="black", vjust=-0.75),
+               axis.text.x=element_blank(), axis.title.y=element_text(size=14, color="black", vjust=-0.75, hjust=0.365),
                panel.background=element_blank(), plot.margin=unit(c(5.5, 5.5, 5.5, -11), "pt"), legend.position="none",
                plot.title=element_text(size=16, hjust=0.13))
 

### 2020-01-31 code/plot_community_bar_plot_desulfobacterota.R

@@ -70,6 +70,7 @@ color <- read_tsv("data/raw/group_colors.csv") %>%
 # Generation of italic names for taxonomic groups
 names <- parse(text=case_when(str_detect(plot$taxon, "uncultured") ~ paste0("plain('Uncultured')~italic('", str_remove(plot$taxon, "uncultured_"), "')"),
                               str_detect(plot$taxon, "unclassified") ~ paste0("italic('", str_remove(plot$taxon, "_unclassified"), "')~plain('(NR)')"),
+                              plot$taxon=="SEEP-SRB4" ~ "plain('SEEP-SRB4')",
                               plot$taxon=="P3OB-42_ge" ~ "plain('P3OB-42')",
                               plot$taxon=="Other_Desulfobacterota" ~ "plain('Other')~italic('Desulfobacterota')",
                               TRUE ~ paste0("italic('", plot$taxon, "')")))
@@ -106,7 +107,7 @@ theme <- theme(text=element_text(family="Times"), line=element_line(color="black
                axis.line.x=element_line(), axis.line.y=element_blank(),
                axis.ticks.x=element_line(), axis.ticks.y.left=element_line(),
                axis.text.y=element_text(size=12, color="black"),
-               axis.text.x=element_blank(), axis.title.y=element_text(size=14, color="black", vjust=-0.75),
+               axis.text.x=element_blank(), axis.title.y=element_text(size=14, color="black", vjust=-0.75, hjust=0.365),
                panel.background=element_blank(), plot.margin=unit(c(5.5, 5.5, 5.5, -11), "pt"), legend.position="none",
                plot.title=element_text(size=16, hjust=0.13))
 

### 2020-01-31 code/plot_community_bar_plot_gammaproteobacteria.R

@@ -109,7 +109,7 @@ theme <- theme(text=element_text(family="Times"), line=element_line(color="black
                axis.line.x=element_line(), axis.line.y=element_blank(),
                axis.ticks.x=element_line(), axis.ticks.y.left=element_line(),
                axis.text.y=element_text(size=12, color="black"),
-               axis.text.x=element_blank(), axis.title.y=element_text(size=14, color="black", vjust=-0.75),
+               axis.text.x=element_blank(), axis.title.y=element_text(size=14, color="black", vjust=-0.75, hjust=0.365),
                panel.background=element_blank(), plot.margin=unit(c(5.5, 5.5, 5.5, -11), "pt"), legend.position="none",
                plot.title=element_text(size=16, hjust=0.13))
 

### 2020-02-03 code/format_taxonomy.R
Silva v138 optimization.

@@ -43,15 +43,6 @@ tax.in <- read.table("data/references/silva.nr_v138.full", header=F, stringsAsFa
 colnames(tax.in) <- c("taxid", "taxlabel")
 
 # Problem corrections observed in the SILVA ARB file (silva.nr_v138.full)
-# Following line corrects the Bacteria;Bacteroidetes;Bacteroidia;Flavobacteriales;Flavobacteriaceae;Polaribacter;Polaribacter; problem
-tax.in$taxlabel <- gsub("Polaribacter;Polaribacter", "Polaribacter", tax.in$taxlabel)
-# Following line corrects the Bacteria;RsaHf231; problem
-# tax.in$taxlabel <- gsub("RsaHf231;", "RsaHF231;", tax.in$taxlabel)
-# Following line corrects the Bacteria;GBS-1;; problem
-# tax.in$taxlabel <- gsub("Bacteria;GBS-1;;", "Bacteria;GBS-1;", tax.in$taxlabel)
-# Following line corrects the Eukaryota;Opisthokonta;Nucletmycea;Fungi;Dikarya;Basidiomycota;Agaricomycotina;Tremellomycetes;Tremellales;Incertae sedis;Sterigmatosporidium; problem
-# tax.in$taxlabel <- gsub(";Incertae sedis;", ";Incertae Sedis;", tax.in$taxlabel)
-# tax.in$taxlabel <- gsub(";[[:space:]]+$", ";", tax.in$taxlabel)
 # Following line corrects the Bacteria;Proteobacteria;Alphaproteobacteria;Rhizobiales;Rhizobiaceae;Mesorhizobium;Mesorhizobium; problem
 tax.in$taxlabel <- gsub("Bacteria;Proteobacteria;Alphaproteobacteria;Rhizobiales;Rhizobiaceae;Mesorhizobium;Mesorhizobium;", "Bacteria;Proteobacteria;Alphaproteobacteria;Rhizobiales;Rhizobiaceae;Mesorhizobium;", tax.in$taxlabel)
 

### 2020-02-03 code/plot_community_bar_plot_alphaproteobacteria.R
Aphaproteobacteria line legend fix

@@ -206,7 +206,7 @@ p <- cowplot::ggdraw() +
   cowplot::draw_plot(fcym, x=0.317, y=0.545, width=0.683, height=0.228) +
   cowplot::draw_plot(fcam, x=0.317, y=0.317, width=0.683, height=0.228) +
   cowplot::draw_plot(fca, x=0.317, y=0, width=0.683, height=0.317) +
-  cowplot::draw_plot(legend, x=0.12, y=0.28, width=0.1, height=0.2) +
+  cowplot::draw_plot(legend, x=0.120, y=0.280, width=0.100, height=0.200) +
   cowplot::draw_line(x=c(0.135, 0.135), y=c(0.288, 0.347), size=0.5) +
   cowplot::draw_label("SAR11 Clade", x=0.145, y=0.317, hjust=0,  fontfamily="Times", size=12)
 ggsave("results/figures/alphaproteobacteria_bar_plot.jpg", p, width=210, height=297, units="mm")

### 2020-02-03 code/plot_community_bar_plot_alphaproteobacteria.R
Legend line problem resolved by packages update.

@@ -207,6 +207,6 @@ p <- cowplot::ggdraw() +
   cowplot::draw_plot(fcam, x=0.317, y=0.317, width=0.683, height=0.228) +
   cowplot::draw_plot(fca, x=0.317, y=0, width=0.683, height=0.317) +
   cowplot::draw_plot(legend, x=0.120, y=0.280, width=0.100, height=0.200) +
-  cowplot::draw_line(x=c(0.135, 0.135), y=c(0.288, 0.347), size=0.5) +
-  cowplot::draw_label("SAR11 Clade", x=0.145, y=0.317, hjust=0,  fontfamily="Times", size=12)
+  cowplot::draw_line(x=c(0.135, 0.135), y=c(0.284, 0.343), size=0.5) +
+  cowplot::draw_label("SAR11 Clade", x=0.145, y=0.313, hjust=0,  fontfamily="Times", size=12)
 ggsave("results/figures/alphaproteobacteria_bar_plot.jpg", p, width=210, height=297, units="mm")

### 2020-02-03 code/plot_community_bar_plot_alphaproteobacteria.R
Legend line adjustment.

@@ -207,6 +207,6 @@ p <- cowplot::ggdraw() +
   cowplot::draw_plot(fcam, x=0.317, y=0.317, width=0.683, height=0.228) +
   cowplot::draw_plot(fca, x=0.317, y=0, width=0.683, height=0.317) +
   cowplot::draw_plot(legend, x=0.120, y=0.280, width=0.100, height=0.200) +
-  cowplot::draw_line(x=c(0.135, 0.135), y=c(0.284, 0.343), size=0.5) +
-  cowplot::draw_label("SAR11 Clade", x=0.145, y=0.313, hjust=0,  fontfamily="Times", size=12)
+  cowplot::draw_line(x=c(0.135, 0.135), y=c(0.285, 0.344), size=0.5) +
+  cowplot::draw_label("SAR11 Clade", x=0.145, y=0.314, hjust=0,  fontfamily="Times", size=12)
 ggsave("results/figures/alphaproteobacteria_bar_plot.jpg", p, width=210, height=297, units="mm")

### 2020-02-04 submission/manuscript.Rmd
Logfile remove logic.

@@
 \newpage
-```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant bacterial sequences.\\label{community}", fig.pos="ht"}
+```{r out.width="80%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant bacterial sequences.\\label{community}", fig.pos="ht"}
 include_graphics("../results/figures/community_bar_plot.jpg")

### 2020-02-14 submission/manuscript.Rmd
Added introduction and supp informations.

@@
 
-\vspace{80mm}
+\vspace{70mm}
 
@@
 ## Introduction
+Marine macrophytes (seagrasses and macroalgae) are important ecosystem engineers that form close associations with microorganism belonging to all three domains of life [@Tarquinio2019; @Egan2013].
+Microbes can live within macrophyte tissue as endophytes or can form epiphytic communities on surfaces of leaves and thalli [@Tarquinio2019; @Egan2013; @Hollants2013].
+Epiphytic and endophytic microbial communities form a close functional relationship with the macrophyte host.
+It was proposed that this close relationship constitutes a holobiont, an integrated community where the macrophyte organism and its symbiotic partners support each other [@Margulis1991; @Tarquinio2019; @Egan2013].
+
+Biofilms formed from microbial epiphytes can contain diverse taxonomic groups and harbor cell densities from 10^2^ to 10^7^ \si{\cells\per\cm\squared} [@Armstrong2000; @Bengtsson2010; @Burke2011b].
+In such an environment a number of positive and negative interactions between the macrophyte and colonizing microorganisms have been described [@Tarquinio2019; @Egan2013; @Hollants2013].
+Macrophytes can promote growth of associated microbes by nutrient exudation, while in return microorganisms may support macrophyte performance through improved nutrient availability, phytohormone production and protection form toxic compounds, oxidative stress, biofouling organisms and pathogens [@Tarquinio2019; @Egan2013; @Hollants2013].
+Beside this positive interactions, macrophytes can  negatively impact the associated microbes such as pathogenic bacteria by producing reactive oxygen species and secondary metabolites [@Tarquinio2019; @Egan2013; @Hollants2013].
+
+All these ecological roles are carried out by a taxonomically diverse community of microorganisms.
+At higher taxonomic ranks a core set of macrophyte epiphytic taxa was described consisting mainly of *Alphaproteobacteria*, *Gammaproteobacteria*, *Desulfobacterota*, *Bacteroidota*, *Cyanobacteria*, *Actinobacteriota*, *Firmicutes*, *Planctomycetota*, *Chloroflexi* and *Verrucomicrobiota* [@Lachnit2011; @Crump2008; @Tujula2010].
+In contrast, at lower taxonomic ranks host specific microbial communities were described [@Roth-Schulze2016; @Lachnit2011].
+Recently, it was shown that even different morphological niches within the same alga had a higher influence on bacterial community variation than biogeography or environmental factors [@Morrissey2019].
+While there is high community variation between host species is was observed that the majority of metagenomes determined functions were conserved both between host species and individuals [@Burke2011a; @Roth-Schulze2016].
+This discrepancy between taxonomic and functional composition could be explained by the lottery hypothesis.
+It postulates that an initial random colonization step is performed from a set of functionally equivalent taxonomic groups resulting in taxonomically different epiphytic communities sharing a core set of functional genes [@Burke2011a; @Roth-Schulze2016].
+In addition, some of the variation in the observed data could be attributed to different techniques used in various studies, such as different protocols for epiphytic cell detachment and/or DNA isolation, as no standard protocol to study epiphytic communities was established [@Ugarelli2019; @Korlevic_in_press, submitted].
+
+The majority of studies describing macrophyte epiphytic communities did not encompass seasonal changes [@Ugarelli2019; @Lachnit2009; @Crump2008; @Burke2011b; @Roth-Schulze2016].
+In addition, if seasonal changes were taken into account low temporal frequency and/or methodologies that do not allow for high taxonomic resolution were used [@Tujula2010; @Lachnit2011; @Miranda2013].
+In the present study we describe the seasonal dynamics of bacterial and archaeal communities on the surfaces of the seagrass *Cymodocea nodosa* and siphonous macroalgae *Caulerpa cylindracea* determined on a mostly monthly scale.
+Bacterial and archaeal epiphytes were sampled in a meadow of *Cymodocea nodosa* invaded by the invasive *Caulerpa cylindracea* and in a locality of only *Caulerpa cylindracea* located in the proximity of the meadow.
+In addition, for comparison the community of the surrounding seawater was characterized.
 
@@
 ### Sampling
-Leaves of *Cymodocea nodosa* were sampled in a *Cymodocea nodosa* meadow in the Bay of Saline (45°7´5˝N, 13°37´20˝E) and in a *Cymodocea nodosa* meadow invaded by *Caulerpa cylindracea* in the proximity of the village of Funtana (45°10´39˝N, 13°35´42˝E).
-Thalli of *Caulerpa cylindracea* were sampled in the same *Cymodocea nodosa* invaded meadow in Funtana and on a locality of only *Caulerpa cylindracea* located in the proximity of the invaded meadow.
-Leaves and thalli were collected on the same day in two contrasting seasons, on 4 December 2017 and 18 June 2018.
-During spring 2018 the *Cymodocea nodosa* meadow in the Bay of Saline decayed to an extent that no leaves could be retrieved (Najdek *et al.*, unpublished data).
+Leaves of *Cymodocea nodosa* were sampled in a *Cymodocea nodosa* meadow located in the proximity of the village of Funtana (\ang{45;10;39} N, \ang{13;35;42} E).
+Thalli of *Caulerpa cylindracea* were sampled in the same *Cymodocea nodosa* invaded meadow in Funtana and on a locality of only *Caulerpa cylindracea* located close to the invaded meadow.
+Sampling of leaves and thalli was performed approximately monthly from December 2017 to October 2018.
 Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with site seawater.
@@
 Leaves and thalli were washed three times with sterile artificial seawater (ASW) to remove loosely attached microbial cells.
+Surrounding seawater was collected in 10 \si{\l} containers by diving and transported to the laboratory where the whole container volume was filtered through a 20 \si{\um} net.
+The filtrate was further sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate membrane filters (Whatman, United Kingdom) using a peristaltic pump.
+Filters were briefly dried at room temperature and stored at \num{-80} \si{\degreeCelsius}.
+Seawater samples were also collected approximately monthly from July 2017 to October 2018.
 
 ### DNA Isolation
-The DNA was isolated according to the protocol for isolation from filters described in @Massana1997.
-This protocol was modified and adapted for DNA isolation from microbial communities from macrophyte surfaces as described below.
-5 \si{\ml} of lysis buffer (40 \si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar} sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or \si{\g} wet-weight of thalli.
-Lysozyme was added (final concentration 1 \si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius} for 30 \si{\minute}.
-Subsequently, proteinase K (final concentration 0.5 \si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were added and the samples were incubated at 55 \si{\degreeCelsius} for 2 \si{\hour}.
-Following the incubation, tubes were vortexed for 10 \si{\minute} and the mixture containing lyzed epiphytic cells was separated from host leaves or thalli by transferring the solution into a clean tube.
-The lysate was extracted twice with a mixture of phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with chloroform:isoamyl alcohol (24:1).
-After each organic solvent mixture addition tubes were slightly vortexed and centrifuged at 4,500 × g for 10 \si{\minute}.
-Following each centrifugation aqueous phases were retrieved.
-After the final extraction 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added.
-DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} °C and centrifuging at 16,000 × g and 4 \si{\degreeCelsius} for 20 \si{\minute}.
-The pellet was washed twice with 1 \si{\ml} of chilled isopropanol and centrifuged after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 10 \si{\minute}.
-After the first washing step duplicate pellets form the same sample were pooled and transferred to a clean 1.5 \si{\ml} tube.
-The dried pellet was resuspended in 100 \si{\ul} of deionized water.
+DNA from surfaces of *Cymodocea nodosa* and *Caulerpa cylindracea* was isolated using a previously modified and adapted protocol that allows a selective epiphytic DNA isolation [@Korlevic_in_press, submitted; @Massana1997].
+Briefly, leaves and thalli are incubated in a lysis buffer and treated with lysozyme and proteinase K.
+Following the incubations, the mixture containing lyzed epiphytic cells is separated from leaves and thalli and extracted using a phenol-chloroform procedure.
+Finally, the extracted DNA is precipitated using isopropanol.
+DNA from seawater picoplankton was isolated from 0.2 \si{\um} polycarbonate filters according to [@Massana1997] with a slight modification.
+Following the phenol-chloroform extraction steps 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added.
+DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 20,000 × g and 4 \si{\degreeCelsius} for 21 \si{\minute}.
+The pellet was washed twice with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}.
+Dried pellets were resuspended in 50 -- 100 \si{\ul} of deionized water.
 
 ### Illumina 16S rRNA Sequencing
-An aliquot of isolated DNA was treated with RNase A (final concentration 200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}.
-The DNA concentration was determined using the Quant-iT PicoGreen dsDNA Assay Kit (Invitrogen, USA) according to the manufacturer’s instructions and diluted to 1 \si{\ng\per\ul}.
+Illumina MiSeq sequencing of the V4 16S rRNA region was performed as described previously [@Korlevic_in_press, submitted].
 The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
@@
 These primers contained on their 5' end a tagged sequence.
-Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained: 1 × Q5 Reaction Buffer , 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template.
-Cycling conditions were: initial denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50 \si{\degreeCelsius} for 60 \si{\s} and elongation at 72 \si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at 72 \si{\degreeCelsius} for 10 \si{\minute}.
-The four parallel reactions volumes were pooled and PCR products were purified using the GeneJET PCR Purification Kit (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and following the protocol that included isopropanol addition for better small DNA fragment yield.
-The column was eluted in 30 \si{\ul} of deionized water.
-Purified PCR products were sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories, Martinsried, Germany.
-Before sequencing, at IMGM the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR.
+Purified PCR products were sent for Illumina MiSeq sequencing at IMGM Laboratories, Martinsried, Germany.
+Before sequencing, the second PCR amplification of the two-step PCR procedure was performed at IMGM using primers targeting the tagged region incorporated in the first PCR.
 In addition, these primers contained adapter and sample-specific index sequences.
-The second PCR was carried out for 8 cycles.
-Beside samples, a positive and negative control were sequenced.
-A negative control was comprised of four parallel PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
+Beside samples, a positive and negative control for each sequencing batch were sequenced.
+A negative control was comprised of PCR reactions without DNA template, while for a positive controls a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
 The sequences obtained in this study have been submitted to the European Nucleotide Archive (ENA) under accession numbers **TO BE ADDED LATER!**.
@@
 Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
-For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 132; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
-Sequences classified as chloroplasts by mothur were exported, aligned using SILVA incremental aligner (SINA, version 1.6.0) [@Pruesse2012] against the same SILVA SSU Ref NR 99 database (release 132) and imported into ARB (version 6.0.6) [@Ludwig2004] for further phylogenetic analysis using the same database.
-Reference sequences close to imported ones were selected and used to calculate a phylogenetic tree using the Maximum Likelihood algorithm RAxML (version 7.0.3) with 1000 tree replicates [@Stamatakis2006].
-Imported partial chloroplast sequences were added to the tree using the maximum parsimony criteria and not allowing changes to tree topology.
-Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019], package tidyverse (version 1.2.1) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Zhu2019; @Allaire2019; @Xie2018].
+For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
+Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019], package vegan (version 2.5-6) [@Oksanen2019], package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014].
 The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (**TO BE ADDED LATER!**).
-Based on the ATCC MSA-1002 mock community included in the analysis a sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
-In addition, the negative control processed together with the samples yielded only `r nc` sequences after sequence quality curation.
-
-### Protein Isolation
-Proteins were isolated according to the protocol for isolation from soil described in @Chourey2010 and modified by @Hultman2015.
-These protocols were further modified and adapted for protein isolation from microbial communities form macrophyte surfaces as described below.
-20 \si{\ml} of protein extraction buffer (4 \si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl [pH 8.0]) was added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of thalli.
-The mixture was incubated in boiling water for 5 \si{\minute}, vortexed for 10 \si{\minute} and incubated again in boiling water for 5 \si{\minute}.
-After a brief vortex the lysate was transferred to a clean tube separating the host leaves or thalli from the mixture containing lyzed epiphytic cells.
-Dithiothreitol (DTT; final concentration 24 \si{\milli\Molar}) was added and proteins were precipitated with chilled 100 \si{\percent} trichloroacetic acid (TCA; final concentration 20 \si{\percent}) overnight at \num{-20} °C.
-Precipitated proteins were centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40 \si{\minute}.
-The obtained protein pellet was washed three times with chilled acetone.
-During the first washing step the pellet was transferred to a clean 1.5 \si{\ml} tube.
-After each washing step samples were centrifuged at 20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}.
-Dried pellets were stored at \num{-80} °C until further analysis.
-
-### Metaproteomics
-Isolated proteins were whole trypsin digested using the FASP (filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK) according to the manufacturer's instructions [@Wisniewski2009] with small modifications.
-Before loading the solution to the column, protein pellets were solubilized in a urea sample buffer included in the kit amended with DTT (final concentration 100 \si{\milli\Molar}) for 45 \si{\minute} at room temperature and centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to remove larger particles.
-The first washing step after protein solution loading was repeated twice.
-In addition, centrifugation steps were prolonged if the column was clogged.
-Trypsin digestion was performed on column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
-The final filtrate containing digested proteins was acidified with 1 \si{\percent} (final concentration) trifluoroacetic acid, freezed at \num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent to the Vienna Metabolomics Center (University of Vienna) for metaproteomic analysis.
-Peptides were resuspended in 1 \si{\percent} (final concentration) trifluoroacetic acid, desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
-Obtained MS/MS spectra were searched against a protein database from metagenomic assembly published in @Burke2011a using SEQUEST-HIT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA).
-The target-decoy approach was used to reduce the probability of false peptide identification.
-Results whose false discovery rate at the peptide level was \SI{< 1}{\percent} were kept.
-For protein identification a minimum of of two peptides and one unique peptide were required.
-For protein quantification, a chromatographic peak area-based free quantitative method was applied.
-
-### Confocal Microscopy
-Host leaves and thalli from DNA and protein isolation steps were washed seven times in deionized water and fixed with formaldehyde (final concentration ~ 3 \si{\percent}).
-In addition, nontreated leaves and thalli, washed three times in ASW to remove loosely attached microbial cells, were fixed in the same concentration of formaldehyde and used as a positive control.
-For long therm storage, fixed leaves and thalli were immersed in a mixture of phosphate-buffered saline (PBS) and ethanol (1:1) and stored at \num{-20} \si{\degreeCelsius}.
-Treated and untreated leaves and thalli segments were stained in a 2 × solution of SYBR Green I and examined under a Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
+Based on the ATCC MSA-1002 mock community included in the analysis an average sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
+In addition, the negative controls processed together with the samples yielded on average only `r nc` sequences after sequence quality curation.
 
 ## Results
+Chloroplast sequences were following the seasonal pattern (\autoref{supp-chloroplast}, \autoref{pcoa}).
 
@@
 
-## Acknowledgements
+## Acknowledgments
 
@@
 ## Figure Captions
+**\autoref{shared}.** \nameref{shared}
+
+**\autoref{pcoa}.** \nameref{pcoa}
+
 **\autoref{community}.** \nameref{community}
 
-\newpage
+**\autoref{cyano}.** \nameref{cyano}
+
+**\autoref{bactero}.** \nameref{bactero}
+
+**\autoref{alpha}.** \nameref{alpha}
+
+**\autoref{gamma}.** \nameref{gamma}
+
+**\autoref{desulfo}.** \nameref{desulfo}
+
 ## Figures
-\newpage
-```{r out.width="80%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant bacterial sequences.\\label{community}", fig.pos="ht"}
+```{r out.width="85%", fig.align="center", fig.cap="Proprotion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\\label{shared}", fig.pos="H"}
+include_graphics("../results/figures/seasonal_shared.jpg")
+```
+
+```{r out.width="100%", fig.align="center", fig.cap="Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+Samples from the same environment or same season are labeld in different colors.
+The proportion of explained variation by each axis is shown on the corresponding axis in parentheses.\\label{pcoa}", fig.pos="H"}
+include_graphics("../results/figures/pcoa_figure.jpg")
+```
+
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant bacterial and archaeal sequences on the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\\label{community}", fig.pos="H"}
 include_graphics("../results/figures/community_bar_plot.jpg")
 ```
+
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant cyanobacterial sequences on the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+The proprotion of cyanobaterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
+NR -- No Relative\\label{cyano}", fig.pos="H"}
+include_graphics("../results/figures/cyanobacteria_bar_plot.jpg")
+```
+
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant sequences within the \\textit{Bacteroidota} on the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+The proprotion of sequences classified as \\textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar.
+NR -- No Relative\\label{bactero}", fig.pos="H"}
+include_graphics("../results/figures/bacteroidota_bar_plot.jpg")
+```
+
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant alphaproteobacterial sequences on the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+The proprotion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
+NR -- No Relative\\label{alpha}", fig.pos="H"}
+include_graphics("../results/figures/alphaproteobacteria_bar_plot.jpg")
+```
+
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant gammaproteobacterial sequences on the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+The proprotion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
+NR -- No Relative\\label{gamma}", fig.pos="H"}
+include_graphics("../results/figures/gammaproteobacteria_bar_plot.jpg")
+
+```
+
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant sequences within the \\textit{Desulfobacterota} on the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+The proprotion of sequences classified as \\textit{Desulfobacterota} in the total bacterial and archaeal community is given above the corresponding bar.
+NR -- No Relative\\label{desulfo}", fig.pos="H"}
+include_graphics("../results/figures/desulfobacterota_bar_plot.jpg")
+```

### 2020-02-14 submission/manuscript.tex
Switch to Silva database v138.
Logfile remove logic.
Added introduction and supp informations.

@@
 \DeclareSIUnit\Molar{\textsc{m}}
+\DeclareSIUnit\cells{\text{cells}}
 
@@
 
+\usepackage{float}
+
+\usepackage{xr}
+\externaldocument[supp-]{supplementary}
+
 \begin{document}
@@
 
+Marine macrophytes (seagrasses and macroalgae) are important ecosystem
+engineers that form close associations with microorganism belonging to
+all three domains of life (Egan \emph{et al.}, 2013; Tarquinio \emph{et
+al.}, 2019). Microbes can live within macrophyte tissue as endophytes or
+can form epiphytic communities on surfaces of leaves and thalli (Egan
+\emph{et al.}, 2013; Hollants \emph{et al.}, 2013; Tarquinio \emph{et
+al.}, 2019). Epiphytic and endophytic microbial communities form a close
+functional relationship with the macrophyte host. It was proposed that
+this close relationship constitutes a holobiont, an integrated community
+where the macrophyte organism and its symbiotic partners support each
+other (Margulis, 1991; Egan \emph{et al.}, 2013; Tarquinio \emph{et
+al.}, 2019).
+
+Biofilms formed from microbial epiphytes can contain diverse taxonomic
+groups and harbor cell densities from 10\textsuperscript{2} to
+10\textsuperscript{7} \si{\cells\per\cm\squared} (Armstrong \emph{et
+al.}, 2000; Bengtsson \emph{et al.}, 2010; Burke \emph{et al.}, 2011b).
+In such an environment a number of positive and negative interactions
+between the macrophyte and colonizing microorganisms have been described
+(Egan \emph{et al.}, 2013; Hollants \emph{et al.}, 2013; Tarquinio
+\emph{et al.}, 2019). Macrophytes can promote growth of associated
+microbes by nutrient exudation, while in return microorganisms may
+support macrophyte performance through improved nutrient availability,
+phytohormone production and protection form toxic compounds, oxidative
+stress, biofouling organisms and pathogens (Egan \emph{et al.}, 2013;
+Hollants \emph{et al.}, 2013; Tarquinio \emph{et al.}, 2019). Beside
+this positive interactions, macrophytes can negatively impact the
+associated microbes such as pathogenic bacteria by producing reactive
+oxygen species and secondary metabolites (Egan \emph{et al.}, 2013;
+Hollants \emph{et al.}, 2013; Tarquinio \emph{et al.}, 2019).
+
+All these ecological roles are carried out by a taxonomically diverse
+community of microorganisms. At higher taxonomic ranks a core set of
+macrophyte epiphytic taxa was described consisting mainly of
+\emph{Alphaproteobacteria}, \emph{Gammaproteobacteria},
+\emph{Desulfobacterota}, \emph{Bacteroidota}, \emph{Cyanobacteria},
+\emph{Actinobacteriota}, \emph{Firmicutes}, \emph{Planctomycetota},
+\emph{Chloroflexi} and \emph{Verrucomicrobiota} (Crump and Koch, 2008;
+Tujula \emph{et al.}, 2010; Lachnit \emph{et al.}, 2011). In contrast,
+at lower taxonomic ranks host specific microbial communities were
+described (Lachnit \emph{et al.}, 2011; Roth-Schulze \emph{et al.},
+2016). Recently, it was shown that even different morphological niches
+within the same alga had a higher influence on bacterial community
+variation than biogeography or environmental factors (Morrissey \emph{et
+al.}, 2019). While there is high community variation between host
+species is was observed that the majority of metagenomes determined
+functions were conserved both between host species and individuals
+(Burke \emph{et al.}, 2011a; Roth-Schulze \emph{et al.}, 2016). This
+discrepancy between taxonomic and functional composition could be
+explained by the lottery hypothesis. It postulates that an initial
+random colonization step is performed from a set of functionally
+equivalent taxonomic groups resulting in taxonomically different
+epiphytic communities sharing a core set of functional genes (Burke
+\emph{et al.}, 2011a; Roth-Schulze \emph{et al.}, 2016). In addition,
+some of the variation in the observed data could be attributed to
+different techniques used in various studies, such as different
+protocols for epiphytic cell detachment and/or DNA isolation, as no
+standard protocol to study epiphytic communities was established
+(Ugarelli \emph{et al.}, 2019; Korlević \emph{et al.}, submitted).
+
+The majority of studies describing macrophyte epiphytic communities did
+not encompass seasonal changes (Crump and Koch, 2008; Lachnit \emph{et
+al.}, 2009; Burke \emph{et al.}, 2011b; Roth-Schulze \emph{et al.},
+2016; Ugarelli \emph{et al.}, 2019). In addition, if seasonal changes
+were taken into account low temporal frequency and/or methodologies that
+do not allow for high taxonomic resolution were used (Tujula \emph{et
+al.}, 2010; Lachnit \emph{et al.}, 2011; Miranda \emph{et al.}, 2013).
+In the present study we describe the seasonal dynamics of bacterial and
+archaeal communities on the surfaces of the seagrass \emph{Cymodocea
+nodosa} and siphonous macroalgae \emph{Caulerpa cylindracea} determined
+on a mostly monthly scale. Bacterial and archaeal epiphytes were sampled
+in a meadow of \emph{Cymodocea nodosa} invaded by the invasive
+\emph{Caulerpa cylindracea} and in a locality of only \emph{Caulerpa
+cylindracea} located in the proximity of the meadow. In addition, for
+comparison the community of the surrounding seawater was characterized.
+
 \newpage
@@
 Leaves of \emph{Cymodocea nodosa} were sampled in a \emph{Cymodocea
-nodosa} meadow in the Bay of Saline (45°7´5˝N, 13°37´20˝E) and in a
-\emph{Cymodocea nodosa} meadow invaded by \emph{Caulerpa cylindracea} in
-the proximity of the village of Funtana (45°10´39˝N, 13°35´42˝E). Thalli
-of \emph{Caulerpa cylindracea} were sampled in the same \emph{Cymodocea
-nodosa} invaded meadow in Funtana and on a locality of only
-\emph{Caulerpa cylindracea} located in the proximity of the invaded
-meadow. Leaves and thalli were collected on the same day in two
-contrasting seasons, on 4 December 2017 and 18 June 2018. During spring
-2018 the \emph{Cymodocea nodosa} meadow in the Bay of Saline decayed to
-an extent that no leaves could be retrieved (Najdek \emph{et al.},
-unpublished data). Leaves and thalli were collected by diving and
-transported to the laboratory in containers placed on ice and filled
-with site seawater. Upon arrival to the laboratory, \emph{Cymodocea
-nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while
-\emph{Caulerpa cylindracea} thalli were cut into 5 -- 8 \si{\cm} long
-sections. Leaves and thalli were washed three times with sterile
-artificial seawater (ASW) to remove loosely attached microbial cells.
+nodosa} meadow located in the proximity of the village of Funtana
+(\ang{45;10;39} N, \ang{13;35;42} E). Thalli of \emph{Caulerpa
+cylindracea} were sampled in the same \emph{Cymodocea nodosa} invaded
+meadow in Funtana and on a locality of only \emph{Caulerpa cylindracea}
+located close to the invaded meadow. Sampling of leaves and thalli was
+performed approximately monthly from December 2017 to October 2018.
+Leaves and thalli were collected by diving and transported to the
+laboratory in containers placed on ice and filled with site seawater.
+Upon arrival to the laboratory, \emph{Cymodocea nodosa} leaves were cut
+into sections of 1 -- 2 \si{\cm}, while \emph{Caulerpa cylindracea}
+thalli were cut into 5 -- 8 \si{\cm} long sections. Leaves and thalli
+were washed three times with sterile artificial seawater (ASW) to remove
+loosely attached microbial cells. Surrounding seawater was collected in
+10 \si{\l} containers by diving and transported to the laboratory where
+the whole container volume was filtered through a 20 \si{\um} net. The
+filtrate was further sequentially filtered through 3 \si{\um} and 0.2
+\si{\um} polycarbonate membrane filters (Whatman, United Kingdom) using
+a peristaltic pump. Filters were briefly dried at room temperature and
+stored at \num{-80} \si{\degreeCelsius}. Seawater samples were also
+collected approximately monthly from July 2017 to October 2018.
 
@@
 
-The DNA was isolated according to the protocol for isolation from
-filters described in Massana \emph{et al.} (1997). This protocol was
-modified and adapted for DNA isolation from microbial communities from
-macrophyte surfaces as described below. 5 \si{\ml} of lysis buffer (40
-\si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar}
-sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or \si{\g}
-wet-weight of thalli. Lysozyme was added (final concentration 1
-\si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius}
-for 30 \si{\minute}. Subsequently, proteinase K (final concentration 0.5
-\si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were
-added and the samples were incubated at 55 \si{\degreeCelsius} for 2
-\si{\hour}. Following the incubation, tubes were vortexed for 10
-\si{\minute} and the mixture containing lyzed epiphytic cells was
-separated from host leaves or thalli by transferring the solution into a
-clean tube. The lysate was extracted twice with a mixture of
-phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with
-chloroform:isoamyl alcohol (24:1). After each organic solvent mixture
-addition tubes were slightly vortexed and centrifuged at 4,500 × g for
-10 \si{\minute}. Following each centrifugation aqueous phases were
-retrieved. After the final extraction 1/10 of chilled 3 \si{\Molar}
-sodium acetate (pH 5.2) was added. DNA was precipitated by adding 1
-volume of chilled isopropanol, incubating the mixtures overnight at
-\num{-20} °C and centrifuging at 16,000 × g and 4 \si{\degreeCelsius}
-for 20 \si{\minute}. The pellet was washed twice with 1 \si{\ml} of
-chilled isopropanol and centrifuged after each washing step at 20,000 ×
-g and 4 \si{\degreeCelsius} for 10 \si{\minute}. After the first washing
-step duplicate pellets form the same sample were pooled and transferred
-to a clean 1.5 \si{\ml} tube. The dried pellet was resuspended in 100
-\si{\ul} of deionized water.
+DNA from surfaces of \emph{Cymodocea nodosa} and \emph{Caulerpa
+cylindracea} was isolated using a previously modified and adapted
+protocol that allows a selective epiphytic DNA isolation (Massana
+\emph{et al.}, 1997; Korlević \emph{et al.}, submitted). Briefly, leaves
+and thalli are incubated in a lysis buffer and treated with lysozyme and
+proteinase K. Following the incubations, the mixture containing lyzed
+epiphytic cells is separated from leaves and thalli and extracted using
+a phenol-chloroform procedure. Finally, the extracted DNA is
+precipitated using isopropanol. DNA from seawater picoplankton was
+isolated from 0.2 \si{\um} polycarbonate filters according to (Massana
+\emph{et al.}, 1997) with a slight modification. Following the
+phenol-chloroform extraction steps 1/10 of chilled 3 \si{\Molar} sodium
+acetate (pH 5.2) was added. DNA was precipitated by adding 1 volume of
+chilled isopropanol, incubating the mixtures overnight at \num{-20}
+\si{\degreeCelsius} and centrifuging at 20,000 × g and 4
+\si{\degreeCelsius} for 21 \si{\minute}. The pellet was washed twice
+with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged
+after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 5
+\si{\minute}. Dried pellets were resuspended in 50 -- 100 \si{\ul} of
+deionized water.
 
@@
 
-An aliquot of isolated DNA was treated with RNase A (final concentration
-200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA
-concentration was determined using the Quant-iT PicoGreen dsDNA Assay
-Kit (Invitrogen, USA) according to the manufacturer's instructions and
-diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was
-amplified using a two-step PCR procedure. In the first PCR the 515F
-(5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3')
-primers from the Earth Microbiome Project
+Illumina MiSeq sequencing of the V4 16S rRNA region was performed as
+described previously (Korlević \emph{et al.}, submitted). The V4 region
+of the 16S rRNA gene was amplified using a two-step PCR procedure. In
+the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R
+(5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project
 (\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
@@
 Apprill \emph{et al.}, 2015; Parada \emph{et al.}, 2016). These primers
-contained on their 5' end a tagged sequence. Each sample was amplified
-in four parallel 25 \si{\ul} reactions of which each contained: 1 × Q5
-Reaction Buffer , 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml}
-BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse
-primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs,
-USA) and 5 \si{\ng} of DNA template. Cycling conditions were: initial
-denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of
-denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50
-\si{\degreeCelsius} for 60 \si{\s} and elongation at 72
-\si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at
-72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions
-volumes were pooled and PCR products were purified using the GeneJET PCR
-Purification Kit (Thermo Fisher Scientific, USA) according to the
-manufacturer's instructions and following the protocol that included
-isopropanol addition for better small DNA fragment yield. The column was
-eluted in 30 \si{\ul} of deionized water. Purified PCR products were
-sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories,
-Martinsried, Germany. Before sequencing, at IMGM the second PCR
-amplification of the two-step PCR procedure was performed using primers
-targeting the tagged region incorporated in the first PCR. In addition,
-these primers contained adapter and sample-specific index sequences. The
-second PCR was carried out for 8 cycles. Beside samples, a positive and
-negative control were sequenced. A negative control was comprised of
-four parallel PCR reactions without DNA template, while for a positive
-control a mock community composed of evenly mixed DNA material
-originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
-used. The sequences obtained in this study have been submitted to the
-European Nucleotide Archive (ENA) under accession numbers \textbf{TO BE
-ADDED LATER!}.
+contained on their 5' end a tagged sequence. Purified PCR products were
+sent for Illumina MiSeq sequencing at IMGM Laboratories, Martinsried,
+Germany. Before sequencing, the second PCR amplification of the two-step
+PCR procedure was performed at IMGM using primers targeting the tagged
+region incorporated in the first PCR. In addition, these primers
+contained adapter and sample-specific index sequences. Beside samples, a
+positive and negative control for each sequencing batch were sequenced.
+A negative control was comprised of PCR reactions without DNA template,
+while for a positive controls a mock community composed of evenly mixed
+DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC,
+USA) was used. The sequences obtained in this study have been submitted
+to the European Nucleotide Archive (ENA) under accession numbers
+\textbf{TO BE ADDED LATER!}.
 
@@
 classification of sequences the SILVA SSU Ref NR 99 database (release
-132; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
-Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
-mothur were exported, aligned using SILVA incremental aligner (SINA,
-version 1.6.0) (Pruesse \emph{et al.}, 2012) against the same SILVA SSU
-Ref NR 99 database (release 132) and imported into ARB (version 6.0.6)
-(Ludwig \emph{et al.}, 2004) for further phylogenetic analysis using the
-same database. Reference sequences close to imported ones were selected
-and used to calculate a phylogenetic tree using the Maximum Likelihood
-algorithm RAxML (version 7.0.3) with 1000 tree replicates (Stamatakis,
-2006). Imported partial chloroplast sequences were added to the tree
-using the maximum parsimony criteria and not allowing changes to tree
-topology. Pipeline data processing and visualization was done using R
-(version 3.6.0) (R Core Team, 2019), package tidyverse (version 1.2.1)
-(Wickham \emph{et al.}, 2019) and multiple other packages (Xie, 2014,
-2015, 2019a, 2019b; Xie \emph{et al.}, 2018; Zhu, 2019; Allaire \emph{et
-al.}, 2019). The detailed analysis procedure including the R Markdown
-file for this paper are available as a GitHub repository (\textbf{TO BE
-ADDED LATER!}). Based on the ATCC MSA-1002 mock community included in
-the analysis a sequencing error rate of 0.01 \si{\percent} was
-determined, which is in line with previously reported values for
-next-generation sequencing data (Kozich \emph{et al.}, 2013; Schloss
-\emph{et al.}, 2016). In addition, the negative control processed
-together with the samples yielded only 2 sequences after sequence
-quality curation.
-
-\hypertarget{protein-isolation}{%
-\subsubsection{Protein Isolation}\label{protein-isolation}}
-
-Proteins were isolated according to the protocol for isolation from soil
-described in Chourey \emph{et al.} (2010) and modified by Hultman
-\emph{et al.} (2015). These protocols were further modified and adapted
-for protein isolation from microbial communities form macrophyte
-surfaces as described below. 20 \si{\ml} of protein extraction buffer (4
-\si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl {[}pH 8.0{]}) was
-added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of
-thalli. The mixture was incubated in boiling water for 5 \si{\minute},
-vortexed for 10 \si{\minute} and incubated again in boiling water for 5
-\si{\minute}. After a brief vortex the lysate was transferred to a clean
-tube separating the host leaves or thalli from the mixture containing
-lyzed epiphytic cells. Dithiothreitol (DTT; final concentration 24
-\si{\milli\Molar}) was added and proteins were precipitated with chilled
-100 \si{\percent} trichloroacetic acid (TCA; final concentration 20
-\si{\percent}) overnight at \num{-20} °C. Precipitated proteins were
-centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40 \si{\minute}.
-The obtained protein pellet was washed three times with chilled acetone.
-During the first washing step the pellet was transferred to a clean 1.5
-\si{\ml} tube. After each washing step samples were centrifuged at
-20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets
-were stored at \num{-80} °C until further analysis.
-
-\hypertarget{metaproteomics}{%
-\subsubsection{Metaproteomics}\label{metaproteomics}}
-
-Isolated proteins were whole trypsin digested using the FASP
-(filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK)
-according to the manufacturer's instructions (Wiśniewski \emph{et al.},
-2009) with small modifications. Before loading the solution to the
-column, protein pellets were solubilized in a urea sample buffer
-included in the kit amended with DTT (final concentration 100
-\si{\milli\Molar}) for 45 \si{\minute} at room temperature and
-centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to
-remove larger particles. The first washing step after protein solution
-loading was repeated twice. In addition, centrifugation steps were
-prolonged if the column was clogged. Trypsin digestion was performed on
-column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
-The final filtrate containing digested proteins was acidified with 1
-\si{\percent} (final concentration) trifluoroacetic acid, freezed at
-\num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent
-to the Vienna Metabolomics Center (University of Vienna) for
-metaproteomic analysis. Peptides were resuspended in 1 \si{\percent}
-(final concentration) trifluoroacetic acid, desalted using the Pierce
-C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's
-instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap
-Mass Spectrometer (Thermo Fisher Scientific, USA). Obtained MS/MS
-spectra were searched against a protein database from metagenomic
-assembly published in Burke \emph{et al.} (2011) using SEQUEST-HIT
-engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo
-Fisher Scientific, USA). The target-decoy approach was used to reduce
-the probability of false peptide identification. Results whose false
-discovery rate at the peptide level was \SI{< 1}{\percent} were kept.
-For protein identification a minimum of of two peptides and one unique
-peptide were required. For protein quantification, a chromatographic
-peak area-based free quantitative method was applied.
-
-\hypertarget{confocal-microscopy}{%
-\subsubsection{Confocal Microscopy}\label{confocal-microscopy}}
-
-Host leaves and thalli from DNA and protein isolation steps were washed
-seven times in deionized water and fixed with formaldehyde (final
-concentration \textasciitilde{} 3 \si{\percent}). In addition,
-nontreated leaves and thalli, washed three times in ASW to remove
-loosely attached microbial cells, were fixed in the same concentration
-of formaldehyde and used as a positive control. For long therm storage,
-fixed leaves and thalli were immersed in a mixture of phosphate-buffered
-saline (PBS) and ethanol (1:1) and stored at \num{-20}
-\si{\degreeCelsius}. Treated and untreated leaves and thalli segments
-were stained in a 2 × solution of SYBR Green I and examined under a
-Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
+138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
+Yilmaz \emph{et al.}, 2014). Pipeline data processing and visualization
+was done using R (version 3.6.0) (R Core Team, 2019), package vegan
+(version 2.5-6) (Oksanen \emph{et al.}, 2019), package tidyverse
+(version 1.3.0) (Wickham \emph{et al.}, 2019) and multiple other
+packages (Xie, 2014, 2015, 2019a, 2019b, 2020; Neuwirth, 2014; Xie
+\emph{et al.}, 2018; Allaire \emph{et al.}, 2019; Zhu, 2019). The
+detailed analysis procedure including the R Markdown file for this paper
+are available as a GitHub repository (\textbf{TO BE ADDED LATER!}).
+Based on the ATCC MSA-1002 mock community included in the analysis an
+average sequencing error rate of 0.01 \si{\percent} was determined,
+which is in line with previously reported values for next-generation
+sequencing data (Kozich \emph{et al.}, 2013; Schloss \emph{et al.},
+2016). In addition, the negative controls processed together with the
+samples yielded on average only 2 sequences after sequence quality
+curation.
 
@@
 
+Chloroplast sequences were following the seasonal pattern
+(\autoref{supp-chloroplast}, \autoref{pcoa}).
+
 \hypertarget{discussion}{%
@@
 
-\hypertarget{acknowledgements}{%
-\subsection{Acknowledgements}\label{acknowledgements}}
+\hypertarget{acknowledgments}{%
+\subsection{Acknowledgments}\label{acknowledgments}}
 
@@
 
+\leavevmode\hypertarget{ref-Armstrong2000}{}%
+Armstrong, E., Rogerson, A., and Leftley, J.W. (2000) The Abundance of
+Heterotrophic Protists Associated with Intertidal Seaweeds.
+\emph{Estuarine, Coastal and Shelf Science} \textbf{50}: 415--424.
+
+\leavevmode\hypertarget{ref-Bengtsson2010}{}%
+Bengtsson, M., Sjøtun, K., and Øvreås, L. (2010) Seasonal dynamics of
+bacterial biofilms on the kelp Laminaria hyperborea. \emph{Aquatic
+Microbial Ecology} \textbf{60}: 71--83.
+
 \leavevmode\hypertarget{ref-Burke2011a}{}%
 Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
-(2011) Bacterial community assembly based on functional genes rather
+(2011a) Bacterial community assembly based on functional genes rather
 than species. \emph{Proceedings of the National Academy of Sciences of
 the United States of America} \textbf{108}: 14288--14293.
+
+\leavevmode\hypertarget{ref-Burke2011b}{}%
+Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
+(2011b) Composition, uniqueness and variability of the epiphytic
+bacterial community of the green alga Ulva australis. \emph{The ISME
+journal} \textbf{5}: 590--600.
 
@@
 
-\leavevmode\hypertarget{ref-Chourey2010}{}%
-Chourey, K., Jansson, J., VerBerkmoes, N., Shah, M., Chavarria, K.L.,
-Tom, L.M., et al. (2010) Direct Cellular Lysis/Protein Extraction
-Protocol for Soil Metaproteomics. \emph{Journal of Proteome Research}
-\textbf{9}: 6615--6622.
+\leavevmode\hypertarget{ref-Crump2008}{}%
+Crump, B.C. and Koch, E.W. (2008) Attached bacterial populations shared
+by four species of aquatic angiosperms. \emph{Applied and environmental
+microbiology} \textbf{74}: 5948--5957.
 
-\leavevmode\hypertarget{ref-Hultman2015}{}%
-Hultman, J., Waldrop, M.P., Mackelprang, R., David, M.M., McFarland, J.,
-Blazewicz, S.J., et al. (2015) Multi-omics of permafrost, active layer
-and thermokarst bog soil microbiomes. \emph{Nature} \textbf{521}:
-208--212.
+\leavevmode\hypertarget{ref-Egan2013}{}%
+Egan, S., Harder, T., Burke, C., Steinberg, P., Kjelleberg, S., and
+Thomas, T. (2013) The seaweed holobiont: understanding seaweed-bacteria
+interactions. \emph{FEMS microbiology reviews} \textbf{37}: 462--476.
+
+\leavevmode\hypertarget{ref-Hollants2013}{}%
+Hollants, J., Leliaert, F., De Clerck, O., and Willems, A. (2013) What
+we can learn from sushi: A review on seaweed-bacterial associations.
+\textbf{83}: 1--16.
+
+\leavevmode\hypertarget{ref-Korlevic_in_press}{}%
+Korlević, M., Markovski, M., Zhao, Z., Herndl, G.J., and Najdek, M.
+Selective DNA and Protein Isolation from Marine Macrophyte Surfaces.
 
@@
 
-\leavevmode\hypertarget{ref-Ludwig2004}{}%
-Ludwig, W., Strunk, O., Westram, R., Richter, L., Meier, H., Yadhukumar,
-et al. (2004) ARB: a software environment for sequence data.
-\emph{Nucleic Acids Research} \textbf{32}: 1363--1371.
+\leavevmode\hypertarget{ref-Lachnit2009}{}%
+Lachnit, T., Blümel, M., Imhoff, J.F., and Wahl, M. (2009) Specific
+epibacterial communities on macroalgae: Phylogeny matters more than
+habitat. \emph{Aquatic Biology} \textbf{5}: 181--186.
+
+\leavevmode\hypertarget{ref-Lachnit2011}{}%
+Lachnit, T., Meske, D., Wahl, M., Harder, T., and Schmitz, R. (2011)
+Epibacterial community patterns on marine macroalgae are host-specific
+but temporally variable. \emph{Environmental Microbiology} \textbf{13}:
+655--665.
+
+\leavevmode\hypertarget{ref-Margulis1991}{}%
+Margulis, L. (1991) Symbiogenesis and symbionticism. In, Margulis,L. and
+Fester,R. (eds), \emph{Symbiosis as a source of evolutionary
+innovation}. Cambridge, Massachusetts: The MIT Press, pp. 1--14.
 
@@
 
+\leavevmode\hypertarget{ref-Miranda2013}{}%
+Miranda, L.N., Hutchison, K., Grossman, A.R., and Brawley, S.H. (2013)
+Diversity and Abundance of the Bacterial Community of the Red Macroalga
+Porphyra umbilicalis: Did Bacterial Farmers Produce Macroalgae?
+\emph{PLoS ONE} \textbf{8}: e58269.
+
+\leavevmode\hypertarget{ref-Morrissey2019}{}%
+Morrissey, K.L., Çavas, L., Willems, A., and De Clerck, O. (2019)
+Disentangling the influence of environment, host specificity and thallus
+differentiation on bacterial communities in siphonous green seaweeds.
+\emph{Frontiers in Microbiology} \textbf{10}:
+
+\leavevmode\hypertarget{ref-Neuwirth2014}{}%
+Neuwirth, E. (2014) RColorBrewer: ColorBrewer Palettes.
+
+\leavevmode\hypertarget{ref-Oksanen2019}{}%
+Oksanen, J., Blanchet, F.G., Friendly, M., Kindt, R., Legendre, P.,
+McGlinn, D., et al. (2019) vegan: Community Ecology Package.
+
 \leavevmode\hypertarget{ref-Parada2016}{}%
@@
 \emph{Environmental Microbiology} \textbf{18}: 1403--1414.
-
-\leavevmode\hypertarget{ref-Pruesse2012}{}%
-Pruesse, E., Peplies, J., and Glöckner, F.O. (2012) SINA: accurate
-high-throughput multiple sequence alignment of ribosomal RNA genes.
-\emph{Bioinformatics (Oxford, England)} \textbf{28}: 1823--1829.
 
@@
 Computing, Vienna, Austria: R Foundation for Statistical Computing.
+
+\leavevmode\hypertarget{ref-Roth-Schulze2016}{}%
+Roth-Schulze, A.J., Zozaya-Valdés, E., Steinberg, P.D., and Thomas, T.
+(2016) Partitioning of functional and taxonomic diversity in
+surface-associated microbial communities. \emph{Environmental
+microbiology} \textbf{18}: 4391--4402.
 
@@
 
-\leavevmode\hypertarget{ref-Stamatakis2006}{}%
-Stamatakis, A. (2006) RAxML-VI-HPC: maximum likelihood-based
-phylogenetic analyses with thousands of taxa and mixed models.
-\emph{Bioinformatics} \textbf{22}: 2688--2690.
+\leavevmode\hypertarget{ref-Tarquinio2019}{}%
+Tarquinio, F., Hyndes, G.A., Laverock, B., Koenders, A., and Säwström,
+C. (2019) The seagrass holobiont: Understanding seagrass-bacteria
+interactions and their role in seagrass ecosystem functioning.
+\emph{FEMS Microbiology Letters} \textbf{366}:
+
+\leavevmode\hypertarget{ref-Tujula2010}{}%
+Tujula, N.A., Crocetti, G.R., Burke, C., Thomas, T., Holmström, C., and
+Kjelleberg, S. (2010) Variability and abundance of the epiphytic
+bacterial community associated with a green marine Ulvacean alga.
+\emph{The ISME Journal} \textbf{4}: 301--311.
+
+\leavevmode\hypertarget{ref-Ugarelli2019}{}%
+Ugarelli, K., Laas, P., and Stingl, U. (2019) The microbial communities
+of leaves and roots associated with turtle grass (Thalassia testudinum)
+and manatee grass (syringodium filliforme) are distinct from seawater
+and sediment communities, but are similar between species and sampling
+sites. \emph{Microorganisms} \textbf{7}:
 
@@
 Software} \textbf{4}: 1686.
-
-\leavevmode\hypertarget{ref-Wisniewski2009}{}%
-Wiśniewski, J.R., Zougman, A., Nagaraj, N., and Mann, M. (2009)
-Universal sample preparation method for proteome analysis. \emph{Nature
-Methods} \textbf{6}: 359--362.
 
@@
 \leavevmode\hypertarget{ref-Xie2019}{}%
-Xie, Y. (2019b) tinytex: Helper Functions to Install and Maintain 'TeX
+Xie, Y. (2019b) TinyTeX: A lightweight, cross-platform, and
+easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
+30--32.
+
+\leavevmode\hypertarget{ref-Xie2020}{}%
+Xie, Y. (2020) tinytex: Helper Functions to Install and Maintain 'TeX
 Live', and Compile 'LaTeX' Documents.
@@
 
+\textbf{\autoref{shared}.} \nameref{shared}
+
+\textbf{\autoref{pcoa}.} \nameref{pcoa}
+
 \textbf{\autoref{community}.} \nameref{community}
 
-\newpage
+\textbf{\autoref{cyano}.} \nameref{cyano}
+
+\textbf{\autoref{bactero}.} \nameref{bactero}
+
+\textbf{\autoref{alpha}.} \nameref{alpha}
+
+\textbf{\autoref{gamma}.} \nameref{gamma}
+
+\textbf{\autoref{desulfo}.} \nameref{desulfo}
 
@@
 
-\newpage
-\begin{figure}[ht]
+\begin{figure}[H]
 
-{\centering \includegraphics[width=1\linewidth]{../results/figures/community_barplot_phylum}
+{\centering \includegraphics[width=0.85\linewidth]{../results/figures/seasonal_shared}
 
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant bacterial sequences.\label{community}}\label{fig:unnamed-chunk-2}
+\caption{Proprotion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{shared}}\label{fig:unnamed-chunk-2}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\linewidth]{../results/figures/pcoa_figure}
+
+}
+
+\caption{Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+Samples from the same environment or same season are labeld in different colors.
+The proportion of explained variation by each axis is shown on the corresponding axis in parentheses.\label{pcoa}}\label{fig:unnamed-chunk-3}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.85\linewidth]{../results/figures/community_bar_plot}
+
+}
+
+\caption{Taxonomic classification and relative contribution of the most abundant bacterial and archaeal sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{community}}\label{fig:unnamed-chunk-4}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.85\linewidth]{../results/figures/cyanobacteria_bar_plot}
+
+}
+
+\caption{Taxonomic classification and relative contribution of the most abundant cyanobacterial sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+The proprotion of cyanobaterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
+NR -- No Relative\label{cyano}}\label{fig:unnamed-chunk-5}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.85\linewidth]{../results/figures/bacteroidota_bar_plot}
+
+}
+
+\caption{Taxonomic classification and relative contribution of the most abundant sequences within the \textit{Bacteroidota} on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+The proprotion of sequences classified as \textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar.
+NR -- No Relative\label{bactero}}\label{fig:unnamed-chunk-6}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.85\linewidth]{../results/figures/alphaproteobacteria_bar_plot}
+
+}
+
+\caption{Taxonomic classification and relative contribution of the most abundant alphaproteobacterial sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+The proprotion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
+NR -- No Relative\label{alpha}}\label{fig:unnamed-chunk-7}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.85\linewidth]{../results/figures/gammaproteobacteria_bar_plot}
+
+}
+
+\caption{Taxonomic classification and relative contribution of the most abundant gammaproteobacterial sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+The proprotion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
+NR -- No Relative\label{gamma}}\label{fig:unnamed-chunk-8}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.85\linewidth]{../results/figures/desulfobacterota_bar_plot}
+
+}
+
+\caption{Taxonomic classification and relative contribution of the most abundant sequences within the \textit{Desulfobacterota} on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+The proprotion of sequences classified as \textit{Desulfobacterota} in the total bacterial and archaeal community is given above the corresponding bar.
+NR -- No Relative\label{desulfo}}\label{fig:unnamed-chunk-9}
 \end{figure}

### 2020-02-14 submission/supplementary.Rmd
Added introduction and supp informations.

---
title: "**Supplementary Information**"
subtitle: "**Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces**"
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: header_supplementary.tex
fontsize: 12pt
geometry: margin=1.0in
csl: citation_style.csl #Get themes at https://github.com/citation-style-language/styles
bibliography: references.bib
---

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE}
opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)
opts_chunk$set("message" = FALSE)

inline_hook <- function(x){
	print(x)

	if(is.list(x)){
		x <- unlist(x)
	}

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=1, nsmall=1, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
knitr::knit_hooks$set(inline=inline_hook)
```

\vspace{70mm}

^1$\dagger$^

\vspace{40mm}

$\dagger$ To whom correspondence should be addressed: marino.korlevic@irb.hr

1\. Ruđer Bošković Institute, Center for Marine Research, G. Paliaga 5, Rovinj, Croatia

2\. University of Vienna, Department of Limnology and Bio-Oceanography, Althanstraße 14, Vienna, Austria

\linenumbers
\sisetup{mode=text}
\setlength\parindent{24pt}

## Supplementary Figures
```{r out.width="85%", fig.align="center", fig.cap="Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\\label{rarefaction}", fig.pos="H"}
include_graphics("../results/figures/rarefaction.jpg")
```

```{r out.width="85%", fig.align="center", fig.cap="Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\\label{calculators}", fig.pos="H"}
include_graphics("../results/figures/calculators.jpg")
```

```{r out.width="85%", fig.align="center", fig.cap="Relative contribution of chloroplast sequences on the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\\label{chloroplast}", fig.pos="H"}
include_graphics("../results/figures/chloroplast_bar_plot.jpg")
```

### 2020-02-14 submission/supplementary.tex
Added introduction and supp informations.

--- 2019-12-18 submission/manuscript.tex
@@
 \hypersetup{unicode=true,
-            pdftitle={Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces},
+            pdftitle={Supplementary Information},
             pdfborder={0 0 0},
@@
 
-  \title{\textbf{Dynamics of Epiphytic Microbial Communities on Marine Macrophyte
-Surfaces}}
+  \title{\textbf{Supplementary Information}}
     \pretitle{\vspace{\droptitle}\centering\huge}
   \posttitle{\par}
+  \subtitle{\textbf{Dynamics of Epiphytic Microbial Communities on Marine Macrophyte
+Surfaces}}
   \author{}
@@
 \usepackage[labelsep=period, labelfont=bf]{caption}
-\renewcommand{\thefigure}{\arabic{figure}}
+\renewcommand{\thefigure}{S\arabic{figure}}
 \captionsetup{justification=raggedright,singlelinecheck=false}
@@
 
+\usepackage{float}
+
 \begin{document}
@@
 2. University of Vienna, Department of Limnology and Bio-Oceanography,
-Althanstraße 14, Vienna, Austria \newpage \linenumbers
-\sisetup{mode=text} \setlength\parindent{24pt}
+Althanstraße 14, Vienna, Austria
 
-\hypertarget{abstract}{%
-\subsection{Abstract}\label{abstract}}
+\linenumbers
+\sisetup{mode=text}
+\setlength\parindent{24pt}
 
-\newpage
+\hypertarget{supplementary-figures}{%
+\subsection{Supplementary Figures}\label{supplementary-figures}}
 
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+\begin{figure}[H]
 
-\newpage
-
-\hypertarget{materials-and-methods}{%
-\subsection{Materials and Methods}\label{materials-and-methods}}
-
-\hypertarget{sampling}{%
-\subsubsection{Sampling}\label{sampling}}
-
-Leaves of \emph{Cymodocea nodosa} were sampled in a \emph{Cymodocea
-nodosa} meadow in the Bay of Saline (45°7´5˝N, 13°37´20˝E) and in a
-\emph{Cymodocea nodosa} meadow invaded by \emph{Caulerpa cylindracea} in
-the proximity of the village of Funtana (45°10´39˝N, 13°35´42˝E). Thalli
-of \emph{Caulerpa cylindracea} were sampled in the same \emph{Cymodocea
-nodosa} invaded meadow in Funtana and on a locality of only
-\emph{Caulerpa cylindracea} located in the proximity of the invaded
-meadow. Leaves and thalli were collected on the same day in two
-contrasting seasons, on 4 December 2017 and 18 June 2018. During spring
-2018 the \emph{Cymodocea nodosa} meadow in the Bay of Saline decayed to
-an extent that no leaves could be retrieved (Najdek \emph{et al.},
-unpublished data). Leaves and thalli were collected by diving and
-transported to the laboratory in containers placed on ice and filled
-with site seawater. Upon arrival to the laboratory, \emph{Cymodocea
-nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while
-\emph{Caulerpa cylindracea} thalli were cut into 5 -- 8 \si{\cm} long
-sections. Leaves and thalli were washed three times with sterile
-artificial seawater (ASW) to remove loosely attached microbial cells.
-
-\hypertarget{dna-isolation}{%
-\subsubsection{DNA Isolation}\label{dna-isolation}}
-
-The DNA was isolated according to the protocol for isolation from
-filters described in Massana \emph{et al.} (1997). This protocol was
-modified and adapted for DNA isolation from microbial communities from
-macrophyte surfaces as described below. 5 \si{\ml} of lysis buffer (40
-\si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar}
-sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or \si{\g}
-wet-weight of thalli. Lysozyme was added (final concentration 1
-\si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius}
-for 30 \si{\minute}. Subsequently, proteinase K (final concentration 0.5
-\si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were
-added and the samples were incubated at 55 \si{\degreeCelsius} for 2
-\si{\hour}. Following the incubation, tubes were vortexed for 10
-\si{\minute} and the mixture containing lyzed epiphytic cells was
-separated from host leaves or thalli by transferring the solution into a
-clean tube. The lysate was extracted twice with a mixture of
-phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with
-chloroform:isoamyl alcohol (24:1). After each organic solvent mixture
-addition tubes were slightly vortexed and centrifuged at 4,500 × g for
-10 \si{\minute}. Following each centrifugation aqueous phases were
-retrieved. After the final extraction 1/10 of chilled 3 \si{\Molar}
-sodium acetate (pH 5.2) was added. DNA was precipitated by adding 1
-volume of chilled isopropanol, incubating the mixtures overnight at
-\num{-20} °C and centrifuging at 16,000 × g and 4 \si{\degreeCelsius}
-for 20 \si{\minute}. The pellet was washed twice with 1 \si{\ml} of
-chilled isopropanol and centrifuged after each washing step at 20,000 ×
-g and 4 \si{\degreeCelsius} for 10 \si{\minute}. After the first washing
-step duplicate pellets form the same sample were pooled and transferred
-to a clean 1.5 \si{\ml} tube. The dried pellet was resuspended in 100
-\si{\ul} of deionized water.
-
-\hypertarget{illumina-16s-rrna-sequencing}{%
-\subsubsection{Illumina 16S rRNA
-Sequencing}\label{illumina-16s-rrna-sequencing}}
-
-An aliquot of isolated DNA was treated with RNase A (final concentration
-200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA
-concentration was determined using the Quant-iT PicoGreen dsDNA Assay
-Kit (Invitrogen, USA) according to the manufacturer's instructions and
-diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was
-amplified using a two-step PCR procedure. In the first PCR the 515F
-(5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3')
-primers from the Earth Microbiome Project
-(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
-were used to amplify the target region (Caporaso \emph{et al.}, 2012;
-Apprill \emph{et al.}, 2015; Parada \emph{et al.}, 2016). These primers
-contained on their 5' end a tagged sequence. Each sample was amplified
-in four parallel 25 \si{\ul} reactions of which each contained: 1 × Q5
-Reaction Buffer , 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml}
-BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse
-primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs,
-USA) and 5 \si{\ng} of DNA template. Cycling conditions were: initial
-denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of
-denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50
-\si{\degreeCelsius} for 60 \si{\s} and elongation at 72
-\si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at
-72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions
-volumes were pooled and PCR products were purified using the GeneJET PCR
-Purification Kit (Thermo Fisher Scientific, USA) according to the
-manufacturer's instructions and following the protocol that included
-isopropanol addition for better small DNA fragment yield. The column was
-eluted in 30 \si{\ul} of deionized water. Purified PCR products were
-sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories,
-Martinsried, Germany. Before sequencing, at IMGM the second PCR
-amplification of the two-step PCR procedure was performed using primers
-targeting the tagged region incorporated in the first PCR. In addition,
-these primers contained adapter and sample-specific index sequences. The
-second PCR was carried out for 8 cycles. Beside samples, a positive and
-negative control were sequenced. A negative control was comprised of
-four parallel PCR reactions without DNA template, while for a positive
-control a mock community composed of evenly mixed DNA material
-originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
-used. The sequences obtained in this study have been submitted to the
-European Nucleotide Archive (ENA) under accession numbers \textbf{TO BE
-ADDED LATER!}.
-
-\hypertarget{sequence-analysis}{%
-\subsubsection{Sequence Analysis}\label{sequence-analysis}}
-
-Obtained sequences were analyzed on the computer cluster Isabella
-(University Computing Center, University of Zagreb) using mothur
-(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
-Standard Operating Procedure (MiSeq SOP;
-\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
-and recommendations given from the Riffomonas project to enhance data
-reproducibility (\url{http://www.riffomonas.org/}). For alignment and
-classification of sequences the SILVA SSU Ref NR 99 database (release
-132; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
-Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
-mothur were exported, aligned using SILVA incremental aligner (SINA,
-version 1.6.0) (Pruesse \emph{et al.}, 2012) against the same SILVA SSU
-Ref NR 99 database (release 132) and imported into ARB (version 6.0.6)
-(Ludwig \emph{et al.}, 2004) for further phylogenetic analysis using the
-same database. Reference sequences close to imported ones were selected
-and used to calculate a phylogenetic tree using the Maximum Likelihood
-algorithm RAxML (version 7.0.3) with 1000 tree replicates (Stamatakis,
-2006). Imported partial chloroplast sequences were added to the tree
-using the maximum parsimony criteria and not allowing changes to tree
-topology. Pipeline data processing and visualization was done using R
-(version 3.6.0) (R Core Team, 2019), package tidyverse (version 1.2.1)
-(Wickham \emph{et al.}, 2019) and multiple other packages (Xie, 2014,
-2015, 2019a, 2019b; Xie \emph{et al.}, 2018; Zhu, 2019; Allaire \emph{et
-al.}, 2019). The detailed analysis procedure including the R Markdown
-file for this paper are available as a GitHub repository (\textbf{TO BE
-ADDED LATER!}). Based on the ATCC MSA-1002 mock community included in
-the analysis a sequencing error rate of 0.01 \si{\percent} was
-determined, which is in line with previously reported values for
-next-generation sequencing data (Kozich \emph{et al.}, 2013; Schloss
-\emph{et al.}, 2016). In addition, the negative control processed
-together with the samples yielded only 2 sequences after sequence
-quality curation.
-
-\hypertarget{protein-isolation}{%
-\subsubsection{Protein Isolation}\label{protein-isolation}}
-
-Proteins were isolated according to the protocol for isolation from soil
-described in Chourey \emph{et al.} (2010) and modified by Hultman
-\emph{et al.} (2015). These protocols were further modified and adapted
-for protein isolation from microbial communities form macrophyte
-surfaces as described below. 20 \si{\ml} of protein extraction buffer (4
-\si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl {[}pH 8.0{]}) was
-added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of
-thalli. The mixture was incubated in boiling water for 5 \si{\minute},
-vortexed for 10 \si{\minute} and incubated again in boiling water for 5
-\si{\minute}. After a brief vortex the lysate was transferred to a clean
-tube separating the host leaves or thalli from the mixture containing
-lyzed epiphytic cells. Dithiothreitol (DTT; final concentration 24
-\si{\milli\Molar}) was added and proteins were precipitated with chilled
-100 \si{\percent} trichloroacetic acid (TCA; final concentration 20
-\si{\percent}) overnight at \num{-20} °C. Precipitated proteins were
-centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40 \si{\minute}.
-The obtained protein pellet was washed three times with chilled acetone.
-During the first washing step the pellet was transferred to a clean 1.5
-\si{\ml} tube. After each washing step samples were centrifuged at
-20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets
-were stored at \num{-80} °C until further analysis.
-
-\hypertarget{metaproteomics}{%
-\subsubsection{Metaproteomics}\label{metaproteomics}}
-
-Isolated proteins were whole trypsin digested using the FASP
-(filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK)
-according to the manufacturer's instructions (Wiśniewski \emph{et al.},
-2009) with small modifications. Before loading the solution to the
-column, protein pellets were solubilized in a urea sample buffer
-included in the kit amended with DTT (final concentration 100
-\si{\milli\Molar}) for 45 \si{\minute} at room temperature and
-centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to
-remove larger particles. The first washing step after protein solution
-loading was repeated twice. In addition, centrifugation steps were
-prolonged if the column was clogged. Trypsin digestion was performed on
-column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
-The final filtrate containing digested proteins was acidified with 1
-\si{\percent} (final concentration) trifluoroacetic acid, freezed at
-\num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent
-to the Vienna Metabolomics Center (University of Vienna) for
-metaproteomic analysis. Peptides were resuspended in 1 \si{\percent}
-(final concentration) trifluoroacetic acid, desalted using the Pierce
-C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's
-instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap
-Mass Spectrometer (Thermo Fisher Scientific, USA). Obtained MS/MS
-spectra were searched against a protein database from metagenomic
-assembly published in Burke \emph{et al.} (2011) using SEQUEST-HIT
-engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo
-Fisher Scientific, USA). The target-decoy approach was used to reduce
-the probability of false peptide identification. Results whose false
-discovery rate at the peptide level was \SI{< 1}{\percent} were kept.
-For protein identification a minimum of of two peptides and one unique
-peptide were required. For protein quantification, a chromatographic
-peak area-based free quantitative method was applied.
-
-\hypertarget{confocal-microscopy}{%
-\subsubsection{Confocal Microscopy}\label{confocal-microscopy}}
-
-Host leaves and thalli from DNA and protein isolation steps were washed
-seven times in deionized water and fixed with formaldehyde (final
-concentration \textasciitilde{} 3 \si{\percent}). In addition,
-nontreated leaves and thalli, washed three times in ASW to remove
-loosely attached microbial cells, were fixed in the same concentration
-of formaldehyde and used as a positive control. For long therm storage,
-fixed leaves and thalli were immersed in a mixture of phosphate-buffered
-saline (PBS) and ethanol (1:1) and stored at \num{-20}
-\si{\degreeCelsius}. Treated and untreated leaves and thalli segments
-were stained in a 2 × solution of SYBR Green I and examined under a
-Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
-
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
-
-\hypertarget{discussion}{%
-\subsection{Discussion}\label{discussion}}
-
-\hypertarget{acknowledgements}{%
-\subsection{Acknowledgements}\label{acknowledgements}}
-
-\newpage
-
-\hypertarget{references}{%
-\subsection{References}\label{references}}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Allaire2019}{}%
-Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
-A., et al. (2019) rmarkdown: Dynamic Documents for R.
-
-\leavevmode\hypertarget{ref-Apprill2015}{}%
-Apprill, A., McNally, S., Parsons, R., and Weber, L. (2015) Minor
-revision to V4 region SSU rRNA 806R gene primer greatly increases
-detection of SAR11 bacterioplankton. \emph{Aquatic Microbial Ecology}
-\textbf{75}: 129--137.
-
-\leavevmode\hypertarget{ref-Burke2011a}{}%
-Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
-(2011) Bacterial community assembly based on functional genes rather
-than species. \emph{Proceedings of the National Academy of Sciences of
-the United States of America} \textbf{108}: 14288--14293.
-
-\leavevmode\hypertarget{ref-Caporaso2012}{}%
-Caporaso, J.G., Lauber, C.L., Walters, W.A., Berg-Lyons, D., Huntley,
-J., Fierer, N., et al. (2012) Ultra-high-throughput microbial community
-analysis on the Illumina HiSeq and MiSeq platforms. \emph{The ISME
-Journal} \textbf{6}: 1621--1624.
-
-\leavevmode\hypertarget{ref-Chourey2010}{}%
-Chourey, K., Jansson, J., VerBerkmoes, N., Shah, M., Chavarria, K.L.,
-Tom, L.M., et al. (2010) Direct Cellular Lysis/Protein Extraction
-Protocol for Soil Metaproteomics. \emph{Journal of Proteome Research}
-\textbf{9}: 6615--6622.
-
-\leavevmode\hypertarget{ref-Hultman2015}{}%
-Hultman, J., Waldrop, M.P., Mackelprang, R., David, M.M., McFarland, J.,
-Blazewicz, S.J., et al. (2015) Multi-omics of permafrost, active layer
-and thermokarst bog soil microbiomes. \emph{Nature} \textbf{521}:
-208--212.
-
-\leavevmode\hypertarget{ref-Kozich2013}{}%
-Kozich, J.J., Westcott, S.L., Baxter, N.T., Highlander, S.K., and
-Schloss, P.D. (2013) Development of a dual-index sequencing strategy and
-curation pipeline for analyzing amplicon sequence data on the MiSeq
-Illumina sequencing platform. \emph{Applied and environmental
-microbiology} \textbf{79}: 5112--5120.
-
-\leavevmode\hypertarget{ref-Ludwig2004}{}%
-Ludwig, W., Strunk, O., Westram, R., Richter, L., Meier, H., Yadhukumar,
-et al. (2004) ARB: a software environment for sequence data.
-\emph{Nucleic Acids Research} \textbf{32}: 1363--1371.
-
-\leavevmode\hypertarget{ref-Massana1997}{}%
-Massana, R., Murray, A.E., Preston, C.M., and DeLong, E.F. (1997)
-Vertical distribution and phylogenetic characterization of marine
-planktonic Archaea in the Santa Barbara Channel. \emph{Applied and
-environmental microbiology} \textbf{63}: 50--56.
-
-\leavevmode\hypertarget{ref-Parada2016}{}%
-Parada, A.E., Needham, D.M., and Fuhrman, J.A. (2016) Every base
-matters: assessing small subunit rRNA primers for marine microbiomes
-with mock communities, time series and global field samples.
-\emph{Environmental Microbiology} \textbf{18}: 1403--1414.
-
-\leavevmode\hypertarget{ref-Pruesse2012}{}%
-Pruesse, E., Peplies, J., and Glöckner, F.O. (2012) SINA: accurate
-high-throughput multiple sequence alignment of ribosomal RNA genes.
-\emph{Bioinformatics (Oxford, England)} \textbf{28}: 1823--1829.
-
-\leavevmode\hypertarget{ref-Quast2013}{}%
-Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
-et al. (2013) The SILVA ribosomal RNA gene database project: improved
-data processing and web-based tools. \emph{Nucleic acids research}
-\textbf{41}: D590--6.
-
-\leavevmode\hypertarget{ref-RCoreTeam2019}{}%
-R Core Team (2019) R: A Language and Environment for Statistical
-Computing, Vienna, Austria: R Foundation for Statistical Computing.
-
-\leavevmode\hypertarget{ref-Schloss2016}{}%
-Schloss, P.D., Girard, R.A., Martin, T., Edwards, J., and Thrash, J.C.
-(2016) Status of the Archaeal and Bacterial Census: an Update.
-\emph{mBio} \textbf{7}: e00201--16.
-
-\leavevmode\hypertarget{ref-Schloss2009}{}%
-Schloss, P.D., Westcott, S.L., Ryabin, T., Hall, J.R., Hartmann, M.,
-Hollister, E.B., et al. (2009) Introducing mothur: open-source,
-platform-independent, community-supported software for describing and
-comparing microbial communities. \emph{Applied and environmental
-microbiology} \textbf{75}: 7537--7541.
-
-\leavevmode\hypertarget{ref-Stamatakis2006}{}%
-Stamatakis, A. (2006) RAxML-VI-HPC: maximum likelihood-based
-phylogenetic analyses with thousands of taxa and mixed models.
-\emph{Bioinformatics} \textbf{22}: 2688--2690.
-
-\leavevmode\hypertarget{ref-Wickham2019}{}%
-Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.D., François,
-R., et al. (2019) Welcome to the tidyverse. \emph{Journal of Open Source
-Software} \textbf{4}: 1686.
-
-\leavevmode\hypertarget{ref-Wisniewski2009}{}%
-Wiśniewski, J.R., Zougman, A., Nagaraj, N., and Mann, M. (2009)
-Universal sample preparation method for proteome analysis. \emph{Nature
-Methods} \textbf{6}: 359--362.
-
-\leavevmode\hypertarget{ref-Xie2015}{}%
-Xie, Y. (2015) Dynamic Documents with \{R\} and knitr, 2nd ed. Boca
-Raton, Florida: Chapman; Hall/CRC.
-
-\leavevmode\hypertarget{ref-Xie2014}{}%
-Xie, Y. (2014) knitr: A Comprehensive Tool for Reproducible Research in
-\{R\}. In, Stodden,V., Leisch,F., and Peng,R.D. (eds),
-\emph{Implementing reproducible computational research}. Chapman;
-Hall/CRC.
-
-\leavevmode\hypertarget{ref-Xie2019a}{}%
-Xie, Y. (2019a) knitr: A General-Purpose Package for Dynamic Report
-Generation in R.
-
-\leavevmode\hypertarget{ref-Xie2019}{}%
-Xie, Y. (2019b) tinytex: Helper Functions to Install and Maintain 'TeX
-Live', and Compile 'LaTeX' Documents.
-
-\leavevmode\hypertarget{ref-Xie2018}{}%
-Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: The
-Definitive Guide, Boca Raton, Florida: Chapman; Hall/CRC.
-
-\leavevmode\hypertarget{ref-Yilmaz2014}{}%
-Yilmaz, P., Parfrey, L.W., Yarza, P., Gerken, J., Pruesse, E., Quast,
-C., et al. (2014) The SILVA and "All-species Living Tree Project (LTP)"
-taxonomic frameworks. \emph{Nucleic acids research} \textbf{42}:
-D643--8.
-
-\leavevmode\hypertarget{ref-Zhu2019}{}%
-Zhu, H. (2019) kableExtra: Construct Complex Table with 'kable' and Pipe
-Syntax.
-
-\newpage 
-\setlength\parindent{0pt}
-
-\hypertarget{figure-captions}{%
-\subsection{Figure Captions}\label{figure-captions}}
-
-\textbf{\autoref{community}.} \nameref{community}
-
-\newpage
-
-\hypertarget{figures}{%
-\subsection{Figures}\label{figures}}
-
-\newpage
-\begin{figure}[ht]
-
-{\centering \includegraphics[width=1\linewidth]{../results/figures/community_barplot_phylum}
+{\centering \includegraphics[width=0.85\linewidth]{../results/figures/rarefaction}
 
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant bacterial sequences.\label{community}}\label{fig:unnamed-chunk-2}
+\caption{Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{rarefaction}}\label{fig:unnamed-chunk-1}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.85\linewidth]{../results/figures/calculators}
+
+}
+
+\caption{Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{calculators}}\label{fig:unnamed-chunk-2}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.85\linewidth]{../results/figures/chloroplast_bar_plot}
+
+}
+
+\caption{Relative contribution of chloroplast sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{chloroplast}}\label{fig:unnamed-chunk-3}
 \end{figure}

### 2020-02-14 code/plot_pcoa.R

@@ -41,8 +41,8 @@ theme <- theme(text=element_text(family="Times"), line=element_line(color="black
          legend.justification=c("left", "top"), plot.title=element_text(size=14, hjust=0.5))
 
 # Plots generation
-# All samlples
-# Filter all sampels
+# All samples
+# Filter all samples
 rarefied_metadata_select <- rarefied_metadata
 
 # Generation of PCoA data

### 2020-02-19 submission/manuscript.Rmd
Intro done, matrix built.

@@
 ---
-title: "**Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces**"
+title: "**Seasonal Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces**"
 output:
@@
 opts_chunk$set("warning" = FALSE)
-opts_chunk$set("cache" = FALSE)
+opts_chunk$set("cache" = T)
 opts_chunk$set("message" = FALSE)
@@
 ## Results
-Chloroplast sequences were following the seasonal pattern (\autoref{supp-chloroplast}, \autoref{pcoa}).
+```{r community}
+# Loading input data containing sequence abundances and subsequent input data customization
+community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
+  filter(!str_detect(taxon, "^Eukaryota")) %>%
+  filter(taxon!="Root")
+
+# Remove Chloroplast and Mitochondria sequences and subtract their number from higher taxonomic levels to which
+# they belong
+chloroplast <- filter(community, str_detect(taxon, "^Chloroplast$"))$rankID
+mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
+
+community <- mutate_at(community, 5:ncol(community), list(~case_when(
+    rankID==str_extract(chloroplast, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Chloroplast"],
+    rankID==str_extract(chloroplast, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Chloroplast"],
+    rankID==str_extract(chloroplast, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Chloroplast"],
+  TRUE ~ .))) %>%
+  filter(!str_detect(taxon, "^Chloroplast")) %>%
+  mutate_at(5:ncol(.), list(~case_when(
+    rankID==str_extract(mitochondria, "(\\d+\\.){4}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Mitochondria"],
+    TRUE ~ .))) %>%
+  filter(!str_detect(taxon, "^Mitochondria")) %>%
+  select(-ATCC_1, -ATCC_5, -NC_1)
+
+million_sequences <- filter(community, taxon=="Archaea" | taxon=="Bacteria") %>%
+  select(total) %>%
+  sum()/1e6
+
+# Loading metadata
+metadata <- read_tsv("data/raw/metadata.csv")
+
+samples_epiphytes <- filter(metadata, !is.na(station) & station!="F") %>%
+  summarise(n=n())
+samples_seawater <- filter(metadata, !is.na(station) & station=="F") %>%
+  summarise(n=n())
+
+# Minimal and maximal number of sequences
+min_max <- filter(community, taxon=="Archaea" | taxon=="Bacteria") %>%
+  summarise_at(6:ncol(.), sum) %>%
+  summarise(min=min(.), max=max(.))
+```
+
+```{r shared}
+# Loading OTU/sample data
+shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")
+
+no_otus <- select(shared, starts_with("Otu")) %>%
+  colnames() %>%
+  length()
+
+# Generation of random rarefied community data
+rarefied <- shared %>%
+  select(-label, -Group, -numOtus) %>%
+  rrarefy(., min(rowSums(.))) %>%
+  as.tibble() %>%
+  add_column("Group"=shared$Group, .before=TRUE) %>%
+  select_if(funs(!is.numeric(.) || sum(.)!=0))
+
+no_otus_rarefied <- select(rarefied, starts_with("Otu")) %>%
+  colnames() %>%
+  length()
+
+otu_min_max <- select(rarefied, starts_with("Otu")) %>%
+  decostand(., method="pa") %>%
+  transmute(sum=rowSums(.))
+```
+Sequencing of the 16S rRNA V4 region yielded a total of `r million_sequences` million sequences after sequence quality curation and exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences.
+A total of `r samples_epiphytes` samples originating from epiphytic archaeal and bacterial communities associated with surfaces of the seagrass *Cymodocea nodosa* and macroalga *Caulerpa cylindracea* were analyzed.
+In addition, `r samples_seawater` samples (one of the samples was sequenced two times) originating from picoplankton archaeal and bacterial communities in the surrounding seawater were also processed.
+The number of reads per sample ranged between `r min_max$min` and `r min_max$max` sequences.
+Even when the highest sequencing effort was applied the rarefaction curves did not level off that is a common observation in high-throughput 16S rRNA amplicon sequencing approaches (\autoref{supp-rarefaction}).
+Following quality curation and exclusion of sequences mentioned before reads were clustered into `r no_otus` different OTUs at a similarity level of 97 \si{\percent}.
+Reads numbers were normalized to the minimum number of sequences, `r min_max$min`, through rarefaction resulting in `r no_otus_rarefied` different OTUs that ranged from `r min(otu_min_max)` to `r max(otu_min_max)` OTUs per sample (\autoref{supp-calculators}).
+To determine seasonal changes of richness and diversity in studied environments the Observed Number of OTUs, Chao1, ACE, Exponential Shannon [@Jost2006] and Inverse Simpson were calculated after normalization through rarefaction.
+
+```{r}
+# Loading OTU/sample data
+shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")
+
+# Generation of random rarefied community data
+rarefied <- shared %>%
+  select(-label, -Group, -numOtus) %>%
+  rrarefy(., min(rowSums(.))) %>%
+  as.tibble() %>%
+  add_column("Group"=shared$Group, .before=TRUE) %>%
+  select_if(funs(!is.numeric(.) || sum(.)!=0))
+
+# Loading metadata 
+metadata <- read_tsv("data/raw/metadata.csv")
+
+# Joining metadata with OTU/sample data and summing sequences from each environment
+rarefied_metadata <- inner_join(rarefied, metadata, by=c("Group"="ID")) %>%
+  group_by(station) %>%
+  select(starts_with("Otu")) %>%
+  summarise_all(sum)
+
+# Copying the sample labels to the rows (input for library vegan)
+row.names(rarefied_metadata) <- rarefied_metadata$station
+
+# Removing column containing sample labels
+rarefied_metadata <- rarefied_metadata %>%
+  select(-station)
+
+# Calculating matrices of similarity indices
+jaccard <- vegdist(rarefied_metadata, method="jaccard", binary=T)
+jaccard <- as.tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(jaccard))) %>%
+  rename(V1=X1, V2=X2, jaccard=as.numeric.jaccard.)
+
+bray <- vegdist(rarefied_metadata, method="bray", binary=F)
+bray <- as.tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(bray))) %>%
+  rename(V1=X1, V2=X2, bray=as.numeric.bray.)
+
+similarity <- inner_join(jaccard, bray, by=c("V1"="V1", "V2"="V2")) %>%
+  mutate(jaccard=1-jaccard) %>%
+  mutate(bray=1-bray)
+```
+
+To determine the proportion of shared archaeal and bacterial OTUs and shared archaeal and bacterial communities sampled in different environments (*Cymodocea nodosa* [Invaded], *Caulerpa cylindracea* [Invaded and Noninvaded] and surrounding seawater) the Jaccard's Similarity Coefficient on presence-absence data and Bray-Curtis Similarity Coefficient were, respectively, calculated.
+Similarity coefficients were determined after normalization through rarefaction and binning of samples from a respective environment.
+The highest proportion of shared OTUs and communities was found between *Caulerpa cylindracea* invaded and noninvaded meadow communities (Jaccard, `r format(filter(similarity, V1=="FCa", V2=="FCaM")$jaccard, digits=2, nsmall=2)`; Bray-Curtis, `r format(filter(similarity, V1=="FCa", V2=="FCaM")$bray, digits=2, nsmall=2)`), while lower shared values were calculated between seawater and epiphytic communities (\autoref{matrix}).
+Shared proportions of OTUs and communities between *Cymodocea nodosa* and *Caulerpa cylindracea* were approximately in the middle between these two extremes.
+
+To assess seasonal changes in the proportion of shared bacterial and archaeal OTUs the Jaccard's Similarity Coefficient was calculated between consecutive sampling points (\autoref{shared}).
 
@@
 ## Figure Captions
+**\autoref{matrix}.** \nameref{matrix}
+
 **\autoref{shared}.** \nameref{shared}
@@
 ## Figures
+```{r out.width="70%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between communties associated with the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and coomunities in the surrounding seawater.\\label{matrix}", fig.pos="H"}
+include_graphics("../results/figures/matrix.jpg")
+```
+
 ```{r out.width="85%", fig.align="center", fig.cap="Proprotion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\\label{shared}", fig.pos="H"}

### 2020-02-19 submission/manuscript.tex
Intro done, matrix built.

@@
 \hypersetup{unicode=true,
-            pdftitle={Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces},
+            pdftitle={Seasonal Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces},
             pdfborder={0 0 0},
@@
 
-  \title{\textbf{Dynamics of Epiphytic Microbial Communities on Marine Macrophyte
-Surfaces}}
+  \title{\textbf{Seasonal Dynamics of Epiphytic Microbial Communities on Marine
+Macrophyte Surfaces}}
     \pretitle{\vspace{\droptitle}\centering\huge}
@@
 
-Chloroplast sequences were following the seasonal pattern
-(\autoref{supp-chloroplast}, \autoref{pcoa}).
+Sequencing of the 16S rRNA V4 region yielded a total of 1.8 million
+sequences after sequence quality curation and exclusion of eukaryotic,
+chloroplast, mitochondrial and no relative sequences. A total of 35
+samples originating from epiphytic archaeal and bacterial communities
+associated with surfaces of the seagrass \emph{Cymodocea nodosa} and
+macroalga \emph{Caulerpa cylindracea} were analyzed. In addition, 18
+samples (one of the samples was sequenced two times) originating from
+picoplankton archaeal and bacterial communities in the surrounding
+seawater were also processed. The number of reads per sample ranged
+between 8,407 and 77,465 sequences. Even when the highest sequencing
+effort was applied the rarefaction curves did not level off that is a
+common observation in high-throughput 16S rRNA amplicon sequencing
+approaches (\autoref{supp-rarefaction}). Following quality curation and
+exclusion of sequences mentioned before reads were clustered into 28,726
+different OTUs at a similarity level of 97 \si{\percent}. Reads numbers
+were normalized to the minimum number of sequences, 8,407, through
+rarefaction resulting in 17,158 different OTUs that ranged from 373 to
+2,088 OTUs per sample (\autoref{supp-calculators}). To determine
+seasonal changes of richness and diversity in studied environments the
+Observed Number of OTUs, Chao1, ACE, Exponential Shannon (Jost, 2006)
+and Inverse Simpson were calculated after normalization through
+rarefaction.
+
+To determine the proportion of shared archaeal and bacterial OTUs and
+shared archaeal and bacterial communities sampled in different
+environments (\emph{Cymodocea nodosa} {[}Invaded{]}, \emph{Caulerpa
+cylindracea} {[}Invaded and Noninvaded{]} and surrounding seawater) the
+Jaccard's Similarity Coefficient on presence-absence data and
+Bray-Curtis Similarity Coefficient were, respectively, calculated.
+Similarity coefficients were determined after normalization through
+rarefaction and binning of samples from a respective environment. The
+highest proportion of shared OTUs and communities was found between
+\emph{Caulerpa cylindracea} invaded and noninvaded meadow communities
+(Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared values were
+calculated between seawater and epiphytic communities
+(\autoref{matrix}). Shared proportions of OTUs and communities between
+\emph{Cymodocea nodosa} and \emph{Caulerpa cylindracea} were
+approximately in the middle between these two extremes.
+
+To assess seasonal changes in the proportion of shared bacterial and
+archaeal OTUs the Jaccard's Similarity Coefficient was calculated
+between consecutive sampling points (\autoref{shared}).
 
@@
 \textbf{83}: 1--16.
+
+\leavevmode\hypertarget{ref-Jost2006}{}%
+Jost, L. (2006) Entropy and diversity. \textbf{113}: 363--375.
 
@@
 
+\textbf{\autoref{matrix}.} \nameref{matrix}
+
 \textbf{\autoref{shared}.} \nameref{shared}
@@
 
+{\centering \includegraphics[width=0.7\linewidth]{../results/figures/matrix}
+
+}
+
+\caption{Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between communties associated with the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and coomunities in the surrounding seawater.\label{matrix}}\label{fig:unnamed-chunk-3}
+\end{figure}
+
+\begin{figure}[H]
+
 {\centering \includegraphics[width=0.85\linewidth]{../results/figures/seasonal_shared}
@@
 
-\caption{Proprotion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{shared}}\label{fig:unnamed-chunk-2}
+\caption{Proprotion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{shared}}\label{fig:unnamed-chunk-4}
 \end{figure}
@@
 Samples from the same environment or same season are labeld in different colors.
-The proportion of explained variation by each axis is shown on the corresponding axis in parentheses.\label{pcoa}}\label{fig:unnamed-chunk-3}
+The proportion of explained variation by each axis is shown on the corresponding axis in parentheses.\label{pcoa}}\label{fig:unnamed-chunk-5}
 \end{figure}
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant bacterial and archaeal sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{community}}\label{fig:unnamed-chunk-4}
+\caption{Taxonomic classification and relative contribution of the most abundant bacterial and archaeal sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{community}}\label{fig:unnamed-chunk-6}
 \end{figure}
@@
 The proprotion of cyanobaterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
-NR -- No Relative\label{cyano}}\label{fig:unnamed-chunk-5}
+NR -- No Relative\label{cyano}}\label{fig:unnamed-chunk-7}
 \end{figure}
@@
 The proprotion of sequences classified as \textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar.
-NR -- No Relative\label{bactero}}\label{fig:unnamed-chunk-6}
+NR -- No Relative\label{bactero}}\label{fig:unnamed-chunk-8}
 \end{figure}
@@
 The proprotion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
-NR -- No Relative\label{alpha}}\label{fig:unnamed-chunk-7}
+NR -- No Relative\label{alpha}}\label{fig:unnamed-chunk-9}
 \end{figure}
@@
 The proprotion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
-NR -- No Relative\label{gamma}}\label{fig:unnamed-chunk-8}
+NR -- No Relative\label{gamma}}\label{fig:unnamed-chunk-10}
 \end{figure}
@@
 The proprotion of sequences classified as \textit{Desulfobacterota} in the total bacterial and archaeal community is given above the corresponding bar.
-NR -- No Relative\label{desulfo}}\label{fig:unnamed-chunk-9}
+NR -- No Relative\label{desulfo}}\label{fig:unnamed-chunk-11}
 \end{figure}

### 2020-02-19 submission/supplementary.Rmd
Intro done, matrix built.

@@
 title: "**Supplementary Information**"
-subtitle: "**Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces**"
+subtitle: "**Seasonal Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces**"
 output:

### 2020-02-19 code/plot_calculators.R

@@ -212,7 +212,7 @@ p2 <- filter(estimators_indices_metadata, station=="FCa") %>%
 fca <- cowplot::plot_grid(p1, p2, nrow=2, ncol=1, rel_heights=c(1,0.75), align="v")
 
 # Generating a plot to extract a common legend
-labels <- c("Observed Number of OTUs", "Chao1", "ACE", "Shannon Exponential", "Inverse Simpson")
+labels <- c("Observed Number of OTUs", "Chao1", "ACE", "Exponential Shannon", "Inverse Simpson")
 p1 <- filter(estimators_indices_metadata, station=="F") %>%
   ggplot(aes(x=date, y=value, linetype=estimator_index, shape=estimator_index,
              fill=estimator_index)) +

### 2020-02-19 code/plot_matrix.R

@@ -1,0 +1,99 @@
+#################################################################################################################
+# code/plot_matrix.R
+# 
+# A script to plot a matrix of similarity indices.
+# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared
+#               data/raw/metadata.csv
+# Produces: results/figures/matrix.jpg
+#
+#################################################################################################################
+
+# Loading OTU/sample data
+shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")
+
+# Generation of random rarefied community data
+rarefied <- shared %>%
+  select(-label, -Group, -numOtus) %>%
+  rrarefy(., min(rowSums(.))) %>%
+  as.tibble() %>%
+  add_column("Group"=shared$Group, .before=TRUE) %>%
+  select_if(funs(!is.numeric(.) || sum(.)!=0))
+
+# Loading metadata 
+metadata <- read_tsv("data/raw/metadata.csv")
+
+# Joining metadata with OTU/sample data and summing sequences from each environment
+rarefied_metadata <- inner_join(rarefied, metadata, by=c("Group"="ID")) %>%
+  group_by(station) %>%
+  select(starts_with("Otu")) %>%
+  summarise_all(sum)
+
+# Copying the sample labels to the rows (input for library vegan)
+row.names(rarefied_metadata) <- rarefied_metadata$station
+
+# Removing column containing sample labels
+rarefied_metadata <- rarefied_metadata %>%
+  select(-station)
+
+# Calculating matrices of similarity indices
+jaccard <- vegdist(rarefied_metadata, method="jaccard", binary=T)
+jaccard <- as.tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(jaccard))) %>%
+  rename(V1=X1, V2=X2, jaccard=as.numeric.jaccard.)
+
+bray <- vegdist(rarefied_metadata, method="bray", binary=F)
+bray <- as.tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(bray))) %>%
+  rename(V1=X1, V2=X2, bray=as.numeric.bray.)
+
[…]

### 2020-02-19 code/plot_pcoa.R

@@ -22,7 +22,7 @@ rarefied <- shared %>%
 # Loading metadata 
 metadata <- read_tsv("data/raw/metadata.csv")
 
-# Joining metadata and point coordiantes
+# Joining metadata and point coordinates
 rarefied_metadata <- inner_join(metadata, rarefied, by=c("ID"="Group"))
 
 # Function for setting the number of decimal places

### 2020-02-19 code/plot_seasonal_shared.R

@@ -11,7 +11,7 @@
 # Loading OTU/sample data
 shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")
 
-# Generation of random rarified community data
+# Generation of random rarefied community data
 rarefied <- shared %>%
   select(-label, -Group, -numOtus) %>%
   rrarefy(., min(rowSums(.))) %>%

### 2020-02-26 submission/manuscript.Rmd
Results done.

--- 2020-02-14 submission/manuscript.Rmd
@@
 Recently, it was shown that even different morphological niches within the same alga had a higher influence on bacterial community variation than biogeography or environmental factors [@Morrissey2019].
-While there is high community variation between host species is was observed that the majority of metagenomes determined functions were conserved both between host species and individuals [@Burke2011a; @Roth-Schulze2016].
+While there is high community variation between host species is was observed that the majority of metagenome determined functions were conserved both between host species and individuals [@Burke2011a; @Roth-Schulze2016].
 This discrepancy between taxonomic and functional composition could be explained by the lottery hypothesis.
@@
 Bacterial and archaeal epiphytes were sampled in a meadow of *Cymodocea nodosa* invaded by the invasive *Caulerpa cylindracea* and in a locality of only *Caulerpa cylindracea* located in the proximity of the meadow.
-In addition, for comparison the community of the surrounding seawater was characterized.
+In addition, for comparison, the community of the surrounding seawater was characterized.
 
@@
 Thalli of *Caulerpa cylindracea* were sampled in the same *Cymodocea nodosa* invaded meadow in Funtana and on a locality of only *Caulerpa cylindracea* located close to the invaded meadow.
-Sampling of leaves and thalli was performed approximately monthly from December 2017 to October 2018.
+Sampling of leaves and thalli was performed approximately monthly from December 2017 to October 2018 (\autoref{supp-nseq_notus}).
 Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with site seawater.
@@
 ### DNA Isolation
-DNA from surfaces of *Cymodocea nodosa* and *Caulerpa cylindracea* was isolated using a previously modified and adapted protocol that allows a selective epiphytic DNA isolation [@Korlevic_in_press, submitted; @Massana1997].
+DNA from surfaces of *Cymodocea nodosa* and *Caulerpa cylindracea* was isolated using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevic_in_press, submitted; @Massana1997].
 Briefly, leaves and thalli are incubated in a lysis buffer and treated with lysozyme and proteinase K.
@@
 The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
-In the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used to amplify the target region [@Caporaso2012; @Apprill2015; @Parada2016].
+In the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used [@Caporaso2012; @Apprill2015; @Parada2016].
 These primers contained on their 5' end a tagged sequence.
 Purified PCR products were sent for Illumina MiSeq sequencing at IMGM Laboratories, Martinsried, Germany.
-Before sequencing, the second PCR amplification of the two-step PCR procedure was performed at IMGM using primers targeting the tagged region incorporated in the first PCR.
+Before sequencing at IMGM, the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR.
 In addition, these primers contained adapter and sample-specific index sequences.
-Beside samples, a positive and negative control for each sequencing batch were sequenced.
-A negative control was comprised of PCR reactions without DNA template, while for a positive controls a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
+Beside samples, a positive and negative control for each sequencing batch was sequenced.
+Negative control was comprised of PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
 The sequences obtained in this study have been submitted to the European Nucleotide Archive (ENA) under accession numbers **TO BE ADDED LATER!**.
@@
 ### Sequence Analysis
-```{r}
+```{r error}
 # Calculation of the error rate based on the ATCC MSA-1002 mock community
@@
 error <- sum(error$mismatches_weight)/sum(error$total_weight)*100
+
 # Calculation of the number of sequences in the negative control
@@
 ```
+
 Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
 For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
-Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019], package vegan (version 2.5-6) [@Oksanen2019], package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014].
+Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019],  packages  vegan (version 2.5-6) [@Oksanen2019], and tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014].
 The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (**TO BE ADDED LATER!**).
@@
 ## Results
-Chloroplast sequences were following the seasonal pattern (\autoref{supp-chloroplast}, \autoref{pcoa}).
+
+```{r community}
+# Loading input data containing sequence abundances and subsequent input data customization
+community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
+  filter(!str_detect(taxon, "^Eukaryota")) %>%
+  filter(taxon!="Root")
+
+# Remove chloroplast and mitochondrial sequences and subtract their number from higher taxonomic levels to which
+# they belong
+chloroplast <- filter(community, str_detect(taxon, "^Chloroplast$"))$rankID
+mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
+
+community <- mutate_at(community, 5:ncol(community), list(~case_when(
+    rankID==str_extract(chloroplast, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Chloroplast"],
+    rankID==str_extract(chloroplast, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Chloroplast"],
+    rankID==str_extract(chloroplast, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Chloroplast"],
+  TRUE ~ .))) %>%
+  filter(!str_detect(taxon, "^Chloroplast")) %>%
+  mutate_at(5:ncol(.), list(~case_when(
+    rankID==str_extract(mitochondria, "(\\d+\\.){4}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Mitochondria"],
+    TRUE ~ .))) %>%
+  filter(!str_detect(taxon, "^Mitochondria")) %>%
+  select(-ATCC_1, -ATCC_5, -NC_1)
+
+# Calculating the number of sequencing in millions
+million_sequences <- filter(community, taxon=="Archaea" | taxon=="Bacteria") %>%
+  select(total) %>%
+  sum()/1e6
+
+# Loading metadata
+metadata <- read_tsv("data/raw/metadata.csv")
+
+# Calculating number of samples
+samples_epiphytes <- filter(metadata, !is.na(station) & station!="F") %>%
+  summarise(n=n())
+samples_seawater <- filter(metadata, !is.na(station) & station=="F") %>%
+  summarise(n=n())
+
+# Calculating the minimal and maximal number of sequences
+min_max <- filter(community, taxon=="Archaea" | taxon=="Bacteria") %>%
+  summarise_at(6:ncol(.), sum) %>%
+  summarise(min=min(.), max=max(.))
+
+# Calculating the relative abundance of Archaea and Bacteria
+archaea_bacteria <- filter(community, taxon=="Archaea" | taxon=="Bacteria") %>%
+  mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
+  gather(key="Group", value="abundance", 6:(ncol(.))) %>%
+  group_by(taxon) %>%
+  summarise(mean=mean(abundance), sd=sd(abundance))
+```
+
+```{r shared}
+# Loading OTU/sample data
+shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")
+
+# Calculating the number of OTUs before rarefaction
+no_otus <- select(shared, starts_with("Otu")) %>%
+  colnames() %>%
+  length()
+
+# Generation of random rarefied community data
+rarefied <- shared %>%
+  select(-label, -Group, -numOtus) %>%
+  rrarefy(., min(rowSums(.))) %>%
+  as_tibble() %>%
+  add_column("Group"=shared$Group, .before=TRUE) %>%
+  select_if(list(~ !is.numeric(.) || sum(.)!=0))
+
+# Calculating the number of OTUs after rarefaction
+no_otus_rarefied <- select(rarefied, starts_with("Otu")) %>%
+  colnames() %>%
+  length()
+
+# Calculating the maximal and minimal number of OTUs after rarefaction
+otu_min_max <- select(rarefied, starts_with("Otu")) %>%
+  decostand(., method="pa") %>%
+  transmute(sum=rowSums(.))
+```
+
+```{r calculators}
+# Copying the sample labels to the rows (input for library vegan)
+row.names(rarefied) <- rarefied$Group
+
+# Removing column containing sample labels
+rarefied <- rarefied %>%
+  select(-Group)
+
+# Calculating Chao1 and ACE species estimators
+estimators <- estimateR(rarefied)
+estimators <- as_tibble(t(estimators)) %>%
+  add_column("Group"=colnames(estimators), .before=TRUE)
+
+# Calculating diversity indices
+shannon <- diversity(rarefied, index="shannon")
+shannon <- tibble(Group=names(shannon), shannon)
+invsimpson <- diversity(rarefied, index="invsimpson")
+invsimpson <- tibble(Group=names(invsimpson), invsimpson)
+
+# Transforming the Shannon entropy to the effective number of species
+# (http://www.loujost.com/Statistics%20and%20Physics/Diversity%20and%20Similarity/EffectiveNumberOfSpecies.htm)
+shannon <- mutate(shannon, shannon=exp(shannon)) %>%
+  rename(eshannon=shannon)
+
+# Joining together estimators and indices
+estimators_indices <- inner_join(estimators, shannon, by="Group") %>%
+  inner_join(., invsimpson, by="Group")
+
+# Loading metadata 
+metadata <- read_tsv("data/raw/metadata.csv")
+
+# Joining metadata with estimators and indices
+estimators_indices_metadata <- inner_join(metadata, estimators_indices, by=c("ID"="Group")) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y"))
+
+# Calculating the mean and standard deviation for each environment
+S.obs_f <- filter(estimators_indices_metadata, station=="F") %>%
+  summarise(mean=mean(S.obs), sd=sd(S.obs))
+S.obs_fcym <- filter(estimators_indices_metadata, station=="FCyM") %>%
+  summarise(mean=mean(S.obs), sd=sd(S.obs))
+S.obs_fcam <- filter(estimators_indices_metadata, station=="FCaM") %>%
+  summarise(mean=mean(S.obs), sd=sd(S.obs))
+S.obs_fca <- filter(estimators_indices_metadata, station=="FCa") %>%
+  summarise(mean=mean(S.obs), sd=sd(S.obs))
+
+# Calculating the mean and standard deviation for each environment and season
+S.obs_season_f <- filter(estimators_indices_metadata, station=="F") %>%
+  group_by(season) %>%
+  summarise(mean=mean(S.obs), sd=sd(S.obs))
+S.obs_season_fcym <- filter(estimators_indices_metadata, station=="FCyM") %>%
+  group_by(season) %>%
+  summarise(mean=mean(S.obs), sd=sd(S.obs))
+S.obs_season_fcam <- filter(estimators_indices_metadata, station=="FCaM") %>%
+  group_by(season) %>%
+  summarise(mean=mean(S.obs), sd=sd(S.obs))
+S.obs_season_fca <- filter(estimators_indices_metadata, station=="FCa") %>%
+  group_by(season) %>%
+  summarise(mean=mean(S.obs), sd=sd(S.obs))
+```
+
+Sequencing of the 16S rRNA V4 region yielded a total of `r million_sequences` million sequences after quality curation and exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences (\autoref{supp-nseq_notus}).
+A total of `r samples_epiphytes` samples originating from epiphytic archaeal and bacterial communities associated with surfaces of the seagrass *Cymodocea nodosa* and macroalga *Caulerpa cylindracea* were analyzed.
+In addition, `r samples_seawater` samples (one of the samples was sequenced two times) originating from picoplankton archaeal and bacterial communities in the surrounding seawater were also processed.
+The number of reads per sample ranged between `r min_max$min` and `r min_max$max` sequences (\autoref{supp-nseq_notus}).
+Even when the highest sequencing effort was applied the rarefaction curves did not level off that is a common observation in high-throughput 16S rRNA amplicon sequencing approaches (\autoref{supp-rarefaction}).
+Following quality curation and exclusion of sequences mentioned before reads were clustered into `r no_otus` different OTUs at a similarity level of 97 \si{\percent}.
+Reads numbers were normalized to the minimum number of sequences, `r min_max$min` (\autoref{supp-nseq_notus}), through rarefaction resulting in `r no_otus_rarefied` different OTUs that ranged from `r min(otu_min_max)` to `r max(otu_min_max)` OTUs per sample (\autoref{supp-calculators}).
+To determine seasonal changes of richness and diversity the Observed Number of OTUs, Chao1, ACE, Exponential Shannon [@Jost2006] and Inverse Simpson were calculated after normalization through rarefaction.
+Generally, richness estimators and diversity indices showed similar trends.
+On average, higher values were found for *Caulerpa cylindracea* (invaded [Number of OTUs, `r S.obs_fcam$mean` ± `r S.obs_fcam$sd` OTUs] and noninvaded [Number of OTUs, `r S.obs_fca$mean` ± `r S.obs_fca$sd` OTUs]), middle vales for *Cymodocea nodosa* (Number of OTUs, `r S.obs_fcym$mean` ± `r S.obs_fcym$sd` OTUs) and lower values for picoplankton communities in the surrounding seawater (Number of OTUs, `r S.obs_f$mean` ± `r S.obs_f$sd` OTUs) (\autoref{supp-calculators}).
+Seasonal changes did not show such large dissimilarities.
+Seawater communities richness was stable during the studied period with the exception of one sampling point in December 2017 when larger values were observed.
+*Cymodocea nodosa* communities showed a slow increase towards the end of the study, while *Caulerpa cylindracea* (invaded and noninvaded) communities were characterized by slightly larger values in Spring and Summer in comparison to Autumn and Winter (\autoref{supp-calculators}).
+
+```{r seasonal_shared}
+# Generation of random rarefied community data
+rarefied <- shared %>%
+  select(-label, -Group, -numOtus) %>%
+  rrarefy(., min(rowSums(.))) %>%
+  as_tibble() %>%
+  add_column("Group"=shared$Group, .before=TRUE) %>%
+  select_if(list(~ !is.numeric(.) || sum(.)!=0))
+
+# Loading metadata 
+metadata <- read_tsv("data/raw/metadata.csv")
+
+# Joining metadata with OTU/sample data and summing sequences from each environment
+rarefied_metadata <- inner_join(rarefied, metadata, by=c("Group"="ID")) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
+  filter(date >= "2017-11-01") %>%
+  select_if(list(~ !is.numeric(.) || sum(.)!=0)) %>%
+  group_by(station) %>%
+  select(starts_with("Otu")) %>%
+  summarise_all(sum)
+
+# Copying the sample labels to the rows (input for library vegan)
+row.names(rarefied_metadata) <- rarefied_metadata$station
+
+# Removing column containing sample labels
+rarefied_metadata <- rarefied_metadata %>%
+  select(-station)
+
+# Calculating matrices of similarity indices
+jaccard <- vegdist(rarefied_metadata, method="jaccard", binary=T)
+jaccard <- as_tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(jaccard))) %>%
+  rename(V1=X1, V2=X2, jaccard=as.numeric.jaccard.)
+
+bray <- vegdist(rarefied_metadata, method="bray", binary=F)
+bray <- as_tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(bray))) %>%
+  rename(V1=X1, V2=X2, bray=as.numeric.bray.)
+
+similarity <- inner_join(jaccard, bray, by=c("V1"="V1", "V2"="V2")) %>%
+  mutate(jaccard=1-jaccard) %>%
+  mutate(bray=1-bray)
+```
+
+```{r anosim}
+# Generation of random rarefied community data and preserving sample names as row names
+rarefied <- shared %>%
+  select(-label, -Group, -numOtus) %>%
+  rrarefy(., min(rowSums(.))) %>%
+  as_tibble() %>%
+  select_if(list(~!is.numeric(.) || sum(.)!=0))
+rownames(rarefied) <- shared$Group
+rarefied <- rarefied %>%
+  add_column("Group"=shared$Group, .before=TRUE)
+
+# Loading metadata 
+metadata_rarefied <- read_tsv("data/raw/metadata.csv") %>%
+  inner_join(., rarefied, by=c("ID"="Group"))
+
+# Calculating ANOSIM
+metadata_rarefied_select <- metadata_rarefied %>%
+  select(starts_with("Otu"))
+grouping <- metadata_rarefied %>%
+  mutate(grouping=str_replace(station, "CaM$", "Ca"))
+anosim_all <- anosim(metadata_rarefied_select, grouping$grouping, permutations=1000, distance="bray")
+anosim_all_signif <- case_when(anosim_all$signif < 0.05 & anosim_all$signif >= 0.01 ~ 0.05,
+                               anosim_all$signif < 0.01 & anosim_all$signif >= 0.001 ~ 0.01,
+                               anosim_all$signif < 0.001 & anosim_all$signif >= 0.0001 ~ 0.001,
+                               anosim_all$signif < 0.0001 ~ 0.0001,
+                               TRUE ~ anosim_all$signif)
+
+metadata_rarefied_select <- metadata_rarefied %>%
+  filter(station=="F") %>%
+  select(starts_with("Otu"))
+grouping <- metadata_rarefied %>%
+  filter(station=="F") %>%
+  mutate(grouping=str_replace(season, "Autumn", "Winter"))
+anosim_f <- anosim(metadata_rarefied_select, grouping$grouping, permutations=1000, distance="bray")
+anosim_f_signif <- case_when(anosim_f$signif < 0.05 & anosim_f$signif >= 0.01 ~ 0.05,
+                             anosim_f$signif < 0.01 & anosim_f$signif >= 0.001 ~ 0.01,
+                             anosim_f$signif < 0.001 & anosim_f$signif >= 0.0001 ~ 0.001,
+                             anosim_f$signif < 0.0001 ~ 0.0001,
+                             TRUE ~ anosim_f$signif)
+
+metadata_rarefied_select <- metadata_rarefied %>%
+  filter(station=="FCyM") %>%
+  select(starts_with("Otu"))
+grouping <- metadata_rarefied %>%
+  filter(station=="FCyM") %>%
+  mutate(grouping=str_replace(season, "Autumn", "Winter"))
+anosim_fcym <- anosim(metadata_rarefied_select, grouping$grouping, permutations=1000, distance="bray")
+anosim_fcym_signif <- case_when(anosim_fcym$signif < 0.05 & anosim_fcym$signif >= 0.01 ~ 0.05,
+                                anosim_fcym$signif < 0.01 & anosim_fcym$signif >= 0.001 ~ 0.01,
+                                anosim_fcym$signif < 0.001 & anosim_fcym$signif >= 0.0001 ~ 0.001,
+                                anosim_fcym$signif < 0.0001 ~ 0.0001,
+                                TRUE ~ anosim_fcym$signif)
+
+metadata_rarefied_select <- metadata_rarefied %>%
+  filter(station=="FCaM" | station=="FCa") %>%
+  select(starts_with("Otu"))
+grouping <- metadata_rarefied %>%
+  filter(station=="FCaM" | station=="FCa") %>%
+  mutate(grouping=str_replace(season, "Autumn", "Winter")) %>%
+  mutate(grouping=str_replace(grouping, "Spring", "Winter"))
+anosim_fca <- anosim(metadata_rarefied_select, grouping$grouping, permutations=1000, distance="bray")
+anosim_fca_signif <- case_when(anosim_fca$signif < 0.05 & anosim_fca$signif >= 0.01 ~ 0.05,
+                                anosim_fca$signif < 0.01 & anosim_fca$signif >= 0.001 ~ 0.01,
+                                anosim_fca$signif < 0.001 & anosim_fca$signif >= 0.0001 ~ 0.001,
+                                anosim_fca$signif < 0.0001 ~ 0.0001,
+                                TRUE ~ anosim_fca$signif)
+```
+
+To determine the proportion of shared archaeal and bacterial OTUs and communities sampled in different environments the Jaccard's Similarity Coefficient on presence-absence data and Bray-Curtis Similarity Coefficient were, respectively, calculated.
+Coefficients were determined after normalization through rarefaction and binning of samples from a particular environment.
+The highest proportion of shared OTUs and community was found between invaded and noninvaded  *Caulerpa cylindracea* (Jaccard, `r format(filter(similarity, V1=="FCa", V2=="FCaM")$jaccard, digits=2, nsmall=2)`; Bray-Curtis, `r format(filter(similarity, V1=="FCa", V2=="FCaM")$bray, digits=2, nsmall=2)`), while lower shared values were calculated between seawater and epiphytic communities (\autoref{matrix}).
+Shared proportion between *Cymodocea nodosa* and *Caulerpa cylindracea* were approximately in the middle between these two extremes.
+To assess seasonal changes in the proportion of shared OTUs and communities the Jaccard's and Bray-Curtis Similarity Coefficients were calculated between consecutive sampling points (\autoref{shared}).
+Both coefficients showed similar trends.
+Temporal proportional changes were more pronounced for seawater in comparison to *Cymodocea nodosa* and especially *Caulerpa cylindracea* associated communities (\autoref{shared}).
+To further disentangle the environmental and seasonal community dissimilarity a Principal Coordinates Analysis (PCoA) based on Bray-Curtis distances and OTU abundances was applied.
+It showed a clear separation between planktonic and surface associated communities (\autoref{pcoa}).
+In addition, a separation of epiphytic bacterial and archaeal communities based on host species was determined.
+This separation was further supported by ANOSIM (R = `r format(anosim_all$statistic, digits=2, nsmall=2)`, *p* < `r anosim_all_signif`).
+Seasonal changes of seawater communities indicated a separation between Spring, Summer and Autumn/Winter samples (ANOSIM, R = `r format(anosim_f$statistic, digits=2, nsmall=2)`, *p* < `r anosim_f_signif`).
+Epiphytic microbial communities associated with *Cymodocea nodosa* showed a similar pattern (ANOSIM, R = `r format(anosim_fcym$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fcym_signif`), while communities from the surfaces of *Caulerpa cylindracea* indicated a non so strongly supported, as in previous cases, separation between Summer and Autumn/Winter/Spring samples (ANOSIM, R = `r format(anosim_fca$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fca_signif`) (\autoref{pcoa}).
+
+```{r chloroplast}
+# Loading input data containing sequence abundances and subsequent input data customization
+community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
+  filter(!str_detect(taxon, "^Eukaryota")) %>%
+  filter(taxon!="Root")
+
+# Remove mitochondrial sequences and subtract their number from higher taxonomic levels to which
+# they belong
+mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
+community <- mutate_at(community, 5:ncol(community), list(~case_when(
+    rankID==str_extract(mitochondria, "(\\d+\\.){4}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Mitochondria"],
+    TRUE ~ .))) %>%
+  filter(!str_detect(taxon, "^Mitochondria")) %>%
+  select(-ATCC_1, -ATCC_5, -NC_1) %>%
+  mutate(`23`=`23_1`+`23_2`) %>%
+  select(-`23_1`, -`23_2`) %>%
+  group_by(taxlevel) %>%
+  mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
+  ungroup()
+
+# Selection of groups for plotting
+plot <- community %>%
+  filter(taxon=="Chloroplast") %>%
+  bind_rows(summarise_all(., list(~ifelse(is.numeric(.), 100-sum(.), paste("Other")))))
+
+# Loading colors for each group on the plot
+color <- read_tsv("data/raw/group_colors.csv") %>%
+  select(-Taxlevel) %>%
+  deframe()
+
+# Generation of italic names for groups
+names <- parse(text=case_when(plot$taxon=="Chloroplast" ~ paste0("plain('", plot$taxon,  "')"),
+                              plot$taxon=="Bacteria_unclassified" ~ "italic('Bacteria')~plain('(No Relative)')",
+                              plot$taxon=="Marinimicrobia_(SAR406_clade)" ~ "italic('Marinimicrobia')",
+                              plot$taxon=="Other" ~ paste0("plain('", plot$taxon, "')"),
+                              plot$taxon=="No Relative" ~ paste0("plain('", plot$taxon, "')"),
+                              TRUE ~ paste0("italic('", plot$taxon, "')")))
+
+# Tidying the sequence abundance data
+plot <- gather(plot, key="Group", value="abundance", 6:(ncol(plot)))
+
+# Loading metadata
+metadata <- read_tsv("data/raw/metadata.csv") %>%
+  filter(ID!="23_1") %>%
+  mutate(ID=str_replace(ID, "23_2", "23")) %>%
+  mutate(label=str_replace(label, "24/4-18 F-2", "24/4-18 F"))
+
+# Joining sequence abundances data and metadata
+plot <- inner_join(metadata, plot, by=c("ID"="Group")) %>%
+  mutate(taxon=factor(taxon, levels=rev(unique(plot$taxon)))) %>%
+  mutate(label=factor(label, levels=metadata$label)) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y"))
+
+# Autumn/Winter and Spring/Summer abundances
+chloroplast <- plot %>%
+  filter(station!="F" & taxon!="Other") %>%
+  mutate(season=str_replace(season, "Autumn", "Winter")) %>%
+  mutate(season=str_replace(season, "Spring", "Summer")) %>%
+  group_by(season) %>%
+  summarise(mean=mean(abundance), sd=sd(abundance))
+```
+
+The microbial taxonomic composition of both, macrophyte associated and seawater communities, was dominated by *Bacteria* (`r filter(archaea_bacteria, taxon=="Bacteria")$mean` ± `r filter(archaea_bacteria, taxon=="Bacteria")$sd` \si{\percent}) over *Archaea* (`r filter(archaea_bacteria, taxon=="Archaea")$mean` ± `r filter(archaea_bacteria, taxon=="Archaea")$sd` \si{\percent}) (\autoref{community}).
+Three phyla were characteristic for *Archaea*: *Crenarchaeota*, *Nanoarchaeota* and *Thermoplasmatota*.
+Their contribution  was larger only in seawater communities during December 2017 (\autoref{community}).
+Higher relative abundances of chloroplast related sequences were only observed in surface associated communities, with higher values in Autumn/Winter (`r filter(chloroplast, season=="Winter")$mean` ± `r filter(chloroplast, season=="Winter")$sd` \si{\percent}) in comparison to Spring/Summer (`r filter(chloroplast, season=="Summer")$mean` ± `r filter(chloroplast, season=="Summer")$sd` \si{\percent}) (\autoref{supp-chloroplast}).
+Generally, at higher taxonomic ranks (phylum-class) epiphytic and seawater microbial communities were composed of similar bacterial taxa.
+Seawater communities were mainly comprised of *Actinobacteriota*, *Bacteroidota*, *Cyanobacteria*, *Alphaproteobacteria*, *Gammaproteobacteria* and *Verrucomicrobiota*.
+Communities associated with *Cymodocea nodosa* were consisted of same groups with the addition of *Planctomycetota* whose contribution was higher in summer 2018.
+In addition, communities from invaded and noninvaded *Caulerpa cylindracea* were similar and characterized by same groups as seawater and *Cymodocea nodosa* communities with the addition of *Desulfobacterota* (\autoref{community}).
+Larger differences between environments and host species could be observed at lower taxonomic ranks (\autoref{cyano} -- \ref{desulfo}).
+
+```{r cyano}
+# Loading input data containing sequence abundances and subsequent input data customization
+community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
+  filter(!str_detect(taxon, "^Eukaryota")) %>%
+  filter(taxon!="Root")
+
+# Remove Chloroplast and Mitochondria sequences and subtract their number from higher taxonomic levels to which
+# they belong
+chloroplast <- filter(community, str_detect(taxon, "^Chloroplast$"))$rankID
+mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
+community <- mutate_at(community, 5:ncol(community), list(~case_when(
+  rankID==str_extract(chloroplast, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Chloroplast"],
+  rankID==str_extract(chloroplast, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Chloroplast"],
+  rankID==str_extract(chloroplast, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Chloroplast"],
+  TRUE ~ .))) %>%
+  filter(!str_detect(taxon, "^Chloroplast")) %>%
+  mutate_at(5:ncol(.), list(~case_when(
+    rankID==str_extract(mitochondria, "(\\d+\\.){4}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Mitochondria"],
+    TRUE ~ .))) %>%
+  filter(!str_detect(taxon, "^Mitochondria")) %>%
+  select(-ATCC_1, -ATCC_5, -NC_1) %>%
+  mutate(`23`=`23_1`+`23_2`) %>%
+  select(-`23_1`, -`23_2`) %>%
+  group_by(taxlevel) %>%
+  mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
+  ungroup()
+
+# Selection of groups for plotting
+select <- filter(community,
+               taxlevel==6 &
+               str_detect(rankID, filter(community, str_detect(taxon, "^Cyanobacteria$"))$rankID)) %>%
+  filter_at(6:ncol(.), any_vars(. >= 1))
+
+plot <- filter(community,
+               taxlevel==6 &
+               str_detect(rankID, filter(community, str_detect(taxon, "^Cyanobacteria$"))$rankID)) %>%
+  mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
+  ungroup() %>%
+  filter(taxon %in% select$taxon) %>%
+  bind_rows(summarise_all(., list(~ifelse(is.numeric(.), 100-sum(.), paste("Other_Cyanobacteria")))))
+
+# Loading colors for each group on the plot
+color <- read_tsv("data/raw/group_colors.csv") %>%
+  select(-Taxlevel) %>%
+  deframe()
+
+# Generation of italic names for taxonomic groups
+names <- parse(text=case_when(plot$taxon=="Cyanobacteriales_unclassified" ~ "italic('Cyanobacteriales')~plain('(NR)')",
+                              plot$taxon=="Spirulina_DRTO-55.2" ~ "italic('Spirulina')",
+                              plot$taxon=="Pleurocapsa_PCC-7319" ~ "italic('Pleurocapsa')",
+                              plot$taxon=="Cyanobacteriia_unclassified" ~
+                                "italic('Cyanobacteriia')~plain('(NR)')",
+                              plot$taxon=="Nodosilineaceae_unclassified" ~ "italic('Nodosilineaceae')~plain('(NR)')",
+                              plot$taxon=="Acrophormium_PCC-7375" ~ "italic('Acrophormium')",
+                              plot$taxon=="Phormidesmis_ANT.LACV5.1" ~ "italic('Phormidesmis')",
+                              plot$taxon=="Phormidium_MBIC10003" ~ "italic('Phormidium')",
+                              plot$taxon=="Cyanobium_PCC-6307" ~ "italic('Cyanobium')",
+                              plot$taxon=="Synechococcus_CC9902" ~ "italic('Synechococcus')",
+                              plot$taxon=="Schizothrix_LEGE_07164" ~ "italic('Schizothrix')",
+                              plot$taxon=="Thermosynechococcales_unclassified" ~ "italic('Thermosynechococcales')~plain('(NR)')",  
+                              plot$taxon=="Other_Cyanobacteria" ~ "plain('Other')~italic('Cyanobacteria')",
+                              TRUE ~ paste0("italic('", plot$taxon, "')")))
+
+# Tidying the sequence abundance data
+plot <- gather(plot, key="Group", value="abundance", 6:ncol(plot))
+
+# Loading metadata
+metadata <- read_tsv("data/raw/metadata.csv") %>%
+  filter(ID!="23_1") %>%
+  mutate(ID=str_replace(ID, "23_2", "23")) %>%
+  mutate(label=str_replace(label, "24/4-18 F-2", "24/4-18 F"))
+
+# Joining sequence abundances data and metadata
+plot <- inner_join(metadata, plot, by=c("ID"="Group")) %>%
+  mutate(taxon=factor(taxon, levels=unique(plot$taxon))) %>%
+  mutate(label=factor(label, levels=metadata$label)) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y"))
+
+# Selecting the relative abundance of the targeted group in the whole community and tidying the obtained data
+whole <- filter(community,
+                taxlevel==2) %>%
+  gather(key="Group", value="abundance", 6:ncol(.)) %>%
+  filter(taxon=="Cyanobacteria")
+
+# Calculating the mean and standard deviation of all samples
+cyano_all_whole <- inner_join(metadata, whole, by=c("ID"="Group")) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
+  summarise(mean=mean(abundance), sd=sd(abundance))
+
+# Calculating the mean and standard deviation for each season and station
+cyano_station_season_whole <- inner_join(metadata, whole, by=c("ID"="Group")) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
+  group_by(station, season) %>%
+  summarise(mean=mean(abundance), sd=sd(abundance))
+```
+
+*Cyanobacteria* related sequences were comprising, on average, `r cyano_all_whole$mean` ± `r cyano_all_whole$sd` \si{\percent} of total sequences (\autoref{cyano}).
+Higher proportions were found for *Cymodocea nodosa* (`r filter(cyano_station_season_whole, station=="FCyM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCyM" & season=="Autumn")$sd` \si{\percent}) and *Caulerpa cylindracea* (invaded [ (`r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$sd` \si{\percent})] and noninvaded [ (`r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$sd` \si{\percent})]) associated communities in autumn and for seawater communities in winter (`r filter(cyano_station_season_whole, station=="F" & season=="Winter")$mean` ± `r filter(cyano_station_season_whole, station=="F" & season=="Winter")$sd` \si{\percent}).
+Large taxonomic differences between surface associated and seawater cyanobacterial communities were observed.
+Seawater communities were mainly comprised of *Cyanobium* and *Synechococcus*, while surface associated communities were consisted of and sequences without known relatives within *Cyanobacteriales* and *Cyanobacteriia* (\autoref{cyano}).
+In addition, seasonal changes in surface associated communities were observed with *Cyanobacteriales* and *Cyanobacteriia* no relative sequences comprising larger proportions in autumn and winter and *Acrophormium*, *Phormidesmis* and no relative *Nodosilineaceae* and in spring and summer (\autoref{cyano}).
+
+```{r bactero}
+# Selecting the relative abundance of the targeted group in the whole community and tidying the obtained data
+whole <- filter(community,
+                taxlevel==2) %>%
+  gather(key="Group", value="abundance", 6:ncol(.)) %>%
+  filter(taxon=="Bacteroidota")
+
+# Calculating the mean and standard deviation of all samples
+bactero_all_whole <- inner_join(metadata, whole, by=c("ID"="Group")) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
+  summarise(mean=mean(abundance), sd=sd(abundance))
+```
+
+Sequences classified as *Bacteroidota* were comprising, on average, `r bactero_all_whole$mean` ± `r bactero_all_whole$sd` \si{\percent} of all sequences (\autoref{bactero}).
+Similarly to *Cyanobacteria*, large differences in the taxonomic composition between seawater and surface associated communities were found (\autoref{bactero}).
+The seawater community was characterized by the NS4 and NS5 marine groups, uncultured *Cryomorphaceae*, uncultured *Flavobacteriaceae*, NS11-12 marine group, *Balneola*, uncultured *Balneolaceae* and *Formosa*.
+In contrast, surface associated communities were characterized by *Lewinella*, *Portibacter*, *Rubidimonas*, no relative *Saprospiraceae*, uncultured *Saprospiraceae*, no relative *Flavobacteriaceae* and uncultured *Rhodothermaceae*.
+Surface associated *Bacteroidota* communities were very diverse as could be observed in the the high proportion of taxa that grouped as other *Bacteroidota* (\autoref{bactero}).
+
+```{r alpha}
+# Selection of groups for plotting
+select <- filter(community,
+               taxlevel==6 &
+               str_detect(rankID, filter(community, str_detect(taxon, "^Alphaproteobacteria$"))$rankID)) %>%
+  filter_at(6:ncol(.), any_vars(. >= 2))
+
+plot <- filter(community,
+               taxlevel==6 &
+               str_detect(rankID, filter(community, str_detect(taxon, "^Alphaproteobacteria$"))$rankID)) %>%
+  mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
+  ungroup() %>%
+  filter(rankID %in% select$rankID) %>%
+  bind_rows(summarise_all(., list(~ifelse(is.numeric(.), 100-sum(.), paste("Other_Alphaproteobacteria"))))) %>%
+# Remove last digit from rankID so for the next step
+  mutate(rankID=if_else(taxon=="uncultured", str_replace(rankID, "\\.\\d+$", ""), rankID))
+
+# Adding data to "uncultured" taxa describing the higher taxonomic level to which they belong
+uncultured <- select(filter(community, rankID %in% filter(plot, taxon=="uncultured")$rankID), rankID, taxon) %>%
+  rename(rankID_uncultured=rankID, taxon_uncultured=taxon)
+
+plot <- left_join(plot, uncultured, by=c("rankID"="rankID_uncultured")) %>%
+  mutate(taxon=if_else(taxon=="uncultured", paste0(taxon, "_", taxon_uncultured), taxon)) %>%
+  select(-taxon_uncultured)
+
+# Loading colors for each group on the plot
+color <- read_tsv("data/raw/group_colors.csv") %>%
+  select(-Taxlevel) %>%
+  deframe()
+
+# Generation of italic names for taxonomic groups
+names <- parse(text=case_when(str_detect(plot$taxon, "uncultured") ~ paste0("plain('Uncultured')~italic('", str_remove(plot$taxon, "uncultured_"), "')"),
+                              str_detect(plot$taxon, "unclassified") ~ paste0("italic('", str_remove(plot$taxon, "_unclassified"), "')~plain('(NR)')"),
+                              plot$taxon=="OCS116_clade_ge" ~ "plain('OCS116 Clade')",
+                              plot$taxon=="Candidatus_Puniceispirillum" ~ "italic('\"Candidatus')~plain('Puniceispirillum\"')",
+                              plot$taxon=="SAR116_clade_ge" ~ "plain('SAR116 Clade')",
+                              plot$taxon=="Stappiaceae_ge" ~ "italic('Stappiaceae')",
+                              plot$taxon=="HIMB11" ~ "plain('HIMB11')",
+                              plot$taxon=="AEGEAN-169_marine_group_ge" ~ "plain('AEGEAN-169 Marine Group')",
+                              plot$taxon=="Clade_Ia" ~ "plain('Clade Ia')",
+                              plot$taxon=="Clade_Ib" ~ "plain('Clade Ib')",
+                              plot$taxon=="Clade_II_ge" ~ "plain('Clade II')",
+                              plot$taxon=="Clade_III_ge" ~ "plain('Clade III')",
+                              plot$taxon=="Other_Alphaproteobacteria" ~ "plain('Other')~italic('Alphaproteobacteria')",
+                              TRUE ~ paste0("italic('", plot$taxon, "')")))
+
+# Tidying the sequence abundance data
+plot <- gather(plot, key="Group", value="abundance", 6:ncol(plot))
+
+# Loading metadata
+metadata <- read_tsv("data/raw/metadata.csv") %>%
+  filter(ID!="23_1") %>%
+  mutate(ID=str_replace(ID, "23_2", "23")) %>%
+  mutate(label=str_replace(label, "24/4-18 F-2", "24/4-18 F"))
+
+# Joining sequence abundances data and metadata
+plot <- inner_join(metadata, plot, by=c("ID"="Group")) %>%
+  mutate(taxon=factor(taxon, levels=unique(plot$taxon))) %>%
+  mutate(label=factor(label, levels=metadata$label)) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y"))
+
+# Selecting the relative abundance of the targeted group in the whole community and tidying the obtained data
+whole <- filter(community,
+                taxlevel==3) %>%
+  gather(key="Group", value="abundance", 6:ncol(.)) %>%
+  filter(taxon=="Alphaproteobacteria")
+
+# Calculating the mean and standard deviation of all samples
+alpha_all_whole <- inner_join(metadata, whole, by=c("ID"="Group")) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
+  summarise(mean=mean(abundance), sd=sd(abundance))
+
+# Calculating the mean abundance and standard deviation of Rhodobacteraceae_unclassified within Alpharoteobacteria
+rhodo <- plot %>%
+  filter(season!="F" & taxon=="Rhodobacteraceae_unclassified") %>%
+  summarise(mean=mean(abundance), sd=sd(abundance))
+```
+
+On average, *Alphaproteobacteria* were in comparison to other high rank taxa the largest taxonomic group, comprising `r alpha_all_whole$mean` ± `r alpha_all_whole$sd` \si{\percent} of all sequences (\autoref{alpha}).
+In accordance to previous taxa, high differences between seawater and surface associated communities were observed.
+Picoplankton communities were composed mainly of the SAR11 clade, AEGEAN-169 marine group, SAR116 clade, no relative *Rhodobacteraceae*, HIMB11 and OCS116 clade, while surface associated communities were composed in high proportion of no relative *Rhodobacteraceae* and to a lesser degree of  *Pseudoahrensia*, no relative *Alphaproteobacteria* and no relative  *Hyphomonadaceae*.
+Representatives of no relative *Rhodobacteraceae* were comprising in on average `r rhodo$mean` ± `r rhodo$sd` \si{\percent} of all alphaproteobacterial sequences from the epiphytic community (\autoref{alpha}).
+
+```{r gamma}
+# Selecting the relative abundance of the targeted group in the whole community and tidying the obtained data
+whole <- filter(community,
+                taxlevel==3) %>%
+  gather(key="Group", value="abundance", 6:ncol(.)) %>%
+  filter(taxon=="Gammaproteobacteria")
+
+# Calculating the mean and standard deviation of all samples
+gamma_all_whole <- inner_join(metadata, whole, by=c("ID"="Group")) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
+  summarise(mean=mean(abundance), sd=sd(abundance))
+```
+
+Sequences related to *Gammaproteobacteria* were comprising, on average, `r gamma_all_whole$mean` ± `r gamma_all_whole$sd` \si{\percent} of all sequences (\autoref{gamma}).
+Similarly to previous taxa, large taxonomic differences between seawater and surface associated communities were found.
+Seawater communities were mainly comprised of the OM60 (NOR5) clade, *Litoricola*, *Acinetobacter* and the SAR86 clade, while epiphytic communities were mainly composed of no relative *Gammaproteobacteria* and *Granulosicocus*.
+Beside these two groups specific to all three epiphytic communities, *Cymodocea nodosa* was characterized by *Arenicella*, no relative *Burkholderiales* and *Methylotenera*, while *Thioploca*, no relative *Cellvibrionaceae* and *Reinekea* were more specific to both invaded and noninvaded *Caulerpa cylindracea*.
+In addition, similarly to *Bacteroidota*, a large proportion of the surface associated community was grouped as other *Gammaproteobacteria* indicating high diversity within this group (\autoref{gamma}).
+
+```{r delta}
+# Selecting the relative abundance of the targeted group in the whole community and tidying the obtained data
+whole <- filter(community,
+                taxlevel==2) %>%
+  gather(key="Group", value="abundance", 6:ncol(.)) %>%
+  filter(taxon=="Desulfobacterota")
+
+# Calculating the mean and standard deviation of all samples
+desulfo_all_whole <- inner_join(metadata, whole, by=c("ID"="Group")) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
+  summarise(mean=mean(abundance), sd=sd(abundance))
+
+# Calculating the mean and standard deviation for each station
+desulfo_station_whole <- inner_join(metadata, whole, by=c("ID"="Group")) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
+  group_by(station) %>%
+  summarise(mean=mean(abundance), sd=sd(abundance))
+
+# Calculating the mean and standard deviation for contrasting season of Caulerpa cylindracea
+desulfo_station_season_whole <- inner_join(metadata, whole, by=c("ID"="Group")) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
+  filter(station=="FCaM" | station=="FCa") %>%
+  mutate(season=str_replace(season, "Winter", "Summer")) %>%
+  mutate(season=str_replace(season, "Spring", "Autumn")) %>%
+  group_by(station, season) %>%
+  summarise(mean=mean(abundance), sd=sd(abundance))
+```
+In contrast to previously described high rank taxa, *Desulfobacterota* were specific to *Caulerpa cylindracea*.
+On average they were comprising `r desulfo_all_whole$mean` ± `r desulfo_all_whole$sd` \si{\percent} of all sequences.
+While seawater and *Cymodocea nodosa* communities were consisted of only `r filter(desulfo_station_whole, station=="F")$mean` ± `r filter(desulfo_station_whole, station=="F")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCyM")$mean` ± `r filter(desulfo_station_whole, station=="FCyM")$sd` \si{\percent} *Desulfobacterota* sequences, respectively, in the invaded and noninvaded *Caulerpa cylindracea* communities their proportion was `r filter(desulfo_station_whole, station=="FCaM")$mean` ± `r filter(desulfo_station_whole, station=="FCaM")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCa")$mean` ± `r filter(desulfo_station_whole, station=="FCa")$sd` \si{\percent}, respectively (\autoref{desulfo}).
+In addition, *Caulerpa cylindracea* associated communities were characterized by higer proportions in Winter and Summer (invaded, `r filter(desulfo_station_season_whole, station=="FCaM", season=="Summer")$mean` ± `r filter(desulfo_station_season_whole, station=="FCaM", season=="Summer")$sd` \si{\percent}; noninvaded, `r filter(desulfo_station_season_whole, station=="FCa", season=="Summer")$mean` ± `r filter(desulfo_station_season_whole, station=="FCa", season=="Summer")$sd` \si{\percent}) in comparison to Autumn and Spring (invaded, `r filter(desulfo_station_season_whole, station=="FCaM", season=="Autumn")$mean` ± `r filter(desulfo_station_season_whole, station=="FCaM", season=="Autumn")$sd` \si{\percent}; noninvaded, `r filter(desulfo_station_season_whole, station=="FCa", season=="Autumn")$mean` ± `r filter(desulfo_station_season_whole, station=="FCa", season=="Autumn")$sd` \si{\percent}).
+The community was mainly consisted of no relative *Desulfobacteraceae*, *Desulfatitalea*, no relative *Desulfobulbaceae*, *Desulfobulbus*, no relative *Desulfocapsaceae*, *Desulfopila*, *Desulforhopalus*, *Desulfotalea*, SEEP-SRB4 and uncultured *Desulfocapsaceae* (\autoref{desulfo}).
 

### 2020-02-26 submission/manuscript.tex
Results done.

--- 2020-02-14 submission/manuscript.tex
@@
 al.}, 2019). While there is high community variation between host
-species is was observed that the majority of metagenomes determined
+species is was observed that the majority of metagenome determined
 functions were conserved both between host species and individuals
@@
 cylindracea} located in the proximity of the meadow. In addition, for
-comparison the community of the surrounding seawater was characterized.
+comparison, the community of the surrounding seawater was characterized.
 
@@
 located close to the invaded meadow. Sampling of leaves and thalli was
-performed approximately monthly from December 2017 to October 2018.
-Leaves and thalli were collected by diving and transported to the
-laboratory in containers placed on ice and filled with site seawater.
-Upon arrival to the laboratory, \emph{Cymodocea nodosa} leaves were cut
-into sections of 1 -- 2 \si{\cm}, while \emph{Caulerpa cylindracea}
-thalli were cut into 5 -- 8 \si{\cm} long sections. Leaves and thalli
-were washed three times with sterile artificial seawater (ASW) to remove
-loosely attached microbial cells. Surrounding seawater was collected in
-10 \si{\l} containers by diving and transported to the laboratory where
-the whole container volume was filtered through a 20 \si{\um} net. The
-filtrate was further sequentially filtered through 3 \si{\um} and 0.2
-\si{\um} polycarbonate membrane filters (Whatman, United Kingdom) using
-a peristaltic pump. Filters were briefly dried at room temperature and
-stored at \num{-80} \si{\degreeCelsius}. Seawater samples were also
-collected approximately monthly from July 2017 to October 2018.
+performed approximately monthly from December 2017 to October 2018
+(\autoref{supp-nseq_notus}). Leaves and thalli were collected by diving
+and transported to the laboratory in containers placed on ice and filled
+with site seawater. Upon arrival to the laboratory, \emph{Cymodocea
+nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while
+\emph{Caulerpa cylindracea} thalli were cut into 5 -- 8 \si{\cm} long
+sections. Leaves and thalli were washed three times with sterile
+artificial seawater (ASW) to remove loosely attached microbial cells.
+Surrounding seawater was collected in 10 \si{\l} containers by diving
+and transported to the laboratory where the whole container volume was
+filtered through a 20 \si{\um} net. The filtrate was further
+sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate
+membrane filters (Whatman, United Kingdom) using a peristaltic pump.
+Filters were briefly dried at room temperature and stored at \num{-80}
+\si{\degreeCelsius}. Seawater samples were also collected approximately
+monthly from July 2017 to October 2018.
 
@@
 cylindracea} was isolated using a previously modified and adapted
-protocol that allows a selective epiphytic DNA isolation (Massana
+protocol that allows for a selective epiphytic DNA isolation (Massana
 \emph{et al.}, 1997; Korlević \emph{et al.}, submitted). Briefly, leaves
@@
 (\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
-were used to amplify the target region (Caporaso \emph{et al.}, 2012;
-Apprill \emph{et al.}, 2015; Parada \emph{et al.}, 2016). These primers
-contained on their 5' end a tagged sequence. Purified PCR products were
-sent for Illumina MiSeq sequencing at IMGM Laboratories, Martinsried,
-Germany. Before sequencing, the second PCR amplification of the two-step
-PCR procedure was performed at IMGM using primers targeting the tagged
-region incorporated in the first PCR. In addition, these primers
-contained adapter and sample-specific index sequences. Beside samples, a
-positive and negative control for each sequencing batch were sequenced.
-A negative control was comprised of PCR reactions without DNA template,
-while for a positive controls a mock community composed of evenly mixed
-DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC,
-USA) was used. The sequences obtained in this study have been submitted
-to the European Nucleotide Archive (ENA) under accession numbers
-\textbf{TO BE ADDED LATER!}.
+were used (Caporaso \emph{et al.}, 2012; Apprill \emph{et al.}, 2015;
+Parada \emph{et al.}, 2016). These primers contained on their 5' end a
+tagged sequence. Purified PCR products were sent for Illumina MiSeq
+sequencing at IMGM Laboratories, Martinsried, Germany. Before sequencing
+at IMGM, the second PCR amplification of the two-step PCR procedure was
+performed using primers targeting the tagged region incorporated in the
+first PCR. In addition, these primers contained adapter and
+sample-specific index sequences. Beside samples, a positive and negative
+control for each sequencing batch was sequenced. Negative control was
+comprised of PCR reactions without DNA template, while for a positive
+control a mock community composed of evenly mixed DNA material
+originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
+used. The sequences obtained in this study have been submitted to the
+European Nucleotide Archive (ENA) under accession numbers \textbf{TO BE
+ADDED LATER!}.
 
@@
 Yilmaz \emph{et al.}, 2014). Pipeline data processing and visualization
-was done using R (version 3.6.0) (R Core Team, 2019), package vegan
-(version 2.5-6) (Oksanen \emph{et al.}, 2019), package tidyverse
-(version 1.3.0) (Wickham \emph{et al.}, 2019) and multiple other
-packages (Xie, 2014, 2015, 2019a, 2019b, 2020; Neuwirth, 2014; Xie
-\emph{et al.}, 2018; Allaire \emph{et al.}, 2019; Zhu, 2019). The
-detailed analysis procedure including the R Markdown file for this paper
-are available as a GitHub repository (\textbf{TO BE ADDED LATER!}).
-Based on the ATCC MSA-1002 mock community included in the analysis an
-average sequencing error rate of 0.01 \si{\percent} was determined,
-which is in line with previously reported values for next-generation
-sequencing data (Kozich \emph{et al.}, 2013; Schloss \emph{et al.},
-2016). In addition, the negative controls processed together with the
-samples yielded on average only 2 sequences after sequence quality
-curation.
+was done using R (version 3.6.0) (R Core Team, 2019), packages vegan
+(version 2.5-6) (Oksanen \emph{et al.}, 2019), and tidyverse (version
+1.3.0) (Wickham \emph{et al.}, 2019) and multiple other packages (Xie,
+2014, 2015, 2019a, 2019b, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018;
+Allaire \emph{et al.}, 2019; Zhu, 2019). The detailed analysis procedure
+including the R Markdown file for this paper are available as a GitHub
+repository (\textbf{TO BE ADDED LATER!}). Based on the ATCC MSA-1002
+mock community included in the analysis an average sequencing error rate
+of 0.01 \si{\percent} was determined, which is in line with previously
+reported values for next-generation sequencing data (Kozich \emph{et
+al.}, 2013; Schloss \emph{et al.}, 2016). In addition, the negative
+controls processed together with the samples yielded on average only 2
+sequences after sequence quality curation.
 
@@
 
-Chloroplast sequences were following the seasonal pattern
-(\autoref{supp-chloroplast}, \autoref{pcoa}).
+Sequencing of the 16S rRNA V4 region yielded a total of 1.8 million
+sequences after quality curation and exclusion of eukaryotic,
+chloroplast, mitochondrial and no relative sequences
+(\autoref{supp-nseq_notus}). A total of 35 samples originating from
+epiphytic archaeal and bacterial communities associated with surfaces of
+the seagrass \emph{Cymodocea nodosa} and macroalga \emph{Caulerpa
+cylindracea} were analyzed. In addition, 18 samples (one of the samples
+was sequenced two times) originating from picoplankton archaeal and
+bacterial communities in the surrounding seawater were also processed.
+The number of reads per sample ranged between 8,407 and 77,465 sequences
+(\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
+applied the rarefaction curves did not level off that is a common
+observation in high-throughput 16S rRNA amplicon sequencing approaches
+(\autoref{supp-rarefaction}). Following quality curation and exclusion
+of sequences mentioned before reads were clustered into 28,726 different
+OTUs at a similarity level of 97 \si{\percent}. Reads numbers were
+normalized to the minimum number of sequences, 8,407
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 16,963
+different OTUs that ranged from 365 to 2,019 OTUs per sample
+(\autoref{supp-calculators}). To determine seasonal changes of richness
+and diversity the Observed Number of OTUs, Chao1, ACE, Exponential
+Shannon (Jost, 2006) and Inverse Simpson were calculated after
+normalization through rarefaction. Generally, richness estimators and
+diversity indices showed similar trends. On average, higher values were
+found for \emph{Caulerpa cylindracea} (invaded {[}Number of OTUs,
+1,686.9 ± 135.8 OTUs{]} and noninvaded {[}Number of OTUs, 1,742.8 ±
+152.4 OTUs{]}), middle vales for \emph{Cymodocea nodosa} (Number of
+OTUs, 1,047.7 ± 212.1 OTUs) and lower values for picoplankton
+communities in the surrounding seawater (Number of OTUs, 528.4 ± 144.5
+OTUs) (\autoref{supp-calculators}). Seasonal changes did not show such
+large dissimilarities. Seawater communities richness was stable during
+the studied period with the exception of one sampling point in December
+2017 when larger values were observed. \emph{Cymodocea nodosa}
+communities showed a slow increase towards the end of the study, while
+\emph{Caulerpa cylindracea} (invaded and noninvaded) communities were
+characterized by slightly larger values in Spring and Summer in
+comparison to Autumn and Winter (\autoref{supp-calculators}).
+
+To determine the proportion of shared archaeal and bacterial OTUs and
+communities sampled in different environments the Jaccard's Similarity
+Coefficient on presence-absence data and Bray-Curtis Similarity
+Coefficient were, respectively, calculated. Coefficients were determined
+after normalization through rarefaction and binning of samples from a
+particular environment. The highest proportion of shared OTUs and
+community was found between invaded and noninvaded \emph{Caulerpa
+cylindracea} (Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared
+values were calculated between seawater and epiphytic communities
+(\autoref{matrix}). Shared proportion between \emph{Cymodocea nodosa}
+and \emph{Caulerpa cylindracea} were approximately in the middle between
+these two extremes. To assess seasonal changes in the proportion of
+shared OTUs and communities the Jaccard's and Bray-Curtis Similarity
+Coefficients were calculated between consecutive sampling points
+(\autoref{shared}). Both coefficients showed similar trends. Temporal
+proportional changes were more pronounced for seawater in comparison to
+\emph{Cymodocea nodosa} and especially \emph{Caulerpa cylindracea}
+associated communities (\autoref{shared}). To further disentangle the
+environmental and seasonal community dissimilarity a Principal
+Coordinates Analysis (PCoA) based on Bray-Curtis distances and OTU
+abundances was applied. It showed a clear separation between planktonic
+and surface associated communities (\autoref{pcoa}). In addition, a
+separation of epiphytic bacterial and archaeal communities based on host
+species was determined. This separation was further supported by ANOSIM
+(R = 0.96, \emph{p} \textless{} 0.001). Seasonal changes of seawater
+communities indicated a separation between Spring, Summer and
+Autumn/Winter samples (ANOSIM, R = 0.64, \emph{p} \textless{} 0.001).
+Epiphytic microbial communities associated with \emph{Cymodocea nodosa}
+showed a similar pattern (ANOSIM, R = 0.54, \emph{p} \textless{} 0.01),
+while communities from the surfaces of \emph{Caulerpa cylindracea}
+indicated a non so strongly supported, as in previous cases, separation
+between Summer and Autumn/Winter/Spring samples (ANOSIM, R = 0.30,
+\emph{p} \textless{} 0.01) (\autoref{pcoa}).
+
+The microbial taxonomic composition of both, macrophyte associated and
+seawater communities, was dominated by \emph{Bacteria} (99.1 ± 2.1
+\si{\percent}) over \emph{Archaea} (0.9 ± 2.1 \si{\percent})
+(\autoref{community}). Three phyla were characteristic for
+\emph{Archaea}: \emph{Crenarchaeota}, \emph{Nanoarchaeota} and
+\emph{Thermoplasmatota}. Their contribution was larger only in seawater
+communities during December 2017 (\autoref{community}). Higher relative
+abundances of chloroplast related sequences were only observed in
+surface associated communities, with higher values in Autumn/Winter
+(37.2 ± 11.2 \si{\percent}) in comparison to Spring/Summer (20.9 ± 9.7
+\si{\percent}) (\autoref{supp-chloroplast}). Generally, at higher
+taxonomic ranks (phylum-class) epiphytic and seawater microbial
+communities were composed of similar bacterial taxa. Seawater
+communities were mainly comprised of \emph{Actinobacteriota},
+\emph{Bacteroidota}, \emph{Cyanobacteria}, \emph{Alphaproteobacteria},
+\emph{Gammaproteobacteria} and \emph{Verrucomicrobiota}. Communities
+associated with \emph{Cymodocea nodosa} were consisted of same groups
+with the addition of \emph{Planctomycetota} whose contribution was
+higher in summer 2018. In addition, communities from invaded and
+noninvaded \emph{Caulerpa cylindracea} were similar and characterized by
+same groups as seawater and \emph{Cymodocea nodosa} communities with the
+addition of \emph{Desulfobacterota} (\autoref{community}). Larger
+differences between environments and host species could be observed at
+lower taxonomic ranks (\autoref{cyano} -- \ref{desulfo}).
+
+\emph{Cyanobacteria} related sequences were comprising, on average, 5.5
+± 4.4 \si{\percent} of total sequences (\autoref{cyano}). Higher
+proportions were found for \emph{Cymodocea nodosa} (16.4 ± 5.3
+\si{\percent}) and \emph{Caulerpa cylindracea} (invaded {[} (7.7 ± 3.9
+\si{\percent}){]} and noninvaded {[} (7.8 ± 2.4 \si{\percent}){]})
+associated communities in autumn and for seawater communities in winter
+(8.8 ± 7.4 \si{\percent}). Large taxonomic differences between surface
+associated and seawater cyanobacterial communities were observed.
+Seawater communities were mainly comprised of \emph{Cyanobium} and
+\emph{Synechococcus}, while surface associated communities were
+consisted of and sequences without known relatives within
+\emph{Cyanobacteriales} and \emph{Cyanobacteriia} (\autoref{cyano}). In
+addition, seasonal changes in surface associated communities were
+observed with \emph{Cyanobacteriales} and \emph{Cyanobacteriia} no
+relative sequences comprising larger proportions in autumn and winter
+and \emph{Acrophormium}, \emph{Phormidesmis} and no relative
+\emph{Nodosilineaceae} and in spring and summer (\autoref{cyano}).
+
+Sequences classified as \emph{Bacteroidota} were comprising, on average,
+19.2 ± 5.5 \si{\percent} of all sequences (\autoref{bactero}). Similarly
+to \emph{Cyanobacteria}, large differences in the taxonomic composition
+between seawater and surface associated communities were found
+(\autoref{bactero}). The seawater community was characterized by the NS4
+and NS5 marine groups, uncultured \emph{Cryomorphaceae}, uncultured
+\emph{Flavobacteriaceae}, NS11-12 marine group, \emph{Balneola},
+uncultured \emph{Balneolaceae} and \emph{Formosa}. In contrast, surface
+associated communities were characterized by \emph{Lewinella},
+\emph{Portibacter}, \emph{Rubidimonas}, no relative
+\emph{Saprospiraceae}, uncultured \emph{Saprospiraceae}, no relative
+\emph{Flavobacteriaceae} and uncultured \emph{Rhodothermaceae}. Surface
+associated \emph{Bacteroidota} communities were very diverse as could be
+observed in the the high proportion of taxa that grouped as other
+\emph{Bacteroidota} (\autoref{bactero}).
+
+On average, \emph{Alphaproteobacteria} were in comparison to other high
+rank taxa the largest taxonomic group, comprising 29.2 ± 12.0
+\si{\percent} of all sequences (\autoref{alpha}). In accordance to
+previous taxa, high differences between seawater and surface associated
+communities were observed. Picoplankton communities were composed mainly
+of the SAR11 clade, AEGEAN-169 marine group, SAR116 clade, no relative
+\emph{Rhodobacteraceae}, HIMB11 and OCS116 clade, while surface
+associated communities were composed in high proportion of no relative
+\emph{Rhodobacteraceae} and to a lesser degree of \emph{Pseudoahrensia},
+no relative \emph{Alphaproteobacteria} and no relative
+\emph{Hyphomonadaceae}. Representatives of no relative
+\emph{Rhodobacteraceae} were comprising in on average 40.6 ± 23.2
+\si{\percent} of all alphaproteobacterial sequences from the epiphytic
+community (\autoref{alpha}).
+
+Sequences related to \emph{Gammaproteobacteria} were comprising, on
+average, 18.7 ± 3.9 \si{\percent} of all sequences (\autoref{gamma}).
+Similarly to previous taxa, large taxonomic differences between seawater
+and surface associated communities were found. Seawater communities were
+mainly comprised of the OM60 (NOR5) clade, \emph{Litoricola},
+\emph{Acinetobacter} and the SAR86 clade, while epiphytic communities
+were mainly composed of no relative \emph{Gammaproteobacteria} and
+\emph{Granulosicocus}. Beside these two groups specific to all three
+epiphytic communities, \emph{Cymodocea nodosa} was characterized by
+\emph{Arenicella}, no relative \emph{Burkholderiales} and
+\emph{Methylotenera}, while \emph{Thioploca}, no relative
+\emph{Cellvibrionaceae} and \emph{Reinekea} were more specific to both
+invaded and noninvaded \emph{Caulerpa cylindracea}. In addition,
+similarly to \emph{Bacteroidota}, a large proportion of the surface
+associated community was grouped as other \emph{Gammaproteobacteria}
+indicating high diversity within this group (\autoref{gamma}).
+
+In contrast to previously described high rank taxa,
+\emph{Desulfobacterota} were specific to \emph{Caulerpa cylindracea}. On
+average they were comprising 11.2 ± 13.3 \si{\percent} of all sequences.
+While seawater and \emph{Cymodocea nodosa} communities were consisted of
+only 0.1 ± 0.08 \si{\percent} and 1.0 ± 0.7 \si{\percent}
+\emph{Desulfobacterota} sequences, respectively, in the invaded and
+noninvaded \emph{Caulerpa cylindracea} communities their proportion was
+25.7 ± 11.3 \si{\percent} and 24.0 ± 4.3 \si{\percent}, respectively
+(\autoref{desulfo}). In addition, \emph{Caulerpa cylindracea} associated
+communities were characterized by higer proportions in Winter and Summer
+(invaded, 30.9 ± 12.4 \si{\percent}; noninvaded, 26.9 ± 3.4
+\si{\percent}) in comparison to Autumn and Spring (invaded, 22.8 ± 10.3
+\si{\percent}; noninvaded, 21.9 ± 3.8 \si{\percent}). The community was
+mainly consisted of no relative \emph{Desulfobacteraceae},
+\emph{Desulfatitalea}, no relative \emph{Desulfobulbaceae},
+\emph{Desulfobulbus}, no relative \emph{Desulfocapsaceae},
+\emph{Desulfopila}, \emph{Desulforhopalus}, \emph{Desulfotalea},
+SEEP-SRB4 and uncultured \emph{Desulfocapsaceae} (\autoref{desulfo}).
 
@@
 \subsection{Figures}\label{figures}}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.7\linewidth]{../results/figures/matrix}
+
+}
+
+\caption{Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between communties associated with the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and coomunities in the surrounding seawater.\label{matrix}}\label{fig:unnamed-chunk-1}
+\end{figure}
 

### 2020-02-26 submission/supplementary.Rmd
Results done.

@@
 
-```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE}
+```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE, message=FALSE}
 opts_chunk$set("tidy" = TRUE)
@@
 knitr::knit_hooks$set(inline=inline_hook)
+
+library("kableExtra")
 ```
@@
 ```
+
+## Supplementary Table
+
+```{r nseq_notus}
+# Loading OTU/sample data
+shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")
+
+# Calculating the number of sequences and OTUs in each sample
+nseq_notus <- shared %>%
+  mutate(nseq=rowSums(select(., starts_with("Otu")))) %>%
+  mutate(notus=rowSums(select(., starts_with("Otu"))!=0)) %>%
+  select(Group, nseq, notus)
+
+# Loading metadata 
+metadata <- read_tsv("data/raw/metadata.csv")
+
+# Joining metadata with the number of sequences and OTUs
+invisible(Sys.setlocale(locale="en_GB.utf8"))
+metadata_nseq_notus <- inner_join(metadata, nseq_notus, by=c("ID"="Group")) %>%
+  select(ID, station, date, season, nseq, notus) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
+  mutate(date=format(date, "%d %B %Y")) %>%
+  mutate(date=str_replace(date, "^0", "")) %>%
+  arrange(match(station, c("F", "FCyM", "FCaM", "FCa"))) %>%
+  mutate(station=str_replace(station, "^F$", "Seawater")) %>%
+  mutate(station=str_replace(station, "^FCyM$", "\\\\textit{Cymodocea nodosa} (Invaded)")) %>%
+  mutate(station=str_replace(station, "^FCaM$", "\\\\textit{Caulerpa cylindracea} (Invaded)")) %>%
+  mutate(station=str_replace(station, "^FCa$", "\\\\textit{Caulerpa cylindracea} (Nonnvaded)")) %>%
+  mutate(ID=str_replace(ID, "^23_1$", "23a")) %>%
+  mutate(ID=str_replace(ID, "^23_2$", "23b")) %>%
+  mutate(nseq=format(nseq, big.mark=","), notus=format(notus, big.mark=","))
+
+# Generating the table
+kable(metadata_nseq_notus, format="latex", booktabs=TRUE, linesep="", longtable=TRUE, escape=FALSE, col.names=c("Sample ID", "Community Type", "Date", "Season", "No.
+of Sequences", "No.
+of OTUs"), caption="Sample ID, Community Type, Sampling Date and Season, No.
+of Sequences and No.
+of OTUs of each sample.\\label{nseq_notus}", align="c") %>%
+  kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9) %>%
+  row_spec(0, bold=TRUE) %>%
+  column_spec(1, width="6em")
+```

### 2020-02-26 submission/supplementary.tex
Results done.

@@
   \posttitle{\par}
-  \subtitle{\textbf{Dynamics of Epiphytic Microbial Communities on Marine Macrophyte
-Surfaces}}
+  \subtitle{\textbf{Seasonal Dynamics of Epiphytic Microbial Communities on Marine
+Macrophyte Surfaces}}
   \author{}
@@
 \renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
 \captionsetup{justification=raggedright,singlelinecheck=false}
@@
 \usepackage{float}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 
+\hypertarget{supplementary-table}{%
+\subsection{Supplementary Table}\label{supplementary-table}}
+
+\begingroup\fontsize{9}{11}\selectfont
+
+\begin{longtable}{>{\centering\arraybackslash}p{6em}ccccc}
+\caption{\label{tab:nseq_notus}Sample ID, Community Type, Sampling Date and Season, No. of Sequences and No. of OTUs of each sample.\label{nseq_notus}}\\
+\toprule
+\textbf{Sample ID} & \textbf{Community Type} & \textbf{Date} & \textbf{Season} & \textbf{No. of Sequences} & \textbf{No. of OTUs}\\
+\midrule
+\endfirsthead
+\caption[]{Sample ID, Community Type, Sampling Date and Season, No. of Sequences and No. of OTUs of each sample.\label{nseq_notus} \textit{(continued)}}\\
+\toprule
+\textbf{Sample ID} & \textbf{Community Type} & \textbf{Date} & \textbf{Season} & \textbf{No. of Sequences} & \textbf{No. of OTUs}\\
+\midrule
+\endhead
+\
+\endfoot
+\bottomrule
+\endlastfoot
+3 & Seawater & 13 July 2017 & Summer & 25,988 & 820\\
+5 & Seawater & 27 July 2017 & Summer & 58,955 & 993\\
+7 & Seawater & 10 August 2017 & Summer & 32,622 & 847\\
+9 & Seawater & 24 August 2017 & Summer & 60,932 & 1,060\\
+11 & Seawater & 19 September 2017 & Summer & 46,106 & 1,168\\
+13 & Seawater & 12 October 2017 & Autumn & 30,941 & 974\\
+15 & Seawater & 23 November 2017 & Autumn & 27,585 & 1,187\\
+17 & Seawater & 14 December 2017 & Autumn & 44,598 & 2,267\\
+19 & Seawater & 13 February 2018 & Winter & 44,185 & 1,313\\
+21 & Seawater & 27 March 2018 & Winter & 46,305 & 1,565\\
+23a & Seawater & 24 April 2018 & Spring & 30,962 & 1,005\\
+23b & Seawater & 24 April 2018 & Spring & 38,556 & 1,201\\
+25 & Seawater & 22 May 2018 & Spring & 53,876 & 1,787\\
+27 & Seawater & 19 June 2018 & Spring & 77,465 & 1,288\\
+29 & Seawater & 10 July 2018 & Summer & 50,799 & 1,452\\
+31 & Seawater & 9 August 2018 & Summer & 39,041 & 1,119\\
+33 & Seawater & 4 September 2018 & Summer & 36,195 & 1,206\\
+35 & Seawater & 5 October 2018 & Autumn & 49,584 & 1,013\\
+37 & \textit{Cymodocea nodosa} (Invaded) & 23 November 2017 & Autumn & 31,236 & 1,554\\
+41 & \textit{Cymodocea nodosa} (Invaded) & 4 December 2017 & Autumn & 24,237 & 1,175\\
+45 & \textit{Cymodocea nodosa} (Invaded) & 14 December 2017 & Autumn & 32,683 & 1,056\\
+49 & \textit{Cymodocea nodosa} (Invaded) & 13 February 2018 & Winter & 9,086 & 1,215\\
+52 & \textit{Cymodocea nodosa} (Invaded) & 27 March 2018 & Winter & 17,004 & 1,217\\
+55 & \textit{Cymodocea nodosa} (Invaded) & 24 April 2018 & Spring & 42,645 & 2,061\\
+58 & \textit{Cymodocea nodosa} (Invaded) & 22 May 2018 & Spring & 21,337 & 1,289\\
+61 & \textit{Cymodocea nodosa} (Invaded) & 19 June 2018 & Spring & 31,729 & 2,099\\
+64 & \textit{Cymodocea nodosa} (Invaded) & 10 July 2018 & Summer & 35,731 & 1,795\\
+67 & \textit{Cymodocea nodosa} (Invaded) & 9 August 2018 & Summer & 26,364 & 2,126\\
+70 & \textit{Cymodocea nodosa} (Invaded) & 4 September 2018 & Summer & 23,278 & 1,728\\
+73 & \textit{Cymodocea nodosa} (Invaded) & 5 October 2018 & Autumn & 29,911 & 2,209\\
+38 & \textit{Caulerpa cylindracea} (Invaded) & 23 November 2017 & Autumn & 36,316 & 2,900\\
+42 & \textit{Caulerpa cylindracea} (Invaded) & 4 December 2017 & Autumn & 28,373 & 3,253\\
+46 & \textit{Caulerpa cylindracea} (Invaded) & 14 December 2017 & Autumn & 34,711 & 3,059\\
+53 & \textit{Caulerpa cylindracea} (Invaded) & 27 March 2018 & Winter & 28,690 & 2,488\\
+56 & \textit{Caulerpa cylindracea} (Invaded) & 24 April 2018 & Spring & 34,784 & 3,067\\
+59 & \textit{Caulerpa cylindracea} (Invaded) & 22 May 2018 & Spring & 23,393 & 2,720\\
+62 & \textit{Caulerpa cylindracea} (Invaded) & 19 June 2018 & Spring & 36,483 & 3,322\\
+65 & \textit{Caulerpa cylindracea} (Invaded) & 10 July 2018 & Summer & 18,487 & 2,196\\
+68 & \textit{Caulerpa cylindracea} (Invaded) & 9 August 2018 & Summer & 31,960 & 3,112\\
+71 & \textit{Caulerpa cylindracea} (Invaded) & 4 September 2018 & Summer & 29,290 & 3,170\\
+74 & \textit{Caulerpa cylindracea} (Invaded) & 5 October 2018 & Autumn & 11,700 & 1,710\\
+39 & \textit{Caulerpa cylindracea} (Nonnvaded) & 23 November 2017 & Autumn & 22,081 & 2,437\\
+43 & \textit{Caulerpa cylindracea} (Nonnvaded) & 4 December 2017 & Autumn & 13,665 & 1,894\\
+47 & \textit{Caulerpa cylindracea} (Nonnvaded) & 14 December 2017 & Autumn & 8,407 & 1,465\\
+51 & \textit{Caulerpa cylindracea} (Nonnvaded) & 13 February 2018 & Winter & 11,674 & 1,909\\
+54 & \textit{Caulerpa cylindracea} (Nonnvaded) & 27 March 2018 & Winter & 39,470 & 3,143\\
+57 & \textit{Caulerpa cylindracea} (Nonnvaded) & 24 April 2018 & Spring & 20,300 & 2,831\\
+60 & \textit{Caulerpa cylindracea} (Nonnvaded) & 22 May 2018 & Spring & 33,042 & 3,312\\
+63 & \textit{Caulerpa cylindracea} (Nonnvaded) & 19 June 2018 & Spring & 41,844 & 3,978\\
+66 & \textit{Caulerpa cylindracea} (Nonnvaded) & 10 July 2018 & Summer & 27,037 & 2,693\\
+69 & \textit{Caulerpa cylindracea} (Nonnvaded) & 9 August 2018 & Summer & 26,738 & 3,142\\
+72 & \textit{Caulerpa cylindracea} (Nonnvaded) & 4 September 2018 & Summer & 31,868 & 3,267\\
+75 & \textit{Caulerpa cylindracea} (Nonnvaded) & 5 October 2018 & Autumn & 33,066 & 3,099\\*
+\end{longtable}
+\endgroup{}
+
 \end{document}

### 2020-02-26 code/plot_community_bar_plot_chloroplast.R

@@ -14,7 +14,7 @@ community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.uniq
   filter(!str_detect(taxon, "^Eukaryota")) %>%
   filter(taxon!="Root")
 
-# Remove and Mitochondria sequences and subtract their number from higher taxonomic levels to which
+# Remove mitochondrial sequences and subtract their number from higher taxonomic levels to which
 # they belong
 mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
 community <- mutate_at(community, 5:ncol(community), list(~case_when(

### 2020-02-26 code/plot_matrix.R

@@ -24,6 +24,9 @@ metadata <- read_tsv("data/raw/metadata.csv")
 
 # Joining metadata with OTU/sample data and summing sequences from each environment
 rarefied_metadata <- inner_join(rarefied, metadata, by=c("Group"="ID")) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
+  filter(date >= "2017-11-01") %>%
+  select_if(funs(!is.numeric(.) || sum(.)!=0)) %>%
   group_by(station) %>%
   select(starts_with("Otu")) %>%
   summarise_all(sum)

### 2020-02-26 code/plot_pcoa.R

@@ -11,11 +11,11 @@
 # Loading OTU/sample data
 shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")
 
-# Generation of random rarified community data
+# Generation of random rarefied community data
 rarefied <- shared %>%
   select(-label, -Group, -numOtus) %>%
   rrarefy(., min(rowSums(.))) %>%
-  as.tibble() %>%
+  as_tibble() %>%
   add_column("Group"=shared$Group, .before=TRUE) %>%
   select_if(list(~!is.numeric(.) || sum(.)!=0))
 
@@ -49,7 +49,7 @@ rarefied_metadata_select <- rarefied_metadata
 spe.bray <- vegdist(select(rarefied_metadata_select, starts_with("Otu")))
 spe.b.pcoa <- cmdscale(spe.bray, k=(nrow(rarefied_metadata_select)-1), eig=TRUE)
 coordinates <- spe.b.pcoa$points[,1:2] %>%
-  as.tibble() %>%
+  as_tibble() %>%
   add_column("Group"=rarefied_metadata_select$label, .before=TRUE)
 coordinates <- inner_join(metadata, coordinates, by=c("label"="Group"))
 
@@ -79,7 +79,7 @@ rarefied_metadata_select <- filter(rarefied_metadata, station=="F") %>%
 spe.bray <- vegdist(select(rarefied_metadata_select, starts_with("Otu")))
 spe.b.pcoa <- cmdscale(spe.bray, k=(nrow(rarefied_metadata_select)-1), eig=TRUE)
 coordinates <- spe.b.pcoa$points[,1:2] %>%
-  as.tibble() %>%
+  as_tibble() %>%
   add_column("Group"=rarefied_metadata_select$label, .before=TRUE)
 coordinates <- inner_join(metadata, coordinates, by=c("label"="Group"))
 
@@ -105,7 +105,7 @@ rarefied_metadata_select <- filter(rarefied_metadata, station=="FCyM") %>%
 spe.bray <- vegdist(select(rarefied_metadata_select, starts_with("Otu")))
 spe.b.pcoa <- cmdscale(spe.bray, k=(nrow(rarefied_metadata_select)-1), eig=TRUE)
 coordinates <- spe.b.pcoa$points[,1:2] %>%
-  as.tibble() %>%
+  as_tibble() %>%
[…]

### 2020-03-02 submission/manuscript.Rmd
Prist part of discussion.

@@
   group_by(season) %>%
-  summarise(mean=mean(S.obs), sd=sd(S.obs))
+  summarise(mean_S.obs=mean(S.obs), sd_S.obs=sd(S.obs),
+            mean_S.chao1=mean(S.chao1), sd_S.chao1=sd(S.chao1),
+            mean_S.ACE=mean(S.ACE), sd_S.ACE=sd(S.ACE),
+            mean_eshannon=mean(eshannon), sd_eshannon=sd(eshannon),
+            mean_invsimpson=mean(invsimpson), sd_invsimpson=sd(invsimpson))
 S.obs_season_fcym <- filter(estimators_indices_metadata, station=="FCyM") %>%
   group_by(season) %>%
-  summarise(mean=mean(S.obs), sd=sd(S.obs))
+  summarise(mean_S.obs=mean(S.obs), sd_S.obs=sd(S.obs),
+            mean_S.chao1=mean(S.chao1), sd_S.chao1=sd(S.chao1),
+            mean_S.ACE=mean(S.ACE), sd_S.ACE=sd(S.ACE),
+            mean_eshannon=mean(eshannon), sd_eshannon=sd(eshannon),
+            mean_invsimpson=mean(invsimpson), sd_invsimpson=sd(invsimpson))
 S.obs_season_fcam <- filter(estimators_indices_metadata, station=="FCaM") %>%
   group_by(season) %>%
-  summarise(mean=mean(S.obs), sd=sd(S.obs))
+  summarise(mean_S.obs=mean(S.obs), sd_S.obs=sd(S.obs),
+            mean_S.chao1=mean(S.chao1), sd_S.chao1=sd(S.chao1),
+            mean_S.ACE=mean(S.ACE), sd_S.ACE=sd(S.ACE),
+            mean_eshannon=mean(eshannon), sd_eshannon=sd(eshannon),
+            mean_invsimpson=mean(invsimpson), sd_invsimpson=sd(invsimpson))
 S.obs_season_fca <- filter(estimators_indices_metadata, station=="FCa") %>%
   group_by(season) %>%
-  summarise(mean=mean(S.obs), sd=sd(S.obs))
+  summarise(mean_S.obs=mean(S.obs), sd_S.obs=sd(S.obs),
+            mean_S.chao1=mean(S.chao1), sd_S.chao1=sd(S.chao1),
+            mean_S.ACE=mean(S.ACE), sd_S.ACE=sd(S.ACE),
+            mean_eshannon=mean(eshannon), sd_eshannon=sd(eshannon),
+            mean_invsimpson=mean(invsimpson), sd_invsimpson=sd(invsimpson))
 ```
@@
 
-The microbial taxonomic composition of both, macrophyte associated and seawater communities, was dominated by *Bacteria* (`r filter(archaea_bacteria, taxon=="Bacteria")$mean` ± `r filter(archaea_bacteria, taxon=="Bacteria")$sd` \si{\percent}) over *Archaea* (`r filter(archaea_bacteria, taxon=="Archaea")$mean` ± `r filter(archaea_bacteria, taxon=="Archaea")$sd` \si{\percent}) (\autoref{community}).
-Three phyla were characteristic for *Archaea*: *Crenarchaeota*, *Nanoarchaeota* and *Thermoplasmatota*.
-Their contribution  was larger only in seawater communities during December 2017 (\autoref{community}).
+The taxonomic composition of both, macrophyte associated and seawater communities, was dominated by *Bacteria* (`r filter(archaea_bacteria, taxon=="Bacteria")$mean` ± `r filter(archaea_bacteria, taxon=="Bacteria")$sd` \si{\percent}) over *Archaea* (`r filter(archaea_bacteria, taxon=="Archaea")$mean` ± `r filter(archaea_bacteria, taxon=="Archaea")$sd` \si{\percent}) (\autoref{community}).
 Higher relative abundances of chloroplast related sequences were only observed in surface associated communities, with higher values in Autumn/Winter (`r filter(chloroplast, season=="Winter")$mean` ± `r filter(chloroplast, season=="Winter")$sd` \si{\percent}) in comparison to Spring/Summer (`r filter(chloroplast, season=="Summer")$mean` ± `r filter(chloroplast, season=="Summer")$sd` \si{\percent}) (\autoref{supp-chloroplast}).
@@
 
-# Remove Chloroplast and Mitochondria sequences and subtract their number from higher taxonomic levels to which
+# Remove chloroplast and mitochondrial sequences and subtract their number from higher taxonomic levels to which
 # they belong
@@
 # Selecting the relative abundance of the targeted group in the whole community and tidying the obtained data
-whole <- filter(community,
-                taxlevel==2) %>%
+whole <- filter(community, taxlevel==2) %>%
   gather(key="Group", value="abundance", 6:ncol(.)) %>%
@@
 Large taxonomic differences between surface associated and seawater cyanobacterial communities were observed.
-Seawater communities were mainly comprised of *Cyanobium* and *Synechococcus*, while surface associated communities were consisted of and sequences without known relatives within *Cyanobacteriales* and *Cyanobacteriia* (\autoref{cyano}).
-In addition, seasonal changes in surface associated communities were observed with *Cyanobacteriales* and *Cyanobacteriia* no relative sequences comprising larger proportions in autumn and winter and *Acrophormium*, *Phormidesmis* and no relative *Nodosilineaceae* and in spring and summer (\autoref{cyano}).
+Seawater communities were mainly comprised of *Cyanobium* and *Synechococcus*, while surface associated communities were consisted of *Pleurocapsa* and sequences without known relatives within *Cyanobacteriia* (\autoref{cyano}).
+In addition, seasonal changes in surface associated communities were observed with *Pleurocapsa* and no relative *Cyanobacteriia* comprising larger proportions in autumn and winter and *Acrophormium*, *Phormidesmis* and no relative *Nodosilineaceae* in spring and summer (\autoref{cyano}).
 
@@
 The seawater community was characterized by the NS4 and NS5 marine groups, uncultured *Cryomorphaceae*, uncultured *Flavobacteriaceae*, NS11-12 marine group, *Balneola*, uncultured *Balneolaceae* and *Formosa*.
-In contrast, surface associated communities were characterized by *Lewinella*, *Portibacter*, *Rubidimonas*, no relative *Saprospiraceae*, uncultured *Saprospiraceae*, no relative *Flavobacteriaceae* and uncultured *Rhodothermaceae*.
+In contrast, in surface associated communities *Lewinella*, *Portibacter*, *Rubidimonas*, no relative *Saprospiraceae*, uncultured *Saprospiraceae*, no relative *Flavobacteriaceae* and uncultured *Rhodothermaceae* were found.
+Some groups showed slight seasonal changes such as no relative *Flavobacteriaceae* that were more pronounced from November 2017 until June 2018.
+In contrast, uncultured *Rhodothermaceae* showed higher proportions from June 2018 until the end of the study period.
 Surface associated *Bacteroidota* communities were very diverse as could be observed in the the high proportion of taxa that grouped as other *Bacteroidota* (\autoref{bactero}).
@@
 # Selecting the relative abundance of the targeted group in the whole community and tidying the obtained data
-whole <- filter(community,
-                taxlevel==3) %>%
+whole <- filter(community, taxlevel==3) %>%
   gather(key="Group", value="abundance", 6:ncol(.)) %>%
@@
 In accordance to previous taxa, high differences between seawater and surface associated communities were observed.
-Picoplankton communities were composed mainly of the SAR11 clade, AEGEAN-169 marine group, SAR116 clade, no relative *Rhodobacteraceae*, HIMB11 and OCS116 clade, while surface associated communities were composed in high proportion of no relative *Rhodobacteraceae* and to a lesser degree of  *Pseudoahrensia*, no relative *Alphaproteobacteria* and no relative  *Hyphomonadaceae*.
-Representatives of no relative *Rhodobacteraceae* were comprising in on average `r rhodo$mean` ± `r rhodo$sd` \si{\percent} of all alphaproteobacterial sequences from the epiphytic community (\autoref{alpha}).
+Picoplankton communities were composed mainly of the SAR11 clade, AEGEAN-169 marine group, SAR116 clade, no relative *Rhodobacteraceae*, HIMB11 and OCS116 clade, while surface associated communities were composed in high proportion of no relative *Rhodobacteraceae* and to a lesser degree of  *Pseudoahrensia*, no relative *Alphaproteobacteria*, no relative  *Hyphomonadaceae* and *Amylibacter*.
+Representatives of no relative *Rhodobacteraceae* were comprising on average `r rhodo$mean` ± `r rhodo$sd` \si{\percent} of all alphaproteobacterial sequences from the epiphytic community (\autoref{alpha}).
+In addition, *Amylibacter* was detected mainly in *Cymodocea nodosa* from November 2017 until March 2018.
 
@@
 Beside these two groups specific to all three epiphytic communities, *Cymodocea nodosa* was characterized by *Arenicella*, no relative *Burkholderiales* and *Methylotenera*, while *Thioploca*, no relative *Cellvibrionaceae* and *Reinekea* were more specific to both invaded and noninvaded *Caulerpa cylindracea*.
+In addition, *Arenicella* was more pronounced in November and December 2017, while no relative *Burkholderiales* and *Methylotenera* were more characteristic for the period form March until May 2018.
+For the *Caulerpa cylindracea* specific taxa no relative *Cellvibrionaceae* and *Reinekea* showed some seasonality and were characterisitic for samples originating from June to October 2018.
 In addition, similarly to *Bacteroidota*, a large proportion of the surface associated community was grouped as other *Gammaproteobacteria* indicating high diversity within this group (\autoref{gamma}).
@@
 ```
+
 In contrast to previously described high rank taxa, *Desulfobacterota* were specific to *Caulerpa cylindracea*.
@@
 ## Discussion
+Surfaces of marine macrophytes are harboring biofilms consisted of diverse microbial taxa [@Egan2013; @Tarquinio2019].
+No standard protocol has been developed to study these macophyte associated microbes [@Ugarelli2019].
+Different procedures of microbial cells removal from host surfaces were described, such as host tissue shaking [@Noges2010], scraping [@Uku2007] and ultrasonication [@Cai2014].
+All these methods showed different removal efficiencies and none was enabling a complete removal of attached microbial cells.
+In the present study, we applied an earlier developed removal protocol [@Korlevic_in_press, submitted], based on a previous idea of direct cellular lysis [@Burke2009], to ensure an almost complete cell detachment.
+The application of a direct lysis procedure coupled with a high frequency sampling protocol and Illumina high resolution amplicon sequencing has enabled us to make a detailed description of bacterial and archaeal communities associated with the surfaces of two marine macrophytes, *Cymodocea nodosa* and *Caulerpa cylindracea*.
+
+Observed highest richness values for *Caulerpa cylindracea* (invaded and noninvaded), middle for *Cymodocea nodosa* and lowest for seawater derived communities.
+Higher values for surface associated communities in comparison to seawater were described earlier for seagrasses [@Ugarelli2019] and could be attributed to a larger set of inhabitable microniches existing on macrophyte surfaces.
+In addition, highest values observed for *Caulerpa cylindracea* are probably a consequence of sediment derived OTUs that were present only in *Caulerpa cylindracea* communities.
+*Caulerpa cylindracea* stolon is attached to surface sediment with rhizoids, so the stolon and rhizoids are in a direct contact with the sediment surface.
+Part of the surface attached *Caulerpa cylnidracea* community is therefore comprised of sediment derived cells that could cause the observed increase in richness.
+In addition, seasonal richness differences observed for surface attached communities showed slightly higher values in spring and summer could be explained by a higher macrophyte growth in these seasons [M.
+Najdek, personal communication; @Zavodnik1998; @Ruitton2005].
+During active periods macrophytes exhibit a more dynamic chemical interaction with the surface community causing an opening of a larger set of microniches.
 

### 2020-03-02 submission/manuscript.tex
Prist part of discussion.

@@
 normalized to the minimum number of sequences, 8,407
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 16,963
-different OTUs that ranged from 365 to 2,019 OTUs per sample
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,007
+different OTUs that ranged from 366 to 1,998 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 found for \emph{Caulerpa cylindracea} (invaded {[}Number of OTUs,
-1,686.9 ± 135.8 OTUs{]} and noninvaded {[}Number of OTUs, 1,742.8 ±
-152.4 OTUs{]}), middle vales for \emph{Cymodocea nodosa} (Number of
-OTUs, 1,047.7 ± 212.1 OTUs) and lower values for picoplankton
-communities in the surrounding seawater (Number of OTUs, 528.4 ± 144.5
+1,688.5 ± 125.6 OTUs{]} and noninvaded {[}Number of OTUs, 1,744.8 ±
+150.8 OTUs{]}), middle vales for \emph{Cymodocea nodosa} (Number of
+OTUs, 1,062.8 ± 209.6 OTUs) and lower values for picoplankton
+communities in the surrounding seawater (Number of OTUs, 528.1 ± 135.8
 OTUs) (\autoref{supp-calculators}). Seasonal changes did not show such
@@
 Epiphytic microbial communities associated with \emph{Cymodocea nodosa}
-showed a similar pattern (ANOSIM, R = 0.54, \emph{p} \textless{} 0.01),
+showed a similar pattern (ANOSIM, R = 0.56, \emph{p} \textless{} 0.01),
 while communities from the surfaces of \emph{Caulerpa cylindracea}
 indicated a non so strongly supported, as in previous cases, separation
-between Summer and Autumn/Winter/Spring samples (ANOSIM, R = 0.30,
-\emph{p} \textless{} 0.01) (\autoref{pcoa}).
+between Summer and Autumn/Winter/Spring samples (ANOSIM, R = 0.31,
+\emph{p} \textless{} 0.05) (\autoref{pcoa}).
 
-The microbial taxonomic composition of both, macrophyte associated and
-seawater communities, was dominated by \emph{Bacteria} (99.1 ± 2.1
-\si{\percent}) over \emph{Archaea} (0.9 ± 2.1 \si{\percent})
-(\autoref{community}). Three phyla were characteristic for
-\emph{Archaea}: \emph{Crenarchaeota}, \emph{Nanoarchaeota} and
-\emph{Thermoplasmatota}. Their contribution was larger only in seawater
-communities during December 2017 (\autoref{community}). Higher relative
-abundances of chloroplast related sequences were only observed in
-surface associated communities, with higher values in Autumn/Winter
-(37.2 ± 11.2 \si{\percent}) in comparison to Spring/Summer (20.9 ± 9.7
-\si{\percent}) (\autoref{supp-chloroplast}). Generally, at higher
-taxonomic ranks (phylum-class) epiphytic and seawater microbial
+The taxonomic composition of both, macrophyte associated and seawater
+communities, was dominated by \emph{Bacteria} (99.1 ± 2.1 \si{\percent})
+over \emph{Archaea} (0.9 ± 2.1 \si{\percent}) (\autoref{community}).
+Higher relative abundances of chloroplast related sequences were only
+observed in surface associated communities, with higher values in
+Autumn/Winter (37.2 ± 11.2 \si{\percent}) in comparison to Spring/Summer
+(20.9 ± 9.7 \si{\percent}) (\autoref{supp-chloroplast}). Generally, at
+higher taxonomic ranks (phylum-class) epiphytic and seawater microbial
 communities were composed of similar bacterial taxa. Seawater
@@
 \emph{Synechococcus}, while surface associated communities were
-consisted of and sequences without known relatives within
-\emph{Cyanobacteriales} and \emph{Cyanobacteriia} (\autoref{cyano}). In
-addition, seasonal changes in surface associated communities were
-observed with \emph{Cyanobacteriales} and \emph{Cyanobacteriia} no
-relative sequences comprising larger proportions in autumn and winter
-and \emph{Acrophormium}, \emph{Phormidesmis} and no relative
-\emph{Nodosilineaceae} and in spring and summer (\autoref{cyano}).
+consisted of \emph{Pleurocapsa} and sequences without known relatives
+within \emph{Cyanobacteriia} (\autoref{cyano}). In addition, seasonal
+changes in surface associated communities were observed with
+\emph{Pleurocapsa} and no relative \emph{Cyanobacteriia} comprising
+larger proportions in autumn and winter and \emph{Acrophormium},
+\emph{Phormidesmis} and no relative \emph{Nodosilineaceae} in spring and
+summer (\autoref{cyano}).
 
@@
 \emph{Flavobacteriaceae}, NS11-12 marine group, \emph{Balneola},
-uncultured \emph{Balneolaceae} and \emph{Formosa}. In contrast, surface
-associated communities were characterized by \emph{Lewinella},
-\emph{Portibacter}, \emph{Rubidimonas}, no relative
-\emph{Saprospiraceae}, uncultured \emph{Saprospiraceae}, no relative
-\emph{Flavobacteriaceae} and uncultured \emph{Rhodothermaceae}. Surface
-associated \emph{Bacteroidota} communities were very diverse as could be
-observed in the the high proportion of taxa that grouped as other
+uncultured \emph{Balneolaceae} and \emph{Formosa}. In contrast, in
+surface associated communities \emph{Lewinella}, \emph{Portibacter},
+\emph{Rubidimonas}, no relative \emph{Saprospiraceae}, uncultured
+\emph{Saprospiraceae}, no relative \emph{Flavobacteriaceae} and
+uncultured \emph{Rhodothermaceae} were found. Some groups showed slight
+seasonal changes such as no relative \emph{Flavobacteriaceae} that were
+more pronounced from November 2017 until June 2018. In contrast,
+uncultured \emph{Rhodothermaceae} showed higher proportions from June
+2018 until the end of the study period. Surface associated
+\emph{Bacteroidota} communities were very diverse as could be observed
+in the the high proportion of taxa that grouped as other
 \emph{Bacteroidota} (\autoref{bactero}).
@@
 \emph{Rhodobacteraceae} and to a lesser degree of \emph{Pseudoahrensia},
-no relative \emph{Alphaproteobacteria} and no relative
-\emph{Hyphomonadaceae}. Representatives of no relative
-\emph{Rhodobacteraceae} were comprising in on average 40.6 ± 23.2
+no relative \emph{Alphaproteobacteria}, no relative
+\emph{Hyphomonadaceae} and \emph{Amylibacter}. Representatives of no
+relative \emph{Rhodobacteraceae} were comprising on average 40.6 ± 23.2
 \si{\percent} of all alphaproteobacterial sequences from the epiphytic
-community (\autoref{alpha}).
+community (\autoref{alpha}). In addition, \emph{Amylibacter} was
+detected mainly in \emph{Cymodocea nodosa} from November 2017 until
+March 2018.
 
@@
 invaded and noninvaded \emph{Caulerpa cylindracea}. In addition,
-similarly to \emph{Bacteroidota}, a large proportion of the surface
-associated community was grouped as other \emph{Gammaproteobacteria}
-indicating high diversity within this group (\autoref{gamma}).
+\emph{Arenicella} was more pronounced in November and December 2017,
+while no relative \emph{Burkholderiales} and \emph{Methylotenera} were
+more characteristic for the period form March until May 2018. For the
+\emph{Caulerpa cylindracea} specific taxa no relative
+\emph{Cellvibrionaceae} and \emph{Reinekea} showed some seasonality and
+were characterisitic for samples originating from June to October 2018.
+In addition, similarly to \emph{Bacteroidota}, a large proportion of the
+surface associated community was grouped as other
+\emph{Gammaproteobacteria} indicating high diversity within this group
+(\autoref{gamma}).
 
@@
 \subsection{Discussion}\label{discussion}}
+
+Surfaces of marine macrophytes are harboring biofilms consisted of
+diverse microbial taxa (Egan \emph{et al.}, 2013; Tarquinio \emph{et
+al.}, 2019). No standard protocol has been developed to study these
+macophyte associated microbes (Ugarelli \emph{et al.}, 2019). Different
+procedures of microbial cells removal from host surfaces were described,
+such as host tissue shaking (Nõges \emph{et al.}, 2010), scraping (Uku
+\emph{et al.}, 2007) and ultrasonication (Cai \emph{et al.}, 2014). All
+these methods showed different removal efficiencies and none was
+enabling a complete removal of attached microbial cells. In the present
+study, we applied an earlier developed removal protocol (Korlević
+\emph{et al.}, submitted), based on a previous idea of direct cellular
+lysis (Burke \emph{et al.}, 2009), to ensure an almost complete cell
+detachment. The application of a direct lysis procedure coupled with a
+high frequency sampling protocol and Illumina high resolution amplicon
+sequencing has enabled us to make a detailed description of bacterial
+and archaeal communities associated with the surfaces of two marine
+macrophytes, \emph{Cymodocea nodosa} and \emph{Caulerpa cylindracea}.
+
+Observed highest richness values for \emph{Caulerpa cylindracea}
+(invaded and noninvaded), middle for \emph{Cymodocea nodosa} and lowest
+for seawater derived communities. Higher values for surface associated
+communities in comparison to seawater were described earlier for
+seagrasses (Ugarelli \emph{et al.}, 2019) and could be attributed to a
+larger set of inhabitable microniches existing on macrophyte surfaces.
+In addition, highest values observed for \emph{Caulerpa cylindracea} are
+probably a consequence of sediment derived OTUs that were present only
+in \emph{Caulerpa cylindracea} communities. \emph{Caulerpa cylindracea}
+stolon is attached to surface sediment with rhizoids, so the stolon and
+rhizoids are in a direct contact with the sediment surface. Part of the
+surface attached \emph{Caulerpa cylnidracea} community is therefore
+comprised of sediment derived cells that could cause the observed
+increase in richness. In addition, seasonal richness differences
+observed for surface attached communities showed slightly higher values
+in spring and summer could be explained by a higher macrophyte growth in
+these seasons (M. Najdek, personal communication; Zavodnik \emph{et
+al.}, 1998; Ruitton \emph{et al.}, 2005). During active periods
+macrophytes exhibit a more dynamic chemical interaction with the surface
+community causing an opening of a larger set of microniches.
 
@@
 
+\leavevmode\hypertarget{ref-Burke2009}{}%
+Burke, C., Kjelleberg, S., and Thomas, T. (2009) Selective extraction of
+bacterial DNA from the surfaces of macroalgae. \emph{Applied and
+environmental microbiology} \textbf{75}: 252--256.
+
 \leavevmode\hypertarget{ref-Burke2011a}{}%
@@
 journal} \textbf{5}: 590--600.
+
+\leavevmode\hypertarget{ref-Cai2014}{}%
+Cai, X., Gao, G., Yang, J., Tang, X., Dai, J., Chen, D., and Song, Y.
+(2014) An ultrasonic method for separation of epiphytic microbes from
+freshwater submerged macrophytes. \emph{Journal of basic microbiology}
+\textbf{54}: 758--761.
 
@@
 
+\leavevmode\hypertarget{ref-Noges2010}{}%
+Nõges, T., Luup, H., and Feldmann, T. (2010) Primary production of
+aquatic macrophytes and their epiphytes in two shallow lakes (Peipsi and
+Võrtsjärv) in Estonia. \emph{Aquatic Ecology} \textbf{44}: 83--92.
+
 \leavevmode\hypertarget{ref-Oksanen2019}{}%
@@
 microbiology} \textbf{18}: 4391--4402.
+
+\leavevmode\hypertarget{ref-Ruitton2005}{}%
+Ruitton, S., Verlaque, M., and Boudouresque, C.F. (2005) Seasonal
+changes of the introduced Caulerpa racemosa var. cylindracea
+(Caulerpales, Chlorophyta) at the northwest limit of its Mediterranean
+range. \emph{Aquatic Botany} \textbf{82}: 55--70.
 
@@
 sites. \emph{Microorganisms} \textbf{7}:
+
+\leavevmode\hypertarget{ref-Uku2007}{}%
+Uku, J., Björk, M., Bergman, B., and Díez, B. (2007) Characterization
+and comparison of prokaryotic epiphytes associated with three East
+African seagrasses. \emph{Journal of Phycology} \textbf{43}: 768--779.
 
@@
 D643--8.
+
+\leavevmode\hypertarget{ref-Zavodnik1998}{}%
+Zavodnik, N., Travizi, A., and De Rosa, S. (1998) Seasonal variations in
+the rate of pliotosynthetic activity and chemical composition of the
+seagrass Cymodocea nodosa (Ucr.) Asch. \emph{Scientia Marina}
+\textbf{62}: 301--309.
 

### 2020-03-02 code/plot_community_bar_plot_cyanobacteria.R

@@ -14,7 +14,7 @@ community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.uniq
   filter(!str_detect(taxon, "^Eukaryota")) %>%
   filter(taxon!="Root")
 
-# Remove Chloroplast and Mitochondria sequences and subtract their number from higher taxonomic levels to which
+# Remove chloroplast and mitochondrial sequences and subtract their number from higher taxonomic levels to which
 # they belong
 chloroplast <- filter(community, str_detect(taxon, "^Chloroplast$"))$rankID
 mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID

### 2020-03-02 submission/manuscript.tex
OTU taxonomy file added to .gitignore.

@@
 normalized to the minimum number of sequences, 8,407
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,007
-different OTUs that ranged from 366 to 1,998 OTUs per sample
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,020
+different OTUs that ranged from 346 to 2,010 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 found for \emph{Caulerpa cylindracea} (invaded {[}Number of OTUs,
-1,688.5 ± 125.6 OTUs{]} and noninvaded {[}Number of OTUs, 1,744.8 ±
-150.8 OTUs{]}), middle vales for \emph{Cymodocea nodosa} (Number of
-OTUs, 1,062.8 ± 209.6 OTUs) and lower values for picoplankton
-communities in the surrounding seawater (Number of OTUs, 528.1 ± 135.8
+1,693.5 ± 122.3 OTUs{]} and noninvaded {[}Number of OTUs, 1,745.5 ±
+150.9 OTUs{]}), middle vales for \emph{Cymodocea nodosa} (Number of
+OTUs, 1,056.7 ± 212.5 OTUs) and lower values for picoplankton
+communities in the surrounding seawater (Number of OTUs, 531.7 ± 142.7
 OTUs) (\autoref{supp-calculators}). Seasonal changes did not show such
@@
 community was found between invaded and noninvaded \emph{Caulerpa
-cylindracea} (Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared
+cylindracea} (Jaccard, 0.36; Bray-Curtis, 0.77), while lower shared
 values were calculated between seawater and epiphytic communities
@@
 communities indicated a separation between Spring, Summer and
-Autumn/Winter samples (ANOSIM, R = 0.64, \emph{p} \textless{} 0.001).
+Autumn/Winter samples (ANOSIM, R = 0.63, \emph{p} \textless{} 0.001).
 Epiphytic microbial communities associated with \emph{Cymodocea nodosa}
-showed a similar pattern (ANOSIM, R = 0.56, \emph{p} \textless{} 0.01),
+showed a similar pattern (ANOSIM, R = 0.54, \emph{p} \textless{} 0.001),
 while communities from the surfaces of \emph{Caulerpa cylindracea}
@@
 \leavevmode\hypertarget{ref-Armstrong2000}{}%
-Armstrong, E., Rogerson, A., and Leftley, J.W. (2000) The Abundance of
+Armstrong, E., Rogerson, A., and Leftley, J. (2000) The Abundance of
 Heterotrophic Protists Associated with Intertidal Seaweeds.

### 2020-03-02 submission/manuscript.Rmd
Desulfobacterota removal checked.

@@
 Najdek, personal communication; @Zavodnik1998; @Ruitton2005].
-During active periods macrophytes exhibit a more dynamic chemical interaction with the surface community causing an opening of a larger set of microniches.
+During active periods macrophytes exhibit a more dynamic chemical interaction with the surface community causing an extension of the microniche number.
 

### 2020-03-02 submission/manuscript.tex
Desulfobacterota removal checked.

--- 2020-03-02 submission/manuscript.tex
@@
 macrophytes exhibit a more dynamic chemical interaction with the surface
-community causing an opening of a larger set of microniches.
+community causing an extension of the microniche number.
 

### 2020-03-02 code/plot_matrix.R

@@ -17,7 +17,7 @@ rarefied <- shared %>%
   rrarefy(., min(rowSums(.))) %>%
   as.tibble() %>%
   add_column("Group"=shared$Group, .before=TRUE) %>%
-  select_if(funs(!is.numeric(.) || sum(.)!=0))
+  select_if(list(~ !is.numeric(.) || sum(.)!=0))
 
 # Loading metadata 
 metadata <- read_tsv("data/raw/metadata.csv")

### 2020-03-02 code/plot_rarefaction.R

@@ -8,7 +8,7 @@
 #
 #################################################################################################################
 
-# Loading input data and selction of values for plotting
+# Loading input data and selection of values for plotting
 rarefaction <- read_tsv(file="data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.groups.rarefaction") %>%
   select(-contains("lci-"), -contains("hci-")) %>%
   gather(-numsampled, key=sample, value=sobs) %>%

### 2020-03-02 code/plot_seasonal_shared.R

@@ -17,7 +17,7 @@ rarefied <- shared %>%
   rrarefy(., min(rowSums(.))) %>%
   as.tibble() %>%
   add_column("Group"=shared$Group, .before=TRUE) %>%
-  select_if(funs(!is.numeric(.) || sum(.)!=0))
+  select_if(list(~ !is.numeric(.) || sum(.)!=0))
 
 # Copying the sample labels to the rows (input for library vegan)
 row.names(rarefied) <- rarefied$Group
@@ -35,7 +35,8 @@ bray <- vegdist(rarefied, method="bray", binary=F)
 bray <- as.tibble(data.frame(t(combn(rownames(rarefied),2)), as.numeric(bray))) %>%
   rename(V1=X1, V2=X2, bray=as.numeric.bray.)
 
-distance <- inner_join(jaccard, bray, by=c("V1"="V1", "V2"="V2"))
+distance <- inner_join(jaccard, bray, by=c("V1"="V1", "V2"="V2")) %>%
+  mutate_at(c("V1", "V2"), list(~ as.character(.)))
 
 # Loading metadata 
 metadata <- read_tsv("data/raw/metadata.csv")
@@ -109,7 +110,7 @@ p2 <- filter(data, index=="jaccard") %>%
   labs(x="", y="%") +
   theme
 
-f <- cowplot::plot_grid(p1, p2, nrow=2, ncol=1, rel_heights=c(1,0.75), align="v")
+f <- cowplot::plot_grid(p1, p2, nrow=2, ncol=1, rel_heights=c(1, 0.75), align="v")
 
 # Plots generation
 # Cymodocea nodosa samples
@@ -147,11 +148,11 @@ p2 <- filter(data, index=="jaccard") %>%
   scale_shape_manual(values=shapes) +
   scale_fill_manual(values=fills) +
   scale_x_date(date_break ="months" , date_labels="%b %Y") +
-  scale_y_continuous(limits=c(20, 40)) +
+  scale_y_continuous(limits=c(20, 50)) +
   labs(x="", y="") +
   theme
 
-fcym <- cowplot::plot_grid(p1, p2, nrow=2, ncol=1, rel_heights=c(1,0.75), align="v")
+fcym <- cowplot::plot_grid(p1, p2, nrow=2, ncol=1, rel_heights=c(1, 0.75), align="v")
 
 # Plots generation
 # Caulerpa cylindracea (Invaded) samples
[…]

### 2020-03-04 submission/manuscript.Rmd
Discussion progress.

@@
 The majority of studies describing macrophyte epiphytic communities did not encompass seasonal changes [@Ugarelli2019; @Lachnit2009; @Crump2008; @Burke2011b; @Roth-Schulze2016].
-In addition, if seasonal changes were taken into account low temporal frequency and/or methodologies that do not allow for high taxonomic resolution were used [@Tujula2010; @Lachnit2011; @Miranda2013].
+In addition, if seasonal changes were taken into account low temporal frequency and/or methodologies that do not allow for high taxonomic resolution were used [@Tujula2010; @Lachnit2011; @Miranda2013; @Michelou2013].
 In the present study we describe the seasonal dynamics of bacterial and archaeal communities on the surfaces of the seagrass *Cymodocea nodosa* and siphonous macroalgae *Caulerpa cylindracea* determined on a mostly monthly scale.
@@
 \newpage
-## Materials and Methods
-### Sampling
-Leaves of *Cymodocea nodosa* were sampled in a *Cymodocea nodosa* meadow located in the proximity of the village of Funtana (\ang{45;10;39} N, \ang{13;35;42} E).
-Thalli of *Caulerpa cylindracea* were sampled in the same *Cymodocea nodosa* invaded meadow in Funtana and on a locality of only *Caulerpa cylindracea* located close to the invaded meadow.
-Sampling of leaves and thalli was performed approximately monthly from December 2017 to October 2018 (\autoref{supp-nseq_notus}).
-Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with site seawater.
-Upon arrival to the laboratory, *Cymodocea nodosa* leaves were cut into sections of 1 -- 2 \si{\cm}, while *Caulerpa cylindracea* thalli were cut into 5 -- 8 \si{\cm} long sections.
-Leaves and thalli were washed three times with sterile artificial seawater (ASW) to remove loosely attached microbial cells.
-Surrounding seawater was collected in 10 \si{\l} containers by diving and transported to the laboratory where the whole container volume was filtered through a 20 \si{\um} net.
-The filtrate was further sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate membrane filters (Whatman, United Kingdom) using a peristaltic pump.
-Filters were briefly dried at room temperature and stored at \num{-80} \si{\degreeCelsius}.
-Seawater samples were also collected approximately monthly from July 2017 to October 2018.
-
-### DNA Isolation
-DNA from surfaces of *Cymodocea nodosa* and *Caulerpa cylindracea* was isolated using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevic_in_press, submitted; @Massana1997].
-Briefly, leaves and thalli are incubated in a lysis buffer and treated with lysozyme and proteinase K.
-Following the incubations, the mixture containing lyzed epiphytic cells is separated from leaves and thalli and extracted using a phenol-chloroform procedure.
-Finally, the extracted DNA is precipitated using isopropanol.
-DNA from seawater picoplankton was isolated from 0.2 \si{\um} polycarbonate filters according to [@Massana1997] with a slight modification.
-Following the phenol-chloroform extraction steps 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added.
-DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 20,000 × g and 4 \si{\degreeCelsius} for 21 \si{\minute}.
-The pellet was washed twice with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}.
-Dried pellets were resuspended in 50 -- 100 \si{\ul} of deionized water.
-
-### Illumina 16S rRNA Sequencing
-Illumina MiSeq sequencing of the V4 16S rRNA region was performed as described previously [@Korlevic_in_press, submitted].
-The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
-In the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used [@Caporaso2012; @Apprill2015; @Parada2016].
-These primers contained on their 5' end a tagged sequence.
-Purified PCR products were sent for Illumina MiSeq sequencing at IMGM Laboratories, Martinsried, Germany.
-Before sequencing at IMGM, the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR.
-In addition, these primers contained adapter and sample-specific index sequences.
-Beside samples, a positive and negative control for each sequencing batch was sequenced.
-Negative control was comprised of PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
-The sequences obtained in this study have been submitted to the European Nucleotide Archive (ENA) under accession numbers **TO BE ADDED LATER!**.
-
-### Sequence Analysis
-```{r error}
-# Calculation of the error rate based on the ATCC MSA-1002 mock community
-error <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.error.summary", na="") %>%
-  filter(numparents==1) %>%
-  mutate(mismatches_weight=weight*mismatches) %>%
-  mutate(total_weight=weight*total)
-error <- sum(error$mismatches_weight)/sum(error$total_weight)*100
-
-# Calculation of the number of sequences in the negative control
-nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
-  filter(taxon=="Root") %>%
-  select(NC_1)
-```
-
-Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
-For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
-Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019],  packages  vegan (version 2.5-6) [@Oksanen2019], and tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014].
-The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (**TO BE ADDED LATER!**).
-Based on the ATCC MSA-1002 mock community included in the analysis an average sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
-In addition, the negative controls processed together with the samples yielded on average only `r nc` sequences after sequence quality curation.
-
 ## Results
@@
 
+```{r core_otus}
+# Joining metadata with OTU/sample data
+metadata_rarefied <- inner_join(metadata, rarefied, by=c("ID"="Group")) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y"))
+
+# Finding the core OTUs for each environment
+core <- metadata_rarefied %>%
+  filter(station=="F") %>%
+  select_if(~ all(. != 0)) %>%
+  select(starts_with("Otu"))
+
+total <- metadata_rarefied %>%
+  filter(station=="F") %>%
+  select(starts_with("Otu"))
+
+core_otus_f <- length(core)/length(total) * 100
+core_seq_f <- sum(core)/sum(total) * 100
+
+core <- metadata_rarefied %>%
+  filter(station=="FCyM") %>%
+  select_if(~ all(. != 0)) %>%
+  select(starts_with("Otu"))
+
+total <- metadata_rarefied %>%
+  filter(station=="FCyM") %>%
+  select(starts_with("Otu"))
+
+core_otus_fcym <- length(core)/length(total) * 100
+core_seq_fcym <- sum(core)/sum(total) * 100
+
+core <- metadata_rarefied %>%
+  filter(station=="FCaM") %>%
+  select_if(~ all(. != 0)) %>%
+  select(starts_with("Otu"))
+
+total <- metadata_rarefied %>%
+  filter(station=="FCaM") %>%
+  select(starts_with("Otu"))
+
+core_otus_fcam <- length(core)/length(total) * 100
+core_seq_fcam <- sum(core)/sum(total) * 100
+
+core <- metadata_rarefied %>%
+  filter(station=="FCa") %>%
+  select_if(~ all(. != 0)) %>%
+  select(starts_with("Otu"))
+
+total <- metadata_rarefied %>%
+  filter(station=="FCa") %>%
+  select(starts_with("Otu"))
+
+core_otus_fca <- length(core)/length(total) * 100
+core_seq_fca <- sum(core)/sum(total) * 100
+
+core <- tibble(core_otus=c(core_otus_fcym, core_otus_fcam, core_otus_fca),
+               core_seq=c(core_seq_fcym, core_seq_fcam, core_seq_fca))
+```
+
 ```{r anosim}
@@
 Temporal proportional changes were more pronounced for seawater in comparison to *Cymodocea nodosa* and especially *Caulerpa cylindracea* associated communities (\autoref{shared}).
+In addition, only `r min(core$core_otus)` -- `r max(core$core_otus)` \si{\percent} of OTUs from each surface associated community were found at every time point.
+These OTUs made also a high proportion of total sequences (`r min(core$core_seq)` -- `r max(core$core_seq)` \si{\percent}).
 To further disentangle the environmental and seasonal community dissimilarity a Principal Coordinates Analysis (PCoA) based on Bray-Curtis distances and OTU abundances was applied.
@@
 
-The taxonomic composition of both, macrophyte associated and seawater communities, was dominated by *Bacteria* (`r filter(archaea_bacteria, taxon=="Bacteria")$mean` ± `r filter(archaea_bacteria, taxon=="Bacteria")$sd` \si{\percent}) over *Archaea* (`r filter(archaea_bacteria, taxon=="Archaea")$mean` ± `r filter(archaea_bacteria, taxon=="Archaea")$sd` \si{\percent}) (\autoref{community}).
+The taxonomic composition of both, macrophyte associated and seawater communities, was dominated by bacterial (`r filter(archaea_bacteria, taxon=="Bacteria")$mean` ± `r filter(archaea_bacteria, taxon=="Bacteria")$sd` \si{\percent}) over archaeal sequences (`r filter(archaea_bacteria, taxon=="Archaea")$mean` ± `r filter(archaea_bacteria, taxon=="Archaea")$sd` \si{\percent}) (\autoref{community}).
 Higher relative abundances of chloroplast related sequences were only observed in surface associated communities, with higher values in Autumn/Winter (`r filter(chloroplast, season=="Winter")$mean` ± `r filter(chloroplast, season=="Winter")$sd` \si{\percent}) in comparison to Spring/Summer (`r filter(chloroplast, season=="Summer")$mean` ± `r filter(chloroplast, season=="Summer")$sd` \si{\percent}) (\autoref{supp-chloroplast}).
@@
 
+\newpage
 ## Discussion
@@
 Different procedures of microbial cells removal from host surfaces were described, such as host tissue shaking [@Noges2010], scraping [@Uku2007] and ultrasonication [@Cai2014].
-All these methods showed different removal efficiencies and none was enabling a complete removal of attached microbial cells.
+All these methods showed different removal efficiencies but none was enabling a complete removal of attached microbial cells.
 In the present study, we applied an earlier developed removal protocol [@Korlevic_in_press, submitted], based on a previous idea of direct cellular lysis [@Burke2009], to ensure an almost complete cell detachment.
@@
 
-Observed highest richness values for *Caulerpa cylindracea* (invaded and noninvaded), middle for *Cymodocea nodosa* and lowest for seawater derived communities.
-Higher values for surface associated communities in comparison to seawater were described earlier for seagrasses [@Ugarelli2019] and could be attributed to a larger set of inhabitable microniches existing on macrophyte surfaces.
-In addition, highest values observed for *Caulerpa cylindracea* are probably a consequence of sediment derived OTUs that were present only in *Caulerpa cylindracea* communities.
-*Caulerpa cylindracea* stolon is attached to surface sediment with rhizoids, so the stolon and rhizoids are in a direct contact with the sediment surface.
-Part of the surface attached *Caulerpa cylnidracea* community is therefore comprised of sediment derived cells that could cause the observed increase in richness.
-In addition, seasonal richness differences observed for surface attached communities showed slightly higher values in spring and summer could be explained by a higher macrophyte growth in these seasons [M.
+In the present study, highest richness values were observed for *Caulerpa cylindracea* (invaded and noninvaded), middle for *Cymodocea nodosa* and lowest for seawater derived communities.
+Higher values for seagrass associated communities in comparison to seawater were described  earlier and could be attributed to a larger set of inhabitable microniches existing on macrophyte surfaces [@Ugarelli2019].
+In addition, highest values observed for *Caulerpa cylindracea* are probably a consequence of part of epiphytic OTUs that are in contact with the sediment.
+*Caulerpa cylindracea* stolon is attached to the sediment surface with rhizoids, so the stolon and rhizoids are in a direct contact with the sediment.
+Part of the surface attached *Caulerpa cylnidracea* community is therefore comprised of OTUs that are in contact with a different environment and that could cause the observed increase in richness.
+In addition, seasonal richness differences observed for surface attached communities showed slightly higher values in spring and summer.
+This pattern could be explained by a higher macrophyte growth in these seasons [M.
 Najdek, personal communication; @Zavodnik1998; @Ruitton2005].
-During active periods macrophytes exhibit a more dynamic chemical interaction with the surface community causing an extension of the microniche number.
+During active periods macrophytes exhibit a more dynamic chemical interaction with the surface community probably causing an increase in the number of inhabitable microniches [@Rickert2016; @Borges2015].
+
+Since the colonization of macrophyte surfaces is performed from a pool of seawater prokaryotic cells it is interesting to see to which extent do these two communities differ.
+We observed a strong differentiation between the surface attached and seawater communities at the level of OTUs that is in agreement with most published studies [@Ugarelli2019; @Roth-Schulze2016; @Burke2011b; @Michelou2013].
+These data indicate that marine macrophytes are selecting, from a pool of seawater microbial taxa, the one that can colonize and proliferate on their surfaces [@Salaun2012; @Michelou2013].
+In contrast to these findings @Fahimipour2017 found, in a global study of *Zostera marina*, similarities between leaves and seawater samples.
+Discrepancies between our data and this study could be explained by differences in studied seagrass species, methodological variations or biogeographic trends as @Fahimipour2017 were analyzing samples from different locations throughout the northern hemisphere.
+It is possible that ambient seawater and leaves communities from the same location are differing but are still more similar to each other when compared to other sampling locations.
+Indeed, it was found that prokaryotic communities vary substantially between different sampling sites [@Bengtsson2017].
+When the taxonomic composition at high ranks was analyzed no such strong differentiation was noticed.
+Phyla and classes such as: *Actinobacteriota*, *Bacteroidota*, *Cyanobacteria*, *Alphaproteobacteria*, *Gammaproteobacteria* and *Verrucomicrobiota*, were described that is in agreement with previously reported data [@Egan2013; @Michelou2013; @Burke2011b].
+In contrast, when low taxonomic ranks were analyzed (i.g.
+family and genus) a strong differentiation was observed.
+A similar differentiation at lower taxonomic ranks was described for other species of macrophytes [@Egan2013; @Michelou2013; @Ugarelli2019].
+
+Beside differences between seawater and surface associated communities, there were discussions if the prokaryotic epiphytic community is host-specific or there are generalists taxa characteristic to all or many macrophytes [@Egan2013].
+Similarly to previously described differences between seawater and surface attached communities, at high taxonomic ranks no strong differentiation between communities associated with different host was observed.
+The only high rank phylum that was differing between *Cymodocea nodosa* ans *Caulerpa cylindracea* was *Desulfobacterota*, whose sequences were more abundant in the *Caulerpa cylindracea* associated community.
+As already mentioned, the rhizoids and part of the stolon are in contact with the sediment, so *Desulfobacterota* are probably a part of the epiphytic community that is in contact with the sediment.
+Similar high rank taxa found in this study were described to be specific for other species of macrophytes [@Burke2011b; @Lachnit2011; @Bengtsson2017].
+In contrast to high taxonomic ranks, a substantial differentiation between host specific communities was found, which supports the host-specific hypothesis.
+Similar host-specificity was observed for different species of macroalgae and seagrasses [@Lachnit2011; @Roth-Schulze2016; @Ugarelli2019; @Morrissey2019].
+Taken together, at high taxonomic ranks a core set of taxa could be described that is characteristic for all or many macrophytes, while at low taxonomic ranks a community specific to host species could be identified [@Egan2013].
+
+Few studies describing microbial epiphytic communities were encompassing seasonal shifts [@Lachnit2011; @Tujula2010; @Michelou2013].
+Seasonal changes in the epiphytic community were substantial as could be observed in the proportion of OTUs that could be find at every sampling time ($\leqslant$ `r max(core$core_otus)` \si{\percent}).
+Interestingly, these OTUs were accounting for a high proportion of sequences ($\leqslant$ `r max(core$core_seq)` \si{\percent}).
+A very similar proportion of persistent OTUs and their sequence contribution was reported in high frequency studies describing seasonal picoplankton changes [@Gilbert2009; @Gilbert2012].
+In comparison to the seawater community, a lower degree of seasonal shifts was observed for the surface associated communities.
+It seems, that microniches on the surfaces of macrophytes are providing more stable conditions in comparison to the seawater.
+At the level of OTUs seasonal changes of *Cymodocea nodosa* and *Caulerpa cylindracea* associated communities were identified that could be linked to the growth cycle of the seagrass and macroalgae (M.
+Najdek, personal communication).
+*Cymodocea nodosa* was characterized by a Spring community during maximum seagrass proliferation, a Summer community during a biomass maximum and a Autumn/Winter community during a biomass decay.
+In contrast, *Caulerpa cylindracea* started to proliferate in late Spring and was characterized only by a Summer community during maximal biomass increase and by a Autumn/Winter/Spring community when the biomass were at the peak and the settlement started to subsequently decay.
+Similar seasonal changes in the epiphytic community was described also for other macroalgae [@Lachnit2011; @Tujula2010].
+The higher temporal stability of *Caulerpa cylindracea* surface communities in comparison to *Cymodocea nodosa* could be also observed in the higher proportion of shared communities between two consecutive sampling points.
+
+\newpage
+## Experimental Procedures
+### Sampling
+Leaves of *Cymodocea nodosa* were sampled in a *Cymodocea nodosa* meadow located in the proximity of the village of Funtana (\ang{45;10;39} N, \ang{13;35;42} E).
+Thalli of *Caulerpa cylindracea* were sampled in the same *Cymodocea nodosa* invaded meadow in Funtana and on a locality of only *Caulerpa cylindracea* located close to the invaded meadow.
+Sampling of leaves and thalli was performed approximately monthly from December 2017 to October 2018 (\autoref{supp-nseq_notus}).
+Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with site seawater.
+Upon arrival to the laboratory, *Cymodocea nodosa* leaves were cut into sections of 1 -- 2 \si{\cm}, while *Caulerpa cylindracea* thalli were cut into 5 -- 8 \si{\cm} long sections.
+Leaves and thalli were washed three times with sterile artificial seawater (ASW) to remove loosely attached microbial cells.
+Surrounding seawater was collected in 10 \si{\l} containers by diving and transported to the laboratory where the whole container volume was filtered through a 20 \si{\um} net.
+The filtrate was further sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate membrane filters (Whatman, United Kingdom) using a peristaltic pump.
+Filters were briefly dried at room temperature and stored at \num{-80} \si{\degreeCelsius}.
+Seawater samples were also collected approximately monthly from July 2017 to October 2018.
+
+### DNA Isolation
+DNA from surfaces of *Cymodocea nodosa* and *Caulerpa cylindracea* was isolated using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevic_in_press, submitted; @Massana1997].
+Briefly, leaves and thalli are incubated in a lysis buffer and treated with lysozyme and proteinase K.
+Following the incubations, the mixture containing lyzed epiphytic cells is separated from leaves and thalli and extracted using a phenol-chloroform procedure.
+Finally, the extracted DNA is precipitated using isopropanol.
+DNA from seawater picoplankton was isolated from 0.2 \si{\um} polycarbonate filters according to [@Massana1997] with a slight modification.
+Following the phenol-chloroform extraction steps 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added.
+DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 20,000 × g and 4 \si{\degreeCelsius} for 21 \si{\minute}.
+The pellet was washed twice with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}.
+Dried pellets were resuspended in 50 -- 100 \si{\ul} of deionized water.
+
+### Illumina 16S rRNA Sequencing
+Illumina MiSeq sequencing of the V4 16S rRNA region was performed as described previously [@Korlevic_in_press, submitted].
+The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
+In the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used [@Caporaso2012; @Apprill2015; @Parada2016].
+These primers contained on their 5' end a tagged sequence.
+Purified PCR products were sent for Illumina MiSeq sequencing at IMGM Laboratories, Martinsried, Germany.
+Before sequencing at IMGM, the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR.
+In addition, these primers contained adapter and sample-specific index sequences.
+Beside samples, a positive and negative control for each sequencing batch was sequenced.
+Negative control was comprised of PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
+The sequences obtained in this study have been submitted to the European Nucleotide Archive (ENA) under accession numbers **TO BE ADDED LATER!**.
+
+### Sequence Analysis
+```{r error}
+# Calculation of the error rate based on the ATCC MSA-1002 mock community
+error <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.error.summary", na="") %>%
+  filter(numparents==1) %>%
+  mutate(mismatches_weight=weight*mismatches) %>%
+  mutate(total_weight=weight*total)
+error <- sum(error$mismatches_weight)/sum(error$total_weight)*100
+
+# Calculation of the number of sequences in the negative control
+nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
+  filter(taxon=="Root") %>%
+  select(NC_1)
+```
+
+Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
+For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
+Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019],  packages  vegan (version 2.5-6) [@Oksanen2019], and tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014].
+The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (**TO BE ADDED LATER!**).
+Based on the ATCC MSA-1002 mock community included in the analysis an average sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
+In addition, the negative controls processed together with the samples yielded on average only `r nc` sequences after sequence quality curation.
 

### 2020-03-04 submission/manuscript.tex
Discussion progress.

@@
 do not allow for high taxonomic resolution were used (Tujula \emph{et
-al.}, 2010; Lachnit \emph{et al.}, 2011; Miranda \emph{et al.}, 2013).
-In the present study we describe the seasonal dynamics of bacterial and
-archaeal communities on the surfaces of the seagrass \emph{Cymodocea
-nodosa} and siphonous macroalgae \emph{Caulerpa cylindracea} determined
-on a mostly monthly scale. Bacterial and archaeal epiphytes were sampled
-in a meadow of \emph{Cymodocea nodosa} invaded by the invasive
-\emph{Caulerpa cylindracea} and in a locality of only \emph{Caulerpa
-cylindracea} located in the proximity of the meadow. In addition, for
-comparison, the community of the surrounding seawater was characterized.
+al.}, 2010; Lachnit \emph{et al.}, 2011; Miranda \emph{et al.}, 2013;
+Michelou \emph{et al.}, 2013). In the present study we describe the
+seasonal dynamics of bacterial and archaeal communities on the surfaces
+of the seagrass \emph{Cymodocea nodosa} and siphonous macroalgae
+\emph{Caulerpa cylindracea} determined on a mostly monthly scale.
+Bacterial and archaeal epiphytes were sampled in a meadow of
+\emph{Cymodocea nodosa} invaded by the invasive \emph{Caulerpa
+cylindracea} and in a locality of only \emph{Caulerpa cylindracea}
+located in the proximity of the meadow. In addition, for comparison, the
+community of the surrounding seawater was characterized.
 
 \newpage
-
-\hypertarget{materials-and-methods}{%
-\subsection{Materials and Methods}\label{materials-and-methods}}
-
-\hypertarget{sampling}{%
-\subsubsection{Sampling}\label{sampling}}
-
-Leaves of \emph{Cymodocea nodosa} were sampled in a \emph{Cymodocea
-nodosa} meadow located in the proximity of the village of Funtana
-(\ang{45;10;39} N, \ang{13;35;42} E). Thalli of \emph{Caulerpa
-cylindracea} were sampled in the same \emph{Cymodocea nodosa} invaded
-meadow in Funtana and on a locality of only \emph{Caulerpa cylindracea}
-located close to the invaded meadow. Sampling of leaves and thalli was
-performed approximately monthly from December 2017 to October 2018
-(\autoref{supp-nseq_notus}). Leaves and thalli were collected by diving
-and transported to the laboratory in containers placed on ice and filled
-with site seawater. Upon arrival to the laboratory, \emph{Cymodocea
-nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while
-\emph{Caulerpa cylindracea} thalli were cut into 5 -- 8 \si{\cm} long
-sections. Leaves and thalli were washed three times with sterile
-artificial seawater (ASW) to remove loosely attached microbial cells.
-Surrounding seawater was collected in 10 \si{\l} containers by diving
-and transported to the laboratory where the whole container volume was
-filtered through a 20 \si{\um} net. The filtrate was further
-sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate
-membrane filters (Whatman, United Kingdom) using a peristaltic pump.
-Filters were briefly dried at room temperature and stored at \num{-80}
-\si{\degreeCelsius}. Seawater samples were also collected approximately
-monthly from July 2017 to October 2018.
-
-\hypertarget{dna-isolation}{%
-\subsubsection{DNA Isolation}\label{dna-isolation}}
-
-DNA from surfaces of \emph{Cymodocea nodosa} and \emph{Caulerpa
-cylindracea} was isolated using a previously modified and adapted
-protocol that allows for a selective epiphytic DNA isolation (Massana
-\emph{et al.}, 1997; Korlević \emph{et al.}, submitted). Briefly, leaves
-and thalli are incubated in a lysis buffer and treated with lysozyme and
-proteinase K. Following the incubations, the mixture containing lyzed
-epiphytic cells is separated from leaves and thalli and extracted using
-a phenol-chloroform procedure. Finally, the extracted DNA is
-precipitated using isopropanol. DNA from seawater picoplankton was
-isolated from 0.2 \si{\um} polycarbonate filters according to (Massana
-\emph{et al.}, 1997) with a slight modification. Following the
-phenol-chloroform extraction steps 1/10 of chilled 3 \si{\Molar} sodium
-acetate (pH 5.2) was added. DNA was precipitated by adding 1 volume of
-chilled isopropanol, incubating the mixtures overnight at \num{-20}
-\si{\degreeCelsius} and centrifuging at 20,000 × g and 4
-\si{\degreeCelsius} for 21 \si{\minute}. The pellet was washed twice
-with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged
-after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 5
-\si{\minute}. Dried pellets were resuspended in 50 -- 100 \si{\ul} of
-deionized water.
-
-\hypertarget{illumina-16s-rrna-sequencing}{%
-\subsubsection{Illumina 16S rRNA
-Sequencing}\label{illumina-16s-rrna-sequencing}}
-
-Illumina MiSeq sequencing of the V4 16S rRNA region was performed as
-described previously (Korlević \emph{et al.}, submitted). The V4 region
-of the 16S rRNA gene was amplified using a two-step PCR procedure. In
-the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R
-(5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project
-(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
-were used (Caporaso \emph{et al.}, 2012; Apprill \emph{et al.}, 2015;
-Parada \emph{et al.}, 2016). These primers contained on their 5' end a
-tagged sequence. Purified PCR products were sent for Illumina MiSeq
-sequencing at IMGM Laboratories, Martinsried, Germany. Before sequencing
-at IMGM, the second PCR amplification of the two-step PCR procedure was
-performed using primers targeting the tagged region incorporated in the
-first PCR. In addition, these primers contained adapter and
-sample-specific index sequences. Beside samples, a positive and negative
-control for each sequencing batch was sequenced. Negative control was
-comprised of PCR reactions without DNA template, while for a positive
-control a mock community composed of evenly mixed DNA material
-originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
-used. The sequences obtained in this study have been submitted to the
-European Nucleotide Archive (ENA) under accession numbers \textbf{TO BE
-ADDED LATER!}.
-
-\hypertarget{sequence-analysis}{%
-\subsubsection{Sequence Analysis}\label{sequence-analysis}}
-
-Obtained sequences were analyzed on the computer cluster Isabella
-(University Computing Center, University of Zagreb) using mothur
-(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
-Standard Operating Procedure (MiSeq SOP;
-\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
-and recommendations given from the Riffomonas project to enhance data
-reproducibility (\url{http://www.riffomonas.org/}). For alignment and
-classification of sequences the SILVA SSU Ref NR 99 database (release
-138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
-Yilmaz \emph{et al.}, 2014). Pipeline data processing and visualization
-was done using R (version 3.6.0) (R Core Team, 2019), packages vegan
-(version 2.5-6) (Oksanen \emph{et al.}, 2019), and tidyverse (version
-1.3.0) (Wickham \emph{et al.}, 2019) and multiple other packages (Xie,
-2014, 2015, 2019a, 2019b, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018;
-Allaire \emph{et al.}, 2019; Zhu, 2019). The detailed analysis procedure
-including the R Markdown file for this paper are available as a GitHub
-repository (\textbf{TO BE ADDED LATER!}). Based on the ATCC MSA-1002
-mock community included in the analysis an average sequencing error rate
-of 0.01 \si{\percent} was determined, which is in line with previously
-reported values for next-generation sequencing data (Kozich \emph{et
-al.}, 2013; Schloss \emph{et al.}, 2016). In addition, the negative
-controls processed together with the samples yielded on average only 2
-sequences after sequence quality curation.
 
@@
 \emph{Cymodocea nodosa} and especially \emph{Caulerpa cylindracea}
-associated communities (\autoref{shared}). To further disentangle the
+associated communities (\autoref{shared}). In addition, only 0.4 -- 1.0
+\si{\percent} of OTUs from each surface associated community were found
+at every time point. These OTUs made also a high proportion of total
+sequences (39.1 -- 51.3 \si{\percent}). To further disentangle the
 environmental and seasonal community dissimilarity a Principal
@@
 The taxonomic composition of both, macrophyte associated and seawater
-communities, was dominated by \emph{Bacteria} (99.1 ± 2.1 \si{\percent})
-over \emph{Archaea} (0.9 ± 2.1 \si{\percent}) (\autoref{community}).
+communities, was dominated by bacterial (99.1 ± 2.1 \si{\percent}) over
+archaeal sequences (0.9 ± 2.1 \si{\percent}) (\autoref{community}).
 Higher relative abundances of chloroplast related sequences were only
@@
 
+\newpage
+
 \hypertarget{discussion}{%
@@
 \emph{et al.}, 2007) and ultrasonication (Cai \emph{et al.}, 2014). All
-these methods showed different removal efficiencies and none was
+these methods showed different removal efficiencies but none was
 enabling a complete removal of attached microbial cells. In the present
@@
 
-Observed highest richness values for \emph{Caulerpa cylindracea}
-(invaded and noninvaded), middle for \emph{Cymodocea nodosa} and lowest
-for seawater derived communities. Higher values for surface associated
-communities in comparison to seawater were described earlier for
-seagrasses (Ugarelli \emph{et al.}, 2019) and could be attributed to a
-larger set of inhabitable microniches existing on macrophyte surfaces.
-In addition, highest values observed for \emph{Caulerpa cylindracea} are
-probably a consequence of sediment derived OTUs that were present only
-in \emph{Caulerpa cylindracea} communities. \emph{Caulerpa cylindracea}
-stolon is attached to surface sediment with rhizoids, so the stolon and
-rhizoids are in a direct contact with the sediment surface. Part of the
-surface attached \emph{Caulerpa cylnidracea} community is therefore
-comprised of sediment derived cells that could cause the observed
-increase in richness. In addition, seasonal richness differences
-observed for surface attached communities showed slightly higher values
-in spring and summer could be explained by a higher macrophyte growth in
-these seasons (M. Najdek, personal communication; Zavodnik \emph{et
-al.}, 1998; Ruitton \emph{et al.}, 2005). During active periods
-macrophytes exhibit a more dynamic chemical interaction with the surface
-community causing an extension of the microniche number.
+In the present study, highest richness values were observed for
+\emph{Caulerpa cylindracea} (invaded and noninvaded), middle for
+\emph{Cymodocea nodosa} and lowest for seawater derived communities.
+Higher values for seagrass associated communities in comparison to
+seawater were described earlier and could be attributed to a larger set
+of inhabitable microniches existing on macrophyte surfaces (Ugarelli
+\emph{et al.}, 2019). In addition, highest values observed for
+\emph{Caulerpa cylindracea} are probably a consequence of part of
+epiphytic OTUs that are in contact with the sediment. \emph{Caulerpa
+cylindracea} stolon is attached to the sediment surface with rhizoids,
+so the stolon and rhizoids are in a direct contact with the sediment.
+Part of the surface attached \emph{Caulerpa cylnidracea} community is
+therefore comprised of OTUs that are in contact with a different
+environment and that could cause the observed increase in richness. In
+addition, seasonal richness differences observed for surface attached
+communities showed slightly higher values in spring and summer. This
+pattern could be explained by a higher macrophyte growth in these
+seasons (M. Najdek, personal communication; Zavodnik \emph{et al.},
+1998; Ruitton \emph{et al.}, 2005). During active periods macrophytes
+exhibit a more dynamic chemical interaction with the surface community
+probably causing an increase in the number of inhabitable microniches
+(Borges and Champenois, 2015; Rickert \emph{et al.}, 2016).
+
+Since the colonization of macrophyte surfaces is performed from a pool
+of seawater prokaryotic cells it is interesting to see to which extent
+do these two communities differ. We observed a strong differentiation
+between the surface attached and seawater communities at the level of
+OTUs that is in agreement with most published studies (Burke \emph{et
+al.}, 2011b; Michelou \emph{et al.}, 2013; Roth-Schulze \emph{et al.},
+2016; Ugarelli \emph{et al.}, 2019). These data indicate that marine
+macrophytes are selecting, from a pool of seawater microbial taxa, the
+one that can colonize and proliferate on their surfaces (Salaün \emph{et
+al.}, 2012; Michelou \emph{et al.}, 2013). In contrast to these findings
+Fahimipour \emph{et al.} (2017) found, in a global study of
+\emph{Zostera marina}, similarities between leaves and seawater samples.
+Discrepancies between our data and this study could be explained by
+differences in studied seagrass species, methodological variations or
+biogeographic trends as Fahimipour \emph{et al.} (2017) were analyzing
+samples from different locations throughout the northern hemisphere. It
+is possible that ambient seawater and leaves communities from the same
+location are differing but are still more similar to each other when
+compared to other sampling locations. Indeed, it was found that
+prokaryotic communities vary substantially between different sampling
+sites (Bengtsson \emph{et al.}, 2017). When the taxonomic composition at
+high ranks was analyzed no such strong differentiation was noticed.
+Phyla and classes such as: \emph{Actinobacteriota}, \emph{Bacteroidota},
+\emph{Cyanobacteria}, \emph{Alphaproteobacteria},
+\emph{Gammaproteobacteria} and \emph{Verrucomicrobiota}, were described
+that is in agreement with previously reported data (Burke \emph{et al.},
+2011b; Egan \emph{et al.}, 2013; Michelou \emph{et al.}, 2013). In
+contrast, when low taxonomic ranks were analyzed (i.g. family and genus)
+a strong differentiation was observed. A similar differentiation at
+lower taxonomic ranks was described for other species of macrophytes
+(Egan \emph{et al.}, 2013; Michelou \emph{et al.}, 2013; Ugarelli
+\emph{et al.}, 2019).
+
+Beside differences between seawater and surface associated communities,
+there were discussions if the prokaryotic epiphytic community is
+host-specific or there are generalists taxa characteristic to all or
+many macrophytes (Egan \emph{et al.}, 2013). Similarly to previously
+described differences between seawater and surface attached communities,
+at high taxonomic ranks no strong differentiation between communities
+associated with different host was observed. The only high rank phylum
+that was differing between \emph{Cymodocea nodosa} ans \emph{Caulerpa
+cylindracea} was \emph{Desulfobacterota}, whose sequences were more
+abundant in the \emph{Caulerpa cylindracea} associated community. As
+already mentioned, the rhizoids and part of the stolon are in contact
+with the sediment, so \emph{Desulfobacterota} are probably a part of the
+epiphytic community that is in contact with the sediment. Similar high
+rank taxa found in this study were described to be specific for other
+species of macrophytes (Burke \emph{et al.}, 2011b; Lachnit \emph{et
+al.}, 2011; Bengtsson \emph{et al.}, 2017). In contrast to high
+taxonomic ranks, a substantial differentiation between host specific
+communities was found, which supports the host-specific hypothesis.
+Similar host-specificity was observed for different species of
+macroalgae and seagrasses (Lachnit \emph{et al.}, 2011; Roth-Schulze
+\emph{et al.}, 2016; Ugarelli \emph{et al.}, 2019; Morrissey \emph{et
+al.}, 2019). Taken together, at high taxonomic ranks a core set of taxa
+could be described that is characteristic for all or many macrophytes,
+while at low taxonomic ranks a community specific to host species could
+be identified (Egan \emph{et al.}, 2013).
+
+Few studies describing microbial epiphytic communities were encompassing
+seasonal shifts (Tujula \emph{et al.}, 2010; Lachnit \emph{et al.},
+2011; Michelou \emph{et al.}, 2013). Seasonal changes in the epiphytic
+community were substantial as could be observed in the proportion of
+OTUs that could be find at every sampling time (\(\leqslant\) 1.0
+\si{\percent}). Interestingly, these OTUs were accounting for a high
+proportion of sequences (\(\leqslant\) 51.3 \si{\percent}). A very
+similar proportion of persistent OTUs and their sequence contribution
+was reported in high frequency studies describing seasonal picoplankton
+changes (Gilbert \emph{et al.}, 2009, 2012). In comparison to the
+seawater community, a lower degree of seasonal shifts was observed for
+the surface associated communities. It seems, that microniches on the
+surfaces of macrophytes are providing more stable conditions in
+comparison to the seawater. At the level of OTUs seasonal changes of
+\emph{Cymodocea nodosa} and \emph{Caulerpa cylindracea} associated
+communities were identified that could be linked to the growth cycle of
+the seagrass and macroalgae (M. Najdek, personal communication).
+\emph{Cymodocea nodosa} was characterized by a Spring community during
+maximum seagrass proliferation, a Summer community during a biomass
+maximum and a Autumn/Winter community during a biomass decay. In
+contrast, \emph{Caulerpa cylindracea} started to proliferate in late
+Spring and was characterized only by a Summer community during maximal
+biomass increase and by a Autumn/Winter/Spring community when the
+biomass were at the peak and the settlement started to subsequently
+decay. Similar seasonal changes in the epiphytic community was described
+also for other macroalgae (Tujula \emph{et al.}, 2010; Lachnit \emph{et
+al.}, 2011). The higher temporal stability of \emph{Caulerpa
+cylindracea} surface communities in comparison to \emph{Cymodocea
+nodosa} could be also observed in the higher proportion of shared
+communities between two consecutive sampling points.
+
+\newpage
+
+\hypertarget{experimental-procedures}{%
+\subsection{Experimental Procedures}\label{experimental-procedures}}
+
+\hypertarget{sampling}{%
+\subsubsection{Sampling}\label{sampling}}
+
+Leaves of \emph{Cymodocea nodosa} were sampled in a \emph{Cymodocea
+nodosa} meadow located in the proximity of the village of Funtana
+(\ang{45;10;39} N, \ang{13;35;42} E). Thalli of \emph{Caulerpa
+cylindracea} were sampled in the same \emph{Cymodocea nodosa} invaded
+meadow in Funtana and on a locality of only \emph{Caulerpa cylindracea}
+located close to the invaded meadow. Sampling of leaves and thalli was
+performed approximately monthly from December 2017 to October 2018
+(\autoref{supp-nseq_notus}). Leaves and thalli were collected by diving
+and transported to the laboratory in containers placed on ice and filled
+with site seawater. Upon arrival to the laboratory, \emph{Cymodocea
+nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while
+\emph{Caulerpa cylindracea} thalli were cut into 5 -- 8 \si{\cm} long
+sections. Leaves and thalli were washed three times with sterile
+artificial seawater (ASW) to remove loosely attached microbial cells.
+Surrounding seawater was collected in 10 \si{\l} containers by diving
+and transported to the laboratory where the whole container volume was
+filtered through a 20 \si{\um} net. The filtrate was further
+sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate
+membrane filters (Whatman, United Kingdom) using a peristaltic pump.
+Filters were briefly dried at room temperature and stored at \num{-80}
+\si{\degreeCelsius}. Seawater samples were also collected approximately
+monthly from July 2017 to October 2018.
+
+\hypertarget{dna-isolation}{%
+\subsubsection{DNA Isolation}\label{dna-isolation}}
+
+DNA from surfaces of \emph{Cymodocea nodosa} and \emph{Caulerpa
+cylindracea} was isolated using a previously modified and adapted
+protocol that allows for a selective epiphytic DNA isolation (Massana
+\emph{et al.}, 1997; Korlević \emph{et al.}, submitted). Briefly, leaves
+and thalli are incubated in a lysis buffer and treated with lysozyme and
+proteinase K. Following the incubations, the mixture containing lyzed
+epiphytic cells is separated from leaves and thalli and extracted using
+a phenol-chloroform procedure. Finally, the extracted DNA is
+precipitated using isopropanol. DNA from seawater picoplankton was
+isolated from 0.2 \si{\um} polycarbonate filters according to (Massana
+\emph{et al.}, 1997) with a slight modification. Following the
+phenol-chloroform extraction steps 1/10 of chilled 3 \si{\Molar} sodium
+acetate (pH 5.2) was added. DNA was precipitated by adding 1 volume of
+chilled isopropanol, incubating the mixtures overnight at \num{-20}
+\si{\degreeCelsius} and centrifuging at 20,000 × g and 4
+\si{\degreeCelsius} for 21 \si{\minute}. The pellet was washed twice
+with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged
+after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 5
+\si{\minute}. Dried pellets were resuspended in 50 -- 100 \si{\ul} of
+deionized water.
+
+\hypertarget{illumina-16s-rrna-sequencing}{%
+\subsubsection{Illumina 16S rRNA
+Sequencing}\label{illumina-16s-rrna-sequencing}}
+
+Illumina MiSeq sequencing of the V4 16S rRNA region was performed as
+described previously (Korlević \emph{et al.}, submitted). The V4 region
+of the 16S rRNA gene was amplified using a two-step PCR procedure. In
+the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R
+(5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project
+(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
+were used (Caporaso \emph{et al.}, 2012; Apprill \emph{et al.}, 2015;
+Parada \emph{et al.}, 2016). These primers contained on their 5' end a
+tagged sequence. Purified PCR products were sent for Illumina MiSeq
+sequencing at IMGM Laboratories, Martinsried, Germany. Before sequencing
+at IMGM, the second PCR amplification of the two-step PCR procedure was
+performed using primers targeting the tagged region incorporated in the
+first PCR. In addition, these primers contained adapter and
+sample-specific index sequences. Beside samples, a positive and negative
+control for each sequencing batch was sequenced. Negative control was
+comprised of PCR reactions without DNA template, while for a positive
+control a mock community composed of evenly mixed DNA material
+originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
+used. The sequences obtained in this study have been submitted to the
+European Nucleotide Archive (ENA) under accession numbers \textbf{TO BE
+ADDED LATER!}.
+
+\hypertarget{sequence-analysis}{%
+\subsubsection{Sequence Analysis}\label{sequence-analysis}}
+
+Obtained sequences were analyzed on the computer cluster Isabella
+(University Computing Center, University of Zagreb) using mothur
+(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
+Standard Operating Procedure (MiSeq SOP;
+\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
+and recommendations given from the Riffomonas project to enhance data
+reproducibility (\url{http://www.riffomonas.org/}). For alignment and
+classification of sequences the SILVA SSU Ref NR 99 database (release
+138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
+Yilmaz \emph{et al.}, 2014). Pipeline data processing and visualization
+was done using R (version 3.6.0) (R Core Team, 2019), packages vegan
+(version 2.5-6) (Oksanen \emph{et al.}, 2019), and tidyverse (version
+1.3.0) (Wickham \emph{et al.}, 2019) and multiple other packages (Xie,
+2014, 2015, 2019a, 2019b, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018;
+Allaire \emph{et al.}, 2019; Zhu, 2019). The detailed analysis procedure
+including the R Markdown file for this paper are available as a GitHub
+repository (\textbf{TO BE ADDED LATER!}). Based on the ATCC MSA-1002
+mock community included in the analysis an average sequencing error rate
+of 0.01 \si{\percent} was determined, which is in line with previously
+reported values for next-generation sequencing data (Kozich \emph{et
+al.}, 2013; Schloss \emph{et al.}, 2016). In addition, the negative
+controls processed together with the samples yielded on average only 2
+sequences after sequence quality curation.
 
@@
 
+\leavevmode\hypertarget{ref-Bengtsson2017}{}%
+Bengtsson, M.M., Bühler, A., Brauer, A., Dahlke, S., Schubert, H., and
+Blindow, I. (2017) Eelgrass Leaf Surface Microbiomes Are Locally
+Variable and Highly Correlated with Epibiotic Eukaryotes.
+\emph{Frontiers in Microbiology} \textbf{8}: 1312.
+
 \leavevmode\hypertarget{ref-Bengtsson2010}{}%
@@
 Microbial Ecology} \textbf{60}: 71--83.
+
+\leavevmode\hypertarget{ref-Borges2015}{}%
+Borges, A.V. and Champenois, W. (2015) Seasonal and spatial variability
+of dimethylsulfoniopropionate (DMSP) in the Mediterranean seagrass
+Posidonia oceanica. \emph{Aquatic Botany} \textbf{125}: 72--79.
 
@@
 interactions. \emph{FEMS microbiology reviews} \textbf{37}: 462--476.
+
+\leavevmode\hypertarget{ref-Fahimipour2017}{}%
+Fahimipour, A.K., Kardish, M.R., Lang, J.M., Green, J.L., Eisen, J.A.,
+and Stachowicz, J.J. (2017) Globalscale structure of the eelgrass
+microbiome. \emph{Applied and Environmental Microbiology} \textbf{83}:
+
+\leavevmode\hypertarget{ref-Gilbert2009}{}%
+Gilbert, J.A., Field, D., Swift, P., Newbold, L., Oliver, A., Smyth, T.,
+et al. (2009) The seasonal structure of microbial communities in the
+Western English Channel. \emph{Environmental Microbiology} \textbf{11}:
+3132--3139.
+
+\leavevmode\hypertarget{ref-Gilbert2012}{}%
+Gilbert, J.A., Steele, J.A., Caporaso, J.G., Steinbrück, L., Reeder, J.,
+Temperton, B., et al. (2012) Defining seasonal marine microbial
+community dynamics. \emph{The ISME Journal} \textbf{6}: 298--308.
 
@@
 
+\leavevmode\hypertarget{ref-Michelou2013}{}%
+Michelou, V.K., Caporaso, J.G., Knight, R., and Palumbi, S.R. (2013) The
+Ecology of Microbial Communities Associated with Macrocystis pyrifera.
+\emph{PLoS ONE} \textbf{8}: e67480.
+
 \leavevmode\hypertarget{ref-Miranda2013}{}%
@@
 
+\leavevmode\hypertarget{ref-Rickert2016}{}%
+Rickert, E., Wahl, M., Link, H., Richter, H., and Pohnert, G. (2016)
+Seasonal variations in surface metabolite composition of fucus
+vesiculosus and fucus serratus from the Baltic Sea. \emph{PLoS ONE}
+\textbf{11}:
+
 \leavevmode\hypertarget{ref-Roth-Schulze2016}{}%
@@
 range. \emph{Aquatic Botany} \textbf{82}: 55--70.
+
+\leavevmode\hypertarget{ref-Salaun2012}{}%
+Salaün, S., Barre, S. la, Santos-Goncalvez, M.D., Potin, P., Haras, D.,
+and Bazire, A. (2012) Influence of Exudates of the Kelp Laminaria
+Digitata on Biofilm Formation of Associated and Exogenous Bacterial
+Epiphytes. \emph{Microbial Ecology} \textbf{64}: 359--369.
 

### 2020-03-04 submission/supplementary.Rmd
Discussion progress.

@@
 of Sequences and No.
-of OTUs of each sample.\\label{nseq_notus}", align="c") %>%
+of OTUs of each sample.
+No.
+of Sequences and OTUs was calculated after the exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences.\\label{nseq_notus}", align="c") %>%
   kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9) %>%

### 2020-03-04 submission/supplementary.tex
OTU taxonomy file added to .gitignore.
Discussion progress.

@@
 \begin{longtable}{>{\centering\arraybackslash}p{6em}ccccc}
-\caption{\label{tab:nseq_notus}Sample ID, Community Type, Sampling Date and Season, No. of Sequences and No. of OTUs of each sample.\label{nseq_notus}}\\
+\caption{\label{tab:nseq_notus}Sample ID, Community Type, Sampling Date and Season, No.
+of Sequences and No.
+of OTUs of each sample.
+No.
+of Sequences and OTUs was calculated after the exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences.\label{nseq_notus}}\\
 \toprule
@@
 \endfirsthead
-\caption[]{Sample ID, Community Type, Sampling Date and Season, No. of Sequences and No. of OTUs of each sample.\label{nseq_notus} \textit{(continued)}}\\
+\caption[]{Sample ID, Community Type, Sampling Date and Season, No.
+of Sequences and No.
+of OTUs of each sample.
+No.
+of Sequences and OTUs was calculated after the exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences.\label{nseq_notus} \textit{(continued)}}\\
 \toprule

### 2020-03-09 submission/manuscript.Rmd
Additional discussion.

@@
 A total of `r samples_epiphytes` samples originating from epiphytic archaeal and bacterial communities associated with surfaces of the seagrass *Cymodocea nodosa* and macroalga *Caulerpa cylindracea* were analyzed.
-In addition, `r samples_seawater` samples (one of the samples was sequenced two times) originating from picoplankton archaeal and bacterial communities in the surrounding seawater were also processed.
+In addition, `r samples_seawater` samples (one of the samples was sequenced two times) originating from picoplankton archaeal and bacterial communities in the ambient seawater were also processed for comparison.
 The number of reads per sample ranged between `r min_max$min` and `r min_max$max` sequences (\autoref{supp-nseq_notus}).
@@
 Generally, richness estimators and diversity indices showed similar trends.
-On average, higher values were found for *Caulerpa cylindracea* (invaded [Number of OTUs, `r S.obs_fcam$mean` ± `r S.obs_fcam$sd` OTUs] and noninvaded [Number of OTUs, `r S.obs_fca$mean` ± `r S.obs_fca$sd` OTUs]), middle vales for *Cymodocea nodosa* (Number of OTUs, `r S.obs_fcym$mean` ± `r S.obs_fcym$sd` OTUs) and lower values for picoplankton communities in the surrounding seawater (Number of OTUs, `r S.obs_f$mean` ± `r S.obs_f$sd` OTUs) (\autoref{supp-calculators}).
+On average, higher values were found for *Caulerpa cylindracea* (invaded [Number of OTUs, `r S.obs_fcam$mean` ± `r S.obs_fcam$sd` OTUs] and noninvaded [Number of OTUs, `r S.obs_fca$mean` ± `r S.obs_fca$sd` OTUs]), middle vales for *Cymodocea nodosa* (Number of OTUs, `r S.obs_fcym$mean` ± `r S.obs_fcym$sd` OTUs) and lower values for picoplankton communities in the ambient seawater (Number of OTUs, `r S.obs_f$mean` ± `r S.obs_f$sd` OTUs) (\autoref{supp-calculators}).
 Seasonal changes did not show such large dissimilarities.
-Seawater communities richness was stable during the studied period with the exception of one sampling point in December 2017 when larger values were observed.
 *Cymodocea nodosa* communities showed a slow increase towards the end of the study, while *Caulerpa cylindracea* (invaded and noninvaded) communities were characterized by slightly larger values in Spring and Summer in comparison to Autumn and Winter (\autoref{supp-calculators}).
@@
 metadata_rarefied_select <- metadata_rarefied %>%
-  filter(station=="F") %>%
-  select(starts_with("Otu"))
-grouping <- metadata_rarefied %>%
-  filter(station=="F") %>%
-  mutate(grouping=str_replace(season, "Autumn", "Winter"))
-anosim_f <- anosim(metadata_rarefied_select, grouping$grouping, permutations=1000, distance="bray")
-anosim_f_signif <- case_when(anosim_f$signif < 0.05 & anosim_f$signif >= 0.01 ~ 0.05,
-                             anosim_f$signif < 0.01 & anosim_f$signif >= 0.001 ~ 0.01,
-                             anosim_f$signif < 0.001 & anosim_f$signif >= 0.0001 ~ 0.001,
-                             anosim_f$signif < 0.0001 ~ 0.0001,
-                             TRUE ~ anosim_f$signif)
-
-metadata_rarefied_select <- metadata_rarefied %>%
   filter(station=="FCyM") %>%
@@
 This separation was further supported by ANOSIM (R = `r format(anosim_all$statistic, digits=2, nsmall=2)`, *p* < `r anosim_all_signif`).
-Seasonal changes of seawater communities indicated a separation between Spring, Summer and Autumn/Winter samples (ANOSIM, R = `r format(anosim_f$statistic, digits=2, nsmall=2)`, *p* < `r anosim_f_signif`).
-Epiphytic microbial communities associated with *Cymodocea nodosa* showed a similar pattern (ANOSIM, R = `r format(anosim_fcym$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fcym_signif`), while communities from the surfaces of *Caulerpa cylindracea* indicated a non so strongly supported, as in previous cases, separation between Summer and Autumn/Winter/Spring samples (ANOSIM, R = `r format(anosim_fca$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fca_signif`) (\autoref{pcoa}).
+Seasonal changes of *Cymodocea nodosa* associated communities indicated a separation between Spring, Summer and Autumn/Winter samples (ANOSIM, R = `r format(anosim_fcym$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fcym_signif`), while communities from the surfaces of *Caulerpa cylindracea* indicated a non so strongly supported, as in previous cases, separation between Summer and Autumn/Winter/Spring samples (ANOSIM, R = `r format(anosim_fca$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fca_signif`) (\autoref{pcoa}).
 
@@
   summarise(mean=mean(abundance), sd=sd(abundance))
-
-# Calculating the mean and standard deviation for contrasting season of Caulerpa cylindracea
-desulfo_station_season_whole <- inner_join(metadata, whole, by=c("ID"="Group")) %>%
-  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
-  filter(station=="FCaM" | station=="FCa") %>%
-  mutate(season=str_replace(season, "Winter", "Summer")) %>%
-  mutate(season=str_replace(season, "Spring", "Autumn")) %>%
-  group_by(station, season) %>%
-  summarise(mean=mean(abundance), sd=sd(abundance))
 ```
@@
 While seawater and *Cymodocea nodosa* communities were consisted of only `r filter(desulfo_station_whole, station=="F")$mean` ± `r filter(desulfo_station_whole, station=="F")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCyM")$mean` ± `r filter(desulfo_station_whole, station=="FCyM")$sd` \si{\percent} *Desulfobacterota* sequences, respectively, in the invaded and noninvaded *Caulerpa cylindracea* communities their proportion was `r filter(desulfo_station_whole, station=="FCaM")$mean` ± `r filter(desulfo_station_whole, station=="FCaM")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCa")$mean` ± `r filter(desulfo_station_whole, station=="FCa")$sd` \si{\percent}, respectively (\autoref{desulfo}).
-In addition, *Caulerpa cylindracea* associated communities were characterized by higer proportions in Winter and Summer (invaded, `r filter(desulfo_station_season_whole, station=="FCaM", season=="Summer")$mean` ± `r filter(desulfo_station_season_whole, station=="FCaM", season=="Summer")$sd` \si{\percent}; noninvaded, `r filter(desulfo_station_season_whole, station=="FCa", season=="Summer")$mean` ± `r filter(desulfo_station_season_whole, station=="FCa", season=="Summer")$sd` \si{\percent}) in comparison to Autumn and Spring (invaded, `r filter(desulfo_station_season_whole, station=="FCaM", season=="Autumn")$mean` ± `r filter(desulfo_station_season_whole, station=="FCaM", season=="Autumn")$sd` \si{\percent}; noninvaded, `r filter(desulfo_station_season_whole, station=="FCa", season=="Autumn")$mean` ± `r filter(desulfo_station_season_whole, station=="FCa", season=="Autumn")$sd` \si{\percent}).
 The community was mainly consisted of no relative *Desulfobacteraceae*, *Desulfatitalea*, no relative *Desulfobulbaceae*, *Desulfobulbus*, no relative *Desulfocapsaceae*, *Desulfopila*, *Desulforhopalus*, *Desulfotalea*, SEEP-SRB4 and uncultured *Desulfocapsaceae* (\autoref{desulfo}).
@@
 Since the colonization of macrophyte surfaces is performed from a pool of seawater prokaryotic cells it is interesting to see to which extent do these two communities differ.
-We observed a strong differentiation between the surface attached and seawater communities at the level of OTUs that is in agreement with most published studies [@Ugarelli2019; @Roth-Schulze2016; @Burke2011b; @Michelou2013].
+We observed a strong differentiation between the surface attached and seawater communities at the level of OTUs that is in agreement with most published studies [@Ugarelli2019; @Roth-Schulze2016; @Burke2011b; @Michelou2013; @Crump2018].
 These data indicate that marine macrophytes are selecting, from a pool of seawater microbial taxa, the one that can colonize and proliferate on their surfaces [@Salaun2012; @Michelou2013].
@@
 
-Few studies describing microbial epiphytic communities were encompassing seasonal shifts [@Lachnit2011; @Tujula2010; @Michelou2013].
-Seasonal changes in the epiphytic community were substantial as could be observed in the proportion of OTUs that could be find at every sampling time ($\leqslant$ `r max(core$core_otus)` \si{\percent}).
-Interestingly, these OTUs were accounting for a high proportion of sequences ($\leqslant$ `r max(core$core_seq)` \si{\percent}).
+Seasonal richness changes in the epiphytic community were substantial as could be observed in the proportion of OTUs that could be find at every sampling time ($\leq$ `r max(core$core_otus)` \si{\percent}).
+Interestingly, these OTUs were accounting for a high proportion of sequences ($\leq$ `r max(core$core_seq)` \si{\percent}).
 A very similar proportion of persistent OTUs and their sequence contribution was reported in high frequency studies describing seasonal picoplankton changes [@Gilbert2009; @Gilbert2012].
@@
 Similar seasonal changes in the epiphytic community was described also for other macroalgae [@Lachnit2011; @Tujula2010].
-The higher temporal stability of *Caulerpa cylindracea* surface communities in comparison to *Cymodocea nodosa* could be also observed in the higher proportion of shared communities between two consecutive sampling points.
+Higher temporal stability of *Caulerpa cylindracea* surface communities in comparison to *Cymodocea nodosa* were also observed in the higher proportion of shared communities between two consecutive sampling points.
+
+Seasonal patterns of high rank epiphytic taxa did not show strong changes.
+Only *Cyanobateria* were characterized by strong seasonal differences.
+Cyanobacterial sequences were more pronounced in the November and December 2017 in comparison to Spring and Summer.
+Interestingly, in these high proportion months the majority of cyanobacterial sequences were classified as *Pleurocapsa*, a group known to colonized different living and nonliving surfaces [@Longford2007; @Burns2004; @Mobberley2012; @Reisser2014].
+It is possible than during periods of low metabolic activity of the i.g.
+seagrass there is no so active interaction and selection of the epiphytic community, so the seagrass leaves are becoming a suitable surface for nonspecific colonizers [@Zavodnik1998].
+*Pleurocapsa* was replaced in Spring ans Summer by *Acrophormium*, *Phormidesmis* and no relative *Nodosilineaceae*.
+A study of coastal microbial mats found also higher proportion of *Nodosilineaceae* sequences in Summer, while *Phormidesmis* sequences were at their peak in Autumn [@Cardoso2019].
+Other high rank taxa did not showed strong successional patterns.
+In each analyzed group low rank taxa
 
@@
 
-```{r out.width="100%", fig.align="center", fig.cap="Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+```{r out.width="55%", fig.align="center", fig.cap="Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
 Samples from the same environment or same season are labeld in different colors.

### 2020-03-09 submission/manuscript.tex
Additional discussion.

@@
 
+\usepackage{txfonts}
+
 \begin{document}
@@
 was sequenced two times) originating from picoplankton archaeal and
-bacterial communities in the surrounding seawater were also processed.
-The number of reads per sample ranged between 8,407 and 77,465 sequences
-(\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
-applied the rarefaction curves did not level off that is a common
-observation in high-throughput 16S rRNA amplicon sequencing approaches
-(\autoref{supp-rarefaction}). Following quality curation and exclusion
-of sequences mentioned before reads were clustered into 28,726 different
-OTUs at a similarity level of 97 \si{\percent}. Reads numbers were
-normalized to the minimum number of sequences, 8,407
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,007
-different OTUs that ranged from 366 to 1,998 OTUs per sample
-(\autoref{supp-calculators}). To determine seasonal changes of richness
-and diversity the Observed Number of OTUs, Chao1, ACE, Exponential
-Shannon (Jost, 2006) and Inverse Simpson were calculated after
-normalization through rarefaction. Generally, richness estimators and
-diversity indices showed similar trends. On average, higher values were
-found for \emph{Caulerpa cylindracea} (invaded {[}Number of OTUs,
+bacterial communities in the ambient seawater were also processed for
+comparison. The number of reads per sample ranged between 8,407 and
+77,465 sequences (\autoref{supp-nseq_notus}). Even when the highest
+sequencing effort was applied the rarefaction curves did not level off
+that is a common observation in high-throughput 16S rRNA amplicon
+sequencing approaches (\autoref{supp-rarefaction}). Following quality
+curation and exclusion of sequences mentioned before reads were
+clustered into 28,726 different OTUs at a similarity level of 97
+\si{\percent}. Reads numbers were normalized to the minimum number of
+sequences, 8,407 (\autoref{supp-nseq_notus}), through rarefaction
+resulting in 17,007 different OTUs that ranged from 366 to 1,998 OTUs
+per sample (\autoref{supp-calculators}). To determine seasonal changes
+of richness and diversity the Observed Number of OTUs, Chao1, ACE,
+Exponential Shannon (Jost, 2006) and Inverse Simpson were calculated
+after normalization through rarefaction. Generally, richness estimators
+and diversity indices showed similar trends. On average, higher values
+were found for \emph{Caulerpa cylindracea} (invaded {[}Number of OTUs,
 1,688.5 ± 125.6 OTUs{]} and noninvaded {[}Number of OTUs, 1,744.8 ±
@@
 OTUs, 1,062.8 ± 209.6 OTUs) and lower values for picoplankton
-communities in the surrounding seawater (Number of OTUs, 528.1 ± 135.8
-OTUs) (\autoref{supp-calculators}). Seasonal changes did not show such
-large dissimilarities. Seawater communities richness was stable during
-the studied period with the exception of one sampling point in December
-2017 when larger values were observed. \emph{Cymodocea nodosa}
-communities showed a slow increase towards the end of the study, while
-\emph{Caulerpa cylindracea} (invaded and noninvaded) communities were
-characterized by slightly larger values in Spring and Summer in
-comparison to Autumn and Winter (\autoref{supp-calculators}).
+communities in the ambient seawater (Number of OTUs, 528.1 ± 135.8 OTUs)
+(\autoref{supp-calculators}). Seasonal changes did not show such large
+dissimilarities. \emph{Cymodocea nodosa} communities showed a slow
+increase towards the end of the study, while \emph{Caulerpa cylindracea}
+(invaded and noninvaded) communities were characterized by slightly
+larger values in Spring and Summer in comparison to Autumn and Winter
+(\autoref{supp-calculators}).
 
@@
 species was determined. This separation was further supported by ANOSIM
-(R = 0.96, \emph{p} \textless{} 0.001). Seasonal changes of seawater
-communities indicated a separation between Spring, Summer and
-Autumn/Winter samples (ANOSIM, R = 0.64, \emph{p} \textless{} 0.001).
-Epiphytic microbial communities associated with \emph{Cymodocea nodosa}
-showed a similar pattern (ANOSIM, R = 0.56, \emph{p} \textless{} 0.01),
-while communities from the surfaces of \emph{Caulerpa cylindracea}
-indicated a non so strongly supported, as in previous cases, separation
-between Summer and Autumn/Winter/Spring samples (ANOSIM, R = 0.31,
-\emph{p} \textless{} 0.05) (\autoref{pcoa}).
+(R = 0.96, \emph{p} \textless{} 0.001). Seasonal changes of
+\emph{Cymodocea nodosa} associated communities indicated a separation
+between Spring, Summer and Autumn/Winter samples (ANOSIM, R = 0.56,
+\emph{p} \textless{} 0.01), while communities from the surfaces of
+\emph{Caulerpa cylindracea} indicated a non so strongly supported, as in
+previous cases, separation between Summer and Autumn/Winter/Spring
+samples (ANOSIM, R = 0.31, \emph{p} \textless{} 0.05) (\autoref{pcoa}).
 
@@
 25.7 ± 11.3 \si{\percent} and 24.0 ± 4.3 \si{\percent}, respectively
-(\autoref{desulfo}). In addition, \emph{Caulerpa cylindracea} associated
-communities were characterized by higer proportions in Winter and Summer
-(invaded, 30.9 ± 12.4 \si{\percent}; noninvaded, 26.9 ± 3.4
-\si{\percent}) in comparison to Autumn and Spring (invaded, 22.8 ± 10.3
-\si{\percent}; noninvaded, 21.9 ± 3.8 \si{\percent}). The community was
-mainly consisted of no relative \emph{Desulfobacteraceae},
-\emph{Desulfatitalea}, no relative \emph{Desulfobulbaceae},
-\emph{Desulfobulbus}, no relative \emph{Desulfocapsaceae},
-\emph{Desulfopila}, \emph{Desulforhopalus}, \emph{Desulfotalea},
-SEEP-SRB4 and uncultured \emph{Desulfocapsaceae} (\autoref{desulfo}).
+(\autoref{desulfo}). The community was mainly consisted of no relative
+\emph{Desulfobacteraceae}, \emph{Desulfatitalea}, no relative
+\emph{Desulfobulbaceae}, \emph{Desulfobulbus}, no relative
+\emph{Desulfocapsaceae}, \emph{Desulfopila}, \emph{Desulforhopalus},
+\emph{Desulfotalea}, SEEP-SRB4 and uncultured \emph{Desulfocapsaceae}
+(\autoref{desulfo}).
 
@@
 al.}, 2011b; Michelou \emph{et al.}, 2013; Roth-Schulze \emph{et al.},
-2016; Ugarelli \emph{et al.}, 2019). These data indicate that marine
-macrophytes are selecting, from a pool of seawater microbial taxa, the
-one that can colonize and proliferate on their surfaces (Salaün \emph{et
-al.}, 2012; Michelou \emph{et al.}, 2013). In contrast to these findings
-Fahimipour \emph{et al.} (2017) found, in a global study of
-\emph{Zostera marina}, similarities between leaves and seawater samples.
-Discrepancies between our data and this study could be explained by
-differences in studied seagrass species, methodological variations or
-biogeographic trends as Fahimipour \emph{et al.} (2017) were analyzing
-samples from different locations throughout the northern hemisphere. It
-is possible that ambient seawater and leaves communities from the same
-location are differing but are still more similar to each other when
-compared to other sampling locations. Indeed, it was found that
-prokaryotic communities vary substantially between different sampling
-sites (Bengtsson \emph{et al.}, 2017). When the taxonomic composition at
-high ranks was analyzed no such strong differentiation was noticed.
-Phyla and classes such as: \emph{Actinobacteriota}, \emph{Bacteroidota},
-\emph{Cyanobacteria}, \emph{Alphaproteobacteria},
-\emph{Gammaproteobacteria} and \emph{Verrucomicrobiota}, were described
-that is in agreement with previously reported data (Burke \emph{et al.},
-2011b; Egan \emph{et al.}, 2013; Michelou \emph{et al.}, 2013). In
-contrast, when low taxonomic ranks were analyzed (i.g. family and genus)
-a strong differentiation was observed. A similar differentiation at
-lower taxonomic ranks was described for other species of macrophytes
-(Egan \emph{et al.}, 2013; Michelou \emph{et al.}, 2013; Ugarelli
-\emph{et al.}, 2019).
+2016; Crump \emph{et al.}, 2018; Ugarelli \emph{et al.}, 2019). These
+data indicate that marine macrophytes are selecting, from a pool of
+seawater microbial taxa, the one that can colonize and proliferate on
+their surfaces (Salaün \emph{et al.}, 2012; Michelou \emph{et al.},
+2013). In contrast to these findings Fahimipour \emph{et al.} (2017)
+found, in a global study of \emph{Zostera marina}, similarities between
+leaves and seawater samples. Discrepancies between our data and this
+study could be explained by differences in studied seagrass species,
+methodological variations or biogeographic trends as Fahimipour \emph{et
+al.} (2017) were analyzing samples from different locations throughout
+the northern hemisphere. It is possible that ambient seawater and leaves
+communities from the same location are differing but are still more
+similar to each other when compared to other sampling locations. Indeed,
+it was found that prokaryotic communities vary substantially between
+different sampling sites (Bengtsson \emph{et al.}, 2017). When the
+taxonomic composition at high ranks was analyzed no such strong
+differentiation was noticed. Phyla and classes such as:
+\emph{Actinobacteriota}, \emph{Bacteroidota}, \emph{Cyanobacteria},
+\emph{Alphaproteobacteria}, \emph{Gammaproteobacteria} and
+\emph{Verrucomicrobiota}, were described that is in agreement with
+previously reported data (Burke \emph{et al.}, 2011b; Egan \emph{et
+al.}, 2013; Michelou \emph{et al.}, 2013). In contrast, when low
+taxonomic ranks were analyzed (i.g. family and genus) a strong
+differentiation was observed. A similar differentiation at lower
+taxonomic ranks was described for other species of macrophytes (Egan
+\emph{et al.}, 2013; Michelou \emph{et al.}, 2013; Ugarelli \emph{et
+al.}, 2019).
 
@@
 
-Few studies describing microbial epiphytic communities were encompassing
-seasonal shifts (Tujula \emph{et al.}, 2010; Lachnit \emph{et al.},
-2011; Michelou \emph{et al.}, 2013). Seasonal changes in the epiphytic
-community were substantial as could be observed in the proportion of
-OTUs that could be find at every sampling time (\(\leqslant\) 1.0
-\si{\percent}). Interestingly, these OTUs were accounting for a high
-proportion of sequences (\(\leqslant\) 51.3 \si{\percent}). A very
-similar proportion of persistent OTUs and their sequence contribution
-was reported in high frequency studies describing seasonal picoplankton
-changes (Gilbert \emph{et al.}, 2009, 2012). In comparison to the
-seawater community, a lower degree of seasonal shifts was observed for
-the surface associated communities. It seems, that microniches on the
-surfaces of macrophytes are providing more stable conditions in
-comparison to the seawater. At the level of OTUs seasonal changes of
-\emph{Cymodocea nodosa} and \emph{Caulerpa cylindracea} associated
-communities were identified that could be linked to the growth cycle of
-the seagrass and macroalgae (M. Najdek, personal communication).
-\emph{Cymodocea nodosa} was characterized by a Spring community during
-maximum seagrass proliferation, a Summer community during a biomass
-maximum and a Autumn/Winter community during a biomass decay. In
-contrast, \emph{Caulerpa cylindracea} started to proliferate in late
-Spring and was characterized only by a Summer community during maximal
-biomass increase and by a Autumn/Winter/Spring community when the
-biomass were at the peak and the settlement started to subsequently
+Seasonal richness changes in the epiphytic community were substantial as
+could be observed in the proportion of OTUs that could be find at every
+sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
+were accounting for a high proportion of sequences (\(\leq\) 51.3
+\si{\percent}). A very similar proportion of persistent OTUs and their
+sequence contribution was reported in high frequency studies describing
+seasonal picoplankton changes (Gilbert \emph{et al.}, 2009, 2012). In
+comparison to the seawater community, a lower degree of seasonal shifts
+was observed for the surface associated communities. It seems, that
+microniches on the surfaces of macrophytes are providing more stable
+conditions in comparison to the seawater. At the level of OTUs seasonal
+changes of \emph{Cymodocea nodosa} and \emph{Caulerpa cylindracea}
+associated communities were identified that could be linked to the
+growth cycle of the seagrass and macroalgae (M. Najdek, personal
+communication). \emph{Cymodocea nodosa} was characterized by a Spring
+community during maximum seagrass proliferation, a Summer community
+during a biomass maximum and a Autumn/Winter community during a biomass
+decay. In contrast, \emph{Caulerpa cylindracea} started to proliferate
+in late Spring and was characterized only by a Summer community during
+maximal biomass increase and by a Autumn/Winter/Spring community when
+the biomass were at the peak and the settlement started to subsequently
 decay. Similar seasonal changes in the epiphytic community was described
 also for other macroalgae (Tujula \emph{et al.}, 2010; Lachnit \emph{et
-al.}, 2011). The higher temporal stability of \emph{Caulerpa
-cylindracea} surface communities in comparison to \emph{Cymodocea
-nodosa} could be also observed in the higher proportion of shared
-communities between two consecutive sampling points.
+al.}, 2011). Higher temporal stability of \emph{Caulerpa cylindracea}
+surface communities in comparison to \emph{Cymodocea nodosa} were also
+observed in the higher proportion of shared communities between two
+consecutive sampling points.
+
+Seasonal patterns of high rank epiphytic taxa did not show strong
+changes. Only \emph{Cyanobateria} were characterized by strong seasonal
+differences. Cyanobacterial sequences were more pronounced in the
+November and December 2017 in comparison to Spring and Summer.
+Interestingly, in these high proportion months the majority of
+cyanobacterial sequences were classified as \emph{Pleurocapsa}, a group
+known to colonized different living and nonliving surfaces (Burns
+\emph{et al.}, 2004; Longford \emph{et al.}, 2007; Mobberley \emph{et
+al.}, 2012; Reisser \emph{et al.}, 2014). It is possible than during
+periods of low metabolic activity of the i.g. seagrass there is no so
+active interaction and selection of the epiphytic community, so the
+seagrass leaves are becoming a suitable surface for nonspecific
+colonizers (Zavodnik \emph{et al.}, 1998). \emph{Pleurocapsa} was
+replaced in Spring ans Summer by \emph{Acrophormium},
+\emph{Phormidesmis} and no relative \emph{Nodosilineaceae}. A study of
+coastal microbial mats found also higher proportion of
+\emph{Nodosilineaceae} sequences in Summer, while \emph{Phormidesmis}
+sequences were at their peak in Autumn (Cardoso \emph{et al.}, 2019).
+Other high rank taxa did not showed strong successional patterns. In
+each analyzed group low rank taxa
 
@@
 
+\leavevmode\hypertarget{ref-Burns2004}{}%
+Burns, B.P., Goh, F., Allen, M., and Neilan, B.A. (2004) Microbial
+diversity of extant stromatolites in the hypersaline marine environment
+of Shark Bay, Australia. \emph{Environmental Microbiology} \textbf{6}:
+1096--1101.
+
 \leavevmode\hypertarget{ref-Cai2014}{}%
@@
 
+\leavevmode\hypertarget{ref-Cardoso2019}{}%
+Cardoso, D.C., Cretoiu, M.S., Stal, L.J., and Bolhuis, H. (2019)
+Seasonal development of a coastal microbial mat. \emph{Scientific
+Reports} \textbf{9}: 1--14.
+
 \leavevmode\hypertarget{ref-Crump2008}{}%
@@
 microbiology} \textbf{74}: 5948--5957.
+
+\leavevmode\hypertarget{ref-Crump2018}{}%
+Crump, B.C., Wojahn, J.M., Tomas, F., and Mueller, R.S. (2018)
+Metatranscriptomics and amplicon sequencing reveal mutualisms in
+seagrass microbiomes. \emph{Frontiers in Microbiology} \textbf{9}:
 
@@
 
+\leavevmode\hypertarget{ref-Longford2007}{}%
+Longford, S., Tujula, N., Crocetti, G., Holmes, A., Holmström, C.,
+Kjelleberg, S., et al. (2007) Comparisons of diversity of bacterial
+communities associated with three sessile marine eukaryotes.
+\emph{Aquatic Microbial Ecology} \textbf{48}: 217--229.
+
 \leavevmode\hypertarget{ref-Margulis1991}{}%
@@
 
+\leavevmode\hypertarget{ref-Mobberley2012}{}%
+Mobberley, J.M., Ortega, M.C., and Foster, J.S. (2012) Comparative
+microbial diversity analyses of modern marine thrombolitic mats by
+barcoded pyrosequencing. \emph{Environmental Microbiology} \textbf{14}:
+82--100.
+
 \leavevmode\hypertarget{ref-Morrissey2019}{}%
@@
 Computing, Vienna, Austria: R Foundation for Statistical Computing.
+
+\leavevmode\hypertarget{ref-Reisser2014}{}%
+Reisser, J., Shaw, J., Hallegraeff, G., Proietti, M., Barnes, D.K.A.,
+Thums, M., et al. (2014) Millimeter-sized marine plastics: a new pelagic
+habitat for microorganisms and invertebrates. \emph{PloS one}
+\textbf{9}: e100289.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{../results/figures/pcoa_figure}
+{\centering \includegraphics[width=0.55\linewidth]{../results/figures/pcoa_figure}
 

### 2020-03-09 code/plot_pcoa.R

@@ -70,32 +70,6 @@ p1 <- ggplot() +
   scale_y_continuous(labels=scaleFUN) +
   theme
 
-# Seawater samples
-# Filter samples from the water column
-rarefied_metadata_select <- filter(rarefied_metadata, station=="F") %>%
-  select_if(list(~!is.numeric(.) || sum(.)!=0))
-
-# Generation of PCoA data
-spe.bray <- vegdist(select(rarefied_metadata_select, starts_with("Otu")))
-spe.b.pcoa <- cmdscale(spe.bray, k=(nrow(rarefied_metadata_select)-1), eig=TRUE)
-coordinates <- spe.b.pcoa$points[,1:2] %>%
-  as_tibble() %>%
-  add_column("Group"=rarefied_metadata_select$label, .before=TRUE)
-coordinates <- inner_join(metadata, coordinates, by=c("label"="Group"))
-
-# PCoA plot generation
-p2 <- ggplot() +
-  geom_point(data=coordinates, aes(x=V1, y=V2, fill=season), shape=21, size=3, stroke=0.5) +
-  scale_fill_manual(name=NULL,
-                    breaks=c("Spring", "Summer", "Autumn", "Winter"),
-                    values=brewer.pal(n=4, name="Set1")) +
-  labs(x=paste0("PCoA I (",format(round(spe.b.pcoa$eig[1]/sum(spe.b.pcoa$eig)*100, digits=2), nsmall=2)," %)"), 
-       y=paste0("PCoA II (",format(round(spe.b.pcoa$eig[2]/sum(spe.b.pcoa$eig)*100, digits=2), nsmall=2)," %)")) +
-  ggtitle(parse(text="bold('Seawater')")) +
-  scale_x_continuous(labels=scaleFUN) +
-  scale_y_continuous(labels=scaleFUN) +
-  theme
-
 # Cymodocea nodosa samples
 # Filter samples from the water column
 rarefied_metadata_select <- filter(rarefied_metadata, station=="FCyM") %>%
@@ -110,7 +84,7 @@ coordinates <- spe.b.pcoa$points[,1:2] %>%
 coordinates <- inner_join(metadata, coordinates, by=c("label"="Group"))
 
 # PCoA plot generation
-p3 <- ggplot() +
+p2 <- ggplot() +
   geom_point(data=coordinates, aes(x=V1, y=V2, fill=season), shape=21, size=3, stroke=0.5) +
   scale_fill_manual(name=NULL,
[…]

### 2020-03-09 submission/manuscript.tex
Added OTU taxonomy.

@@
 bacterial communities in the ambient seawater were also processed for
-comparison. The number of reads per sample ranged between 8,407 and
+comparison. The number of reads per sample ranged between 8,410 and
 77,465 sequences (\autoref{supp-nseq_notus}). Even when the highest
@@
 curation and exclusion of sequences mentioned before reads were
-clustered into 28,726 different OTUs at a similarity level of 97
+clustered into 28,729 different OTUs at a similarity level of 97
 \si{\percent}. Reads numbers were normalized to the minimum number of
-sequences, 8,407 (\autoref{supp-nseq_notus}), through rarefaction
-resulting in 17,007 different OTUs that ranged from 366 to 1,998 OTUs
+sequences, 8,410 (\autoref{supp-nseq_notus}), through rarefaction
+resulting in 17,025 different OTUs that ranged from 351 to 2,059 OTUs
 per sample (\autoref{supp-calculators}). To determine seasonal changes
@@
 were found for \emph{Caulerpa cylindracea} (invaded {[}Number of OTUs,
-1,688.5 ± 125.6 OTUs{]} and noninvaded {[}Number of OTUs, 1,744.8 ±
-150.8 OTUs{]}), middle vales for \emph{Cymodocea nodosa} (Number of
-OTUs, 1,062.8 ± 209.6 OTUs) and lower values for picoplankton
-communities in the ambient seawater (Number of OTUs, 528.1 ± 135.8 OTUs)
+1,694.5 ± 133.7 OTUs{]} and noninvaded {[}Number of OTUs, 1,738.2 ±
+168.5 OTUs{]}), middle vales for \emph{Cymodocea nodosa} (Number of
+OTUs, 1,057.7 ± 207.3 OTUs) and lower values for picoplankton
+communities in the ambient seawater (Number of OTUs, 519 ± 139.9 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 community was found between invaded and noninvaded \emph{Caulerpa
-cylindracea} (Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared
+cylindracea} (Jaccard, 0.35; Bray-Curtis, 0.78), while lower shared
 values were calculated between seawater and epiphytic communities
@@
 \emph{Cymodocea nodosa} and especially \emph{Caulerpa cylindracea}
-associated communities (\autoref{shared}). In addition, only 0.4 -- 1.0
+associated communities (\autoref{shared}). In addition, only 0.4 -- 1.1
 \si{\percent} of OTUs from each surface associated community were found
 at every time point. These OTUs made also a high proportion of total
-sequences (39.1 -- 51.3 \si{\percent}). To further disentangle the
+sequences (35.4 -- 53.1 \si{\percent}). To further disentangle the
 environmental and seasonal community dissimilarity a Principal
@@
 \emph{Cymodocea nodosa} associated communities indicated a separation
-between Spring, Summer and Autumn/Winter samples (ANOSIM, R = 0.56,
+between Spring, Summer and Autumn/Winter samples (ANOSIM, R = 0.55,
 \emph{p} \textless{} 0.01), while communities from the surfaces of
@@
 previous cases, separation between Summer and Autumn/Winter/Spring
-samples (ANOSIM, R = 0.31, \emph{p} \textless{} 0.05) (\autoref{pcoa}).
+samples (ANOSIM, R = 0.32, \emph{p} \textless{} 0.01) (\autoref{pcoa}).
 
@@
 noninvaded \emph{Caulerpa cylindracea} communities their proportion was
-25.7 ± 11.3 \si{\percent} and 24.0 ± 4.3 \si{\percent}, respectively
+25.7 ± 11.2 \si{\percent} and 24.0 ± 4.3 \si{\percent}, respectively
 (\autoref{desulfo}). The community was mainly consisted of no relative
@@
 could be observed in the proportion of OTUs that could be find at every
-sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 51.3
+sampling time (\(\leq\) 1.1 \si{\percent}). Interestingly, these OTUs
+were accounting for a high proportion of sequences (\(\leq\) 53.1
 \si{\percent}). A very similar proportion of persistent OTUs and their

### 2020-03-09 submission/supplementary.tex
Added OTU taxonomy.

--- 2020-02-14 submission/supplementary.tex
@@
 
+\hypertarget{supplementary-table}{%
+\subsection{Supplementary Table}\label{supplementary-table}}
+
+\begingroup\fontsize{9}{11}\selectfont
+
+\begin{longtable}[t]{>{\centering\arraybackslash}p{6em}ccccc}
+\caption{\label{tab:nseq_notus}Sample ID, Community Type, Sampling Date and Season, No.
+of Sequences and No.
+of OTUs of each sample.
+No.
+of Sequences and OTUs was calculated after the exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences.\label{nseq_notus}}\\
+\toprule
+\textbf{Sample ID} & \textbf{Community Type} & \textbf{Date} & \textbf{Season} & \textbf{No. of Sequences} & \textbf{No. of OTUs}\\
+\midrule
+\endfirsthead
+\caption[]{Sample ID, Community Type, Sampling Date and Season, No.
+of Sequences and No.
+of OTUs of each sample.
+No.
+of Sequences and OTUs was calculated after the exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences.\label{nseq_notus} \textit{(continued)}}\\
+\toprule
+\textbf{Sample ID} & \textbf{Community Type} & \textbf{Date} & \textbf{Season} & \textbf{No. of Sequences} & \textbf{No. of OTUs}\\
+\midrule
+\endhead
+\
+\endfoot
+\bottomrule
+\endlastfoot
+3 & Seawater & 13 July 2017 & Summer & 26,008 & 824\\
+5 & Seawater & 27 July 2017 & Summer & 58,955 & 993\\
+7 & Seawater & 10 August 2017 & Summer & 32,626 & 852\\
+9 & Seawater & 24 August 2017 & Summer & 60,929 & 1,058\\
+11 & Seawater & 19 September 2017 & Summer & 46,111 & 1,171\\
+13 & Seawater & 12 October 2017 & Autumn & 30,937 & 976\\
+15 & Seawater & 23 November 2017 & Autumn & 27,581 & 1,183\\
+17 & Seawater & 14 December 2017 & Autumn & 44,595 & 2,267\\
+19 & Seawater & 13 February 2018 & Winter & 44,186 & 1,309\\
+21 & Seawater & 27 March 2018 & Winter & 46,307 & 1,559\\
+23a & Seawater & 24 April 2018 & Spring & 30,970 & 998\\
+23b & Seawater & 24 April 2018 & Spring & 38,560 & 1,200\\
+25 & Seawater & 22 May 2018 & Spring & 53,880 & 1,783\\
+27 & Seawater & 19 June 2018 & Spring & 77,465 & 1,297\\
+29 & Seawater & 10 July 2018 & Summer & 50,794 & 1,451\\
+31 & Seawater & 9 August 2018 & Summer & 39,049 & 1,129\\
+33 & Seawater & 4 September 2018 & Summer & 36,206 & 1,205\\
+35 & Seawater & 5 October 2018 & Autumn & 49,580 & 1,013\\
+37 & \textit{Cymodocea nodosa} (Invaded) & 23 November 2017 & Autumn & 31,235 & 1,548\\
+41 & \textit{Cymodocea nodosa} (Invaded) & 4 December 2017 & Autumn & 24,237 & 1,176\\
+45 & \textit{Cymodocea nodosa} (Invaded) & 14 December 2017 & Autumn & 32,683 & 1,051\\
+49 & \textit{Cymodocea nodosa} (Invaded) & 13 February 2018 & Winter & 9,087 & 1,218\\
+52 & \textit{Cymodocea nodosa} (Invaded) & 27 March 2018 & Winter & 17,003 & 1,218\\
+55 & \textit{Cymodocea nodosa} (Invaded) & 24 April 2018 & Spring & 42,640 & 2,054\\
+58 & \textit{Cymodocea nodosa} (Invaded) & 22 May 2018 & Spring & 21,337 & 1,292\\
+61 & \textit{Cymodocea nodosa} (Invaded) & 19 June 2018 & Spring & 31,735 & 2,104\\
+64 & \textit{Cymodocea nodosa} (Invaded) & 10 July 2018 & Summer & 35,733 & 1,792\\
+67 & \textit{Cymodocea nodosa} (Invaded) & 9 August 2018 & Summer & 26,362 & 2,119\\
+70 & \textit{Cymodocea nodosa} (Invaded) & 4 September 2018 & Summer & 23,278 & 1,715\\
+73 & \textit{Cymodocea nodosa} (Invaded) & 5 October 2018 & Autumn & 29,909 & 2,205\\
+38 & \textit{Caulerpa cylindracea} (Invaded) & 23 November 2017 & Autumn & 36,321 & 2,889\\
+42 & \textit{Caulerpa cylindracea} (Invaded) & 4 December 2017 & Autumn & 28,370 & 3,249\\
+46 & \textit{Caulerpa cylindracea} (Invaded) & 14 December 2017 & Autumn & 34,720 & 3,054\\
+53 & \textit{Caulerpa cylindracea} (Invaded) & 27 March 2018 & Winter & 28,696 & 2,472\\
+56 & \textit{Caulerpa cylindracea} (Invaded) & 24 April 2018 & Spring & 34,773 & 3,054\\
+59 & \textit{Caulerpa cylindracea} (Invaded) & 22 May 2018 & Spring & 23,397 & 2,706\\
+62 & \textit{Caulerpa cylindracea} (Invaded) & 19 June 2018 & Spring & 36,477 & 3,299\\
+65 & \textit{Caulerpa cylindracea} (Invaded) & 10 July 2018 & Summer & 18,485 & 2,188\\
+68 & \textit{Caulerpa cylindracea} (Invaded) & 9 August 2018 & Summer & 31,940 & 3,103\\
+71 & \textit{Caulerpa cylindracea} (Invaded) & 4 September 2018 & Summer & 29,281 & 3,160\\
+74 & \textit{Caulerpa cylindracea} (Invaded) & 5 October 2018 & Autumn & 11,705 & 1,705\\
+39 & \textit{Caulerpa cylindracea} (Nonnvaded) & 23 November 2017 & Autumn & 22,064 & 2,438\\
+43 & \textit{Caulerpa cylindracea} (Nonnvaded) & 4 December 2017 & Autumn & 13,661 & 1,887\\
+47 & \textit{Caulerpa cylindracea} (Nonnvaded) & 14 December 2017 & Autumn & 8,410 & 1,451\\
+51 & \textit{Caulerpa cylindracea} (Nonnvaded) & 13 February 2018 & Winter & 11,676 & 1,901\\
+54 & \textit{Caulerpa cylindracea} (Nonnvaded) & 27 March 2018 & Winter & 39,472 & 3,136\\
+57 & \textit{Caulerpa cylindracea} (Nonnvaded) & 24 April 2018 & Spring & 20,298 & 2,822\\
+60 & \textit{Caulerpa cylindracea} (Nonnvaded) & 22 May 2018 & Spring & 33,047 & 3,297\\
+63 & \textit{Caulerpa cylindracea} (Nonnvaded) & 19 June 2018 & Spring & 41,833 & 3,968\\
+66 & \textit{Caulerpa cylindracea} (Nonnvaded) & 10 July 2018 & Summer & 27,036 & 2,674\\
+69 & \textit{Caulerpa cylindracea} (Nonnvaded) & 9 August 2018 & Summer & 26,737 & 3,130\\
+72 & \textit{Caulerpa cylindracea} (Nonnvaded) & 4 September 2018 & Summer & 31,869 & 3,240\\
+75 & \textit{Caulerpa cylindracea} (Nonnvaded) & 5 October 2018 & Autumn & 33,068 & 3,086\\*
+\end{longtable}
+\endgroup{}
+
 \end{document}

### 2020-03-09 code/plot_calculators.R

@@ -15,7 +15,7 @@ shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.
 rarefied <- shared %>%
   select(-label, -Group, -numOtus) %>%
   rrarefy(., min(rowSums(.))) %>%
-  as.tibble() %>%
+  as_tibble() %>%
   add_column("Group"=shared$Group, .before=TRUE) %>%
   select_if(list(~!is.numeric(.) || sum(.)!=0))
 
@@ -28,7 +28,7 @@ rarefied <- rarefied %>%
 
 # Calculating Chao1 and ACE species estimators
 estimators <- estimateR(rarefied)
-estimators <- as.tibble(t(estimators)) %>%
+estimators <- as_tibble(t(estimators)) %>%
   add_column("Group"=colnames(estimators), .before=TRUE)
 
 # Calculating diversity indices

### 2020-03-09 code/plot_seasonal_shared.R

@@ -15,7 +15,7 @@ shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.
 rarefied <- shared %>%
   select(-label, -Group, -numOtus) %>%
   rrarefy(., min(rowSums(.))) %>%
-  as.tibble() %>%
+  as_tibble() %>%
   add_column("Group"=shared$Group, .before=TRUE) %>%
   select_if(list(~ !is.numeric(.) || sum(.)!=0))
 
@@ -28,11 +28,11 @@ rarefied <- rarefied %>%
 
 # Calculating dissimilarity indices
 jaccard <- vegdist(rarefied, method="jaccard", binary=T)
-jaccard <- as.tibble(data.frame(t(combn(rownames(rarefied),2)), as.numeric(jaccard))) %>%
+jaccard <- as_tibble(data.frame(t(combn(rownames(rarefied),2)), as.numeric(jaccard))) %>%
   rename(V1=X1, V2=X2, jaccard=as.numeric.jaccard.)
 
 bray <- vegdist(rarefied, method="bray", binary=F)
-bray <- as.tibble(data.frame(t(combn(rownames(rarefied),2)), as.numeric(bray))) %>%
+bray <- as_tibble(data.frame(t(combn(rownames(rarefied),2)), as.numeric(bray))) %>%
   rename(V1=X1, V2=X2, bray=as.numeric.bray.)
 
 distance <- inner_join(jaccard, bray, by=c("V1"="V1", "V2"="V2")) %>%

### 2020-03-12 submission/manuscript.Rmd
Manuscript 1st version complete.

@@
 opts_chunk$set("warning" = FALSE)
-opts_chunk$set("cache" = T)
+opts_chunk$set("cache" = FALSE)
 opts_chunk$set("message" = FALSE)
@@
 
-\vspace{70mm}
-
-^1$\dagger$^
+\vspace{50mm}
+Marino Korlević^1$*$^, Marsej Markovski^1^, Zihao Zhao^2^, Gerhard J. Herndl^2,3^, Mirjana Najdek^1^
 
@@
 
-$\dagger$ To whom correspondence should be addressed: marino.korlevic@irb.hr
+^$*$^To whom correspondence should be addressed: marino.korlevic@irb.hr
 
-1\. Ruđer Bošković Institute, Center for Marine Research, G. Paliaga 5, Rovinj, Croatia
+1\. Center for Marine Research, Ruđer Bošković Institute, Croatia
 
-2\. University of Vienna, Department of Limnology and Bio-Oceanography, Althanstraße 14, Vienna, Austria
+2\. Department of Functional and Evolutionary Ecology, University of Vienna, Austria
+
+3\. Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research, Utrecht University, The Netherlands
+
 \newpage
@@
 ## Abstract
+Surfaces of marine macrophytes (seagrasses and macroalgae) are inhabited by diverse microbial communities.
+Most studies focusing on macrophyte epiphytic communities did not take into account temporal changes or applied low sampling frequency approaches.
+Illumina sequencing of the V4 16S rRNA region was performed to determine the seasonal dynamics of epiphytic communities sampled from the surfaces of the seagrass *Cymodocea nodosa* and invasive macroalga *Caulerpa cylindracea*.
+Leaves and thalli were sampled in a meadow of *Cymodocea nodosa* invaded by the invasive *Caulerpa cylindracea* and in a monospecific settlement of *Caulerpa cylindracea* located in the proximity of the meadow at monthly intervals.
+For comparison the ambient prokaryotic plankton community was also characterized.
+Sequencing results at the OTU level showed a clear differentiation between ambient water and epiphytic communities and a host-specific community assemblage.
+In addition, successional changes were observed that could be connected to the macrophyte growth cycle.
+Taxonomic analysis showed similar high rank groups in the ambient water and epiphytic communities, with the exception of *Desulfobacterota* that were found only on *Caulerpa cylindracea*.
+Only *Cyanobacteria* showed seasonal change, while other high rank taxa were present throughout the year.
+In every analyzed high rank taxa, phylogenetic groups present throughout the year and comprising most of the sequences could be identified together with low proportion taxa showing seasonal patterns connected to the macrophyte growth cycle.
+Taken together, epiphytic microbial communities of the seagrass *Cymodocea nodosa* and the macroalgae *Caulerpa cylindracea* appear to be host-specific and contain taxa that undergo successional changes.
 
@@
 It postulates that an initial random colonization step is performed from a set of functionally equivalent taxonomic groups resulting in taxonomically different epiphytic communities sharing a core set of functional genes [@Burke2011a; @Roth-Schulze2016].
-In addition, some of the variation in the observed data could be attributed to different techniques used in various studies, such as different protocols for epiphytic cell detachment and/or DNA isolation, as no standard protocol to study epiphytic communities was established [@Ugarelli2019; @Korlevic_in_press, submitted].
+In addition, some of the variation in the observed data could be attributed to different techniques used in various studies, such as different protocols for epiphytic cell detachment and/or DNA isolation, as no standard protocol to study epiphytic communities was established [@Ugarelli2019; @Korlevic_submitted, submitted].
 
@@
 In the present study we describe the seasonal dynamics of bacterial and archaeal communities on the surfaces of the seagrass *Cymodocea nodosa* and siphonous macroalgae *Caulerpa cylindracea* determined on a mostly monthly scale.
-Bacterial and archaeal epiphytes were sampled in a meadow of *Cymodocea nodosa* invaded by the invasive *Caulerpa cylindracea* and in a locality of only *Caulerpa cylindracea* located in the proximity of the meadow.
-In addition, for comparison, the community of the surrounding seawater was characterized.
+Bacterial and archaeal epiphytes were sampled in a meadow of *C.
+nodosa* invaded by the invasive *C.
+cylindracea* and in a locality of only *C.
+cylindracea* located in the proximity of the meadow.
+In addition, for comparison, the community of the ambient seawater was characterized.
 
@@
   rrarefy(., min(rowSums(.))) %>%
-  as_tibble() %>%
+  as_tibble(.name_repair="unique") %>%
   add_column("Group"=shared$Group, .before=TRUE) %>%
@@
 estimators <- estimateR(rarefied)
-estimators <- as_tibble(t(estimators)) %>%
+estimators <- as_tibble(t(estimators), .name_repair="unique") %>%
   add_column("Group"=colnames(estimators), .before=TRUE)
@@
 Sequencing of the 16S rRNA V4 region yielded a total of `r million_sequences` million sequences after quality curation and exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences (\autoref{supp-nseq_notus}).
-A total of `r samples_epiphytes` samples originating from epiphytic archaeal and bacterial communities associated with surfaces of the seagrass *Cymodocea nodosa* and macroalga *Caulerpa cylindracea* were analyzed.
+A total of `r samples_epiphytes` samples originating from epiphytic archaeal and bacterial communities associated with surfaces of the seagrass *C.
+nodosa* and macroalga *C.
+cylindracea* were analyzed.
 In addition, `r samples_seawater` samples (one of the samples was sequenced two times) originating from picoplankton archaeal and bacterial communities in the ambient seawater were also processed for comparison.
@@
 Generally, richness estimators and diversity indices showed similar trends.
-On average, higher values were found for *Caulerpa cylindracea* (invaded [Number of OTUs, `r S.obs_fcam$mean` ± `r S.obs_fcam$sd` OTUs] and noninvaded [Number of OTUs, `r S.obs_fca$mean` ± `r S.obs_fca$sd` OTUs]), middle vales for *Cymodocea nodosa* (Number of OTUs, `r S.obs_fcym$mean` ± `r S.obs_fcym$sd` OTUs) and lower values for picoplankton communities in the ambient seawater (Number of OTUs, `r S.obs_f$mean` ± `r S.obs_f$sd` OTUs) (\autoref{supp-calculators}).
+On average, higher values were found for *C.
+cylindracea* (invaded [Number of OTUs, `r S.obs_fcam$mean` ± `r S.obs_fcam$sd` OTUs] and noninvaded [Number of OTUs, `r S.obs_fca$mean` ± `r S.obs_fca$sd` OTUs]), middle vales for *C.
+nodosa* (Number of OTUs, `r S.obs_fcym$mean` ± `r S.obs_fcym$sd` OTUs) and lower values for picoplankton communities in the ambient seawater (Number of OTUs, `r S.obs_f$mean` ± `r S.obs_f$sd` OTUs) (\autoref{supp-calculators}).
 Seasonal changes did not show such large dissimilarities.
-*Cymodocea nodosa* communities showed a slow increase towards the end of the study, while *Caulerpa cylindracea* (invaded and noninvaded) communities were characterized by slightly larger values in Spring and Summer in comparison to Autumn and Winter (\autoref{supp-calculators}).
+*C.
+nodosa* communities showed a slow increase towards the end of the study, while *C.
+cylindracea* (invaded and noninvaded) communities were characterized by slightly larger values in Spring and Summer in comparison to Autumn and Winter (\autoref{supp-calculators}).
 
@@
   group_by(station) %>%
-  select(starts_with("Otu")) %>%
-  summarise_all(sum)
+  summarise_at(vars(starts_with("Otu")), sum)
 
@@
 jaccard <- vegdist(rarefied_metadata, method="jaccard", binary=T)
-jaccard <- as_tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(jaccard))) %>%
-  rename(V1=X1, V2=X2, jaccard=as.numeric.jaccard.)
+jaccard <- as_tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(jaccard)),
+                     .name_repair= ~c("V1", "V2", "jaccard"))
 
 bray <- vegdist(rarefied_metadata, method="bray", binary=F)
-bray <- as_tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(bray))) %>%
-  rename(V1=X1, V2=X2, bray=as.numeric.bray.)
+bray <- as_tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(bray)),
+                  .name_repair= ~c("V1", "V2", "bray"))
 
@@
 # Finding the core OTUs for each environment
-core <- metadata_rarefied %>%
-  filter(station=="F") %>%
-  select_if(~ all(. != 0)) %>%
-  select(starts_with("Otu"))
-
-total <- metadata_rarefied %>%
-  filter(station=="F") %>%
-  select(starts_with("Otu"))
-
-core_otus_f <- length(core)/length(total) * 100
-core_seq_f <- sum(core)/sum(total) * 100
-
 core <- metadata_rarefied %>%
@@
   rrarefy(., min(rowSums(.))) %>%
-  as_tibble() %>%
+  as_tibble(.name_repair="unique") %>%
   select_if(list(~!is.numeric(.) || sum(.)!=0))
@@
 Coefficients were determined after normalization through rarefaction and binning of samples from a particular environment.
-The highest proportion of shared OTUs and community was found between invaded and noninvaded  *Caulerpa cylindracea* (Jaccard, `r format(filter(similarity, V1=="FCa", V2=="FCaM")$jaccard, digits=2, nsmall=2)`; Bray-Curtis, `r format(filter(similarity, V1=="FCa", V2=="FCaM")$bray, digits=2, nsmall=2)`), while lower shared values were calculated between seawater and epiphytic communities (\autoref{matrix}).
-Shared proportion between *Cymodocea nodosa* and *Caulerpa cylindracea* were approximately in the middle between these two extremes.
+The highest proportion of shared OTUs and community was found between invaded and noninvaded  *C.
+cylindracea* (Jaccard, `r format(filter(similarity, V1=="FCa", V2=="FCaM")$jaccard, digits=2, nsmall=2)`; Bray-Curtis, `r format(filter(similarity, V1=="FCa", V2=="FCaM")$bray, digits=2, nsmall=2)`), while lower shared values were calculated between seawater and epiphytic communities (\autoref{matrix}).
+Shared proportion between *C.
+nodosa* and *C.
+cylindracea* were approximately in the middle between these two extremes.
 To assess seasonal changes in the proportion of shared OTUs and communities the Jaccard's and Bray-Curtis Similarity Coefficients were calculated between consecutive sampling points (\autoref{shared}).
 Both coefficients showed similar trends.
-Temporal proportional changes were more pronounced for seawater in comparison to *Cymodocea nodosa* and especially *Caulerpa cylindracea* associated communities (\autoref{shared}).
+Temporal proportional changes were more pronounced for seawater in comparison to *C.
+nodosa* and especially *C.
+cylindracea* associated communities (\autoref{shared}).
 In addition, only `r min(core$core_otus)` -- `r max(core$core_otus)` \si{\percent} of OTUs from each surface associated community were found at every time point.
@@
 This separation was further supported by ANOSIM (R = `r format(anosim_all$statistic, digits=2, nsmall=2)`, *p* < `r anosim_all_signif`).
-Seasonal changes of *Cymodocea nodosa* associated communities indicated a separation between Spring, Summer and Autumn/Winter samples (ANOSIM, R = `r format(anosim_fcym$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fcym_signif`), while communities from the surfaces of *Caulerpa cylindracea* indicated a non so strongly supported, as in previous cases, separation between Summer and Autumn/Winter/Spring samples (ANOSIM, R = `r format(anosim_fca$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fca_signif`) (\autoref{pcoa}).
+Seasonal changes of *C.
+nodosa* associated communities indicated a separation between Spring, Summer and Autumn/Winter samples (ANOSIM, R = `r format(anosim_fcym$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fcym_signif`), while communities from the surfaces of *C.
+cylindracea* indicated a non so strongly supported, as in previous cases, separation between Summer and Autumn/Winter/Spring samples (ANOSIM, R = `r format(anosim_fca$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fca_signif`) (\autoref{pcoa}).
 
@@
 Seawater communities were mainly comprised of *Actinobacteriota*, *Bacteroidota*, *Cyanobacteria*, *Alphaproteobacteria*, *Gammaproteobacteria* and *Verrucomicrobiota*.
-Communities associated with *Cymodocea nodosa* were consisted of same groups with the addition of *Planctomycetota* whose contribution was higher in summer 2018.
-In addition, communities from invaded and noninvaded *Caulerpa cylindracea* were similar and characterized by same groups as seawater and *Cymodocea nodosa* communities with the addition of *Desulfobacterota* (\autoref{community}).
+Communities associated with *C.
+nodosa* were consisted of same groups with the addition of *Planctomycetota* whose contribution was higher in summer 2018.
+In addition, communities from invaded and noninvaded *C.
+cylindracea* were similar and characterized by same groups as seawater and *C.
+nodosa* communities with the addition of *Desulfobacterota* (\autoref{community}).
 Larger differences between environments and host species could be observed at lower taxonomic ranks (\autoref{cyano} -- \ref{desulfo}).
@@
 *Cyanobacteria* related sequences were comprising, on average, `r cyano_all_whole$mean` ± `r cyano_all_whole$sd` \si{\percent} of total sequences (\autoref{cyano}).
-Higher proportions were found for *Cymodocea nodosa* (`r filter(cyano_station_season_whole, station=="FCyM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCyM" & season=="Autumn")$sd` \si{\percent}) and *Caulerpa cylindracea* (invaded [ (`r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$sd` \si{\percent})] and noninvaded [ (`r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$sd` \si{\percent})]) associated communities in autumn and for seawater communities in winter (`r filter(cyano_station_season_whole, station=="F" & season=="Winter")$mean` ± `r filter(cyano_station_season_whole, station=="F" & season=="Winter")$sd` \si{\percent}).
+Higher proportions were found for *C.
+nodosa* (`r filter(cyano_station_season_whole, station=="FCyM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCyM" & season=="Autumn")$sd` \si{\percent}) and *C.
+cylindracea* (invaded [ (`r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$sd` \si{\percent})] and noninvaded [ (`r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$sd` \si{\percent})]) associated communities in autumn and for seawater communities in winter (`r filter(cyano_station_season_whole, station=="F" & season=="Winter")$mean` ± `r filter(cyano_station_season_whole, station=="F" & season=="Winter")$sd` \si{\percent}).
 Large taxonomic differences between surface associated and seawater cyanobacterial communities were observed.
@@
 Representatives of no relative *Rhodobacteraceae* were comprising on average `r rhodo$mean` ± `r rhodo$sd` \si{\percent} of all alphaproteobacterial sequences from the epiphytic community (\autoref{alpha}).
-In addition, *Amylibacter* was detected mainly in *Cymodocea nodosa* from November 2017 until March 2018.
+In addition, *Amylibacter* was detected mainly in *C.
+nodosa* from November 2017 until March 2018.
 
@@
 Seawater communities were mainly comprised of the OM60 (NOR5) clade, *Litoricola*, *Acinetobacter* and the SAR86 clade, while epiphytic communities were mainly composed of no relative *Gammaproteobacteria* and *Granulosicocus*.
-Beside these two groups specific to all three epiphytic communities, *Cymodocea nodosa* was characterized by *Arenicella*, no relative *Burkholderiales* and *Methylotenera*, while *Thioploca*, no relative *Cellvibrionaceae* and *Reinekea* were more specific to both invaded and noninvaded *Caulerpa cylindracea*.
+Beside these two groups specific to all three epiphytic communities, *C.
+nodosa* was characterized by *Arenicella*, no relative *Burkholderiales* and *Methylotenera*, while *Thioploca*, no relative *Cellvibrionaceae* and *Reinekea* were more specific to both invaded and noninvaded *C.
+cylindracea*.
 In addition, *Arenicella* was more pronounced in November and December 2017, while no relative *Burkholderiales* and *Methylotenera* were more characteristic for the period form March until May 2018.
-For the *Caulerpa cylindracea* specific taxa no relative *Cellvibrionaceae* and *Reinekea* showed some seasonality and were characterisitic for samples originating from June to October 2018.
+For the *C.
+cylindracea* specific taxa no relative *Cellvibrionaceae* and *Reinekea* showed some seasonality and were characterisitic for samples originating from June to October 2018.
 In addition, similarly to *Bacteroidota*, a large proportion of the surface associated community was grouped as other *Gammaproteobacteria* indicating high diversity within this group (\autoref{gamma}).
@@
 
-In contrast to previously described high rank taxa, *Desulfobacterota* were specific to *Caulerpa cylindracea*.
+In contrast to previously described high rank taxa, *Desulfobacterota* were specific to *C.
+cylindracea*.
 On average they were comprising `r desulfo_all_whole$mean` ± `r desulfo_all_whole$sd` \si{\percent} of all sequences.
-While seawater and *Cymodocea nodosa* communities were consisted of only `r filter(desulfo_station_whole, station=="F")$mean` ± `r filter(desulfo_station_whole, station=="F")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCyM")$mean` ± `r filter(desulfo_station_whole, station=="FCyM")$sd` \si{\percent} *Desulfobacterota* sequences, respectively, in the invaded and noninvaded *Caulerpa cylindracea* communities their proportion was `r filter(desulfo_station_whole, station=="FCaM")$mean` ± `r filter(desulfo_station_whole, station=="FCaM")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCa")$mean` ± `r filter(desulfo_station_whole, station=="FCa")$sd` \si{\percent}, respectively (\autoref{desulfo}).
+While seawater and *C.
+nodosa* communities were consisted of only `r filter(desulfo_station_whole, station=="F")$mean` ± `r filter(desulfo_station_whole, station=="F")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCyM")$mean` ± `r filter(desulfo_station_whole, station=="FCyM")$sd` \si{\percent} *Desulfobacterota* sequences, respectively, in the invaded and noninvaded *C.
+cylindracea* communities their proportion was `r filter(desulfo_station_whole, station=="FCaM")$mean` ± `r filter(desulfo_station_whole, station=="FCaM")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCa")$mean` ± `r filter(desulfo_station_whole, station=="FCa")$sd` \si{\percent}, respectively (\autoref{desulfo}).
 The community was mainly consisted of no relative *Desulfobacteraceae*, *Desulfatitalea*, no relative *Desulfobulbaceae*, *Desulfobulbus*, no relative *Desulfocapsaceae*, *Desulfopila*, *Desulforhopalus*, *Desulfotalea*, SEEP-SRB4 and uncultured *Desulfocapsaceae* (\autoref{desulfo}).
@@
 All these methods showed different removal efficiencies but none was enabling a complete removal of attached microbial cells.
-In the present study, we applied an earlier developed removal protocol [@Korlevic_in_press, submitted], based on a previous idea of direct cellular lysis [@Burke2009], to ensure an almost complete cell detachment.
-The application of a direct lysis procedure coupled with a high frequency sampling protocol and Illumina high resolution amplicon sequencing has enabled us to make a detailed description of bacterial and archaeal communities associated with the surfaces of two marine macrophytes, *Cymodocea nodosa* and *Caulerpa cylindracea*.
+In the present study, we applied an earlier developed removal protocol [@Korlevic_submitted, submitted], based on a previous idea of direct cellular lysis [@Burke2009], to ensure an almost complete cell detachment.
+The application of a direct lysis procedure coupled with a high frequency sampling protocol and Illumina high resolution amplicon sequencing has enabled us to make a detailed description of bacterial and archaeal communities associated with the surfaces of two marine macrophytes, *C.
+nodosa* and *C.
+cylindracea*.
 
-In the present study, highest richness values were observed for *Caulerpa cylindracea* (invaded and noninvaded), middle for *Cymodocea nodosa* and lowest for seawater derived communities.
+In the present study, highest richness values were observed for *C.
+cylindracea* (invaded and noninvaded), middle for *C.
+nodosa* and lowest for seawater derived communities.
 Higher values for seagrass associated communities in comparison to seawater were described  earlier and could be attributed to a larger set of inhabitable microniches existing on macrophyte surfaces [@Ugarelli2019].
-In addition, highest values observed for *Caulerpa cylindracea* are probably a consequence of part of epiphytic OTUs that are in contact with the sediment.
-*Caulerpa cylindracea* stolon is attached to the sediment surface with rhizoids, so the stolon and rhizoids are in a direct contact with the sediment.
+In addition, highest values observed for *C.
+cylindracea* are probably a consequence of part of epiphytic OTUs that are in contact with the sediment.
+*C.
+cylindracea* stolon is attached to the sediment surface with rhizoids, so the stolon and rhizoids are in a direct contact with the sediment.
 Part of the surface attached *Caulerpa cylnidracea* community is therefore comprised of OTUs that are in contact with a different environment and that could cause the observed increase in richness.
@@
 Similarly to previously described differences between seawater and surface attached communities, at high taxonomic ranks no strong differentiation between communities associated with different host was observed.
-The only high rank phylum that was differing between *Cymodocea nodosa* ans *Caulerpa cylindracea* was *Desulfobacterota*, whose sequences were more abundant in the *Caulerpa cylindracea* associated community.
+The only high rank phylum that was differing between *C.
+nodosa* ans *C.
+cylindracea* was *Desulfobacterota*, whose sequences were more abundant in the *C.
+cylindracea* associated community.
 As already mentioned, the rhizoids and part of the stolon are in contact with the sediment, so *Desulfobacterota* are probably a part of the epiphytic community that is in contact with the sediment.
@@
 It seems, that microniches on the surfaces of macrophytes are providing more stable conditions in comparison to the seawater.
-At the level of OTUs seasonal changes of *Cymodocea nodosa* and *Caulerpa cylindracea* associated communities were identified that could be linked to the growth cycle of the seagrass and macroalgae (M.
+At the level of OTUs seasonal changes of *C.
+nodosa* and *C.
+cylindracea* associated communities were identified that could be linked to the growth cycle of the seagrass and macroalgae (M.
 Najdek, personal communication).
-*Cymodocea nodosa* was characterized by a Spring community during maximum seagrass proliferation, a Summer community during a biomass maximum and a Autumn/Winter community during a biomass decay.
-In contrast, *Caulerpa cylindracea* started to proliferate in late Spring and was characterized only by a Summer community during maximal biomass increase and by a Autumn/Winter/Spring community when the biomass were at the peak and the settlement started to subsequently decay.
+*C.
+nodosa* was characterized by a Spring community during maximum seagrass proliferation, a Summer community during a biomass maximum and a Autumn/Winter community during a biomass decay.
+In contrast, *C.
+cylindracea* started to proliferate in late Spring and was characterized only by a Summer community during maximal biomass increase and by a Autumn/Winter/Spring community when the biomass were at the peak and the settlement started to subsequently decay.
 Similar seasonal changes in the epiphytic community was described also for other macroalgae [@Lachnit2011; @Tujula2010].
-Higher temporal stability of *Caulerpa cylindracea* surface communities in comparison to *Cymodocea nodosa* were also observed in the higher proportion of shared communities between two consecutive sampling points.
+Higher temporal stability of *C.
+cylindracea* surface communities in comparison to *C.
+nodosa* were also observed in the higher proportion of shared communities between two consecutive sampling points.
 
-Seasonal patterns of high rank epiphytic taxa did not show strong changes.
-Only *Cyanobateria* were characterized by strong seasonal differences.
-Cyanobacterial sequences were more pronounced in the November and December 2017 in comparison to Spring and Summer.
+Analysis of seasonal chloroplast sequence abundances showed higher values in Autumn/Winter in comparison to Spring/Summer.
+This pattern is not surprising as e.g seagrasses are known to harbor more epiphytes during Autumn/Winter [@Reyes2001] and we used an adapted DNA isolation protocol that is known to partially coextract DNA from planktonic eukaryotes [@Korlevic2015].
+Strong seasonal fluctuations of high rank epiphytic taxa were not observed, with the exception of *Cyanobateria*.
+Cyanobacterial sequences were more pronounced in November and December in comparison to Spring and Summer.
 Interestingly, in these high proportion months the majority of cyanobacterial sequences were classified as *Pleurocapsa*, a group known to colonized different living and nonliving surfaces [@Longford2007; @Burns2004; @Mobberley2012; @Reisser2014].
-It is possible than during periods of low metabolic activity of the i.g.
-seagrass there is no so active interaction and selection of the epiphytic community, so the seagrass leaves are becoming a suitable surface for nonspecific colonizers [@Zavodnik1998].
-*Pleurocapsa* was replaced in Spring ans Summer by *Acrophormium*, *Phormidesmis* and no relative *Nodosilineaceae*.
+It is possible than during periods of low metabolic activity there is no so active interaction and selection of the epiphytic community by the seagrass, causing leaves to become a suitable surface for nonspecific colonizers [@Zavodnik1998].
+*Pleurocapsa* was replaced in Spring and Summer by *Acrophormium*, *Phormidesmis* and no relative *Nodosilineaceae*.
 A study of coastal microbial mats found also higher proportion of *Nodosilineaceae* sequences in Summer, while *Phormidesmis* sequences were at their peak in Autumn [@Cardoso2019].
-Other high rank taxa did not showed strong successional patterns.
-In each analyzed group low rank taxa
+Other high rank taxa did not showed strong successional patterns, but in every analyzed group, with the exception of *Desulfobacterota*, taxa present throughout the year in similar proportions and season specific taxa could be identified.
+Within *Bacteroidota* different groups withing the family *Saprospiraceae* (i.g.
+*Lewinella*, *Portibacter* and *Rubidimonas*) were detected through the year.
+Members of this family are often found in association with macrophytes and it is suggested that are involved in the hydrolysis and utilization of complex carbon sources [@Burke2011b; @Crump2018; @McIlroy2014].
+On the other hand, families *Flavobacteriaceae* and *Rhodothermaceae* showed seasonal patterns, with *Flavobacteriaceae* being more pronounced from November to June and *Rhodothermaceae* from June to October.
+Within *Alphaproteobacteria* the family *Rhodobacteraceae* was comprising the majority of sequences thought the year.
+This metabolically versatile family is often associated with macrophyte surfaces and usually is one of the most abundant groups [@Burke2011b; @Michelou2013; @Pujalte2014].
+In addition, *Hyphomonadaceae* were found in all samples.
+Interestingly, some of the species within this group contain stalks on their cells which can be used to attach to the macrophyte surface [@Weidner2000; @Abraham2014].
+Within the *Gammaproteobacteria*, sequences without known representatives were the most pronounced group present throughout the year.
+In addition, *Granulosicoccus* was also found in almost all samples.
+*Gammaproteobacteria* are often a major constituent of macrophyte epiphytic communities [@Michelou2013; @Burke2011b; @Crump2018].
+Beside these two groups other less pronounced taxa showed seasonal and host-specific patterns.
+In example, *C.
+cylindracea* was characterized by *Thioploca*, a known sulfur sediment bacteria and *Cellvibrionaceae*, a family whose cultured members are known polysaccharides degraders [@Jorgensen1999; @Xie2017].
+*Desulfobacterota* were found only associated with *C.
+cylindracea* and no group within this phylum showed seasonal patterns.
+The presence of this phylum only on *C.
+cylindracea* is to be expected as part of the epiphytic community is in a direct contact with the sediment.
+The *Desulfobacterota* community was dominated by *Desulfatitalea* and no relative *Desulfocapsaceae*, known sulfate sediment groups [@Higashioka2015; @Kuever2014].
+
+In temperate zones marine macrophytes are exhibiting growth cycles, so it  is not surprising that the associated epiphytic microbial community is undergoing partial seasonal changes.
+In the present study, we could, in every analyzed high rank taxa, identify phylogenetic groups that were present throughout the year and that were comprising most of the sequences and lower proportion taxa showing seasonal patterns connected to the macrophyte growth cycle.
+Studies focusing on functional comparisons between communities associated with different hosts showed that the majortiy of functions could be found in every community, indicating functional redundacy [@Roth-Schulze2016].
+This difference between taxonomic and functionl discrepancy was expalined by the lottery hypothesis that hypothesize an initial random colonization step performed from a set of functionally equivalent taxonomic groups [@Burke2011a; @Roth-Schulze2016].
+It is possible that functional redundancy is a characteristic of high abundance taxa detected to be present throughout the year, while seasonal and/or host-specific functions are an attribute of taxa displaying successional patterns.
+Further studies connecting taxonomy with functional properties will be required to elucidate the degree of functional redundancy or specificity in epiphyitc microbial communites.
 
@@
 ### Sampling
-Leaves of *Cymodocea nodosa* were sampled in a *Cymodocea nodosa* meadow located in the proximity of the village of Funtana (\ang{45;10;39} N, \ang{13;35;42} E).
-Thalli of *Caulerpa cylindracea* were sampled in the same *Cymodocea nodosa* invaded meadow in Funtana and on a locality of only *Caulerpa cylindracea* located close to the invaded meadow.
+Leaves of *C.
+nodosa* were sampled in a *C.
+nodosa* meadow located in the proximity of the village of Funtana (\ang{45;10;39} N, \ang{13;35;42} E).
+Thalli of *C.
+cylindracea* were sampled in the same *C.
+nodosa* invaded meadow in Funtana and on a locality of only *C.
+cylindracea* located close to the invaded meadow.
 Sampling of leaves and thalli was performed approximately monthly from December 2017 to October 2018 (\autoref{supp-nseq_notus}).
 Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with site seawater.
-Upon arrival to the laboratory, *Cymodocea nodosa* leaves were cut into sections of 1 -- 2 \si{\cm}, while *Caulerpa cylindracea* thalli were cut into 5 -- 8 \si{\cm} long sections.
+Upon arrival to the laboratory, *C.
+nodosa* leaves were cut into sections of 1 -- 2 \si{\cm}, while *C.
+cylindracea* thalli were cut into 5 -- 8 \si{\cm} long sections.
 Leaves and thalli were washed three times with sterile artificial seawater (ASW) to remove loosely attached microbial cells.
-Surrounding seawater was collected in 10 \si{\l} containers by diving and transported to the laboratory where the whole container volume was filtered through a 20 \si{\um} net.
+Ambient seawater was collected in 10 \si{\l} containers by diving and transported to the laboratory where the whole container volume was filtered through a 20 \si{\um} net.
 The filtrate was further sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate membrane filters (Whatman, United Kingdom) using a peristaltic pump.
@@
 ### DNA Isolation
-DNA from surfaces of *Cymodocea nodosa* and *Caulerpa cylindracea* was isolated using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevic_in_press, submitted; @Massana1997].
+DNA from surfaces of *C.
+nodosa* and *C.
+cylindracea* was isolated using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevic_submitted, submitted; @Massana1997].
 Briefly, leaves and thalli are incubated in a lysis buffer and treated with lysozyme and proteinase K.
@@
 ### Illumina 16S rRNA Sequencing
-Illumina MiSeq sequencing of the V4 16S rRNA region was performed as described previously [@Korlevic_in_press, submitted].
+Illumina MiSeq sequencing of the V4 16S rRNA region was performed as described previously [@Korlevic_submitted, submitted].
 The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
@@
 Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019],  packages  vegan (version 2.5-6) [@Oksanen2019], and tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014].
-The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (**TO BE ADDED LATER!**).
-Based on the ATCC MSA-1002 mock community included in the analysis an average sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
+The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/mkorlevic/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020).
+Based on the ATCC MSA-1002 mock community included in the analysis an average sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016a].
 In addition, the negative controls processed together with the samples yielded on average only `r nc` sequences after sequence quality curation.
@@
 ## Acknowledgments
+This work was founded by the Croatian Science Foundation through the	MICRO-SEAGRASS project (IP-2016-06-7118).
+We would like to thank Margareta Buterer for technical support, Paolo Paliaga for help during sampling and the University Computing Center of the University of Zagreb for access to the cluster Isabella.
 
@@
 ## Figures
-```{r out.width="70%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between communties associated with the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and coomunities in the surrounding seawater.\\label{matrix}", fig.pos="H"}
+```{r out.width="70%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between communties associated with the surfaces of macrophytes (\\textit{C.
+nodosa} [Invaded] and \\textit{C.
+cylindracea} [Invaded and Noninvaded]) and coomunities in the ambient seawater.\\label{matrix}", fig.pos="H"}
 include_graphics("../results/figures/matrix.jpg")
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Proprotion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\\label{shared}", fig.pos="H"}
+```{r out.width="85%", fig.align="center", fig.cap="Proprotion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\\textit{C.
+nodosa} [Invaded] and \\textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\\label{shared}", fig.pos="H"}
 include_graphics("../results/figures/seasonal_shared.jpg")
@@
 
-```{r out.width="55%", fig.align="center", fig.cap="Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+```{r out.width="55%", fig.align="center", fig.cap="Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{C.
+nodosa} [Invaded] and \\textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
 Samples from the same environment or same season are labeld in different colors.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant bacterial and archaeal sequences on the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\\label{community}", fig.pos="H"}
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) bacterial and archaeal sequences on the surfaces of macrophytes (\\textit{C.
+nodosa} [Invaded] and \\textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\\label{community}", fig.pos="H"}
 include_graphics("../results/figures/community_bar_plot.jpg")
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant cyanobacterial sequences on the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) cyanobacterial sequences on the surfaces of macrophytes (\\textit{C.
+nodosa} [Invaded] and \\textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
 The proprotion of cyanobaterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant sequences within the \\textit{Bacteroidota} on the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) sequences within the \\textit{Bacteroidota} on the surfaces of macrophytes (\\textit{C.
+nodosa} [Invaded] and \\textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
 The proprotion of sequences classified as \\textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant alphaproteobacterial sequences on the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) alphaproteobacterial sequences on the surfaces of macrophytes (\\textit{C.
+nodosa} [Invaded] and \\textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
 The proprotion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant gammaproteobacterial sequences on the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) gammaproteobacterial sequences on the surfaces of macrophytes (\\textit{C.
+nodosa} [Invaded] and \\textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
 The proprotion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant sequences within the \\textit{Desulfobacterota} on the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) sequences within the \\textit{Desulfobacterota} on the surfaces of macrophytes (\\textit{C.
+nodosa} [Invaded] and \\textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
 The proprotion of sequences classified as \\textit{Desulfobacterota} in the total bacterial and archaeal community is given above the corresponding bar.

### 2020-03-12 submission/manuscript.tex
Manuscript 1st version complete.

--- 2020-03-09 submission/manuscript.tex
@@
 
-\vspace{70mm}
-
-\textsuperscript{1\(\dagger\)}
+\vspace{50mm}
+
+Marino Korlević\textsuperscript{1\(*\)}, Marsej
+Markovski\textsuperscript{1}, Zihao Zhao\textsuperscript{2}, Gerhard J.
+Herndl\textsuperscript{2,3}, Mirjana Najdek\textsuperscript{1}
 
@@
 
-\(\dagger\) To whom correspondence should be addressed:
+\textsuperscript{\(*\)}To whom correspondence should be addressed:
 \href{mailto:marino.korlevic@irb.hr}{\nolinkurl{marino.korlevic@irb.hr}}
 
-1. Ruđer Bošković Institute, Center for Marine Research, G. Paliaga 5,
-Rovinj, Croatia
-
-2. University of Vienna, Department of Limnology and Bio-Oceanography,
-Althanstraße 14, Vienna, Austria \newpage \linenumbers
-\sisetup{mode=text} \setlength\parindent{24pt}
+1. Center for Marine Research, Ruđer Bošković Institute, Croatia
+
+2. Department of Functional and Evolutionary Ecology, University of
+Vienna, Austria
+
+3. Department of Marine Microbiology and Biogeochemistry, Royal
+Netherlands Institute for Sea Research, Utrecht University, The
+Netherlands
+
+\newpage
+\linenumbers
+\sisetup{mode=text}
+\setlength\parindent{24pt}
 
@@
 \subsection{Abstract}\label{abstract}}
+
+Surfaces of marine macrophytes (seagrasses and macroalgae) are inhabited
+by diverse microbial communities. Most studies focusing on macrophyte
+epiphytic communities did not take into account temporal changes or
+applied low sampling frequency approaches. Illumina sequencing of the V4
+16S rRNA region was performed to determine the seasonal dynamics of
+epiphytic communities sampled from the surfaces of the seagrass
+\emph{Cymodocea nodosa} and invasive macroalga \emph{Caulerpa
+cylindracea}. Leaves and thalli were sampled in a meadow of
+\emph{Cymodocea nodosa} invaded by the invasive \emph{Caulerpa
+cylindracea} and in a monospecific settlement of \emph{Caulerpa
+cylindracea} located in the proximity of the meadow at monthly
+intervals. For comparison the ambient prokaryotic plankton community was
+also characterized. Sequencing results at the OTU level showed a clear
+differentiation between ambient water and epiphytic communities and a
+host-specific community assemblage. In addition, successional changes
+were observed that could be connected to the macrophyte growth cycle.
+Taxonomic analysis showed similar high rank groups in the ambient water
+and epiphytic communities, with the exception of \emph{Desulfobacterota}
+that were found only on \emph{Caulerpa cylindracea}. Only
+\emph{Cyanobacteria} showed seasonal change, while other high rank taxa
+were present throughout the year. In every analyzed high rank taxa,
+phylogenetic groups present throughout the year and comprising most of
+the sequences could be identified together with low proportion taxa
+showing seasonal patterns connected to the macrophyte growth cycle.
+Taken together, epiphytic microbial communities of the seagrass
+\emph{Cymodocea nodosa} and the macroalgae \emph{Caulerpa cylindracea}
+appear to be host-specific and contain taxa that undergo successional
+changes.
 
@@
 \emph{Caulerpa cylindracea} determined on a mostly monthly scale.
-Bacterial and archaeal epiphytes were sampled in a meadow of
-\emph{Cymodocea nodosa} invaded by the invasive \emph{Caulerpa
-cylindracea} and in a locality of only \emph{Caulerpa cylindracea}
-located in the proximity of the meadow. In addition, for comparison, the
-community of the surrounding seawater was characterized.
+Bacterial and archaeal epiphytes were sampled in a meadow of \emph{C.
+nodosa} invaded by the invasive \emph{C. cylindracea} and in a locality
+of only \emph{C. cylindracea} located in the proximity of the meadow. In
+addition, for comparison, the community of the ambient seawater was
+characterized.
 
@@
 epiphytic archaeal and bacterial communities associated with surfaces of
-the seagrass \emph{Cymodocea nodosa} and macroalga \emph{Caulerpa
-cylindracea} were analyzed. In addition, 18 samples (one of the samples
-was sequenced two times) originating from picoplankton archaeal and
-bacterial communities in the ambient seawater were also processed for
-comparison. The number of reads per sample ranged between 8,407 and
-77,465 sequences (\autoref{supp-nseq_notus}). Even when the highest
-sequencing effort was applied the rarefaction curves did not level off
-that is a common observation in high-throughput 16S rRNA amplicon
-sequencing approaches (\autoref{supp-rarefaction}). Following quality
-curation and exclusion of sequences mentioned before reads were
-clustered into 28,726 different OTUs at a similarity level of 97
-\si{\percent}. Reads numbers were normalized to the minimum number of
-sequences, 8,407 (\autoref{supp-nseq_notus}), through rarefaction
-resulting in 17,007 different OTUs that ranged from 366 to 1,998 OTUs
-per sample (\autoref{supp-calculators}). To determine seasonal changes
-of richness and diversity the Observed Number of OTUs, Chao1, ACE,
-Exponential Shannon (Jost, 2006) and Inverse Simpson were calculated
-after normalization through rarefaction. Generally, richness estimators
-and diversity indices showed similar trends. On average, higher values
-were found for \emph{Caulerpa cylindracea} (invaded {[}Number of OTUs,
-1,688.5 ± 125.6 OTUs{]} and noninvaded {[}Number of OTUs, 1,744.8 ±
-150.8 OTUs{]}), middle vales for \emph{Cymodocea nodosa} (Number of
-OTUs, 1,062.8 ± 209.6 OTUs) and lower values for picoplankton
-communities in the ambient seawater (Number of OTUs, 528.1 ± 135.8 OTUs)
+the seagrass \emph{C. nodosa} and macroalga \emph{C. cylindracea} were
+analyzed. In addition, 18 samples (one of the samples was sequenced two
+times) originating from picoplankton archaeal and bacterial communities
+in the ambient seawater were also processed for comparison. The number
+of reads per sample ranged between 8,410 and 77,465 sequences
+(\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
+applied the rarefaction curves did not level off that is a common
+observation in high-throughput 16S rRNA amplicon sequencing approaches
+(\autoref{supp-rarefaction}). Following quality curation and exclusion
+of sequences mentioned before reads were clustered into 28,729 different
+OTUs at a similarity level of 97 \si{\percent}. Reads numbers were
+normalized to the minimum number of sequences, 8,410
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,019
+different OTUs that ranged from 355 to 2,018 OTUs per sample
+(\autoref{supp-calculators}). To determine seasonal changes of richness
+and diversity the Observed Number of OTUs, Chao1, ACE, Exponential
+Shannon (Jost, 2006) and Inverse Simpson were calculated after
+normalization through rarefaction. Generally, richness estimators and
+diversity indices showed similar trends. On average, higher values were
+found for \emph{C. cylindracea} (invaded {[}Number of OTUs, 1,690.5 ±
+152.3 OTUs{]} and noninvaded {[}Number of OTUs, 1,733.8 ± 163.5
+OTUs{]}), middle vales for \emph{C. nodosa} (Number of OTUs, 1,066.9 ±
+224.5 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 526 ± 145.6 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
-dissimilarities. \emph{Cymodocea nodosa} communities showed a slow
-increase towards the end of the study, while \emph{Caulerpa cylindracea}
-(invaded and noninvaded) communities were characterized by slightly
-larger values in Spring and Summer in comparison to Autumn and Winter
+dissimilarities. \emph{C. nodosa} communities showed a slow increase
+towards the end of the study, while \emph{C. cylindracea} (invaded and
+noninvaded) communities were characterized by slightly larger values in
+Spring and Summer in comparison to Autumn and Winter
 (\autoref{supp-calculators}).
@@
 particular environment. The highest proportion of shared OTUs and
-community was found between invaded and noninvaded \emph{Caulerpa
-cylindracea} (Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared
-values were calculated between seawater and epiphytic communities
-(\autoref{matrix}). Shared proportion between \emph{Cymodocea nodosa}
-and \emph{Caulerpa cylindracea} were approximately in the middle between
-these two extremes. To assess seasonal changes in the proportion of
-shared OTUs and communities the Jaccard's and Bray-Curtis Similarity
-Coefficients were calculated between consecutive sampling points
-(\autoref{shared}). Both coefficients showed similar trends. Temporal
-proportional changes were more pronounced for seawater in comparison to
-\emph{Cymodocea nodosa} and especially \emph{Caulerpa cylindracea}
-associated communities (\autoref{shared}). In addition, only 0.4 -- 1.0
-\si{\percent} of OTUs from each surface associated community were found
-at every time point. These OTUs made also a high proportion of total
-sequences (39.1 -- 51.3 \si{\percent}). To further disentangle the
-environmental and seasonal community dissimilarity a Principal
-Coordinates Analysis (PCoA) based on Bray-Curtis distances and OTU
-abundances was applied. It showed a clear separation between planktonic
-and surface associated communities (\autoref{pcoa}). In addition, a
-separation of epiphytic bacterial and archaeal communities based on host
-species was determined. This separation was further supported by ANOSIM
-(R = 0.96, \emph{p} \textless{} 0.001). Seasonal changes of
-\emph{Cymodocea nodosa} associated communities indicated a separation
-between Spring, Summer and Autumn/Winter samples (ANOSIM, R = 0.56,
-\emph{p} \textless{} 0.01), while communities from the surfaces of
-\emph{Caulerpa cylindracea} indicated a non so strongly supported, as in
-previous cases, separation between Summer and Autumn/Winter/Spring
-samples (ANOSIM, R = 0.31, \emph{p} \textless{} 0.05) (\autoref{pcoa}).
+community was found between invaded and noninvaded \emph{C. cylindracea}
+(Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared values were
+calculated between seawater and epiphytic communities
+(\autoref{matrix}). Shared proportion between \emph{C. nodosa} and
+\emph{C. cylindracea} were approximately in the middle between these two
+extremes. To assess seasonal changes in the proportion of shared OTUs
+and communities the Jaccard's and Bray-Curtis Similarity Coefficients
+were calculated between consecutive sampling points (\autoref{shared}).
+Both coefficients showed similar trends. Temporal proportional changes
+were more pronounced for seawater in comparison to \emph{C. nodosa} and
+especially \emph{C. cylindracea} associated communities
+(\autoref{shared}). In addition, only 0.4 -- 1.0 \si{\percent} of OTUs
+from each surface associated community were found at every time point.
+These OTUs made also a high proportion of total sequences (39.9 -- 52.8
+\si{\percent}). To further disentangle the environmental and seasonal
+community dissimilarity a Principal Coordinates Analysis (PCoA) based on
+Bray-Curtis distances and OTU abundances was applied. It showed a clear
+separation between planktonic and surface associated communities
+(\autoref{pcoa}). In addition, a separation of epiphytic bacterial and
+archaeal communities based on host species was determined. This
+separation was further supported by ANOSIM (R = 0.96, \emph{p}
+\textless{} 0.001). Seasonal changes of \emph{C. nodosa} associated
+communities indicated a separation between Spring, Summer and
+Autumn/Winter samples (ANOSIM, R = 0.54, \emph{p} \textless{} 0.01),
+while communities from the surfaces of \emph{C. cylindracea} indicated a
+non so strongly supported, as in previous cases, separation between
+Summer and Autumn/Winter/Spring samples (ANOSIM, R = 0.31, \emph{p}
+\textless{} 0.05) (\autoref{pcoa}).
 
@@
 \emph{Gammaproteobacteria} and \emph{Verrucomicrobiota}. Communities
-associated with \emph{Cymodocea nodosa} were consisted of same groups
-with the addition of \emph{Planctomycetota} whose contribution was
-higher in summer 2018. In addition, communities from invaded and
-noninvaded \emph{Caulerpa cylindracea} were similar and characterized by
-same groups as seawater and \emph{Cymodocea nodosa} communities with the
-addition of \emph{Desulfobacterota} (\autoref{community}). Larger
-differences between environments and host species could be observed at
-lower taxonomic ranks (\autoref{cyano} -- \ref{desulfo}).
+associated with \emph{C. nodosa} were consisted of same groups with the
+addition of \emph{Planctomycetota} whose contribution was higher in
+summer 2018. In addition, communities from invaded and noninvaded
+\emph{C. cylindracea} were similar and characterized by same groups as
+seawater and \emph{C. nodosa} communities with the addition of
+\emph{Desulfobacterota} (\autoref{community}). Larger differences
+between environments and host species could be observed at lower
+taxonomic ranks (\autoref{cyano} -- \ref{desulfo}).
 
@@
 ± 4.4 \si{\percent} of total sequences (\autoref{cyano}). Higher
-proportions were found for \emph{Cymodocea nodosa} (16.4 ± 5.3
-\si{\percent}) and \emph{Caulerpa cylindracea} (invaded {[} (7.7 ± 3.9
-\si{\percent}){]} and noninvaded {[} (7.8 ± 2.4 \si{\percent}){]})
-associated communities in autumn and for seawater communities in winter
-(8.8 ± 7.4 \si{\percent}). Large taxonomic differences between surface
-associated and seawater cyanobacterial communities were observed.
-Seawater communities were mainly comprised of \emph{Cyanobium} and
-\emph{Synechococcus}, while surface associated communities were
-consisted of \emph{Pleurocapsa} and sequences without known relatives
-within \emph{Cyanobacteriia} (\autoref{cyano}). In addition, seasonal
-changes in surface associated communities were observed with
-\emph{Pleurocapsa} and no relative \emph{Cyanobacteriia} comprising
-larger proportions in autumn and winter and \emph{Acrophormium},
-\emph{Phormidesmis} and no relative \emph{Nodosilineaceae} in spring and
-summer (\autoref{cyano}).
+proportions were found for \emph{C. nodosa} (16.4 ± 5.3 \si{\percent})
+and \emph{C. cylindracea} (invaded {[} (7.7 ± 3.9 \si{\percent}){]} and
+noninvaded {[} (7.8 ± 2.4 \si{\percent}){]}) associated communities in
+autumn and for seawater communities in winter (8.8 ± 7.4 \si{\percent}).
+Large taxonomic differences between surface associated and seawater
+cyanobacterial communities were observed. Seawater communities were
+mainly comprised of \emph{Cyanobium} and \emph{Synechococcus}, while
+surface associated communities were consisted of \emph{Pleurocapsa} and
+sequences without known relatives within \emph{Cyanobacteriia}
+(\autoref{cyano}). In addition, seasonal changes in surface associated
+communities were observed with \emph{Pleurocapsa} and no relative
+\emph{Cyanobacteriia} comprising larger proportions in autumn and winter
+and \emph{Acrophormium}, \emph{Phormidesmis} and no relative
+\emph{Nodosilineaceae} in spring and summer (\autoref{cyano}).
 
@@
 community (\autoref{alpha}). In addition, \emph{Amylibacter} was
-detected mainly in \emph{Cymodocea nodosa} from November 2017 until
-March 2018.
+detected mainly in \emph{C. nodosa} from November 2017 until March 2018.
 
@@
 \emph{Granulosicocus}. Beside these two groups specific to all three
-epiphytic communities, \emph{Cymodocea nodosa} was characterized by
+epiphytic communities, \emph{C. nodosa} was characterized by
 \emph{Arenicella}, no relative \emph{Burkholderiales} and
@@
 \emph{Cellvibrionaceae} and \emph{Reinekea} were more specific to both
-invaded and noninvaded \emph{Caulerpa cylindracea}. In addition,
+invaded and noninvaded \emph{C. cylindracea}. In addition,
 \emph{Arenicella} was more pronounced in November and December 2017,
@@
 more characteristic for the period form March until May 2018. For the
-\emph{Caulerpa cylindracea} specific taxa no relative
-\emph{Cellvibrionaceae} and \emph{Reinekea} showed some seasonality and
-were characterisitic for samples originating from June to October 2018.
-In addition, similarly to \emph{Bacteroidota}, a large proportion of the
-surface associated community was grouped as other
-\emph{Gammaproteobacteria} indicating high diversity within this group
-(\autoref{gamma}).
+\emph{C. cylindracea} specific taxa no relative \emph{Cellvibrionaceae}
+and \emph{Reinekea} showed some seasonality and were characterisitic for
+samples originating from June to October 2018. In addition, similarly to
+\emph{Bacteroidota}, a large proportion of the surface associated
+community was grouped as other \emph{Gammaproteobacteria} indicating
+high diversity within this group (\autoref{gamma}).
 
 In contrast to previously described high rank taxa,
-\emph{Desulfobacterota} were specific to \emph{Caulerpa cylindracea}. On
+\emph{Desulfobacterota} were specific to \emph{C. cylindracea}. On
 average they were comprising 11.2 ± 13.3 \si{\percent} of all sequences.
-While seawater and \emph{Cymodocea nodosa} communities were consisted of
-only 0.1 ± 0.08 \si{\percent} and 1.0 ± 0.7 \si{\percent}
+While seawater and \emph{C. nodosa} communities were consisted of only
+0.1 ± 0.08 \si{\percent} and 1.0 ± 0.7 \si{\percent}
 \emph{Desulfobacterota} sequences, respectively, in the invaded and
-noninvaded \emph{Caulerpa cylindracea} communities their proportion was
-25.7 ± 11.3 \si{\percent} and 24.0 ± 4.3 \si{\percent}, respectively
+noninvaded \emph{C. cylindracea} communities their proportion was 25.7 ±
+11.2 \si{\percent} and 24.0 ± 4.3 \si{\percent}, respectively
 (\autoref{desulfo}). The community was mainly consisted of no relative
@@
 and archaeal communities associated with the surfaces of two marine
-macrophytes, \emph{Cymodocea nodosa} and \emph{Caulerpa cylindracea}.
-
-In the present study, highest richness values were observed for
-\emph{Caulerpa cylindracea} (invaded and noninvaded), middle for
-\emph{Cymodocea nodosa} and lowest for seawater derived communities.
-Higher values for seagrass associated communities in comparison to
-seawater were described earlier and could be attributed to a larger set
-of inhabitable microniches existing on macrophyte surfaces (Ugarelli
-\emph{et al.}, 2019). In addition, highest values observed for
-\emph{Caulerpa cylindracea} are probably a consequence of part of
-epiphytic OTUs that are in contact with the sediment. \emph{Caulerpa
-cylindracea} stolon is attached to the sediment surface with rhizoids,
-so the stolon and rhizoids are in a direct contact with the sediment.
-Part of the surface attached \emph{Caulerpa cylnidracea} community is
-therefore comprised of OTUs that are in contact with a different
-environment and that could cause the observed increase in richness. In
-addition, seasonal richness differences observed for surface attached
-communities showed slightly higher values in spring and summer. This
-pattern could be explained by a higher macrophyte growth in these
-seasons (M. Najdek, personal communication; Zavodnik \emph{et al.},
-1998; Ruitton \emph{et al.}, 2005). During active periods macrophytes
-exhibit a more dynamic chemical interaction with the surface community
-probably causing an increase in the number of inhabitable microniches
-(Borges and Champenois, 2015; Rickert \emph{et al.}, 2016).
+macrophytes, \emph{C. nodosa} and \emph{C. cylindracea}.
+
+In the present study, highest richness values were observed for \emph{C.
+cylindracea} (invaded and noninvaded), middle for \emph{C. nodosa} and
+lowest for seawater derived communities. Higher values for seagrass
+associated communities in comparison to seawater were described earlier
+and could be attributed to a larger set of inhabitable microniches
+existing on macrophyte surfaces (Ugarelli \emph{et al.}, 2019). In
+addition, highest values observed for \emph{C. cylindracea} are probably
+a consequence of part of epiphytic OTUs that are in contact with the
+sediment. \emph{C. cylindracea} stolon is attached to the sediment
+surface with rhizoids, so the stolon and rhizoids are in a direct
+contact with the sediment. Part of the surface attached \emph{Caulerpa
+cylnidracea} community is therefore comprised of OTUs that are in
+contact with a different environment and that could cause the observed
+increase in richness. In addition, seasonal richness differences
+observed for surface attached communities showed slightly higher values
+in spring and summer. This pattern could be explained by a higher
+macrophyte growth in these seasons (M. Najdek, personal communication;
+Zavodnik \emph{et al.}, 1998; Ruitton \emph{et al.}, 2005). During
+active periods macrophytes exhibit a more dynamic chemical interaction
+with the surface community probably causing an increase in the number of
+inhabitable microniches (Borges and Champenois, 2015; Rickert \emph{et
+al.}, 2016).
 
@@
 associated with different host was observed. The only high rank phylum
-that was differing between \emph{Cymodocea nodosa} ans \emph{Caulerpa
-cylindracea} was \emph{Desulfobacterota}, whose sequences were more
-abundant in the \emph{Caulerpa cylindracea} associated community. As
-already mentioned, the rhizoids and part of the stolon are in contact
-with the sediment, so \emph{Desulfobacterota} are probably a part of the
-epiphytic community that is in contact with the sediment. Similar high
-rank taxa found in this study were described to be specific for other
-species of macrophytes (Burke \emph{et al.}, 2011b; Lachnit \emph{et
-al.}, 2011; Bengtsson \emph{et al.}, 2017). In contrast to high
-taxonomic ranks, a substantial differentiation between host specific
-communities was found, which supports the host-specific hypothesis.
-Similar host-specificity was observed for different species of
-macroalgae and seagrasses (Lachnit \emph{et al.}, 2011; Roth-Schulze
-\emph{et al.}, 2016; Ugarelli \emph{et al.}, 2019; Morrissey \emph{et
-al.}, 2019). Taken together, at high taxonomic ranks a core set of taxa
-could be described that is characteristic for all or many macrophytes,
-while at low taxonomic ranks a community specific to host species could
-be identified (Egan \emph{et al.}, 2013).
+that was differing between \emph{C. nodosa} ans \emph{C. cylindracea}
+was \emph{Desulfobacterota}, whose sequences were more abundant in the
+\emph{C. cylindracea} associated community. As already mentioned, the
+rhizoids and part of the stolon are in contact with the sediment, so
+\emph{Desulfobacterota} are probably a part of the epiphytic community
+that is in contact with the sediment. Similar high rank taxa found in
+this study were described to be specific for other species of
+macrophytes (Burke \emph{et al.}, 2011b; Lachnit \emph{et al.}, 2011;
+Bengtsson \emph{et al.}, 2017). In contrast to high taxonomic ranks, a
+substantial differentiation between host specific communities was found,
+which supports the host-specific hypothesis. Similar host-specificity
+was observed for different species of macroalgae and seagrasses (Lachnit
+\emph{et al.}, 2011; Roth-Schulze \emph{et al.}, 2016; Ugarelli \emph{et
+al.}, 2019; Morrissey \emph{et al.}, 2019). Taken together, at high
+taxonomic ranks a core set of taxa could be described that is
+characteristic for all or many macrophytes, while at low taxonomic ranks
+a community specific to host species could be identified (Egan \emph{et
+al.}, 2013).
 
@@
 sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 51.3
+were accounting for a high proportion of sequences (\(\leq\) 52.8
 \si{\percent}). A very similar proportion of persistent OTUs and their
@@
 conditions in comparison to the seawater. At the level of OTUs seasonal
-changes of \emph{Cymodocea nodosa} and \emph{Caulerpa cylindracea}
-associated communities were identified that could be linked to the
-growth cycle of the seagrass and macroalgae (M. Najdek, personal
-communication). \emph{Cymodocea nodosa} was characterized by a Spring
-community during maximum seagrass proliferation, a Summer community
-during a biomass maximum and a Autumn/Winter community during a biomass
-decay. In contrast, \emph{Caulerpa cylindracea} started to proliferate
-in late Spring and was characterized only by a Summer community during
-maximal biomass increase and by a Autumn/Winter/Spring community when
-the biomass were at the peak and the settlement started to subsequently
-decay. Similar seasonal changes in the epiphytic community was described
-also for other macroalgae (Tujula \emph{et al.}, 2010; Lachnit \emph{et
-al.}, 2011). Higher temporal stability of \emph{Caulerpa cylindracea}
-surface communities in comparison to \emph{Cymodocea nodosa} were also
-observed in the higher proportion of shared communities between two
-consecutive sampling points.
-
-Seasonal patterns of high rank epiphytic taxa did not show strong
-changes. Only \emph{Cyanobateria} were characterized by strong seasonal
-differences. Cyanobacterial sequences were more pronounced in the
-November and December 2017 in comparison to Spring and Summer.
+changes of \emph{C. nodosa} and \emph{C. cylindracea} associated
+communities were identified that could be linked to the growth cycle of
+the seagrass and macroalgae (M. Najdek, personal communication).
+\emph{C. nodosa} was characterized by a Spring community during maximum
+seagrass proliferation, a Summer community during a biomass maximum and
+a Autumn/Winter community during a biomass decay. In contrast, \emph{C.
+cylindracea} started to proliferate in late Spring and was characterized
+only by a Summer community during maximal biomass increase and by a
+Autumn/Winter/Spring community when the biomass were at the peak and the
+settlement started to subsequently decay. Similar seasonal changes in
+the epiphytic community was described also for other macroalgae (Tujula
+\emph{et al.}, 2010; Lachnit \emph{et al.}, 2011). Higher temporal
+stability of \emph{C. cylindracea} surface communities in comparison to
+\emph{C. nodosa} were also observed in the higher proportion of shared
+communities between two consecutive sampling points.
+
+Analysis of seasonal chloroplast sequence abundances showed higher
+values in Autumn/Winter in comparison to Spring/Summer. This pattern is
+not surprising as e.g seagrasses are known to harbor more epiphytes
+during Autumn/Winter (Reyes and Sansón, 2001) and we used an adapted DNA
+isolation protocol that is known to partially coextract DNA from
+planktonic eukaryotes (Korlević \emph{et al.}, 2015). Strong seasonal
+fluctuations of high rank epiphytic taxa were not observed, with the
+exception of \emph{Cyanobateria}. Cyanobacterial sequences were more
+pronounced in November and December in comparison to Spring and Summer.
 Interestingly, in these high proportion months the majority of
@@
 al.}, 2012; Reisser \emph{et al.}, 2014). It is possible than during
-periods of low metabolic activity of the i.g. seagrass there is no so
-active interaction and selection of the epiphytic community, so the
-seagrass leaves are becoming a suitable surface for nonspecific
-colonizers (Zavodnik \emph{et al.}, 1998). \emph{Pleurocapsa} was
-replaced in Spring ans Summer by \emph{Acrophormium},
-\emph{Phormidesmis} and no relative \emph{Nodosilineaceae}. A study of
-coastal microbial mats found also higher proportion of
-\emph{Nodosilineaceae} sequences in Summer, while \emph{Phormidesmis}
-sequences were at their peak in Autumn (Cardoso \emph{et al.}, 2019).
-Other high rank taxa did not showed strong successional patterns. In
-each analyzed group low rank taxa
+periods of low metabolic activity there is no so active interaction and
+selection of the epiphytic community by the seagrass, causing leaves to
+become a suitable surface for nonspecific colonizers (Zavodnik \emph{et
+al.}, 1998). \emph{Pleurocapsa} was replaced in Spring and Summer by
+\emph{Acrophormium}, \emph{Phormidesmis} and no relative
+\emph{Nodosilineaceae}. A study of coastal microbial mats found also
+higher proportion of \emph{Nodosilineaceae} sequences in Summer, while
+\emph{Phormidesmis} sequences were at their peak in Autumn (Cardoso
+\emph{et al.}, 2019). Other high rank taxa did not showed strong
+successional patterns, but in every analyzed group, with the exception
+of \emph{Desulfobacterota}, taxa present throughout the year in similar
+proportions and season specific taxa could be identified. Within
+\emph{Bacteroidota} different groups withing the family
+\emph{Saprospiraceae} (i.g. \emph{Lewinella}, \emph{Portibacter} and
+\emph{Rubidimonas}) were detected through the year. Members of this
+family are often found in association with macrophytes and it is
+suggested that are involved in the hydrolysis and utilization of complex
+carbon sources (Burke \emph{et al.}, 2011b; McIlroy and Nielsen, 2014;
+Crump \emph{et al.}, 2018). On the other hand, families
+\emph{Flavobacteriaceae} and \emph{Rhodothermaceae} showed seasonal
+patterns, with \emph{Flavobacteriaceae} being more pronounced from
+November to June and \emph{Rhodothermaceae} from June to October. Within
+\emph{Alphaproteobacteria} the family \emph{Rhodobacteraceae} was
+comprising the majority of sequences thought the year. This
+metabolically versatile family is often associated with macrophyte
+surfaces and usually is one of the most abundant groups (Burke \emph{et
+al.}, 2011b; Michelou \emph{et al.}, 2013; Pujalte \emph{et al.}, 2014).
+In addition, \emph{Hyphomonadaceae} were found in all samples.
+Interestingly, some of the species within this group contain stalks on
+their cells which can be used to attach to the macrophyte surface
+(Weidner \emph{et al.}, 2000; Abraham and Rohde, 2014). Within the
+\emph{Gammaproteobacteria}, sequences without known representatives were
+the most pronounced group present throughout the year. In addition,
+\emph{Granulosicoccus} was also found in almost all samples.
+\emph{Gammaproteobacteria} are often a major constituent of macrophyte
+epiphytic communities (Burke \emph{et al.}, 2011b; Michelou \emph{et
+al.}, 2013; Crump \emph{et al.}, 2018). Beside these two groups other
+less pronounced taxa showed seasonal and host-specific patterns. In
+example, \emph{C. cylindracea} was characterized by \emph{Thioploca}, a
+known sulfur sediment bacteria and \emph{Cellvibrionaceae}, a family
+whose cultured members are known polysaccharides degraders (Jørgensen
+and Gallardo, 1999; Xie \emph{et al.}, 2017). \emph{Desulfobacterota}
+were found only associated with \emph{C. cylindracea} and no group
+within this phylum showed seasonal patterns. The presence of this phylum
+only on \emph{C. cylindracea} is to be expected as part of the epiphytic
+community is in a direct contact with the sediment. The
+\emph{Desulfobacterota} community was dominated by \emph{Desulfatitalea}
+and no relative \emph{Desulfocapsaceae}, known sulfate sediment groups
+(Kuever, 2014; Higashioka \emph{et al.}, 2015).
+
+In temperate zones marine macrophytes are exhibiting growth cycles, so
+it is not surprising that the associated epiphytic microbial community
+is undergoing partial seasonal changes. In the present study, we could,
+in every analyzed high rank taxa, identify phylogenetic groups that were
+present throughout the year and that were comprising most of the
+sequences and lower proportion taxa showing seasonal patterns connected
+to the macrophyte growth cycle. Studies focusing on functional
+comparisons between communities associated with different hosts showed
+that the majortiy of functions could be found in every community,
+indicating functional redundacy (Roth-Schulze \emph{et al.}, 2016). This
+difference between taxonomic and functionl discrepancy was expalined by
+the lottery hypothesis that hypothesize an initial random colonization
+step performed from a set of functionally equivalent taxonomic groups
+(Burke \emph{et al.}, 2011a; Roth-Schulze \emph{et al.}, 2016). It is
+possible that functional redundancy is a characteristic of high
+abundance taxa detected to be present throughout the year, while
+seasonal and/or host-specific functions are an attribute of taxa
+displaying successional patterns. Further studies connecting taxonomy
+with functional properties will be required to elucidate the degree of
+functional redundancy or specificity in epiphyitc microbial communites.
 
@@
 
-Leaves of \emph{Cymodocea nodosa} were sampled in a \emph{Cymodocea
-nodosa} meadow located in the proximity of the village of Funtana
-(\ang{45;10;39} N, \ang{13;35;42} E). Thalli of \emph{Caulerpa
-cylindracea} were sampled in the same \emph{Cymodocea nodosa} invaded
-meadow in Funtana and on a locality of only \emph{Caulerpa cylindracea}
-located close to the invaded meadow. Sampling of leaves and thalli was
-performed approximately monthly from December 2017 to October 2018
-(\autoref{supp-nseq_notus}). Leaves and thalli were collected by diving
-and transported to the laboratory in containers placed on ice and filled
-with site seawater. Upon arrival to the laboratory, \emph{Cymodocea
-nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while
-\emph{Caulerpa cylindracea} thalli were cut into 5 -- 8 \si{\cm} long
-sections. Leaves and thalli were washed three times with sterile
-artificial seawater (ASW) to remove loosely attached microbial cells.
-Surrounding seawater was collected in 10 \si{\l} containers by diving
+Leaves of \emph{C. nodosa} were sampled in a \emph{C. nodosa} meadow
+located in the proximity of the village of Funtana (\ang{45;10;39} N,
+\ang{13;35;42} E). Thalli of \emph{C. cylindracea} were sampled in the
+same \emph{C. nodosa} invaded meadow in Funtana and on a locality of
+only \emph{C. cylindracea} located close to the invaded meadow. Sampling
+of leaves and thalli was performed approximately monthly from December
+2017 to October 2018 (\autoref{supp-nseq_notus}). Leaves and thalli were
+collected by diving and transported to the laboratory in containers
+placed on ice and filled with site seawater. Upon arrival to the
+laboratory, \emph{C. nodosa} leaves were cut into sections of 1 -- 2
+\si{\cm}, while \emph{C. cylindracea} thalli were cut into 5 -- 8
+\si{\cm} long sections. Leaves and thalli were washed three times with
+sterile artificial seawater (ASW) to remove loosely attached microbial
+cells. Ambient seawater was collected in 10 \si{\l} containers by diving
 and transported to the laboratory where the whole container volume was
@@
 
-DNA from surfaces of \emph{Cymodocea nodosa} and \emph{Caulerpa
-cylindracea} was isolated using a previously modified and adapted
-protocol that allows for a selective epiphytic DNA isolation (Massana
-\emph{et al.}, 1997; Korlević \emph{et al.}, submitted). Briefly, leaves
-and thalli are incubated in a lysis buffer and treated with lysozyme and
-proteinase K. Following the incubations, the mixture containing lyzed
-epiphytic cells is separated from leaves and thalli and extracted using
-a phenol-chloroform procedure. Finally, the extracted DNA is
-precipitated using isopropanol. DNA from seawater picoplankton was
-isolated from 0.2 \si{\um} polycarbonate filters according to (Massana
-\emph{et al.}, 1997) with a slight modification. Following the
-phenol-chloroform extraction steps 1/10 of chilled 3 \si{\Molar} sodium
-acetate (pH 5.2) was added. DNA was precipitated by adding 1 volume of
-chilled isopropanol, incubating the mixtures overnight at \num{-20}
+DNA from surfaces of \emph{C. nodosa} and \emph{C. cylindracea} was
+isolated using a previously modified and adapted protocol that allows
+for a selective epiphytic DNA isolation (Massana \emph{et al.}, 1997;
+Korlević \emph{et al.}, submitted). Briefly, leaves and thalli are
+incubated in a lysis buffer and treated with lysozyme and proteinase K.
+Following the incubations, the mixture containing lyzed epiphytic cells
+is separated from leaves and thalli and extracted using a
+phenol-chloroform procedure. Finally, the extracted DNA is precipitated
+using isopropanol. DNA from seawater picoplankton was isolated from 0.2
+\si{\um} polycarbonate filters according to (Massana \emph{et al.},
+1997) with a slight modification. Following the phenol-chloroform
+extraction steps 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2)
+was added. DNA was precipitated by adding 1 volume of chilled
+isopropanol, incubating the mixtures overnight at \num{-20}
 \si{\degreeCelsius} and centrifuging at 20,000 × g and 4
@@
 including the R Markdown file for this paper are available as a GitHub
-repository (\textbf{TO BE ADDED LATER!}). Based on the ATCC MSA-1002
-mock community included in the analysis an average sequencing error rate
-of 0.01 \si{\percent} was determined, which is in line with previously
-reported values for next-generation sequencing data (Kozich \emph{et
-al.}, 2013; Schloss \emph{et al.}, 2016). In addition, the negative
-controls processed together with the samples yielded on average only 2
-sequences after sequence quality curation.
+repository
+(\url{https://github.com/mkorlevic/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020}).
+Based on the ATCC MSA-1002 mock community included in the analysis an
+average sequencing error rate of 0.01 \si{\percent} was determined,
+which is in line with previously reported values for next-generation
+sequencing data (Kozich \emph{et al.}, 2013; Schloss \emph{et al.},
+2016). In addition, the negative controls processed together with the
+samples yielded on average only 2 sequences after sequence quality
+curation.
 
@@
 
+This work was founded by the Croatian Science Foundation through the
+MICRO-SEAGRASS project (IP-2016-06-7118). We would like to thank
+Margareta Buterer for technical support, Paolo Paliaga for help during
+sampling and the University Computing Center of the University of Zagreb
+for access to the cluster Isabella.
+
 \newpage
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Abraham2014}{}%
+Abraham, W.R. and Rohde, M. (2014) The Family Hyphomonadaceae. In,
+\emph{The prokaryotes: Alphaproteobacteria and betaproteobacteria}.
+Springer-Verlag Berlin Heidelberg, pp. 283--299.
+
 \leavevmode\hypertarget{ref-Allaire2019}{}%
@@
 
+\leavevmode\hypertarget{ref-Higashioka2015}{}%
+Higashioka, Y., Kojima, H., Watanabe, T., and Fukui, M. (2015) Draft
+genome sequence of Desulfatitalea tepidiphila S28bFT. \emph{Genome
+Announcements} \textbf{3}:
+
 \leavevmode\hypertarget{ref-Hollants2013}{}%
@@
 
-\leavevmode\hypertarget{ref-Korlevic_in_press}{}%
+\leavevmode\hypertarget{ref-Jorgensen1999}{}%
+Jørgensen, B.B. and Gallardo, V.A. (1999) Thioploca spp.: filamentous
+sulfur bacteria with nitrate vacuoles. \emph{FEMS Microbiology Ecology}
+\textbf{28}: 301--313.
+
+\leavevmode\hypertarget{ref-Korlevic_submitted}{}%
 Korlević, M., Markovski, M., Zhao, Z., Herndl, G.J., and Najdek, M.
 Selective DNA and Protein Isolation from Marine Macrophyte Surfaces.
+
+\leavevmode\hypertarget{ref-Korlevic2015}{}%
+Korlević, M., Pop Ristova, P., Garić, R., Amann, R., and Orlić, S.
+(2015) Bacterial Diversity in the South Adriatic Sea during a Strong,
+Deep Winter Convection Year. \emph{Applied and Environmental
+Microbiology} \textbf{81}: 1715--1726.
 
@@
 
+\leavevmode\hypertarget{ref-Kuever2014}{}%
+Kuever, J. (2014) The Family Desulfobulbaceae. In, \emph{The
+prokaryotes: Deltaproteobacteria and epsilonproteobacteria}.
+Springer-Verlag Berlin Heidelberg, pp. 75--86.
+
 \leavevmode\hypertarget{ref-Lachnit2009}{}%
@@
 
+\leavevmode\hypertarget{ref-McIlroy2014}{}%
+McIlroy, S.J. and Nielsen, P.H. (2014) The Family Saprospiraceae. In,
+\emph{The prokaryotes: Other major lineages of bacteria and the
+archaea}. Springer-Verlag Berlin Heidelberg, pp. 863--889.
+
 \leavevmode\hypertarget{ref-Michelou2013}{}%
@@
 
+\leavevmode\hypertarget{ref-Pujalte2014}{}%
+Pujalte, M.J., Lucena, T., Ruvira, M.A., Arahal, D.R., and Macián, M.C.
+(2014) The family Rhodobacteraceae. In, \emph{The prokaryotes:
+Alphaproteobacteria and betaproteobacteria}. Springer-Verlag Berlin
+Heidelberg, pp. 439--512.
+
 \leavevmode\hypertarget{ref-Quast2013}{}%
@@
 
+\leavevmode\hypertarget{ref-Reyes2001}{}%
+Reyes, J. and Sansón, M. (2001) Biomass and production of the epiphytes
+on the leaves of Cymodocea nodosa in the Canary Islands. \emph{Botanica
+Marina} \textbf{44}: 307--313.
+
 \leavevmode\hypertarget{ref-Rickert2016}{}%
@@
 
-\leavevmode\hypertarget{ref-Schloss2016}{}%
-Schloss, P.D., Girard, R.A., Martin, T., Edwards, J., and Thrash, J.C.
-(2016) Status of the Archaeal and Bacterial Census: an Update.
-\emph{mBio} \textbf{7}: e00201--16.
+\leavevmode\hypertarget{ref-Schloss2016a}{}%
+Schloss, P.D., Jenior, M.L., Koumpouras, C.C., Westcott, S.L., and
+Highlander, S.K. (2016) Sequencing 16S rRNA gene fragments using the
+PacBio SMRT DNA sequencing system. \emph{PeerJ} \textbf{4}: e1869.
 
@@
 
+\leavevmode\hypertarget{ref-Weidner2000}{}%
+Weidner, Arnold, Stackebrandt, and Pühler (2000) Phylogenetic Analysis
+of Bacterial Communities Associated with Leaves of the Seagrass
+Halophila stipulacea by a Culture-Independent Small-Subunit rRNA Gene
+Approach. \emph{Microbial ecology} \textbf{39}: 22--31.
+
 \leavevmode\hypertarget{ref-Wickham2019}{}%
@@
 
+\leavevmode\hypertarget{ref-Xie2017}{}%
+Xie, Z., Lin, W., and Luo, J. (2017) Comparative Phenotype and Genome
+Analysis of Cellvibrio sp. PR1, a Xylanolytic and Agarolytic Bacterium
+from the Pearl River. \emph{BioMed research international}
+\textbf{2017}: 6304248.
+
 \leavevmode\hypertarget{ref-Yilmaz2014}{}%
@@
 
-\caption{Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between communties associated with the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and coomunities in the surrounding seawater.\label{matrix}}\label{fig:unnamed-chunk-1}
+\caption{Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between communties associated with the surfaces of macrophytes (\textit{C.
+nodosa} [Invaded] and \textit{C.
+cylindracea} [Invaded and Noninvaded]) and coomunities in the ambient seawater.\label{matrix}}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 
-\caption{Proprotion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{shared}}\label{fig:unnamed-chunk-2}
+\caption{Proprotion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\textit{C.
+nodosa} [Invaded] and \textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\label{shared}}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 
-\caption{Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+\caption{Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of macrophytes (\textit{C.
+nodosa} [Invaded] and \textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
 Samples from the same environment or same season are labeld in different colors.
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant bacterial and archaeal sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{community}}\label{fig:unnamed-chunk-4}
+\caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) bacterial and archaeal sequences on the surfaces of macrophytes (\textit{C.
+nodosa} [Invaded] and \textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\label{community}}\label{fig:unnamed-chunk-4}
 \end{figure}
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant cyanobacterial sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+\caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) cyanobacterial sequences on the surfaces of macrophytes (\textit{C.
+nodosa} [Invaded] and \textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
 The proprotion of cyanobaterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant sequences within the \textit{Bacteroidota} on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+\caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) sequences within the \textit{Bacteroidota} on the surfaces of macrophytes (\textit{C.
+nodosa} [Invaded] and \textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
 The proprotion of sequences classified as \textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar.
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant alphaproteobacterial sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+\caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) alphaproteobacterial sequences on the surfaces of macrophytes (\textit{C.
+nodosa} [Invaded] and \textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
 The proprotion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant gammaproteobacterial sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+\caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) gammaproteobacterial sequences on the surfaces of macrophytes (\textit{C.
+nodosa} [Invaded] and \textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
 The proprotion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant sequences within the \textit{Desulfobacterota} on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.
+\caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) sequences within the \textit{Desulfobacterota} on the surfaces of macrophytes (\textit{C.
+nodosa} [Invaded] and \textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
 The proprotion of sequences classified as \textit{Desulfobacterota} in the total bacterial and archaeal community is given above the corresponding bar.

### 2020-03-12 submission/supplementary.Rmd
Manuscript 1st version complete.

@@
 
-\vspace{70mm}
-
-^1$\dagger$^
+\vspace{60mm}
+Marino Korlević^1$*$^, Marsej Markovski^1^, Zihao Zhao^2^, Gerhard J. Herndl^2,3^, Mirjana Najdek^1^
 
@@
 
-$\dagger$ To whom correspondence should be addressed: marino.korlevic@irb.hr
+^$*$^To whom correspondence should be addressed: marino.korlevic@irb.hr
 
-1\. Ruđer Bošković Institute, Center for Marine Research, G. Paliaga 5, Rovinj, Croatia
+1\. Center for Marine Research, Ruđer Bošković Institute, Croatia
 
-2\. University of Vienna, Department of Limnology and Bio-Oceanography, Althanstraße 14, Vienna, Austria
+2\. Department of Functional and Evolutionary Ecology, University of Vienna, Austria
 
-\linenumbers
+3\. Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research, Utrecht University, The Netherlands
+
 \sisetup{mode=text}
@@
 ## Supplementary Figures
-```{r out.width="85%", fig.align="center", fig.cap="Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\\label{rarefaction}", fig.pos="H"}
+
+```{r out.width="85%", fig.align="center", fig.cap="Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{C.
+nodosa} [Invaded] and \\textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\\label{rarefaction}", fig.pos="H"}
 include_graphics("../results/figures/rarefaction.jpg")
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\\label{calculators}", fig.pos="H"}
+```{r out.width="85%", fig.align="center", fig.cap="Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{C.
+nodosa} [Invaded] and \\textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\\label{calculators}", fig.pos="H"}
 include_graphics("../results/figures/calculators.jpg")
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Relative contribution of chloroplast sequences on the surfaces of macrophytes (\\textit{Cymodocea nodosa} [Invaded] and \\textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\\label{chloroplast}", fig.pos="H"}
+```{r out.width="85%", fig.align="center", fig.cap="Relative contribution of chloroplast sequences on the surfaces of macrophytes (\\textit{C.
+nodosa} [Invaded] and \\textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\\label{chloroplast}", fig.pos="H"}
 include_graphics("../results/figures/chloroplast_bar_plot.jpg")

### 2020-03-12 submission/supplementary.tex
Manuscript 1st version complete.

@@
 
-\vspace{70mm}
+\vspace{60mm}
 
-\textsuperscript{1\(\dagger\)}
+Marino Korlević\textsuperscript{1\(*\)}, Marsej
+Markovski\textsuperscript{1}, Zihao Zhao\textsuperscript{2}, Gerhard J.
+Herndl\textsuperscript{2,3}, Mirjana Najdek\textsuperscript{1}
 
@@
 
-\(\dagger\) To whom correspondence should be addressed:
+\textsuperscript{\(*\)}To whom correspondence should be addressed:
 \href{mailto:marino.korlevic@irb.hr}{\nolinkurl{marino.korlevic@irb.hr}}
 
-1. Ruđer Bošković Institute, Center for Marine Research, G. Paliaga 5,
-Rovinj, Croatia
+1. Center for Marine Research, Ruđer Bošković Institute, Croatia
 
-2. University of Vienna, Department of Limnology and Bio-Oceanography,
-Althanstraße 14, Vienna, Austria
+2. Department of Functional and Evolutionary Ecology, University of
+Vienna, Austria
 
-\linenumbers
+3. Department of Marine Microbiology and Biogeochemistry, Royal
+Netherlands Institute for Sea Research, Utrecht University, The
+Netherlands
+
 \sisetup{mode=text}
@@
 
-\caption{Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{rarefaction}}\label{fig:unnamed-chunk-1}
+\caption{Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes (\textit{C.
+nodosa} [Invaded] and \textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\label{rarefaction}}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 
-\caption{Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{calculators}}\label{fig:unnamed-chunk-2}
+\caption{Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of macrophytes (\textit{C.
+nodosa} [Invaded] and \textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\label{calculators}}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 
-\caption{Relative contribution of chloroplast sequences on the surfaces of macrophytes (\textit{Cymodocea nodosa} [Invaded] and \textit{Caulerpa cylindracea} [Invaded and Noninvaded]) and in the surrounding seawater.\label{chloroplast}}\label{fig:unnamed-chunk-3}
+\caption{Relative contribution of chloroplast sequences on the surfaces of macrophytes (\textit{C.
+nodosa} [Invaded] and \textit{C.
+cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\label{chloroplast}}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 
-\begin{longtable}[t]{>{\centering\arraybackslash}p{6em}ccccc}
+\begin{longtable}{>{\centering\arraybackslash}p{6em}ccccc}
 \caption{\label{tab:nseq_notus}Sample ID, Community Type, Sampling Date and Season, No.

### 2020-03-12 code/plot_calculators.R

@@ -11,11 +11,11 @@
 # Loading OTU/sample data
 shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")
 
-# Generation of random rarified community data
+# Generation of random rarefied community data
 rarefied <- shared %>%
   select(-label, -Group, -numOtus) %>%
   rrarefy(., min(rowSums(.))) %>%
-  as_tibble() %>%
+  as_tibble(.name_repair="unique") %>%
   add_column("Group"=shared$Group, .before=TRUE) %>%
   select_if(list(~!is.numeric(.) || sum(.)!=0))
 

### 2020-03-12 code/plot_matrix.R

@@ -15,7 +15,7 @@ shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.
 rarefied <- shared %>%
   select(-label, -Group, -numOtus) %>%
   rrarefy(., min(rowSums(.))) %>%
-  as.tibble() %>%
+  as_tibble(.name_repair="unique") %>%
   add_column("Group"=shared$Group, .before=TRUE) %>%
   select_if(list(~ !is.numeric(.) || sum(.)!=0))
 
@@ -28,8 +28,7 @@ rarefied_metadata <- inner_join(rarefied, metadata, by=c("Group"="ID")) %>%
   filter(date >= "2017-11-01") %>%
   select_if(funs(!is.numeric(.) || sum(.)!=0)) %>%
   group_by(station) %>%
-  select(starts_with("Otu")) %>%
-  summarise_all(sum)
+  summarise_at(vars(starts_with("Otu")), sum)
 
 # Copying the sample labels to the rows (input for library vegan)
 row.names(rarefied_metadata) <- rarefied_metadata$station
@@ -40,12 +39,12 @@ rarefied_metadata <- rarefied_metadata %>%
 
 # Calculating matrices of similarity indices
 jaccard <- vegdist(rarefied_metadata, method="jaccard", binary=T)
-jaccard <- as.tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(jaccard))) %>%
-  rename(V1=X1, V2=X2, jaccard=as.numeric.jaccard.)
+jaccard <- as_tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(jaccard)),
+                     .name_repair= ~c("V1", "V2", "jaccard"))
 
 bray <- vegdist(rarefied_metadata, method="bray", binary=F)
-bray <- as.tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(bray))) %>%
-  rename(V1=X1, V2=X2, bray=as.numeric.bray.)
+bray <- as_tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(bray)),
+                  .name_repair= ~c("V1", "V2", "bray"))
 
 similarity <- inner_join(jaccard, bray, by=c("V1"="V1", "V2"="V2")) %>%
   mutate(jaccard=1-jaccard) %>%

### 2020-03-12 code/plot_pcoa.R

@@ -15,7 +15,7 @@ shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.
 rarefied <- shared %>%
   select(-label, -Group, -numOtus) %>%
   rrarefy(., min(rowSums(.))) %>%
-  as_tibble() %>%
+  as_tibble(.name_repair="unique") %>%
   add_column("Group"=shared$Group, .before=TRUE) %>%
   select_if(list(~!is.numeric(.) || sum(.)!=0))
 
@@ -49,13 +49,13 @@ rarefied_metadata_select <- rarefied_metadata
 spe.bray <- vegdist(select(rarefied_metadata_select, starts_with("Otu")))
 spe.b.pcoa <- cmdscale(spe.bray, k=(nrow(rarefied_metadata_select)-1), eig=TRUE)
 coordinates <- spe.b.pcoa$points[,1:2] %>%
-  as_tibble() %>%
+  as_tibble(.name_repair= ~c("A1", "A2")) %>%
   add_column("Group"=rarefied_metadata_select$label, .before=TRUE)
 coordinates <- inner_join(metadata, coordinates, by=c("label"="Group"))
 
 # PCoA plot generation
 p1 <- ggplot() +
-  geom_point(data=coordinates, aes(x=V1, y=V2, fill=station), shape=21, size=3, stroke=0.5) +
+  geom_point(data=coordinates, aes(x=A1, y=A2, fill=station), shape=21, size=3, stroke=0.5) +
   scale_fill_manual(name=NULL,
                       values=c("#66A61E", "#E6AB02", "#A6761D", "#666666"),
                       breaks=c("F", "FCyM", "FCaM", "FCa"),
@@ -79,13 +79,13 @@ rarefied_metadata_select <- filter(rarefied_metadata, station=="FCyM") %>%
 spe.bray <- vegdist(select(rarefied_metadata_select, starts_with("Otu")))
 spe.b.pcoa <- cmdscale(spe.bray, k=(nrow(rarefied_metadata_select)-1), eig=TRUE)
 coordinates <- spe.b.pcoa$points[,1:2] %>%
-  as_tibble() %>%
+  as_tibble(.name_repair= ~c("A1", "A2")) %>%
   add_column("Group"=rarefied_metadata_select$label, .before=TRUE)
 coordinates <- inner_join(metadata, coordinates, by=c("label"="Group"))
 
 # PCoA plot generation
 p2 <- ggplot() +
[…]

### 2020-03-12 code/plot_seasonal_shared.R

@@ -15,7 +15,7 @@ shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.
 rarefied <- shared %>%
   select(-label, -Group, -numOtus) %>%
   rrarefy(., min(rowSums(.))) %>%
-  as_tibble() %>%
+  as_tibble(.name_repair="unique") %>%
   add_column("Group"=shared$Group, .before=TRUE) %>%
   select_if(list(~ !is.numeric(.) || sum(.)!=0))
 
@@ -28,12 +28,12 @@ rarefied <- rarefied %>%
 
 # Calculating dissimilarity indices
 jaccard <- vegdist(rarefied, method="jaccard", binary=T)
-jaccard <- as_tibble(data.frame(t(combn(rownames(rarefied),2)), as.numeric(jaccard))) %>%
-  rename(V1=X1, V2=X2, jaccard=as.numeric.jaccard.)
+jaccard <- as_tibble(data.frame(t(combn(rownames(rarefied), 2)), as.numeric(jaccard)),
+                     .name_repair= ~c("V1", "V2", "jaccard"))
 
 bray <- vegdist(rarefied, method="bray", binary=F)
-bray <- as_tibble(data.frame(t(combn(rownames(rarefied),2)), as.numeric(bray))) %>%
-  rename(V1=X1, V2=X2, bray=as.numeric.bray.)
+bray <- as_tibble(data.frame(t(combn(rownames(rarefied), 2)), as.numeric(bray)),
+                  .name_repair= ~c("V1", "V2", "bray"))
 
 distance <- inner_join(jaccard, bray, by=c("V1"="V1", "V2"="V2")) %>%
   mutate_at(c("V1", "V2"), list(~ as.character(.)))

### 2020-03-17 submission/manuscript.tex
Analysis rerun.

@@
 in the ambient seawater were also processed for comparison. The number
-of reads per sample ranged between 8,410 and 77,465 sequences
+of reads per sample ranged between 8,408 and 77,465 sequences
 (\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
@@
 (\autoref{supp-rarefaction}). Following quality curation and exclusion
-of sequences mentioned before reads were clustered into 28,729 different
+of sequences mentioned before reads were clustered into 28,736 different
 OTUs at a similarity level of 97 \si{\percent}. Reads numbers were
-normalized to the minimum number of sequences, 8,410
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,019
-different OTUs that ranged from 355 to 2,018 OTUs per sample
+normalized to the minimum number of sequences, 8,408
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,114
+different OTUs that ranged from 379 to 2,016 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (invaded {[}Number of OTUs, 1,690.5 ±
-152.3 OTUs{]} and noninvaded {[}Number of OTUs, 1,733.8 ± 163.5
-OTUs{]}), middle vales for \emph{C. nodosa} (Number of OTUs, 1,066.9 ±
-224.5 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 526 ± 145.6 OTUs)
+found for \emph{C. cylindracea} (invaded {[}Number of OTUs, 1,676.9 ±
+140.3 OTUs{]} and noninvaded {[}Number of OTUs, 1,745.9 ± 156.3
+OTUs{]}), middle vales for \emph{C. nodosa} (Number of OTUs, 1,059.5 ±
+212.3 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 530.8 ± 143.1 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 from each surface associated community were found at every time point.
-These OTUs made also a high proportion of total sequences (39.9 -- 52.8
+These OTUs made also a high proportion of total sequences (40.3 -- 52.4
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 communities indicated a separation between Spring, Summer and
-Autumn/Winter samples (ANOSIM, R = 0.54, \emph{p} \textless{} 0.01),
+Autumn/Winter samples (ANOSIM, R = 0.55, \emph{p} \textless{} 0.001),
 while communities from the surfaces of \emph{C. cylindracea} indicated a
@@
 Summer and Autumn/Winter/Spring samples (ANOSIM, R = 0.31, \emph{p}
-\textless{} 0.05) (\autoref{pcoa}).
+\textless{} 0.01) (\autoref{pcoa}).
 
@@
 noninvaded {[} (7.8 ± 2.4 \si{\percent}){]}) associated communities in
-autumn and for seawater communities in winter (8.8 ± 7.4 \si{\percent}).
+autumn and for seawater communities in winter (8.8 ± 7.5 \si{\percent}).
 Large taxonomic differences between surface associated and seawater
@@
 Sequences related to \emph{Gammaproteobacteria} were comprising, on
-average, 18.7 ± 3.9 \si{\percent} of all sequences (\autoref{gamma}).
+average, 18.6 ± 3.9 \si{\percent} of all sequences (\autoref{gamma}).
 Similarly to previous taxa, large taxonomic differences between seawater
@@
 sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 52.8
+were accounting for a high proportion of sequences (\(\leq\) 52.4
 \si{\percent}). A very similar proportion of persistent OTUs and their
@@
 originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
-used. The sequences obtained in this study have been submitted to the
-European Nucleotide Archive (ENA) under accession numbers \textbf{TO BE
-ADDED LATER!}.
+used. Sequences obtained in this study have been deposited in the
+European Nucleotide Archive (ENA) at EMBL-EBI under accession number
+PRJEB37267.
 

### 2020-03-18 submission/manuscript.Rmd
README done
Changed to Mixed and Monospecific.

@@
 Only *Cyanobacteria* showed seasonal change, while other high rank taxa were present throughout the year.
-In every analyzed high rank taxa, phylogenetic groups present throughout the year and comprising most of the sequences could be identified together with low proportion taxa showing seasonal patterns connected to the macrophyte growth cycle.
+In every analyzed high rank taxa, phylogenetic groups present throughout the year comprised most of the sequences and could be identified together with low proportion taxa showing seasonal patterns connected to the macrophyte growth cycle.
 Taken together, epiphytic microbial communities of the seagrass *Cymodocea nodosa* and the macroalgae *Caulerpa cylindracea* appear to be host-specific and contain taxa that undergo successional changes.
@@
 Recently, it was shown that even different morphological niches within the same alga had a higher influence on bacterial community variation than biogeography or environmental factors [@Morrissey2019].
-While there is high community variation between host species is was observed that the majority of metagenome determined functions were conserved both between host species and individuals [@Burke2011a; @Roth-Schulze2016].
+While there is high community variation between host species it was observed that the majority of metagenome determined functions were conserved both between host species and individuals [@Burke2011a; @Roth-Schulze2016].
 This discrepancy between taxonomic and functional composition could be explained by the lottery hypothesis.
@@
 The majority of studies describing macrophyte epiphytic communities did not encompass seasonal changes [@Ugarelli2019; @Lachnit2009; @Crump2008; @Burke2011b; @Roth-Schulze2016].
-In addition, if seasonal changes were taken into account low temporal frequency and/or methodologies that do not allow for high taxonomic resolution were used [@Tujula2010; @Lachnit2011; @Miranda2013; @Michelou2013].
+In addition, if seasonal changes were taken into account low temporal frequency and/or methodologies that do not allow high taxonomic resolution were used [@Tujula2010; @Lachnit2011; @Miranda2013; @Michelou2013].
 In the present study we describe the seasonal dynamics of bacterial and archaeal communities on the surfaces of the seagrass *Cymodocea nodosa* and siphonous macroalgae *Caulerpa cylindracea* determined on a mostly monthly scale.
@@
 Following quality curation and exclusion of sequences mentioned before reads were clustered into `r no_otus` different OTUs at a similarity level of 97 \si{\percent}.
-Reads numbers were normalized to the minimum number of sequences, `r min_max$min` (\autoref{supp-nseq_notus}), through rarefaction resulting in `r no_otus_rarefied` different OTUs that ranged from `r min(otu_min_max)` to `r max(otu_min_max)` OTUs per sample (\autoref{supp-calculators}).
+Read numbers were normalized to the minimum number of sequences, `r min_max$min` (\autoref{supp-nseq_notus}), through rarefaction resulting in `r no_otus_rarefied` different OTUs that ranged from `r min(otu_min_max)` to `r max(otu_min_max)` OTUs per sample (\autoref{supp-calculators}).
 To determine seasonal changes of richness and diversity the Observed Number of OTUs, Chao1, ACE, Exponential Shannon [@Jost2006] and Inverse Simpson were calculated after normalization through rarefaction.
@@
 On average, higher values were found for *C.
-cylindracea* (invaded [Number of OTUs, `r S.obs_fcam$mean` ± `r S.obs_fcam$sd` OTUs] and noninvaded [Number of OTUs, `r S.obs_fca$mean` ± `r S.obs_fca$sd` OTUs]), middle vales for *C.
+cylindracea* (mixed [Number of OTUs, `r S.obs_fcam$mean` ± `r S.obs_fcam$sd` OTUs] and monospecific [Number of OTUs, `r S.obs_fca$mean` ± `r S.obs_fca$sd` OTUs]), middle values for *C.
 nodosa* (Number of OTUs, `r S.obs_fcym$mean` ± `r S.obs_fcym$sd` OTUs) and lower values for picoplankton communities in the ambient seawater (Number of OTUs, `r S.obs_f$mean` ± `r S.obs_f$sd` OTUs) (\autoref{supp-calculators}).
@@
 nodosa* communities showed a slow increase towards the end of the study, while *C.
-cylindracea* (invaded and noninvaded) communities were characterized by slightly larger values in Spring and Summer in comparison to Autumn and Winter (\autoref{supp-calculators}).
+cylindracea* (mixed and monospecific) communities were characterized by slightly larger values in Spring and Summer in comparison to Autumn and Winter (\autoref{supp-calculators}).
 
@@
 Coefficients were determined after normalization through rarefaction and binning of samples from a particular environment.
-The highest proportion of shared OTUs and community was found between invaded and noninvaded  *C.
+The highest proportion of shared OTUs and community was found between mixed and monospecific  *C.
 cylindracea* (Jaccard, `r format(filter(similarity, V1=="FCa", V2=="FCaM")$jaccard, digits=2, nsmall=2)`; Bray-Curtis, `r format(filter(similarity, V1=="FCa", V2=="FCaM")$bray, digits=2, nsmall=2)`), while lower shared values were calculated between seawater and epiphytic communities (\autoref{matrix}).
@@
 In addition, only `r min(core$core_otus)` -- `r max(core$core_otus)` \si{\percent} of OTUs from each surface associated community were found at every time point.
-These OTUs made also a high proportion of total sequences (`r min(core$core_seq)` -- `r max(core$core_seq)` \si{\percent}).
+These OTUs also made a high proportion of total sequences (`r min(core$core_seq)` -- `r max(core$core_seq)` \si{\percent}).
 To further disentangle the environmental and seasonal community dissimilarity a Principal Coordinates Analysis (PCoA) based on Bray-Curtis distances and OTU abundances was applied.
@@
 Seasonal changes of *C.
-nodosa* associated communities indicated a separation between Spring, Summer and Autumn/Winter samples (ANOSIM, R = `r format(anosim_fcym$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fcym_signif`), while communities from the surfaces of *C.
-cylindracea* indicated a non so strongly supported, as in previous cases, separation between Summer and Autumn/Winter/Spring samples (ANOSIM, R = `r format(anosim_fca$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fca_signif`) (\autoref{pcoa}).
+nodosa* associated communities indicated a separation between Spring, Summer and Autumn/Winter samples (ANOSIM, R = `r format(anosim_fcym$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fcym_signif`).
+For *C.
+cylindracea* associated communities a separation between Summer and Autumn/Winter/Spring samples was observed that was not so strongly supported (ANOSIM, R = `r format(anosim_fca$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fca_signif`) (\autoref{pcoa}).
 
@@
 Communities associated with *C.
-nodosa* were consisted of same groups with the addition of *Planctomycetota* whose contribution was higher in summer 2018.
-In addition, communities from invaded and noninvaded *C.
+nodosa* consisted of same groups with the addition of *Planctomycetota* whose contribution was higher in summer 2018.
+In addition, communities from mixed and monospecific *C.
 cylindracea* were similar and characterized by same groups as seawater and *C.
@@
 nodosa* (`r filter(cyano_station_season_whole, station=="FCyM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCyM" & season=="Autumn")$sd` \si{\percent}) and *C.
-cylindracea* (invaded [ (`r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$sd` \si{\percent})] and noninvaded [ (`r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$sd` \si{\percent})]) associated communities in autumn and for seawater communities in winter (`r filter(cyano_station_season_whole, station=="F" & season=="Winter")$mean` ± `r filter(cyano_station_season_whole, station=="F" & season=="Winter")$sd` \si{\percent}).
+cylindracea* (mixed [ (`r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$sd` \si{\percent})] and monospecific [ (`r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$sd` \si{\percent})]) associated communities in autumn and for seawater communities in winter (`r filter(cyano_station_season_whole, station=="F" & season=="Winter")$mean` ± `r filter(cyano_station_season_whole, station=="F" & season=="Winter")$sd` \si{\percent}).
 Large taxonomic differences between surface associated and seawater cyanobacterial communities were observed.
@@
 In contrast, uncultured *Rhodothermaceae* showed higher proportions from June 2018 until the end of the study period.
-Surface associated *Bacteroidota* communities were very diverse as could be observed in the the high proportion of taxa that grouped as other *Bacteroidota* (\autoref{bactero}).
+Surface associated *Bacteroidota* communities were very diverse as could be observed in the high proportion of taxa that grouped as other *Bacteroidota* (\autoref{bactero}).
 
@@
 Similarly to previous taxa, large taxonomic differences between seawater and surface associated communities were found.
-Seawater communities were mainly comprised of the OM60 (NOR5) clade, *Litoricola*, *Acinetobacter* and the SAR86 clade, while epiphytic communities were mainly composed of no relative *Gammaproteobacteria* and *Granulosicocus*.
+Seawater communities were mainly comprised of the OM60 (NOR5) clade, *Litoricola*, *Acinetobacter* and the SAR86 clade, while epiphytic communities were mainly composed of no relative *Gammaproteobacteria* and *Granulosicoccus*.
 Beside these two groups specific to all three epiphytic communities, *C.
-nodosa* was characterized by *Arenicella*, no relative *Burkholderiales* and *Methylotenera*, while *Thioploca*, no relative *Cellvibrionaceae* and *Reinekea* were more specific to both invaded and noninvaded *C.
+nodosa* was characterized by *Arenicella*, no relative *Burkholderiales* and *Methylotenera*, while *Thioploca*, no relative *Cellvibrionaceae* and *Reinekea* were more specific to both mixed and monospecific *C.
 cylindracea*.
@@
 For the *C.
-cylindracea* specific taxa no relative *Cellvibrionaceae* and *Reinekea* showed some seasonality and were characterisitic for samples originating from June to October 2018.
+cylindracea* specific taxa no relative *Cellvibrionaceae* and *Reinekea* showed some seasonality and were characteristic for samples originating from June to October 2018.
 In addition, similarly to *Bacteroidota*, a large proportion of the surface associated community was grouped as other *Gammaproteobacteria* indicating high diversity within this group (\autoref{gamma}).
@@
 cylindracea*.
-On average they were comprising `r desulfo_all_whole$mean` ± `r desulfo_all_whole$sd` \si{\percent} of all sequences.
-While seawater and *C.
-nodosa* communities were consisted of only `r filter(desulfo_station_whole, station=="F")$mean` ± `r filter(desulfo_station_whole, station=="F")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCyM")$mean` ± `r filter(desulfo_station_whole, station=="FCyM")$sd` \si{\percent} *Desulfobacterota* sequences, respectively, in the invaded and noninvaded *C.
+On average they comprised `r desulfo_all_whole$mean` ± `r desulfo_all_whole$sd` \si{\percent} of all sequences.
+Seawater and *C.
+nodosa* communities consisted of only `r filter(desulfo_station_whole, station=="F")$mean` ± `r filter(desulfo_station_whole, station=="F")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCyM")$mean` ± `r filter(desulfo_station_whole, station=="FCyM")$sd` \si{\percent} *Desulfobacterota* sequences, respectively.
+In the mixed and monospecific *C.
 cylindracea* communities their proportion was `r filter(desulfo_station_whole, station=="FCaM")$mean` ± `r filter(desulfo_station_whole, station=="FCaM")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCa")$mean` ± `r filter(desulfo_station_whole, station=="FCa")$sd` \si{\percent}, respectively (\autoref{desulfo}).
-The community was mainly consisted of no relative *Desulfobacteraceae*, *Desulfatitalea*, no relative *Desulfobulbaceae*, *Desulfobulbus*, no relative *Desulfocapsaceae*, *Desulfopila*, *Desulforhopalus*, *Desulfotalea*, SEEP-SRB4 and uncultured *Desulfocapsaceae* (\autoref{desulfo}).
+The community consisted mainly of no relative *Desulfobacteraceae*, *Desulfatitalea*, no relative *Desulfobulbaceae*, *Desulfobulbus*, no relative *Desulfocapsaceae*, *Desulfopila*, *Desulforhopalus*, *Desulfotalea*, SEEP-SRB4 and uncultured *Desulfocapsaceae* (\autoref{desulfo}).
 
@@
 ## Discussion
-Surfaces of marine macrophytes are harboring biofilms consisted of diverse microbial taxa [@Egan2013; @Tarquinio2019].
+Surfaces of marine macrophytes harbor biofilms consisting of diverse microbial taxa [@Egan2013; @Tarquinio2019].
 No standard protocol has been developed to study these macophyte associated microbes [@Ugarelli2019].
-Different procedures of microbial cells removal from host surfaces were described, such as host tissue shaking [@Noges2010], scraping [@Uku2007] and ultrasonication [@Cai2014].
+Different procedures for removal of microbial cells from host surfaces were described, such as host tissue shaking [@Noges2010], scraping [@Uku2007] and ultrasonication [@Cai2014].
 All these methods showed different removal efficiencies but none was enabling a complete removal of attached microbial cells.
@@
 In the present study, highest richness values were observed for *C.
-cylindracea* (invaded and noninvaded), middle for *C.
+cylindracea* (mixed and monospecific), middle for *C.
 nodosa* and lowest for seawater derived communities.
@@
 In addition, highest values observed for *C.
-cylindracea* are probably a consequence of part of epiphytic OTUs that are in contact with the sediment.
+cylindracea* are partly due to its contact with the sediment.
 *C.
 cylindracea* stolon is attached to the sediment surface with rhizoids, so the stolon and rhizoids are in a direct contact with the sediment.
-Part of the surface attached *Caulerpa cylnidracea* community is therefore comprised of OTUs that are in contact with a different environment and that could cause the observed increase in richness.
 In addition, seasonal richness differences observed for surface attached communities showed slightly higher values in spring and summer.
@@
 
-Since the colonization of macrophyte surfaces is performed from a pool of seawater prokaryotic cells it is interesting to see to which extent do these two communities differ.
+Since the colonization of macrophyte surfaces is performed from a pool of prokaryotic cells from the ambient seawater, it was interesting to see to which extent these two communities differ.
 We observed a strong differentiation between the surface attached and seawater communities at the level of OTUs that is in agreement with most published studies [@Ugarelli2019; @Roth-Schulze2016; @Burke2011b; @Michelou2013; @Crump2018].
@@
 When the taxonomic composition at high ranks was analyzed no such strong differentiation was noticed.
-Phyla and classes such as: *Actinobacteriota*, *Bacteroidota*, *Cyanobacteria*, *Alphaproteobacteria*, *Gammaproteobacteria* and *Verrucomicrobiota*, were described that is in agreement with previously reported data [@Egan2013; @Michelou2013; @Burke2011b].
+Phyla and classes such as: *Actinobacteriota*, *Bacteroidota*, *Cyanobacteria*, *Alphaproteobacteria*, *Gammaproteobacteria* and *Verrucomicrobiota* were described that is in agreement with previously reported data [@Egan2013; @Michelou2013; @Burke2011b].
 In contrast, when low taxonomic ranks were analyzed (i.g.
@@
 The only high rank phylum that was differing between *C.
-nodosa* ans *C.
+nodosa* and *C.
 cylindracea* was *Desulfobacterota*, whose sequences were more abundant in the *C.
 cylindracea* associated community.
-As already mentioned, the rhizoids and part of the stolon are in contact with the sediment, so *Desulfobacterota* are probably a part of the epiphytic community that is in contact with the sediment.
+As already mentioned, the rhizoids and part of the stolon are in contact with the sediment, so *Desulfobacterota* are probably a part of the epiphytic community that was in contact with the sediment.
 Similar high rank taxa found in this study were described to be specific for other species of macrophytes [@Burke2011b; @Lachnit2011; @Bengtsson2017].
@@
 
-Seasonal richness changes in the epiphytic community were substantial as could be observed in the proportion of OTUs that could be find at every sampling time ($\leq$ `r max(core$core_otus)` \si{\percent}).
+Seasonal richness changes in the epiphytic community were substantial as could be observed in the proportion of OTUs that could be found at every sampling time ($\leq$ `r max(core$core_otus)` \si{\percent}).
 Interestingly, these OTUs were accounting for a high proportion of sequences ($\leq$ `r max(core$core_seq)` \si{\percent}).
@@
 In comparison to the seawater community, a lower degree of seasonal shifts was observed for the surface associated communities.
-It seems, that microniches on the surfaces of macrophytes are providing more stable conditions in comparison to the seawater.
+It seems, microniches on the surfaces of macrophytes are providing more stable conditions in comparison to the seawater.
 At the level of OTUs seasonal changes of *C.
@@
 *C.
-nodosa* was characterized by a Spring community during maximum seagrass proliferation, a Summer community during a biomass maximum and a Autumn/Winter community during a biomass decay.
+nodosa* was characterized by a Spring community during maximum seagrass proliferation, a Summer community during a biomass maximum and a Autumn/Winter community during a biomass senescence.
 In contrast, *C.
-cylindracea* started to proliferate in late Spring and was characterized only by a Summer community during maximal biomass increase and by a Autumn/Winter/Spring community when the biomass were at the peak and the settlement started to subsequently decay.
-Similar seasonal changes in the epiphytic community was described also for other macroalgae [@Lachnit2011; @Tujula2010].
+cylindracea* started to proliferate in late Spring and was characterized only by a Summer community during maximal biomass increase and by a Autumn/Winter/Spring community when the biomass was at the peak and the settlement started to subsequently decay.
+Similar seasonal changes in the epiphytic community was also described for other macroalgae [@Lachnit2011; @Tujula2010].
 Higher temporal stability of *C.
@@
 Analysis of seasonal chloroplast sequence abundances showed higher values in Autumn/Winter in comparison to Spring/Summer.
-This pattern is not surprising as e.g seagrasses are known to harbor more epiphytes during Autumn/Winter [@Reyes2001] and we used an adapted DNA isolation protocol that is known to partially coextract DNA from planktonic eukaryotes [@Korlevic2015].
+This pattern is not surprising as seagrasses are known to harbor more epiphytes during Autumn/Winter [@Reyes2001].
+Furthermore, we used an adapted DNA isolation protocol that is known to partially coextract DNA from planktonic eukaryotes [@Korlevic2015].
 Strong seasonal fluctuations of high rank epiphytic taxa were not observed, with the exception of *Cyanobateria*.
@@
 Interestingly, in these high proportion months the majority of cyanobacterial sequences were classified as *Pleurocapsa*, a group known to colonized different living and nonliving surfaces [@Longford2007; @Burns2004; @Mobberley2012; @Reisser2014].
-It is possible than during periods of low metabolic activity there is no so active interaction and selection of the epiphytic community by the seagrass, causing leaves to become a suitable surface for nonspecific colonizers [@Zavodnik1998].
+It is possible that during periods of low metabolic activity there is a reduced interaction and selection of the epiphytic community by the seagrass, causing leaves to become a suitable surface for nonspecific colonizers [@Zavodnik1998].
 *Pleurocapsa* was replaced in Spring and Summer by *Acrophormium*, *Phormidesmis* and no relative *Nodosilineaceae*.
 A study of coastal microbial mats found also higher proportion of *Nodosilineaceae* sequences in Summer, while *Phormidesmis* sequences were at their peak in Autumn [@Cardoso2019].
-Other high rank taxa did not showed strong successional patterns, but in every analyzed group, with the exception of *Desulfobacterota*, taxa present throughout the year in similar proportions and season specific taxa could be identified.
+Other high rank taxa did not show strong successional patterns.
+In every analyzed group, with the exception of *Desulfobacterota*, taxa present throughout the year in similar proportions and season specific taxa could be identified.
 Within *Bacteroidota* different groups withing the family *Saprospiraceae* (i.g.
 *Lewinella*, *Portibacter* and *Rubidimonas*) were detected through the year.
-Members of this family are often found in association with macrophytes and it is suggested that are involved in the hydrolysis and utilization of complex carbon sources [@Burke2011b; @Crump2018; @McIlroy2014].
+Members of this family are often found in association with macrophytes and it is suggested that they are involved in the hydrolysis and utilization of complex carbon sources [@Burke2011b; @Crump2018; @McIlroy2014].
 On the other hand, families *Flavobacteriaceae* and *Rhodothermaceae* showed seasonal patterns, with *Flavobacteriaceae* being more pronounced from November to June and *Rhodothermaceae* from June to October.
-Within *Alphaproteobacteria* the family *Rhodobacteraceae* was comprising the majority of sequences thought the year.
+Within *Alphaproteobacteria* the family *Rhodobacteraceae* was comprising the majority of sequences throughout the year.
 This metabolically versatile family is often associated with macrophyte surfaces and usually is one of the most abundant groups [@Burke2011b; @Michelou2013; @Pujalte2014].
@@
 Beside these two groups other less pronounced taxa showed seasonal and host-specific patterns.
-In example, *C.
-cylindracea* was characterized by *Thioploca*, a known sulfur sediment bacteria and *Cellvibrionaceae*, a family whose cultured members are known polysaccharides degraders [@Jorgensen1999; @Xie2017].
+For example, *C.
+cylindracea* was characterized by *Thioploca*, a known sulfur sediment bacteria and *Cellvibrionaceae*, a family whose cultured members are known polysaccharide degraders [@Jorgensen1999; @Xie2017].
 *Desulfobacterota* were found only associated with *C.
@@
 The presence of this phylum only on *C.
-cylindracea* is to be expected as part of the epiphytic community is in a direct contact with the sediment.
+cylindracea* is to be expected as part of the epiphytic community is directly in contact with the sediment.
 The *Desulfobacterota* community was dominated by *Desulfatitalea* and no relative *Desulfocapsaceae*, known sulfate sediment groups [@Higashioka2015; @Kuever2014].
@@
 In the present study, we could, in every analyzed high rank taxa, identify phylogenetic groups that were present throughout the year and that were comprising most of the sequences and lower proportion taxa showing seasonal patterns connected to the macrophyte growth cycle.
-Studies focusing on functional comparisons between communities associated with different hosts showed that the majortiy of functions could be found in every community, indicating functional redundacy [@Roth-Schulze2016].
-This difference between taxonomic and functionl discrepancy was expalined by the lottery hypothesis that hypothesize an initial random colonization step performed from a set of functionally equivalent taxonomic groups [@Burke2011a; @Roth-Schulze2016].
+Studies focusing on functional comparisons between communities associated with different hosts showed that the majority of functions could be found in every community, indicating functional redundancy [@Roth-Schulze2016].
+This difference between taxonomic and functional discrepancy was explained by the lottery hypothesis that hypothesizes an initial random colonization step performed from a set of functionally equivalent taxonomic groups [@Burke2011a; @Roth-Schulze2016].
 It is possible that functional redundancy is a characteristic of high abundance taxa detected to be present throughout the year, while seasonal and/or host-specific functions are an attribute of taxa displaying successional patterns.
-Further studies connecting taxonomy with functional properties will be required to elucidate the degree of functional redundancy or specificity in epiphyitc microbial communites.
+Further studies connecting taxonomy with functional properties will be required to elucidate the degree of functional redundancy or specificity in epiphytic microbial communities.
 
@@
 ### Sampling
-Leaves of *C.
-nodosa* were sampled in a *C.
-nodosa* meadow located in the proximity of the village of Funtana (\ang{45;10;39} N, \ang{13;35;42} E).
-Thalli of *C.
-cylindracea* were sampled in the same *C.
-nodosa* invaded meadow in Funtana and on a locality of only *C.
-cylindracea* located close to the invaded meadow.
-Sampling of leaves and thalli was performed approximately monthly from December 2017 to October 2018 (\autoref{supp-nseq_notus}).
+Sampling was performed in the Bay of Funtana, northern Adriatic Sea (\ang{45;10;39} N, \ang{13;35;42} E).
+Leaves and thalli were sampled in a meadow of *Cymodocea nodosa* invaded by the invasive *Caulerpa cylindracea* (mixed settlement) and in a monospecific settlement of *Caulerpa cylindracea* located in the proximity of the meadow at approximately monthly intervals from December 2017 to October 2018 (\autoref{supp-nseq_notus}).
 Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with site seawater.
@@
 Finally, the extracted DNA is precipitated using isopropanol.
-DNA from seawater picoplankton was isolated from 0.2 \si{\um} polycarbonate filters according to [@Massana1997] with a slight modification.
+DNA from seawater picoplankton was isolated from 0.2 \si{\um} polycarbonate filters according to @Massana1997 with a slight modification.
 Following the phenol-chloroform extraction steps 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added.
@@
 Negative control was comprised of PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
-The sequences obtained in this study have been submitted to the European Nucleotide Archive (ENA) under accession numbers **TO BE ADDED LATER!**.
+Sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession number PRJEB37267.
 
@@
 ## Figures
-```{r out.width="70%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between communties associated with the surfaces of macrophytes (\\textit{C.
-nodosa} [Invaded] and \\textit{C.
-cylindracea} [Invaded and Noninvaded]) and coomunities in the ambient seawater.\\label{matrix}", fig.pos="H"}
+```{r out.width="70%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between assemblages associated with the surfaces of macrophytes (\\textit{C.
+nodosa} [Mixed] and \\textit{C.
+cylindracea} [Mixed and Monospecific]) and comunities in the ambient seawater.\\label{matrix}", fig.pos="H"}
 include_graphics("../results/figures/matrix.jpg")
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Proprotion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\\textit{C.
-nodosa} [Invaded] and \\textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\\label{shared}", fig.pos="H"}
+```{r out.width="85%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\\textit{C.
+nodosa} [Mixed] and \\textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\\label{shared}", fig.pos="H"}
 include_graphics("../results/figures/seasonal_shared.jpg")
@@
 ```{r out.width="55%", fig.align="center", fig.cap="Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{C.
-nodosa} [Invaded] and \\textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
-Samples from the same environment or same season are labeld in different colors.
+nodosa} [Mixed] and \\textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+Samples from the same environment or same season are labeled in different colors.
 The proportion of explained variation by each axis is shown on the corresponding axis in parentheses.\\label{pcoa}", fig.pos="H"}
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) bacterial and archaeal sequences on the surfaces of macrophytes (\\textit{C.
-nodosa} [Invaded] and \\textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\\label{community}", fig.pos="H"}
+nodosa} [Mixed] and \\textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+NR -- No Relative\\label{community}", fig.pos="H"}
 include_graphics("../results/figures/community_bar_plot.jpg")
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) cyanobacterial sequences on the surfaces of macrophytes (\\textit{C.
-nodosa} [Invaded] and \\textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
-The proprotion of cyanobaterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
+nodosa} [Mixed] and \\textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+The proportion of cyanobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
 NR -- No Relative\\label{cyano}", fig.pos="H"}
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) sequences within the \\textit{Bacteroidota} on the surfaces of macrophytes (\\textit{C.
-nodosa} [Invaded] and \\textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
-The proprotion of sequences classified as \\textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar.
+nodosa} [Mixed] and \\textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+The proportion of sequences classified as \\textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar.
 NR -- No Relative\\label{bactero}", fig.pos="H"}
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) alphaproteobacterial sequences on the surfaces of macrophytes (\\textit{C.
-nodosa} [Invaded] and \\textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
-The proprotion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
+nodosa} [Mixed] and \\textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+The proportion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
 NR -- No Relative\\label{alpha}", fig.pos="H"}
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) gammaproteobacterial sequences on the surfaces of macrophytes (\\textit{C.
-nodosa} [Invaded] and \\textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
-The proprotion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
+nodosa} [Mixed] and \\textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+The proportion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
 NR -- No Relative\\label{gamma}", fig.pos="H"}
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) sequences within the \\textit{Desulfobacterota} on the surfaces of macrophytes (\\textit{C.
-nodosa} [Invaded] and \\textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
-The proprotion of sequences classified as \\textit{Desulfobacterota} in the total bacterial and archaeal community is given above the corresponding bar.
+nodosa} [Mixed] and \\textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+The proportion of sequences classified as \\textit{Desulfobacterota} in the total bacterial and archaeal community is given above the corresponding bar.
 NR -- No Relative\\label{desulfo}", fig.pos="H"}

### 2020-03-18 submission/supplementary.Rmd
Changed to Mixed and Monospecific.

@@
 ```{r out.width="85%", fig.align="center", fig.cap="Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{C.
-nodosa} [Invaded] and \\textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\\label{rarefaction}", fig.pos="H"}
+nodosa} [Mixed] and \\textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\\label{rarefaction}", fig.pos="H"}
 include_graphics("../results/figures/rarefaction.jpg")
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{C.
-nodosa} [Invaded] and \\textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\\label{calculators}", fig.pos="H"}
+nodosa} [Mixed] and \\textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\\label{calculators}", fig.pos="H"}
 include_graphics("../results/figures/calculators.jpg")
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Relative contribution of chloroplast sequences on the surfaces of macrophytes (\\textit{C.
-nodosa} [Invaded] and \\textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\\label{chloroplast}", fig.pos="H"}
+nodosa} [Mixed] and \\textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\\label{chloroplast}", fig.pos="H"}
 include_graphics("../results/figures/chloroplast_bar_plot.jpg")
@@
   mutate(station=str_replace(station, "^F$", "Seawater")) %>%
-  mutate(station=str_replace(station, "^FCyM$", "\\\\textit{Cymodocea nodosa} (Invaded)")) %>%
-  mutate(station=str_replace(station, "^FCaM$", "\\\\textit{Caulerpa cylindracea} (Invaded)")) %>%
-  mutate(station=str_replace(station, "^FCa$", "\\\\textit{Caulerpa cylindracea} (Nonnvaded)")) %>%
+  mutate(station=str_replace(station, "^FCyM$", "\\\\textit{Cymodocea nodosa} (Mixed)")) %>%
+  mutate(station=str_replace(station, "^FCaM$", "\\\\textit{Caulerpa cylindracea} (Mixed)")) %>%
+  mutate(station=str_replace(station, "^FCa$", "\\\\textit{Caulerpa cylindracea} (Monospecific)")) %>%
   mutate(ID=str_replace(ID, "^23_1$", "23a")) %>%

### 2020-03-18 code/plot_calculators.R

@@ -124,7 +124,7 @@ p1 <- filter(estimators_indices_metadata, station=="FCyM") %>%
   scale_fill_manual(values=fills_p1) +
   scale_y_continuous(limits=c(300, 2500)) +
   labs(x="", y="") +
-  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Mixed)')")) +
   theme +
   theme(axis.ticks.x=element_blank(), axis.text.x=element_blank())
 
@@ -145,7 +145,7 @@ p2 <- filter(estimators_indices_metadata, station=="FCyM") %>%
 fcym <- cowplot::plot_grid(p1, p2, nrow=2, ncol=1, rel_heights=c(1,0.75), align="v")
 
 # Plots generation
-# Caulerpa cylindracea (Invaded) samples
+# Caulerpa cylindracea (Mixed) samples
 p1 <- filter(estimators_indices_metadata, station=="FCaM") %>%
   filter(estimator_index=="S.obs" | estimator_index=="S.chao1" | estimator_index=="S.ACE") %>%
   ggplot(aes(x=date, y=value, linetype=estimator_index, shape=estimator_index,
@@ -157,7 +157,7 @@ p1 <- filter(estimators_indices_metadata, station=="FCaM") %>%
   scale_fill_manual(values=fills_p1) +
   scale_y_continuous(limits=c(1300, 4300)) +
   labs(x="", y="") +
-  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Mixed)')")) +
   theme +
   theme(axis.ticks.x=element_blank(), axis.text.x=element_blank())
 
@@ -179,7 +179,7 @@ fcam <- cowplot::plot_grid(p1, p2, nrow=2, ncol=1, rel_heights=c(1,0.75), align=
   cowplot_theme
 
 # Plots generation
-# Caulerpa cylindracea (Noninvaded) samples
+# Caulerpa cylindracea (Monospecific) samples
 p1 <- filter(estimators_indices_metadata, station=="FCa") %>%
   filter(estimator_index=="S.obs" | estimator_index=="S.chao1" | estimator_index=="S.ACE") %>%
[…]

### 2020-03-18 code/plot_community_bar_plot.R

@@ -61,7 +61,7 @@ names <- parse(text=case_when(plot$taxon=="Chloroplast" ~ paste0("plain('", plot
                               plot$taxon=="Bacteria_unclassified" ~ "italic('Bacteria')~plain('(No Relative)')",
                               plot$taxon=="Marinimicrobia_(SAR406_clade)" ~ "italic('Marinimicrobia')",
                               plot$taxon=="Other" ~ paste0("plain('", plot$taxon, "')"),
-                              plot$taxon=="No_Relative" ~ "plain('No Relative')",
+                              plot$taxon=="No_Relative" ~ "plain('NR')",
                               TRUE ~ paste0("italic('", plot$taxon, "')")))
 
 # Tidying the sequence abundance data
@@ -111,7 +111,7 @@ fcym <- filter(plot, station=="FCyM") %>%
   geom_bar(aes(x=date, y=abundance, fill=taxon), stat="identity", colour="black", size=0.3, width=8) +
   scale_fill_manual(values=color, labels=names) + 
   labs(x=NULL, y="%") +
-  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Mixed)')")) +
   scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=c(0, 0), breaks=seq(0, 100, by=10)) +
@@ -122,7 +122,7 @@ fcam <- filter(plot, station=="FCaM") %>%
   geom_bar(aes(x=date, y=abundance, fill=taxon), stat="identity", colour="black", size=0.3, width=8) +
   scale_fill_manual(values=color, labels=names) + 
   labs(x=NULL, y="%") +
-  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Mixed)')")) +
   scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=c(0, 0), breaks=seq(0, 100, by=10)) +
[…]

### 2020-03-18 code/plot_community_bar_plot_alphaproteobacteria.R

@@ -145,7 +145,7 @@ fcym <- filter(plot, station=="FCyM") %>%
   geom_text(data=filter(whole, station=="FCyM"), aes(x=date, y=75, label=paste(format(round(abundance, 1), nsmall=1), "%")),
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
-  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Mixed)')")) +
   scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=expand_scale(mult=c(0, 0.33)), breaks=seq(0, 100, by=10)) +
@@ -159,7 +159,7 @@ fcam <- filter(plot, station=="FCaM") %>%
   geom_text(data=filter(whole, station=="FCaM"), aes(x=date, y=75, label=paste(format(round(abundance, 1), nsmall=1), "%")),
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
-  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Mixed)')")) +
   scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=expand_scale(mult=c(0, 0.33)), breaks=seq(0, 100, by=10)) +
@@ -173,7 +173,7 @@ fca <- filter(plot, station=="FCa") %>%
   geom_text(data=filter(whole, station=="FCa"), aes(x=date, y=75, label=paste(format(round(abundance, 1), nsmall=1), "%")),
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x="Date", y="%") +
-  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Noninvaded)')")) +
+  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Monospecific)')")) +
[…]

### 2020-03-18 code/plot_community_bar_plot_bacteroidota.R

@@ -1,7 +1,7 @@
 #################################################################################################################
 # plot_community_bar_plot_bacteroidota.R
 # 
-# A script to plot sequence abundances of groups from the phylum Bacteoidetes of each sample.
+# A script to plot sequence abundances of groups from the phylum Bacteroidota of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary
 #               data/raw/metadata.csv
 #               data/raw/group_colors.csv
@@ -139,7 +139,7 @@ fcym <- filter(plot, station=="FCyM") %>%
   geom_text(data=filter(whole, station=="FCyM"), aes(x=date, y=75, label=paste(format(round(abundance, 1), nsmall=1), "%")),
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
-  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Mixed)')")) +
   scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=expand_scale(mult=c(0, 0.33)), breaks=seq(0, 100, by=10)) +
@@ -153,7 +153,7 @@ fcam <- filter(plot, station=="FCaM") %>%
   geom_text(data=filter(whole, station=="FCaM"), aes(x=date, y=75, label=paste(format(round(abundance, 1), nsmall=1), "%")),
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
-  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Mixed)')")) +
   scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=expand_scale(mult=c(0, 0.33)), breaks=seq(0, 100, by=10)) +
@@ -167,7 +167,7 @@ fca <- filter(plot, station=="FCa") %>%
[…]

### 2020-03-18 code/plot_community_bar_plot_chloroplast.R

@@ -96,7 +96,7 @@ fcym <- filter(plot, station=="FCyM") %>%
   geom_bar(aes(x=date, y=abundance, fill=taxon), stat="identity", colour="black", size=0.3, width=8) +
   scale_fill_manual(values=color, labels=names) + 
   labs(x=NULL, y="%") +
-  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Mixed)')")) +
   scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=c(0, 0), breaks=seq(0, 100, by=10)) +
@@ -107,7 +107,7 @@ fcam <- filter(plot, station=="FCaM") %>%
   geom_bar(aes(x=date, y=abundance, fill=taxon), stat="identity", colour="black", size=0.3, width=8) +
   scale_fill_manual(values=color, labels=names) + 
   labs(x=NULL, y="%") +
-  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Mixed)')")) +
   scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=c(0, 0), breaks=seq(0, 100, by=10)) +
@@ -118,7 +118,7 @@ fca <- filter(plot, station=="FCa") %>%
   geom_bar(aes(x=date, y=abundance, fill=taxon), stat="identity", colour="black", size=0.3, width=8) +
   scale_fill_manual(values=color, labels=names) + 
   labs(x="Date", y="%") +
-  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Noninvaded)')")) +
+  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Monospecific)')")) +
   scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=c(0, 0), breaks=seq(0, 100, by=10)) +

### 2020-03-18 code/plot_community_bar_plot_cyanobacteria.R

@@ -136,7 +136,7 @@ fcym <- filter(plot, station=="FCyM") %>%
   geom_text(data=filter(whole, station=="FCyM"), aes(x=date, y=75, label=paste(format(round(abundance, 1), nsmall=1), "%")),
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
-  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Mixed)')")) +
   scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=expand_scale(mult=c(0, 0.33)), breaks=seq(0, 100, by=10)) +
@@ -150,7 +150,7 @@ fcam <- filter(plot, station=="FCaM") %>%
   geom_text(data=filter(whole, station=="FCaM"), aes(x=date, y=75, label=paste(format(round(abundance, 1), nsmall=1), "%")),
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
-  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Mixed)')")) +
   scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=expand_scale(mult=c(0, 0.33)), breaks=seq(0, 100, by=10)) +
@@ -164,7 +164,7 @@ fca <- filter(plot, station=="FCa") %>%
   geom_text(data=filter(whole, station=="FCa"), aes(x=date, y=80, label=paste(format(round(abundance, 1), nsmall=1), "%")),
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x="Date", y="%") +
-  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Noninvaded)')")) +
+  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Monospecific)')")) +
[…]

### 2020-03-18 code/plot_community_bar_plot_desulfobacterota.R

@@ -137,7 +137,7 @@ fcym <- filter(plot, station=="FCyM") %>%
   geom_text(data=filter(whole, station=="FCyM"), aes(x=date, y=80, label=paste(format(round(abundance, 1), nsmall=1), "%")),
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
-  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Mixed)')")) +
   scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=expand_scale(mult=c(0, 0.33)), breaks=seq(0, 100, by=10)) +
@@ -151,7 +151,7 @@ fcam <- filter(plot, station=="FCaM") %>%
   geom_text(data=filter(whole, station=="FCaM"), aes(x=date, y=75, label=paste(format(round(abundance, 1), nsmall=1), "%")),
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
-  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Mixed)')")) +
   scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=expand_scale(mult=c(0, 0.33)), breaks=seq(0, 100, by=10)) +
@@ -165,7 +165,7 @@ fca <- filter(plot, station=="FCa") %>%
   geom_text(data=filter(whole, station=="FCa"), aes(x=date, y=75, label=paste(format(round(abundance, 1), nsmall=1), "%")),
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x="Date", y="%") +
-  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Noninvaded)')")) +
+  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Monospecific)')")) +
[…]

### 2020-03-18 code/plot_community_bar_plot_gammaproteobacteria.R

@@ -139,7 +139,7 @@ fcym <- filter(plot, station=="FCyM") %>%
   geom_text(data=filter(whole, station=="FCyM"), aes(x=date, y=75, label=paste(format(round(abundance, 1), nsmall=1), "%")),
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
-  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Mixed)')")) +
   scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=expand_scale(mult=c(0, 0.33)), breaks=seq(0, 100, by=10)) +
@@ -153,7 +153,7 @@ fcam <- filter(plot, station=="FCaM") %>%
   geom_text(data=filter(whole, station=="FCaM"), aes(x=date, y=75, label=paste(format(round(abundance, 1), nsmall=1), "%")),
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
-  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Mixed)')")) +
   scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=expand_scale(mult=c(0, 0.33)), breaks=seq(0, 100, by=10)) +
@@ -167,7 +167,7 @@ fca <- filter(plot, station=="FCa") %>%
   geom_text(data=filter(whole, station=="FCa"), aes(x=date, y=75, label=paste(format(round(abundance, 1), nsmall=1), "%")),
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x="Date", y="%") +
-  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Noninvaded)')")) +
+  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Monospecific)')")) +
[…]

### 2020-03-18 code/plot_matrix.R

@@ -69,9 +69,9 @@ theme <- theme(text=element_text(family="Times"), line=element_line(color="black
 
 # Plot generation
 group_name <- c("F"=expression(bold("Seawater")),
-                "FCyM"=expression(atop(bolditalic("Cymodocea nodosa"), bold("(Invaded)"))),
-                "FCaM"=expression(atop(bolditalic("Caulerpa cylindracea"), bold("(Invaded)"))),
-                "FCa"=expression(atop(bolditalic("Caulerpa cylindracea"), bold("(Noninvaded)"))))
+                "FCyM"=expression(atop(bolditalic("Cymodocea nodosa"), bold("(Mixed)"))),
+                "FCaM"=expression(atop(bolditalic("Caulerpa cylindracea"), bold("(Mixed)"))),
+                "FCa"=expression(atop(bolditalic("Caulerpa cylindracea"), bold("(Monospecific)"))))
 
 jaccard <- ggplot(similarity, aes(x=V2, y=V1, fill=jaccard)) + 
   geom_tile(colour="black", size=0.5) +

### 2020-03-18 code/plot_pcoa.R

@@ -60,9 +60,9 @@ p1 <- ggplot() +
                       values=c("#66A61E", "#E6AB02", "#A6761D", "#666666"),
                       breaks=c("F", "FCyM", "FCaM", "FCa"),
                       labels=c("Seawater",
-                               parse(text="italic('Cymodocea nodosa')~'(Invaded)'"),
-                               parse(text="italic('Caulerpa cylindracea')~'(Invaded)'"),
-                               parse(text="italic('Caulerpa cylindracea')~'(Noninvaded)'"))) +
+                               parse(text="italic('Cymodocea nodosa')~'(Mixed)'"),
+                               parse(text="italic('Caulerpa cylindracea')~'(Mixed)'"),
+                               parse(text="italic('Caulerpa cylindracea')~'(Monospecific)'"))) +
   labs(x=paste0("PCoA I (",format(round(spe.b.pcoa$eig[1]/sum(spe.b.pcoa$eig)*100, digits=2), nsmall=2)," %)"), 
        y=paste0("PCoA II (",format(round(spe.b.pcoa$eig[2]/sum(spe.b.pcoa$eig)*100, digits=2), nsmall=2)," %)")) +
   ggtitle(parse(text="bold('Seawater and Epiphytes')")) +
@@ -91,7 +91,7 @@ p2 <- ggplot() +
                     values=brewer.pal(n=4, name="Set1")) +
   labs(x=paste0("PCoA I (",format(round(spe.b.pcoa$eig[1]/sum(spe.b.pcoa$eig)*100, digits=2), nsmall=2)," %)"), 
        y=paste0("PCoA II (",format(round(spe.b.pcoa$eig[2]/sum(spe.b.pcoa$eig)*100, digits=2), nsmall=2)," %)")) +
-  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Mixed)')")) +
   scale_x_continuous(labels=scaleFUN) +
   scale_y_continuous(labels=scaleFUN) +
   theme
@@ -118,11 +118,11 @@ p3 <- ggplot() +
   scale_shape_manual(name=NULL,
                      values=c(21, 23),
                      breaks=c("FCaM", "FCa"),
-                     labels=c(parse(text="italic('Caulerpa cylindracea')~'(Invaded)'"),
[…]

### 2020-03-18 code/plot_rarefaction.R

@@ -75,7 +75,7 @@ p2 <- filter(metadata_rarefaction, str_detect(label, "FCyM$")) %>%
   scale_colour_manual(values=set_names(color_type$color, color_type$date)) +
   scale_linetype_manual(values=set_names(color_type$line_type, color_type$date)) +
   labs(x="", y="") +
-  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Mixed)')")) +
   theme
 
 p3 <- filter(metadata_rarefaction, str_detect(label, "FCaM$")) %>%
@@ -87,7 +87,7 @@ p3 <- filter(metadata_rarefaction, str_detect(label, "FCaM$")) %>%
   scale_x_continuous(limits=c(0, 40000)) +
   scale_y_continuous(limits=c(0, 4000)) +
   labs(x="Number of Sequences", y="Number of OTUs") +
-  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Mixed)')")) +
   theme
 
 p4 <- filter(metadata_rarefaction, str_detect(label, "FCa$")) %>%
@@ -96,7 +96,7 @@ p4 <- filter(metadata_rarefaction, str_detect(label, "FCa$")) %>%
   geom_line(size=1.5) +
   scale_colour_manual(values=set_names(color_type$color, color_type$date)) +
   scale_linetype_manual(values=set_names(color_type$line_type, color_type$date)) +
-  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Noninvaded)')")) +
+  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Monospecific)')")) +
   labs(x="Number of Sequences", y="") +
   theme
 

### 2020-03-18 code/plot_seasonal_shared.R

@@ -136,7 +136,7 @@ p1 <- filter(data, index=="bray") %>%
   scale_fill_manual(values=fills) +
   scale_y_continuous(limits=c(0, 100)) +
   labs(x="", y="") +
-  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Mixed)')")) +
   theme +
   theme(axis.ticks.x=element_blank(), axis.text.x=element_blank())
 
@@ -155,7 +155,7 @@ p2 <- filter(data, index=="jaccard") %>%
 fcym <- cowplot::plot_grid(p1, p2, nrow=2, ncol=1, rel_heights=c(1, 0.75), align="v")
 
 # Plots generation
-# Caulerpa cylindracea (Invaded) samples
+# Caulerpa cylindracea (Mixed) samples
 data <- filter(distance_metadata, station.x=="FCaM" & station.y=="FCaM") %>%
   filter((date.x=="2017-11-23" & date.y=="2017-12-04") |
          (date.x=="2017-12-04" & date.y=="2017-12-14") |
@@ -178,7 +178,7 @@ p1 <- filter(data, index=="bray") %>%
   scale_fill_manual(values=fills) +
   scale_y_continuous(limits=c(0, 100)) +
   labs(x="", y="%") +
-  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Invaded)')")) +
+  ggtitle(parse(text="bolditalic('Caulerpa cylindracea')~bold('(Mixed)')")) +
   theme +
   theme(axis.ticks.x=element_blank(), axis.text.x=element_blank())
 
@@ -197,7 +197,7 @@ p2 <- filter(data, index=="jaccard") %>%
 fcam <- cowplot::plot_grid(p1, p2, nrow=2, ncol=1, rel_heights=c(1, 0.75), align="v")
 
 # Plots generation
-# Caulerpa cylindracea (Noninvaded) samples
+# Caulerpa cylindracea (Monospecific) samples
 data <- filter(distance_metadata, station.x=="FCa" & station.y=="FCa") %>%
   filter((date.x=="2017-11-23" & date.y=="2017-12-04") |
          (date.x=="2017-12-04" & date.y=="2017-12-14") |
@@ -220,7 +220,7 @@ p1 <- filter(data, index=="bray") %>%
[…]

### 2020-03-19 submission/manuscript.tex
Pipeline optimization.

@@
 were present throughout the year. In every analyzed high rank taxa,
-phylogenetic groups present throughout the year and comprising most of
-the sequences could be identified together with low proportion taxa
+phylogenetic groups present throughout the year comprised most of the
+sequences and could be identified together with low proportion taxa
 showing seasonal patterns connected to the macrophyte growth cycle.
@@
 al.}, 2019). While there is high community variation between host
-species is was observed that the majority of metagenome determined
+species it was observed that the majority of metagenome determined
 functions were conserved both between host species and individuals
@@
 were taken into account low temporal frequency and/or methodologies that
-do not allow for high taxonomic resolution were used (Tujula \emph{et
-al.}, 2010; Lachnit \emph{et al.}, 2011; Miranda \emph{et al.}, 2013;
-Michelou \emph{et al.}, 2013). In the present study we describe the
-seasonal dynamics of bacterial and archaeal communities on the surfaces
-of the seagrass \emph{Cymodocea nodosa} and siphonous macroalgae
-\emph{Caulerpa cylindracea} determined on a mostly monthly scale.
-Bacterial and archaeal epiphytes were sampled in a meadow of \emph{C.
-nodosa} invaded by the invasive \emph{C. cylindracea} and in a locality
-of only \emph{C. cylindracea} located in the proximity of the meadow. In
-addition, for comparison, the community of the ambient seawater was
-characterized.
+do not allow high taxonomic resolution were used (Tujula \emph{et al.},
+2010; Lachnit \emph{et al.}, 2011; Miranda \emph{et al.}, 2013; Michelou
+\emph{et al.}, 2013). In the present study we describe the seasonal
+dynamics of bacterial and archaeal communities on the surfaces of the
+seagrass \emph{Cymodocea nodosa} and siphonous macroalgae \emph{Caulerpa
+cylindracea} determined on a mostly monthly scale. Bacterial and
+archaeal epiphytes were sampled in a meadow of \emph{C. nodosa} invaded
+by the invasive \emph{C. cylindracea} and in a locality of only \emph{C.
+cylindracea} located in the proximity of the meadow. In addition, for
+comparison, the community of the ambient seawater was characterized.
 
@@
 in the ambient seawater were also processed for comparison. The number
-of reads per sample ranged between 8,408 and 77,465 sequences
+of reads per sample ranged between 8,409 and 77,465 sequences
 (\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
@@
 (\autoref{supp-rarefaction}). Following quality curation and exclusion
-of sequences mentioned before reads were clustered into 28,736 different
-OTUs at a similarity level of 97 \si{\percent}. Reads numbers were
-normalized to the minimum number of sequences, 8,408
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,114
-different OTUs that ranged from 379 to 2,016 OTUs per sample
+of sequences mentioned before reads were clustered into 28,691 different
+OTUs at a similarity level of 97 \si{\percent}. Read numbers were
+normalized to the minimum number of sequences, 8,409
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,071
+different OTUs that ranged from 386 to 2,009 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (invaded {[}Number of OTUs, 1,676.9 ±
-140.3 OTUs{]} and noninvaded {[}Number of OTUs, 1,745.9 ± 156.3
-OTUs{]}), middle vales for \emph{C. nodosa} (Number of OTUs, 1,059.5 ±
-212.3 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 530.8 ± 143.1 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,679.5 ±
+127.4 OTUs{]} and monospecific {[}Number of OTUs, 1,736.3 ± 156.5
+OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,059 ±
+216.2 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 528.2 ± 135.4 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
 dissimilarities. \emph{C. nodosa} communities showed a slow increase
-towards the end of the study, while \emph{C. cylindracea} (invaded and
-noninvaded) communities were characterized by slightly larger values in
-Spring and Summer in comparison to Autumn and Winter
+towards the end of the study, while \emph{C. cylindracea} (mixed and
+monospecific) communities were characterized by slightly larger values
+in Spring and Summer in comparison to Autumn and Winter
 (\autoref{supp-calculators}).
@@
 particular environment. The highest proportion of shared OTUs and
-community was found between invaded and noninvaded \emph{C. cylindracea}
+community was found between mixed and monospecific \emph{C. cylindracea}
 (Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared values were
@@
 from each surface associated community were found at every time point.
-These OTUs made also a high proportion of total sequences (40.3 -- 52.4
+These OTUs also made a high proportion of total sequences (39.9 -- 51.8
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 communities indicated a separation between Spring, Summer and
-Autumn/Winter samples (ANOSIM, R = 0.55, \emph{p} \textless{} 0.001),
-while communities from the surfaces of \emph{C. cylindracea} indicated a
-non so strongly supported, as in previous cases, separation between
-Summer and Autumn/Winter/Spring samples (ANOSIM, R = 0.31, \emph{p}
-\textless{} 0.01) (\autoref{pcoa}).
+Autumn/Winter samples (ANOSIM, R = 0.53, \emph{p} \textless{} 0.01). For
+\emph{C. cylindracea} associated communities a separation between Summer
+and Autumn/Winter/Spring samples was observed that was not so strongly
+supported (ANOSIM, R = 0.30, \emph{p} \textless{} 0.05)
+(\autoref{pcoa}).
 
@@
 \emph{Gammaproteobacteria} and \emph{Verrucomicrobiota}. Communities
-associated with \emph{C. nodosa} were consisted of same groups with the
+associated with \emph{C. nodosa} consisted of same groups with the
 addition of \emph{Planctomycetota} whose contribution was higher in
-summer 2018. In addition, communities from invaded and noninvaded
+summer 2018. In addition, communities from mixed and monospecific
 \emph{C. cylindracea} were similar and characterized by same groups as
@@
 proportions were found for \emph{C. nodosa} (16.4 ± 5.3 \si{\percent})
-and \emph{C. cylindracea} (invaded {[} (7.7 ± 3.9 \si{\percent}){]} and
-noninvaded {[} (7.8 ± 2.4 \si{\percent}){]}) associated communities in
-autumn and for seawater communities in winter (8.8 ± 7.5 \si{\percent}).
+and \emph{C. cylindracea} (mixed {[} (7.7 ± 3.9 \si{\percent}){]} and
+monospecific {[} (7.8 ± 2.4 \si{\percent}){]}) associated communities in
+autumn and for seawater communities in winter (8.8 ± 7.4 \si{\percent}).
 Large taxonomic differences between surface associated and seawater
@@
 \emph{Bacteroidota} communities were very diverse as could be observed
-in the the high proportion of taxa that grouped as other
-\emph{Bacteroidota} (\autoref{bactero}).
+in the high proportion of taxa that grouped as other \emph{Bacteroidota}
+(\autoref{bactero}).
 
@@
 were mainly composed of no relative \emph{Gammaproteobacteria} and
-\emph{Granulosicocus}. Beside these two groups specific to all three
+\emph{Granulosicoccus}. Beside these two groups specific to all three
 epiphytic communities, \emph{C. nodosa} was characterized by
@@
 \emph{Cellvibrionaceae} and \emph{Reinekea} were more specific to both
-invaded and noninvaded \emph{C. cylindracea}. In addition,
+mixed and monospecific \emph{C. cylindracea}. In addition,
 \emph{Arenicella} was more pronounced in November and December 2017,
@@
 \emph{C. cylindracea} specific taxa no relative \emph{Cellvibrionaceae}
-and \emph{Reinekea} showed some seasonality and were characterisitic for
+and \emph{Reinekea} showed some seasonality and were characteristic for
 samples originating from June to October 2018. In addition, similarly to
@@
 \emph{Desulfobacterota} were specific to \emph{C. cylindracea}. On
-average they were comprising 11.2 ± 13.3 \si{\percent} of all sequences.
-While seawater and \emph{C. nodosa} communities were consisted of only
-0.1 ± 0.08 \si{\percent} and 1.0 ± 0.7 \si{\percent}
-\emph{Desulfobacterota} sequences, respectively, in the invaded and
-noninvaded \emph{C. cylindracea} communities their proportion was 25.7 ±
-11.2 \si{\percent} and 24.0 ± 4.3 \si{\percent}, respectively
-(\autoref{desulfo}). The community was mainly consisted of no relative
-\emph{Desulfobacteraceae}, \emph{Desulfatitalea}, no relative
-\emph{Desulfobulbaceae}, \emph{Desulfobulbus}, no relative
-\emph{Desulfocapsaceae}, \emph{Desulfopila}, \emph{Desulforhopalus},
-\emph{Desulfotalea}, SEEP-SRB4 and uncultured \emph{Desulfocapsaceae}
-(\autoref{desulfo}).
+average they comprised 11.2 ± 13.3 \si{\percent} of all sequences.
+Seawater and \emph{C. nodosa} communities consisted of only 0.1 ± 0.08
+\si{\percent} and 1.0 ± 0.7 \si{\percent} \emph{Desulfobacterota}
+sequences, respectively. In the mixed and monospecific \emph{C.
+cylindracea} communities their proportion was 25.7 ± 11.2 \si{\percent}
+and 24.0 ± 4.3 \si{\percent}, respectively (\autoref{desulfo}). The
+community consisted mainly of no relative \emph{Desulfobacteraceae},
+\emph{Desulfatitalea}, no relative \emph{Desulfobulbaceae},
+\emph{Desulfobulbus}, no relative \emph{Desulfocapsaceae},
+\emph{Desulfopila}, \emph{Desulforhopalus}, \emph{Desulfotalea},
+SEEP-SRB4 and uncultured \emph{Desulfocapsaceae} (\autoref{desulfo}).
 
@@
 
-Surfaces of marine macrophytes are harboring biofilms consisted of
-diverse microbial taxa (Egan \emph{et al.}, 2013; Tarquinio \emph{et
-al.}, 2019). No standard protocol has been developed to study these
-macophyte associated microbes (Ugarelli \emph{et al.}, 2019). Different
-procedures of microbial cells removal from host surfaces were described,
-such as host tissue shaking (Nõges \emph{et al.}, 2010), scraping (Uku
+Surfaces of marine macrophytes harbor biofilms consisting of diverse
+microbial taxa (Egan \emph{et al.}, 2013; Tarquinio \emph{et al.},
+2019). No standard protocol has been developed to study these macophyte
+associated microbes (Ugarelli \emph{et al.}, 2019). Different procedures
+for removal of microbial cells from host surfaces were described, such
+as host tissue shaking (Nõges \emph{et al.}, 2010), scraping (Uku
 \emph{et al.}, 2007) and ultrasonication (Cai \emph{et al.}, 2014). All
@@
 In the present study, highest richness values were observed for \emph{C.
-cylindracea} (invaded and noninvaded), middle for \emph{C. nodosa} and
+cylindracea} (mixed and monospecific), middle for \emph{C. nodosa} and
 lowest for seawater derived communities. Higher values for seagrass
@@
 existing on macrophyte surfaces (Ugarelli \emph{et al.}, 2019). In
-addition, highest values observed for \emph{C. cylindracea} are probably
-a consequence of part of epiphytic OTUs that are in contact with the
-sediment. \emph{C. cylindracea} stolon is attached to the sediment
-surface with rhizoids, so the stolon and rhizoids are in a direct
-contact with the sediment. Part of the surface attached \emph{Caulerpa
-cylnidracea} community is therefore comprised of OTUs that are in
-contact with a different environment and that could cause the observed
-increase in richness. In addition, seasonal richness differences
-observed for surface attached communities showed slightly higher values
-in spring and summer. This pattern could be explained by a higher
-macrophyte growth in these seasons (M. Najdek, personal communication;
-Zavodnik \emph{et al.}, 1998; Ruitton \emph{et al.}, 2005). During
-active periods macrophytes exhibit a more dynamic chemical interaction
-with the surface community probably causing an increase in the number of
-inhabitable microniches (Borges and Champenois, 2015; Rickert \emph{et
-al.}, 2016).
+addition, highest values observed for \emph{C. cylindracea} are partly
+due to its contact with the sediment. \emph{C. cylindracea} stolon is
+attached to the sediment surface with rhizoids, so the stolon and
+rhizoids are in a direct contact with the sediment. In addition,
+seasonal richness differences observed for surface attached communities
+showed slightly higher values in spring and summer. This pattern could
+be explained by a higher macrophyte growth in these seasons (M. Najdek,
+personal communication; Zavodnik \emph{et al.}, 1998; Ruitton \emph{et
+al.}, 2005). During active periods macrophytes exhibit a more dynamic
+chemical interaction with the surface community probably causing an
+increase in the number of inhabitable microniches (Borges and
+Champenois, 2015; Rickert \emph{et al.}, 2016).
 
 Since the colonization of macrophyte surfaces is performed from a pool
-of seawater prokaryotic cells it is interesting to see to which extent
-do these two communities differ. We observed a strong differentiation
-between the surface attached and seawater communities at the level of
-OTUs that is in agreement with most published studies (Burke \emph{et
-al.}, 2011b; Michelou \emph{et al.}, 2013; Roth-Schulze \emph{et al.},
-2016; Crump \emph{et al.}, 2018; Ugarelli \emph{et al.}, 2019). These
-data indicate that marine macrophytes are selecting, from a pool of
-seawater microbial taxa, the one that can colonize and proliferate on
-their surfaces (Salaün \emph{et al.}, 2012; Michelou \emph{et al.},
-2013). In contrast to these findings Fahimipour \emph{et al.} (2017)
-found, in a global study of \emph{Zostera marina}, similarities between
-leaves and seawater samples. Discrepancies between our data and this
-study could be explained by differences in studied seagrass species,
-methodological variations or biogeographic trends as Fahimipour \emph{et
-al.} (2017) were analyzing samples from different locations throughout
-the northern hemisphere. It is possible that ambient seawater and leaves
-communities from the same location are differing but are still more
-similar to each other when compared to other sampling locations. Indeed,
-it was found that prokaryotic communities vary substantially between
-different sampling sites (Bengtsson \emph{et al.}, 2017). When the
-taxonomic composition at high ranks was analyzed no such strong
-differentiation was noticed. Phyla and classes such as:
-\emph{Actinobacteriota}, \emph{Bacteroidota}, \emph{Cyanobacteria},
-\emph{Alphaproteobacteria}, \emph{Gammaproteobacteria} and
-\emph{Verrucomicrobiota}, were described that is in agreement with
-previously reported data (Burke \emph{et al.}, 2011b; Egan \emph{et
-al.}, 2013; Michelou \emph{et al.}, 2013). In contrast, when low
-taxonomic ranks were analyzed (i.g. family and genus) a strong
-differentiation was observed. A similar differentiation at lower
-taxonomic ranks was described for other species of macrophytes (Egan
-\emph{et al.}, 2013; Michelou \emph{et al.}, 2013; Ugarelli \emph{et
-al.}, 2019).
+of prokaryotic cells from the ambient seawater, it was interesting to
+see to which extent these two communities differ. We observed a strong
+differentiation between the surface attached and seawater communities at
+the level of OTUs that is in agreement with most published studies
+(Burke \emph{et al.}, 2011b; Michelou \emph{et al.}, 2013; Roth-Schulze
+\emph{et al.}, 2016; Crump \emph{et al.}, 2018; Ugarelli \emph{et al.},
+2019). These data indicate that marine macrophytes are selecting, from a
+pool of seawater microbial taxa, the one that can colonize and
+proliferate on their surfaces (Salaün \emph{et al.}, 2012; Michelou
+\emph{et al.}, 2013). In contrast to these findings Fahimipour \emph{et
+al.} (2017) found, in a global study of \emph{Zostera marina},
+similarities between leaves and seawater samples. Discrepancies between
+our data and this study could be explained by differences in studied
+seagrass species, methodological variations or biogeographic trends as
+Fahimipour \emph{et al.} (2017) were analyzing samples from different
+locations throughout the northern hemisphere. It is possible that
+ambient seawater and leaves communities from the same location are
+differing but are still more similar to each other when compared to
+other sampling locations. Indeed, it was found that prokaryotic
+communities vary substantially between different sampling sites
+(Bengtsson \emph{et al.}, 2017). When the taxonomic composition at high
+ranks was analyzed no such strong differentiation was noticed. Phyla and
+classes such as: \emph{Actinobacteriota}, \emph{Bacteroidota},
+\emph{Cyanobacteria}, \emph{Alphaproteobacteria},
+\emph{Gammaproteobacteria} and \emph{Verrucomicrobiota} were described
+that is in agreement with previously reported data (Burke \emph{et al.},
+2011b; Egan \emph{et al.}, 2013; Michelou \emph{et al.}, 2013). In
+contrast, when low taxonomic ranks were analyzed (i.g. family and genus)
+a strong differentiation was observed. A similar differentiation at
+lower taxonomic ranks was described for other species of macrophytes
+(Egan \emph{et al.}, 2013; Michelou \emph{et al.}, 2013; Ugarelli
+\emph{et al.}, 2019).
 
@@
 associated with different host was observed. The only high rank phylum
-that was differing between \emph{C. nodosa} ans \emph{C. cylindracea}
+that was differing between \emph{C. nodosa} and \emph{C. cylindracea}
 was \emph{Desulfobacterota}, whose sequences were more abundant in the
@@
 \emph{Desulfobacterota} are probably a part of the epiphytic community
-that is in contact with the sediment. Similar high rank taxa found in
+that was in contact with the sediment. Similar high rank taxa found in
 this study were described to be specific for other species of
@@
 Seasonal richness changes in the epiphytic community were substantial as
-could be observed in the proportion of OTUs that could be find at every
+could be observed in the proportion of OTUs that could be found at every
 sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 52.4
+were accounting for a high proportion of sequences (\(\leq\) 51.8
 \si{\percent}). A very similar proportion of persistent OTUs and their
@@
 comparison to the seawater community, a lower degree of seasonal shifts
-was observed for the surface associated communities. It seems, that
+was observed for the surface associated communities. It seems,
 microniches on the surfaces of macrophytes are providing more stable
@@
 seagrass proliferation, a Summer community during a biomass maximum and
-a Autumn/Winter community during a biomass decay. In contrast, \emph{C.
-cylindracea} started to proliferate in late Spring and was characterized
-only by a Summer community during maximal biomass increase and by a
-Autumn/Winter/Spring community when the biomass were at the peak and the
-settlement started to subsequently decay. Similar seasonal changes in
-the epiphytic community was described also for other macroalgae (Tujula
-\emph{et al.}, 2010; Lachnit \emph{et al.}, 2011). Higher temporal
-stability of \emph{C. cylindracea} surface communities in comparison to
-\emph{C. nodosa} were also observed in the higher proportion of shared
-communities between two consecutive sampling points.
+a Autumn/Winter community during a biomass senescence. In contrast,
+\emph{C. cylindracea} started to proliferate in late Spring and was
+characterized only by a Summer community during maximal biomass increase
+and by a Autumn/Winter/Spring community when the biomass was at the peak
+and the settlement started to subsequently decay. Similar seasonal
+changes in the epiphytic community was also described for other
+macroalgae (Tujula \emph{et al.}, 2010; Lachnit \emph{et al.}, 2011).
+Higher temporal stability of \emph{C. cylindracea} surface communities
+in comparison to \emph{C. nodosa} were also observed in the higher
+proportion of shared communities between two consecutive sampling
+points.
 
@@
 values in Autumn/Winter in comparison to Spring/Summer. This pattern is
-not surprising as e.g seagrasses are known to harbor more epiphytes
-during Autumn/Winter (Reyes and Sansón, 2001) and we used an adapted DNA
-isolation protocol that is known to partially coextract DNA from
+not surprising as seagrasses are known to harbor more epiphytes during
+Autumn/Winter (Reyes and Sansón, 2001). Furthermore, we used an adapted
+DNA isolation protocol that is known to partially coextract DNA from
 planktonic eukaryotes (Korlević \emph{et al.}, 2015). Strong seasonal
@@
 \emph{et al.}, 2004; Longford \emph{et al.}, 2007; Mobberley \emph{et
-al.}, 2012; Reisser \emph{et al.}, 2014). It is possible than during
-periods of low metabolic activity there is no so active interaction and
+al.}, 2012; Reisser \emph{et al.}, 2014). It is possible that during
+periods of low metabolic activity there is a reduced interaction and
 selection of the epiphytic community by the seagrass, causing leaves to
@@
 \emph{Phormidesmis} sequences were at their peak in Autumn (Cardoso
-\emph{et al.}, 2019). Other high rank taxa did not showed strong
-successional patterns, but in every analyzed group, with the exception
-of \emph{Desulfobacterota}, taxa present throughout the year in similar
+\emph{et al.}, 2019). Other high rank taxa did not show strong
+successional patterns. In every analyzed group, with the exception of
+\emph{Desulfobacterota}, taxa present throughout the year in similar
 proportions and season specific taxa could be identified. Within
@@
 family are often found in association with macrophytes and it is
-suggested that are involved in the hydrolysis and utilization of complex
-carbon sources (Burke \emph{et al.}, 2011b; McIlroy and Nielsen, 2014;
-Crump \emph{et al.}, 2018). On the other hand, families
+suggested that they are involved in the hydrolysis and utilization of
+complex carbon sources (Burke \emph{et al.}, 2011b; McIlroy and Nielsen,
+2014; Crump \emph{et al.}, 2018). On the other hand, families
 \emph{Flavobacteriaceae} and \emph{Rhodothermaceae} showed seasonal
@@
 \emph{Alphaproteobacteria} the family \emph{Rhodobacteraceae} was
-comprising the majority of sequences thought the year. This
+comprising the majority of sequences throughout the year. This
 metabolically versatile family is often associated with macrophyte
@@
 al.}, 2013; Crump \emph{et al.}, 2018). Beside these two groups other
-less pronounced taxa showed seasonal and host-specific patterns. In
+less pronounced taxa showed seasonal and host-specific patterns. For
 example, \emph{C. cylindracea} was characterized by \emph{Thioploca}, a
 known sulfur sediment bacteria and \emph{Cellvibrionaceae}, a family
-whose cultured members are known polysaccharides degraders (Jørgensen
-and Gallardo, 1999; Xie \emph{et al.}, 2017). \emph{Desulfobacterota}
-were found only associated with \emph{C. cylindracea} and no group
-within this phylum showed seasonal patterns. The presence of this phylum
-only on \emph{C. cylindracea} is to be expected as part of the epiphytic
-community is in a direct contact with the sediment. The
+whose cultured members are known polysaccharide degraders (Jørgensen and
+Gallardo, 1999; Xie \emph{et al.}, 2017). \emph{Desulfobacterota} were
+found only associated with \emph{C. cylindracea} and no group within
+this phylum showed seasonal patterns. The presence of this phylum only
+on \emph{C. cylindracea} is to be expected as part of the epiphytic
+community is directly in contact with the sediment. The
 \emph{Desulfobacterota} community was dominated by \emph{Desulfatitalea}
@@
 comparisons between communities associated with different hosts showed
-that the majortiy of functions could be found in every community,
-indicating functional redundacy (Roth-Schulze \emph{et al.}, 2016). This
-difference between taxonomic and functionl discrepancy was expalined by
-the lottery hypothesis that hypothesize an initial random colonization
-step performed from a set of functionally equivalent taxonomic groups
-(Burke \emph{et al.}, 2011a; Roth-Schulze \emph{et al.}, 2016). It is
-possible that functional redundancy is a characteristic of high
-abundance taxa detected to be present throughout the year, while
-seasonal and/or host-specific functions are an attribute of taxa
-displaying successional patterns. Further studies connecting taxonomy
-with functional properties will be required to elucidate the degree of
-functional redundancy or specificity in epiphyitc microbial communites.
+that the majority of functions could be found in every community,
+indicating functional redundancy (Roth-Schulze \emph{et al.}, 2016).
+This difference between taxonomic and functional discrepancy was
+explained by the lottery hypothesis that hypothesizes an initial random
+colonization step performed from a set of functionally equivalent
+taxonomic groups (Burke \emph{et al.}, 2011a; Roth-Schulze \emph{et
+al.}, 2016). It is possible that functional redundancy is a
+characteristic of high abundance taxa detected to be present throughout
+the year, while seasonal and/or host-specific functions are an attribute
+of taxa displaying successional patterns. Further studies connecting
+taxonomy with functional properties will be required to elucidate the
+degree of functional redundancy or specificity in epiphytic microbial
+communities.
 
@@
 
-Leaves of \emph{C. nodosa} were sampled in a \emph{C. nodosa} meadow
-located in the proximity of the village of Funtana (\ang{45;10;39} N,
-\ang{13;35;42} E). Thalli of \emph{C. cylindracea} were sampled in the
-same \emph{C. nodosa} invaded meadow in Funtana and on a locality of
-only \emph{C. cylindracea} located close to the invaded meadow. Sampling
-of leaves and thalli was performed approximately monthly from December
-2017 to October 2018 (\autoref{supp-nseq_notus}). Leaves and thalli were
+Sampling was performed in the Bay of Funtana, northern Adriatic Sea
+(\ang{45;10;39} N, \ang{13;35;42} E). Leaves and thalli were sampled in
+a meadow of \emph{Cymodocea nodosa} invaded by the invasive
+\emph{Caulerpa cylindracea} (mixed settlement) and in a monospecific
+settlement of \emph{Caulerpa cylindracea} located in the proximity of
+the meadow at approximately monthly intervals from December 2017 to
+October 2018 (\autoref{supp-nseq_notus}). Leaves and thalli were
 collected by diving and transported to the laboratory in containers
@@
 using isopropanol. DNA from seawater picoplankton was isolated from 0.2
-\si{\um} polycarbonate filters according to (Massana \emph{et al.},
-1997) with a slight modification. Following the phenol-chloroform
-extraction steps 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2)
-was added. DNA was precipitated by adding 1 volume of chilled
-isopropanol, incubating the mixtures overnight at \num{-20}
-\si{\degreeCelsius} and centrifuging at 20,000 × g and 4
-\si{\degreeCelsius} for 21 \si{\minute}. The pellet was washed twice
-with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged
-after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 5
-\si{\minute}. Dried pellets were resuspended in 50 -- 100 \si{\ul} of
-deionized water.
+\si{\um} polycarbonate filters according to Massana \emph{et al.} (1997)
+with a slight modification. Following the phenol-chloroform extraction
+steps 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added.
+DNA was precipitated by adding 1 volume of chilled isopropanol,
+incubating the mixtures overnight at \num{-20} \si{\degreeCelsius} and
+centrifuging at 20,000 × g and 4 \si{\degreeCelsius} for 21
+\si{\minute}. The pellet was washed twice with 500 \si{\ul} of chilled
+70 \si{\percent} ethanol and centrifuged after each washing step at
+20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets
+were resuspended in 50 -- 100 \si{\ul} of deionized water.
 
@@
 \leavevmode\hypertarget{ref-Abraham2014}{}%
-Abraham, W.R. and Rohde, M. (2014) The Family Hyphomonadaceae. In,
-\emph{The prokaryotes: Alphaproteobacteria and betaproteobacteria}.
+Abraham, W.R. and Rohde, M. (2014) The Family
+\textless{}i\textgreater{}Hyphomonadaceae\textless{}i/\textgreater{}.
+In, \emph{The prokaryotes:
+\textless{}i\textgreater{}Alphaproteobacteria\textless{}i/\textgreater{}
+and
+\textless{}i\textgreater{}Betaproteobacteria\textless{}i/\textgreater{}}.
 Springer-Verlag Berlin Heidelberg, pp. 283--299.
@@
 Bengtsson, M., Sjøtun, K., and Øvreås, L. (2010) Seasonal dynamics of
-bacterial biofilms on the kelp Laminaria hyperborea. \emph{Aquatic
-Microbial Ecology} \textbf{60}: 71--83.
+bacterial biofilms on the kelp \textless{}i\textgreater{}Laminaria
+hyperborea\textless{}i/\textgreater{}. \emph{Aquatic Microbial Ecology}
+\textbf{60}: 71--83.
 
@@
 of dimethylsulfoniopropionate (DMSP) in the Mediterranean seagrass
-Posidonia oceanica. \emph{Aquatic Botany} \textbf{125}: 72--79.
+\textless{}i\textgreater{}Posidonia oceanica\textless{}i/\textgreater{}.
+\emph{Aquatic Botany} \textbf{125}: 72--79.
 
 \leavevmode\hypertarget{ref-Burke2009}{}%
-Burke, C., Kjelleberg, S., and Thomas, T. (2009) Selective extraction of
-bacterial DNA from the surfaces of macroalgae. \emph{Applied and
-environmental microbiology} \textbf{75}: 252--256.
+Burke, C., Kjelleberg, S., and Thomas, T. (2009) Selective Extraction of
+Bacterial DNA from the Surfaces of Macroalgae. \emph{Applied and
+Environmental Microbiology} \textbf{75}: 252--256.
 
@@
 (2011b) Composition, uniqueness and variability of the epiphytic
-bacterial community of the green alga Ulva australis. \emph{The ISME
-journal} \textbf{5}: 590--600.
+bacterial community of the green alga \textless{}i\textgreater{}Ulva
+australis\textless{}i/\textgreater{}. \emph{The ISME Journal}
+\textbf{5}: 590--600.
 
@@
 (2014) An ultrasonic method for separation of epiphytic microbes from
-freshwater submerged macrophytes. \emph{Journal of basic microbiology}
+freshwater submerged macrophytes. \emph{Journal of Basic Microbiology}
 \textbf{54}: 758--761.
@@
 Seasonal development of a coastal microbial mat. \emph{Scientific
-Reports} \textbf{9}: 1--14.
+Reports} \textbf{9}: 9035.
 
 \leavevmode\hypertarget{ref-Crump2008}{}%
-Crump, B.C. and Koch, E.W. (2008) Attached bacterial populations shared
-by four species of aquatic angiosperms. \emph{Applied and environmental
-microbiology} \textbf{74}: 5948--5957.
+Crump, B.C. and Koch, E.W. (2008) Attached Bacterial Populations Shared
+by Four Species of Aquatic Angiosperms. \emph{Applied and Environmental
+Microbiology} \textbf{74}: 5948--5957.
 
@@
 Crump, B.C., Wojahn, J.M., Tomas, F., and Mueller, R.S. (2018)
-Metatranscriptomics and amplicon sequencing reveal mutualisms in
-seagrass microbiomes. \emph{Frontiers in Microbiology} \textbf{9}:
+Metatranscriptomics and Amplicon Sequencing Reveal Mutualisms in
+Seagrass Microbiomes. \emph{Frontiers in Microbiology} \textbf{9}: 388.
 
@@
 Thomas, T. (2013) The seaweed holobiont: understanding seaweed-bacteria
-interactions. \emph{FEMS microbiology reviews} \textbf{37}: 462--476.
+interactions. \emph{FEMS Microbiology Reviews} \textbf{37}: 462--476.
 
@@
 Fahimipour, A.K., Kardish, M.R., Lang, J.M., Green, J.L., Eisen, J.A.,
-and Stachowicz, J.J. (2017) Globalscale structure of the eelgrass
-microbiome. \emph{Applied and Environmental Microbiology} \textbf{83}:
+and Stachowicz, J.J. (2017) Global-Scale Structure of the Eelgrass
+Microbiome. \emph{Applied and Environmental Microbiology} \textbf{83}:
+e03391--16.
 
@@
 Higashioka, Y., Kojima, H., Watanabe, T., and Fukui, M. (2015) Draft
-genome sequence of Desulfatitalea tepidiphila S28bFT. \emph{Genome
-Announcements} \textbf{3}:
+genome sequence of \textless{}i\textgreater{}Desulfatitalea
+tepidiphila\textless{}/i\textgreater{} S28bF\(^{T}\). \emph{Genome
+Announcements} \textbf{3}: e01326--15.
 
@@
 Hollants, J., Leliaert, F., De Clerck, O., and Willems, A. (2013) What
-we can learn from sushi: A review on seaweed-bacterial associations.
+we can learn from sushi: a review on seaweed-bacterial associations.
 \textbf{83}: 1--16.
@@
 \leavevmode\hypertarget{ref-Jost2006}{}%
-Jost, L. (2006) Entropy and diversity. \textbf{113}: 363--375.
+Jost, L. (2006) Entropy and diversity. \emph{Oikos} \textbf{113}:
+363--375.
 
 \leavevmode\hypertarget{ref-Jorgensen1999}{}%
-Jørgensen, B.B. and Gallardo, V.A. (1999) Thioploca spp.: filamentous
-sulfur bacteria with nitrate vacuoles. \emph{FEMS Microbiology Ecology}
-\textbf{28}: 301--313.
+Jørgensen, B.B. and Gallardo, V.A. (1999)
+\textless{}i\textgreater{}Thioploca\textless{}/i\textgreater{} spp.:
+filamentous sulfur bacteria with nitrate vacuoles. \emph{FEMS
+Microbiology Ecology} \textbf{28}: 301--313.
 
@@
 Kozich, J.J., Westcott, S.L., Baxter, N.T., Highlander, S.K., and
-Schloss, P.D. (2013) Development of a dual-index sequencing strategy and
-curation pipeline for analyzing amplicon sequence data on the MiSeq
-Illumina sequencing platform. \emph{Applied and environmental
+Schloss, P.D. (2013) Development of a Dual-Index Sequencing Strategy and
+Curation Pipeline for Analyzing Amplicon Sequence Data on the MiSeq
+Illumina Sequencing Platform. \emph{Applied and environmental
 microbiology} \textbf{79}: 5112--5120.
@@
 Kuever, J. (2014) The Family Desulfobulbaceae. In, \emph{The
-prokaryotes: Deltaproteobacteria and epsilonproteobacteria}.
+prokaryotes:
+\textless{}i\textgreater{}Deltaproteobacteria\textless{}/i\textgreater{}
+and
+\textless{}i\textgreater{}Epsilonproteobacteria\textless{}/i\textgreater{}}.
 Springer-Verlag Berlin Heidelberg, pp. 75--86.
@@
 Lachnit, T., Blümel, M., Imhoff, J.F., and Wahl, M. (2009) Specific
-epibacterial communities on macroalgae: Phylogeny matters more than
+epibacterial communities on macroalgae: phylogeny matters more than
 habitat. \emph{Aquatic Biology} \textbf{5}: 181--186.
@@
 \leavevmode\hypertarget{ref-Margulis1991}{}%
-Margulis, L. (1991) Symbiogenesis and symbionticism. In, Margulis,L. and
-Fester,R. (eds), \emph{Symbiosis as a source of evolutionary
-innovation}. Cambridge, Massachusetts: The MIT Press, pp. 1--14.
+Margulis, L. (1991) Symbiogenesis and Symbionticism. In, Margulis,L. and
+Fester,R. (eds), \emph{Symbiosis as a source of evolutionary innovation:
+Speciation and morphogenesis}. Cambridge, Massachusetts: The MIT Press,
+pp. 1--14.
 
@@
 Massana, R., Murray, A.E., Preston, C.M., and DeLong, E.F. (1997)
-Vertical distribution and phylogenetic characterization of marine
-planktonic Archaea in the Santa Barbara Channel. \emph{Applied and
-environmental microbiology} \textbf{63}: 50--56.
+Vertical Distribution and Phylogenetic Characterization of Marine
+Planktonic \textless{}i\textgreater{}Archaea\textless{}/i\textgreater{}
+in the Santa Barbara Channel. \emph{Applied and environmental
+microbiology} \textbf{63}: 50--56.
 
 \leavevmode\hypertarget{ref-McIlroy2014}{}%
-McIlroy, S.J. and Nielsen, P.H. (2014) The Family Saprospiraceae. In,
-\emph{The prokaryotes: Other major lineages of bacteria and the
-archaea}. Springer-Verlag Berlin Heidelberg, pp. 863--889.
+McIlroy, S.J. and Nielsen, P.H. (2014) The Family
+\textless{}i\textgreater{}Saprospiraceae\textless{}/i\textgreater{}. In,
+\emph{The prokaryotes: Other major lineages of
+\textless{}i\textgreater{}Bacteria\textless{}/i\textgreater{} and the
+\textless{}i\textgreater{}Archaea\textless{}/i\textgreater{}}.
+Springer-Verlag Berlin Heidelberg, pp. 863--889.
 
@@
 Ecology of Microbial Communities Associated with Macrocystis pyrifera.
-\emph{PLoS ONE} \textbf{8}: e67480.
+\emph{PloS one} \textbf{8}: e67480.
 
@@
 Porphyra umbilicalis: Did Bacterial Farmers Produce Macroalgae?
-\emph{PLoS ONE} \textbf{8}: e58269.
+\emph{PloS one} \textbf{8}: e58269.
 
@@
 Morrissey, K.L., Çavas, L., Willems, A., and De Clerck, O. (2019)
-Disentangling the influence of environment, host specificity and thallus
-differentiation on bacterial communities in siphonous green seaweeds.
-\emph{Frontiers in Microbiology} \textbf{10}:
+Disentangling the Influence of Environment, Host Specificity and Thallus
+Differentiation on Bacterial Communities in Siphonous Green Seaweeds.
+\emph{Frontiers in Microbiology} \textbf{10}: 717.
 
@@
 Pujalte, M.J., Lucena, T., Ruvira, M.A., Arahal, D.R., and Macián, M.C.
-(2014) The family Rhodobacteraceae. In, \emph{The prokaryotes:
-Alphaproteobacteria and betaproteobacteria}. Springer-Verlag Berlin
-Heidelberg, pp. 439--512.
+(2014) The Family Rhodobacteraceae. In, \emph{The prokaryotes:
+\textless{}i\textgreater{}Alphaproteobacteria\textless{}/i\textgreater{}
+and
+\textless{}i\textgreater{}Betaproteobacteria\textless{}/i\textgreater{}}.
+Springer-Verlag Berlin Heidelberg, pp. 439--512.
 
@@
 et al. (2013) The SILVA ribosomal RNA gene database project: improved
-data processing and web-based tools. \emph{Nucleic acids research}
-\textbf{41}: D590--6.
+data processing and web-based tools. \emph{Nucleic Acids Research}
+\textbf{41}: D590--D596.
 
@@
 Reisser, J., Shaw, J., Hallegraeff, G., Proietti, M., Barnes, D.K.A.,
-Thums, M., et al. (2014) Millimeter-sized marine plastics: a new pelagic
-habitat for microorganisms and invertebrates. \emph{PloS one}
+Thums, M., et al. (2014) Millimeter-Sized Marine Plastics: A New Pelagic
+Habitat for Microorganisms and Invertebrates. \emph{PloS one}
 \textbf{9}: e100289.
@@
 \leavevmode\hypertarget{ref-Reyes2001}{}%
-Reyes, J. and Sansón, M. (2001) Biomass and production of the epiphytes
-on the leaves of Cymodocea nodosa in the Canary Islands. \emph{Botanica
+Reyes, J. and Sansón, M. (2001) Biomass and Production of the Epiphytes
+on the Leaves of \textless{}i\textgreater{}Cymodocea
+nodosa\textless{}/i\textgreater{} in the Canary Islands. \emph{Botanica
 Marina} \textbf{44}: 307--313.
@@
 Rickert, E., Wahl, M., Link, H., Richter, H., and Pohnert, G. (2016)
-Seasonal variations in surface metabolite composition of fucus
-vesiculosus and fucus serratus from the Baltic Sea. \emph{PLoS ONE}
-\textbf{11}:
+Seasonal Variations in Surface Metabolite Composition of Fucus
+vesiculosus and Fucus serratus from the Baltic Sea. \emph{PloS one}
+\textbf{11}: e0168196.
 
@@
 Schloss, P.D., Westcott, S.L., Ryabin, T., Hall, J.R., Hartmann, M.,
-Hollister, E.B., et al. (2009) Introducing mothur: open-source,
-platform-independent, community-supported software for describing and
-comparing microbial communities. \emph{Applied and environmental
+Hollister, E.B., et al. (2009) Introducing mothur: Open-Source,
+Platform-Independent, Community-Supported Software for Describing and
+Comparing Microbial Communities. \emph{Applied and environmental
 microbiology} \textbf{75}: 7537--7541.
@@
 Tarquinio, F., Hyndes, G.A., Laverock, B., Koenders, A., and Säwström,
-C. (2019) The seagrass holobiont: Understanding seagrass-bacteria
+C. (2019) The seagrass holobiont: understanding seagrass-bacteria
 interactions and their role in seagrass ecosystem functioning.
-\emph{FEMS Microbiology Letters} \textbf{366}:
+\emph{FEMS Microbiology Letters} \textbf{366}: fnz057.
 
@@
 Kjelleberg, S. (2010) Variability and abundance of the epiphytic
-bacterial community associated with a green marine Ulvacean alga.
+bacterial community associated with a green marine
+\textless{}i\textgreater{}Ulvacean\textless{}/i\textgreater{} alga.
 \emph{The ISME Journal} \textbf{4}: 301--311.
@@
 \leavevmode\hypertarget{ref-Ugarelli2019}{}%
-Ugarelli, K., Laas, P., and Stingl, U. (2019) The microbial communities
-of leaves and roots associated with turtle grass (Thalassia testudinum)
-and manatee grass (syringodium filliforme) are distinct from seawater
-and sediment communities, but are similar between species and sampling
-sites. \emph{Microorganisms} \textbf{7}:
+Ugarelli, K., Laas, P., and Stingl, U. (2019) The Microbial Communities
+of Leaves and Roots Associated with Turtle Grass
+(\textless{}i\textgreater{}Thalassia
+testudinum\textless{}/i\textgreater{}) and Manatee Grass
+(\textless{}i\textgreater{}Syringodium
+filliforme\textless{}/i\textgreater{}) are Distinct from Seawater and
+Sediment Communities, but Are Similar between Species and Sampling Si.
+\emph{Microorganisms} \textbf{7}: 4.
 
 \leavevmode\hypertarget{ref-Uku2007}{}%
-Uku, J., Björk, M., Bergman, B., and Díez, B. (2007) Characterization
-and comparison of prokaryotic epiphytes associated with three East
-African seagrasses. \emph{Journal of Phycology} \textbf{43}: 768--779.
+Uku, J., Björk, M., Bergman, B., and Díez, B. (2007) CHARACTERIZATION
+AND COMPARISON OF PROKARYOTIC EPIPHYTES ASSOCIATED WITH THREE EAST
+AFRICAN SEAGRASSES. \emph{Journal of Phycology} \textbf{43}: 768--779.
 
@@
 of Bacterial Communities Associated with Leaves of the Seagrass
-Halophila stipulacea by a Culture-Independent Small-Subunit rRNA Gene
-Approach. \emph{Microbial ecology} \textbf{39}: 22--31.
+\textless{}i\textgreater{}Halophila
+stipulacea\textless{}/i\textgreater{} by a Culture-Independent
+Small-Subunit rRNA Gene Approach. \emph{Microbial Ecology} \textbf{39}:
+22--31.
 
@@
 Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.D., François,
-R., et al. (2019) Welcome to the tidyverse. \emph{Journal of Open Source
+R., et al. (2019) Welcome to the Tidyverse. \emph{Journal of Open Source
 Software} \textbf{4}: 1686.
@@
 \leavevmode\hypertarget{ref-Xie2015}{}%
-Xie, Y. (2015) Dynamic Documents with \{R\} and knitr, 2nd ed. Boca
-Raton, Florida: Chapman; Hall/CRC.
+Xie, Y. (2015) Dynamic Documents with R and knitr, 2nd ed. Boca Raton,
+Florida: Chapman; Hall/CRC.
 
@@
 Xie, Y. (2014) knitr: A Comprehensive Tool for Reproducible Research in
-\{R\}. In, Stodden,V., Leisch,F., and Peng,R.D. (eds),
-\emph{Implementing reproducible computational research}. Chapman;
-Hall/CRC.
+R. In, Stodden,V., Leisch,F., and Peng,R.D. (eds), \emph{Implementing
+reproducible computational research}. Chapman; Hall/CRC.
 
@@
 Xie, Z., Lin, W., and Luo, J. (2017) Comparative Phenotype and Genome
-Analysis of Cellvibrio sp. PR1, a Xylanolytic and Agarolytic Bacterium
-from the Pearl River. \emph{BioMed research international}
-\textbf{2017}: 6304248.
+Analysis of
+\textless{}i\textgreater{}Cellvibrio\textless{}/i\textgreater{} sp. PR1,
+a Xylanolytic and Agarolytic Bacterium from the Pearl River.
+\emph{BioMed Research International} \textbf{2017}: 6304248.
 
@@
 C., et al. (2014) The SILVA and "All-species Living Tree Project (LTP)"
-taxonomic frameworks. \emph{Nucleic acids research} \textbf{42}:
-D643--8.
+taxonomic frameworks. \emph{Nucleic Acids Research} \textbf{42}:
+D643--D648.
 
@@
 Zavodnik, N., Travizi, A., and De Rosa, S. (1998) Seasonal variations in
-the rate of pliotosynthetic activity and chemical composition of the
-seagrass Cymodocea nodosa (Ucr.) Asch. \emph{Scientia Marina}
+the rate of photosynthetic activity and chemical composition of the
+seagrass \textless{}i\textgreater{}Cymodocea
+nodosa\textless{}/i\textgreater{} (Ucr.) Asch. \emph{Scientia Marina}
 \textbf{62}: 301--309.
@@
 
-\caption{Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between communties associated with the surfaces of macrophytes (\textit{C.
-nodosa} [Invaded] and \textit{C.
-cylindracea} [Invaded and Noninvaded]) and coomunities in the ambient seawater.\label{matrix}}\label{fig:unnamed-chunk-1}
+\caption{Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between assemblages associated with the surfaces of macrophytes (\textit{C.
+nodosa} [Mixed] and \textit{C.
+cylindracea} [Mixed and Monospecific]) and comunities in the ambient seawater.\label{matrix}}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 
-\caption{Proprotion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\textit{C.
-nodosa} [Invaded] and \textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\label{shared}}\label{fig:unnamed-chunk-2}
+\caption{Proportion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\textit{C.
+nodosa} [Mixed] and \textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\label{shared}}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 \caption{Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of macrophytes (\textit{C.
-nodosa} [Invaded] and \textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
-Samples from the same environment or same season are labeld in different colors.
+nodosa} [Mixed] and \textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+Samples from the same environment or same season are labeled in different colors.
 The proportion of explained variation by each axis is shown on the corresponding axis in parentheses.\label{pcoa}}\label{fig:unnamed-chunk-3}
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) bacterial and archaeal sequences on the surfaces of macrophytes (\textit{C.
-nodosa} [Invaded] and \textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\label{community}}\label{fig:unnamed-chunk-4}
+nodosa} [Mixed] and \textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+NR -- No Relative\label{community}}\label{fig:unnamed-chunk-4}
 \end{figure}
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) cyanobacterial sequences on the surfaces of macrophytes (\textit{C.
-nodosa} [Invaded] and \textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
-The proprotion of cyanobaterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
+nodosa} [Mixed] and \textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+The proportion of cyanobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
 NR -- No Relative\label{cyano}}\label{fig:unnamed-chunk-5}
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) sequences within the \textit{Bacteroidota} on the surfaces of macrophytes (\textit{C.
-nodosa} [Invaded] and \textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
-The proprotion of sequences classified as \textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar.
+nodosa} [Mixed] and \textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+The proportion of sequences classified as \textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar.
 NR -- No Relative\label{bactero}}\label{fig:unnamed-chunk-6}
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) alphaproteobacterial sequences on the surfaces of macrophytes (\textit{C.
-nodosa} [Invaded] and \textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
-The proprotion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
+nodosa} [Mixed] and \textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+The proportion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
 NR -- No Relative\label{alpha}}\label{fig:unnamed-chunk-7}
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) gammaproteobacterial sequences on the surfaces of macrophytes (\textit{C.
-nodosa} [Invaded] and \textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
-The proprotion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
+nodosa} [Mixed] and \textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+The proportion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
 NR -- No Relative\label{gamma}}\label{fig:unnamed-chunk-8}
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) sequences within the \textit{Desulfobacterota} on the surfaces of macrophytes (\textit{C.
-nodosa} [Invaded] and \textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.
-The proprotion of sequences classified as \textit{Desulfobacterota} in the total bacterial and archaeal community is given above the corresponding bar.
+nodosa} [Mixed] and \textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+The proportion of sequences classified as \textit{Desulfobacterota} in the total bacterial and archaeal community is given above the corresponding bar.
 NR -- No Relative\label{desulfo}}\label{fig:unnamed-chunk-9}

### 2020-03-19 submission/supplementary.tex
Analysis rerun.
Pipeline optimization.

@@
 \caption{Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes (\textit{C.
-nodosa} [Invaded] and \textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\label{rarefaction}}\label{fig:unnamed-chunk-1}
+nodosa} [Mixed] and \textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\label{rarefaction}}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 \caption{Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of macrophytes (\textit{C.
-nodosa} [Invaded] and \textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\label{calculators}}\label{fig:unnamed-chunk-2}
+nodosa} [Mixed] and \textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\label{calculators}}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 \caption{Relative contribution of chloroplast sequences on the surfaces of macrophytes (\textit{C.
-nodosa} [Invaded] and \textit{C.
-cylindracea} [Invaded and Noninvaded]) and in the ambient seawater.\label{chloroplast}}\label{fig:unnamed-chunk-3}
+nodosa} [Mixed] and \textit{C.
+cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\label{chloroplast}}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 
-\begin{longtable}{>{\centering\arraybackslash}p{6em}ccccc}
+\begin{longtable}[t]{>{\centering\arraybackslash}p{6em}ccccc}
 \caption{\label{tab:nseq_notus}Sample ID, Community Type, Sampling Date and Season, No.
@@
 \endlastfoot
-3 & Seawater & 13 July 2017 & Summer & 26,008 & 824\\
-5 & Seawater & 27 July 2017 & Summer & 58,955 & 993\\
-7 & Seawater & 10 August 2017 & Summer & 32,626 & 852\\
-9 & Seawater & 24 August 2017 & Summer & 60,929 & 1,058\\
-11 & Seawater & 19 September 2017 & Summer & 46,111 & 1,171\\
-13 & Seawater & 12 October 2017 & Autumn & 30,937 & 976\\
-15 & Seawater & 23 November 2017 & Autumn & 27,581 & 1,183\\
-17 & Seawater & 14 December 2017 & Autumn & 44,595 & 2,267\\
-19 & Seawater & 13 February 2018 & Winter & 44,186 & 1,309\\
-21 & Seawater & 27 March 2018 & Winter & 46,307 & 1,559\\
-23a & Seawater & 24 April 2018 & Spring & 30,970 & 998\\
-23b & Seawater & 24 April 2018 & Spring & 38,560 & 1,200\\
-25 & Seawater & 22 May 2018 & Spring & 53,880 & 1,783\\
-27 & Seawater & 19 June 2018 & Spring & 77,465 & 1,297\\
-29 & Seawater & 10 July 2018 & Summer & 50,794 & 1,451\\
-31 & Seawater & 9 August 2018 & Summer & 39,049 & 1,129\\
-33 & Seawater & 4 September 2018 & Summer & 36,206 & 1,205\\
-35 & Seawater & 5 October 2018 & Autumn & 49,580 & 1,013\\
-37 & \textit{Cymodocea nodosa} (Invaded) & 23 November 2017 & Autumn & 31,235 & 1,548\\
-41 & \textit{Cymodocea nodosa} (Invaded) & 4 December 2017 & Autumn & 24,237 & 1,176\\
-45 & \textit{Cymodocea nodosa} (Invaded) & 14 December 2017 & Autumn & 32,683 & 1,051\\
-49 & \textit{Cymodocea nodosa} (Invaded) & 13 February 2018 & Winter & 9,087 & 1,218\\
-52 & \textit{Cymodocea nodosa} (Invaded) & 27 March 2018 & Winter & 17,003 & 1,218\\
-55 & \textit{Cymodocea nodosa} (Invaded) & 24 April 2018 & Spring & 42,640 & 2,054\\
-58 & \textit{Cymodocea nodosa} (Invaded) & 22 May 2018 & Spring & 21,337 & 1,292\\
-61 & \textit{Cymodocea nodosa} (Invaded) & 19 June 2018 & Spring & 31,735 & 2,104\\
-64 & \textit{Cymodocea nodosa} (Invaded) & 10 July 2018 & Summer & 35,733 & 1,792\\
-67 & \textit{Cymodocea nodosa} (Invaded) & 9 August 2018 & Summer & 26,362 & 2,119\\
-70 & \textit{Cymodocea nodosa} (Invaded) & 4 September 2018 & Summer & 23,278 & 1,715\\
-73 & \textit{Cymodocea nodosa} (Invaded) & 5 October 2018 & Autumn & 29,909 & 2,205\\
-38 & \textit{Caulerpa cylindracea} (Invaded) & 23 November 2017 & Autumn & 36,321 & 2,889\\
-42 & \textit{Caulerpa cylindracea} (Invaded) & 4 December 2017 & Autumn & 28,370 & 3,249\\
-46 & \textit{Caulerpa cylindracea} (Invaded) & 14 December 2017 & Autumn & 34,720 & 3,054\\
-53 & \textit{Caulerpa cylindracea} (Invaded) & 27 March 2018 & Winter & 28,696 & 2,472\\
-56 & \textit{Caulerpa cylindracea} (Invaded) & 24 April 2018 & Spring & 34,773 & 3,054\\
-59 & \textit{Caulerpa cylindracea} (Invaded) & 22 May 2018 & Spring & 23,397 & 2,706\\
-62 & \textit{Caulerpa cylindracea} (Invaded) & 19 June 2018 & Spring & 36,477 & 3,299\\
-65 & \textit{Caulerpa cylindracea} (Invaded) & 10 July 2018 & Summer & 18,485 & 2,188\\
-68 & \textit{Caulerpa cylindracea} (Invaded) & 9 August 2018 & Summer & 31,940 & 3,103\\
-71 & \textit{Caulerpa cylindracea} (Invaded) & 4 September 2018 & Summer & 29,281 & 3,160\\
-74 & \textit{Caulerpa cylindracea} (Invaded) & 5 October 2018 & Autumn & 11,705 & 1,705\\
-39 & \textit{Caulerpa cylindracea} (Nonnvaded) & 23 November 2017 & Autumn & 22,064 & 2,438\\
-43 & \textit{Caulerpa cylindracea} (Nonnvaded) & 4 December 2017 & Autumn & 13,661 & 1,887\\
-47 & \textit{Caulerpa cylindracea} (Nonnvaded) & 14 December 2017 & Autumn & 8,410 & 1,451\\
-51 & \textit{Caulerpa cylindracea} (Nonnvaded) & 13 February 2018 & Winter & 11,676 & 1,901\\
-54 & \textit{Caulerpa cylindracea} (Nonnvaded) & 27 March 2018 & Winter & 39,472 & 3,136\\
-57 & \textit{Caulerpa cylindracea} (Nonnvaded) & 24 April 2018 & Spring & 20,298 & 2,822\\
-60 & \textit{Caulerpa cylindracea} (Nonnvaded) & 22 May 2018 & Spring & 33,047 & 3,297\\
-63 & \textit{Caulerpa cylindracea} (Nonnvaded) & 19 June 2018 & Spring & 41,833 & 3,968\\
-66 & \textit{Caulerpa cylindracea} (Nonnvaded) & 10 July 2018 & Summer & 27,036 & 2,674\\
-69 & \textit{Caulerpa cylindracea} (Nonnvaded) & 9 August 2018 & Summer & 26,737 & 3,130\\
-72 & \textit{Caulerpa cylindracea} (Nonnvaded) & 4 September 2018 & Summer & 31,869 & 3,240\\
-75 & \textit{Caulerpa cylindracea} (Nonnvaded) & 5 October 2018 & Autumn & 33,068 & 3,086\\*
+3 & Seawater & 13 July 2017 & Summer & 25,996 & 821\\
+5 & Seawater & 27 July 2017 & Summer & 58,954 & 993\\
+7 & Seawater & 10 August 2017 & Summer & 32,626 & 853\\
+9 & Seawater & 24 August 2017 & Summer & 60,935 & 1,064\\
+11 & Seawater & 19 September 2017 & Summer & 46,107 & 1,179\\
+13 & Seawater & 12 October 2017 & Autumn & 30,911 & 971\\
+15 & Seawater & 23 November 2017 & Autumn & 27,585 & 1,197\\
+17 & Seawater & 14 December 2017 & Autumn & 44,595 & 2,275\\
+19 & Seawater & 13 February 2018 & Winter & 44,207 & 1,306\\
+21 & Seawater & 27 March 2018 & Winter & 46,313 & 1,561\\
+23a & Seawater & 24 April 2018 & Spring & 30,973 & 999\\
+23b & Seawater & 24 April 2018 & Spring & 38,563 & 1,195\\
+25 & Seawater & 22 May 2018 & Spring & 53,877 & 1,787\\
+27 & Seawater & 19 June 2018 & Spring & 77,465 & 1,290\\
+29 & Seawater & 10 July 2018 & Summer & 50,793 & 1,445\\
+31 & Seawater & 9 August 2018 & Summer & 39,041 & 1,123\\
+33 & Seawater & 4 September 2018 & Summer & 36,206 & 1,206\\
+35 & Seawater & 5 October 2018 & Autumn & 49,585 & 1,019\\
+37 & \textit{Cymodocea nodosa} (Mixed) & 23 November 2017 & Autumn & 31,237 & 1,555\\
+41 & \textit{Cymodocea nodosa} (Mixed) & 4 December 2017 & Autumn & 24,234 & 1,174\\
+45 & \textit{Cymodocea nodosa} (Mixed) & 14 December 2017 & Autumn & 32,683 & 1,061\\
+49 & \textit{Cymodocea nodosa} (Mixed) & 13 February 2018 & Winter & 9,088 & 1,216\\
+52 & \textit{Cymodocea nodosa} (Mixed) & 27 March 2018 & Winter & 17,005 & 1,216\\
+55 & \textit{Cymodocea nodosa} (Mixed) & 24 April 2018 & Spring & 42,645 & 2,053\\
+58 & \textit{Cymodocea nodosa} (Mixed) & 22 May 2018 & Spring & 21,337 & 1,290\\
+61 & \textit{Cymodocea nodosa} (Mixed) & 19 June 2018 & Spring & 31,726 & 2,098\\
+64 & \textit{Cymodocea nodosa} (Mixed) & 10 July 2018 & Summer & 35,729 & 1,793\\
+67 & \textit{Cymodocea nodosa} (Mixed) & 9 August 2018 & Summer & 26,356 & 2,116\\
+70 & \textit{Cymodocea nodosa} (Mixed) & 4 September 2018 & Summer & 23,275 & 1,724\\
+73 & \textit{Cymodocea nodosa} (Mixed) & 5 October 2018 & Autumn & 29,907 & 2,211\\
+38 & \textit{Caulerpa cylindracea} (Mixed) & 23 November 2017 & Autumn & 36,309 & 2,896\\
+42 & \textit{Caulerpa cylindracea} (Mixed) & 4 December 2017 & Autumn & 28,383 & 3,249\\
+46 & \textit{Caulerpa cylindracea} (Mixed) & 14 December 2017 & Autumn & 34,718 & 3,067\\
+53 & \textit{Caulerpa cylindracea} (Mixed) & 27 March 2018 & Winter & 28,689 & 2,481\\
+56 & \textit{Caulerpa cylindracea} (Mixed) & 24 April 2018 & Spring & 34,784 & 3,057\\
+59 & \textit{Caulerpa cylindracea} (Mixed) & 22 May 2018 & Spring & 23,392 & 2,712\\
+62 & \textit{Caulerpa cylindracea} (Mixed) & 19 June 2018 & Spring & 36,488 & 3,311\\
+65 & \textit{Caulerpa cylindracea} (Mixed) & 10 July 2018 & Summer & 18,483 & 2,188\\
+68 & \textit{Caulerpa cylindracea} (Mixed) & 9 August 2018 & Summer & 31,943 & 3,097\\
+71 & \textit{Caulerpa cylindracea} (Mixed) & 4 September 2018 & Summer & 29,273 & 3,152\\
+74 & \textit{Caulerpa cylindracea} (Mixed) & 5 October 2018 & Autumn & 11,707 & 1,711\\
+39 & \textit{Caulerpa cylindracea} (Monospecific) & 23 November 2017 & Autumn & 22,057 & 2,428\\
+43 & \textit{Caulerpa cylindracea} (Monospecific) & 4 December 2017 & Autumn & 13,661 & 1,892\\
+47 & \textit{Caulerpa cylindracea} (Monospecific) & 14 December 2017 & Autumn & 8,409 & 1,451\\
+51 & \textit{Caulerpa cylindracea} (Monospecific) & 13 February 2018 & Winter & 11,673 & 1,899\\
+54 & \textit{Caulerpa cylindracea} (Monospecific) & 27 March 2018 & Winter & 39,471 & 3,142\\
+57 & \textit{Caulerpa cylindracea} (Monospecific) & 24 April 2018 & Spring & 20,304 & 2,836\\
+60 & \textit{Caulerpa cylindracea} (Monospecific) & 22 May 2018 & Spring & 33,035 & 3,314\\
+63 & \textit{Caulerpa cylindracea} (Monospecific) & 19 June 2018 & Spring & 41,841 & 3,976\\
+66 & \textit{Caulerpa cylindracea} (Monospecific) & 10 July 2018 & Summer & 27,038 & 2,669\\
+69 & \textit{Caulerpa cylindracea} (Monospecific) & 9 August 2018 & Summer & 26,734 & 3,123\\
+72 & \textit{Caulerpa cylindracea} (Monospecific) & 4 September 2018 & Summer & 31,874 & 3,248\\
+75 & \textit{Caulerpa cylindracea} (Monospecific) & 5 October 2018 & Autumn & 33,073 & 3,078\\*
 \end{longtable}

### 2020-03-19 submission/manuscript.Rmd
Pipeline optimization.
Added changes to manuscipt.

@@
 ```{r out.width="70%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between assemblages associated with the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed] and \\textit{C.
-cylindracea} [Mixed and Monospecific]) and comunities in the ambient seawater.\\label{matrix}", fig.pos="H"}
+nodosa} [Mixed Settlement] and \\textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and comunities in the ambient seawater.\\label{matrix}", fig.pos="H"}
 include_graphics("../results/figures/matrix.jpg")
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed] and \\textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\\label{shared}", fig.pos="H"}
+nodosa} [Mixed Settlement] and \\textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.\\label{shared}", fig.pos="H"}
 include_graphics("../results/figures/seasonal_shared.jpg")
@@
 ```{r out.width="55%", fig.align="center", fig.cap="Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed] and \\textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+nodosa} [Mixed Settlement] and \\textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
 Samples from the same environment or same season are labeled in different colors.
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) bacterial and archaeal sequences on the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed] and \\textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+nodosa} [Mixed Settlement] and \\textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
 NR -- No Relative\\label{community}", fig.pos="H"}
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) cyanobacterial sequences on the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed] and \\textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+nodosa} [Mixed Settlement] and \\textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
 The proportion of cyanobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) sequences within the \\textit{Bacteroidota} on the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed] and \\textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+nodosa} [Mixed Settlement] and \\textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
 The proportion of sequences classified as \\textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar.
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) alphaproteobacterial sequences on the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed] and \\textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+nodosa} [Mixed Settlement] and \\textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
 The proportion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) gammaproteobacterial sequences on the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed] and \\textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+nodosa} [Mixed Settlement] and \\textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
 The proportion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) sequences within the \\textit{Desulfobacterota} on the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed] and \\textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+nodosa} [Mixed Settlement] and \\textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
 The proportion of sequences classified as \\textit{Desulfobacterota} in the total bacterial and archaeal community is given above the corresponding bar.

### 2020-03-19 submission/supplementary.Rmd
Added changes to manuscipt.

@@
 ```{r out.width="85%", fig.align="center", fig.cap="Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed] and \\textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\\label{rarefaction}", fig.pos="H"}
+nodosa} [Mixed Settlement] and \\textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.\\label{rarefaction}", fig.pos="H"}
 include_graphics("../results/figures/rarefaction.jpg")
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed] and \\textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\\label{calculators}", fig.pos="H"}
+nodosa} [Mixed Settlement] and \\textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.\\label{calculators}", fig.pos="H"}
 include_graphics("../results/figures/calculators.jpg")
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Relative contribution of chloroplast sequences on the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed] and \\textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\\label{chloroplast}", fig.pos="H"}
+nodosa} [Mixed Settlement] and \\textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.\\label{chloroplast}", fig.pos="H"}
 include_graphics("../results/figures/chloroplast_bar_plot.jpg")

### 2020-03-20 submission/manuscript.tex
References update.

@@
 10\textsuperscript{7} \si{\cells\per\cm\squared} (Armstrong \emph{et
-al.}, 2000; Bengtsson \emph{et al.}, 2010; Burke \emph{et al.}, 2011b).
-In such an environment a number of positive and negative interactions
-between the macrophyte and colonizing microorganisms have been described
-(Egan \emph{et al.}, 2013; Hollants \emph{et al.}, 2013; Tarquinio
-\emph{et al.}, 2019). Macrophytes can promote growth of associated
-microbes by nutrient exudation, while in return microorganisms may
-support macrophyte performance through improved nutrient availability,
-phytohormone production and protection form toxic compounds, oxidative
-stress, biofouling organisms and pathogens (Egan \emph{et al.}, 2013;
-Hollants \emph{et al.}, 2013; Tarquinio \emph{et al.}, 2019). Beside
-this positive interactions, macrophytes can negatively impact the
-associated microbes such as pathogenic bacteria by producing reactive
-oxygen species and secondary metabolites (Egan \emph{et al.}, 2013;
-Hollants \emph{et al.}, 2013; Tarquinio \emph{et al.}, 2019).
+al.}, 2000; Bengtsson \emph{et al.}, 2010; Burke and Thomas \emph{et
+al.}, 2011). In such an environment a number of positive and negative
+interactions between the macrophyte and colonizing microorganisms have
+been described (Egan \emph{et al.}, 2013; Hollants \emph{et al.}, 2013;
+Tarquinio \emph{et al.}, 2019). Macrophytes can promote growth of
+associated microbes by nutrient exudation, while in return
+microorganisms may support macrophyte performance through improved
+nutrient availability, phytohormone production and protection form toxic
+compounds, oxidative stress, biofouling organisms and pathogens (Egan
+\emph{et al.}, 2013; Hollants \emph{et al.}, 2013; Tarquinio \emph{et
+al.}, 2019). Beside this positive interactions, macrophytes can
+negatively impact the associated microbes such as pathogenic bacteria by
+producing reactive oxygen species and secondary metabolites (Egan
+\emph{et al.}, 2013; Hollants \emph{et al.}, 2013; Tarquinio \emph{et
+al.}, 2019).
 
@@
 functions were conserved both between host species and individuals
-(Burke \emph{et al.}, 2011a; Roth-Schulze \emph{et al.}, 2016). This
-discrepancy between taxonomic and functional composition could be
-explained by the lottery hypothesis. It postulates that an initial
-random colonization step is performed from a set of functionally
-equivalent taxonomic groups resulting in taxonomically different
-epiphytic communities sharing a core set of functional genes (Burke
-\emph{et al.}, 2011a; Roth-Schulze \emph{et al.}, 2016). In addition,
-some of the variation in the observed data could be attributed to
-different techniques used in various studies, such as different
-protocols for epiphytic cell detachment and/or DNA isolation, as no
-standard protocol to study epiphytic communities was established
-(Ugarelli \emph{et al.}, 2019; Korlević \emph{et al.}, submitted).
+(Burke and Peter Steinberg \emph{et al.}, 2011; Roth-Schulze \emph{et
+al.}, 2016). This discrepancy between taxonomic and functional
+composition could be explained by the lottery hypothesis. It postulates
+that an initial random colonization step is performed from a set of
+functionally equivalent taxonomic groups resulting in taxonomically
+different epiphytic communities sharing a core set of functional genes
+(Burke and Peter Steinberg \emph{et al.}, 2011; Roth-Schulze \emph{et
+al.}, 2016). In addition, some of the variation in the observed data
+could be attributed to different techniques used in various studies,
+such as different protocols for epiphytic cell detachment and/or DNA
+isolation, as no standard protocol to study epiphytic communities was
+established (Ugarelli \emph{et al.}, 2019; Korlević \emph{et al.},
+submitted).
 
@@
 not encompass seasonal changes (Crump and Koch, 2008; Lachnit \emph{et
-al.}, 2009; Burke \emph{et al.}, 2011b; Roth-Schulze \emph{et al.},
-2016; Ugarelli \emph{et al.}, 2019). In addition, if seasonal changes
-were taken into account low temporal frequency and/or methodologies that
-do not allow high taxonomic resolution were used (Tujula \emph{et al.},
-2010; Lachnit \emph{et al.}, 2011; Miranda \emph{et al.}, 2013; Michelou
-\emph{et al.}, 2013). In the present study we describe the seasonal
-dynamics of bacterial and archaeal communities on the surfaces of the
-seagrass \emph{Cymodocea nodosa} and siphonous macroalgae \emph{Caulerpa
-cylindracea} determined on a mostly monthly scale. Bacterial and
-archaeal epiphytes were sampled in a meadow of \emph{C. nodosa} invaded
-by the invasive \emph{C. cylindracea} and in a locality of only \emph{C.
-cylindracea} located in the proximity of the meadow. In addition, for
-comparison, the community of the ambient seawater was characterized.
+al.}, 2009; Burke and Thomas \emph{et al.}, 2011; Roth-Schulze \emph{et
+al.}, 2016; Ugarelli \emph{et al.}, 2019). In addition, if seasonal
+changes were taken into account low temporal frequency and/or
+methodologies that do not allow high taxonomic resolution were used
+(Tujula \emph{et al.}, 2010; Lachnit \emph{et al.}, 2011; Miranda
+\emph{et al.}, 2013; Michelou \emph{et al.}, 2013). In the present study
+we describe the seasonal dynamics of bacterial and archaeal communities
+on the surfaces of the seagrass \emph{Cymodocea nodosa} and siphonous
+macroalgae \emph{Caulerpa cylindracea} determined on a mostly monthly
+scale. Bacterial and archaeal epiphytes were sampled in a meadow of
+\emph{C. nodosa} invaded by the invasive \emph{C. cylindracea} and in a
+locality of only \emph{C. cylindracea} located in the proximity of the
+meadow. In addition, for comparison, the community of the ambient
+seawater was characterized.
 
@@
 in the ambient seawater were also processed for comparison. The number
-of reads per sample ranged between 8,409 and 77,465 sequences
+of reads per sample ranged between 8,409 and 77,463 sequences
 (\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
@@
 (\autoref{supp-rarefaction}). Following quality curation and exclusion
-of sequences mentioned before reads were clustered into 28,691 different
+of sequences mentioned before reads were clustered into 28,702 different
 OTUs at a similarity level of 97 \si{\percent}. Read numbers were
 normalized to the minimum number of sequences, 8,409
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,071
-different OTUs that ranged from 386 to 2,009 OTUs per sample
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,172
+different OTUs that ranged from 366 to 2,043 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,679.5 ±
-127.4 OTUs{]} and monospecific {[}Number of OTUs, 1,736.3 ± 156.5
-OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,059 ±
-216.2 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 528.2 ± 135.4 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,683.2 ±
+130.9 OTUs{]} and monospecific {[}Number of OTUs, 1,737.2 ± 166.4
+OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,058.4 ±
+216.8 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 529.5 ± 143.0 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 community was found between mixed and monospecific \emph{C. cylindracea}
-(Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared values were
+(Jaccard, 0.35; Bray-Curtis, 0.78), while lower shared values were
 calculated between seawater and epiphytic communities
@@
 from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (39.9 -- 51.8
+These OTUs also made a high proportion of total sequences (40.8 -- 52.2
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 and Autumn/Winter/Spring samples was observed that was not so strongly
-supported (ANOSIM, R = 0.30, \emph{p} \textless{} 0.05)
+supported (ANOSIM, R = 0.30, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 monospecific {[} (7.8 ± 2.4 \si{\percent}){]}) associated communities in
-autumn and for seawater communities in winter (8.8 ± 7.4 \si{\percent}).
+autumn and for seawater communities in winter (8.8 ± 7.5 \si{\percent}).
 Large taxonomic differences between surface associated and seawater
@@
 the level of OTUs that is in agreement with most published studies
-(Burke \emph{et al.}, 2011b; Michelou \emph{et al.}, 2013; Roth-Schulze
-\emph{et al.}, 2016; Crump \emph{et al.}, 2018; Ugarelli \emph{et al.},
-2019). These data indicate that marine macrophytes are selecting, from a
-pool of seawater microbial taxa, the one that can colonize and
-proliferate on their surfaces (Salaün \emph{et al.}, 2012; Michelou
-\emph{et al.}, 2013). In contrast to these findings Fahimipour \emph{et
-al.} (2017) found, in a global study of \emph{Zostera marina},
+(Burke and Thomas \emph{et al.}, 2011; Michelou \emph{et al.}, 2013;
+Roth-Schulze \emph{et al.}, 2016; Crump \emph{et al.}, 2018; Ugarelli
+\emph{et al.}, 2019). These data indicate that marine macrophytes are
+selecting, from a pool of seawater microbial taxa, the one that can
+colonize and proliferate on their surfaces (Salaün \emph{et al.}, 2012;
+Michelou \emph{et al.}, 2013). In contrast to these findings Fahimipour
+\emph{et al.} (2017) found, in a global study of \emph{Zostera marina},
 similarities between leaves and seawater samples. Discrepancies between
@@
 \emph{Gammaproteobacteria} and \emph{Verrucomicrobiota} were described
-that is in agreement with previously reported data (Burke \emph{et al.},
-2011b; Egan \emph{et al.}, 2013; Michelou \emph{et al.}, 2013). In
-contrast, when low taxonomic ranks were analyzed (i.g. family and genus)
-a strong differentiation was observed. A similar differentiation at
-lower taxonomic ranks was described for other species of macrophytes
-(Egan \emph{et al.}, 2013; Michelou \emph{et al.}, 2013; Ugarelli
-\emph{et al.}, 2019).
+that is in agreement with previously reported data (Burke and Thomas
+\emph{et al.}, 2011; Egan \emph{et al.}, 2013; Michelou \emph{et al.},
+2013). In contrast, when low taxonomic ranks were analyzed (i.g. family
+and genus) a strong differentiation was observed. A similar
+differentiation at lower taxonomic ranks was described for other species
+of macrophytes (Egan \emph{et al.}, 2013; Michelou \emph{et al.}, 2013;
+Ugarelli \emph{et al.}, 2019).
 
@@
 this study were described to be specific for other species of
-macrophytes (Burke \emph{et al.}, 2011b; Lachnit \emph{et al.}, 2011;
-Bengtsson \emph{et al.}, 2017). In contrast to high taxonomic ranks, a
-substantial differentiation between host specific communities was found,
-which supports the host-specific hypothesis. Similar host-specificity
-was observed for different species of macroalgae and seagrasses (Lachnit
-\emph{et al.}, 2011; Roth-Schulze \emph{et al.}, 2016; Ugarelli \emph{et
-al.}, 2019; Morrissey \emph{et al.}, 2019). Taken together, at high
-taxonomic ranks a core set of taxa could be described that is
-characteristic for all or many macrophytes, while at low taxonomic ranks
-a community specific to host species could be identified (Egan \emph{et
-al.}, 2013).
+macrophytes (Burke and Thomas \emph{et al.}, 2011; Lachnit \emph{et
+al.}, 2011; Bengtsson \emph{et al.}, 2017). In contrast to high
+taxonomic ranks, a substantial differentiation between host specific
+communities was found, which supports the host-specific hypothesis.
+Similar host-specificity was observed for different species of
+macroalgae and seagrasses (Lachnit \emph{et al.}, 2011; Roth-Schulze
+\emph{et al.}, 2016; Ugarelli \emph{et al.}, 2019; Morrissey \emph{et
+al.}, 2019). Taken together, at high taxonomic ranks a core set of taxa
+could be described that is characteristic for all or many macrophytes,
+while at low taxonomic ranks a community specific to host species could
+be identified (Egan \emph{et al.}, 2013).
 
@@
 sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 51.8
+were accounting for a high proportion of sequences (\(\leq\) 52.2
 \si{\percent}). A very similar proportion of persistent OTUs and their
@@
 suggested that they are involved in the hydrolysis and utilization of
-complex carbon sources (Burke \emph{et al.}, 2011b; McIlroy and Nielsen,
-2014; Crump \emph{et al.}, 2018). On the other hand, families
-\emph{Flavobacteriaceae} and \emph{Rhodothermaceae} showed seasonal
-patterns, with \emph{Flavobacteriaceae} being more pronounced from
-November to June and \emph{Rhodothermaceae} from June to October. Within
-\emph{Alphaproteobacteria} the family \emph{Rhodobacteraceae} was
+complex carbon sources (Burke and Thomas \emph{et al.}, 2011; McIlroy
+and Nielsen, 2014; Crump \emph{et al.}, 2018). On the other hand,
+families \emph{Flavobacteriaceae} and \emph{Rhodothermaceae} showed
+seasonal patterns, with \emph{Flavobacteriaceae} being more pronounced
+from November to June and \emph{Rhodothermaceae} from June to October.
+Within \emph{Alphaproteobacteria} the family \emph{Rhodobacteraceae} was
 comprising the majority of sequences throughout the year. This
 metabolically versatile family is often associated with macrophyte
-surfaces and usually is one of the most abundant groups (Burke \emph{et
-al.}, 2011b; Michelou \emph{et al.}, 2013; Pujalte \emph{et al.}, 2014).
-In addition, \emph{Hyphomonadaceae} were found in all samples.
-Interestingly, some of the species within this group contain stalks on
-their cells which can be used to attach to the macrophyte surface
-(Weidner \emph{et al.}, 2000; Abraham and Rohde, 2014). Within the
-\emph{Gammaproteobacteria}, sequences without known representatives were
-the most pronounced group present throughout the year. In addition,
-\emph{Granulosicoccus} was also found in almost all samples.
-\emph{Gammaproteobacteria} are often a major constituent of macrophyte
-epiphytic communities (Burke \emph{et al.}, 2011b; Michelou \emph{et
-al.}, 2013; Crump \emph{et al.}, 2018). Beside these two groups other
-less pronounced taxa showed seasonal and host-specific patterns. For
-example, \emph{C. cylindracea} was characterized by \emph{Thioploca}, a
-known sulfur sediment bacteria and \emph{Cellvibrionaceae}, a family
-whose cultured members are known polysaccharide degraders (Jørgensen and
-Gallardo, 1999; Xie \emph{et al.}, 2017). \emph{Desulfobacterota} were
-found only associated with \emph{C. cylindracea} and no group within
-this phylum showed seasonal patterns. The presence of this phylum only
-on \emph{C. cylindracea} is to be expected as part of the epiphytic
-community is directly in contact with the sediment. The
-\emph{Desulfobacterota} community was dominated by \emph{Desulfatitalea}
-and no relative \emph{Desulfocapsaceae}, known sulfate sediment groups
-(Kuever, 2014; Higashioka \emph{et al.}, 2015).
+surfaces and usually is one of the most abundant groups (Burke and
+Thomas \emph{et al.}, 2011; Michelou \emph{et al.}, 2013; Pujalte
+\emph{et al.}, 2014). In addition, \emph{Hyphomonadaceae} were found in
+all samples. Interestingly, some of the species within this group
+contain stalks on their cells which can be used to attach to the
+macrophyte surface (Weidner \emph{et al.}, 2000; Abraham and Rohde,
+2014). Within the \emph{Gammaproteobacteria}, sequences without known
+representatives were the most pronounced group present throughout the
+year. In addition, \emph{Granulosicoccus} was also found in almost all
+samples. \emph{Gammaproteobacteria} are often a major constituent of
+macrophyte epiphytic communities (Burke and Thomas \emph{et al.}, 2011;
+Michelou \emph{et al.}, 2013; Crump \emph{et al.}, 2018). Beside these
+two groups other less pronounced taxa showed seasonal and host-specific
+patterns. For example, \emph{C. cylindracea} was characterized by
+\emph{Thioploca}, a known sulfur sediment bacteria and
+\emph{Cellvibrionaceae}, a family whose cultured members are known
+polysaccharide degraders (Jørgensen and Gallardo, 1999; Xie \emph{et
+al.}, 2017). \emph{Desulfobacterota} were found only associated with
+\emph{C. cylindracea} and no group within this phylum showed seasonal
+patterns. The presence of this phylum only on \emph{C. cylindracea} is
+to be expected as part of the epiphytic community is directly in contact
+with the sediment. The \emph{Desulfobacterota} community was dominated
+by \emph{Desulfatitalea} and no relative \emph{Desulfocapsaceae}, known
+sulfate sediment groups (Kuever, 2014; Higashioka \emph{et al.}, 2015).
 
@@
 colonization step performed from a set of functionally equivalent
-taxonomic groups (Burke \emph{et al.}, 2011a; Roth-Schulze \emph{et
-al.}, 2016). It is possible that functional redundancy is a
-characteristic of high abundance taxa detected to be present throughout
-the year, while seasonal and/or host-specific functions are an attribute
-of taxa displaying successional patterns. Further studies connecting
-taxonomy with functional properties will be required to elucidate the
-degree of functional redundancy or specificity in epiphytic microbial
-communities.
+taxonomic groups (Burke and Peter Steinberg \emph{et al.}, 2011;
+Roth-Schulze \emph{et al.}, 2016). It is possible that functional
+redundancy is a characteristic of high abundance taxa detected to be
+present throughout the year, while seasonal and/or host-specific
+functions are an attribute of taxa displaying successional patterns.
+Further studies connecting taxonomy with functional properties will be
+required to elucidate the degree of functional redundancy or specificity
+in epiphytic microbial communities.
 
@@
 1.3.0) (Wickham \emph{et al.}, 2019) and multiple other packages (Xie,
-2014, 2015, 2019a, 2019b, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018;
-Allaire \emph{et al.}, 2019; Zhu, 2019). The detailed analysis procedure
-including the R Markdown file for this paper are available as a GitHub
-repository
+2014, 2015, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018; Y. Xie,
+2019a, 2019b; Allaire \emph{et al.}, 2019; Zhu, 2019). The detailed
+analysis procedure including the R Markdown file for this paper are
+available as a GitHub repository
 (\url{https://github.com/mkorlevic/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020}).
@@
 \leavevmode\hypertarget{ref-Abraham2014}{}%
-Abraham, W.R. and Rohde, M. (2014) The Family
-\textless{}i\textgreater{}Hyphomonadaceae\textless{}i/\textgreater{}.
-In, \emph{The prokaryotes:
-\textless{}i\textgreater{}Alphaproteobacteria\textless{}i/\textgreater{}
-and
-\textless{}i\textgreater{}Betaproteobacteria\textless{}i/\textgreater{}}.
-Springer-Verlag Berlin Heidelberg, pp. 283--299.
+Abraham, W.R. and Rohde, M. (2014) The Family \emph{Hyphomonadaceae}.
+In, Rosenberg,E., DeLong,E.F., Lory,S., Stackebrandt,E., and Thompson,F.
+(eds), \emph{The Prokaryotes: Alphaproteobacteria and
+Betaproteobacteria}. Berlin, Heidelberg: Springer-Verlag, pp. 283--299.
 
@@
 revision to V4 region SSU rRNA 806R gene primer greatly increases
-detection of SAR11 bacterioplankton. \emph{Aquatic Microbial Ecology}
+detection of SAR11 bacterioplankton. \emph{Aquat Microb Ecol}
 \textbf{75}: 129--137.
@@
 Armstrong, E., Rogerson, A., and Leftley, J. (2000) The Abundance of
-Heterotrophic Protists Associated with Intertidal Seaweeds.
-\emph{Estuarine, Coastal and Shelf Science} \textbf{50}: 415--424.
+Heterotrophic Protists Associated with Intertidal Seaweeds. \emph{Estuar
+Coast Shelf Sci} \textbf{50}: 415--424.
 
@@
 Blindow, I. (2017) Eelgrass Leaf Surface Microbiomes Are Locally
-Variable and Highly Correlated with Epibiotic Eukaryotes.
-\emph{Frontiers in Microbiology} \textbf{8}: 1312.
+Variable and Highly Correlated with Epibiotic Eukaryotes. \emph{Front
+Microbiol} \textbf{8}: 1312.
 
@@
 Bengtsson, M., Sjøtun, K., and Øvreås, L. (2010) Seasonal dynamics of
-bacterial biofilms on the kelp \textless{}i\textgreater{}Laminaria
-hyperborea\textless{}i/\textgreater{}. \emph{Aquatic Microbial Ecology}
-\textbf{60}: 71--83.
+bacterial biofilms on the kelp \emph{Laminaria hyperborea}. \emph{Aquat
+Microb Ecol} \textbf{60}: 71--83.
 
@@
 of dimethylsulfoniopropionate (DMSP) in the Mediterranean seagrass
-\textless{}i\textgreater{}Posidonia oceanica\textless{}i/\textgreater{}.
-\emph{Aquatic Botany} \textbf{125}: 72--79.
+\emph{Posidonia oceanica}. \emph{Aquat Bot} \textbf{125}: 72--79.
 
@@
 Burke, C., Kjelleberg, S., and Thomas, T. (2009) Selective Extraction of
-Bacterial DNA from the Surfaces of Macroalgae. \emph{Applied and
-Environmental Microbiology} \textbf{75}: 252--256.
+Bacterial DNA from the Surfaces of Macroalgae. \emph{Appl Environ
+Microbiol} \textbf{75}: 252--256.
 
@@
 Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
-(2011a) Bacterial community assembly based on functional genes rather
-than species. \emph{Proceedings of the National Academy of Sciences of
-the United States of America} \textbf{108}: 14288--14293.
+(2011) Bacterial community assembly based on functional genes rather
+than species. \emph{Proc Natl Acad Sci U S A} \textbf{108}:
+14288--14293.
 
@@
 Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
-(2011b) Composition, uniqueness and variability of the epiphytic
-bacterial community of the green alga \textless{}i\textgreater{}Ulva
-australis\textless{}i/\textgreater{}. \emph{The ISME Journal}
-\textbf{5}: 590--600.
+(2011) Composition, uniqueness and variability of the epiphytic
+bacterial community of the green alga \emph{Ulva australis}. \emph{ISME
+J} \textbf{5}: 590--600.
 
@@
 diversity of extant stromatolites in the hypersaline marine environment
-of Shark Bay, Australia. \emph{Environmental Microbiology} \textbf{6}:
+of Shark Bay, Australia. \emph{Environ Microbiol} \textbf{6}:
 1096--1101.
@@
 (2014) An ultrasonic method for separation of epiphytic microbes from
-freshwater submerged macrophytes. \emph{Journal of Basic Microbiology}
-\textbf{54}: 758--761.
+freshwater submerged macrophytes. \emph{J Basic Microbiol} \textbf{54}:
+758--761.
 
@@
 J., Fierer, N., et al. (2012) Ultra-high-throughput microbial community
-analysis on the Illumina HiSeq and MiSeq platforms. \emph{The ISME
-Journal} \textbf{6}: 1621--1624.
+analysis on the Illumina HiSeq and MiSeq platforms. \emph{ISME J}
+\textbf{6}: 1621--1624.
 
@@
 Cardoso, D.C., Cretoiu, M.S., Stal, L.J., and Bolhuis, H. (2019)
-Seasonal development of a coastal microbial mat. \emph{Scientific
-Reports} \textbf{9}: 9035.
+Seasonal development of a coastal microbial mat. \emph{Sci Rep}
+\textbf{9}: 9035.
 
@@
 Crump, B.C. and Koch, E.W. (2008) Attached Bacterial Populations Shared
-by Four Species of Aquatic Angiosperms. \emph{Applied and Environmental
-Microbiology} \textbf{74}: 5948--5957.
+by Four Species of Aquatic Angiosperms. \emph{Appl Environ Microbiol}
+\textbf{74}: 5948--5957.
 
@@
 Metatranscriptomics and Amplicon Sequencing Reveal Mutualisms in
-Seagrass Microbiomes. \emph{Frontiers in Microbiology} \textbf{9}: 388.
+Seagrass Microbiomes. \emph{Front Microbiol} \textbf{9}: 388.
 
@@
 Thomas, T. (2013) The seaweed holobiont: understanding seaweed-bacteria
-interactions. \emph{FEMS Microbiology Reviews} \textbf{37}: 462--476.
+interactions. \emph{FEMS Microbiol Rev} \textbf{37}: 462--476.
 
@@
 and Stachowicz, J.J. (2017) Global-Scale Structure of the Eelgrass
-Microbiome. \emph{Applied and Environmental Microbiology} \textbf{83}:
-e03391--16.
+Microbiome. \emph{Appl Environ Microbiol} \textbf{83}: e03391--16.
 
@@
 et al. (2009) The seasonal structure of microbial communities in the
-Western English Channel. \emph{Environmental Microbiology} \textbf{11}:
+Western English Channel. \emph{Environ Microbiol} \textbf{11}:
 3132--3139.
@@
 Temperton, B., et al. (2012) Defining seasonal marine microbial
-community dynamics. \emph{The ISME Journal} \textbf{6}: 298--308.
+community dynamics. \emph{ISME J} \textbf{6}: 298--308.
 
@@
 Higashioka, Y., Kojima, H., Watanabe, T., and Fukui, M. (2015) Draft
-genome sequence of \textless{}i\textgreater{}Desulfatitalea
-tepidiphila\textless{}/i\textgreater{} S28bF\(^{T}\). \emph{Genome
-Announcements} \textbf{3}: e01326--15.
+Genome Sequence of \emph{Desulfatitalea tepidiphila}
+S28bF\textsuperscript{T}. \emph{Genome Announc} \textbf{3}: e01326--15.
 
@@
 \leavevmode\hypertarget{ref-Jorgensen1999}{}%
-Jørgensen, B.B. and Gallardo, V.A. (1999)
-\textless{}i\textgreater{}Thioploca\textless{}/i\textgreater{} spp.:
-filamentous sulfur bacteria with nitrate vacuoles. \emph{FEMS
-Microbiology Ecology} \textbf{28}: 301--313.
+Jørgensen, B.B. and Gallardo, V.A. (1999) \emph{Thioploca} spp.:
+filamentous sulfur bacteria with nitrate vacuoles. \emph{FEMS Microbiol
+Ecol} \textbf{28}: 301--313.
 
@@
 Selective DNA and Protein Isolation from Marine Macrophyte Surfaces.
+Manuscript submitted for publication.
 
@@
 (2015) Bacterial Diversity in the South Adriatic Sea during a Strong,
-Deep Winter Convection Year. \emph{Applied and Environmental
-Microbiology} \textbf{81}: 1715--1726.
+Deep Winter Convection Year. \emph{Appl Environ Microbiol} \textbf{81}:
+1715--1726.
 
@@
 Curation Pipeline for Analyzing Amplicon Sequence Data on the MiSeq
-Illumina Sequencing Platform. \emph{Applied and environmental
-microbiology} \textbf{79}: 5112--5120.
+Illumina Sequencing Platform. \emph{Appl Environ Microbiol} \textbf{79}:
+5112--5120.
 
 \leavevmode\hypertarget{ref-Kuever2014}{}%
-Kuever, J. (2014) The Family Desulfobulbaceae. In, \emph{The
-prokaryotes:
-\textless{}i\textgreater{}Deltaproteobacteria\textless{}/i\textgreater{}
-and
-\textless{}i\textgreater{}Epsilonproteobacteria\textless{}/i\textgreater{}}.
-Springer-Verlag Berlin Heidelberg, pp. 75--86.
+Kuever, J. (2014) The Family \emph{Desulfobulbaceae}. In, Rosenberg,E.,
+DeLong,E.F., Lory,S., Stackebrandt,E., and Thompson,F. (eds), \emph{The
+Prokaryotes: Deltaproteobacteria and Epsilonproteobacteria}. Berlin,
+Heidelberg: Springer-Verlag, pp. 75--86.
 
@@
 epibacterial communities on macroalgae: phylogeny matters more than
-habitat. \emph{Aquatic Biology} \textbf{5}: 181--186.
+habitat. \emph{Aquat Biol} \textbf{5}: 181--186.
 
@@
 Epibacterial community patterns on marine macroalgae are host-specific
-but temporally variable. \emph{Environmental Microbiology} \textbf{13}:
-655--665.
+but temporally variable. \emph{Environ Microbiol} \textbf{13}: 655--665.
 
@@
 Kjelleberg, S., et al. (2007) Comparisons of diversity of bacterial
-communities associated with three sessile marine eukaryotes.
-\emph{Aquatic Microbial Ecology} \textbf{48}: 217--229.
+communities associated with three sessile marine eukaryotes. \emph{Aquat
+Microb Ecol} \textbf{48}: 217--229.
 
@@
 Margulis, L. (1991) Symbiogenesis and Symbionticism. In, Margulis,L. and
-Fester,R. (eds), \emph{Symbiosis as a source of evolutionary innovation:
-Speciation and morphogenesis}. Cambridge, Massachusetts: The MIT Press,
+Fester,R. (eds), \emph{Symbiosis as a Source of Evolutionary Innovation:
+Speciation and Morphogenesis}. Cambridge, Massachusetts: The MIT Press,
 pp. 1--14.
@@
 Vertical Distribution and Phylogenetic Characterization of Marine
-Planktonic \textless{}i\textgreater{}Archaea\textless{}/i\textgreater{}
-in the Santa Barbara Channel. \emph{Applied and environmental
-microbiology} \textbf{63}: 50--56.
+Planktonic \emph{Archaea} in the Santa Barbara Channel. \emph{Appl
+Environ Microbiol} \textbf{63}: 50--56.
 
 \leavevmode\hypertarget{ref-McIlroy2014}{}%
-McIlroy, S.J. and Nielsen, P.H. (2014) The Family
-\textless{}i\textgreater{}Saprospiraceae\textless{}/i\textgreater{}. In,
-\emph{The prokaryotes: Other major lineages of
-\textless{}i\textgreater{}Bacteria\textless{}/i\textgreater{} and the
-\textless{}i\textgreater{}Archaea\textless{}/i\textgreater{}}.
-Springer-Verlag Berlin Heidelberg, pp. 863--889.
+McIlroy, S.J. and Nielsen, P.H. (2014) The Family \emph{Saprospiraceae}.
+In, Rosenberg,E., DeLong,E.F., Lory,S., Stackebrandt,E., and Thompson,F.
+(eds), \emph{The Prokaryotes: Other Major Lineages of Bacteria and The
+Archaea}. Berlin, Heidelberg: Springer-Verlag, pp. 863--889.
 
@@
 Michelou, V.K., Caporaso, J.G., Knight, R., and Palumbi, S.R. (2013) The
-Ecology of Microbial Communities Associated with Macrocystis pyrifera.
-\emph{PloS one} \textbf{8}: e67480.
+Ecology of Microbial Communities Associated with \emph{Macrocystis
+pyrifera}. \emph{PLoS One} \textbf{8}: e67480.
 
@@
 Diversity and Abundance of the Bacterial Community of the Red Macroalga
-Porphyra umbilicalis: Did Bacterial Farmers Produce Macroalgae?
-\emph{PloS one} \textbf{8}: e58269.
+\emph{Porphyra umbilicalis}: Did Bacterial Farmers Produce Macroalgae?
+\emph{PLoS One} \textbf{8}: e58269.
 
@@
 microbial diversity analyses of modern marine thrombolitic mats by
-barcoded pyrosequencing. \emph{Environmental Microbiology} \textbf{14}:
-82--100.
+barcoded pyrosequencing. \emph{Environ Microbiol} \textbf{14}: 82--100.
 
@@
 Differentiation on Bacterial Communities in Siphonous Green Seaweeds.
-\emph{Frontiers in Microbiology} \textbf{10}: 717.
+\emph{Front Microbiol} \textbf{10}: 717.
 
@@
 aquatic macrophytes and their epiphytes in two shallow lakes (Peipsi and
-Võrtsjärv) in Estonia. \emph{Aquatic Ecology} \textbf{44}: 83--92.
+Võrtsjärv) in Estonia. \emph{Aquat Ecol} \textbf{44}: 83--92.
 
@@
 with mock communities, time series and global field samples.
-\emph{Environmental Microbiology} \textbf{18}: 1403--1414.
+\emph{Environ Microbiol} \textbf{18}: 1403--1414.
 
@@
 Pujalte, M.J., Lucena, T., Ruvira, M.A., Arahal, D.R., and Macián, M.C.
-(2014) The Family Rhodobacteraceae. In, \emph{The prokaryotes:
-\textless{}i\textgreater{}Alphaproteobacteria\textless{}/i\textgreater{}
-and
-\textless{}i\textgreater{}Betaproteobacteria\textless{}/i\textgreater{}}.
-Springer-Verlag Berlin Heidelberg, pp. 439--512.
+(2014) The Family \emph{Rhodobacteraceae}. In, Rosenberg,E.,
+DeLong,E.F., Lory,S., Stackebrandt,E., and Thompson,F. (eds), \emph{The
+Prokaryotes: Alphaproteobacteria and Betaproteobacteria}. Berlin,
+Heidelberg: Springer-Verlag, pp. 439--512.
 
@@
 et al. (2013) The SILVA ribosomal RNA gene database project: improved
-data processing and web-based tools. \emph{Nucleic Acids Research}
+data processing and web-based tools. \emph{Nucleic Acids Res}
 \textbf{41}: D590--D596.
@@
 Thums, M., et al. (2014) Millimeter-Sized Marine Plastics: A New Pelagic
-Habitat for Microorganisms and Invertebrates. \emph{PloS one}
+Habitat for Microorganisms and Invertebrates. \emph{PLoS One}
 \textbf{9}: e100289.
@@
 Reyes, J. and Sansón, M. (2001) Biomass and Production of the Epiphytes
-on the Leaves of \textless{}i\textgreater{}Cymodocea
-nodosa\textless{}/i\textgreater{} in the Canary Islands. \emph{Botanica
-Marina} \textbf{44}: 307--313.
+on the Leaves of \emph{Cymodocea nodosa} in the Canary Islands.
+\emph{Bot Mar} \textbf{44}: 307--313.
 
@@
 Rickert, E., Wahl, M., Link, H., Richter, H., and Pohnert, G. (2016)
-Seasonal Variations in Surface Metabolite Composition of Fucus
-vesiculosus and Fucus serratus from the Baltic Sea. \emph{PloS one}
-\textbf{11}: e0168196.
+Seasonal Variations in Surface Metabolite Composition of \emph{Fucus
+vesiculosus} and \emph{Fucus serratus} from the Baltic Sea. \emph{PLoS
+One} \textbf{11}: e0168196.
 
@@
 (2016) Partitioning of functional and taxonomic diversity in
-surface-associated microbial communities. \emph{Environmental
-microbiology} \textbf{18}: 4391--4402.
+surface-associated microbial communities. \emph{Environ Microbiol}
+\textbf{18}: 4391--4402.
 
@@
 Ruitton, S., Verlaque, M., and Boudouresque, C.F. (2005) Seasonal
-changes of the introduced Caulerpa racemosa var. cylindracea
-(Caulerpales, Chlorophyta) at the northwest limit of its Mediterranean
-range. \emph{Aquatic Botany} \textbf{82}: 55--70.
+changes of the introduced \emph{Caulerpa racemosa} var.
+\emph{cylindracea} (Caulerpales, Chlorophyta) at the northwest limit of
+its Mediterranean range. \emph{Aquat Bot} \textbf{82}: 55--70.
 
@@
 Salaün, S., Barre, S. la, Santos-Goncalvez, M.D., Potin, P., Haras, D.,
-and Bazire, A. (2012) Influence of Exudates of the Kelp Laminaria
-Digitata on Biofilm Formation of Associated and Exogenous Bacterial
-Epiphytes. \emph{Microbial Ecology} \textbf{64}: 359--369.
+and Bazire, A. (2012) Influence of Exudates of the Kelp \emph{Laminaria
+Digitata} on Biofilm Formation of Associated and Exogenous Bacterial
+Epiphytes. \emph{Microb Ecol} \textbf{64}: 359--369.
 
@@
 Platform-Independent, Community-Supported Software for Describing and
-Comparing Microbial Communities. \emph{Applied and environmental
-microbiology} \textbf{75}: 7537--7541.
+Comparing Microbial Communities. \emph{Appl Environ Microbiol}
+\textbf{75}: 7537--7541.
 
@@
 interactions and their role in seagrass ecosystem functioning.
-\emph{FEMS Microbiology Letters} \textbf{366}: fnz057.
+\emph{FEMS Microbiol Lett} \textbf{366}: fnz057.
 
@@
 Kjelleberg, S. (2010) Variability and abundance of the epiphytic
-bacterial community associated with a green marine
-\textless{}i\textgreater{}Ulvacean\textless{}/i\textgreater{} alga.
-\emph{The ISME Journal} \textbf{4}: 301--311.
+bacterial community associated with a green marine \emph{Ulvacean} alga.
+\emph{ISME J} \textbf{4}: 301--311.
 
@@
 Ugarelli, K., Laas, P., and Stingl, U. (2019) The Microbial Communities
-of Leaves and Roots Associated with Turtle Grass
-(\textless{}i\textgreater{}Thalassia
-testudinum\textless{}/i\textgreater{}) and Manatee Grass
-(\textless{}i\textgreater{}Syringodium
-filliforme\textless{}/i\textgreater{}) are Distinct from Seawater and
-Sediment Communities, but Are Similar between Species and Sampling Si.
-\emph{Microorganisms} \textbf{7}: 4.
+of Leaves and Roots Associated with Turtle Grass (Thalassia testudinum)
+and Manatee Grass (Syringodium filliforme) are Distinct from Seawater
+and Sediment Communities, but Are Similar between Species and Sampling
+Sites. \emph{Microorganisms} \textbf{7}: 4.
 
@@
 AND COMPARISON OF PROKARYOTIC EPIPHYTES ASSOCIATED WITH THREE EAST
-AFRICAN SEAGRASSES. \emph{Journal of Phycology} \textbf{43}: 768--779.
+AFRICAN SEAGRASSES. \emph{J Phycol} \textbf{43}: 768--779.
 
@@
 of Bacterial Communities Associated with Leaves of the Seagrass
-\textless{}i\textgreater{}Halophila
-stipulacea\textless{}/i\textgreater{} by a Culture-Independent
-Small-Subunit rRNA Gene Approach. \emph{Microbial Ecology} \textbf{39}:
-22--31.
+\emph{Halophila stipulacea} by a Culture-Independent Small-Subunit rRNA
+Gene Approach. \emph{Microb Ecol} \textbf{39}: 22--31.
 
@@
 Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.D., François,
-R., et al. (2019) Welcome to the Tidyverse. \emph{Journal of Open Source
-Software} \textbf{4}: 1686.
+R., et al. (2019) Welcome to the Tidyverse. \emph{J Open Source Softw}
+\textbf{4}: 1686.
 
@@
 Xie, Y. (2014) knitr: A Comprehensive Tool for Reproducible Research in
-R. In, Stodden,V., Leisch,F., and Peng,R.D. (eds), \emph{Implementing
-reproducible computational research}. Chapman; Hall/CRC.
+R. In, Stodden,V., Leisch,F., and Peng,R.D. (eds), \emph{Implement.
+Reprod. Comput. Res.} Chapman; Hall/CRC.
 
@@
 Xie, Z., Lin, W., and Luo, J. (2017) Comparative Phenotype and Genome
-Analysis of
-\textless{}i\textgreater{}Cellvibrio\textless{}/i\textgreater{} sp. PR1,
-a Xylanolytic and Agarolytic Bacterium from the Pearl River.
-\emph{BioMed Research International} \textbf{2017}: 6304248.
+Analysis of \emph{Cellvibrio} sp. PR1, a Xylanolytic and Agarolytic
+Bacterium from the Pearl River. \emph{Biomed Res Int} \textbf{2017}:
+6304248.
 
@@
 C., et al. (2014) The SILVA and "All-species Living Tree Project (LTP)"
-taxonomic frameworks. \emph{Nucleic Acids Research} \textbf{42}:
-D643--D648.
+taxonomic frameworks. \emph{Nucleic Acids Res} \textbf{42}: D643--D648.
 
@@
 the rate of photosynthetic activity and chemical composition of the
-seagrass \textless{}i\textgreater{}Cymodocea
-nodosa\textless{}/i\textgreater{} (Ucr.) Asch. \emph{Scientia Marina}
+seagrass \emph{Cymodocea nodosa} (Ucr.) Asch. \emph{Sci Mar}
 \textbf{62}: 301--309.
@@
 \caption{Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between assemblages associated with the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed] and \textit{C.
-cylindracea} [Mixed and Monospecific]) and comunities in the ambient seawater.\label{matrix}}\label{fig:unnamed-chunk-1}
+nodosa} [Mixed Settlement] and \textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and comunities in the ambient seawater.\label{matrix}}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 \caption{Proportion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed] and \textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\label{shared}}\label{fig:unnamed-chunk-2}
+nodosa} [Mixed Settlement] and \textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.\label{shared}}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 \caption{Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed] and \textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+nodosa} [Mixed Settlement] and \textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
 Samples from the same environment or same season are labeled in different colors.
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) bacterial and archaeal sequences on the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed] and \textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+nodosa} [Mixed Settlement] and \textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
 NR -- No Relative\label{community}}\label{fig:unnamed-chunk-4}
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) cyanobacterial sequences on the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed] and \textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+nodosa} [Mixed Settlement] and \textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
 The proportion of cyanobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) sequences within the \textit{Bacteroidota} on the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed] and \textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+nodosa} [Mixed Settlement] and \textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
 The proportion of sequences classified as \textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar.
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) alphaproteobacterial sequences on the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed] and \textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+nodosa} [Mixed Settlement] and \textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
 The proportion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) gammaproteobacterial sequences on the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed] and \textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+nodosa} [Mixed Settlement] and \textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
 The proportion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) sequences within the \textit{Desulfobacterota} on the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed] and \textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.
+nodosa} [Mixed Settlement] and \textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
 The proportion of sequences classified as \textit{Desulfobacterota} in the total bacterial and archaeal community is given above the corresponding bar.

### 2020-03-20 submission/supplementary.tex
References update.

@@
 \caption{Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed] and \textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\label{rarefaction}}\label{fig:unnamed-chunk-1}
+nodosa} [Mixed Settlement] and \textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.\label{rarefaction}}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 \caption{Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed] and \textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\label{calculators}}\label{fig:unnamed-chunk-2}
+nodosa} [Mixed Settlement] and \textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.\label{calculators}}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 \caption{Relative contribution of chloroplast sequences on the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed] and \textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\label{chloroplast}}\label{fig:unnamed-chunk-3}
+nodosa} [Mixed Settlement] and \textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.\label{chloroplast}}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 \endlastfoot
-3 & Seawater & 13 July 2017 & Summer & 25,996 & 821\\
-5 & Seawater & 27 July 2017 & Summer & 58,954 & 993\\
-7 & Seawater & 10 August 2017 & Summer & 32,626 & 853\\
-9 & Seawater & 24 August 2017 & Summer & 60,935 & 1,064\\
-11 & Seawater & 19 September 2017 & Summer & 46,107 & 1,179\\
-13 & Seawater & 12 October 2017 & Autumn & 30,911 & 971\\
-15 & Seawater & 23 November 2017 & Autumn & 27,585 & 1,197\\
-17 & Seawater & 14 December 2017 & Autumn & 44,595 & 2,275\\
-19 & Seawater & 13 February 2018 & Winter & 44,207 & 1,306\\
-21 & Seawater & 27 March 2018 & Winter & 46,313 & 1,561\\
-23a & Seawater & 24 April 2018 & Spring & 30,973 & 999\\
-23b & Seawater & 24 April 2018 & Spring & 38,563 & 1,195\\
-25 & Seawater & 22 May 2018 & Spring & 53,877 & 1,787\\
-27 & Seawater & 19 June 2018 & Spring & 77,465 & 1,290\\
-29 & Seawater & 10 July 2018 & Summer & 50,793 & 1,445\\
-31 & Seawater & 9 August 2018 & Summer & 39,041 & 1,123\\
-33 & Seawater & 4 September 2018 & Summer & 36,206 & 1,206\\
-35 & Seawater & 5 October 2018 & Autumn & 49,585 & 1,019\\
-37 & \textit{Cymodocea nodosa} (Mixed) & 23 November 2017 & Autumn & 31,237 & 1,555\\
-41 & \textit{Cymodocea nodosa} (Mixed) & 4 December 2017 & Autumn & 24,234 & 1,174\\
-45 & \textit{Cymodocea nodosa} (Mixed) & 14 December 2017 & Autumn & 32,683 & 1,061\\
-49 & \textit{Cymodocea nodosa} (Mixed) & 13 February 2018 & Winter & 9,088 & 1,216\\
-52 & \textit{Cymodocea nodosa} (Mixed) & 27 March 2018 & Winter & 17,005 & 1,216\\
-55 & \textit{Cymodocea nodosa} (Mixed) & 24 April 2018 & Spring & 42,645 & 2,053\\
-58 & \textit{Cymodocea nodosa} (Mixed) & 22 May 2018 & Spring & 21,337 & 1,290\\
-61 & \textit{Cymodocea nodosa} (Mixed) & 19 June 2018 & Spring & 31,726 & 2,098\\
-64 & \textit{Cymodocea nodosa} (Mixed) & 10 July 2018 & Summer & 35,729 & 1,793\\
-67 & \textit{Cymodocea nodosa} (Mixed) & 9 August 2018 & Summer & 26,356 & 2,116\\
-70 & \textit{Cymodocea nodosa} (Mixed) & 4 September 2018 & Summer & 23,275 & 1,724\\
-73 & \textit{Cymodocea nodosa} (Mixed) & 5 October 2018 & Autumn & 29,907 & 2,211\\
-38 & \textit{Caulerpa cylindracea} (Mixed) & 23 November 2017 & Autumn & 36,309 & 2,896\\
-42 & \textit{Caulerpa cylindracea} (Mixed) & 4 December 2017 & Autumn & 28,383 & 3,249\\
-46 & \textit{Caulerpa cylindracea} (Mixed) & 14 December 2017 & Autumn & 34,718 & 3,067\\
-53 & \textit{Caulerpa cylindracea} (Mixed) & 27 March 2018 & Winter & 28,689 & 2,481\\
-56 & \textit{Caulerpa cylindracea} (Mixed) & 24 April 2018 & Spring & 34,784 & 3,057\\
-59 & \textit{Caulerpa cylindracea} (Mixed) & 22 May 2018 & Spring & 23,392 & 2,712\\
-62 & \textit{Caulerpa cylindracea} (Mixed) & 19 June 2018 & Spring & 36,488 & 3,311\\
-65 & \textit{Caulerpa cylindracea} (Mixed) & 10 July 2018 & Summer & 18,483 & 2,188\\
-68 & \textit{Caulerpa cylindracea} (Mixed) & 9 August 2018 & Summer & 31,943 & 3,097\\
-71 & \textit{Caulerpa cylindracea} (Mixed) & 4 September 2018 & Summer & 29,273 & 3,152\\
-74 & \textit{Caulerpa cylindracea} (Mixed) & 5 October 2018 & Autumn & 11,707 & 1,711\\
-39 & \textit{Caulerpa cylindracea} (Monospecific) & 23 November 2017 & Autumn & 22,057 & 2,428\\
+3 & Seawater & 13 July 2017 & Summer & 25,996 & 824\\
+5 & Seawater & 27 July 2017 & Summer & 58,950 & 989\\
+7 & Seawater & 10 August 2017 & Summer & 32,631 & 851\\
+9 & Seawater & 24 August 2017 & Summer & 60,935 & 1,060\\
+11 & Seawater & 19 September 2017 & Summer & 46,115 & 1,170\\
+13 & Seawater & 12 October 2017 & Autumn & 30,913 & 973\\
+15 & Seawater & 23 November 2017 & Autumn & 27,586 & 1,187\\
+17 & Seawater & 14 December 2017 & Autumn & 44,584 & 2,270\\
+19 & Seawater & 13 February 2018 & Winter & 44,185 & 1,313\\
+21 & Seawater & 27 March 2018 & Winter & 46,375 & 1,573\\
+23a & Seawater & 24 April 2018 & Spring & 30,968 & 996\\
+23b & Seawater & 24 April 2018 & Spring & 38,572 & 1,200\\
+25 & Seawater & 22 May 2018 & Spring & 53,871 & 1,796\\
+27 & Seawater & 19 June 2018 & Spring & 77,463 & 1,292\\
+29 & Seawater & 10 July 2018 & Summer & 50,811 & 1,444\\
+31 & Seawater & 9 August 2018 & Summer & 39,051 & 1,120\\
+33 & Seawater & 4 September 2018 & Summer & 36,200 & 1,202\\
+35 & Seawater & 5 October 2018 & Autumn & 49,568 & 1,012\\
+37 & \textit{Cymodocea nodosa} (Mixed) & 23 November 2017 & Autumn & 31,236 & 1,552\\
+41 & \textit{Cymodocea nodosa} (Mixed) & 4 December 2017 & Autumn & 24,252 & 1,184\\
+45 & \textit{Cymodocea nodosa} (Mixed) & 14 December 2017 & Autumn & 32,681 & 1,053\\
+49 & \textit{Cymodocea nodosa} (Mixed) & 13 February 2018 & Winter & 9,089 & 1,213\\
+52 & \textit{Cymodocea nodosa} (Mixed) & 27 March 2018 & Winter & 17,002 & 1,211\\
+55 & \textit{Cymodocea nodosa} (Mixed) & 24 April 2018 & Spring & 42,642 & 2,058\\
+58 & \textit{Cymodocea nodosa} (Mixed) & 22 May 2018 & Spring & 21,337 & 1,294\\
+61 & \textit{Cymodocea nodosa} (Mixed) & 19 June 2018 & Spring & 31,724 & 2,111\\
+64 & \textit{Cymodocea nodosa} (Mixed) & 10 July 2018 & Summer & 35,738 & 1,794\\
+67 & \textit{Cymodocea nodosa} (Mixed) & 9 August 2018 & Summer & 26,355 & 2,119\\
+70 & \textit{Cymodocea nodosa} (Mixed) & 4 September 2018 & Summer & 23,275 & 1,723\\
+73 & \textit{Cymodocea nodosa} (Mixed) & 5 October 2018 & Autumn & 29,911 & 2,216\\
+38 & \textit{Caulerpa cylindracea} (Mixed) & 23 November 2017 & Autumn & 36,311 & 2,886\\
+42 & \textit{Caulerpa cylindracea} (Mixed) & 4 December 2017 & Autumn & 28,381 & 3,235\\
+46 & \textit{Caulerpa cylindracea} (Mixed) & 14 December 2017 & Autumn & 34,724 & 3,074\\
+53 & \textit{Caulerpa cylindracea} (Mixed) & 27 March 2018 & Winter & 28,694 & 2,483\\
+56 & \textit{Caulerpa cylindracea} (Mixed) & 24 April 2018 & Spring & 34,780 & 3,061\\
+59 & \textit{Caulerpa cylindracea} (Mixed) & 22 May 2018 & Spring & 23,393 & 2,720\\
+62 & \textit{Caulerpa cylindracea} (Mixed) & 19 June 2018 & Spring & 36,491 & 3,316\\
+65 & \textit{Caulerpa cylindracea} (Mixed) & 10 July 2018 & Summer & 18,485 & 2,186\\
+68 & \textit{Caulerpa cylindracea} (Mixed) & 9 August 2018 & Summer & 31,956 & 3,112\\
+71 & \textit{Caulerpa cylindracea} (Mixed) & 4 September 2018 & Summer & 29,273 & 3,153\\
+74 & \textit{Caulerpa cylindracea} (Mixed) & 5 October 2018 & Autumn & 11,692 & 1,703\\
+39 & \textit{Caulerpa cylindracea} (Monospecific) & 23 November 2017 & Autumn & 22,081 & 2,432\\
 43 & \textit{Caulerpa cylindracea} (Monospecific) & 4 December 2017 & Autumn & 13,661 & 1,892\\
-47 & \textit{Caulerpa cylindracea} (Monospecific) & 14 December 2017 & Autumn & 8,409 & 1,451\\
-51 & \textit{Caulerpa cylindracea} (Monospecific) & 13 February 2018 & Winter & 11,673 & 1,899\\
-54 & \textit{Caulerpa cylindracea} (Monospecific) & 27 March 2018 & Winter & 39,471 & 3,142\\
-57 & \textit{Caulerpa cylindracea} (Monospecific) & 24 April 2018 & Spring & 20,304 & 2,836\\
-60 & \textit{Caulerpa cylindracea} (Monospecific) & 22 May 2018 & Spring & 33,035 & 3,314\\
-63 & \textit{Caulerpa cylindracea} (Monospecific) & 19 June 2018 & Spring & 41,841 & 3,976\\
-66 & \textit{Caulerpa cylindracea} (Monospecific) & 10 July 2018 & Summer & 27,038 & 2,669\\
-69 & \textit{Caulerpa cylindracea} (Monospecific) & 9 August 2018 & Summer & 26,734 & 3,123\\
-72 & \textit{Caulerpa cylindracea} (Monospecific) & 4 September 2018 & Summer & 31,874 & 3,248\\
-75 & \textit{Caulerpa cylindracea} (Monospecific) & 5 October 2018 & Autumn & 33,073 & 3,078\\*
+47 & \textit{Caulerpa cylindracea} (Monospecific) & 14 December 2017 & Autumn & 8,409 & 1,454\\
+51 & \textit{Caulerpa cylindracea} (Monospecific) & 13 February 2018 & Winter & 11,674 & 1,908\\
+54 & \textit{Caulerpa cylindracea} (Monospecific) & 27 March 2018 & Winter & 39,470 & 3,133\\
+57 & \textit{Caulerpa cylindracea} (Monospecific) & 24 April 2018 & Spring & 20,296 & 2,833\\
+60 & \textit{Caulerpa cylindracea} (Monospecific) & 22 May 2018 & Spring & 33,034 & 3,315\\
+63 & \textit{Caulerpa cylindracea} (Monospecific) & 19 June 2018 & Spring & 41,842 & 3,970\\
+66 & \textit{Caulerpa cylindracea} (Monospecific) & 10 July 2018 & Summer & 27,040 & 2,675\\
+69 & \textit{Caulerpa cylindracea} (Monospecific) & 9 August 2018 & Summer & 26,735 & 3,120\\
+72 & \textit{Caulerpa cylindracea} (Monospecific) & 4 September 2018 & Summer & 31,871 & 3,264\\
+75 & \textit{Caulerpa cylindracea} (Monospecific) & 5 October 2018 & Autumn & 33,077 & 3,078\\*
 \end{longtable}

### 2020-03-20 code/plot_community_bar_plot.R

@@ -58,10 +58,10 @@ color <- read_tsv("data/raw/group_colors.csv") %>%
 
 # Generation of italic names for groups
 names <- parse(text=case_when(plot$taxon=="Chloroplast" ~ paste0("plain('", plot$taxon,  "')"),
-                              plot$taxon=="Bacteria_unclassified" ~ "italic('Bacteria')~plain('(No Relative)')",
+                              plot$taxon=="Bacteria_unclassified" ~ "italic('Bacteria')~plain('(NR)')",
                               plot$taxon=="Marinimicrobia_(SAR406_clade)" ~ "italic('Marinimicrobia')",
                               plot$taxon=="Other" ~ paste0("plain('", plot$taxon, "')"),
-                              plot$taxon=="No_Relative" ~ "plain('NR')",
+                              plot$taxon=="No_Relative" ~ "plain('No Relative')",
                               TRUE ~ paste0("italic('", plot$taxon, "')")))
 
 # Tidying the sequence abundance data

### 2020-03-24 submission/manuscript.Rmd
References checked.

@@
 
-\vspace{50mm}
+\vspace{20mm}
 Marino Korlević^1$*$^, Marsej Markovski^1^, Zihao Zhao^2^, Gerhard J. Herndl^2,3^, Mirjana Najdek^1^
 
-\vspace{40mm}
-
-^$*$^To whom correspondence should be addressed: marino.korlevic@irb.hr
-
 1\. Center for Marine Research, Ruđer Bošković Institute, Croatia
@@
 3\. Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research, Utrecht University, The Netherlands
+
+^$*$^To whom correspondence should be addressed:
+
+Marino Korlević
+
+G. Paliaga 5, 52210 Rovinj, Croatia
+
+Tel.: +385 52 804 768
+
+Fax: +385 52 804 780
+
+e-mail: marino.korlevic@irb.hr
+
+Running title: Seasonal dynamics of epiphytic communities
 
@@
 
-## Abstract
+## Summary
 Surfaces of marine macrophytes (seagrasses and macroalgae) are inhabited by diverse microbial communities.
@@
 It postulates that an initial random colonization step is performed from a set of functionally equivalent taxonomic groups resulting in taxonomically different epiphytic communities sharing a core set of functional genes [@Burke2011a; @Roth-Schulze2016].
-In addition, some of the variation in the observed data could be attributed to different techniques used in various studies, such as different protocols for epiphytic cell detachment and/or DNA isolation, as no standard protocol to study epiphytic communities was established [@Ugarelli2019; @Korlevic_submitted, submitted].
+In addition, some of the variation in the observed data could be attributed to different techniques used in various studies, such as different protocols for epiphytic cell detachment and/or DNA isolation, as no standard protocol to study epiphytic communities was established [Korlević *et al.*, unpublished; @Ugarelli2019].
 
@@
 All these methods showed different removal efficiencies but none was enabling a complete removal of attached microbial cells.
-In the present study, we applied an earlier developed removal protocol [@Korlevic_submitted, submitted], based on a previous idea of direct cellular lysis [@Burke2009], to ensure an almost complete cell detachment.
+In the present study, we applied an earlier developed removal protocol (Korlević *et al.*, unpublished), based on a previous idea of direct cellular lysis [@Burke2009], to ensure an almost complete cell detachment.
 The application of a direct lysis procedure coupled with a high frequency sampling protocol and Illumina high resolution amplicon sequencing has enabled us to make a detailed description of bacterial and archaeal communities associated with the surfaces of two marine macrophytes, *C.
@@
 \newpage
-## Experimental Procedures
+## Experimental procedures
 ### Sampling
@@
 
-### DNA Isolation
+### DNA isolation
 DNA from surfaces of *C.
 nodosa* and *C.
-cylindracea* was isolated using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevic_submitted, submitted; @Massana1997].
+cylindracea* was isolated using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [Korlević *et al.*, unpublished; @Massana1997].
 Briefly, leaves and thalli are incubated in a lysis buffer and treated with lysozyme and proteinase K.
@@
 
-### Illumina 16S rRNA Sequencing
-Illumina MiSeq sequencing of the V4 16S rRNA region was performed as described previously [@Korlevic_submitted, submitted].
+### Illumina 16S rRNA sequencing
+Illumina MiSeq sequencing of the V4 16S rRNA region was performed as described previously (Korlević *et al.*, unpublished).
 The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
@@
 
-### Sequence Analysis
+### Sequence analysis
 ```{r error}
@@
 
-## Figure Captions
+## Figure legends
 **\autoref{matrix}.** \nameref{matrix}

### 2020-03-24 submission/manuscript.tex
References checked.

@@
 \renewcommand{\thefigure}{\arabic{figure}}
+\renewcommand{\figurename}{Fig.}
 \captionsetup{justification=raggedright,singlelinecheck=false}
@@
 
+\renewcommand{\figureautorefname}{Fig.}
+
 \begin{document}
@@
 
-\vspace{50mm}
+\vspace{20mm}
 
@@
 
-\vspace{40mm}
-
-\textsuperscript{\(*\)}To whom correspondence should be addressed:
-\href{mailto:marino.korlevic@irb.hr}{\nolinkurl{marino.korlevic@irb.hr}}
-
 1. Center for Marine Research, Ruđer Bošković Institute, Croatia
@@
 Netherlands
+
+\textsuperscript{\(*\)}To whom correspondence should be addressed:
+
+Marino Korlević
+
+G. Paliaga 5, 52210 Rovinj, Croatia
+
+Tel.: +385 52 804 768
+
+Fax: +385 52 804 780
+
+e-mail:
+\href{mailto:marino.korlevic@irb.hr}{\nolinkurl{marino.korlevic@irb.hr}}
+
+Running title: Seasonal dynamics of epiphytic communities
 
@@
 
-\hypertarget{abstract}{%
-\subsection{Abstract}\label{abstract}}
+\hypertarget{summary}{%
+\subsection{Summary}\label{summary}}
 
@@
 isolation, as no standard protocol to study epiphytic communities was
-established (Ugarelli \emph{et al.}, 2019; Korlević \emph{et al.},
-submitted).
+established (Korlević \emph{et al.}, unpublished; Ugarelli \emph{et
+al.}, 2019).
 
@@
 normalized to the minimum number of sequences, 8,409
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,172
-different OTUs that ranged from 366 to 2,043 OTUs per sample
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,055
+different OTUs that ranged from 370 to 2,033 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,683.2 ±
-130.9 OTUs{]} and monospecific {[}Number of OTUs, 1,737.2 ± 166.4
-OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,058.4 ±
-216.8 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 529.5 ± 143.0 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,691.5 ±
+144.4 OTUs{]} and monospecific {[}Number of OTUs, 1,733.3 ± 156.8
+OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,060.9 ±
+216.1 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 527.5 ± 147.8 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 community was found between mixed and monospecific \emph{C. cylindracea}
-(Jaccard, 0.35; Bray-Curtis, 0.78), while lower shared values were
+(Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared values were
 calculated between seawater and epiphytic communities
@@
 from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (40.8 -- 52.2
+These OTUs also made a high proportion of total sequences (40.3 -- 53.7
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 communities indicated a separation between Spring, Summer and
-Autumn/Winter samples (ANOSIM, R = 0.53, \emph{p} \textless{} 0.01). For
+Autumn/Winter samples (ANOSIM, R = 0.54, \emph{p} \textless{} 0.01). For
 \emph{C. cylindracea} associated communities a separation between Summer
 and Autumn/Winter/Spring samples was observed that was not so strongly
-supported (ANOSIM, R = 0.30, \emph{p} \textless{} 0.01)
+supported (ANOSIM, R = 0.31, \emph{p} \textless{} 0.05)
 (\autoref{pcoa}).
@@
 study, we applied an earlier developed removal protocol (Korlević
-\emph{et al.}, submitted), based on a previous idea of direct cellular
+\emph{et al.}, unpublished), based on a previous idea of direct cellular
 lysis (Burke \emph{et al.}, 2009), to ensure an almost complete cell
@@
 sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 52.2
+were accounting for a high proportion of sequences (\(\leq\) 53.7
 \si{\percent}). A very similar proportion of persistent OTUs and their
@@
 \hypertarget{experimental-procedures}{%
-\subsection{Experimental Procedures}\label{experimental-procedures}}
+\subsection{Experimental procedures}\label{experimental-procedures}}
 
@@
 \hypertarget{dna-isolation}{%
-\subsubsection{DNA Isolation}\label{dna-isolation}}
+\subsubsection{DNA isolation}\label{dna-isolation}}
 
@@
 isolated using a previously modified and adapted protocol that allows
-for a selective epiphytic DNA isolation (Massana \emph{et al.}, 1997;
-Korlević \emph{et al.}, submitted). Briefly, leaves and thalli are
-incubated in a lysis buffer and treated with lysozyme and proteinase K.
-Following the incubations, the mixture containing lyzed epiphytic cells
-is separated from leaves and thalli and extracted using a
+for a selective epiphytic DNA isolation (Korlević \emph{et al.},
+unpublished; Massana \emph{et al.}, 1997). Briefly, leaves and thalli
+are incubated in a lysis buffer and treated with lysozyme and proteinase
+K. Following the incubations, the mixture containing lyzed epiphytic
+cells is separated from leaves and thalli and extracted using a
 phenol-chloroform procedure. Finally, the extracted DNA is precipitated
@@
 \subsubsection{Illumina 16S rRNA
-Sequencing}\label{illumina-16s-rrna-sequencing}}
+sequencing}\label{illumina-16s-rrna-sequencing}}
 
 Illumina MiSeq sequencing of the V4 16S rRNA region was performed as
-described previously (Korlević \emph{et al.}, submitted). The V4 region
-of the 16S rRNA gene was amplified using a two-step PCR procedure. In
-the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R
-(5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project
+described previously (Korlević \emph{et al.}, unpublished). The V4
+region of the 16S rRNA gene was amplified using a two-step PCR
+procedure. In the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and
+806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome
+Project
 (\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
@@
 \hypertarget{sequence-analysis}{%
-\subsubsection{Sequence Analysis}\label{sequence-analysis}}
+\subsubsection{Sequence analysis}\label{sequence-analysis}}
 
@@
 \leavevmode\hypertarget{ref-Abraham2014}{}%
-Abraham, W.R. and Rohde, M. (2014) The Family \emph{Hyphomonadaceae}.
-In, Rosenberg,E., DeLong,E.F., Lory,S., Stackebrandt,E., and Thompson,F.
-(eds), \emph{The Prokaryotes: Alphaproteobacteria and
-Betaproteobacteria}. Berlin, Heidelberg: Springer-Verlag, pp. 283--299.
+Abraham, W.R. and Rohde, M. (2014) The family \emph{Hyphomonadaceae}. In
+\emph{The prokaryotes: Alphaproteobacteria and Betaproteobacteria}.
+Rosenberg, E., DeLong, E.F., Lory, S., Stackebrandt, E., and Thompson,
+F. (eds). Berlin, Heidelberg: Springer-Verlag, pp. 283--299.
 
@@
 Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
-A., et al. (2019) rmarkdown: Dynamic Documents for R.
+A., et al. (2019) Rmarkdown: dynamic documents for R.
 
@@
 \leavevmode\hypertarget{ref-Armstrong2000}{}%
-Armstrong, E., Rogerson, A., and Leftley, J. (2000) The Abundance of
-Heterotrophic Protists Associated with Intertidal Seaweeds. \emph{Estuar
+Armstrong, E., Rogerson, A., and Leftley, J. (2000) The abundance of
+heterotrophic protists associated with intertidal seaweeds. \emph{Estuar
 Coast Shelf Sci} \textbf{50}: 415--424.
@@
 Bengtsson, M.M., Bühler, A., Brauer, A., Dahlke, S., Schubert, H., and
-Blindow, I. (2017) Eelgrass Leaf Surface Microbiomes Are Locally
-Variable and Highly Correlated with Epibiotic Eukaryotes. \emph{Front
+Blindow, I. (2017) Eelgrass leaf surface microbiomes are locally
+variable and highly correlated with epibiotic eukaryotes. \emph{Front
 Microbiol} \textbf{8}: 1312.
@@
 \leavevmode\hypertarget{ref-Burke2009}{}%
-Burke, C., Kjelleberg, S., and Thomas, T. (2009) Selective Extraction of
-Bacterial DNA from the Surfaces of Macroalgae. \emph{Appl Environ
+Burke, C., Kjelleberg, S., and Thomas, T. (2009) Selective extraction of
+bacterial DNA from the surfaces of macroalgae. \emph{Appl Environ
 Microbiol} \textbf{75}: 252--256.
@@
 \leavevmode\hypertarget{ref-Crump2008}{}%
-Crump, B.C. and Koch, E.W. (2008) Attached Bacterial Populations Shared
-by Four Species of Aquatic Angiosperms. \emph{Appl Environ Microbiol}
+Crump, B.C. and Koch, E.W. (2008) Attached bacterial populations shared
+by four species of aquatic angiosperms. \emph{Appl Environ Microbiol}
 \textbf{74}: 5948--5957.
@@
 Crump, B.C., Wojahn, J.M., Tomas, F., and Mueller, R.S. (2018)
-Metatranscriptomics and Amplicon Sequencing Reveal Mutualisms in
-Seagrass Microbiomes. \emph{Front Microbiol} \textbf{9}: 388.
+Metatranscriptomics and amplicon sequencing reveal mutualisms in
+seagrass microbiomes. \emph{Front Microbiol} \textbf{9}: 388.
 
@@
 Fahimipour, A.K., Kardish, M.R., Lang, J.M., Green, J.L., Eisen, J.A.,
-and Stachowicz, J.J. (2017) Global-Scale Structure of the Eelgrass
-Microbiome. \emph{Appl Environ Microbiol} \textbf{83}: e03391--16.
+and Stachowicz, J.J. (2017) Global-scale structure of the eelgrass
+microbiome. \emph{Appl Environ Microbiol} \textbf{83}: e03391--16.
 
@@
 Higashioka, Y., Kojima, H., Watanabe, T., and Fukui, M. (2015) Draft
-Genome Sequence of \emph{Desulfatitalea tepidiphila}
+genome sequence of \emph{Desulfatitalea tepidiphila}
 S28bF\textsuperscript{T}. \emph{Genome Announc} \textbf{3}: e01326--15.
@@
 
-\leavevmode\hypertarget{ref-Korlevic_submitted}{}%
-Korlević, M., Markovski, M., Zhao, Z., Herndl, G.J., and Najdek, M.
-Selective DNA and Protein Isolation from Marine Macrophyte Surfaces.
-Manuscript submitted for publication.
-
 \leavevmode\hypertarget{ref-Korlevic2015}{}%
 Korlević, M., Pop Ristova, P., Garić, R., Amann, R., and Orlić, S.
-(2015) Bacterial Diversity in the South Adriatic Sea during a Strong,
-Deep Winter Convection Year. \emph{Appl Environ Microbiol} \textbf{81}:
+(2015) Bacterial diversity in the South Adriatic Sea during a strong,
+deep winter convection year. \emph{Appl Environ Microbiol} \textbf{81}:
 1715--1726.
@@
 Kozich, J.J., Westcott, S.L., Baxter, N.T., Highlander, S.K., and
-Schloss, P.D. (2013) Development of a Dual-Index Sequencing Strategy and
-Curation Pipeline for Analyzing Amplicon Sequence Data on the MiSeq
-Illumina Sequencing Platform. \emph{Appl Environ Microbiol} \textbf{79}:
+Schloss, P.D. (2013) Development of a dual-index sequencing strategy and
+curation pipeline for analyzing amplicon sequence data on the MiSeq
+Illumina sequencing platform. \emph{Appl Environ Microbiol} \textbf{79}:
 5112--5120.
@@
 \leavevmode\hypertarget{ref-Kuever2014}{}%
-Kuever, J. (2014) The Family \emph{Desulfobulbaceae}. In, Rosenberg,E.,
-DeLong,E.F., Lory,S., Stackebrandt,E., and Thompson,F. (eds), \emph{The
-Prokaryotes: Deltaproteobacteria and Epsilonproteobacteria}. Berlin,
-Heidelberg: Springer-Verlag, pp. 75--86.
+Kuever, J. (2014) The family \emph{Desulfobulbaceae}. In \emph{The
+prokaryotes: Deltaproteobacteria and Epsilonproteobacteria}. Rosenberg,
+E., DeLong, E.F., Lory, S., Stackebrandt, E., and Thompson, F. (eds).
+Berlin, Heidelberg: Springer-Verlag, pp. 75--86.
 
@@
 \leavevmode\hypertarget{ref-Margulis1991}{}%
-Margulis, L. (1991) Symbiogenesis and Symbionticism. In, Margulis,L. and
-Fester,R. (eds), \emph{Symbiosis as a Source of Evolutionary Innovation:
-Speciation and Morphogenesis}. Cambridge, Massachusetts: The MIT Press,
-pp. 1--14.
+Margulis, L. (1991) Symbiogenesis and symbionticism. In \emph{Symbiosis
+as a source of evolutionary innovation: Speciation and morphogenesis}.
+Margulis, L. and Fester, R. (eds). Cambridge, Massachusetts: The MIT
+Press, pp. 1--14.
 
@@
 Massana, R., Murray, A.E., Preston, C.M., and DeLong, E.F. (1997)
-Vertical Distribution and Phylogenetic Characterization of Marine
-Planktonic \emph{Archaea} in the Santa Barbara Channel. \emph{Appl
+Vertical distribution and phylogenetic characterization of marine
+planktonic \emph{Archaea} in the Santa Barbara Channel. \emph{Appl
 Environ Microbiol} \textbf{63}: 50--56.
@@
 \leavevmode\hypertarget{ref-McIlroy2014}{}%
-McIlroy, S.J. and Nielsen, P.H. (2014) The Family \emph{Saprospiraceae}.
-In, Rosenberg,E., DeLong,E.F., Lory,S., Stackebrandt,E., and Thompson,F.
-(eds), \emph{The Prokaryotes: Other Major Lineages of Bacteria and The
-Archaea}. Berlin, Heidelberg: Springer-Verlag, pp. 863--889.
+McIlroy, S.J. and Nielsen, P.H. (2014) The family \emph{Saprospiraceae}.
+In \emph{The prokaryotes: Other major lineages of Bacteria and the
+Archaea}. Rosenberg, E., DeLong, E.F., Lory, S., Stackebrandt, E., and
+Thompson, F. (eds). Berlin, Heidelberg: Springer-Verlag, pp. 863--889.
 
@@
 Michelou, V.K., Caporaso, J.G., Knight, R., and Palumbi, S.R. (2013) The
-Ecology of Microbial Communities Associated with \emph{Macrocystis
+ecology of microbial communities associated with \emph{Macrocystis
 pyrifera}. \emph{PLoS One} \textbf{8}: e67480.
@@
 Miranda, L.N., Hutchison, K., Grossman, A.R., and Brawley, S.H. (2013)
-Diversity and Abundance of the Bacterial Community of the Red Macroalga
-\emph{Porphyra umbilicalis}: Did Bacterial Farmers Produce Macroalgae?
+Diversity and abundance of the bacterial community of the red macroalga
+\emph{Porphyra umbilicalis}: did bacterial farmers produce macroalgae?
 \emph{PLoS One} \textbf{8}: e58269.
@@
 Morrissey, K.L., Çavas, L., Willems, A., and De Clerck, O. (2019)
-Disentangling the Influence of Environment, Host Specificity and Thallus
-Differentiation on Bacterial Communities in Siphonous Green Seaweeds.
+Disentangling the influence of environment, host specificity and thallus
+differentiation on bacterial communities in siphonous green seaweeds.
 \emph{Front Microbiol} \textbf{10}: 717.
@@
 \leavevmode\hypertarget{ref-Neuwirth2014}{}%
-Neuwirth, E. (2014) RColorBrewer: ColorBrewer Palettes.
+Neuwirth, E. (2014) RColorBrewer: ColorBrewer palettes.
 
@@
 Oksanen, J., Blanchet, F.G., Friendly, M., Kindt, R., Legendre, P.,
-McGlinn, D., et al. (2019) vegan: Community Ecology Package.
+McGlinn, D., et al. (2019) Vegan: community ecology package.
 
@@
 Pujalte, M.J., Lucena, T., Ruvira, M.A., Arahal, D.R., and Macián, M.C.
-(2014) The Family \emph{Rhodobacteraceae}. In, Rosenberg,E.,
-DeLong,E.F., Lory,S., Stackebrandt,E., and Thompson,F. (eds), \emph{The
-Prokaryotes: Alphaproteobacteria and Betaproteobacteria}. Berlin,
+(2014) The family \emph{Rhodobacteraceae}. In \emph{The prokaryotes:
+Alphaproteobacteria and Betaproteobacteria}. Rosenberg, E., DeLong,
+E.F., Lory, S., Stackebrandt, E., and Thompson, F. (eds). Berlin,
 Heidelberg: Springer-Verlag, pp. 439--512.
@@
 \leavevmode\hypertarget{ref-RCoreTeam2019}{}%
-R Core Team (2019) R: A Language and Environment for Statistical
-Computing, Vienna, Austria: R Foundation for Statistical Computing.
+R Core Team (2019) A language and environment for statistical computing.
 
@@
 Reisser, J., Shaw, J., Hallegraeff, G., Proietti, M., Barnes, D.K.A.,
-Thums, M., et al. (2014) Millimeter-Sized Marine Plastics: A New Pelagic
-Habitat for Microorganisms and Invertebrates. \emph{PLoS One}
+Thums, M., et al. (2014) Millimeter-sized marine plastics: a new pelagic
+habitat for microorganisms and invertebrates. \emph{PLoS One}
 \textbf{9}: e100289.
@@
 \leavevmode\hypertarget{ref-Reyes2001}{}%
-Reyes, J. and Sansón, M. (2001) Biomass and Production of the Epiphytes
-on the Leaves of \emph{Cymodocea nodosa} in the Canary Islands.
+Reyes, J. and Sansón, M. (2001) Biomass and production of the epiphytes
+on the leaves of \emph{Cymodocea nodosa} in the Canary Islands.
 \emph{Bot Mar} \textbf{44}: 307--313.
@@
 Rickert, E., Wahl, M., Link, H., Richter, H., and Pohnert, G. (2016)
-Seasonal Variations in Surface Metabolite Composition of \emph{Fucus
+Seasonal variations in surface metabolite composition of \emph{Fucus
 vesiculosus} and \emph{Fucus serratus} from the Baltic Sea. \emph{PLoS
@@
 \leavevmode\hypertarget{ref-Salaun2012}{}%
-Salaün, S., Barre, S. la, Santos-Goncalvez, M.D., Potin, P., Haras, D.,
-and Bazire, A. (2012) Influence of Exudates of the Kelp \emph{Laminaria
-Digitata} on Biofilm Formation of Associated and Exogenous Bacterial
-Epiphytes. \emph{Microb Ecol} \textbf{64}: 359--369.
+Salaün, S., La Barre, S., Santos-Goncalvez, M.D., Potin, P., Haras, D.,
+and Bazire, A. (2012) Influence of exudates of the kelp \emph{Laminaria
+digitata} on biofilm formation of associated and exogenous bacterial
+epiphytes. \emph{Microb Ecol} \textbf{64}: 359--369.
 
@@
 Schloss, P.D., Westcott, S.L., Ryabin, T., Hall, J.R., Hartmann, M.,
-Hollister, E.B., et al. (2009) Introducing mothur: Open-Source,
-Platform-Independent, Community-Supported Software for Describing and
-Comparing Microbial Communities. \emph{Appl Environ Microbiol}
+Hollister, E.B., et al. (2009) Introducing mothur: open-source,
+platform-independent, community-supported software for describing and
+comparing microbial communities. \emph{Appl Environ Microbiol}
 \textbf{75}: 7537--7541.
@@
 \leavevmode\hypertarget{ref-Ugarelli2019}{}%
-Ugarelli, K., Laas, P., and Stingl, U. (2019) The Microbial Communities
-of Leaves and Roots Associated with Turtle Grass (Thalassia testudinum)
-and Manatee Grass (Syringodium filliforme) are Distinct from Seawater
-and Sediment Communities, but Are Similar between Species and Sampling
-Sites. \emph{Microorganisms} \textbf{7}: 4.
+Ugarelli, K., Laas, P., and Stingl, U. (2019) The microbial communities
+of leaves and roots associated with turtle grass (\emph{Thalassia
+testudinum}) and manatee grass (\emph{Syringodium filliforme}) are
+distinct from seawater and sediment communities, but are similar between
+species and sampling sites. \emph{Microorganisms} \textbf{7}: 4.
 
 \leavevmode\hypertarget{ref-Uku2007}{}%
-Uku, J., Björk, M., Bergman, B., and Díez, B. (2007) CHARACTERIZATION
-AND COMPARISON OF PROKARYOTIC EPIPHYTES ASSOCIATED WITH THREE EAST
-AFRICAN SEAGRASSES. \emph{J Phycol} \textbf{43}: 768--779.
+Uku, J., Björk, M., Bergman, B., and Díez, B. (2007) Characterization
+and comparison of prokaryotic epiphytes associated with three East
+African seagrasses. \emph{J Phycol} \textbf{43}: 768--779.
 
 \leavevmode\hypertarget{ref-Weidner2000}{}%
-Weidner, Arnold, Stackebrandt, and Pühler (2000) Phylogenetic Analysis
-of Bacterial Communities Associated with Leaves of the Seagrass
-\emph{Halophila stipulacea} by a Culture-Independent Small-Subunit rRNA
-Gene Approach. \emph{Microb Ecol} \textbf{39}: 22--31.
+Weidner, S., Arnold, W., Stackebrandt, E., and Pühler, A. (2000)
+Phylogenetic analysis of bacterial communities associated with leaves of
+the seagrass \emph{Halophila stipulacea} by a culture-independent
+small-subunit rRNA gene approach. \emph{Microb Ecol} \textbf{39}:
+22--31.
 
@@
 Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.D., François,
-R., et al. (2019) Welcome to the Tidyverse. \emph{J Open Source Softw}
+R., et al. (2019) Welcome to the tidyverse. \emph{J Open Source Softw}
 \textbf{4}: 1686.
@@
 \leavevmode\hypertarget{ref-Xie2015}{}%
-Xie, Y. (2015) Dynamic Documents with R and knitr, 2nd ed. Boca Raton,
+Xie, Y. (2015) Dynamic documents with R and knitr, 2nd ed. Boca Raton,
 Florida: Chapman; Hall/CRC.
@@
 \leavevmode\hypertarget{ref-Xie2014}{}%
-Xie, Y. (2014) knitr: A Comprehensive Tool for Reproducible Research in
-R. In, Stodden,V., Leisch,F., and Peng,R.D. (eds), \emph{Implement.
-Reprod. Comput. Res.} Chapman; Hall/CRC.
+Xie, Y. (2014) Knitr: a comprehensive tool for reproducible research in
+R. In \emph{Implementing reproducible computational research}. Stodden,
+V., Leisch, F., and Peng, R.D. (eds). New York: Chapman; Hall/CRC, pp.
+3--32.
 
 \leavevmode\hypertarget{ref-Xie2019a}{}%
-Xie, Y. (2019a) knitr: A General-Purpose Package for Dynamic Report
-Generation in R.
+Xie, Y. (2019a) Knitr: a general-purpose package for dynamic report
+generation in R.
 
 \leavevmode\hypertarget{ref-Xie2019}{}%
-Xie, Y. (2019b) TinyTeX: A lightweight, cross-platform, and
+Xie, Y. (2019b) TinyTeX: a lightweight, cross-platform, and
 easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
-30--32.
+\textbf{40}: 30--32.
 
 \leavevmode\hypertarget{ref-Xie2020}{}%
-Xie, Y. (2020) tinytex: Helper Functions to Install and Maintain 'TeX
-Live', and Compile 'LaTeX' Documents.
+Xie, Y. (2020) TinyTex: helper functions to install and maintain 'TeX
+Live', and compile 'LaTeX' documents.
 
 \leavevmode\hypertarget{ref-Xie2018}{}%
-Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: The
-Definitive Guide, Boca Raton, Florida: Chapman; Hall/CRC.
+Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: the
+definitive guide, 1st ed. Boca Raton, Florida: Chapman; Hall/CRC.
 
 \leavevmode\hypertarget{ref-Xie2017}{}%
-Xie, Z., Lin, W., and Luo, J. (2017) Comparative Phenotype and Genome
-Analysis of \emph{Cellvibrio} sp. PR1, a Xylanolytic and Agarolytic
-Bacterium from the Pearl River. \emph{Biomed Res Int} \textbf{2017}:
+Xie, Z., Lin, W., and Luo, J. (2017) Comparative phenotype and genome
+analysis of \emph{Cellvibrio} sp. PR1, a xylanolytic and agarolytic
+bacterium from the Pearl River. \emph{Biomed Res Int} \textbf{2017}:
 6304248.
@@
 Yilmaz, P., Parfrey, L.W., Yarza, P., Gerken, J., Pruesse, E., Quast,
-C., et al. (2014) The SILVA and "All-species Living Tree Project (LTP)"
+C., et al. (2014) The SILVA and "All-Species Living Tree Project (LTP)"
 taxonomic frameworks. \emph{Nucleic Acids Res} \textbf{42}: D643--D648.
@@
 \leavevmode\hypertarget{ref-Zhu2019}{}%
-Zhu, H. (2019) kableExtra: Construct Complex Table with 'kable' and Pipe
-Syntax.
+Zhu, H. (2019) KableExtra: construct complex table with 'kable' and pipe
+syntax.
 
@@
 
-\hypertarget{figure-captions}{%
-\subsection{Figure Captions}\label{figure-captions}}
+\hypertarget{figure-legends}{%
+\subsection{Figure legends}\label{figure-legends}}
 

### 2020-03-25 submission/manuscript.tex
Reference update.

@@
 normalized to the minimum number of sequences, 8,409
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,055
-different OTUs that ranged from 370 to 2,033 OTUs per sample
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,022
+different OTUs that ranged from 376 to 2,013 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,691.5 ±
-144.4 OTUs{]} and monospecific {[}Number of OTUs, 1,733.3 ± 156.8
-OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,060.9 ±
-216.1 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 527.5 ± 147.8 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,692.4 ±
+123.7 OTUs{]} and monospecific {[}Number of OTUs, 1,727.2 ± 156.3
+OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,049.2 ±
+206.9 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 529.4 ± 143.8 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 especially \emph{C. cylindracea} associated communities
-(\autoref{shared}). In addition, only 0.4 -- 1.0 \si{\percent} of OTUs
+(\autoref{shared}). In addition, only 0.5 -- 1.0 \si{\percent} of OTUs
 from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (40.3 -- 53.7
+These OTUs also made a high proportion of total sequences (41.6 -- 53.0
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 53.7
+were accounting for a high proportion of sequences (\(\leq\) 53.0
 \si{\percent}). A very similar proportion of persistent OTUs and their
@@
 we can learn from sushi: a review on seaweed-bacterial associations.
-\textbf{83}: 1--16.
+\emph{FEMS Microbiol Ecol} \textbf{83}: 1--16.
 

### 2020-04-01 submission/manuscript.tex
Process optimzation

@@
 in the ambient seawater were also processed for comparison. The number
-of reads per sample ranged between 8,409 and 77,463 sequences
+of reads per sample ranged between 8,408 and 77,463 sequences
 (\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
@@
 (\autoref{supp-rarefaction}). Following quality curation and exclusion
-of sequences mentioned before reads were clustered into 28,702 different
+of sequences mentioned before reads were clustered into 28,750 different
 OTUs at a similarity level of 97 \si{\percent}. Read numbers were
-normalized to the minimum number of sequences, 8,409
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,022
-different OTUs that ranged from 376 to 2,013 OTUs per sample
+normalized to the minimum number of sequences, 8,408
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,222
+different OTUs that ranged from 364 to 2,005 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,692.4 ±
-123.7 OTUs{]} and monospecific {[}Number of OTUs, 1,727.2 ± 156.3
-OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,049.2 ±
-206.9 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 529.4 ± 143.8 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,693.8 ±
+137.1 OTUs{]} and monospecific {[}Number of OTUs, 1,733.2 ± 161.8
+OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,051 ±
+201.5 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 531.2 ± 140.7 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 especially \emph{C. cylindracea} associated communities
-(\autoref{shared}). In addition, only 0.5 -- 1.0 \si{\percent} of OTUs
+(\autoref{shared}). In addition, only 0.4 -- 1.1 \si{\percent} of OTUs
 from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (41.6 -- 53.0
+These OTUs also made a high proportion of total sequences (40.2 -- 53.1
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 communities indicated a separation between Spring, Summer and
-Autumn/Winter samples (ANOSIM, R = 0.54, \emph{p} \textless{} 0.01). For
+Autumn/Winter samples (ANOSIM, R = 0.53, \emph{p} \textless{} 0.01). For
 \emph{C. cylindracea} associated communities a separation between Summer
 and Autumn/Winter/Spring samples was observed that was not so strongly
-supported (ANOSIM, R = 0.31, \emph{p} \textless{} 0.05)
+supported (ANOSIM, R = 0.32, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 could be observed in the proportion of OTUs that could be found at every
-sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 53.0
+sampling time (\(\leq\) 1.1 \si{\percent}). Interestingly, these OTUs
+were accounting for a high proportion of sequences (\(\leq\) 53.1
 \si{\percent}). A very similar proportion of persistent OTUs and their

### 2020-04-02 code/plot_calculators.R
Typing error corrected.

@@ -209,7 +209,7 @@ p2 <- filter(estimators_indices_metadata, station=="FCa") %>%
   labs(x="Date", y="") +
   theme
 
-fca <- cowplot::plot_grid(p1, p2, nrow=2, ncol=1, rel_heights=c(1,0.75), align="v")
+fca <- cowplot::plot_grid(p1, p2, nrow=2, ncol=1, rel_heights=c(1, 0.75), align="v")
 
 # Generating a plot to extract a common legend
 labels <- c("Observed Number of OTUs", "Chao1", "ACE", "Exponential Shannon", "Inverse Simpson")

### 2020-04-19 submission/manuscript.Rmd
References optimization.

@@
 
-Biofilms formed from microbial epiphytes can contain diverse taxonomic groups and harbor cell densities from 10^2^ to 10^7^ \si{\cells\per\cm\squared} [@Armstrong2000; @Bengtsson2010; @Burke2011b].
+Biofilms formed from microbial epiphytes can contain diverse taxonomic groups and harbor cell densities from 10^2^ to 10^7^ \si{\cells\per\cm\squared} [@Armstrong2000; @Bengtsson2010; @Burke2011a].
 In such an environment a number of positive and negative interactions between the macrophyte and colonizing microorganisms have been described [@Tarquinio2019; @Egan2013; @Hollants2013].
@@
 Recently, it was shown that even different morphological niches within the same alga had a higher influence on bacterial community variation than biogeography or environmental factors [@Morrissey2019].
-While there is high community variation between host species it was observed that the majority of metagenome determined functions were conserved both between host species and individuals [@Burke2011a; @Roth-Schulze2016].
+While there is high community variation between host species it was observed that the majority of metagenome determined functions were conserved both between host species and individuals [@Burke2011; @Roth-Schulze2016].
 This discrepancy between taxonomic and functional composition could be explained by the lottery hypothesis.
-It postulates that an initial random colonization step is performed from a set of functionally equivalent taxonomic groups resulting in taxonomically different epiphytic communities sharing a core set of functional genes [@Burke2011a; @Roth-Schulze2016].
+It postulates that an initial random colonization step is performed from a set of functionally equivalent taxonomic groups resulting in taxonomically different epiphytic communities sharing a core set of functional genes [@Burke2011; @Roth-Schulze2016].
 In addition, some of the variation in the observed data could be attributed to different techniques used in various studies, such as different protocols for epiphytic cell detachment and/or DNA isolation, as no standard protocol to study epiphytic communities was established [Korlević *et al.*, unpublished; @Ugarelli2019].
 
-The majority of studies describing macrophyte epiphytic communities did not encompass seasonal changes [@Ugarelli2019; @Lachnit2009; @Crump2008; @Burke2011b; @Roth-Schulze2016].
+The majority of studies describing macrophyte epiphytic communities did not encompass seasonal changes [@Ugarelli2019; @Lachnit2009; @Crump2008; @Burke2011a; @Roth-Schulze2016].
 In addition, if seasonal changes were taken into account low temporal frequency and/or methodologies that do not allow high taxonomic resolution were used [@Tujula2010; @Lachnit2011; @Miranda2013; @Michelou2013].
@@
 Since the colonization of macrophyte surfaces is performed from a pool of prokaryotic cells from the ambient seawater, it was interesting to see to which extent these two communities differ.
-We observed a strong differentiation between the surface attached and seawater communities at the level of OTUs that is in agreement with most published studies [@Ugarelli2019; @Roth-Schulze2016; @Burke2011b; @Michelou2013; @Crump2018].
+We observed a strong differentiation between the surface attached and seawater communities at the level of OTUs that is in agreement with most published studies [@Ugarelli2019; @Roth-Schulze2016; @Burke2011a; @Michelou2013; @Crump2018].
 These data indicate that marine macrophytes are selecting, from a pool of seawater microbial taxa, the one that can colonize and proliferate on their surfaces [@Salaun2012; @Michelou2013].
@@
 When the taxonomic composition at high ranks was analyzed no such strong differentiation was noticed.
-Phyla and classes such as: *Actinobacteriota*, *Bacteroidota*, *Cyanobacteria*, *Alphaproteobacteria*, *Gammaproteobacteria* and *Verrucomicrobiota* were described that is in agreement with previously reported data [@Egan2013; @Michelou2013; @Burke2011b].
+Phyla and classes such as: *Actinobacteriota*, *Bacteroidota*, *Cyanobacteria*, *Alphaproteobacteria*, *Gammaproteobacteria* and *Verrucomicrobiota* were described that is in agreement with previously reported data [@Egan2013; @Michelou2013; @Burke2011a].
 In contrast, when low taxonomic ranks were analyzed (i.g.
@@
 As already mentioned, the rhizoids and part of the stolon are in contact with the sediment, so *Desulfobacterota* are probably a part of the epiphytic community that was in contact with the sediment.
-Similar high rank taxa found in this study were described to be specific for other species of macrophytes [@Burke2011b; @Lachnit2011; @Bengtsson2017].
+Similar high rank taxa found in this study were described to be specific for other species of macrophytes [@Burke2011a; @Lachnit2011; @Bengtsson2017].
 In contrast to high taxonomic ranks, a substantial differentiation between host specific communities was found, which supports the host-specific hypothesis.
@@
 *Lewinella*, *Portibacter* and *Rubidimonas*) were detected through the year.
-Members of this family are often found in association with macrophytes and it is suggested that they are involved in the hydrolysis and utilization of complex carbon sources [@Burke2011b; @Crump2018; @McIlroy2014].
+Members of this family are often found in association with macrophytes and it is suggested that they are involved in the hydrolysis and utilization of complex carbon sources [@Burke2011a; @Crump2018; @McIlroy2014].
 On the other hand, families *Flavobacteriaceae* and *Rhodothermaceae* showed seasonal patterns, with *Flavobacteriaceae* being more pronounced from November to June and *Rhodothermaceae* from June to October.
 Within *Alphaproteobacteria* the family *Rhodobacteraceae* was comprising the majority of sequences throughout the year.
-This metabolically versatile family is often associated with macrophyte surfaces and usually is one of the most abundant groups [@Burke2011b; @Michelou2013; @Pujalte2014].
+This metabolically versatile family is often associated with macrophyte surfaces and usually is one of the most abundant groups [@Burke2011a; @Michelou2013; @Pujalte2014].
 In addition, *Hyphomonadaceae* were found in all samples.
@@
 In addition, *Granulosicoccus* was also found in almost all samples.
-*Gammaproteobacteria* are often a major constituent of macrophyte epiphytic communities [@Michelou2013; @Burke2011b; @Crump2018].
+*Gammaproteobacteria* are often a major constituent of macrophyte epiphytic communities [@Michelou2013; @Burke2011a; @Crump2018].
 Beside these two groups other less pronounced taxa showed seasonal and host-specific patterns.
@@
 Studies focusing on functional comparisons between communities associated with different hosts showed that the majority of functions could be found in every community, indicating functional redundancy [@Roth-Schulze2016].
-This difference between taxonomic and functional discrepancy was explained by the lottery hypothesis that hypothesizes an initial random colonization step performed from a set of functionally equivalent taxonomic groups [@Burke2011a; @Roth-Schulze2016].
+This difference between taxonomic and functional discrepancy was explained by the lottery hypothesis that hypothesizes an initial random colonization step performed from a set of functionally equivalent taxonomic groups [@Burke2011; @Roth-Schulze2016].
 It is possible that functional redundancy is a characteristic of high abundance taxa detected to be present throughout the year, while seasonal and/or host-specific functions are an attribute of taxa displaying successional patterns.
@@
 The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/mkorlevic/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020).
-Based on the ATCC MSA-1002 mock community included in the analysis an average sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016a].
+Based on the ATCC MSA-1002 mock community included in the analysis an average sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
 In addition, the negative controls processed together with the samples yielded on average only `r nc` sequences after sequence quality curation.

### 2020-04-19 submission/manuscript.tex
References optimization.

--- 2020-03-24 submission/manuscript.tex
@@
 (\autoref{supp-rarefaction}). Following quality curation and exclusion
-of sequences mentioned before reads were clustered into 28,702 different
+of sequences mentioned before reads were clustered into 28,750 different
 OTUs at a similarity level of 97 \si{\percent}. Read numbers were
-normalized to the minimum number of sequences, 8,409
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,055
-different OTUs that ranged from 370 to 2,033 OTUs per sample
+normalized to the minimum number of sequences, 8,408
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,115
+different OTUs that ranged from 372 to 2,084 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,691.5 ±
-144.4 OTUs{]} and monospecific {[}Number of OTUs, 1,733.3 ± 156.8
-OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,060.9 ±
-216.1 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 527.5 ± 147.8 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,684.6 ±
+140.0 OTUs{]} and monospecific {[}Number of OTUs, 1,735 ± 172.0
+OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,059.7 ±
+208.8 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 531.7 ± 141.4 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (40.3 -- 53.7
+These OTUs also made a high proportion of total sequences (40.1 -- 51.4
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 and Autumn/Winter/Spring samples was observed that was not so strongly
-supported (ANOSIM, R = 0.31, \emph{p} \textless{} 0.05)
+supported (ANOSIM, R = 0.32, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 53.7
+were accounting for a high proportion of sequences (\(\leq\) 51.4
 \si{\percent}). A very similar proportion of persistent OTUs and their
@@
 2014, 2015, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018; Y. Xie,
-2019a, 2019b; Allaire \emph{et al.}, 2019; Zhu, 2019). The detailed
+2019b, 2019a; Allaire \emph{et al.}, 2019; Zhu, 2019). The detailed
 analysis procedure including the R Markdown file for this paper are
@@
 Abraham, W.R. and Rohde, M. (2014) The family \emph{Hyphomonadaceae}. In
-\emph{The prokaryotes: Alphaproteobacteria and Betaproteobacteria}.
+\emph{The Prokaryotes: Alphaproteobacteria and Betaproteobacteria}.
 Rosenberg, E., DeLong, E.F., Lory, S., Stackebrandt, E., and Thompson,
@@
 Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
-A., et al. (2019) Rmarkdown: dynamic documents for R.
+A., et al. (2019) Rmarkdown: Dynamic documents for R.
 
@@
 \leavevmode\hypertarget{ref-Bengtsson2010}{}%
-Bengtsson, M., Sjøtun, K., and Øvreås, L. (2010) Seasonal dynamics of
-bacterial biofilms on the kelp \emph{Laminaria hyperborea}. \emph{Aquat
-Microb Ecol} \textbf{60}: 71--83.
+Bengtsson, M., Sjøtun, K., and Øvre\textbackslash{}aas, L. (2010)
+Seasonal dynamics of bacterial biofilms on the kelp \emph{Laminaria
+hyperborea}. \emph{Aquat Microb Ecol} \textbf{60}: 71--83.
 
@@
 
-\leavevmode\hypertarget{ref-Burke2011a}{}%
+\leavevmode\hypertarget{ref-Burke2011}{}%
 Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
@@
 
-\leavevmode\hypertarget{ref-Burke2011b}{}%
+\leavevmode\hypertarget{ref-Burke2011a}{}%
 Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
@@
 Egan, S., Harder, T., Burke, C., Steinberg, P., Kjelleberg, S., and
-Thomas, T. (2013) The seaweed holobiont: understanding seaweed-bacteria
+Thomas, T. (2013) The seaweed holobiont: Understanding seaweed-bacteria
 interactions. \emph{FEMS Microbiol Rev} \textbf{37}: 462--476.
@@
 Hollants, J., Leliaert, F., De Clerck, O., and Willems, A. (2013) What
-we can learn from sushi: a review on seaweed-bacterial associations.
-\textbf{83}: 1--16.
+we can learn from sushi: A review on seaweed-bacterial associations.
+\emph{FEMS Microbiol Ecol} \textbf{83}:
 
@@
 Kuever, J. (2014) The family \emph{Desulfobulbaceae}. In \emph{The
-prokaryotes: Deltaproteobacteria and Epsilonproteobacteria}. Rosenberg,
+Prokaryotes: Deltaproteobacteria and Epsilonproteobacteria}. Rosenberg,
 E., DeLong, E.F., Lory, S., Stackebrandt, E., and Thompson, F. (eds).
@@
 Lachnit, T., Blümel, M., Imhoff, J.F., and Wahl, M. (2009) Specific
-epibacterial communities on macroalgae: phylogeny matters more than
+epibacterial communities on macroalgae: Phylogeny matters more than
 habitat. \emph{Aquat Biol} \textbf{5}: 181--186.
@@
 Margulis, L. (1991) Symbiogenesis and symbionticism. In \emph{Symbiosis
-as a source of evolutionary innovation: Speciation and morphogenesis}.
+as a Source of Evolutionary Innovation: Speciation and Morphogenesis}.
 Margulis, L. and Fester, R. (eds). Cambridge, Massachusetts: The MIT
@@
 McIlroy, S.J. and Nielsen, P.H. (2014) The family \emph{Saprospiraceae}.
-In \emph{The prokaryotes: Other major lineages of Bacteria and the
+In \emph{The Prokaryotes: Other Major Lineages of Bacteria and the
 Archaea}. Rosenberg, E., DeLong, E.F., Lory, S., Stackebrandt, E., and
@@
 ecology of microbial communities associated with \emph{Macrocystis
-pyrifera}. \emph{PLoS One} \textbf{8}: e67480.
+pyrifera}. \emph{PloS One} \textbf{8}: e67480.
 
@@
 Diversity and abundance of the bacterial community of the red macroalga
-\emph{Porphyra umbilicalis}: did bacterial farmers produce macroalgae?
-\emph{PLoS One} \textbf{8}: e58269.
+\emph{Porphyra umbilicalis}: Did bacterial farmers produce macroalgae?
+\emph{PloS One} \textbf{8}: e58269.
 
@@
 Oksanen, J., Blanchet, F.G., Friendly, M., Kindt, R., Legendre, P.,
-McGlinn, D., et al. (2019) Vegan: community ecology package.
+McGlinn, D., et al. (2019) Vegan: Community ecology package.
 
@@
 Parada, A.E., Needham, D.M., and Fuhrman, J.A. (2016) Every base
-matters: assessing small subunit rRNA primers for marine microbiomes
+matters: Assessing small subunit rRNA primers for marine microbiomes
 with mock communities, time series and global field samples.
@@
 Pujalte, M.J., Lucena, T., Ruvira, M.A., Arahal, D.R., and Macián, M.C.
-(2014) The family \emph{Rhodobacteraceae}. In \emph{The prokaryotes:
+(2014) The family \emph{Rhodobacteraceae}. In \emph{The Prokaryotes:
 Alphaproteobacteria and Betaproteobacteria}. Rosenberg, E., DeLong,
@@
 Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
-et al. (2013) The SILVA ribosomal RNA gene database project: improved
+et al. (2013) The SILVA ribosomal RNA gene database project: Improved
 data processing and web-based tools. \emph{Nucleic Acids Res}
@@
 \leavevmode\hypertarget{ref-RCoreTeam2019}{}%
-R Core Team (2019) A language and environment for statistical computing.
+R Core Team (2019) A language and environment for statistical computing,
+Vienna, Austria: R Foundation for Statistical Computing.
 
@@
 Reisser, J., Shaw, J., Hallegraeff, G., Proietti, M., Barnes, D.K.A.,
-Thums, M., et al. (2014) Millimeter-sized marine plastics: a new pelagic
-habitat for microorganisms and invertebrates. \emph{PLoS One}
+Thums, M., et al. (2014) Millimeter-sized marine plastics: A new pelagic
+habitat for microorganisms and invertebrates. \emph{PloS One}
 \textbf{9}: e100289.
@@
 Seasonal variations in surface metabolite composition of \emph{Fucus
-vesiculosus} and \emph{Fucus serratus} from the Baltic Sea. \emph{PLoS
+vesiculosus} and \emph{Fucus serratus} from the Baltic Sea. \emph{PloS
 One} \textbf{11}: e0168196.
@@
 Salaün, S., La Barre, S., Santos-Goncalvez, M.D., Potin, P., Haras, D.,
-and Bazire, A. (2012) Influence of exudates of the kelp \emph{Laminaria
-digitata} on biofilm formation of associated and exogenous bacterial
+and Bazire, A. (2012) Influence of exudates of the kelp Laminaria
+digitata on biofilm formation of associated and exogenous bacterial
 epiphytes. \emph{Microb Ecol} \textbf{64}: 359--369.
 
-\leavevmode\hypertarget{ref-Schloss2016a}{}%
+\leavevmode\hypertarget{ref-Schloss2016}{}%
 Schloss, P.D., Jenior, M.L., Koumpouras, C.C., Westcott, S.L., and
@@
 Schloss, P.D., Westcott, S.L., Ryabin, T., Hall, J.R., Hartmann, M.,
-Hollister, E.B., et al. (2009) Introducing mothur: open-source,
+Hollister, E.B., et al. (2009) Introducing mothur: Open-source,
 platform-independent, community-supported software for describing and
@@
 Tarquinio, F., Hyndes, G.A., Laverock, B., Koenders, A., and Säwström,
-C. (2019) The seagrass holobiont: understanding seagrass-bacteria
+C. (2019) The seagrass holobiont: Understanding seagrass-bacteria
 interactions and their role in seagrass ecosystem functioning.
@@
 \leavevmode\hypertarget{ref-Xie2015}{}%
-Xie, Y. (2015) Dynamic documents with R and knitr, 2nd ed. Boca Raton,
-Florida: Chapman; Hall/CRC.
+Xie, Y. (2015) Dynamic Documents with R and knitr, 2nd ed. Boca Raton,
+Florida: Chapman and Hall/CRC.
 
 \leavevmode\hypertarget{ref-Xie2014}{}%
-Xie, Y. (2014) Knitr: a comprehensive tool for reproducible research in
-R. In \emph{Implementing reproducible computational research}. Stodden,
-V., Leisch, F., and Peng, R.D. (eds). New York: Chapman; Hall/CRC, pp.
-3--32.
+Xie, Y. (2014) Knitr: A comprehensive tool for reproducible research in
+R. In \emph{Implementing Reproducible Computational Research}. Stodden,
+V., Leisch, F., and Peng, R.D. (eds). New York: Chapman and Hall/CRC,
+pp. 3--32.
+
+\leavevmode\hypertarget{ref-Xie2019}{}%
+Xie, Y. (2019a) Knitr: A general-purpose package for dynamic report
+generation in R.
 
 \leavevmode\hypertarget{ref-Xie2019a}{}%
-Xie, Y. (2019a) Knitr: a general-purpose package for dynamic report
-generation in R.
-
-\leavevmode\hypertarget{ref-Xie2019}{}%
-Xie, Y. (2019b) TinyTeX: a lightweight, cross-platform, and
+Xie, Y. (2019b) TinyTeX: A lightweight, cross-platform, and
 easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
@@
 \leavevmode\hypertarget{ref-Xie2020}{}%
-Xie, Y. (2020) TinyTex: helper functions to install and maintain 'TeX
+Xie, Y. (2020) TinyTex: Helper functions to install and maintain 'TeX
 Live', and compile 'LaTeX' documents.
@@
 \leavevmode\hypertarget{ref-Xie2018}{}%
-Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: the
-definitive guide, 1st ed. Boca Raton, Florida: Chapman; Hall/CRC.
+Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: The
+Definitive Guide, 1st ed. Boca Raton, Florida: Chapman and Hall/CRC.
 
@@
 analysis of \emph{Cellvibrio} sp. PR1, a xylanolytic and agarolytic
-bacterium from the Pearl River. \emph{Biomed Res Int} \textbf{2017}:
+bacterium from the Pearl River. \emph{BioMed Res Int} \textbf{2017}:
 6304248.
@@
 \leavevmode\hypertarget{ref-Zhu2019}{}%
-Zhu, H. (2019) KableExtra: construct complex table with 'kable' and pipe
+Zhu, H. (2019) KableExtra: Construct complex table with 'kable' and pipe
 syntax.

### 2020-04-19 submission/manuscript.tex
References update.

@@
 normalized to the minimum number of sequences, 8,408
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,115
-different OTUs that ranged from 372 to 2,084 OTUs per sample
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,073
+different OTUs that ranged from 348 to 2,005 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,684.6 ±
-140.0 OTUs{]} and monospecific {[}Number of OTUs, 1,735 ± 172.0
-OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,059.7 ±
-208.8 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 531.7 ± 141.4 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,686.3 ±
+132.0 OTUs{]} and monospecific {[}Number of OTUs, 1,726.3 ± 158.4
+OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,054.6 ±
+216.8 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 522.1 ± 140.8 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (40.1 -- 51.4
+These OTUs also made a high proportion of total sequences (38.5 -- 53.4
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 communities indicated a separation between Spring, Summer and
-Autumn/Winter samples (ANOSIM, R = 0.54, \emph{p} \textless{} 0.01). For
+Autumn/Winter samples (ANOSIM, R = 0.52, \emph{p} \textless{} 0.01). For
 \emph{C. cylindracea} associated communities a separation between Summer
@@
 sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 51.4
+were accounting for a high proportion of sequences (\(\leq\) 53.4
 \si{\percent}). A very similar proportion of persistent OTUs and their

### 2020-04-19 submission/manuscript.tex
Reference update.

@@
 normalized to the minimum number of sequences, 8,408
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,073
-different OTUs that ranged from 348 to 2,005 OTUs per sample
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,180
+different OTUs that ranged from 381 to 1,997 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,686.3 ±
-132.0 OTUs{]} and monospecific {[}Number of OTUs, 1,726.3 ± 158.4
-OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,054.6 ±
-216.8 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 522.1 ± 140.8 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,693.8 ±
+134.8 OTUs{]} and monospecific {[}Number of OTUs, 1,726.2 ± 156.0
+OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,053.4 ±
+210.7 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 535.7 ± 137.8 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (38.5 -- 53.4
+These OTUs also made a high proportion of total sequences (40.5 -- 51.3
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 and Autumn/Winter/Spring samples was observed that was not so strongly
-supported (ANOSIM, R = 0.32, \emph{p} \textless{} 0.01)
+supported (ANOSIM, R = 0.29, \emph{p} \textless{} 0.05)
 (\autoref{pcoa}).
@@
 sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 53.4
+were accounting for a high proportion of sequences (\(\leq\) 51.3
 \si{\percent}). A very similar proportion of persistent OTUs and their
@@
 \leavevmode\hypertarget{ref-Bengtsson2010}{}%
-Bengtsson, M., Sjøtun, K., and Øvre\textbackslash{}aas, L. (2010)
-Seasonal dynamics of bacterial biofilms on the kelp \emph{Laminaria
-hyperborea}. \emph{Aquat Microb Ecol} \textbf{60}: 71--83.
+Bengtsson, M., Sjøtun, K., and Øvreås, L. (2010) Seasonal dynamics of
+bacterial biofilms on the kelp \emph{Laminaria hyperborea}. \emph{Aquat
+Microb Ecol} \textbf{60}: 71--83.
 

### 2020-04-19 submission/manuscript.tex
Reference optimization.

@@
 normalized to the minimum number of sequences, 8,408
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,180
-different OTUs that ranged from 381 to 1,997 OTUs per sample
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 16,972
+different OTUs that ranged from 351 to 1,996 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,693.8 ±
-134.8 OTUs{]} and monospecific {[}Number of OTUs, 1,726.2 ± 156.0
-OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,053.4 ±
-210.7 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 535.7 ± 137.8 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,682.5 ±
+134.4 OTUs{]} and monospecific {[}Number of OTUs, 1,727.4 ± 154.7
+OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,056 ±
+201.8 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 523.8 ± 149.4 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (40.5 -- 51.3
+These OTUs also made a high proportion of total sequences (39.0 -- 51.6
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 communities indicated a separation between Spring, Summer and
-Autumn/Winter samples (ANOSIM, R = 0.52, \emph{p} \textless{} 0.01). For
+Autumn/Winter samples (ANOSIM, R = 0.54, \emph{p} \textless{} 0.01). For
 \emph{C. cylindracea} associated communities a separation between Summer
 and Autumn/Winter/Spring samples was observed that was not so strongly
-supported (ANOSIM, R = 0.29, \emph{p} \textless{} 0.05)
+supported (ANOSIM, R = 0.30, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 51.3
+were accounting for a high proportion of sequences (\(\leq\) 51.6
 \si{\percent}). A very similar proportion of persistent OTUs and their
@@
 Salaün, S., La Barre, S., Santos-Goncalvez, M.D., Potin, P., Haras, D.,
-and Bazire, A. (2012) Influence of exudates of the kelp Laminaria
-digitata on biofilm formation of associated and exogenous bacterial
+and Bazire, A. (2012) Influence of exudates of the kelp \emph{Laminaria
+digitata} on biofilm formation of associated and exogenous bacterial
 epiphytes. \emph{Microb Ecol} \textbf{64}: 359--369.

### 2020-04-19 submission/manuscript.tex
References update.

@@
 normalized to the minimum number of sequences, 8,408
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 16,972
-different OTUs that ranged from 351 to 1,996 OTUs per sample
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,203
+different OTUs that ranged from 376 to 2,027 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,682.5 ±
-134.4 OTUs{]} and monospecific {[}Number of OTUs, 1,727.4 ± 154.7
-OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,056 ±
-201.8 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 523.8 ± 149.4 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,689.6 ±
+136.3 OTUs{]} and monospecific {[}Number of OTUs, 1,739.3 ± 156.0
+OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,055.3 ±
+217.2 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 526.5 ± 141.2 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (39.0 -- 51.6
+These OTUs also made a high proportion of total sequences (40.1 -- 51.5
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 communities indicated a separation between Spring, Summer and
-Autumn/Winter samples (ANOSIM, R = 0.54, \emph{p} \textless{} 0.01). For
+Autumn/Winter samples (ANOSIM, R = 0.55, \emph{p} \textless{} 0.01). For
 \emph{C. cylindracea} associated communities a separation between Summer
 and Autumn/Winter/Spring samples was observed that was not so strongly
-supported (ANOSIM, R = 0.30, \emph{p} \textless{} 0.01)
+supported (ANOSIM, R = 0.33, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 51.6
+were accounting for a high proportion of sequences (\(\leq\) 51.5
 \si{\percent}). A very similar proportion of persistent OTUs and their

### 2020-04-19 submission/manuscript.tex
Reference update.

--- 2020-04-19 submission/manuscript.tex
@@
 normalized to the minimum number of sequences, 8,408
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,180
-different OTUs that ranged from 381 to 1,997 OTUs per sample
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,131
+different OTUs that ranged from 371 to 2,000 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,693.8 ±
-134.8 OTUs{]} and monospecific {[}Number of OTUs, 1,726.2 ± 156.0
-OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,053.4 ±
-210.7 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 535.7 ± 137.8 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,691.8 ±
+130.0 OTUs{]} and monospecific {[}Number of OTUs, 1,734.7 ± 149.6
+OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,052.8 ±
+217.5 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 528.1 ± 140.9 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (40.5 -- 51.3
+These OTUs also made a high proportion of total sequences (41.1 -- 53.3
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 and Autumn/Winter/Spring samples was observed that was not so strongly
-supported (ANOSIM, R = 0.29, \emph{p} \textless{} 0.05)
+supported (ANOSIM, R = 0.31, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 51.3
+were accounting for a high proportion of sequences (\(\leq\) 53.3
 \si{\percent}). A very similar proportion of persistent OTUs and their

### 2020-04-19 submission/manuscript.tex
References update.

--- 2020-04-19 submission/manuscript.tex
@@
 normalized to the minimum number of sequences, 8,408
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 16,972
-different OTUs that ranged from 351 to 1,996 OTUs per sample
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,039
+different OTUs that ranged from 380 to 2,023 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,682.5 ±
-134.4 OTUs{]} and monospecific {[}Number of OTUs, 1,727.4 ± 154.7
-OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,056 ±
-201.8 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 523.8 ± 149.4 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,687.6 ±
+121.1 OTUs{]} and monospecific {[}Number of OTUs, 1,738.4 ± 160.6
+OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,062 ±
+218.2 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 524.8 ± 146.3 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 especially \emph{C. cylindracea} associated communities
-(\autoref{shared}). In addition, only 0.4 -- 1.0 \si{\percent} of OTUs
+(\autoref{shared}). In addition, only 0.4 -- 1.1 \si{\percent} of OTUs
 from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (39.0 -- 51.6
+These OTUs also made a high proportion of total sequences (39.6 -- 52.6
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 and Autumn/Winter/Spring samples was observed that was not so strongly
-supported (ANOSIM, R = 0.30, \emph{p} \textless{} 0.01)
+supported (ANOSIM, R = 0.32, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 could be observed in the proportion of OTUs that could be found at every
-sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 51.6
+sampling time (\(\leq\) 1.1 \si{\percent}). Interestingly, these OTUs
+were accounting for a high proportion of sequences (\(\leq\) 52.6
 \si{\percent}). A very similar proportion of persistent OTUs and their
@@
 \leavevmode\hypertarget{ref-Zavodnik1998}{}%
-Zavodnik, N., Travizi, A., and De Rosa, S. (1998) Seasonal variations in
+Zavodnik, N., Travizi, A., and de Rosa, S. (1998) Seasonal variations in
 the rate of photosynthetic activity and chemical composition of the

### 2020-04-20 submission/manuscript.tex
Discussion update.

--- 2020-04-19 submission/manuscript.tex
@@
 normalized to the minimum number of sequences, 8,408
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,131
-different OTUs that ranged from 371 to 2,000 OTUs per sample
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,043
+different OTUs that ranged from 352 to 1,999 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,691.8 ±
-130.0 OTUs{]} and monospecific {[}Number of OTUs, 1,734.7 ± 149.6
-OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,052.8 ±
-217.5 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 528.1 ± 140.9 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,673.1 ±
+132.8 OTUs{]} and monospecific {[}Number of OTUs, 1,739.1 ± 161.9
+OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,045.2 ±
+209.6 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 526.1 ± 140.4 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (41.1 -- 53.3
+These OTUs also made a high proportion of total sequences (39.4 -- 52.0
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 communities indicated a separation between Spring, Summer and
-Autumn/Winter samples (ANOSIM, R = 0.52, \emph{p} \textless{} 0.01). For
+Autumn/Winter samples (ANOSIM, R = 0.56, \emph{p} \textless{} 0.01). For
 \emph{C. cylindracea} associated communities a separation between Summer
 and Autumn/Winter/Spring samples was observed that was not so strongly
-supported (ANOSIM, R = 0.31, \emph{p} \textless{} 0.01)
+supported (ANOSIM, R = 0.30, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 53.3
+were accounting for a high proportion of sequences (\(\leq\) 52.0
 \si{\percent}). A very similar proportion of persistent OTUs and their
@@
 values in Autumn/Winter in comparison to Spring/Summer. This pattern is
-not surprising as seagrasses are known to harbor more epiphytes during
-Autumn/Winter (Reyes and Sansón, 2001). Furthermore, we used an adapted
-DNA isolation protocol that is known to partially coextract DNA from
-planktonic eukaryotes (Korlević \emph{et al.}, 2015). Strong seasonal
-fluctuations of high rank epiphytic taxa were not observed, with the
-exception of \emph{Cyanobateria}. Cyanobacterial sequences were more
-pronounced in November and December in comparison to Spring and Summer.
-Interestingly, in these high proportion months the majority of
+not surprising as seagrasses are known to harbor more algal epiphytes
+during Autumn/Winter (Reyes and Sansón, 2001). Furthermore, we used an
+adapted DNA isolation protocol that is known to partially coextract DNA
+from planktonic eukaryotes (Korlević \emph{et al.}, 2015). Strong
+seasonal fluctuations of high rank epiphytic taxa were not observed,
+with the exception of \emph{Cyanobateria}. Cyanobacterial sequences were
+more pronounced in November and December in comparison to Spring and
+Summer. Interestingly, in these high proportion months the majority of
 cyanobacterial sequences were classified as \emph{Pleurocapsa}, a group

### 2020-05-18 submission/manuscript.Rmd
Discussion update.
References changed.

@@
 It postulates that an initial random colonization step is performed from a set of functionally equivalent taxonomic groups resulting in taxonomically different epiphytic communities sharing a core set of functional genes [@Burke2011; @Roth-Schulze2016].
-In addition, some of the variation in the observed data could be attributed to different techniques used in various studies, such as different protocols for epiphytic cell detachment and/or DNA isolation, as no standard protocol to study epiphytic communities was established [Korlević *et al.*, unpublished; @Ugarelli2019].
+In addition, some of the variation in the observed data could be attributed to different techniques used in various studies, such as different protocols for epiphytic cell detachment and/or DNA isolation, as no standard protocol to study epiphytic communities was established [@Korlevic2020, submitted; @Ugarelli2019].
 
@@
 
-# Calculating the number of sequencing in millions
+# Calculating the number of sequences in millions
 million_sequences <- filter(community, taxon=="Archaea" | taxon=="Bacteria") %>%
@@
 All these methods showed different removal efficiencies but none was enabling a complete removal of attached microbial cells.
-In the present study, we applied an earlier developed removal protocol (Korlević *et al.*, unpublished), based on a previous idea of direct cellular lysis [@Burke2009], to ensure an almost complete cell detachment.
+In the present study, we applied an earlier developed removal protocol [@Korlevic2020, submitted], based on a previous idea of direct cellular lysis [@Burke2009], to ensure an almost complete cell detachment.
 The application of a direct lysis procedure coupled with a high frequency sampling protocol and Illumina high resolution amplicon sequencing has enabled us to make a detailed description of bacterial and archaeal communities associated with the surfaces of two marine macrophytes, *C.
@@
 Analysis of seasonal chloroplast sequence abundances showed higher values in Autumn/Winter in comparison to Spring/Summer.
-This pattern is not surprising as seagrasses are known to harbor more epiphytes during Autumn/Winter [@Reyes2001].
+This pattern is not surprising as seagrasses are known to harbor more algal epiphytes during Autumn/Winter [@Reyes2001].
 Furthermore, we used an adapted DNA isolation protocol that is known to partially coextract DNA from planktonic eukaryotes [@Korlevic2015].
@@
 nodosa* and *C.
-cylindracea* was isolated using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [Korlević *et al.*, unpublished; @Massana1997].
+cylindracea* was isolated using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevic2020, submitted; @Massana1997].
 Briefly, leaves and thalli are incubated in a lysis buffer and treated with lysozyme and proteinase K.
@@
 ### Illumina 16S rRNA sequencing
-Illumina MiSeq sequencing of the V4 16S rRNA region was performed as described previously (Korlević *et al.*, unpublished).
+Illumina MiSeq sequencing of the V4 16S rRNA region was performed as described previously [@Korlevic2020, submitted].
 The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.

### 2020-05-18 submission/manuscript.tex
Optimization.

@@
 isolation, as no standard protocol to study epiphytic communities was
-established (Korlević \emph{et al.}, unpublished; Ugarelli \emph{et
-al.}, 2019).
+established (Ugarelli \emph{et al.}, 2019; Korlević \emph{et al.}, 2020,
+submitted).
 
@@
 normalized to the minimum number of sequences, 8,408
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,043
-different OTUs that ranged from 352 to 1,999 OTUs per sample
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,196
+different OTUs that ranged from 371 to 2,070 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,673.1 ±
-132.8 OTUs{]} and monospecific {[}Number of OTUs, 1,739.1 ± 161.9
-OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,045.2 ±
-209.6 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 526.1 ± 140.4 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,694.9 ±
+130.8 OTUs{]} and monospecific {[}Number of OTUs, 1,737.2 ± 171.4
+OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,056.2 ±
+209.3 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 527.4 ± 147.5 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (39.4 -- 52.0
+These OTUs also made a high proportion of total sequences (41.3 -- 52.5
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 communities indicated a separation between Spring, Summer and
-Autumn/Winter samples (ANOSIM, R = 0.56, \emph{p} \textless{} 0.01). For
+Autumn/Winter samples (ANOSIM, R = 0.54, \emph{p} \textless{} 0.01). For
 \emph{C. cylindracea} associated communities a separation between Summer
 and Autumn/Winter/Spring samples was observed that was not so strongly
-supported (ANOSIM, R = 0.30, \emph{p} \textless{} 0.01)
+supported (ANOSIM, R = 0.32, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 study, we applied an earlier developed removal protocol (Korlević
-\emph{et al.}, unpublished), based on a previous idea of direct cellular
-lysis (Burke \emph{et al.}, 2009), to ensure an almost complete cell
-detachment. The application of a direct lysis procedure coupled with a
-high frequency sampling protocol and Illumina high resolution amplicon
-sequencing has enabled us to make a detailed description of bacterial
-and archaeal communities associated with the surfaces of two marine
-macrophytes, \emph{C. nodosa} and \emph{C. cylindracea}.
+\emph{et al.}, 2020, submitted), based on a previous idea of direct
+cellular lysis (Burke \emph{et al.}, 2009), to ensure an almost complete
+cell detachment. The application of a direct lysis procedure coupled
+with a high frequency sampling protocol and Illumina high resolution
+amplicon sequencing has enabled us to make a detailed description of
+bacterial and archaeal communities associated with the surfaces of two
+marine macrophytes, \emph{C. nodosa} and \emph{C. cylindracea}.
 
@@
 sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 52.0
+were accounting for a high proportion of sequences (\(\leq\) 52.5
 \si{\percent}). A very similar proportion of persistent OTUs and their
@@
 isolated using a previously modified and adapted protocol that allows
-for a selective epiphytic DNA isolation (Korlević \emph{et al.},
-unpublished; Massana \emph{et al.}, 1997). Briefly, leaves and thalli
-are incubated in a lysis buffer and treated with lysozyme and proteinase
-K. Following the incubations, the mixture containing lyzed epiphytic
-cells is separated from leaves and thalli and extracted using a
+for a selective epiphytic DNA isolation (Massana \emph{et al.}, 1997;
+Korlević \emph{et al.}, 2020, submitted). Briefly, leaves and thalli are
+incubated in a lysis buffer and treated with lysozyme and proteinase K.
+Following the incubations, the mixture containing lyzed epiphytic cells
+is separated from leaves and thalli and extracted using a
 phenol-chloroform procedure. Finally, the extracted DNA is precipitated
@@
 Illumina MiSeq sequencing of the V4 16S rRNA region was performed as
-described previously (Korlević \emph{et al.}, unpublished). The V4
+described previously (Korlević \emph{et al.}, 2020, submitted). The V4
 region of the 16S rRNA gene was amplified using a two-step PCR
@@
 
+\leavevmode\hypertarget{ref-Korlevic2020}{}%
+Korlević, M., Markovski, M., Zhao, Z., Herndl, G.J., and Najdek, M.
+(2020) Selective DNA and protein isolation from marine macrophyte
+surfaces.
+
 \leavevmode\hypertarget{ref-Korlevic2015}{}%

### 2020-05-18 submission/manuscript.Rmd
Removed line numbers.

@@
 \newpage
-\linenumbers
 \sisetup{mode=text}
@@
 It postulates that an initial random colonization step is performed from a set of functionally equivalent taxonomic groups resulting in taxonomically different epiphytic communities sharing a core set of functional genes [@Burke2011; @Roth-Schulze2016].
-In addition, some of the variation in the observed data could be attributed to different techniques used in various studies, such as different protocols for epiphytic cell detachment and/or DNA isolation, as no standard protocol to study epiphytic communities was established [@Korlevic2020, submitted; @Ugarelli2019].
+In addition, some of the variation in the observed data could be attributed to different techniques used in various studies, such as different protocols for epiphytic cell detachment and/or DNA isolation, as no standard protocol to study epiphytic communities was established [@Korlevica, submitted; @Ugarelli2019].
 
@@
 All these methods showed different removal efficiencies but none was enabling a complete removal of attached microbial cells.
-In the present study, we applied an earlier developed removal protocol [@Korlevic2020, submitted], based on a previous idea of direct cellular lysis [@Burke2009], to ensure an almost complete cell detachment.
+In the present study, we applied an earlier developed removal protocol [@Korlevica, submitted], based on a previous idea of direct cellular lysis [@Burke2009], to ensure an almost complete cell detachment.
 The application of a direct lysis procedure coupled with a high frequency sampling protocol and Illumina high resolution amplicon sequencing has enabled us to make a detailed description of bacterial and archaeal communities associated with the surfaces of two marine macrophytes, *C.
@@
 nodosa* and *C.
-cylindracea* was isolated using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevic2020, submitted; @Massana1997].
+cylindracea* was isolated using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevica, submitted; @Massana1997].
 Briefly, leaves and thalli are incubated in a lysis buffer and treated with lysozyme and proteinase K.
@@
 ### Illumina 16S rRNA sequencing
-Illumina MiSeq sequencing of the V4 16S rRNA region was performed as described previously [@Korlevic2020, submitted].
+Illumina MiSeq sequencing of the V4 16S rRNA region was performed as described previously [@Korlevica, submitted].
 The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.

### 2020-05-18 submission/manuscript.tex
Removed line numbers.

@@
 isolation, as no standard protocol to study epiphytic communities was
-established (Ugarelli \emph{et al.}, 2019; Korlević \emph{et al.}, 2020,
+established (Ugarelli \emph{et al.}, 2019; Korlević \emph{et al.},
 submitted).
@@
 normalized to the minimum number of sequences, 8,408
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,196
-different OTUs that ranged from 371 to 2,070 OTUs per sample
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,084
+different OTUs that ranged from 369 to 2,027 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,694.9 ±
-130.8 OTUs{]} and monospecific {[}Number of OTUs, 1,737.2 ± 171.4
-OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,056.2 ±
-209.3 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 527.4 ± 147.5 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,692.7 ±
+142.2 OTUs{]} and monospecific {[}Number of OTUs, 1,729.2 ± 163.4
+OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,051.8 ±
+212.2 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 530.7 ± 140.9 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (41.3 -- 52.5
+These OTUs also made a high proportion of total sequences (40.6 -- 51.4
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 study, we applied an earlier developed removal protocol (Korlević
-\emph{et al.}, 2020, submitted), based on a previous idea of direct
-cellular lysis (Burke \emph{et al.}, 2009), to ensure an almost complete
-cell detachment. The application of a direct lysis procedure coupled
-with a high frequency sampling protocol and Illumina high resolution
-amplicon sequencing has enabled us to make a detailed description of
-bacterial and archaeal communities associated with the surfaces of two
-marine macrophytes, \emph{C. nodosa} and \emph{C. cylindracea}.
+\emph{et al.}, submitted), based on a previous idea of direct cellular
+lysis (Burke \emph{et al.}, 2009), to ensure an almost complete cell
+detachment. The application of a direct lysis procedure coupled with a
+high frequency sampling protocol and Illumina high resolution amplicon
+sequencing has enabled us to make a detailed description of bacterial
+and archaeal communities associated with the surfaces of two marine
+macrophytes, \emph{C. nodosa} and \emph{C. cylindracea}.
 
@@
 sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 52.5
+were accounting for a high proportion of sequences (\(\leq\) 51.4
 \si{\percent}). A very similar proportion of persistent OTUs and their
@@
 for a selective epiphytic DNA isolation (Massana \emph{et al.}, 1997;
-Korlević \emph{et al.}, 2020, submitted). Briefly, leaves and thalli are
+Korlević \emph{et al.}, submitted). Briefly, leaves and thalli are
 incubated in a lysis buffer and treated with lysozyme and proteinase K.
@@
 Illumina MiSeq sequencing of the V4 16S rRNA region was performed as
-described previously (Korlević \emph{et al.}, 2020, submitted). The V4
-region of the 16S rRNA gene was amplified using a two-step PCR
-procedure. In the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and
-806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome
-Project
+described previously (Korlević \emph{et al.}, submitted). The V4 region
+of the 16S rRNA gene was amplified using a two-step PCR procedure. In
+the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R
+(5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project
 (\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
@@
 
-\leavevmode\hypertarget{ref-Korlevic2020}{}%
+\leavevmode\hypertarget{ref-Korlevica}{}%
 Korlević, M., Markovski, M., Zhao, Z., Herndl, G.J., and Najdek, M.
-(2020) Selective DNA and protein isolation from marine macrophyte
-surfaces.
+Selective DNA and protein isolation from marine macrophyte surfaces.
 

### 2020-05-19 submission/manuscript.tex
Optimized front page.

@@
 normalized to the minimum number of sequences, 8,408
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,084
-different OTUs that ranged from 369 to 2,027 OTUs per sample
+(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,017
+different OTUs that ranged from 360 to 2,108 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes of richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,692.7 ±
-142.2 OTUs{]} and monospecific {[}Number of OTUs, 1,729.2 ± 163.4
-OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,051.8 ±
-212.2 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 530.7 ± 140.9 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,675.9 ±
+144.0 OTUs{]} and monospecific {[}Number of OTUs, 1,735.6 ± 175.0
+OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,052.6 ±
+204.9 OTUs) and lower values for picoplankton communities in the ambient
+seawater (Number of OTUs, 525.3 ± 146.6 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not show such large
@@
 community was found between mixed and monospecific \emph{C. cylindracea}
-(Jaccard, 0.35; Bray-Curtis, 0.78), while lower shared values were
+(Jaccard, 0.36; Bray-Curtis, 0.78), while lower shared values were
 calculated between seawater and epiphytic communities
@@
 from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (40.6 -- 51.4
+These OTUs also made a high proportion of total sequences (40.6 -- 52.6
 \si{\percent}). To further disentangle the environmental and seasonal
@@
 communities indicated a separation between Spring, Summer and
-Autumn/Winter samples (ANOSIM, R = 0.52, \emph{p} \textless{} 0.01). For
+Autumn/Winter samples (ANOSIM, R = 0.53, \emph{p} \textless{} 0.01). For
 \emph{C. cylindracea} associated communities a separation between Summer
 and Autumn/Winter/Spring samples was observed that was not so strongly
-supported (ANOSIM, R = 0.32, \emph{p} \textless{} 0.01)
+supported (ANOSIM, R = 0.31, \emph{p} \textless{} 0.05)
 (\autoref{pcoa}).
@@
 sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 51.4
+were accounting for a high proportion of sequences (\(\leq\) 52.6
 \si{\percent}). A very similar proportion of persistent OTUs and their

### 2020-05-19 submission/supplementary.Rmd
Changes made to front page.
Optimized front page.

@@
 
-\vspace{60mm}
+\vspace{20mm}
 Marino Korlević^1$*$^, Marsej Markovski^1^, Zihao Zhao^2^, Gerhard J. Herndl^2,3^, Mirjana Najdek^1^
-
-\vspace{40mm}
-
-^$*$^To whom correspondence should be addressed: marino.korlevic@irb.hr
 

### 2020-05-28 submission/manuscript.Rmd
First review.

@@
 
-3\. Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research, Utrecht University, The Netherlands
+3\. NIOZ, Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research, Utrecht University, The Netherlands
 
@@
 Surfaces of marine macrophytes (seagrasses and macroalgae) are inhabited by diverse microbial communities.
-Most studies focusing on macrophyte epiphytic communities did not take into account temporal changes or applied low sampling frequency approaches.
-Illumina sequencing of the V4 16S rRNA region was performed to determine the seasonal dynamics of epiphytic communities sampled from the surfaces of the seagrass *Cymodocea nodosa* and invasive macroalga *Caulerpa cylindracea*.
+Most studies focusing on epiphytic communities of macrophytes did not take into account temporal changes or applied low sampling frequency approaches.
+Illumina sequencing of the V4 16S rRNA region was performed to determine the seasonal dynamics of epiphytic communities of the seagrass *Cymodocea nodosa* and invasive macroalga *Caulerpa cylindracea*.
 Leaves and thalli were sampled in a meadow of *Cymodocea nodosa* invaded by the invasive *Caulerpa cylindracea* and in a monospecific settlement of *Caulerpa cylindracea* located in the proximity of the meadow at monthly intervals.
 For comparison the ambient prokaryotic plankton community was also characterized.
-Sequencing results at the OTU level showed a clear differentiation between ambient water and epiphytic communities and a host-specific community assemblage.
-In addition, successional changes were observed that could be connected to the macrophyte growth cycle.
-Taxonomic analysis showed similar high rank groups in the ambient water and epiphytic communities, with the exception of *Desulfobacterota* that were found only on *Caulerpa cylindracea*.
-Only *Cyanobacteria* showed seasonal change, while other high rank taxa were present throughout the year.
-In every analyzed high rank taxa, phylogenetic groups present throughout the year comprised most of the sequences and could be identified together with low proportion taxa showing seasonal patterns connected to the macrophyte growth cycle.
+At the OTU level, the microbial community composition differed between the ambient water and the epiphytic communities exhibiting host-specificity.
+Also, successional changes were observed connected to the macrophyte growth cycle.
+Taxonomic analysis, however, showed similar high rank groups in the ambient water and the epiphytic communities, with the exception of *Desulfobacterota*, which were only found on *Caulerpa cylindracea*.
+*Cyanobacteria* showed seasonal changes while other high rank taxa were present throughout the year.
+Phylogenetic groups present throughout the year constituted most of the sequences, while less abundant taxa showed seasonal patterns connected to the macrophyte growth cycle.
 Taken together, epiphytic microbial communities of the seagrass *Cymodocea nodosa* and the macroalgae *Caulerpa cylindracea* appear to be host-specific and contain taxa that undergo successional changes.
@@
 ## Introduction
-Marine macrophytes (seagrasses and macroalgae) are important ecosystem engineers that form close associations with microorganism belonging to all three domains of life [@Tarquinio2019; @Egan2013].
+Marine macrophytes (seagrasses and macroalgae) are important ecosystem engineers forming close associations with microorganism belonging to all three domains of life [@Tarquinio2019; @Egan2013].
 Microbes can live within macrophyte tissue as endophytes or can form epiphytic communities on surfaces of leaves and thalli [@Tarquinio2019; @Egan2013; @Hollants2013].
-Epiphytic and endophytic microbial communities form a close functional relationship with the macrophyte host.
+Epiphytic and endophytic microbial communities exhibit a close functional relationship with the macrophyte host.
 It was proposed that this close relationship constitutes a holobiont, an integrated community where the macrophyte organism and its symbiotic partners support each other [@Margulis1991; @Tarquinio2019; @Egan2013].
 
-Biofilms formed from microbial epiphytes can contain diverse taxonomic groups and harbor cell densities from 10^2^ to 10^7^ \si{\cells\per\cm\squared} [@Armstrong2000; @Bengtsson2010; @Burke2011a].
-In such an environment a number of positive and negative interactions between the macrophyte and colonizing microorganisms have been described [@Tarquinio2019; @Egan2013; @Hollants2013].
-Macrophytes can promote growth of associated microbes by nutrient exudation, while in return microorganisms may support macrophyte performance through improved nutrient availability, phytohormone production and protection form toxic compounds, oxidative stress, biofouling organisms and pathogens [@Tarquinio2019; @Egan2013; @Hollants2013].
-Beside this positive interactions, macrophytes can  negatively impact the associated microbes such as pathogenic bacteria by producing reactive oxygen species and secondary metabolites [@Tarquinio2019; @Egan2013; @Hollants2013].
+Biofilms of microbial epiphytes can contain diverse taxonomic groups and harbor cell abundances from 10^2^ to 10^7^ \si{\cells\per\cm\squared} [@Armstrong2000; @Bengtsson2010; @Burke2011a].
+In such an environment a number of positive and negative interactions between the macrophyte and the colonizing microorganisms have been described [@Tarquinio2019; @Egan2013; @Hollants2013].
+Macrophytes can promote growth of associated microbes by nutrient exudation, while in return microorganisms may support macrophyte performance through improved nutrient availability, phytohormone production and protection from toxic compounds, oxidative stress, biofouling organisms and pathogens [@Tarquinio2019; @Egan2013; @Hollants2013].
+Besides these positive interactions, macrophytes can  negatively impact the associated microbes such as pathogenic bacteria by producing reactive oxygen species and secondary metabolites [@Tarquinio2019; @Egan2013; @Hollants2013].
 
@@
 At higher taxonomic ranks a core set of macrophyte epiphytic taxa was described consisting mainly of *Alphaproteobacteria*, *Gammaproteobacteria*, *Desulfobacterota*, *Bacteroidota*, *Cyanobacteria*, *Actinobacteriota*, *Firmicutes*, *Planctomycetota*, *Chloroflexi* and *Verrucomicrobiota* [@Lachnit2011; @Crump2008; @Tujula2010].
-In contrast, at lower taxonomic ranks host specific microbial communities were described [@Roth-Schulze2016; @Lachnit2011].
-Recently, it was shown that even different morphological niches within the same alga had a higher influence on bacterial community variation than biogeography or environmental factors [@Morrissey2019].
-While there is high community variation between host species it was observed that the majority of metagenome determined functions were conserved both between host species and individuals [@Burke2011; @Roth-Schulze2016].
-This discrepancy between taxonomic and functional composition could be explained by the lottery hypothesis.
-It postulates that an initial random colonization step is performed from a set of functionally equivalent taxonomic groups resulting in taxonomically different epiphytic communities sharing a core set of functional genes [@Burke2011; @Roth-Schulze2016].
-In addition, some of the variation in the observed data could be attributed to different techniques used in various studies, such as different protocols for epiphytic cell detachment and/or DNA isolation, as no standard protocol to study epiphytic communities was established [@Korlevica, submitted; @Ugarelli2019].
-
-The majority of studies describing macrophyte epiphytic communities did not encompass seasonal changes [@Ugarelli2019; @Lachnit2009; @Crump2008; @Burke2011a; @Roth-Schulze2016].
-In addition, if seasonal changes were taken into account low temporal frequency and/or methodologies that do not allow high taxonomic resolution were used [@Tujula2010; @Lachnit2011; @Miranda2013; @Michelou2013].
+In contrast, at lower taxonomic ranks host specific microbial communities were found [@Roth-Schulze2016; @Lachnit2011].
+Recently, it was shown that even different morphological niches within the same alga had a higher influence on the composition of the bacterial community than biogeography or environmental factors [@Morrissey2019].
+While the microbial community composition varies between host species, metagenomic analyses revealed that the majority of the microbial functions are conserved [@Burke2011; @Roth-Schulze2016].
+This discrepancy between the microbial taxonomic and functional composition might be explained by the lottery hypothesis.
+It postulates that an initial random colonization step takes places from a set of functionally equivalent taxonomic groups resulting in taxonomically different epiphytic communities sharing a core set of functional genes [@Burke2011; @Roth-Schulze2016].
+In addition, some of the variation in the reported data could be attributed to different techniques used in these studies, such as different protocols for epiphytic cell detachment and/or DNA isolation, as no standard protocol has been yet established to study epiphytic communities [@Korlevica, submitted; @Ugarelli2019].
+
+The majority of studies describing macrophyte epiphytic microbial communities did not include possible seasonal changes [@Ugarelli2019; @Lachnit2009; @Crump2008; @Burke2011a; @Roth-Schulze2016].
+If seasonal changes were taken into account, low temporal frequency, applied methodologies and/or limited number of analyzed host species did not allow a high taxonomic resolution [@Tujula2010; @Lachnit2011; @Miranda2013; @Michelou2013; @Mancuso2016].
 In the present study we describe the seasonal dynamics of bacterial and archaeal communities on the surfaces of the seagrass *Cymodocea nodosa* and siphonous macroalgae *Caulerpa cylindracea* determined on a mostly monthly scale.
@@
 cylindracea* and in a locality of only *C.
-cylindracea* located in the proximity of the meadow.
-In addition, for comparison, the community of the ambient seawater was characterized.
+cylindracea* located in the proximity of the seagrass meadow.
+For comparison, the microbial community of the ambient seawater was also characterized.
 
@@
 
-Sequencing of the 16S rRNA V4 region yielded a total of `r million_sequences` million sequences after quality curation and exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences (\autoref{supp-nseq_notus}).
+Sequencing of the 16S rRNA V4 region yielded a total of `r million_sequences` million sequences after quality curation and exclusion of sequences without known relatives (no relative sequences) and eukaryotic, chloroplast and mitochondrial sequences (\autoref{supp-nseq_notus}).
 A total of `r samples_epiphytes` samples originating from epiphytic archaeal and bacterial communities associated with surfaces of the seagrass *C.
-nodosa* and macroalga *C.
+nodosa* and the macroalga *C.
 cylindracea* were analyzed.
-In addition, `r samples_seawater` samples (one of the samples was sequenced two times) originating from picoplankton archaeal and bacterial communities in the ambient seawater were also processed for comparison.
+In addition, `r samples_seawater` samples (one of the samples was sequenced twice) originating from the ambient seawater were also processed for comparison.
 The number of reads per sample ranged between `r min_max$min` and `r min_max$max` sequences (\autoref{supp-nseq_notus}).
-Even when the highest sequencing effort was applied the rarefaction curves did not level off that is a common observation in high-throughput 16S rRNA amplicon sequencing approaches (\autoref{supp-rarefaction}).
-Following quality curation and exclusion of sequences mentioned before reads were clustered into `r no_otus` different OTUs at a similarity level of 97 \si{\percent}.
-Read numbers were normalized to the minimum number of sequences, `r min_max$min` (\autoref{supp-nseq_notus}), through rarefaction resulting in `r no_otus_rarefied` different OTUs that ranged from `r min(otu_min_max)` to `r max(otu_min_max)` OTUs per sample (\autoref{supp-calculators}).
-To determine seasonal changes of richness and diversity the Observed Number of OTUs, Chao1, ACE, Exponential Shannon [@Jost2006] and Inverse Simpson were calculated after normalization through rarefaction.
+Even when the highest sequencing effort was applied the rarefaction curves did not level off as commonly observed in high-throughput 16S rRNA amplicon sequencing approaches (\autoref{supp-rarefaction}).
+Following quality curation and exclusion of sequences as mentioned above reads were clustered into `r no_otus` different OTUs at a similarity level of 97 \si{\percent}.
+Read numbers were normalized to the minimum number of sequences (`r min_max$min`, \autoref{supp-nseq_notus}) through rarefaction resulting in `r no_otus_rarefied` different OTUs with `r min(otu_min_max)` to `r max(otu_min_max)` OTUs per sample (\autoref{supp-calculators}).
+To determine seasonal changes in richness and diversity the observed number of OTUs, Chao1, ACE, Exponential Shannon [@Jost2006] and Inverse Simpson were calculated after normalization through rarefaction.
 Generally, richness estimators and diversity indices showed similar trends.
 On average, higher values were found for *C.
-cylindracea* (mixed [Number of OTUs, `r S.obs_fcam$mean` ± `r S.obs_fcam$sd` OTUs] and monospecific [Number of OTUs, `r S.obs_fca$mean` ± `r S.obs_fca$sd` OTUs]), middle values for *C.
-nodosa* (Number of OTUs, `r S.obs_fcym$mean` ± `r S.obs_fcym$sd` OTUs) and lower values for picoplankton communities in the ambient seawater (Number of OTUs, `r S.obs_f$mean` ± `r S.obs_f$sd` OTUs) (\autoref{supp-calculators}).
-Seasonal changes did not show such large dissimilarities.
+cylindracea* (mixed [Number of OTUs, `r S.obs_fcam$mean` ± `r S.obs_fcam$sd` OTUs] and monospecific [Number of OTUs, `r S.obs_fca$mean` ± `r S.obs_fca$sd` OTUs]) than for *C.
+nodosa* (Number of OTUs, `r S.obs_fcym$mean` ± `r S.obs_fcym$sd` OTUs) and lowest values were obtained for the microbial community of the ambient seawater (Number of OTUs, `r S.obs_f$mean` ± `r S.obs_f$sd` OTUs) (\autoref{supp-calculators}).
+Seasonal changes did not reveal such large dissimilarities.
 *C.
 nodosa* communities showed a slow increase towards the end of the study, while *C.
-cylindracea* (mixed and monospecific) communities were characterized by slightly larger values in Spring and Summer in comparison to Autumn and Winter (\autoref{supp-calculators}).
+cylindracea* (mixed and monospecific) communities were characterized by slightly higher values in spring and summer than in autumn and winter (\autoref{supp-calculators}).
 
@@
 
-To determine the proportion of shared archaeal and bacterial OTUs and communities sampled in different environments the Jaccard's Similarity Coefficient on presence-absence data and Bray-Curtis Similarity Coefficient were, respectively, calculated.
-Coefficients were determined after normalization through rarefaction and binning of samples from a particular environment.
+To determine the proportion of shared archaeal and bacterial OTUs and communities sampled in different environments the Jaccard's Similarity Coefficient on presence-absence data and Bray-Curtis Similarity Coefficient, respectively, were calculated.
+Coefficients were determined after normalization through rarefaction and binning of samples from the particular environment.
 The highest proportion of shared OTUs and community was found between mixed and monospecific  *C.
-cylindracea* (Jaccard, `r format(filter(similarity, V1=="FCa", V2=="FCaM")$jaccard, digits=2, nsmall=2)`; Bray-Curtis, `r format(filter(similarity, V1=="FCa", V2=="FCaM")$bray, digits=2, nsmall=2)`), while lower shared values were calculated between seawater and epiphytic communities (\autoref{matrix}).
+cylindracea* (Jaccard, `r format(filter(similarity, V1=="FCa", V2=="FCaM")$jaccard, digits=2, nsmall=2)`; Bray-Curtis, `r format(filter(similarity, V1=="FCa", V2=="FCaM")$bray, digits=2, nsmall=2)`), while lower shared values were calculated between ambient seawater and epiphytic communities (\autoref{matrix}).
 Shared proportion between *C.
 nodosa* and *C.
-cylindracea* were approximately in the middle between these two extremes.
+cylindracea* were approximately in-between the values of mixed and monospecific *C.
+cylindracea*.
 To assess seasonal changes in the proportion of shared OTUs and communities the Jaccard's and Bray-Curtis Similarity Coefficients were calculated between consecutive sampling points (\autoref{shared}).
 Both coefficients showed similar trends.
-Temporal proportional changes were more pronounced for seawater in comparison to *C.
+Temporal proportional changes were more pronounced for ambient seawater than for *C.
 nodosa* and especially *C.
 cylindracea* associated communities (\autoref{shared}).
-In addition, only `r min(core$core_otus)` -- `r max(core$core_otus)` \si{\percent} of OTUs from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (`r min(core$core_seq)` -- `r max(core$core_seq)` \si{\percent}).
+In addition, only `r min(core$core_otus)` -- `r max(core$core_otus)` \si{\percent} of OTUs from each surface associated community were present at all seasons.
+These persistent OTUs constituted a high proportion of total sequences (`r min(core$core_seq)` -- `r max(core$core_seq)` \si{\percent}).
 To further disentangle the environmental and seasonal community dissimilarity a Principal Coordinates Analysis (PCoA) based on Bray-Curtis distances and OTU abundances was applied.
-It showed a clear separation between planktonic and surface associated communities (\autoref{pcoa}).
-In addition, a separation of epiphytic bacterial and archaeal communities based on host species was determined.
+A clear separation between ambient seawater and surface associated communities was found (\autoref{pcoa}).
+In addition, a separation of epiphytic bacterial and archaeal communities based on host species was detected.
 This separation was further supported by ANOSIM (R = `r format(anosim_all$statistic, digits=2, nsmall=2)`, *p* < `r anosim_all_signif`).
 Seasonal changes of *C.
-nodosa* associated communities indicated a separation between Spring, Summer and Autumn/Winter samples (ANOSIM, R = `r format(anosim_fcym$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fcym_signif`).
+nodosa* associated communities indicated a separation between spring, summer and autumn/winter samples (ANOSIM, R = `r format(anosim_fcym$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fcym_signif`).
 For *C.
-cylindracea* associated communities a separation between Summer and Autumn/Winter/Spring samples was observed that was not so strongly supported (ANOSIM, R = `r format(anosim_fca$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fca_signif`) (\autoref{pcoa}).
+cylindracea* associated communities a separation between summer and autumn/winter/spring samples was observed that was, however, not as strong as for *C.
+nodosa* associated communities (ANOSIM, R = `r format(anosim_fca$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fca_signif`) (\autoref{pcoa}).
 
@@
 
-The taxonomic composition of both, macrophyte associated and seawater communities, was dominated by bacterial (`r filter(archaea_bacteria, taxon=="Bacteria")$mean` ± `r filter(archaea_bacteria, taxon=="Bacteria")$sd` \si{\percent}) over archaeal sequences (`r filter(archaea_bacteria, taxon=="Archaea")$mean` ± `r filter(archaea_bacteria, taxon=="Archaea")$sd` \si{\percent}) (\autoref{community}).
-Higher relative abundances of chloroplast related sequences were only observed in surface associated communities, with higher values in Autumn/Winter (`r filter(chloroplast, season=="Winter")$mean` ± `r filter(chloroplast, season=="Winter")$sd` \si{\percent}) in comparison to Spring/Summer (`r filter(chloroplast, season=="Summer")$mean` ± `r filter(chloroplast, season=="Summer")$sd` \si{\percent}) (\autoref{supp-chloroplast}).
-Generally, at higher taxonomic ranks (phylum-class) epiphytic and seawater microbial communities were composed of similar bacterial taxa.
-Seawater communities were mainly comprised of *Actinobacteriota*, *Bacteroidota*, *Cyanobacteria*, *Alphaproteobacteria*, *Gammaproteobacteria* and *Verrucomicrobiota*.
+The taxonomic composition of both, macrophyte associated and ambient seawater communities was dominated by bacterial (`r filter(archaea_bacteria, taxon=="Bacteria")$mean` ± `r filter(archaea_bacteria, taxon=="Bacteria")$sd` \si{\percent}) over archaeal sequences (`r filter(archaea_bacteria, taxon=="Archaea")$mean` ± `r filter(archaea_bacteria, taxon=="Archaea")$sd` \si{\percent}) (\autoref{community}).
+Higher relative abundances of chloroplast related sequences were only observed in surface associated communities, with higher values in autumn/winter (`r filter(chloroplast, season=="Winter")$mean` ± `r filter(chloroplast, season=="Winter")$sd` \si{\percent}) than in spring/summer (`r filter(chloroplast, season=="Summer")$mean` ± `r filter(chloroplast, season=="Summer")$sd` \si{\percent}) (\autoref{supp-chloroplast}).
+Generally, at higher taxonomic ranks (phylum-class), epiphytic and ambient seawater microbial communities were composed of similar bacterial taxa.
+Ambient seawater communities were mainly comprised of *Actinobacteriota*, *Bacteroidota*, *Cyanobacteria*, *Alphaproteobacteria*, *Gammaproteobacteria* and *Verrucomicrobiota*.
 Communities associated with *C.
-nodosa* consisted of same groups with the addition of *Planctomycetota* whose contribution was higher in summer 2018.
+nodosa* consisted additionally of *Planctomycetota* contributing more in summer 2018 than in other seasons.
 In addition, communities from mixed and monospecific *C.
-cylindracea* were similar and characterized by same groups as seawater and *C.
+cylindracea* were similar and characterized by the same groups as ambient seawater and *C.
 nodosa* communities with the addition of *Desulfobacterota* (\autoref{community}).
-Larger differences between environments and host species could be observed at lower taxonomic ranks (\autoref{cyano} -- \ref{desulfo}).
+Larger differences between environments and host species were observed at lower taxonomic ranks (Figs.
+\ref{cyano} -- \ref{desulfo}).
 
@@
 
-*Cyanobacteria* related sequences were comprising, on average, `r cyano_all_whole$mean` ± `r cyano_all_whole$sd` \si{\percent} of total sequences (\autoref{cyano}).
+*Cyanobacteria* related sequences comprised, on average, `r cyano_all_whole$mean` ± `r cyano_all_whole$sd` \si{\percent} of total sequences (\autoref{cyano}).
 Higher proportions were found for *C.
 nodosa* (`r filter(cyano_station_season_whole, station=="FCyM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCyM" & season=="Autumn")$sd` \si{\percent}) and *C.
-cylindracea* (mixed [ (`r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$sd` \si{\percent})] and monospecific [ (`r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$sd` \si{\percent})]) associated communities in autumn and for seawater communities in winter (`r filter(cyano_station_season_whole, station=="F" & season=="Winter")$mean` ± `r filter(cyano_station_season_whole, station=="F" & season=="Winter")$sd` \si{\percent}).
-Large taxonomic differences between surface associated and seawater cyanobacterial communities were observed.
-Seawater communities were mainly comprised of *Cyanobium* and *Synechococcus*, while surface associated communities were consisted of *Pleurocapsa* and sequences without known relatives within *Cyanobacteriia* (\autoref{cyano}).
-In addition, seasonal changes in surface associated communities were observed with *Pleurocapsa* and no relative *Cyanobacteriia* comprising larger proportions in autumn and winter and *Acrophormium*, *Phormidesmis* and no relative *Nodosilineaceae* in spring and summer (\autoref{cyano}).
+cylindracea* mixed (`r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$sd` \si{\percent}) and monospecific (`r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$sd` \si{\percent}) associated communities in autumn and for ambient seawater communities in winter (`r filter(cyano_station_season_whole, station=="F" & season=="Winter")$mean` ± `r filter(cyano_station_season_whole, station=="F" & season=="Winter")$sd` \si{\percent}).
+Large taxonomic differences between surface associated and ambient seawater cyanobacterial communities were observed.
+Ambient seawater communities were mainly comprised of *Cyanobium* and *Synechococcus*, while surface associated communities were comprised of *Pleurocapsa* and sequences within the class *Cyanobacteriia* that could not be further classified (no relative *Cyanobacteriia*) (\autoref{cyano}).
+In addition, seasonal changes in surface associated communities were observed in *Pleurocapsa* and no relative *Cyanobacteriia* comprising larger proportions in autumn and winter and *Acrophormium*, *Phormidesmis* and sequences without known relatives within the *Nodosilineaceae* (no relative *Nodosilineaceae*) in spring and summer (\autoref{cyano}).
 
@@
 
-Sequences classified as *Bacteroidota* were comprising, on average, `r bactero_all_whole$mean` ± `r bactero_all_whole$sd` \si{\percent} of all sequences (\autoref{bactero}).
-Similarly to *Cyanobacteria*, large differences in the taxonomic composition between seawater and surface associated communities were found (\autoref{bactero}).
-The seawater community was characterized by the NS4 and NS5 marine groups, uncultured *Cryomorphaceae*, uncultured *Flavobacteriaceae*, NS11-12 marine group, *Balneola*, uncultured *Balneolaceae* and *Formosa*.
-In contrast, in surface associated communities *Lewinella*, *Portibacter*, *Rubidimonas*, no relative *Saprospiraceae*, uncultured *Saprospiraceae*, no relative *Flavobacteriaceae* and uncultured *Rhodothermaceae* were found.
-Some groups showed slight seasonal changes such as no relative *Flavobacteriaceae* that were more pronounced from November 2017 until June 2018.
+Sequences classified as *Bacteroidota* comprised, on average, `r bactero_all_whole$mean` ± `r bactero_all_whole$sd` \si{\percent} of all sequences (\autoref{bactero}).
+Similarly to *Cyanobacteria*, large differences in the taxonomic composition between ambient seawater and surface associated communities were found (\autoref{bactero}).
+The ambient seawater community was characterized by the NS4 and NS5 marine groups, uncultured *Cryomorphaceae*, uncultured *Flavobacteriaceae*, NS11-12 marine group, *Balneola*, uncultured *Balneolaceae* and *Formosa*.
+In contrast, in surface associated communities *Lewinella*, *Portibacter*, *Rubidimonas*, sequences without known relatives within the *Saprospiraceae* (no relative *Saprospiraceae*), uncultured *Saprospiraceae*, (sequences without known relatives within the *Flavobacteriaceae* (no relative *Flavobacteriaceae*)and uncultured *Rhodothermaceae* were found.
+Some groups showed minor seasonal changes such as no relative *Flavobacteriaceae* whose sequences were more abundant from November 2017 until June 2018.
 In contrast, uncultured *Rhodothermaceae* showed higher proportions from June 2018 until the end of the study period.
-Surface associated *Bacteroidota* communities were very diverse as could be observed in the high proportion of taxa that grouped as other *Bacteroidota* (\autoref{bactero}).
+Surface associated *Bacteroidota* communities were very diverse as observed in the high proportion of taxa clustering as other *Bacteroidota* (\autoref{bactero}).
 
@@
 
-On average, *Alphaproteobacteria* were in comparison to other high rank taxa the largest taxonomic group, comprising `r alpha_all_whole$mean` ± `r alpha_all_whole$sd` \si{\percent} of all sequences (\autoref{alpha}).
-In accordance to previous taxa, high differences between seawater and surface associated communities were observed.
-Picoplankton communities were composed mainly of the SAR11 clade, AEGEAN-169 marine group, SAR116 clade, no relative *Rhodobacteraceae*, HIMB11 and OCS116 clade, while surface associated communities were composed in high proportion of no relative *Rhodobacteraceae* and to a lesser degree of  *Pseudoahrensia*, no relative *Alphaproteobacteria*, no relative  *Hyphomonadaceae* and *Amylibacter*.
-Representatives of no relative *Rhodobacteraceae* were comprising on average `r rhodo$mean` ± `r rhodo$sd` \si{\percent} of all alphaproteobacterial sequences from the epiphytic community (\autoref{alpha}).
+On average, *Alphaproteobacteria* were in comparison to the other high rank taxa the largest taxonomic group, comprising `r alpha_all_whole$mean` ± `r alpha_all_whole$sd` \si{\percent} of all sequences (\autoref{alpha}).
+In accordance to the above described taxa, large differences between ambient seawater and surface associated communities were observed.
+Ambient seawater communities were composed mainly of the SAR11 clade, AEGEAN-169 marine group, SAR116 clade, sequences without known relatives within the *Rhodobacteraceae* (no relative *Rhodobacteraceae*), HIMB11 and the OCS116 clade, while surface associated communities were composed mainly of no relative *Rhodobacteraceae* and to a lesser degree of *Pseudoahrensia*, *Amylibacter* and sequences without known relatives within the *Alphaproteobacteria* (no relative *Alphaproteobacteria*) and *Hyphomonadaceae* (no relative  *Hyphomonadaceae*).
+Representatives of no relative *Rhodobacteraceae* comprised on average `r rhodo$mean` ± `r rhodo$sd` \si{\percent} of all alphaproteobacterial sequences in the epiphytic community (\autoref{alpha}).
 In addition, *Amylibacter* was detected mainly in *C.
@@
 
-Sequences related to *Gammaproteobacteria* were comprising, on average, `r gamma_all_whole$mean` ± `r gamma_all_whole$sd` \si{\percent} of all sequences (\autoref{gamma}).
-Similarly to previous taxa, large taxonomic differences between seawater and surface associated communities were found.
-Seawater communities were mainly comprised of the OM60 (NOR5) clade, *Litoricola*, *Acinetobacter* and the SAR86 clade, while epiphytic communities were mainly composed of no relative *Gammaproteobacteria* and *Granulosicoccus*.
+Sequences related to *Gammaproteobacteria* comprised on average `r gamma_all_whole$mean` ± `r gamma_all_whole$sd` \si{\percent} of all sequences (\autoref{gamma}).
+Similar to above mentioned taxa, large taxonomic differences between ambient seawater and surface associated communities were found.
+Ambient seawater communities were mainly comprised of the OM60 (NOR5) clade, *Litoricola*, *Acinetobacter* and the SAR86 clade, while epiphytic communities were mainly composed of sequences without known relatives within the *Gammaproteobacteria* (no relative *Gammaproteobacteria*) and *Granulosicoccus*.
 Beside these two groups specific to all three epiphytic communities, *C.
-nodosa* was characterized by *Arenicella*, no relative *Burkholderiales* and *Methylotenera*, while *Thioploca*, no relative *Cellvibrionaceae* and *Reinekea* were more specific to both mixed and monospecific *C.
+nodosa* was characterized by *Arenicella*, *Methylotenera* and sequences without known relatives within the *Burkholderiales* (no relative *Burkholderiales*), while *Thioploca*, *Reinekea* and sequences without known relatives within *Cellvibrionaceae* (no relative *Cellvibrionaceae*) were more specific to both mixed and monospecific *C.
 cylindracea*.
-In addition, *Arenicella* was more pronounced in November and December 2017, while no relative *Burkholderiales* and *Methylotenera* were more characteristic for the period form March until May 2018.
+In addition, *Arenicella* was more pronounced in November and December 2017, while no relative *Burkholderiales* and *Methylotenera* were characteristic for the period from March until May 2018.
 For the *C.
-cylindracea* specific taxa no relative *Cellvibrionaceae* and *Reinekea* showed some seasonality and were characteristic for samples originating from June to October 2018.
-In addition, similarly to *Bacteroidota*, a large proportion of the surface associated community was grouped as other *Gammaproteobacteria* indicating high diversity within this group (\autoref{gamma}).
+cylindracea* specific taxa no relative *Cellvibrionaceae* and *Reinekea* showed seasonality and were characteristic for samples originating from June to October 2018.
+In addition, similar to *Bacteroidota*, a large proportion of the surface associated community was grouped as other *Gammaproteobacteria* indicating high diversity within this group (\autoref{gamma}).
 
@@
 
-# Calculating the mean and standard deviation of all samples
-desulfo_all_whole <- inner_join(metadata, whole, by=c("ID"="Group")) %>%
-  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
-  summarise(mean=mean(abundance), sd=sd(abundance))
-
 # Calculating the mean and standard deviation for each station
@@
 
-In contrast to previously described high rank taxa, *Desulfobacterota* were specific to *C.
+*Desulfobacterota* were specific for *C.
 cylindracea*.
-On average they comprised `r desulfo_all_whole$mean` ± `r desulfo_all_whole$sd` \si{\percent} of all sequences.
-Seawater and *C.
-nodosa* communities consisted of only `r filter(desulfo_station_whole, station=="F")$mean` ± `r filter(desulfo_station_whole, station=="F")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCyM")$mean` ± `r filter(desulfo_station_whole, station=="FCyM")$sd` \si{\percent} *Desulfobacterota* sequences, respectively.
 In the mixed and monospecific *C.
-cylindracea* communities their proportion was `r filter(desulfo_station_whole, station=="FCaM")$mean` ± `r filter(desulfo_station_whole, station=="FCaM")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCa")$mean` ± `r filter(desulfo_station_whole, station=="FCa")$sd` \si{\percent}, respectively (\autoref{desulfo}).
-The community consisted mainly of no relative *Desulfobacteraceae*, *Desulfatitalea*, no relative *Desulfobulbaceae*, *Desulfobulbus*, no relative *Desulfocapsaceae*, *Desulfopila*, *Desulforhopalus*, *Desulfotalea*, SEEP-SRB4 and uncultured *Desulfocapsaceae* (\autoref{desulfo}).
+cylindracea* communities the proportion of *Desulfobacterota* was `r filter(desulfo_station_whole, station=="FCaM")$mean` ± `r filter(desulfo_station_whole, station=="FCaM")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCa")$mean` ± `r filter(desulfo_station_whole, station=="FCa")$sd` \si{\percent}, respectively (\autoref{desulfo}).
+In contrast, in ambient seawater and *C.
+nodosa* communities the contribution of *Desulfobacterota* was only `r filter(desulfo_station_whole, station=="F")$mean` ± `r filter(desulfo_station_whole, station=="F")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCyM")$mean` ± `r filter(desulfo_station_whole, station=="FCyM")$sd` \si{\percent}, respectively.
+In *C.
+cylindracea* the community consisted mainly of *Desulfatitalea*, *Desulfobulbus*, *Desulfopila*, *Desulforhopalus*, *Desulfotalea*, SEEP-SRB4, uncultured *Desulfocapsaceae* and sequences without known relatives within the *Desulfobacteraceae* (no relative *Desulfobacteraceae*), *Desulfobulbaceae* (no relative *Desulfobulbaceae*) and *Desulfocapsaceae* (no relative *Desulfocapsaceae*) (\autoref{desulfo}).
 
@@
 Surfaces of marine macrophytes harbor biofilms consisting of diverse microbial taxa [@Egan2013; @Tarquinio2019].
-No standard protocol has been developed to study these macophyte associated microbes [@Ugarelli2019].
-Different procedures for removal of microbial cells from host surfaces were described, such as host tissue shaking [@Noges2010], scraping [@Uku2007] and ultrasonication [@Cai2014].
-All these methods showed different removal efficiencies but none was enabling a complete removal of attached microbial cells.
-In the present study, we applied an earlier developed removal protocol [@Korlevica, submitted], based on a previous idea of direct cellular lysis [@Burke2009], to ensure an almost complete cell detachment.
-The application of a direct lysis procedure coupled with a high frequency sampling protocol and Illumina high resolution amplicon sequencing has enabled us to make a detailed description of bacterial and archaeal communities associated with the surfaces of two marine macrophytes, *C.
+No standard protocol has been developed to study these macophyte-associated microbes [@Ugarelli2019].
+Different procedures for removal of microbial cells from host surfaces are described, such as host tissue shaking [@Noges2010], scraping [@Uku2007], swabbing [@Mancuso2016] and ultrasonication [@Cai2014].
+All these methods result in different removal efficiencies but none was enabling a complete removal of attached microbial cells based on our experience.
+In the present study, we applied a removal protocol [@Korlevica, submitted] based on direct cellular lysis [@Burke2009].
+The application of a direct lysis procedure coupled with a high sampling frequency and Illumina amplicon sequencing has enabled us to described in detail the bacterial and archaeal communities associated with the surfaces of two marine macrophytes, *C.
 nodosa* and *C.
@@
 
-In the present study, highest richness values were observed for *C.
-cylindracea* (mixed and monospecific), middle for *C.
-nodosa* and lowest for seawater derived communities.
-Higher values for seagrass associated communities in comparison to seawater were described  earlier and could be attributed to a larger set of inhabitable microniches existing on macrophyte surfaces [@Ugarelli2019].
-In addition, highest values observed for *C.
-cylindracea* are partly due to its contact with the sediment.
-*C.
-cylindracea* stolon is attached to the sediment surface with rhizoids, so the stolon and rhizoids are in a direct contact with the sediment.
-In addition, seasonal richness differences observed for surface attached communities showed slightly higher values in spring and summer.
-This pattern could be explained by a higher macrophyte growth in these seasons [M.
+In the present study, highest richness was observed for *C.
+cylindracea* (mixed and monospecific) followed by *C.
+nodosa* and lowest richness was found in ambient seawater microbial communities.
+Higher richness of microbial communities associated with seagrasses than in ambient seawater was described earlier and could be attributed to a larger set of inhabitable microniches existing on macrophyte surfaces than in the ambient seawater [@Ugarelli2019].
+The highest richness observed for *C.
+cylindracea* might be partly due to its contact with the sediment.
+The stolon of *C.
+cylindracea* is attached to the sediment surface with rhizoids and thus, the stolon and rhizoids are in a direct contact with the sediment.
+In addition, seasonal differences in richness observed for surface attached communities indicated a slightly higher richness in spring and summer.
+This pattern could be explained by a higher macrophyte growth in these two seasons than in autumn and winter [M.
 Najdek, personal communication; @Zavodnik1998; @Ruitton2005].
-During active periods macrophytes exhibit a more dynamic chemical interaction with the surface community probably causing an increase in the number of inhabitable microniches [@Rickert2016; @Borges2015].
-
-Since the colonization of macrophyte surfaces is performed from a pool of prokaryotic cells from the ambient seawater, it was interesting to see to which extent these two communities differ.
-We observed a strong differentiation between the surface attached and seawater communities at the level of OTUs that is in agreement with most published studies [@Ugarelli2019; @Roth-Schulze2016; @Burke2011a; @Michelou2013; @Crump2018].
-These data indicate that marine macrophytes are selecting, from a pool of seawater microbial taxa, the one that can colonize and proliferate on their surfaces [@Salaun2012; @Michelou2013].
-In contrast to these findings @Fahimipour2017 found, in a global study of *Zostera marina*, similarities between leaves and seawater samples.
-Discrepancies between our data and this study could be explained by differences in studied seagrass species, methodological variations or biogeographic trends as @Fahimipour2017 were analyzing samples from different locations throughout the northern hemisphere.
-It is possible that ambient seawater and leaves communities from the same location are differing but are still more similar to each other when compared to other sampling locations.
+During their main growth season in spring and summer macrophytes exhibit a more dynamic chemical interaction with the surface community probably causing an increase in the number of inhabitable microniches [@Rickert2016; @Borges2015].
+
+We observed a strong differentiation between the surface attached and ambient seawater communities at the level of OTUs, in agreement with most published studies [@Ugarelli2019; @Roth-Schulze2016; @Burke2011a; @Michelou2013; @Crump2018; @Mancuso2016].
+This indicates that marine macrophytes are a selecting factor from the pool of microbial taxa present in the ambient seawater, modifying the microbial community once the macrophyte associated microbial biofilm develops [@Salaun2012; @Michelou2013].
+In contrast, @Fahimipour2017 report in a global study of *Zostera marina*, similarities between the microbial community developed on leaves and in the ambient seawater.
+The discrepancy between our data and the study of @Fahimipour2017 could be explained by different seagrass species, methodological variations or biogeographic trends as @Fahimipour2017 analyzed samples from different locations throughout the northern hemisphere.
+It is possible that the microbial communities in ambient seawater and on leaves from the same location are differing but are still more similar to each other when compared to other sampling locations.
 Indeed, it was found that prokaryotic communities vary substantially between different sampling sites [@Bengtsson2017].
 When the taxonomic composition at high ranks was analyzed no such strong differentiation was noticed.
-Phyla and classes such as: *Actinobacteriota*, *Bacteroidota*, *Cyanobacteria*, *Alphaproteobacteria*, *Gammaproteobacteria* and *Verrucomicrobiota* were described that is in agreement with previously reported data [@Egan2013; @Michelou2013; @Burke2011a].
-In contrast, when low taxonomic ranks were analyzed (i.g.
-family and genus) a strong differentiation was observed.
-A similar differentiation at lower taxonomic ranks was described for other species of macrophytes [@Egan2013; @Michelou2013; @Ugarelli2019].
-
-Beside differences between seawater and surface associated communities, there were discussions if the prokaryotic epiphytic community is host-specific or there are generalists taxa characteristic to all or many macrophytes [@Egan2013].
-Similarly to previously described differences between seawater and surface attached communities, at high taxonomic ranks no strong differentiation between communities associated with different host was observed.
+Phyla and classes such as *Actinobacteriota*, *Bacteroidota*, *Cyanobacteria*, *Alphaproteobacteria*, *Gammaproteobacteria* and *Verrucomicrobiota* were found in both ambient seawater as well as macrophyte associated, in agreement with previous studies [@Egan2013; @Michelou2013; @Burke2011a].
+In contrast, when low taxonomic ranks were analyzed (i.e., family and genus) a strong differentiation was observed (Figs.
+\ref{cyano} -- \ref{desulfo}).
+A similar differentiation at lower microbial taxonomic ranks between ambient seawater and macrophytes was described for other macrophyte species as well [@Egan2013; @Michelou2013; @Ugarelli2019; @Mancuso2016].
+
+Beside differences between ambient seawater and surface associated microbial communities, it is unclear whether the prokaryotic epiphytic community is host-specific or whether there are generalist taxa characteristic to all or many macrophytes [@Egan2013].
+Similar to previously described differences between microbial communities in the ambient seawater and on macrophytes, at high taxonomic ranks no major difference between the microbial communities associated with different hosts was observed.
 The only high rank phylum that was differing between *C.
 nodosa* and *C.
-cylindracea* was *Desulfobacterota*, whose sequences were more abundant in the *C.
+cylindracea* was *Desulfobacterota*, with more abundant sequences in the *C.
 cylindracea* associated community.
-As already mentioned, the rhizoids and part of the stolon are in contact with the sediment, so *Desulfobacterota* are probably a part of the epiphytic community that was in contact with the sediment.
-Similar high rank taxa found in this study were described to be specific for other species of macrophytes [@Burke2011a; @Lachnit2011; @Bengtsson2017].
-In contrast to high taxonomic ranks, a substantial differentiation between host specific communities was found, which supports the host-specific hypothesis.
-Similar host-specificity was observed for different species of macroalgae and seagrasses [@Lachnit2011; @Roth-Schulze2016; @Ugarelli2019; @Morrissey2019].
-Taken together, at high taxonomic ranks a core set of taxa could be described that is characteristic for all or many macrophytes, while at low taxonomic ranks a community specific to host species could be identified [@Egan2013].
-
-Seasonal richness changes in the epiphytic community were substantial as could be observed in the proportion of OTUs that could be found at every sampling time ($\leq$ `r max(core$core_otus)` \si{\percent}).
-Interestingly, these OTUs were accounting for a high proportion of sequences ($\leq$ `r max(core$core_seq)` \si{\percent}).
-A very similar proportion of persistent OTUs and their sequence contribution was reported in high frequency studies describing seasonal picoplankton changes [@Gilbert2009; @Gilbert2012].
-In comparison to the seawater community, a lower degree of seasonal shifts was observed for the surface associated communities.
-It seems, microniches on the surfaces of macrophytes are providing more stable conditions in comparison to the seawater.
+As already mentioned, the rhizoids and part of the stolon are in contact with the sediment.
+Thus *Desulfobacterota* are probably a part of the epiphytic community that was in contact with the sediment.
+Similar high rank taxa found in this study were described to be specific for other species of macrophytes [@Burke2011a; @Lachnit2011; @Bengtsson2017; @Mancuso2016].
+In contrast to high taxonomic ranks, a substantial differentiation between host specific communities was found supporting the notion that macrophyte associated microbial communities might be host-specific.
+Host-specificity was also observed for other species of macroalgae and seagrasses [@Lachnit2011; @Roth-Schulze2016; @Ugarelli2019; @Morrissey2019].
+Taken together, at high taxonomic ranks a core set of taxa could be described that is characteristic for all or many macrophytes, while at low taxonomic ranks a community specific to host species was identified (Figs.
+\ref{pcoa} and \ref{community}) [@Egan2013].
+
+Seasonal changes in richness in the epiphytic community were substantial as indicated by the proportion of OTUs ($\leq$ `r max(core$core_otus)` \si{\percent}) present at every sampling date.
+These persistent OTUs, however, were accounting for a high proportion of sequences ($\leq$ `r max(core$core_seq)` \si{\percent}) (\autoref{shared}).
+A very similar proportion of persistent OTUs was reported in high-frequency sampling studies describing seasonal changes in picoplankton [@Gilbert2009; @Gilbert2012].
+In comparison to the seawater community, a lower degree of seasonal shifts was observed for the macrophyte surface associated communities.
+It appears that microniches at the surfaces of macrophytes are providing more stable conditions than the ambient seawater.
 At the level of OTUs seasonal changes of *C.
@@
 *C.
-nodosa* was characterized by a Spring community during maximum seagrass proliferation, a Summer community during a biomass maximum and a Autumn/Winter community during a biomass senescence.
+nodosa* was characterized by a spring community during maximum seagrass proliferation, a summer community during the highest standing stock of *C.nodosa* and an autumn/winter community during the decay of seagrass biomass.
 In contrast, *C.
-cylindracea* started to proliferate in late Spring and was characterized only by a Summer community during maximal biomass increase and by a Autumn/Winter/Spring community when the biomass was at the peak and the settlement started to subsequently decay.
-Similar seasonal changes in the epiphytic community was also described for other macroalgae [@Lachnit2011; @Tujula2010].
-Higher temporal stability of *C.
-cylindracea* surface communities in comparison to *C.
-nodosa* were also observed in the higher proportion of shared communities between two consecutive sampling points.
-
-Analysis of seasonal chloroplast sequence abundances showed higher values in Autumn/Winter in comparison to Spring/Summer.
-This pattern is not surprising as seagrasses are known to harbor more algal epiphytes during Autumn/Winter [@Reyes2001].
-Furthermore, we used an adapted DNA isolation protocol that is known to partially coextract DNA from planktonic eukaryotes [@Korlevic2015].
-Strong seasonal fluctuations of high rank epiphytic taxa were not observed, with the exception of *Cyanobateria*.
-Cyanobacterial sequences were more pronounced in November and December in comparison to Spring and Summer.
-Interestingly, in these high proportion months the majority of cyanobacterial sequences were classified as *Pleurocapsa*, a group known to colonized different living and nonliving surfaces [@Longford2007; @Burns2004; @Mobberley2012; @Reisser2014].
+cylindracea* started to proliferate in late spring and was characterized only by a summer community during high growth rates and by an autumn/winter/spring community when the biomass was at the peak and decaying thereafter.
+Similar seasonal changes in the epiphytic community were also described for other macroalgae [@Lachnit2011; @Tujula2010].
+Higher seasonal stability of *C.
+cylindracea* surface communities than in *C.
+nodosa* was also observed in the higher proportion of shared communities between two consecutive sampling dates in *C.
+cylindracea*.
+
+Chloroplast sequence abundances were higher in autumn/winter than in spring/summer.
+This pattern is not surprising as seagrasses harbor more algal epiphytes during autumn/winter than in spring/summer [@Reyes2001].
+Furthermore, we used an adapted DNA isolation protocol that is known to partially co-extract DNA from planktonic eukaryotes [@Korlevic2015].
+Strong seasonal fluctuations of high rank epiphytic taxa were not observed, with the exception of *Cyanobacteria*.
+Cyanobacterial sequences were more pronounced in November and December than in spring and summer.
+In the months of high cyanobacterial sequence abundances the majority of sequences from this group were classified as *Pleurocapsa*, a group known to colonized different living and non-living surfaces [@Longford2007; @Burns2004; @Mobberley2012; @Reisser2014].
 It is possible that during periods of low metabolic activity there is a reduced interaction and selection of the epiphytic community by the seagrass, causing leaves to become a suitable surface for nonspecific colonizers [@Zavodnik1998].
-*Pleurocapsa* was replaced in Spring and Summer by *Acrophormium*, *Phormidesmis* and no relative *Nodosilineaceae*.
-A study of coastal microbial mats found also higher proportion of *Nodosilineaceae* sequences in Summer, while *Phormidesmis* sequences were at their peak in Autumn [@Cardoso2019].
-Other high rank taxa did not show strong successional patterns.
-In every analyzed group, with the exception of *Desulfobacterota*, taxa present throughout the year in similar proportions and season specific taxa could be identified.
-Within *Bacteroidota* different groups withing the family *Saprospiraceae* (i.g.
-*Lewinella*, *Portibacter* and *Rubidimonas*) were detected through the year.
+*Pleurocapsa* was replaced in spring and summer by *Acrophormium*, *Phormidesmis* and sequences without known relatives within the *Nodosilineaceae*.
+A study of coastal microbial mats found also a higher proportion of *Nodosilineaceae* sequences in summer, while *Phormidesmis* sequences were at their peak in autumn [@Cardoso2019].
+Other high rank taxa did not exhibit strong successional patterns.
+In every analyzed group, with the exception of *Desulfobacterota*, taxa present throughout the year in similar proportions and season specific taxa could be identified (Figs.
+\ref{bactero} and \ref{desulfo}).
+Within *Bacteroidota* different groups within the family *Saprospiraceae* (e.g.
+*Lewinella*, *Portibacter* and *Rubidimonas*) were detected across all seasons.
 Members of this family are often found in association with macrophytes and it is suggested that they are involved in the hydrolysis and utilization of complex carbon sources [@Burke2011a; @Crump2018; @McIlroy2014].
-On the other hand, families *Flavobacteriaceae* and *Rhodothermaceae* showed seasonal patterns, with *Flavobacteriaceae* being more pronounced from November to June and *Rhodothermaceae* from June to October.
-Within *Alphaproteobacteria* the family *Rhodobacteraceae* was comprising the majority of sequences throughout the year.
-This metabolically versatile family is often associated with macrophyte surfaces and usually is one of the most abundant groups [@Burke2011a; @Michelou2013; @Pujalte2014].
+In contrast, the families *Flavobacteriaceae* and *Rhodothermaceae* showed seasonal patterns, with *Flavobacteriaceae* being more pronounced from November to June and *Rhodothermaceae* from June to October (\autoref{bactero}).
+Within *Alphaproteobacteria* the family *Rhodobacteraceae* comprised the majority of sequences throughout the year (\autoref{alpha}).
+This metabolically versatile family is often associated with macrophyte surfaces and usually is one of the most abundant groups [@Burke2011a; @Michelou2013; @Pujalte2014; @Manucos2016].
 In addition, *Hyphomonadaceae* were found in all samples.
-Interestingly, some of the species within this group contain stalks on their cells which can be used to attach to the macrophyte surface [@Weidner2000; @Abraham2014].
-Within the *Gammaproteobacteria*, sequences without known representatives were the most pronounced group present throughout the year.
+Interestingly, some of the species within this group contain stalks on their cells, which can be used to attach to the macrophyte surface [@Weidner2000; @Abraham2014].
+Within the *Gammaproteobacteria*, sequences without known representatives were the most pronounced group present throughout the year (\autoref{gamma}).
 In addition, *Granulosicoccus* was also found in almost all samples.
 *Gammaproteobacteria* are often a major constituent of macrophyte epiphytic communities [@Michelou2013; @Burke2011a; @Crump2018].
-Beside these two groups other less pronounced taxa showed seasonal and host-specific patterns.
+Beside these two groups, other less abundant, taxa showed seasonal and host-specific patterns.
 For example, *C.
-cylindracea* was characterized by *Thioploca*, a known sulfur sediment bacteria and *Cellvibrionaceae*, a family whose cultured members are known polysaccharide degraders [@Jorgensen1999; @Xie2017].
+cylindracea* harbored *Thioploca*, a known sulfur sediment bacteria and *Cellvibrionaceae*, a family with cultured members known as polysaccharide degraders [@Jorgensen1999; @Xie2017].
 *Desulfobacterota* were found only associated with *C.
-cylindracea* and no group within this phylum showed seasonal patterns.
+cylindracea* and no group within this phylum showed seasonal patterns (\autoref{desulfo}).
 The presence of this phylum only on *C.
-cylindracea* is to be expected as part of the epiphytic community is directly in contact with the sediment.
+cylindracea* is to be expected as part of the epiphytic community is in direct contact with the sediment.
 The *Desulfobacterota* community was dominated by *Desulfatitalea* and no relative *Desulfocapsaceae*, known sulfate sediment groups [@Higashioka2015; @Kuever2014].
 
-In temperate zones marine macrophytes are exhibiting growth cycles, so it  is not surprising that the associated epiphytic microbial community is undergoing partial seasonal changes.
-In the present study, we could, in every analyzed high rank taxa, identify phylogenetic groups that were present throughout the year and that were comprising most of the sequences and lower proportion taxa showing seasonal patterns connected to the macrophyte growth cycle.
+In temperate zones, marine macrophytes are exhibiting growth cycles, so it  is not surprising that the associated epiphytic microbial community is undergoing partial seasonal changes.
+In the present study, we could identify in every analyzed high rank taxa phylogenetic groups present throughout the year, comprising most of the sequences and a lower proportion of taxa showing seasonal patterns connected to the macrophyte growth cycle (Figs.
+\ref{community} and \ref{desulfo}).
 Studies focusing on functional comparisons between communities associated with different hosts showed that the majority of functions could be found in every community, indicating functional redundancy [@Roth-Schulze2016].
-This difference between taxonomic and functional discrepancy was explained by the lottery hypothesis that hypothesizes an initial random colonization step performed from a set of functionally equivalent taxonomic groups [@Burke2011; @Roth-Schulze2016].
+This difference between phylogenetic varaibility and functional stability was explained by the lottery hypothesis assuming an initial random colonization step performed by a set of functionally equivalent taxonomic groups [@Burke2011; @Roth-Schulze2016].
 It is possible that functional redundancy is a characteristic of high abundance taxa detected to be present throughout the year, while seasonal and/or host-specific functions are an attribute of taxa displaying successional patterns.
@@
 Leaves and thalli were sampled in a meadow of *Cymodocea nodosa* invaded by the invasive *Caulerpa cylindracea* (mixed settlement) and in a monospecific settlement of *Caulerpa cylindracea* located in the proximity of the meadow at approximately monthly intervals from December 2017 to October 2018 (\autoref{supp-nseq_notus}).
-Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with site seawater.
+Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with seawater collected at the sampling site.
 Upon arrival to the laboratory, *C.
@@
 Briefly, leaves and thalli are incubated in a lysis buffer and treated with lysozyme and proteinase K.
-Following the incubations, the mixture containing lyzed epiphytic cells is separated from leaves and thalli and extracted using a phenol-chloroform procedure.
-Finally, the extracted DNA is precipitated using isopropanol.
-DNA from seawater picoplankton was isolated from 0.2 \si{\um} polycarbonate filters according to @Massana1997 with a slight modification.
-Following the phenol-chloroform extraction steps 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added.
-DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 20,000 × g and 4 \si{\degreeCelsius} for 21 \si{\minute}.
-The pellet was washed twice with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}.
-Dried pellets were resuspended in 50 -- 100 \si{\ul} of deionized water.
+Following the incubations, the mixture containing lysed epiphytic cells was separated from the leaves and thalli and extracted using phenol-chloroform.
+Finally, the extracted DNA was precipitated using isopropanol.
+DNA from seawater picoplankton was extracted from 0.2 \si{\um} polycarbonate filters according to @Massana1997 with a slight modification.
+Following the phenol-chloroform extraction, 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added.
+DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 20,000 × *g* and 4 \si{\degreeCelsius} for 21 \si{\minute}.
+The pellet was washed twice with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged after each washing step at 20,000 × *g* and 4 \si{\degreeCelsius} for 5 \si{\minute}.
+Dried pellets were re-suspended in 50 -- 100 \si{\ul} of deionized water.
 
 ### Illumina 16S rRNA sequencing
-Illumina MiSeq sequencing of the V4 16S rRNA region was performed as described previously [@Korlevica, submitted].
+Illumina MiSeq sequencing of the V4 region of the 16S rRNA gene was performed as described previously [@Korlevica, submitted].
 The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
-In the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used [@Caporaso2012; @Apprill2015; @Parada2016].
+In the first PCR, the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used [@Caporaso2012; @Apprill2015; @Parada2016].
 These primers contained on their 5' end a tagged sequence.
 Purified PCR products were sent for Illumina MiSeq sequencing at IMGM Laboratories, Martinsried, Germany.
-Before sequencing at IMGM, the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR.
+Prior to sequencing at IMGM, the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR.
 In addition, these primers contained adapter and sample-specific index sequences.
 Beside samples, a positive and negative control for each sequencing batch was sequenced.
-Negative control was comprised of PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
+The negative control comprised PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
 Sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession number PRJEB37267.
@@
 
-Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
+Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations provided by the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
 For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
-Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019],  packages  vegan (version 2.5-6) [@Oksanen2019], and tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014].
-The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/mkorlevic/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020).
+Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019] combined with  packages  vegan (version 2.5-6) [@Oksanen2019], tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014].
+The detailed analysis procedure including the R Markdown file are available in the GitHub repository (https://github.com/mkorlevic/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020).
 Based on the ATCC MSA-1002 mock community included in the analysis an average sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
@@
 This work was founded by the Croatian Science Foundation through the MICRO-SEAGRASS project (IP-2016-06-7118).
-We would like to thank Margareta Buterer for technical support, Paolo Paliaga for help during sampling and the University Computing Center of the University of Zagreb for access to the cluster Isabella.
+We would like to thank Margareta Buterer for technical support, Paolo Paliaga for help during sampling and the University Computing Center of the University of Zagreb for access to the computer cluster Isabella.
 
@@
 ## Figures
-```{r out.width="70%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between assemblages associated with the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed Settlement] and \\textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and comunities in the ambient seawater.\\label{matrix}", fig.pos="H"}
+```{r out.width="70%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between assemblages associated with the surfaces of the macrophytes \\textit{C.
+nodosa} (Mixed Settlement) and \\textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and communities in the ambient seawater.\\label{matrix}", fig.pos="H"}
 include_graphics("../results/figures/matrix.jpg")
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed Settlement] and \\textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.\\label{shared}", fig.pos="H"}
+```{r out.width="85%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling dates and from the surfaces of the macrophytes \\textit{C.
+nodosa} (Mixed Settlement) and \\textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.\\label{shared}", fig.pos="H"}
 include_graphics("../results/figures/seasonal_shared.jpg")
@@
 
-```{r out.width="55%", fig.align="center", fig.cap="Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed Settlement] and \\textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
+```{r out.width="55%", fig.align="center", fig.cap="Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of the macrophytes \\textit{C.
+nodosa} (Mixed Settlement) and \\textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
 Samples from the same environment or same season are labeled in different colors.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) bacterial and archaeal sequences on the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed Settlement] and \\textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
-NR -- No Relative\\label{community}", fig.pos="H"}
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) bacterial and archaeal sequences on the surfaces of the macrophytes \\textit{C.
+nodosa} (Mixed Settlement) and \\textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
+NR -- No Relative (sequences without known relatives within the corresponding group)\\label{community}", fig.pos="H"}
 include_graphics("../results/figures/community_bar_plot.jpg")
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) cyanobacterial sequences on the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed Settlement] and \\textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) cyanobacterial sequences on the surfaces of the macrophytes \\textit{C.
+nodosa} (Mixed Settlement) and \\textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
 The proportion of cyanobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
-NR -- No Relative\\label{cyano}", fig.pos="H"}
+NR -- No Relative (sequences without known relatives within the corresponding group)\\label{cyano}", fig.pos="H"}
 include_graphics("../results/figures/cyanobacteria_bar_plot.jpg")
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) sequences within the \\textit{Bacteroidota} on the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed Settlement] and \\textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) sequences within the \\textit{Bacteroidota} on the surfaces of the macrophytes \\textit{C.
+nodosa} (Mixed Settlement) and \\textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
 The proportion of sequences classified as \\textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar.
-NR -- No Relative\\label{bactero}", fig.pos="H"}
+NR -- No Relative (sequences without known relatives within the corresponding group)\\label{bactero}", fig.pos="H"}
 include_graphics("../results/figures/bacteroidota_bar_plot.jpg")
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) alphaproteobacterial sequences on the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed Settlement] and \\textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) alphaproteobacterial sequences on the surfaces of the macrophytes \\textit{C.
+nodosa} (Mixed Settlement) and \\textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
 The proportion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
-NR -- No Relative\\label{alpha}", fig.pos="H"}
+NR -- No Relative (sequences without known relatives within the corresponding group)\\label{alpha}", fig.pos="H"}
 include_graphics("../results/figures/alphaproteobacteria_bar_plot.jpg")
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) gammaproteobacterial sequences on the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed Settlement] and \\textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) gammaproteobacterial sequences on the surfaces of the macrophytes \\textit{C.
+nodosa} (Mixed Settlement) and \\textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
 The proportion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
-NR -- No Relative\\label{gamma}", fig.pos="H"}
+NR -- No Relative (sequences without known relatives within the corresponding group)\\label{gamma}", fig.pos="H"}
 include_graphics("../results/figures/gammaproteobacteria_bar_plot.jpg")
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) sequences within the \\textit{Desulfobacterota} on the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed Settlement] and \\textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) sequences within the \\textit{Desulfobacterota} on the surfaces of the macrophytes \\textit{C.
+nodosa} (Mixed Settlement) and \\textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
 The proportion of sequences classified as \\textit{Desulfobacterota} in the total bacterial and archaeal community is given above the corresponding bar.
-NR -- No Relative\\label{desulfo}", fig.pos="H"}
+NR -- No Relative (sequences without known relatives within the corresponding group)\\label{desulfo}", fig.pos="H"}
 include_graphics("../results/figures/desulfobacterota_bar_plot.jpg")

### 2020-05-28 submission/supplementary.Rmd
First review.

@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed Settlement] and \\textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.\\label{rarefaction}", fig.pos="H"}
+```{r out.width="85%", fig.align="center", fig.cap="Rarefaction curves of bacterial and archaeal communities from the surfaces of the macrophytes \\textit{C.
+nodosa} (Mixed Settlement) and \\textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.\\label{rarefaction}", fig.pos="H"}
 include_graphics("../results/figures/rarefaction.jpg")
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed Settlement] and \\textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.\\label{calculators}", fig.pos="H"}
+```{r out.width="85%", fig.align="center", fig.cap="Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of the macrophytes \\textit{C.
+nodosa} (Mixed Settlement) and \\textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.\\label{calculators}", fig.pos="H"}
 include_graphics("../results/figures/calculators.jpg")
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Relative contribution of chloroplast sequences on the surfaces of macrophytes (\\textit{C.
-nodosa} [Mixed Settlement] and \\textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.\\label{chloroplast}", fig.pos="H"}
+```{r out.width="85%", fig.align="center", fig.cap="Relative contribution of chloroplast sequences on the surfaces of the macrophytes \\textit{C.
+nodosa} (Mixed Settlement) and \\textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.\\label{chloroplast}", fig.pos="H"}
 include_graphics("../results/figures/chloroplast_bar_plot.jpg")
@@
   mutate(station=str_replace(station, "^FCa$", "\\\\textit{Caulerpa cylindracea} (Monospecific)")) %>%
+  mutate(season=str_replace(season, "^Summer$", "summer")) %>%
+  mutate(season=str_replace(season, "^Autumn$", "autumn")) %>%
+  mutate(season=str_replace(season, "^Winter$", "winter")) %>%
+  mutate(season=str_replace(season, "^Spring$", "spring")) %>%
   mutate(ID=str_replace(ID, "^23_1$", "23a")) %>%
@@
 of OTUs of each sample.
-No.
-of Sequences and OTUs was calculated after the exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences.\\label{nseq_notus}", align="c") %>%
+The number of sequences and OTUs was calculated after exclusion of sequences without known relatives (no relative sequences) and eukaryotic, chloroplast and mitochondrial sequences.\\label{nseq_notus}", align="c") %>%
   kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9) %>%

### 2020-05-28 submission/manuscript.Rmd
Review optimization.

@@
 \ref{cyano} -- \ref{desulfo}).
-A similar differentiation at lower microbial taxonomic ranks between ambient seawater and macrophytes was described for other macrophyte species as well [@Egan2013; @Michelou2013; @Ugarelli2019; @Mancuso2016].
+A similar differentiation at lower microbial taxonomic ranks between ambient seawater and macrophytes was described for other macrophyte species as well [@Egan2013; @Michelou2013; @Ugarelli2019].
 

### 2020-05-28 submission/manuscript.tex
Review optimization.

@@
 
-3. Department of Marine Microbiology and Biogeochemistry, Royal
+3. NIOZ, Department of Marine Microbiology and Biogeochemistry, Royal
 Netherlands Institute for Sea Research, Utrecht University, The
@@
 Surfaces of marine macrophytes (seagrasses and macroalgae) are inhabited
-by diverse microbial communities. Most studies focusing on macrophyte
-epiphytic communities did not take into account temporal changes or
+by diverse microbial communities. Most studies focusing on epiphytic
+communities of macrophytes did not take into account temporal changes or
 applied low sampling frequency approaches. Illumina sequencing of the V4
 16S rRNA region was performed to determine the seasonal dynamics of
-epiphytic communities sampled from the surfaces of the seagrass
-\emph{Cymodocea nodosa} and invasive macroalga \emph{Caulerpa
-cylindracea}. Leaves and thalli were sampled in a meadow of
-\emph{Cymodocea nodosa} invaded by the invasive \emph{Caulerpa
-cylindracea} and in a monospecific settlement of \emph{Caulerpa
-cylindracea} located in the proximity of the meadow at monthly
-intervals. For comparison the ambient prokaryotic plankton community was
-also characterized. Sequencing results at the OTU level showed a clear
-differentiation between ambient water and epiphytic communities and a
-host-specific community assemblage. In addition, successional changes
-were observed that could be connected to the macrophyte growth cycle.
-Taxonomic analysis showed similar high rank groups in the ambient water
-and epiphytic communities, with the exception of \emph{Desulfobacterota}
-that were found only on \emph{Caulerpa cylindracea}. Only
-\emph{Cyanobacteria} showed seasonal change, while other high rank taxa
-were present throughout the year. In every analyzed high rank taxa,
-phylogenetic groups present throughout the year comprised most of the
-sequences and could be identified together with low proportion taxa
-showing seasonal patterns connected to the macrophyte growth cycle.
-Taken together, epiphytic microbial communities of the seagrass
-\emph{Cymodocea nodosa} and the macroalgae \emph{Caulerpa cylindracea}
-appear to be host-specific and contain taxa that undergo successional
-changes.
+epiphytic communities of the seagrass \emph{Cymodocea nodosa} and
+invasive macroalga \emph{Caulerpa cylindracea}. Leaves and thalli were
+sampled in a meadow of \emph{Cymodocea nodosa} invaded by the invasive
+\emph{Caulerpa cylindracea} and in a monospecific settlement of
+\emph{Caulerpa cylindracea} located in the proximity of the meadow at
+monthly intervals. For comparison the ambient prokaryotic plankton
+community was also characterized. At the OTU level, the microbial
+community composition differed between the ambient water and the
+epiphytic communities exhibiting host-specificity. Also, successional
+changes were observed connected to the macrophyte growth cycle.
+Taxonomic analysis, however, showed similar high rank groups in the
+ambient water and the epiphytic communities, with the exception of
+\emph{Desulfobacterota}, which were only found on \emph{Caulerpa
+cylindracea}. \emph{Cyanobacteria} showed seasonal changes while other
+high rank taxa were present throughout the year. Phylogenetic groups
+present throughout the year constituted most of the sequences, while
+less abundant taxa showed seasonal patterns connected to the macrophyte
+growth cycle. Taken together, epiphytic microbial communities of the
+seagrass \emph{Cymodocea nodosa} and the macroalgae \emph{Caulerpa
+cylindracea} appear to be host-specific and contain taxa that undergo
+successional changes.
 
@@
 Marine macrophytes (seagrasses and macroalgae) are important ecosystem
-engineers that form close associations with microorganism belonging to
-all three domains of life (Egan \emph{et al.}, 2013; Tarquinio \emph{et
+engineers forming close associations with microorganism belonging to all
+three domains of life (Egan \emph{et al.}, 2013; Tarquinio \emph{et
 al.}, 2019). Microbes can live within macrophyte tissue as endophytes or
@@
 \emph{et al.}, 2013; Hollants \emph{et al.}, 2013; Tarquinio \emph{et
-al.}, 2019). Epiphytic and endophytic microbial communities form a close
-functional relationship with the macrophyte host. It was proposed that
-this close relationship constitutes a holobiont, an integrated community
-where the macrophyte organism and its symbiotic partners support each
-other (Margulis, 1991; Egan \emph{et al.}, 2013; Tarquinio \emph{et
-al.}, 2019).
-
-Biofilms formed from microbial epiphytes can contain diverse taxonomic
-groups and harbor cell densities from 10\textsuperscript{2} to
+al.}, 2019). Epiphytic and endophytic microbial communities exhibit a
+close functional relationship with the macrophyte host. It was proposed
+that this close relationship constitutes a holobiont, an integrated
+community where the macrophyte organism and its symbiotic partners
+support each other (Margulis, 1991; Egan \emph{et al.}, 2013; Tarquinio
+\emph{et al.}, 2019).
+
+Biofilms of microbial epiphytes can contain diverse taxonomic groups and
+harbor cell abundances from 10\textsuperscript{2} to
 10\textsuperscript{7} \si{\cells\per\cm\squared} (Armstrong \emph{et
@@
 al.}, 2011). In such an environment a number of positive and negative
-interactions between the macrophyte and colonizing microorganisms have
-been described (Egan \emph{et al.}, 2013; Hollants \emph{et al.}, 2013;
-Tarquinio \emph{et al.}, 2019). Macrophytes can promote growth of
+interactions between the macrophyte and the colonizing microorganisms
+have been described (Egan \emph{et al.}, 2013; Hollants \emph{et al.},
+2013; Tarquinio \emph{et al.}, 2019). Macrophytes can promote growth of
 associated microbes by nutrient exudation, while in return
 microorganisms may support macrophyte performance through improved
-nutrient availability, phytohormone production and protection form toxic
+nutrient availability, phytohormone production and protection from toxic
 compounds, oxidative stress, biofouling organisms and pathogens (Egan
 \emph{et al.}, 2013; Hollants \emph{et al.}, 2013; Tarquinio \emph{et
-al.}, 2019). Beside this positive interactions, macrophytes can
+al.}, 2019). Besides these positive interactions, macrophytes can
 negatively impact the associated microbes such as pathogenic bacteria by
@@
 Tujula \emph{et al.}, 2010; Lachnit \emph{et al.}, 2011). In contrast,
-at lower taxonomic ranks host specific microbial communities were
-described (Lachnit \emph{et al.}, 2011; Roth-Schulze \emph{et al.},
-2016). Recently, it was shown that even different morphological niches
-within the same alga had a higher influence on bacterial community
-variation than biogeography or environmental factors (Morrissey \emph{et
-al.}, 2019). While there is high community variation between host
-species it was observed that the majority of metagenome determined
-functions were conserved both between host species and individuals
-(Burke and Peter Steinberg \emph{et al.}, 2011; Roth-Schulze \emph{et
-al.}, 2016). This discrepancy between taxonomic and functional
-composition could be explained by the lottery hypothesis. It postulates
-that an initial random colonization step is performed from a set of
-functionally equivalent taxonomic groups resulting in taxonomically
-different epiphytic communities sharing a core set of functional genes
-(Burke and Peter Steinberg \emph{et al.}, 2011; Roth-Schulze \emph{et
-al.}, 2016). In addition, some of the variation in the observed data
-could be attributed to different techniques used in various studies,
-such as different protocols for epiphytic cell detachment and/or DNA
-isolation, as no standard protocol to study epiphytic communities was
-established (Ugarelli \emph{et al.}, 2019; Korlević \emph{et al.},
+at lower taxonomic ranks host specific microbial communities were found
+(Lachnit \emph{et al.}, 2011; Roth-Schulze \emph{et al.}, 2016).
+Recently, it was shown that even different morphological niches within
+the same alga had a higher influence on the composition of the bacterial
+community than biogeography or environmental factors (Morrissey \emph{et
+al.}, 2019). While the microbial community composition varies between
+host species, metagenomic analyses revealed that the majority of the
+microbial functions are conserved (Burke and Peter Steinberg \emph{et
+al.}, 2011; Roth-Schulze \emph{et al.}, 2016). This discrepancy between
+the microbial taxonomic and functional composition might be explained by
+the lottery hypothesis. It postulates that an initial random
+colonization step takes places from a set of functionally equivalent
+taxonomic groups resulting in taxonomically different epiphytic
+communities sharing a core set of functional genes (Burke and Peter
+Steinberg \emph{et al.}, 2011; Roth-Schulze \emph{et al.}, 2016). In
+addition, some of the variation in the reported data could be attributed
+to different techniques used in these studies, such as different
+protocols for epiphytic cell detachment and/or DNA isolation, as no
+standard protocol has been yet established to study epiphytic
+communities (Ugarelli \emph{et al.}, 2019; Korlević \emph{et al.},
 submitted).
 
-The majority of studies describing macrophyte epiphytic communities did
-not encompass seasonal changes (Crump and Koch, 2008; Lachnit \emph{et
-al.}, 2009; Burke and Thomas \emph{et al.}, 2011; Roth-Schulze \emph{et
-al.}, 2016; Ugarelli \emph{et al.}, 2019). In addition, if seasonal
-changes were taken into account low temporal frequency and/or
-methodologies that do not allow high taxonomic resolution were used
-(Tujula \emph{et al.}, 2010; Lachnit \emph{et al.}, 2011; Miranda
-\emph{et al.}, 2013; Michelou \emph{et al.}, 2013). In the present study
+The majority of studies describing macrophyte epiphytic microbial
+communities did not include possible seasonal changes (Crump and Koch,
+2008; Lachnit \emph{et al.}, 2009; Burke and Thomas \emph{et al.}, 2011;
+Roth-Schulze \emph{et al.}, 2016; Ugarelli \emph{et al.}, 2019). If
+seasonal changes were taken into account, low temporal frequency,
+applied methodologies and/or limited number of analyzed host species did
+not allow a high taxonomic resolution (Tujula \emph{et al.}, 2010;
+Lachnit \emph{et al.}, 2011; Miranda \emph{et al.}, 2013; Michelou
+\emph{et al.}, 2013; Mancuso \emph{et al.}, 2016). In the present study
 we describe the seasonal dynamics of bacterial and archaeal communities
@@
 locality of only \emph{C. cylindracea} located in the proximity of the
-meadow. In addition, for comparison, the community of the ambient
-seawater was characterized.
+seagrass meadow. For comparison, the microbial community of the ambient
+seawater was also characterized.
 
@@
 Sequencing of the 16S rRNA V4 region yielded a total of 1.8 million
-sequences after quality curation and exclusion of eukaryotic,
-chloroplast, mitochondrial and no relative sequences
-(\autoref{supp-nseq_notus}). A total of 35 samples originating from
-epiphytic archaeal and bacterial communities associated with surfaces of
-the seagrass \emph{C. nodosa} and macroalga \emph{C. cylindracea} were
-analyzed. In addition, 18 samples (one of the samples was sequenced two
-times) originating from picoplankton archaeal and bacterial communities
-in the ambient seawater were also processed for comparison. The number
-of reads per sample ranged between 8,408 and 77,463 sequences
+sequences after quality curation and exclusion of sequences without
+known relatives (no relative sequences) and eukaryotic, chloroplast and
+mitochondrial sequences (\autoref{supp-nseq_notus}). A total of 35
+samples originating from epiphytic archaeal and bacterial communities
+associated with surfaces of the seagrass \emph{C. nodosa} and the
+macroalga \emph{C. cylindracea} were analyzed. In addition, 18 samples
+(one of the samples was sequenced twice) originating from the ambient
+seawater were also processed for comparison. The number of reads per
+sample ranged between 8,408 and 77,463 sequences
 (\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
-applied the rarefaction curves did not level off that is a common
-observation in high-throughput 16S rRNA amplicon sequencing approaches
+applied the rarefaction curves did not level off as commonly observed in
+high-throughput 16S rRNA amplicon sequencing approaches
 (\autoref{supp-rarefaction}). Following quality curation and exclusion
-of sequences mentioned before reads were clustered into 28,750 different
-OTUs at a similarity level of 97 \si{\percent}. Read numbers were
-normalized to the minimum number of sequences, 8,408
-(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,017
-different OTUs that ranged from 360 to 2,108 OTUs per sample
-(\autoref{supp-calculators}). To determine seasonal changes of richness
-and diversity the Observed Number of OTUs, Chao1, ACE, Exponential
+of sequences as mentioned above reads were clustered into 28,750
+different OTUs at a similarity level of 97 \si{\percent}. Read numbers
+were normalized to the minimum number of sequences (8,408,
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,135
+different OTUs with 373 to 2,025 OTUs per sample
+(\autoref{supp-calculators}). To determine seasonal changes in richness
+and diversity the observed number of OTUs, Chao1, ACE, Exponential
 Shannon (Jost, 2006) and Inverse Simpson were calculated after
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,675.9 ±
-144.0 OTUs{]} and monospecific {[}Number of OTUs, 1,735.6 ± 175.0
-OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,052.6 ±
-204.9 OTUs) and lower values for picoplankton communities in the ambient
-seawater (Number of OTUs, 525.3 ± 146.6 OTUs)
-(\autoref{supp-calculators}). Seasonal changes did not show such large
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,694.2 ±
+133.7 OTUs{]} and monospecific {[}Number of OTUs, 1,737.9 ± 160.5
+OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,055.3 ± 209.2
+OTUs) and lowest values were obtained for the microbial community of the
+ambient seawater (Number of OTUs, 532.6 ± 141.9 OTUs)
+(\autoref{supp-calculators}). Seasonal changes did not reveal such large
 dissimilarities. \emph{C. nodosa} communities showed a slow increase
 towards the end of the study, while \emph{C. cylindracea} (mixed and
-monospecific) communities were characterized by slightly larger values
-in Spring and Summer in comparison to Autumn and Winter
+monospecific) communities were characterized by slightly higher values
+in spring and summer than in autumn and winter
 (\autoref{supp-calculators}).
@@
 Coefficient on presence-absence data and Bray-Curtis Similarity
-Coefficient were, respectively, calculated. Coefficients were determined
-after normalization through rarefaction and binning of samples from a
+Coefficient, respectively, were calculated. Coefficients were determined
+after normalization through rarefaction and binning of samples from the
 particular environment. The highest proportion of shared OTUs and
 community was found between mixed and monospecific \emph{C. cylindracea}
-(Jaccard, 0.36; Bray-Curtis, 0.78), while lower shared values were
-calculated between seawater and epiphytic communities
+(Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared values were
+calculated between ambient seawater and epiphytic communities
 (\autoref{matrix}). Shared proportion between \emph{C. nodosa} and
-\emph{C. cylindracea} were approximately in the middle between these two
-extremes. To assess seasonal changes in the proportion of shared OTUs
-and communities the Jaccard's and Bray-Curtis Similarity Coefficients
-were calculated between consecutive sampling points (\autoref{shared}).
-Both coefficients showed similar trends. Temporal proportional changes
-were more pronounced for seawater in comparison to \emph{C. nodosa} and
-especially \emph{C. cylindracea} associated communities
-(\autoref{shared}). In addition, only 0.4 -- 1.0 \si{\percent} of OTUs
-from each surface associated community were found at every time point.
-These OTUs also made a high proportion of total sequences (40.6 -- 52.6
-\si{\percent}). To further disentangle the environmental and seasonal
-community dissimilarity a Principal Coordinates Analysis (PCoA) based on
-Bray-Curtis distances and OTU abundances was applied. It showed a clear
-separation between planktonic and surface associated communities
-(\autoref{pcoa}). In addition, a separation of epiphytic bacterial and
-archaeal communities based on host species was determined. This
-separation was further supported by ANOSIM (R = 0.96, \emph{p}
-\textless{} 0.001). Seasonal changes of \emph{C. nodosa} associated
-communities indicated a separation between Spring, Summer and
-Autumn/Winter samples (ANOSIM, R = 0.53, \emph{p} \textless{} 0.01). For
-\emph{C. cylindracea} associated communities a separation between Summer
-and Autumn/Winter/Spring samples was observed that was not so strongly
-supported (ANOSIM, R = 0.31, \emph{p} \textless{} 0.05)
+\emph{C. cylindracea} were approximately in-between the values of mixed
+and monospecific \emph{C. cylindracea}. To assess seasonal changes in
+the proportion of shared OTUs and communities the Jaccard's and
+Bray-Curtis Similarity Coefficients were calculated between consecutive
+sampling points (\autoref{shared}). Both coefficients showed similar
+trends. Temporal proportional changes were more pronounced for ambient
+seawater than for \emph{C. nodosa} and especially \emph{C. cylindracea}
+associated communities (\autoref{shared}). In addition, only 0.4 -- 1.0
+\si{\percent} of OTUs from each surface associated community were
+present at all seasons. These persistent OTUs constituted a high
+proportion of total sequences (39.4 -- 51.9 \si{\percent}). To further
+disentangle the environmental and seasonal community dissimilarity a
+Principal Coordinates Analysis (PCoA) based on Bray-Curtis distances and
+OTU abundances was applied. A clear separation between ambient seawater
+and surface associated communities was found (\autoref{pcoa}). In
+addition, a separation of epiphytic bacterial and archaeal communities
+based on host species was detected. This separation was further
+supported by ANOSIM (R = 0.96, \emph{p} \textless{} 0.001). Seasonal
+changes of \emph{C. nodosa} associated communities indicated a
+separation between spring, summer and autumn/winter samples (ANOSIM, R =
+0.54, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
+communities a separation between summer and autumn/winter/spring samples
+was observed that was, however, not as strong as for \emph{C. nodosa}
+associated communities (ANOSIM, R = 0.31, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
 
-The taxonomic composition of both, macrophyte associated and seawater
-communities, was dominated by bacterial (99.1 ± 2.1 \si{\percent}) over
-archaeal sequences (0.9 ± 2.1 \si{\percent}) (\autoref{community}).
-Higher relative abundances of chloroplast related sequences were only
-observed in surface associated communities, with higher values in
-Autumn/Winter (37.2 ± 11.2 \si{\percent}) in comparison to Spring/Summer
-(20.9 ± 9.7 \si{\percent}) (\autoref{supp-chloroplast}). Generally, at
-higher taxonomic ranks (phylum-class) epiphytic and seawater microbial
-communities were composed of similar bacterial taxa. Seawater
-communities were mainly comprised of \emph{Actinobacteriota},
-\emph{Bacteroidota}, \emph{Cyanobacteria}, \emph{Alphaproteobacteria},
-\emph{Gammaproteobacteria} and \emph{Verrucomicrobiota}. Communities
-associated with \emph{C. nodosa} consisted of same groups with the
-addition of \emph{Planctomycetota} whose contribution was higher in
-summer 2018. In addition, communities from mixed and monospecific
-\emph{C. cylindracea} were similar and characterized by same groups as
-seawater and \emph{C. nodosa} communities with the addition of
-\emph{Desulfobacterota} (\autoref{community}). Larger differences
-between environments and host species could be observed at lower
-taxonomic ranks (\autoref{cyano} -- \ref{desulfo}).
-
-\emph{Cyanobacteria} related sequences were comprising, on average, 5.5
-± 4.4 \si{\percent} of total sequences (\autoref{cyano}). Higher
-proportions were found for \emph{C. nodosa} (16.4 ± 5.3 \si{\percent})
-and \emph{C. cylindracea} (mixed {[} (7.7 ± 3.9 \si{\percent}){]} and
-monospecific {[} (7.8 ± 2.4 \si{\percent}){]}) associated communities in
-autumn and for seawater communities in winter (8.8 ± 7.5 \si{\percent}).
-Large taxonomic differences between surface associated and seawater
-cyanobacterial communities were observed. Seawater communities were
-mainly comprised of \emph{Cyanobium} and \emph{Synechococcus}, while
-surface associated communities were consisted of \emph{Pleurocapsa} and
-sequences without known relatives within \emph{Cyanobacteriia}
+The taxonomic composition of both, macrophyte associated and ambient
+seawater communities was dominated by bacterial (99.1 ± 2.1
+\si{\percent}) over archaeal sequences (0.9 ± 2.1 \si{\percent})
+(\autoref{community}). Higher relative abundances of chloroplast related
+sequences were only observed in surface associated communities, with
+higher values in autumn/winter (37.2 ± 11.2 \si{\percent}) than in
+spring/summer (20.9 ± 9.7 \si{\percent}) (\autoref{supp-chloroplast}).
+Generally, at higher taxonomic ranks (phylum-class), epiphytic and
+ambient seawater microbial communities were composed of similar
+bacterial taxa. Ambient seawater communities were mainly comprised of
+\emph{Actinobacteriota}, \emph{Bacteroidota}, \emph{Cyanobacteria},
+\emph{Alphaproteobacteria}, \emph{Gammaproteobacteria} and
+\emph{Verrucomicrobiota}. Communities associated with \emph{C. nodosa}
+consisted additionally of \emph{Planctomycetota} contributing more in
+summer 2018 than in other seasons. In addition, communities from mixed
+and monospecific \emph{C. cylindracea} were similar and characterized by
+the same groups as ambient seawater and \emph{C. nodosa} communities
+with the addition of \emph{Desulfobacterota} (\autoref{community}).
+Larger differences between environments and host species were observed
+at lower taxonomic ranks (Figs. \ref{cyano} -- \ref{desulfo}).
+
+\emph{Cyanobacteria} related sequences comprised, on average, 5.5 ± 4.4
+\si{\percent} of total sequences (\autoref{cyano}). Higher proportions
+were found for \emph{C. nodosa} (16.4 ± 5.3 \si{\percent}) and \emph{C.
+cylindracea} mixed (7.7 ± 3.9 \si{\percent}) and monospecific (7.8 ± 2.4
+\si{\percent}) associated communities in autumn and for ambient seawater
+communities in winter (8.8 ± 7.5 \si{\percent}). Large taxonomic
+differences between surface associated and ambient seawater
+cyanobacterial communities were observed. Ambient seawater communities
+were mainly comprised of \emph{Cyanobium} and \emph{Synechococcus},
+while surface associated communities were comprised of
+\emph{Pleurocapsa} and sequences within the class \emph{Cyanobacteriia}
+that could not be further classified (no relative \emph{Cyanobacteriia})
 (\autoref{cyano}). In addition, seasonal changes in surface associated
-communities were observed with \emph{Pleurocapsa} and no relative
+communities were observed in \emph{Pleurocapsa} and no relative
 \emph{Cyanobacteriia} comprising larger proportions in autumn and winter
-and \emph{Acrophormium}, \emph{Phormidesmis} and no relative
-\emph{Nodosilineaceae} in spring and summer (\autoref{cyano}).
-
-Sequences classified as \emph{Bacteroidota} were comprising, on average,
-19.2 ± 5.5 \si{\percent} of all sequences (\autoref{bactero}). Similarly
-to \emph{Cyanobacteria}, large differences in the taxonomic composition
-between seawater and surface associated communities were found
-(\autoref{bactero}). The seawater community was characterized by the NS4
-and NS5 marine groups, uncultured \emph{Cryomorphaceae}, uncultured
-\emph{Flavobacteriaceae}, NS11-12 marine group, \emph{Balneola},
-uncultured \emph{Balneolaceae} and \emph{Formosa}. In contrast, in
-surface associated communities \emph{Lewinella}, \emph{Portibacter},
-\emph{Rubidimonas}, no relative \emph{Saprospiraceae}, uncultured
-\emph{Saprospiraceae}, no relative \emph{Flavobacteriaceae} and
-uncultured \emph{Rhodothermaceae} were found. Some groups showed slight
-seasonal changes such as no relative \emph{Flavobacteriaceae} that were
-more pronounced from November 2017 until June 2018. In contrast,
-uncultured \emph{Rhodothermaceae} showed higher proportions from June
-2018 until the end of the study period. Surface associated
-\emph{Bacteroidota} communities were very diverse as could be observed
-in the high proportion of taxa that grouped as other \emph{Bacteroidota}
-(\autoref{bactero}).
-
-On average, \emph{Alphaproteobacteria} were in comparison to other high
-rank taxa the largest taxonomic group, comprising 29.2 ± 12.0
-\si{\percent} of all sequences (\autoref{alpha}). In accordance to
-previous taxa, high differences between seawater and surface associated
-communities were observed. Picoplankton communities were composed mainly
-of the SAR11 clade, AEGEAN-169 marine group, SAR116 clade, no relative
-\emph{Rhodobacteraceae}, HIMB11 and OCS116 clade, while surface
-associated communities were composed in high proportion of no relative
-\emph{Rhodobacteraceae} and to a lesser degree of \emph{Pseudoahrensia},
-no relative \emph{Alphaproteobacteria}, no relative
-\emph{Hyphomonadaceae} and \emph{Amylibacter}. Representatives of no
-relative \emph{Rhodobacteraceae} were comprising on average 40.6 ± 23.2
-\si{\percent} of all alphaproteobacterial sequences from the epiphytic
-community (\autoref{alpha}). In addition, \emph{Amylibacter} was
-detected mainly in \emph{C. nodosa} from November 2017 until March 2018.
-
-Sequences related to \emph{Gammaproteobacteria} were comprising, on
-average, 18.6 ± 3.9 \si{\percent} of all sequences (\autoref{gamma}).
-Similarly to previous taxa, large taxonomic differences between seawater
-and surface associated communities were found. Seawater communities were
-mainly comprised of the OM60 (NOR5) clade, \emph{Litoricola},
-\emph{Acinetobacter} and the SAR86 clade, while epiphytic communities
-were mainly composed of no relative \emph{Gammaproteobacteria} and
-\emph{Granulosicoccus}. Beside these two groups specific to all three
-epiphytic communities, \emph{C. nodosa} was characterized by
-\emph{Arenicella}, no relative \emph{Burkholderiales} and
-\emph{Methylotenera}, while \emph{Thioploca}, no relative
-\emph{Cellvibrionaceae} and \emph{Reinekea} were more specific to both
-mixed and monospecific \emph{C. cylindracea}. In addition,
-\emph{Arenicella} was more pronounced in November and December 2017,
-while no relative \emph{Burkholderiales} and \emph{Methylotenera} were
-more characteristic for the period form March until May 2018. For the
-\emph{C. cylindracea} specific taxa no relative \emph{Cellvibrionaceae}
-and \emph{Reinekea} showed some seasonality and were characteristic for
-samples originating from June to October 2018. In addition, similarly to
-\emph{Bacteroidota}, a large proportion of the surface associated
-community was grouped as other \emph{Gammaproteobacteria} indicating
-high diversity within this group (\autoref{gamma}).
-
-In contrast to previously described high rank taxa,
-\emph{Desulfobacterota} were specific to \emph{C. cylindracea}. On
-average they comprised 11.2 ± 13.3 \si{\percent} of all sequences.
-Seawater and \emph{C. nodosa} communities consisted of only 0.1 ± 0.08
-\si{\percent} and 1.0 ± 0.7 \si{\percent} \emph{Desulfobacterota}
-sequences, respectively. In the mixed and monospecific \emph{C.
-cylindracea} communities their proportion was 25.7 ± 11.2 \si{\percent}
-and 24.0 ± 4.3 \si{\percent}, respectively (\autoref{desulfo}). The
-community consisted mainly of no relative \emph{Desulfobacteraceae},
-\emph{Desulfatitalea}, no relative \emph{Desulfobulbaceae},
-\emph{Desulfobulbus}, no relative \emph{Desulfocapsaceae},
+and \emph{Acrophormium}, \emph{Phormidesmis} and sequences without known
+relatives within the \emph{Nodosilineaceae} (no relative
+\emph{Nodosilineaceae}) in spring and summer (\autoref{cyano}).
+
+Sequences classified as \emph{Bacteroidota} comprised, on average, 19.2
+± 5.5 \si{\percent} of all sequences (\autoref{bactero}). Similarly to
+\emph{Cyanobacteria}, large differences in the taxonomic composition
+between ambient seawater and surface associated communities were found
+(\autoref{bactero}). The ambient seawater community was characterized by
+the NS4 and NS5 marine groups, uncultured \emph{Cryomorphaceae},
+uncultured \emph{Flavobacteriaceae}, NS11-12 marine group,
+\emph{Balneola}, uncultured \emph{Balneolaceae} and \emph{Formosa}. In
+contrast, in surface associated communities \emph{Lewinella},
+\emph{Portibacter}, \emph{Rubidimonas}, sequences without known
+relatives within the \emph{Saprospiraceae} (no relative
+\emph{Saprospiraceae}), uncultured \emph{Saprospiraceae}, (sequences
+without known relatives within the \emph{Flavobacteriaceae} (no relative
+\emph{Flavobacteriaceae})and uncultured \emph{Rhodothermaceae} were
+found. Some groups showed minor seasonal changes such as no relative
+\emph{Flavobacteriaceae} whose sequences were more abundant from
+November 2017 until June 2018. In contrast, uncultured
+\emph{Rhodothermaceae} showed higher proportions from June 2018 until
+the end of the study period. Surface associated \emph{Bacteroidota}
+communities were very diverse as observed in the high proportion of taxa
+clustering as other \emph{Bacteroidota} (\autoref{bactero}).
+
+On average, \emph{Alphaproteobacteria} were in comparison to the other
+high rank taxa the largest taxonomic group, comprising 29.2 ± 12.0
+\si{\percent} of all sequences (\autoref{alpha}). In accordance to the
+above described taxa, large differences between ambient seawater and
+surface associated communities were observed. Ambient seawater
+communities were composed mainly of the SAR11 clade, AEGEAN-169 marine
+group, SAR116 clade, sequences without known relatives within the
+\emph{Rhodobacteraceae} (no relative \emph{Rhodobacteraceae}), HIMB11
+and the OCS116 clade, while surface associated communities were composed
+mainly of no relative \emph{Rhodobacteraceae} and to a lesser degree of
+\emph{Pseudoahrensia}, \emph{Amylibacter} and sequences without known
+relatives within the \emph{Alphaproteobacteria} (no relative
+\emph{Alphaproteobacteria}) and \emph{Hyphomonadaceae} (no relative
+\emph{Hyphomonadaceae}). Representatives of no relative
+\emph{Rhodobacteraceae} comprised on average 40.6 ± 23.2 \si{\percent}
+of all alphaproteobacterial sequences in the epiphytic community
+(\autoref{alpha}). In addition, \emph{Amylibacter} was detected mainly
+in \emph{C. nodosa} from November 2017 until March 2018.
+
+Sequences related to \emph{Gammaproteobacteria} comprised on average
+18.6 ± 3.9 \si{\percent} of all sequences (\autoref{gamma}). Similar to
+above mentioned taxa, large taxonomic differences between ambient
+seawater and surface associated communities were found. Ambient seawater
+communities were mainly comprised of the OM60 (NOR5) clade,
+\emph{Litoricola}, \emph{Acinetobacter} and the SAR86 clade, while
+epiphytic communities were mainly composed of sequences without known
+relatives within the \emph{Gammaproteobacteria} (no relative
+\emph{Gammaproteobacteria}) and \emph{Granulosicoccus}. Beside these two
+groups specific to all three epiphytic communities, \emph{C. nodosa} was
+characterized by \emph{Arenicella}, \emph{Methylotenera} and sequences
+without known relatives within the \emph{Burkholderiales} (no relative
+\emph{Burkholderiales}), while \emph{Thioploca}, \emph{Reinekea} and
+sequences without known relatives within \emph{Cellvibrionaceae} (no
+relative \emph{Cellvibrionaceae}) were more specific to both mixed and
+monospecific \emph{C. cylindracea}. In addition, \emph{Arenicella} was
+more pronounced in November and December 2017, while no relative
+\emph{Burkholderiales} and \emph{Methylotenera} were characteristic for
+the period from March until May 2018. For the \emph{C. cylindracea}
+specific taxa no relative \emph{Cellvibrionaceae} and \emph{Reinekea}
+showed seasonality and were characteristic for samples originating from
+June to October 2018. In addition, similar to \emph{Bacteroidota}, a
+large proportion of the surface associated community was grouped as
+other \emph{Gammaproteobacteria} indicating high diversity within this
+group (\autoref{gamma}).
+
+\emph{Desulfobacterota} were specific for \emph{C. cylindracea}. In the
+mixed and monospecific \emph{C. cylindracea} communities the proportion
+of \emph{Desulfobacterota} was 25.7 ± 11.2 \si{\percent} and 24.0 ± 4.3
+\si{\percent}, respectively (\autoref{desulfo}). In contrast, in ambient
+seawater and \emph{C. nodosa} communities the contribution of
+\emph{Desulfobacterota} was only 0.1 ± 0.08 \si{\percent} and 1.0 ± 0.7
+\si{\percent}, respectively. In \emph{C. cylindracea} the community
+consisted mainly of \emph{Desulfatitalea}, \emph{Desulfobulbus},
 \emph{Desulfopila}, \emph{Desulforhopalus}, \emph{Desulfotalea},
-SEEP-SRB4 and uncultured \emph{Desulfocapsaceae} (\autoref{desulfo}).
+SEEP-SRB4, uncultured \emph{Desulfocapsaceae} and sequences without
+known relatives within the \emph{Desulfobacteraceae} (no relative
+\emph{Desulfobacteraceae}), \emph{Desulfobulbaceae} (no relative
+\emph{Desulfobulbaceae}) and \emph{Desulfocapsaceae} (no relative
+\emph{Desulfocapsaceae}) (\autoref{desulfo}).
 
@@
 microbial taxa (Egan \emph{et al.}, 2013; Tarquinio \emph{et al.},
-2019). No standard protocol has been developed to study these macophyte
-associated microbes (Ugarelli \emph{et al.}, 2019). Different procedures
-for removal of microbial cells from host surfaces were described, such
-as host tissue shaking (Nõges \emph{et al.}, 2010), scraping (Uku
-\emph{et al.}, 2007) and ultrasonication (Cai \emph{et al.}, 2014). All
-these methods showed different removal efficiencies but none was
-enabling a complete removal of attached microbial cells. In the present
-study, we applied an earlier developed removal protocol (Korlević
-\emph{et al.}, submitted), based on a previous idea of direct cellular
-lysis (Burke \emph{et al.}, 2009), to ensure an almost complete cell
-detachment. The application of a direct lysis procedure coupled with a
-high frequency sampling protocol and Illumina high resolution amplicon
-sequencing has enabled us to make a detailed description of bacterial
-and archaeal communities associated with the surfaces of two marine
-macrophytes, \emph{C. nodosa} and \emph{C. cylindracea}.
-
-In the present study, highest richness values were observed for \emph{C.
-cylindracea} (mixed and monospecific), middle for \emph{C. nodosa} and
-lowest for seawater derived communities. Higher values for seagrass
-associated communities in comparison to seawater were described earlier
-and could be attributed to a larger set of inhabitable microniches
-existing on macrophyte surfaces (Ugarelli \emph{et al.}, 2019). In
-addition, highest values observed for \emph{C. cylindracea} are partly
-due to its contact with the sediment. \emph{C. cylindracea} stolon is
-attached to the sediment surface with rhizoids, so the stolon and
+2019). No standard protocol has been developed to study these
+macophyte-associated microbes (Ugarelli \emph{et al.}, 2019). Different
+procedures for removal of microbial cells from host surfaces are
+described, such as host tissue shaking (Nõges \emph{et al.}, 2010),
+scraping (Uku \emph{et al.}, 2007), swabbing (Mancuso \emph{et al.},
+2016) and ultrasonication (Cai \emph{et al.}, 2014). All these methods
+result in different removal efficiencies but none was enabling a
+complete removal of attached microbial cells based on our experience. In
+the present study, we applied a removal protocol (Korlević \emph{et
+al.}, submitted) based on direct cellular lysis (Burke \emph{et al.},
+2009). The application of a direct lysis procedure coupled with a high
+sampling frequency and Illumina amplicon sequencing has enabled us to
+described in detail the bacterial and archaeal communities associated
+with the surfaces of two marine macrophytes, \emph{C. nodosa} and
+\emph{C. cylindracea}.
+
+In the present study, highest richness was observed for \emph{C.
+cylindracea} (mixed and monospecific) followed by \emph{C. nodosa} and
+lowest richness was found in ambient seawater microbial communities.
+Higher richness of microbial communities associated with seagrasses than
+in ambient seawater was described earlier and could be attributed to a
+larger set of inhabitable microniches existing on macrophyte surfaces
+than in the ambient seawater (Ugarelli \emph{et al.}, 2019). The highest
+richness observed for \emph{C. cylindracea} might be partly due to its
+contact with the sediment. The stolon of \emph{C. cylindracea} is
+attached to the sediment surface with rhizoids and thus, the stolon and
 rhizoids are in a direct contact with the sediment. In addition,
-seasonal richness differences observed for surface attached communities
-showed slightly higher values in spring and summer. This pattern could
-be explained by a higher macrophyte growth in these seasons (M. Najdek,
-personal communication; Zavodnik \emph{et al.}, 1998; Ruitton \emph{et
-al.}, 2005). During active periods macrophytes exhibit a more dynamic
-chemical interaction with the surface community probably causing an
-increase in the number of inhabitable microniches (Borges and
-Champenois, 2015; Rickert \emph{et al.}, 2016).
-
-Since the colonization of macrophyte surfaces is performed from a pool
-of prokaryotic cells from the ambient seawater, it was interesting to
-see to which extent these two communities differ. We observed a strong
-differentiation between the surface attached and seawater communities at
-the level of OTUs that is in agreement with most published studies
-(Burke and Thomas \emph{et al.}, 2011; Michelou \emph{et al.}, 2013;
-Roth-Schulze \emph{et al.}, 2016; Crump \emph{et al.}, 2018; Ugarelli
-\emph{et al.}, 2019). These data indicate that marine macrophytes are
-selecting, from a pool of seawater microbial taxa, the one that can
-colonize and proliferate on their surfaces (Salaün \emph{et al.}, 2012;
-Michelou \emph{et al.}, 2013). In contrast to these findings Fahimipour
-\emph{et al.} (2017) found, in a global study of \emph{Zostera marina},
-similarities between leaves and seawater samples. Discrepancies between
-our data and this study could be explained by differences in studied
-seagrass species, methodological variations or biogeographic trends as
-Fahimipour \emph{et al.} (2017) were analyzing samples from different
-locations throughout the northern hemisphere. It is possible that
-ambient seawater and leaves communities from the same location are
-differing but are still more similar to each other when compared to
-other sampling locations. Indeed, it was found that prokaryotic
-communities vary substantially between different sampling sites
-(Bengtsson \emph{et al.}, 2017). When the taxonomic composition at high
-ranks was analyzed no such strong differentiation was noticed. Phyla and
-classes such as: \emph{Actinobacteriota}, \emph{Bacteroidota},
-\emph{Cyanobacteria}, \emph{Alphaproteobacteria},
-\emph{Gammaproteobacteria} and \emph{Verrucomicrobiota} were described
-that is in agreement with previously reported data (Burke and Thomas
-\emph{et al.}, 2011; Egan \emph{et al.}, 2013; Michelou \emph{et al.},
-2013). In contrast, when low taxonomic ranks were analyzed (i.g. family
-and genus) a strong differentiation was observed. A similar
-differentiation at lower taxonomic ranks was described for other species
-of macrophytes (Egan \emph{et al.}, 2013; Michelou \emph{et al.}, 2013;
+seasonal differences in richness observed for surface attached
+communities indicated a slightly higher richness in spring and summer.
+This pattern could be explained by a higher macrophyte growth in these
+two seasons than in autumn and winter (M. Najdek, personal
+communication; Zavodnik \emph{et al.}, 1998; Ruitton \emph{et al.},
+2005). During their main growth season in spring and summer macrophytes
+exhibit a more dynamic chemical interaction with the surface community
+probably causing an increase in the number of inhabitable microniches
+(Borges and Champenois, 2015; Rickert \emph{et al.}, 2016).
+
+We observed a strong differentiation between the surface attached and
+ambient seawater communities at the level of OTUs, in agreement with
+most published studies (Burke and Thomas \emph{et al.}, 2011; Michelou
+\emph{et al.}, 2013; Roth-Schulze \emph{et al.}, 2016; Mancuso \emph{et
+al.}, 2016; Crump \emph{et al.}, 2018; Ugarelli \emph{et al.}, 2019).
+This indicates that marine macrophytes are a selecting factor from the
+pool of microbial taxa present in the ambient seawater, modifying the
+microbial community once the macrophyte associated microbial biofilm
+develops (Salaün \emph{et al.}, 2012; Michelou \emph{et al.}, 2013). In
+contrast, Fahimipour \emph{et al.} (2017) report in a global study of
+\emph{Zostera marina}, similarities between the microbial community
+developed on leaves and in the ambient seawater. The discrepancy between
+our data and the study of Fahimipour \emph{et al.} (2017) could be
+explained by different seagrass species, methodological variations or
+biogeographic trends as Fahimipour \emph{et al.} (2017) analyzed samples
+from different locations throughout the northern hemisphere. It is
+possible that the microbial communities in ambient seawater and on
+leaves from the same location are differing but are still more similar
+to each other when compared to other sampling locations. Indeed, it was
+found that prokaryotic communities vary substantially between different
+sampling sites (Bengtsson \emph{et al.}, 2017). When the taxonomic
+composition at high ranks was analyzed no such strong differentiation
+was noticed. Phyla and classes such as \emph{Actinobacteriota},
+\emph{Bacteroidota}, \emph{Cyanobacteria}, \emph{Alphaproteobacteria},
+\emph{Gammaproteobacteria} and \emph{Verrucomicrobiota} were found in
+both ambient seawater as well as macrophyte associated, in agreement
+with previous studies (Burke and Thomas \emph{et al.}, 2011; Egan
+\emph{et al.}, 2013; Michelou \emph{et al.}, 2013). In contrast, when
+low taxonomic ranks were analyzed (i.e., family and genus) a strong
+differentiation was observed (Figs. \ref{cyano} -- \ref{desulfo}). A
+similar differentiation at lower microbial taxonomic ranks between
+ambient seawater and macrophytes was described for other macrophyte
+species as well (Egan \emph{et al.}, 2013; Michelou \emph{et al.}, 2013;
 Ugarelli \emph{et al.}, 2019).
 
-Beside differences between seawater and surface associated communities,
-there were discussions if the prokaryotic epiphytic community is
-host-specific or there are generalists taxa characteristic to all or
-many macrophytes (Egan \emph{et al.}, 2013). Similarly to previously
-described differences between seawater and surface attached communities,
-at high taxonomic ranks no strong differentiation between communities
-associated with different host was observed. The only high rank phylum
+Beside differences between ambient seawater and surface associated
+microbial communities, it is unclear whether the prokaryotic epiphytic
+community is host-specific or whether there are generalist taxa
+characteristic to all or many macrophytes (Egan \emph{et al.}, 2013).
+Similar to previously described differences between microbial
+communities in the ambient seawater and on macrophytes, at high
+taxonomic ranks no major difference between the microbial communities
+associated with different hosts was observed. The only high rank phylum
 that was differing between \emph{C. nodosa} and \emph{C. cylindracea}
-was \emph{Desulfobacterota}, whose sequences were more abundant in the
+was \emph{Desulfobacterota}, with more abundant sequences in the
 \emph{C. cylindracea} associated community. As already mentioned, the
-rhizoids and part of the stolon are in contact with the sediment, so
+rhizoids and part of the stolon are in contact with the sediment. Thus
 \emph{Desulfobacterota} are probably a part of the epiphytic community
@@
 macrophytes (Burke and Thomas \emph{et al.}, 2011; Lachnit \emph{et
-al.}, 2011; Bengtsson \emph{et al.}, 2017). In contrast to high
-taxonomic ranks, a substantial differentiation between host specific
-communities was found, which supports the host-specific hypothesis.
-Similar host-specificity was observed for different species of
-macroalgae and seagrasses (Lachnit \emph{et al.}, 2011; Roth-Schulze
-\emph{et al.}, 2016; Ugarelli \emph{et al.}, 2019; Morrissey \emph{et
-al.}, 2019). Taken together, at high taxonomic ranks a core set of taxa
-could be described that is characteristic for all or many macrophytes,
-while at low taxonomic ranks a community specific to host species could
-be identified (Egan \emph{et al.}, 2013).
-
-Seasonal richness changes in the epiphytic community were substantial as
-could be observed in the proportion of OTUs that could be found at every
-sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
-were accounting for a high proportion of sequences (\(\leq\) 52.6
-\si{\percent}). A very similar proportion of persistent OTUs and their
-sequence contribution was reported in high frequency studies describing
-seasonal picoplankton changes (Gilbert \emph{et al.}, 2009, 2012). In
-comparison to the seawater community, a lower degree of seasonal shifts
-was observed for the surface associated communities. It seems,
-microniches on the surfaces of macrophytes are providing more stable
-conditions in comparison to the seawater. At the level of OTUs seasonal
-changes of \emph{C. nodosa} and \emph{C. cylindracea} associated
-communities were identified that could be linked to the growth cycle of
-the seagrass and macroalgae (M. Najdek, personal communication).
-\emph{C. nodosa} was characterized by a Spring community during maximum
-seagrass proliferation, a Summer community during a biomass maximum and
-a Autumn/Winter community during a biomass senescence. In contrast,
-\emph{C. cylindracea} started to proliferate in late Spring and was
-characterized only by a Summer community during maximal biomass increase
-and by a Autumn/Winter/Spring community when the biomass was at the peak
-and the settlement started to subsequently decay. Similar seasonal
-changes in the epiphytic community was also described for other
-macroalgae (Tujula \emph{et al.}, 2010; Lachnit \emph{et al.}, 2011).
-Higher temporal stability of \emph{C. cylindracea} surface communities
-in comparison to \emph{C. nodosa} were also observed in the higher
-proportion of shared communities between two consecutive sampling
-points.
-
-Analysis of seasonal chloroplast sequence abundances showed higher
-values in Autumn/Winter in comparison to Spring/Summer. This pattern is
-not surprising as seagrasses are known to harbor more algal epiphytes
-during Autumn/Winter (Reyes and Sansón, 2001). Furthermore, we used an
-adapted DNA isolation protocol that is known to partially coextract DNA
-from planktonic eukaryotes (Korlević \emph{et al.}, 2015). Strong
-seasonal fluctuations of high rank epiphytic taxa were not observed,
-with the exception of \emph{Cyanobateria}. Cyanobacterial sequences were
-more pronounced in November and December in comparison to Spring and
-Summer. Interestingly, in these high proportion months the majority of
-cyanobacterial sequences were classified as \emph{Pleurocapsa}, a group
-known to colonized different living and nonliving surfaces (Burns
-\emph{et al.}, 2004; Longford \emph{et al.}, 2007; Mobberley \emph{et
-al.}, 2012; Reisser \emph{et al.}, 2014). It is possible that during
-periods of low metabolic activity there is a reduced interaction and
-selection of the epiphytic community by the seagrass, causing leaves to
-become a suitable surface for nonspecific colonizers (Zavodnik \emph{et
-al.}, 1998). \emph{Pleurocapsa} was replaced in Spring and Summer by
-\emph{Acrophormium}, \emph{Phormidesmis} and no relative
-\emph{Nodosilineaceae}. A study of coastal microbial mats found also
-higher proportion of \emph{Nodosilineaceae} sequences in Summer, while
-\emph{Phormidesmis} sequences were at their peak in Autumn (Cardoso
-\emph{et al.}, 2019). Other high rank taxa did not show strong
-successional patterns. In every analyzed group, with the exception of
-\emph{Desulfobacterota}, taxa present throughout the year in similar
-proportions and season specific taxa could be identified. Within
-\emph{Bacteroidota} different groups withing the family
-\emph{Saprospiraceae} (i.g. \emph{Lewinella}, \emph{Portibacter} and
-\emph{Rubidimonas}) were detected through the year. Members of this
-family are often found in association with macrophytes and it is
-suggested that they are involved in the hydrolysis and utilization of
-complex carbon sources (Burke and Thomas \emph{et al.}, 2011; McIlroy
-and Nielsen, 2014; Crump \emph{et al.}, 2018). On the other hand,
-families \emph{Flavobacteriaceae} and \emph{Rhodothermaceae} showed
-seasonal patterns, with \emph{Flavobacteriaceae} being more pronounced
-from November to June and \emph{Rhodothermaceae} from June to October.
-Within \emph{Alphaproteobacteria} the family \emph{Rhodobacteraceae} was
-comprising the majority of sequences throughout the year. This
+al.}, 2011; Mancuso \emph{et al.}, 2016; Bengtsson \emph{et al.}, 2017).
+In contrast to high taxonomic ranks, a substantial differentiation
+between host specific communities was found supporting the notion that
+macrophyte associated microbial communities might be host-specific.
+Host-specificity was also observed for other species of macroalgae and
+seagrasses (Lachnit \emph{et al.}, 2011; Roth-Schulze \emph{et al.},
+2016; Ugarelli \emph{et al.}, 2019; Morrissey \emph{et al.}, 2019).
+Taken together, at high taxonomic ranks a core set of taxa could be
+described that is characteristic for all or many macrophytes, while at
+low taxonomic ranks a community specific to host species was identified
+(Figs. \ref{pcoa} and \ref{community}) (Egan \emph{et al.}, 2013).
+
+Seasonal changes in richness in the epiphytic community were substantial
+as indicated by the proportion of OTUs (\(\leq\) 1.0 \si{\percent})
+present at every sampling date. These persistent OTUs, however, were
+accounting for a high proportion of sequences (\(\leq\) 51.9
+\si{\percent}) (\autoref{shared}). A very similar proportion of
+persistent OTUs was reported in high-frequency sampling studies
+describing seasonal changes in picoplankton (Gilbert \emph{et al.},
+2009, 2012). In comparison to the seawater community, a lower degree of
+seasonal shifts was observed for the macrophyte surface associated
+communities. It appears that microniches at the surfaces of macrophytes
+are providing more stable conditions than the ambient seawater. At the
+level of OTUs seasonal changes of \emph{C. nodosa} and \emph{C.
+cylindracea} associated communities were identified that could be linked
+to the growth cycle of the seagrass and macroalgae (M. Najdek, personal
+communication). \emph{C. nodosa} was characterized by a spring community
+during maximum seagrass proliferation, a summer community during the
+highest standing stock of \emph{C.nodosa} and an autumn/winter community
+during the decay of seagrass biomass. In contrast, \emph{C. cylindracea}
+started to proliferate in late spring and was characterized only by a
+summer community during high growth rates and by an autumn/winter/spring
+community when the biomass was at the peak and decaying thereafter.
+Similar seasonal changes in the epiphytic community were also described
+for other macroalgae (Tujula \emph{et al.}, 2010; Lachnit \emph{et al.},
+2011). Higher seasonal stability of \emph{C. cylindracea} surface
+communities than in \emph{C. nodosa} was also observed in the higher
+proportion of shared communities between two consecutive sampling dates
+in \emph{C. cylindracea}.
+
+Chloroplast sequence abundances were higher in autumn/winter than in
+spring/summer. This pattern is not surprising as seagrasses harbor more
+algal epiphytes during autumn/winter than in spring/summer (Reyes and
+Sansón, 2001). Furthermore, we used an adapted DNA isolation protocol
+that is known to partially co-extract DNA from planktonic eukaryotes
+(Korlević \emph{et al.}, 2015). Strong seasonal fluctuations of high
+rank epiphytic taxa were not observed, with the exception of
+\emph{Cyanobacteria}. Cyanobacterial sequences were more pronounced in
+November and December than in spring and summer. In the months of high
+cyanobacterial sequence abundances the majority of sequences from this
+group were classified as \emph{Pleurocapsa}, a group known to colonized
+different living and non-living surfaces (Burns \emph{et al.}, 2004;
+Longford \emph{et al.}, 2007; Mobberley \emph{et al.}, 2012; Reisser
+\emph{et al.}, 2014). It is possible that during periods of low
+metabolic activity there is a reduced interaction and selection of the
+epiphytic community by the seagrass, causing leaves to become a suitable
+surface for nonspecific colonizers (Zavodnik \emph{et al.}, 1998).
+\emph{Pleurocapsa} was replaced in spring and summer by
+\emph{Acrophormium}, \emph{Phormidesmis} and sequences without known
+relatives within the \emph{Nodosilineaceae}. A study of coastal
+microbial mats found also a higher proportion of \emph{Nodosilineaceae}
+sequences in summer, while \emph{Phormidesmis} sequences were at their
+peak in autumn (Cardoso \emph{et al.}, 2019). Other high rank taxa did
+not exhibit strong successional patterns. In every analyzed group, with
+the exception of \emph{Desulfobacterota}, taxa present throughout the
+year in similar proportions and season specific taxa could be identified
+(Figs. \ref{bactero} and \ref{desulfo}). Within \emph{Bacteroidota}
+different groups within the family \emph{Saprospiraceae} (e.g.
+\emph{Lewinella}, \emph{Portibacter} and \emph{Rubidimonas}) were
+detected across all seasons. Members of this family are often found in
+association with macrophytes and it is suggested that they are involved
+in the hydrolysis and utilization of complex carbon sources (Burke and
+Thomas \emph{et al.}, 2011; McIlroy and Nielsen, 2014; Crump \emph{et
+al.}, 2018). In contrast, the families \emph{Flavobacteriaceae} and
+\emph{Rhodothermaceae} showed seasonal patterns, with
+\emph{Flavobacteriaceae} being more pronounced from November to June and
+\emph{Rhodothermaceae} from June to October (\autoref{bactero}). Within
+\emph{Alphaproteobacteria} the family \emph{Rhodobacteraceae} comprised
+the majority of sequences throughout the year (\autoref{alpha}). This
 metabolically versatile family is often associated with macrophyte
-surfaces and usually is one of the most abundant groups (Burke and
-Thomas \emph{et al.}, 2011; Michelou \emph{et al.}, 2013; Pujalte
-\emph{et al.}, 2014). In addition, \emph{Hyphomonadaceae} were found in
-all samples. Interestingly, some of the species within this group
-contain stalks on their cells which can be used to attach to the
+surfaces and usually is one of the most abundant groups ({\textbf{???}};
+Burke and Thomas \emph{et al.}, 2011; Michelou \emph{et al.}, 2013;
+Pujalte \emph{et al.}, 2014). In addition, \emph{Hyphomonadaceae} were
+found in all samples. Interestingly, some of the species within this
+group contain stalks on their cells, which can be used to attach to the
 macrophyte surface (Weidner \emph{et al.}, 2000; Abraham and Rohde,
@@
 representatives were the most pronounced group present throughout the
-year. In addition, \emph{Granulosicoccus} was also found in almost all
-samples. \emph{Gammaproteobacteria} are often a major constituent of
-macrophyte epiphytic communities (Burke and Thomas \emph{et al.}, 2011;
-Michelou \emph{et al.}, 2013; Crump \emph{et al.}, 2018). Beside these
-two groups other less pronounced taxa showed seasonal and host-specific
-patterns. For example, \emph{C. cylindracea} was characterized by
-\emph{Thioploca}, a known sulfur sediment bacteria and
-\emph{Cellvibrionaceae}, a family whose cultured members are known
+year (\autoref{gamma}). In addition, \emph{Granulosicoccus} was also
+found in almost all samples. \emph{Gammaproteobacteria} are often a
+major constituent of macrophyte epiphytic communities (Burke and Thomas
+\emph{et al.}, 2011; Michelou \emph{et al.}, 2013; Crump \emph{et al.},
+2018). Beside these two groups, other less abundant, taxa showed
+seasonal and host-specific patterns. For example, \emph{C. cylindracea}
+harbored \emph{Thioploca}, a known sulfur sediment bacteria and
+\emph{Cellvibrionaceae}, a family with cultured members known as
 polysaccharide degraders (Jørgensen and Gallardo, 1999; Xie \emph{et
@@
 \emph{C. cylindracea} and no group within this phylum showed seasonal
-patterns. The presence of this phylum only on \emph{C. cylindracea} is
-to be expected as part of the epiphytic community is directly in contact
-with the sediment. The \emph{Desulfobacterota} community was dominated
-by \emph{Desulfatitalea} and no relative \emph{Desulfocapsaceae}, known
-sulfate sediment groups (Kuever, 2014; Higashioka \emph{et al.}, 2015).
-
-In temperate zones marine macrophytes are exhibiting growth cycles, so
+patterns (\autoref{desulfo}). The presence of this phylum only on
+\emph{C. cylindracea} is to be expected as part of the epiphytic
+community is in direct contact with the sediment. The
+\emph{Desulfobacterota} community was dominated by \emph{Desulfatitalea}
+and no relative \emph{Desulfocapsaceae}, known sulfate sediment groups
+(Kuever, 2014; Higashioka \emph{et al.}, 2015).
+
+In temperate zones, marine macrophytes are exhibiting growth cycles, so
 it is not surprising that the associated epiphytic microbial community
-is undergoing partial seasonal changes. In the present study, we could,
-in every analyzed high rank taxa, identify phylogenetic groups that were
-present throughout the year and that were comprising most of the
-sequences and lower proportion taxa showing seasonal patterns connected
-to the macrophyte growth cycle. Studies focusing on functional
-comparisons between communities associated with different hosts showed
-that the majority of functions could be found in every community,
-indicating functional redundancy (Roth-Schulze \emph{et al.}, 2016).
-This difference between taxonomic and functional discrepancy was
-explained by the lottery hypothesis that hypothesizes an initial random
-colonization step performed from a set of functionally equivalent
+is undergoing partial seasonal changes. In the present study, we could
+identify in every analyzed high rank taxa phylogenetic groups present
+throughout the year, comprising most of the sequences and a lower
+proportion of taxa showing seasonal patterns connected to the macrophyte
+growth cycle (Figs. \ref{community} and \ref{desulfo}). Studies focusing
+on functional comparisons between communities associated with different
+hosts showed that the majority of functions could be found in every
+community, indicating functional redundancy (Roth-Schulze \emph{et al.},
+2016). This difference between phylogenetic varaibility and functional
+stability was explained by the lottery hypothesis assuming an initial
+random colonization step performed by a set of functionally equivalent
 taxonomic groups (Burke and Peter Steinberg \emph{et al.}, 2011;
@@
 collected by diving and transported to the laboratory in containers
-placed on ice and filled with site seawater. Upon arrival to the
-laboratory, \emph{C. nodosa} leaves were cut into sections of 1 -- 2
-\si{\cm}, while \emph{C. cylindracea} thalli were cut into 5 -- 8
-\si{\cm} long sections. Leaves and thalli were washed three times with
-sterile artificial seawater (ASW) to remove loosely attached microbial
-cells. Ambient seawater was collected in 10 \si{\l} containers by diving
-and transported to the laboratory where the whole container volume was
-filtered through a 20 \si{\um} net. The filtrate was further
+placed on ice and filled with seawater collected at the sampling site.
+Upon arrival to the laboratory, \emph{C. nodosa} leaves were cut into
+sections of 1 -- 2 \si{\cm}, while \emph{C. cylindracea} thalli were cut
+into 5 -- 8 \si{\cm} long sections. Leaves and thalli were washed three
+times with sterile artificial seawater (ASW) to remove loosely attached
+microbial cells. Ambient seawater was collected in 10 \si{\l} containers
+by diving and transported to the laboratory where the whole container
+volume was filtered through a 20 \si{\um} net. The filtrate was further
 sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate
@@
 incubated in a lysis buffer and treated with lysozyme and proteinase K.
-Following the incubations, the mixture containing lyzed epiphytic cells
-is separated from leaves and thalli and extracted using a
-phenol-chloroform procedure. Finally, the extracted DNA is precipitated
-using isopropanol. DNA from seawater picoplankton was isolated from 0.2
+Following the incubations, the mixture containing lysed epiphytic cells
+was separated from the leaves and thalli and extracted using
+phenol-chloroform. Finally, the extracted DNA was precipitated using
+isopropanol. DNA from seawater picoplankton was extracted from 0.2
 \si{\um} polycarbonate filters according to Massana \emph{et al.} (1997)
-with a slight modification. Following the phenol-chloroform extraction
-steps 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added.
-DNA was precipitated by adding 1 volume of chilled isopropanol,
-incubating the mixtures overnight at \num{-20} \si{\degreeCelsius} and
-centrifuging at 20,000 × g and 4 \si{\degreeCelsius} for 21
-\si{\minute}. The pellet was washed twice with 500 \si{\ul} of chilled
-70 \si{\percent} ethanol and centrifuged after each washing step at
-20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets
-were resuspended in 50 -- 100 \si{\ul} of deionized water.
+with a slight modification. Following the phenol-chloroform extraction,
+1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added. DNA was
+precipitated by adding 1 volume of chilled isopropanol, incubating the
+mixtures overnight at \num{-20} \si{\degreeCelsius} and centrifuging at
+20,000 × \emph{g} and 4 \si{\degreeCelsius} for 21 \si{\minute}. The
+pellet was washed twice with 500 \si{\ul} of chilled 70 \si{\percent}
+ethanol and centrifuged after each washing step at 20,000 × \emph{g} and
+4 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets were
+re-suspended in 50 -- 100 \si{\ul} of deionized water.
 
@@
 
-Illumina MiSeq sequencing of the V4 16S rRNA region was performed as
-described previously (Korlević \emph{et al.}, submitted). The V4 region
-of the 16S rRNA gene was amplified using a two-step PCR procedure. In
-the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R
-(5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project
+Illumina MiSeq sequencing of the V4 region of the 16S rRNA gene was
+performed as described previously (Korlević \emph{et al.}, submitted).
+The V4 region of the 16S rRNA gene was amplified using a two-step PCR
+procedure. In the first PCR, the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and
+806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome
+Project
 (\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
@@
 tagged sequence. Purified PCR products were sent for Illumina MiSeq
-sequencing at IMGM Laboratories, Martinsried, Germany. Before sequencing
-at IMGM, the second PCR amplification of the two-step PCR procedure was
-performed using primers targeting the tagged region incorporated in the
-first PCR. In addition, these primers contained adapter and
-sample-specific index sequences. Beside samples, a positive and negative
-control for each sequencing batch was sequenced. Negative control was
-comprised of PCR reactions without DNA template, while for a positive
-control a mock community composed of evenly mixed DNA material
-originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
-used. Sequences obtained in this study have been deposited in the
-European Nucleotide Archive (ENA) at EMBL-EBI under accession number
+sequencing at IMGM Laboratories, Martinsried, Germany. Prior to
+sequencing at IMGM, the second PCR amplification of the two-step PCR
+procedure was performed using primers targeting the tagged region
+incorporated in the first PCR. In addition, these primers contained
+adapter and sample-specific index sequences. Beside samples, a positive
+and negative control for each sequencing batch was sequenced. The
+negative control comprised PCR reactions without DNA template, while for
+a positive control a mock community composed of evenly mixed DNA
+material originating from 20 bacterial strains (ATCC MSA-1002, ATCC,
+USA) was used. Sequences obtained in this study have been deposited in
+the European Nucleotide Archive (ENA) at EMBL-EBI under accession number
 PRJEB37267.
@@
 \url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
-and recommendations given from the Riffomonas project to enhance data
+and recommendations provided by the Riffomonas project to enhance data
 reproducibility (\url{http://www.riffomonas.org/}). For alignment and
@@
 Yilmaz \emph{et al.}, 2014). Pipeline data processing and visualization
-was done using R (version 3.6.0) (R Core Team, 2019), packages vegan
-(version 2.5-6) (Oksanen \emph{et al.}, 2019), and tidyverse (version
-1.3.0) (Wickham \emph{et al.}, 2019) and multiple other packages (Xie,
-2014, 2015, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018; Y. Xie,
-2019b, 2019a; Allaire \emph{et al.}, 2019; Zhu, 2019). The detailed
-analysis procedure including the R Markdown file for this paper are
-available as a GitHub repository
+was done using R (version 3.6.0) (R Core Team, 2019) combined with
+packages vegan (version 2.5-6) (Oksanen \emph{et al.}, 2019), tidyverse
+(version 1.3.0) (Wickham \emph{et al.}, 2019) and multiple other
+packages (Xie, 2014, 2015, 2020; Neuwirth, 2014; Xie \emph{et al.},
+2018; Y. Xie, 2019b, 2019a; Allaire \emph{et al.}, 2019; Zhu, 2019). The
+detailed analysis procedure including the R Markdown file are available
+in the GitHub repository
 (\url{https://github.com/mkorlevic/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020}).
@@
 sampling and the University Computing Center of the University of Zagreb
-for access to the cluster Isabella.
+for access to the computer cluster Isabella.
 
@@
 
+\leavevmode\hypertarget{ref-Mancuso2016}{}%
+Mancuso, F.P., D'Hondt, S., Willems, A., Airoldi, L., and De Clerck, O.
+(2016) Diversity and temporal dynamics of the epiphytic bacterial
+communities associated with the canopy-forming seaweed \emph{Cystoseira
+compressa} (Esper) Gerloff and Nizamuddin. \emph{Front Microbiol}
+\textbf{7}: 476.
+
 \leavevmode\hypertarget{ref-Margulis1991}{}%
@@
 
-\caption{Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between assemblages associated with the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed Settlement] and \textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and comunities in the ambient seawater.\label{matrix}}\label{fig:unnamed-chunk-1}
+\caption{Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between assemblages associated with the surfaces of the macrophytes \textit{C.
+nodosa} (Mixed Settlement) and \textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and communities in the ambient seawater.\label{matrix}}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 
-\caption{Proportion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling points and from the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed Settlement] and \textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.\label{shared}}\label{fig:unnamed-chunk-2}
+\caption{Proportion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling dates and from the surfaces of the macrophytes \textit{C.
+nodosa} (Mixed Settlement) and \textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.\label{shared}}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 
-\caption{Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed Settlement] and \textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
+\caption{Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of the macrophytes \textit{C.
+nodosa} (Mixed Settlement) and \textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
 Samples from the same environment or same season are labeled in different colors.
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) bacterial and archaeal sequences on the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed Settlement] and \textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
-NR -- No Relative\label{community}}\label{fig:unnamed-chunk-4}
+\caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) bacterial and archaeal sequences on the surfaces of the macrophytes \textit{C.
+nodosa} (Mixed Settlement) and \textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
+NR -- No Relative (sequences without known relatives within the corresponding group)\label{community}}\label{fig:unnamed-chunk-4}
 \end{figure}
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) cyanobacterial sequences on the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed Settlement] and \textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
+\caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) cyanobacterial sequences on the surfaces of the macrophytes \textit{C.
+nodosa} (Mixed Settlement) and \textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
 The proportion of cyanobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
-NR -- No Relative\label{cyano}}\label{fig:unnamed-chunk-5}
+NR -- No Relative (sequences without known relatives within the corresponding group)\label{cyano}}\label{fig:unnamed-chunk-5}
 \end{figure}
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) sequences within the \textit{Bacteroidota} on the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed Settlement] and \textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
+\caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) sequences within the \textit{Bacteroidota} on the surfaces of the macrophytes \textit{C.
+nodosa} (Mixed Settlement) and \textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
 The proportion of sequences classified as \textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar.
-NR -- No Relative\label{bactero}}\label{fig:unnamed-chunk-6}
+NR -- No Relative (sequences without known relatives within the corresponding group)\label{bactero}}\label{fig:unnamed-chunk-6}
 \end{figure}
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) alphaproteobacterial sequences on the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed Settlement] and \textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
+\caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) alphaproteobacterial sequences on the surfaces of the macrophytes \textit{C.
+nodosa} (Mixed Settlement) and \textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
 The proportion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
-NR -- No Relative\label{alpha}}\label{fig:unnamed-chunk-7}
+NR -- No Relative (sequences without known relatives within the corresponding group)\label{alpha}}\label{fig:unnamed-chunk-7}
 \end{figure}
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) gammaproteobacterial sequences on the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed Settlement] and \textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
+\caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) gammaproteobacterial sequences on the surfaces of the macrophytes \textit{C.
+nodosa} (Mixed Settlement) and \textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
 The proportion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
-NR -- No Relative\label{gamma}}\label{fig:unnamed-chunk-8}
+NR -- No Relative (sequences without known relatives within the corresponding group)\label{gamma}}\label{fig:unnamed-chunk-8}
 \end{figure}
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) sequences within the \textit{Desulfobacterota} on the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed Settlement] and \textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.
+\caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) sequences within the \textit{Desulfobacterota} on the surfaces of the macrophytes \textit{C.
+nodosa} (Mixed Settlement) and \textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
 The proportion of sequences classified as \textit{Desulfobacterota} in the total bacterial and archaeal community is given above the corresponding bar.
-NR -- No Relative\label{desulfo}}\label{fig:unnamed-chunk-9}
+NR -- No Relative (sequences without known relatives within the corresponding group)\label{desulfo}}\label{fig:unnamed-chunk-9}
 \end{figure}

### 2020-05-28 submission/supplementary.Rmd
Review optimization.

@@
   mutate(station=str_replace(station, "^FCa$", "\\\\textit{Caulerpa cylindracea} (Monospecific)")) %>%
-  mutate(season=str_replace(season, "^Summer$", "summer")) %>%
-  mutate(season=str_replace(season, "^Autumn$", "autumn")) %>%
-  mutate(season=str_replace(season, "^Winter$", "winter")) %>%
-  mutate(season=str_replace(season, "^Spring$", "spring")) %>%
   mutate(ID=str_replace(ID, "^23_1$", "23a")) %>%
@@
 of Sequences", "No.
-of OTUs"), caption="Sample ID, Community Type, Sampling Date and Season, No.
-of Sequences and No.
+of OTUs"), caption="Sample ID, community type, sampling date and season, no.
+of sequences and no.
 of OTUs of each sample.

### 2020-05-28 submission/supplementary.tex
Process optimzation
Optimized front page.
Review optimization.

--- 2020-03-19 submission/supplementary.tex
@@
 \renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\figurename}{Fig.}
 \renewcommand{\thetable}{S\arabic{table}}
@@
 \usepackage{float}
+
+\renewcommand{\figureautorefname}{Fig.}
 \usepackage{booktabs}
@@
 
-\vspace{60mm}
+\vspace{20mm}
 
@@
 Herndl\textsuperscript{2,3}, Mirjana Najdek\textsuperscript{1}
-
-\vspace{40mm}
-
-\textsuperscript{\(*\)}To whom correspondence should be addressed:
-\href{mailto:marino.korlevic@irb.hr}{\nolinkurl{marino.korlevic@irb.hr}}
 
@@
 
-3. Department of Marine Microbiology and Biogeochemistry, Royal
+3. NIOZ, Department of Marine Microbiology and Biogeochemistry, Royal
 Netherlands Institute for Sea Research, Utrecht University, The
 Netherlands
+
+\textsuperscript{\(*\)}To whom correspondence should be addressed:
+
+Marino Korlević
+
+G. Paliaga 5, 52210 Rovinj, Croatia
+
+Tel.: +385 52 804 768
+
+Fax: +385 52 804 780
+
+e-mail:
+\href{mailto:marino.korlevic@irb.hr}{\nolinkurl{marino.korlevic@irb.hr}}
+
+Running title: Seasonal dynamics of epiphytic communities
 
@@
 
-\caption{Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed] and \textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\label{rarefaction}}\label{fig:unnamed-chunk-1}
+\caption{Rarefaction curves of bacterial and archaeal communities from the surfaces of the macrophytes \textit{C.
+nodosa} (Mixed Settlement) and \textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.\label{rarefaction}}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 
-\caption{Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed] and \textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\label{calculators}}\label{fig:unnamed-chunk-2}
+\caption{Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of the macrophytes \textit{C.
+nodosa} (Mixed Settlement) and \textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.\label{calculators}}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 
-\caption{Relative contribution of chloroplast sequences on the surfaces of macrophytes (\textit{C.
-nodosa} [Mixed] and \textit{C.
-cylindracea} [Mixed and Monospecific]) and in the ambient seawater.\label{chloroplast}}\label{fig:unnamed-chunk-3}
+\caption{Relative contribution of chloroplast sequences on the surfaces of the macrophytes \textit{C.
+nodosa} (Mixed Settlement) and \textit{C.
+cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.\label{chloroplast}}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 \begin{longtable}[t]{>{\centering\arraybackslash}p{6em}ccccc}
-\caption{\label{tab:nseq_notus}Sample ID, Community Type, Sampling Date and Season, No.
-of Sequences and No.
+\caption{\label{tab:nseq_notus}Sample ID, community type, sampling date and season, no.
+of sequences and no.
 of OTUs of each sample.
-No.
-of Sequences and OTUs was calculated after the exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences.\label{nseq_notus}}\\
+The number of sequences and OTUs was calculated after exclusion of sequences without known relatives (no relative sequences) and eukaryotic, chloroplast and mitochondrial sequences.\label{nseq_notus}}\\
 \toprule
@@
 \endfirsthead
-\caption[]{Sample ID, Community Type, Sampling Date and Season, No.
-of Sequences and No.
+\caption[]{Sample ID, community type, sampling date and season, no.
+of sequences and no.
 of OTUs of each sample.
-No.
-of Sequences and OTUs was calculated after the exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences.\label{nseq_notus} \textit{(continued)}}\\
+The number of sequences and OTUs was calculated after exclusion of sequences without known relatives (no relative sequences) and eukaryotic, chloroplast and mitochondrial sequences.\label{nseq_notus} \textit{(continued)}}\\
 \toprule
@@
 \endlastfoot
-3 & Seawater & 13 July 2017 & Summer & 25,996 & 821\\
-5 & Seawater & 27 July 2017 & Summer & 58,954 & 993\\
-7 & Seawater & 10 August 2017 & Summer & 32,626 & 853\\
-9 & Seawater & 24 August 2017 & Summer & 60,935 & 1,064\\
-11 & Seawater & 19 September 2017 & Summer & 46,107 & 1,179\\
-13 & Seawater & 12 October 2017 & Autumn & 30,911 & 971\\
-15 & Seawater & 23 November 2017 & Autumn & 27,585 & 1,197\\
-17 & Seawater & 14 December 2017 & Autumn & 44,595 & 2,275\\
-19 & Seawater & 13 February 2018 & Winter & 44,207 & 1,306\\
-21 & Seawater & 27 March 2018 & Winter & 46,313 & 1,561\\
-23a & Seawater & 24 April 2018 & Spring & 30,973 & 999\\
-23b & Seawater & 24 April 2018 & Spring & 38,563 & 1,195\\
-25 & Seawater & 22 May 2018 & Spring & 53,877 & 1,787\\
-27 & Seawater & 19 June 2018 & Spring & 77,465 & 1,290\\
-29 & Seawater & 10 July 2018 & Summer & 50,793 & 1,445\\
-31 & Seawater & 9 August 2018 & Summer & 39,041 & 1,123\\
-33 & Seawater & 4 September 2018 & Summer & 36,206 & 1,206\\
-35 & Seawater & 5 October 2018 & Autumn & 49,585 & 1,019\\
-37 & \textit{Cymodocea nodosa} (Mixed) & 23 November 2017 & Autumn & 31,237 & 1,555\\
-41 & \textit{Cymodocea nodosa} (Mixed) & 4 December 2017 & Autumn & 24,234 & 1,174\\
-45 & \textit{Cymodocea nodosa} (Mixed) & 14 December 2017 & Autumn & 32,683 & 1,061\\
-49 & \textit{Cymodocea nodosa} (Mixed) & 13 February 2018 & Winter & 9,088 & 1,216\\
-52 & \textit{Cymodocea nodosa} (Mixed) & 27 March 2018 & Winter & 17,005 & 1,216\\
-55 & \textit{Cymodocea nodosa} (Mixed) & 24 April 2018 & Spring & 42,645 & 2,053\\
-58 & \textit{Cymodocea nodosa} (Mixed) & 22 May 2018 & Spring & 21,337 & 1,290\\
-61 & \textit{Cymodocea nodosa} (Mixed) & 19 June 2018 & Spring & 31,726 & 2,098\\
-64 & \textit{Cymodocea nodosa} (Mixed) & 10 July 2018 & Summer & 35,729 & 1,793\\
-67 & \textit{Cymodocea nodosa} (Mixed) & 9 August 2018 & Summer & 26,356 & 2,116\\
-70 & \textit{Cymodocea nodosa} (Mixed) & 4 September 2018 & Summer & 23,275 & 1,724\\
-73 & \textit{Cymodocea nodosa} (Mixed) & 5 October 2018 & Autumn & 29,907 & 2,211\\
-38 & \textit{Caulerpa cylindracea} (Mixed) & 23 November 2017 & Autumn & 36,309 & 2,896\\
-42 & \textit{Caulerpa cylindracea} (Mixed) & 4 December 2017 & Autumn & 28,383 & 3,249\\
-46 & \textit{Caulerpa cylindracea} (Mixed) & 14 December 2017 & Autumn & 34,718 & 3,067\\
-53 & \textit{Caulerpa cylindracea} (Mixed) & 27 March 2018 & Winter & 28,689 & 2,481\\
-56 & \textit{Caulerpa cylindracea} (Mixed) & 24 April 2018 & Spring & 34,784 & 3,057\\
-59 & \textit{Caulerpa cylindracea} (Mixed) & 22 May 2018 & Spring & 23,392 & 2,712\\
-62 & \textit{Caulerpa cylindracea} (Mixed) & 19 June 2018 & Spring & 36,488 & 3,311\\
-65 & \textit{Caulerpa cylindracea} (Mixed) & 10 July 2018 & Summer & 18,483 & 2,188\\
-68 & \textit{Caulerpa cylindracea} (Mixed) & 9 August 2018 & Summer & 31,943 & 3,097\\
-71 & \textit{Caulerpa cylindracea} (Mixed) & 4 September 2018 & Summer & 29,273 & 3,152\\
-74 & \textit{Caulerpa cylindracea} (Mixed) & 5 October 2018 & Autumn & 11,707 & 1,711\\
-39 & \textit{Caulerpa cylindracea} (Monospecific) & 23 November 2017 & Autumn & 22,057 & 2,428\\
-43 & \textit{Caulerpa cylindracea} (Monospecific) & 4 December 2017 & Autumn & 13,661 & 1,892\\
-47 & \textit{Caulerpa cylindracea} (Monospecific) & 14 December 2017 & Autumn & 8,409 & 1,451\\
-51 & \textit{Caulerpa cylindracea} (Monospecific) & 13 February 2018 & Winter & 11,673 & 1,899\\
-54 & \textit{Caulerpa cylindracea} (Monospecific) & 27 March 2018 & Winter & 39,471 & 3,142\\
-57 & \textit{Caulerpa cylindracea} (Monospecific) & 24 April 2018 & Spring & 20,304 & 2,836\\
-60 & \textit{Caulerpa cylindracea} (Monospecific) & 22 May 2018 & Spring & 33,035 & 3,314\\
-63 & \textit{Caulerpa cylindracea} (Monospecific) & 19 June 2018 & Spring & 41,841 & 3,976\\
-66 & \textit{Caulerpa cylindracea} (Monospecific) & 10 July 2018 & Summer & 27,038 & 2,669\\
-69 & \textit{Caulerpa cylindracea} (Monospecific) & 9 August 2018 & Summer & 26,734 & 3,123\\
-72 & \textit{Caulerpa cylindracea} (Monospecific) & 4 September 2018 & Summer & 31,874 & 3,248\\
-75 & \textit{Caulerpa cylindracea} (Monospecific) & 5 October 2018 & Autumn & 33,073 & 3,078\\*
+3 & Seawater & 13 July 2017 & Summer & 26,006 & 825\\
+5 & Seawater & 27 July 2017 & Summer & 58,951 & 995\\
+7 & Seawater & 10 August 2017 & Summer & 32,623 & 856\\
+9 & Seawater & 24 August 2017 & Summer & 60,938 & 1,059\\
+11 & Seawater & 19 September 2017 & Summer & 46,106 & 1,163\\
+13 & Seawater & 12 October 2017 & Autumn & 30,905 & 973\\
+15 & Seawater & 23 November 2017 & Autumn & 27,582 & 1,180\\
+17 & Seawater & 14 December 2017 & Autumn & 44,591 & 2,267\\
+19 & Seawater & 13 February 2018 & Winter & 44,193 & 1,312\\
+21 & Seawater & 27 March 2018 & Winter & 46,352 & 1,583\\
+23a & Seawater & 24 April 2018 & Spring & 30,976 & 995\\
+23b & Seawater & 24 April 2018 & Spring & 38,565 & 1,196\\
+25 & Seawater & 22 May 2018 & Spring & 53,872 & 1,791\\
+27 & Seawater & 19 June 2018 & Spring & 77,463 & 1,291\\
+29 & Seawater & 10 July 2018 & Summer & 50,786 & 1,443\\
+31 & Seawater & 9 August 2018 & Summer & 39,039 & 1,120\\
+33 & Seawater & 4 September 2018 & Summer & 36,201 & 1,205\\
+35 & Seawater & 5 October 2018 & Autumn & 49,585 & 1,014\\
+37 & \textit{Cymodocea nodosa} (Mixed) & 23 November 2017 & Autumn & 31,241 & 1,560\\
+41 & \textit{Cymodocea nodosa} (Mixed) & 4 December 2017 & Autumn & 24,241 & 1,178\\
+45 & \textit{Cymodocea nodosa} (Mixed) & 14 December 2017 & Autumn & 32,686 & 1,058\\
+49 & \textit{Cymodocea nodosa} (Mixed) & 13 February 2018 & Winter & 9,091 & 1,213\\
+52 & \textit{Cymodocea nodosa} (Mixed) & 27 March 2018 & Winter & 17,000 & 1,215\\
+55 & \textit{Cymodocea nodosa} (Mixed) & 24 April 2018 & Spring & 42,653 & 2,063\\
+58 & \textit{Cymodocea nodosa} (Mixed) & 22 May 2018 & Spring & 21,337 & 1,278\\
+61 & \textit{Cymodocea nodosa} (Mixed) & 19 June 2018 & Spring & 31,726 & 2,097\\
+64 & \textit{Cymodocea nodosa} (Mixed) & 10 July 2018 & Summer & 35,746 & 1,793\\
+67 & \textit{Cymodocea nodosa} (Mixed) & 9 August 2018 & Summer & 26,360 & 2,113\\
+70 & \textit{Cymodocea nodosa} (Mixed) & 4 September 2018 & Summer & 23,276 & 1,713\\
+73 & \textit{Cymodocea nodosa} (Mixed) & 5 October 2018 & Autumn & 29,910 & 2,216\\
+38 & \textit{Caulerpa cylindracea} (Mixed) & 23 November 2017 & Autumn & 36,318 & 2,895\\
+42 & \textit{Caulerpa cylindracea} (Mixed) & 4 December 2017 & Autumn & 28,388 & 3,251\\
+46 & \textit{Caulerpa cylindracea} (Mixed) & 14 December 2017 & Autumn & 34,721 & 3,055\\
+53 & \textit{Caulerpa cylindracea} (Mixed) & 27 March 2018 & Winter & 28,688 & 2,478\\
+56 & \textit{Caulerpa cylindracea} (Mixed) & 24 April 2018 & Spring & 34,765 & 3,060\\
+59 & \textit{Caulerpa cylindracea} (Mixed) & 22 May 2018 & Spring & 23,403 & 2,723\\
+62 & \textit{Caulerpa cylindracea} (Mixed) & 19 June 2018 & Spring & 36,487 & 3,310\\
+65 & \textit{Caulerpa cylindracea} (Mixed) & 10 July 2018 & Summer & 18,486 & 2,192\\
+68 & \textit{Caulerpa cylindracea} (Mixed) & 9 August 2018 & Summer & 31,953 & 3,099\\
+71 & \textit{Caulerpa cylindracea} (Mixed) & 4 September 2018 & Summer & 29,280 & 3,152\\
+74 & \textit{Caulerpa cylindracea} (Mixed) & 5 October 2018 & Autumn & 11,698 & 1,702\\
+39 & \textit{Caulerpa cylindracea} (Monospecific) & 23 November 2017 & Autumn & 22,086 & 2,435\\
+43 & \textit{Caulerpa cylindracea} (Monospecific) & 4 December 2017 & Autumn & 13,661 & 1,890\\
+47 & \textit{Caulerpa cylindracea} (Monospecific) & 14 December 2017 & Autumn & 8,408 & 1,454\\
+51 & \textit{Caulerpa cylindracea} (Monospecific) & 13 February 2018 & Winter & 11,673 & 1,902\\
+54 & \textit{Caulerpa cylindracea} (Monospecific) & 27 March 2018 & Winter & 39,469 & 3,131\\
+57 & \textit{Caulerpa cylindracea} (Monospecific) & 24 April 2018 & Spring & 20,299 & 2,832\\
+60 & \textit{Caulerpa cylindracea} (Monospecific) & 22 May 2018 & Spring & 33,042 & 3,305\\
+63 & \textit{Caulerpa cylindracea} (Monospecific) & 19 June 2018 & Spring & 41,852 & 3,964\\
+66 & \textit{Caulerpa cylindracea} (Monospecific) & 10 July 2018 & Summer & 27,036 & 2,673\\
+69 & \textit{Caulerpa cylindracea} (Monospecific) & 9 August 2018 & Summer & 26,736 & 3,114\\
+72 & \textit{Caulerpa cylindracea} (Monospecific) & 4 September 2018 & Summer & 31,872 & 3,246\\
+75 & \textit{Caulerpa cylindracea} (Monospecific) & 5 October 2018 & Autumn & 33,086 & 3,076\\*
 \end{longtable}

### 2020-05-29 submission/manuscript.Rmd
Review optimization.

@@
 Within *Alphaproteobacteria* the family *Rhodobacteraceae* comprised the majority of sequences throughout the year (\autoref{alpha}).
-This metabolically versatile family is often associated with macrophyte surfaces and usually is one of the most abundant groups [@Burke2011a; @Michelou2013; @Pujalte2014; @Manucos2016].
+This metabolically versatile family is often associated with macrophyte surfaces and usually is one of the most abundant groups [@Burke2011a; @Michelou2013; @Pujalte2014; @Mancuso2016].
 In addition, *Hyphomonadaceae* were found in all samples.

### 2020-05-29 submission/manuscript.tex
Review optimization.

@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,135
-different OTUs with 373 to 2,025 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,036
+different OTUs with 362 to 2,029 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,694.2 ±
-133.7 OTUs{]} and monospecific {[}Number of OTUs, 1,737.9 ± 160.5
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,055.3 ± 209.2
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,684.5 ±
+131.3 OTUs{]} and monospecific {[}Number of OTUs, 1,727.6 ± 161.5
+OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,060.2 ± 212.3
 OTUs) and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 532.6 ± 141.9 OTUs)
+ambient seawater (Number of OTUs, 524.5 ± 137.4 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (39.4 -- 51.9 \si{\percent}). To further
+proportion of total sequences (39.1 -- 52.4 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 separation between spring, summer and autumn/winter samples (ANOSIM, R =
-0.54, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
+0.53, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
 communities a separation between summer and autumn/winter/spring samples
@@
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\leq\) 51.9
+accounting for a high proportion of sequences (\(\leq\) 52.4
 \si{\percent}) (\autoref{shared}). A very similar proportion of
@@
 metabolically versatile family is often associated with macrophyte
-surfaces and usually is one of the most abundant groups ({\textbf{???}};
-Burke and Thomas \emph{et al.}, 2011; Michelou \emph{et al.}, 2013;
-Pujalte \emph{et al.}, 2014). In addition, \emph{Hyphomonadaceae} were
-found in all samples. Interestingly, some of the species within this
-group contain stalks on their cells, which can be used to attach to the
-macrophyte surface (Weidner \emph{et al.}, 2000; Abraham and Rohde,
-2014). Within the \emph{Gammaproteobacteria}, sequences without known
-representatives were the most pronounced group present throughout the
-year (\autoref{gamma}). In addition, \emph{Granulosicoccus} was also
-found in almost all samples. \emph{Gammaproteobacteria} are often a
-major constituent of macrophyte epiphytic communities (Burke and Thomas
-\emph{et al.}, 2011; Michelou \emph{et al.}, 2013; Crump \emph{et al.},
-2018). Beside these two groups, other less abundant, taxa showed
-seasonal and host-specific patterns. For example, \emph{C. cylindracea}
-harbored \emph{Thioploca}, a known sulfur sediment bacteria and
-\emph{Cellvibrionaceae}, a family with cultured members known as
-polysaccharide degraders (Jørgensen and Gallardo, 1999; Xie \emph{et
-al.}, 2017). \emph{Desulfobacterota} were found only associated with
-\emph{C. cylindracea} and no group within this phylum showed seasonal
-patterns (\autoref{desulfo}). The presence of this phylum only on
-\emph{C. cylindracea} is to be expected as part of the epiphytic
-community is in direct contact with the sediment. The
+surfaces and usually is one of the most abundant groups (Burke and
+Thomas \emph{et al.}, 2011; Michelou \emph{et al.}, 2013; Pujalte
+\emph{et al.}, 2014; Mancuso \emph{et al.}, 2016). In addition,
+\emph{Hyphomonadaceae} were found in all samples. Interestingly, some of
+the species within this group contain stalks on their cells, which can
+be used to attach to the macrophyte surface (Weidner \emph{et al.},
+2000; Abraham and Rohde, 2014). Within the \emph{Gammaproteobacteria},
+sequences without known representatives were the most pronounced group
+present throughout the year (\autoref{gamma}). In addition,
+\emph{Granulosicoccus} was also found in almost all samples.
+\emph{Gammaproteobacteria} are often a major constituent of macrophyte
+epiphytic communities (Burke and Thomas \emph{et al.}, 2011; Michelou
+\emph{et al.}, 2013; Crump \emph{et al.}, 2018). Beside these two
+groups, other less abundant, taxa showed seasonal and host-specific
+patterns. For example, \emph{C. cylindracea} harbored \emph{Thioploca},
+a known sulfur sediment bacteria and \emph{Cellvibrionaceae}, a family
+with cultured members known as polysaccharide degraders (Jørgensen and
+Gallardo, 1999; Xie \emph{et al.}, 2017). \emph{Desulfobacterota} were
+found only associated with \emph{C. cylindracea} and no group within
+this phylum showed seasonal patterns (\autoref{desulfo}). The presence
+of this phylum only on \emph{C. cylindracea} is to be expected as part
+of the epiphytic community is in direct contact with the sediment. The
 \emph{Desulfobacterota} community was dominated by \emph{Desulfatitalea}

### 2020-05-29 submission/manuscript.Rmd
Line break optimization.

@@
 The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
-In the first PCR, the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used [@Caporaso2012; @Apprill2015; @Parada2016].
+In the first PCR, the 515F\linebreak(5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used [@Caporaso2012; @Apprill2015; @Parada2016].
 These primers contained on their 5' end a tagged sequence.

### 2020-05-29 submission/manuscript.tex
Line break optimization.

@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,036
-different OTUs with 362 to 2,029 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,041
+different OTUs with 359 to 1,995 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,684.5 ±
-131.3 OTUs{]} and monospecific {[}Number of OTUs, 1,727.6 ± 161.5
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,060.2 ± 212.3
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,698.7 ±
+142.1 OTUs{]} and monospecific {[}Number of OTUs, 1,731.1 ± 151.9
+OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,053.7 ± 209.6
 OTUs) and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 524.5 ± 137.4 OTUs)
+ambient seawater (Number of OTUs, 530.6 ± 144.6 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 community was found between mixed and monospecific \emph{C. cylindracea}
-(Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared values were
+(Jaccard, 0.35; Bray-Curtis, 0.78), while lower shared values were
 calculated between ambient seawater and epiphytic communities
@@
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (39.1 -- 52.4 \si{\percent}). To further
+proportion of total sequences (40.6 -- 53.1 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 was observed that was, however, not as strong as for \emph{C. nodosa}
-associated communities (ANOSIM, R = 0.31, \emph{p} \textless{} 0.01)
+associated communities (ANOSIM, R = 0.33, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\leq\) 52.4
+accounting for a high proportion of sequences (\(\leq\) 53.1
 \si{\percent}) (\autoref{shared}). A very similar proportion of
@@
 The V4 region of the 16S rRNA gene was amplified using a two-step PCR
-procedure. In the first PCR, the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and
-806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome
-Project
+procedure. In the first PCR, the
+515F\linebreak(5'-GTGYCAGCMGCCGCGGTAA-3') and 806R
+(5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project
 (\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})

### 2020-05-29 submission/manuscript.tex
reference optimization.

--- 2020-05-29 submission/manuscript.tex
@@
 
+\usepackage{microtype}
+\usepackage{biblatex}
+\appto{\bibsetup}{\raggedright}
+
 \begin{document}
@@
 \newpage
+\linenumbers
 \sisetup{mode=text}
@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,036
-different OTUs with 362 to 2,029 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,048
+different OTUs with 374 to 2,035 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,684.5 ±
-131.3 OTUs{]} and monospecific {[}Number of OTUs, 1,727.6 ± 161.5
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,060.2 ± 212.3
-OTUs) and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 524.5 ± 137.4 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,684.7 ±
+130.7 OTUs{]} and monospecific {[}Number of OTUs, 1,733.4 ± 152.0
+OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,050 ± 210.1 OTUs)
+and lowest values were obtained for the microbial community of the
+ambient seawater (Number of OTUs, 528.4 ± 133.2 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 seawater than for \emph{C. nodosa} and especially \emph{C. cylindracea}
-associated communities (\autoref{shared}). In addition, only 0.4 -- 1.0
+associated communities (\autoref{shared}). In addition, only 0.5 -- 1.0
 \si{\percent} of OTUs from each surface associated community were
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (39.1 -- 52.4 \si{\percent}). To further
+proportion of total sequences (42.4 -- 50.7 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 separation between spring, summer and autumn/winter samples (ANOSIM, R =
-0.53, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
+0.54, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
 communities a separation between summer and autumn/winter/spring samples
 was observed that was, however, not as strong as for \emph{C. nodosa}
-associated communities (ANOSIM, R = 0.31, \emph{p} \textless{} 0.01)
+associated communities (ANOSIM, R = 0.30, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\leq\) 52.4
+accounting for a high proportion of sequences (\(\leq\) 50.7
 \si{\percent}) (\autoref{shared}). A very similar proportion of

### 2020-05-30 submission/manuscript.tex
\emergencystretch setup.

@@
 \usepackage{lineno}
+\renewcommand{\linenumberfont}{\normalfont\tiny}
 
@@
 \usepackage{microtype}
-\usepackage{biblatex}
-\appto{\bibsetup}{\raggedright}
 
@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,048
-different OTUs with 374 to 2,035 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,087
+different OTUs with 361 to 2,028 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,684.7 ±
-130.7 OTUs{]} and monospecific {[}Number of OTUs, 1,733.4 ± 152.0
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,050 ± 210.1 OTUs)
-and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 528.4 ± 133.2 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,681.8 ±
+124.2 OTUs{]} and monospecific {[}Number of OTUs, 1,746 ± 161.4 OTUs{]})
+than for \emph{C. nodosa} (Number of OTUs, 1,052.8 ± 209.5 OTUs) and
+lowest values were obtained for the microbial community of the ambient
+seawater (Number of OTUs, 527.8 ± 140.4 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 seawater than for \emph{C. nodosa} and especially \emph{C. cylindracea}
-associated communities (\autoref{shared}). In addition, only 0.5 -- 1.0
+associated communities (\autoref{shared}). In addition, only 0.4 -- 1.0
 \si{\percent} of OTUs from each surface associated community were
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (42.4 -- 50.7 \si{\percent}). To further
+proportion of total sequences (39.5 -- 52.4 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 was observed that was, however, not as strong as for \emph{C. nodosa}
-associated communities (ANOSIM, R = 0.30, \emph{p} \textless{} 0.01)
+associated communities (ANOSIM, R = 0.31, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\leq\) 50.7
+accounting for a high proportion of sequences (\(\leq\) 52.4
 \si{\percent}) (\autoref{shared}). A very similar proportion of
@@
 
+\interlinepenalty=10000 \emergencystretch=3.5em
+
 \hypertarget{refs}{}

### 2020-05-30 submission/manuscript.tex
\setlength{\emergencystretch} setup.

@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,087
-different OTUs with 361 to 2,028 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,016
+different OTUs with 368 to 2,042 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,681.8 ±
-124.2 OTUs{]} and monospecific {[}Number of OTUs, 1,746 ± 161.4 OTUs{]})
-than for \emph{C. nodosa} (Number of OTUs, 1,052.8 ± 209.5 OTUs) and
-lowest values were obtained for the microbial community of the ambient
-seawater (Number of OTUs, 527.8 ± 140.4 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,691.2 ±
+131.0 OTUs{]} and monospecific {[}Number of OTUs, 1,730.6 ± 169.4
+OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,055.5 ± 209.1
+OTUs) and lowest values were obtained for the microbial community of the
+ambient seawater (Number of OTUs, 525.2 ± 144.9 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 community was found between mixed and monospecific \emph{C. cylindracea}
-(Jaccard, 0.35; Bray-Curtis, 0.78), while lower shared values were
+(Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared values were
 calculated between ambient seawater and epiphytic communities
@@
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (39.5 -- 52.4 \si{\percent}). To further
+proportion of total sequences (37.2 -- 51.2 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 was observed that was, however, not as strong as for \emph{C. nodosa}
-associated communities (ANOSIM, R = 0.31, \emph{p} \textless{} 0.01)
+associated communities (ANOSIM, R = 0.32, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\leq\) 52.4
+accounting for a high proportion of sequences (\(\leq\) 51.2
 \si{\percent}) (\autoref{shared}). A very similar proportion of
@@
 
-\interlinepenalty=10000 \emergencystretch=3.5em
+\interlinepenalty=10000 \setlength{\emergencystretch}{3.5em}
 

### 2020-06-01 submission/manuscript.Rmd
reference optimization.
\emergencystretch setup.
\setlength{\emergencystretch} setup.
Capitalizaion changed.

--- 2020-05-29 submission/manuscript.Rmd
@@
 ## References
+\interlinepenalty=10000
+\setlength{\emergencystretch}{3.5em}
 <div id="refs"></div>
@@
 ## Figures
-```{r out.width="70%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between assemblages associated with the surfaces of the macrophytes \\textit{C.
-nodosa} (Mixed Settlement) and \\textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and communities in the ambient seawater.\\label{matrix}", fig.pos="H"}
+```{r out.width="70%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal OTUs (Jaccard's similarity coefficient) and shared bacterial and archaeal communities (Bray-Curtis similarity coefficient) between assemblages associated with the surfaces of the macrophytes \\textit{C.
+nodosa} (mixed settlement) and \\textit{C.
+cylindracea} (mixed and monospecific settlement) and communities in the ambient seawater.\\label{matrix}", fig.pos="H"}
 include_graphics("../results/figures/matrix.jpg")
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling dates and from the surfaces of the macrophytes \\textit{C.
-nodosa} (Mixed Settlement) and \\textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.\\label{shared}", fig.pos="H"}
+```{r out.width="85%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal communities (Bray-Curtis similarity coefficient) and shared bacterial and archaeal OTUs (Jaccard's similarity coefficient) between consecutive sampling dates and from the surfaces of the macrophytes \\textit{C.
+nodosa} (mixed settlement) and \\textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\\label{shared}", fig.pos="H"}
 include_graphics("../results/figures/seasonal_shared.jpg")
@@
 ```{r out.width="55%", fig.align="center", fig.cap="Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of the macrophytes \\textit{C.
-nodosa} (Mixed Settlement) and \\textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
+nodosa} (mixed settlement) and \\textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
 Samples from the same environment or same season are labeled in different colors.
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) bacterial and archaeal sequences on the surfaces of the macrophytes \\textit{C.
-nodosa} (Mixed Settlement) and \\textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
+nodosa} (mixed settlement) and \\textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
 NR -- No Relative (sequences without known relatives within the corresponding group)\\label{community}", fig.pos="H"}
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) cyanobacterial sequences on the surfaces of the macrophytes \\textit{C.
-nodosa} (Mixed Settlement) and \\textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
+nodosa} (mixed settlement) and \\textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
 The proportion of cyanobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) sequences within the \\textit{Bacteroidota} on the surfaces of the macrophytes \\textit{C.
-nodosa} (Mixed Settlement) and \\textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
+nodosa} (mixed settlement) and \\textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
 The proportion of sequences classified as \\textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar.
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) alphaproteobacterial sequences on the surfaces of the macrophytes \\textit{C.
-nodosa} (Mixed Settlement) and \\textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
+nodosa} (mixed settlement) and \\textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
 The proportion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) gammaproteobacterial sequences on the surfaces of the macrophytes \\textit{C.
-nodosa} (Mixed Settlement) and \\textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
+nodosa} (mixed settlement) and \\textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
 The proportion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) sequences within the \\textit{Desulfobacterota} on the surfaces of the macrophytes \\textit{C.
-nodosa} (Mixed Settlement) and \\textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
+nodosa} (mixed settlement) and \\textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
 The proportion of sequences classified as \\textit{Desulfobacterota} in the total bacterial and archaeal community is given above the corresponding bar.

### 2020-06-01 submission/manuscript.tex
Capitalizaion optimization.

@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,016
-different OTUs with 368 to 2,042 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,157
+different OTUs with 371 to 2,071 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,691.2 ±
-131.0 OTUs{]} and monospecific {[}Number of OTUs, 1,730.6 ± 169.4
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,055.5 ± 209.1
-OTUs) and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 525.2 ± 144.9 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,685 ± 127.6
+OTUs{]} and monospecific {[}Number of OTUs, 1,735.2 ± 166.6 OTUs{]})
+than for \emph{C. nodosa} (Number of OTUs, 1,055.2 ± 212.0 OTUs) and
+lowest values were obtained for the microbial community of the ambient
+seawater (Number of OTUs, 533.4 ± 142.1 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (37.2 -- 51.2 \si{\percent}). To further
+proportion of total sequences (40.7 -- 51.7 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 was observed that was, however, not as strong as for \emph{C. nodosa}
-associated communities (ANOSIM, R = 0.32, \emph{p} \textless{} 0.01)
+associated communities (ANOSIM, R = 0.29, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\leq\) 51.2
+accounting for a high proportion of sequences (\(\leq\) 51.7
 \si{\percent}) (\autoref{shared}). A very similar proportion of
@@
 
-\caption{Proportion of shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) and shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) between assemblages associated with the surfaces of the macrophytes \textit{C.
-nodosa} (Mixed Settlement) and \textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and communities in the ambient seawater.\label{matrix}}\label{fig:unnamed-chunk-1}
+\caption{Proportion of shared bacterial and archaeal OTUs (Jaccard's similarity coefficient) and shared bacterial and archaeal communities (Bray-Curtis similarity coefficient) between assemblages associated with the surfaces of the macrophytes \textit{C.
+nodosa} (mixed settlement) and \textit{C.
+cylindracea} (mixed and monospecific settlement) and communities in the ambient seawater.\label{matrix}}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 
-\caption{Proportion of shared bacterial and archaeal communities (Bray-Curtis Similarity Coefficient) and shared bacterial and archaeal OTUs (Jaccard's Similarity Coefficient) between consecutive sampling dates and from the surfaces of the macrophytes \textit{C.
-nodosa} (Mixed Settlement) and \textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.\label{shared}}\label{fig:unnamed-chunk-2}
+\caption{Proportion of shared bacterial and archaeal communities (Bray-Curtis similarity coefficient) and shared bacterial and archaeal OTUs (Jaccard's similarity coefficient) between consecutive sampling dates and from the surfaces of the macrophytes \textit{C.
+nodosa} (mixed settlement) and \textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\label{shared}}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 \caption{Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of the macrophytes \textit{C.
-nodosa} (Mixed Settlement) and \textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
+nodosa} (mixed settlement) and \textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
 Samples from the same environment or same season are labeled in different colors.
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) bacterial and archaeal sequences on the surfaces of the macrophytes \textit{C.
-nodosa} (Mixed Settlement) and \textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
+nodosa} (mixed settlement) and \textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
 NR -- No Relative (sequences without known relatives within the corresponding group)\label{community}}\label{fig:unnamed-chunk-4}
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) cyanobacterial sequences on the surfaces of the macrophytes \textit{C.
-nodosa} (Mixed Settlement) and \textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
+nodosa} (mixed settlement) and \textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
 The proportion of cyanobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) sequences within the \textit{Bacteroidota} on the surfaces of the macrophytes \textit{C.
-nodosa} (Mixed Settlement) and \textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
+nodosa} (mixed settlement) and \textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
 The proportion of sequences classified as \textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar.
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) alphaproteobacterial sequences on the surfaces of the macrophytes \textit{C.
-nodosa} (Mixed Settlement) and \textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
+nodosa} (mixed settlement) and \textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
 The proportion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) gammaproteobacterial sequences on the surfaces of the macrophytes \textit{C.
-nodosa} (Mixed Settlement) and \textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
+nodosa} (mixed settlement) and \textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
 The proportion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar.
@@
 \caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) sequences within the \textit{Desulfobacterota} on the surfaces of the macrophytes \textit{C.
-nodosa} (Mixed Settlement) and \textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.
+nodosa} (mixed settlement) and \textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
 The proportion of sequences classified as \textit{Desulfobacterota} in the total bacterial and archaeal community is given above the corresponding bar.

### 2020-06-01 submission/manuscript.tex
Capitalization update.

@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,157
-different OTUs with 371 to 2,071 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,119
+different OTUs with 373 to 2,033 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,685 ± 127.6
-OTUs{]} and monospecific {[}Number of OTUs, 1,735.2 ± 166.6 OTUs{]})
-than for \emph{C. nodosa} (Number of OTUs, 1,055.2 ± 212.0 OTUs) and
-lowest values were obtained for the microbial community of the ambient
-seawater (Number of OTUs, 533.4 ± 142.1 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,682.3 ±
+132.1 OTUs{]} and monospecific {[}Number of OTUs, 1,741.2 ± 163.6
+OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,050.8 ± 202.7
+OTUs) and lowest values were obtained for the microbial community of the
+ambient seawater (Number of OTUs, 532.9 ± 141.3 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (40.7 -- 51.7 \si{\percent}). To further
+proportion of total sequences (39.1 -- 52.8 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 separation between spring, summer and autumn/winter samples (ANOSIM, R =
-0.54, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
+0.55, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
 communities a separation between summer and autumn/winter/spring samples
 was observed that was, however, not as strong as for \emph{C. nodosa}
-associated communities (ANOSIM, R = 0.29, \emph{p} \textless{} 0.01)
+associated communities (ANOSIM, R = 0.31, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\leq\) 51.7
+accounting for a high proportion of sequences (\(\leq\) 52.8
 \si{\percent}) (\autoref{shared}). A very similar proportion of

### 2020-06-01 submission/supplementary.Rmd
reference optimization.
Capitalization update.

@@
 
+\linenumbers
 \sisetup{mode=text}
@@
 
-## Supplementary Figures
+## Supplementary figures
 
 ```{r out.width="85%", fig.align="center", fig.cap="Rarefaction curves of bacterial and archaeal communities from the surfaces of the macrophytes \\textit{C.
-nodosa} (Mixed Settlement) and \\textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.\\label{rarefaction}", fig.pos="H"}
+nodosa} (mixed settlement) and \\textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\\label{rarefaction}", fig.pos="H"}
 include_graphics("../results/figures/rarefaction.jpg")
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of the macrophytes \\textit{C.
-nodosa} (Mixed Settlement) and \\textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.\\label{calculators}", fig.pos="H"}
+```{r out.width="85%", fig.align="center", fig.cap="Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon diversity index and Inverse Simpson index of bacterial and archaeal communities from the surfaces of the macrophytes \\textit{C.
+nodosa} (mixed settlement) and \\textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\\label{calculators}", fig.pos="H"}
 include_graphics("../results/figures/calculators.jpg")
@@
 ```{r out.width="85%", fig.align="center", fig.cap="Relative contribution of chloroplast sequences on the surfaces of the macrophytes \\textit{C.
-nodosa} (Mixed Settlement) and \\textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.\\label{chloroplast}", fig.pos="H"}
+nodosa} (mixed settlement) and \\textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\\label{chloroplast}", fig.pos="H"}
 include_graphics("../results/figures/chloroplast_bar_plot.jpg")
@@
 
-## Supplementary Table
+## Supplementary table
 

### 2020-06-01 submission/supplementary.tex
reference optimization.
\emergencystretch setup.
Capitalization update.

@@
 
+\usepackage{txfonts}
+
 \renewcommand{\figureautorefname}{Fig.}
+
+\usepackage{microtype}
 \usepackage{booktabs}
@@
 \hypertarget{supplementary-figures}{%
-\subsection{Supplementary Figures}\label{supplementary-figures}}
+\subsection{Supplementary figures}\label{supplementary-figures}}
 
@@
 \caption{Rarefaction curves of bacterial and archaeal communities from the surfaces of the macrophytes \textit{C.
-nodosa} (Mixed Settlement) and \textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.\label{rarefaction}}\label{fig:unnamed-chunk-1}
+nodosa} (mixed settlement) and \textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\label{rarefaction}}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 
-\caption{Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of the macrophytes \textit{C.
-nodosa} (Mixed Settlement) and \textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.\label{calculators}}\label{fig:unnamed-chunk-2}
+\caption{Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon diversity index and Inverse Simpson index of bacterial and archaeal communities from the surfaces of the macrophytes \textit{C.
+nodosa} (mixed settlement) and \textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\label{calculators}}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 \caption{Relative contribution of chloroplast sequences on the surfaces of the macrophytes \textit{C.
-nodosa} (Mixed Settlement) and \textit{C.
-cylindracea} (Mixed and Monospecific Settlement) and in the ambient seawater.\label{chloroplast}}\label{fig:unnamed-chunk-3}
+nodosa} (mixed settlement) and \textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\label{chloroplast}}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 \hypertarget{supplementary-table}{%
-\subsection{Supplementary Table}\label{supplementary-table}}
+\subsection{Supplementary table}\label{supplementary-table}}
 

### 2020-06-03 submission/manuscript.tex
Rendering optimization.

@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,119
-different OTUs with 373 to 2,033 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,043
+different OTUs with 352 to 2,002 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,682.3 ±
-132.1 OTUs{]} and monospecific {[}Number of OTUs, 1,741.2 ± 163.6
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,050.8 ± 202.7
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,680.8 ±
+130.0 OTUs{]} and monospecific {[}Number of OTUs, 1,731.2 ± 156.7
+OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,055.4 ± 212.2
 OTUs) and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 532.9 ± 141.3 OTUs)
+ambient seawater (Number of OTUs, 527.6 ± 142.7 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (39.1 -- 52.8 \si{\percent}). To further
+proportion of total sequences (38.7 -- 52.2 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 separation between spring, summer and autumn/winter samples (ANOSIM, R =
-0.55, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
+0.53, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
 communities a separation between summer and autumn/winter/spring samples
@@
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\leq\) 52.8
+accounting for a high proportion of sequences (\(\leq\) 52.2
 \si{\percent}) (\autoref{shared}). A very similar proportion of
@@
 \leavevmode\hypertarget{ref-RCoreTeam2019}{}%
-R Core Team (2019) A language and environment for statistical computing,
-Vienna, Austria: R Foundation for Statistical Computing.
+R Core Team (2019) R: A language and environment for statistical
+computing, Vienna, Austria: R Foundation for Statistical Computing.
 
@@
 \leavevmode\hypertarget{ref-Xie2020}{}%
-Xie, Y. (2020) TinyTex: Helper functions to install and maintain 'TeX
+Xie, Y. (2020) Tinytex: Helper functions to install and maintain 'TeX
 Live', and compile 'LaTeX' documents.
@@
 \leavevmode\hypertarget{ref-Zhu2019}{}%
-Zhu, H. (2019) KableExtra: Construct complex table with 'kable' and pipe
+Zhu, H. (2019) kableExtra: Construct complex table with 'kable' and pipe
 syntax.

### 2020-06-03 submission/manuscript.tex
References optimization.

@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,043
-different OTUs with 352 to 2,002 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,145
+different OTUs with 365 to 2,036 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,680.8 ±
-130.0 OTUs{]} and monospecific {[}Number of OTUs, 1,731.2 ± 156.7
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,055.4 ± 212.2
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,680.3 ±
+125.3 OTUs{]} and monospecific {[}Number of OTUs, 1,737.9 ± 164.5
+OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,059.6 ± 211.0
 OTUs) and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 527.6 ± 142.7 OTUs)
+ambient seawater (Number of OTUs, 523.4 ± 136.0 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (38.7 -- 52.2 \si{\percent}). To further
+proportion of total sequences (40.4 -- 53.9 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\leq\) 52.2
+accounting for a high proportion of sequences (\(\leq\) 53.9
 \si{\percent}) (\autoref{shared}). A very similar proportion of
@@
 Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
-A., et al. (2019) Rmarkdown: Dynamic documents for R.
+A., et al. (2019) rmarkdown: Dynamic documents for R.
 
@@
 Oksanen, J., Blanchet, F.G., Friendly, M., Kindt, R., Legendre, P.,
-McGlinn, D., et al. (2019) Vegan: Community ecology package.
+McGlinn, D., et al. (2019) vegan: Community ecology package.
 
@@
 \leavevmode\hypertarget{ref-Xie2014}{}%
-Xie, Y. (2014) Knitr: A comprehensive tool for reproducible research in
+Xie, Y. (2014) knitr: A comprehensive tool for reproducible research in
 R. In \emph{Implementing Reproducible Computational Research}. Stodden,
@@
 \leavevmode\hypertarget{ref-Xie2019}{}%
-Xie, Y. (2019a) Knitr: A general-purpose package for dynamic report
+Xie, Y. (2019a) knitr: A general-purpose package for dynamic report
 generation in R.
@@
 \leavevmode\hypertarget{ref-Xie2020}{}%
-Xie, Y. (2020) Tinytex: Helper functions to install and maintain 'TeX
+Xie, Y. (2020) tinytex: Helper functions to install and maintain 'TeX
 Live', and compile 'LaTeX' documents.

### 2020-06-08 submission/manuscript.Rmd
Changed repository address, addresses issue_001

@@
 Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019] combined with  packages  vegan (version 2.5-6) [@Oksanen2019], tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014].
-The detailed analysis procedure including the R Markdown file are available in the GitHub repository (https://github.com/mkorlevic/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020).
+The detailed analysis procedure including the R Markdown file are available in the GitHub repository (https://github.com/MicrobesRovinj/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020).
 Based on the ATCC MSA-1002 mock community included in the analysis an average sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].

### 2020-06-08 submission/manuscript.tex

@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,145
-different OTUs with 365 to 2,036 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,093
+different OTUs with 369 to 2,055 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,680.3 ±
-125.3 OTUs{]} and monospecific {[}Number of OTUs, 1,737.9 ± 164.5
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,059.6 ± 211.0
+141.6 OTUs{]} and monospecific {[}Number of OTUs, 1,732.7 ± 166.4
+OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,054.4 ± 209.4
 OTUs) and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 523.4 ± 136.0 OTUs)
+ambient seawater (Number of OTUs, 523.6 ± 142.4 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 community was found between mixed and monospecific \emph{C. cylindracea}
-(Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared values were
+(Jaccard, 0.36; Bray-Curtis, 0.77), while lower shared values were
 calculated between ambient seawater and epiphytic communities
@@
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (40.4 -- 53.9 \si{\percent}). To further
+proportion of total sequences (40.1 -- 53.6 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 was observed that was, however, not as strong as for \emph{C. nodosa}
-associated communities (ANOSIM, R = 0.31, \emph{p} \textless{} 0.01)
+associated communities (ANOSIM, R = 0.30, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\leq\) 53.9
+accounting for a high proportion of sequences (\(\leq\) 53.6
 \si{\percent}) (\autoref{shared}). A very similar proportion of
@@
 in the GitHub repository
-(\url{https://github.com/mkorlevic/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020}).
+(\url{https://github.com/MicrobesRovinj/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020}).
 Based on the ATCC MSA-1002 mock community included in the analysis an

### 2020-09-02 submission/manuscript.Rmd
Added cowplot citation.

@@
 For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
-Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019] combined with  packages  vegan (version 2.5-6) [@Oksanen2019], tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014].
+Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019] combined with  packages  vegan (version 2.5-6) [@Oksanen2019], tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Wilke2018].
 The detailed analysis procedure including the R Markdown file are available in the GitHub repository (https://github.com/MicrobesRovinj/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020).

### 2020-09-02 submission/manuscript.tex
Rendering cowplot reference.

@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,093
-different OTUs with 369 to 2,055 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,126
+different OTUs with 364 to 1,982 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,680.3 ±
-141.6 OTUs{]} and monospecific {[}Number of OTUs, 1,732.7 ± 166.4
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,054.4 ± 209.4
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,686.4 ±
+138.5 OTUs{]} and monospecific {[}Number of OTUs, 1,728.1 ± 145.2
+OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,049.3 ± 203.8
 OTUs) and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 523.6 ± 142.4 OTUs)
+ambient seawater (Number of OTUs, 533.2 ± 140.8 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 community was found between mixed and monospecific \emph{C. cylindracea}
-(Jaccard, 0.36; Bray-Curtis, 0.77), while lower shared values were
+(Jaccard, 0.35; Bray-Curtis, 0.78), while lower shared values were
 calculated between ambient seawater and epiphytic communities
@@
 seawater than for \emph{C. nodosa} and especially \emph{C. cylindracea}
-associated communities (\autoref{shared}). In addition, only 0.4 -- 1.0
+associated communities (\autoref{shared}). In addition, only 0.4 -- 1.1
 \si{\percent} of OTUs from each surface associated community were
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (40.1 -- 53.6 \si{\percent}). To further
+proportion of total sequences (41.2 -- 54.1 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 separation between spring, summer and autumn/winter samples (ANOSIM, R =
-0.53, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
+0.52, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
 communities a separation between summer and autumn/winter/spring samples
 was observed that was, however, not as strong as for \emph{C. nodosa}
-associated communities (ANOSIM, R = 0.30, \emph{p} \textless{} 0.01)
+associated communities (ANOSIM, R = 0.29, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 Seasonal changes in richness in the epiphytic community were substantial
-as indicated by the proportion of OTUs (\(\leq\) 1.0 \si{\percent})
+as indicated by the proportion of OTUs (\(\leq\) 1.1 \si{\percent})
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\leq\) 53.6
+accounting for a high proportion of sequences (\(\leq\) 54.1
 \si{\percent}) (\autoref{shared}). A very similar proportion of
@@
 packages (Xie, 2014, 2015, 2020; Neuwirth, 2014; Xie \emph{et al.},
-2018; Y. Xie, 2019b, 2019a; Allaire \emph{et al.}, 2019; Zhu, 2019). The
-detailed analysis procedure including the R Markdown file are available
-in the GitHub repository
+2018; Wilke, 2018; Y. Xie, 2019b, 2019a; Allaire \emph{et al.}, 2019;
+Zhu, 2019). The detailed analysis procedure including the R Markdown
+file are available in the GitHub repository
 (\url{https://github.com/MicrobesRovinj/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020}).
@@
 
+\leavevmode\hypertarget{ref-Wilke2018}{}%
+Wilke, C.O. (2018) cowplot: Streamlined plot theme and plot annotations
+for 'ggplot2'.
+
 \leavevmode\hypertarget{ref-Xie2015}{}%

### 2020-09-15 submission/manuscript.Rmd
Reference update.

--- 2020-06-08 submission/manuscript.Rmd
@@
 For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
-Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019] combined with  packages  vegan (version 2.5-6) [@Oksanen2019], tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014].
+Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019] combined with  packages  vegan (version 2.5-6) [@Oksanen2019], tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Wilke2018].
 The detailed analysis procedure including the R Markdown file are available in the GitHub repository (https://github.com/MicrobesRovinj/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020).

### 2020-09-15 submission/manuscript.tex
Rendering with changed references.

--- 2020-06-08 submission/manuscript.tex
@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,093
-different OTUs with 369 to 2,055 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,071
+different OTUs with 382 to 2,031 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,680.3 ±
-141.6 OTUs{]} and monospecific {[}Number of OTUs, 1,732.7 ± 166.4
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,054.4 ± 209.4
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,686.1 ±
+142.8 OTUs{]} and monospecific {[}Number of OTUs, 1,729.5 ± 159.6
+OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,064.1 ± 210.0
 OTUs) and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 523.6 ± 142.4 OTUs)
+ambient seawater (Number of OTUs, 525.4 ± 145.1 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (40.1 -- 53.6 \si{\percent}). To further
+proportion of total sequences (41.4 -- 52.4 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 separation between spring, summer and autumn/winter samples (ANOSIM, R =
-0.53, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
+0.54, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
 communities a separation between summer and autumn/winter/spring samples
@@
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\leq\) 53.6
+accounting for a high proportion of sequences (\(\leq\) 52.4
 \si{\percent}) (\autoref{shared}). A very similar proportion of
@@
 (version 1.3.0) (Wickham \emph{et al.}, 2019) and multiple other
-packages (Xie, 2014, 2015, 2020; Neuwirth, 2014; Xie \emph{et al.},
-2018; Y. Xie, 2019b, 2019a; Allaire \emph{et al.}, 2019; Zhu, 2019). The
+packages (Xie, 2014, 2015, 2019, 2020; Neuwirth, 2014; Xie \emph{et
+al.}, 2018; Wilke, 2018; Allaire \emph{et al.}, 2019; Zhu, 2019). The
 detailed analysis procedure including the R Markdown file are available
@@
 
-\leavevmode\hypertarget{ref-Xie2019}{}%
-Xie, Y. (2019a) knitr: A general-purpose package for dynamic report
-generation in R.
-
 \leavevmode\hypertarget{ref-Xie2019a}{}%
-Xie, Y. (2019b) TinyTeX: A lightweight, cross-platform, and
+Xie, Y. (2019) TinyTeX: A lightweight, cross-platform, and
 easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}

### 2020-11-16 submission/manuscript.Rmd
Coauthors review done.

@@
 \newpage
-\linenumbers
+[]: # (\linenumbers)
 \sisetup{mode=text}
@@
 It postulates that an initial random colonization step takes places from a set of functionally equivalent taxonomic groups resulting in taxonomically different epiphytic communities sharing a core set of functional genes [@Burke2011; @Roth-Schulze2016].
-In addition, some of the variation in the reported data could be attributed to different techniques used in these studies, such as different protocols for epiphytic cell detachment and/or DNA isolation, as no standard protocol has been yet established to study epiphytic communities [@Korlevica, submitted; @Ugarelli2019].
+In addition, some of the variation in the reported data could be attributed to different techniques used in these studies, such as different protocols for epiphytic cell detachment and/or DNA isolation, as no standard protocol has been yet established to study epiphytic communities [@Korlevic, submitted; @Ugarelli2019].
 
@@
 Read numbers were normalized to the minimum number of sequences (`r min_max$min`, \autoref{supp-nseq_notus}) through rarefaction resulting in `r no_otus_rarefied` different OTUs with `r min(otu_min_max)` to `r max(otu_min_max)` OTUs per sample (\autoref{supp-calculators}).
-To determine seasonal changes in richness and diversity the observed number of OTUs, Chao1, ACE, Exponential Shannon [@Jost2006] and Inverse Simpson were calculated after normalization through rarefaction.
+To determine seasonal changes in richness and diversity the observed number of OTUs, Chao1, ACE, Exponential Shannon and Inverse Simpson [@Jost2006] were calculated after normalization through rarefaction.
 Generally, richness estimators and diversity indices showed similar trends.
@@
 # Calculating matrices of similarity indices
-jaccard <- vegdist(rarefied_metadata, method="jaccard", binary=T)
-jaccard <- as_tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(jaccard)),
-                     .name_repair= ~c("V1", "V2", "jaccard"))
-
-bray <- vegdist(rarefied_metadata, method="bray", binary=F)
-bray <- as_tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(bray)),
-                  .name_repair= ~c("V1", "V2", "bray"))
+jaccard <- as.matrix(vegdist(rarefied_metadata, method="jaccard", binary=T))
+jaccard[lower.tri(jaccard, diag=TRUE)] <- NA
+jaccard <- as_tibble(jaccard) %>%
+  add_column("V1"=rownames(jaccard), .before=TRUE) %>%
+  gather(key="V2", value="jaccard", 2:ncol(.)) %>%
+  filter(!is.na(jaccard))
+
+bray <- as.matrix(vegdist(rarefied_metadata, method="bray", binary=F))
+bray[lower.tri(bray, diag=TRUE)] <- NA
+bray <- as_tibble(bray) %>%
+  add_column("V1"=rownames(bray), .before=TRUE) %>%
+  gather(key="V2", value="bray", 2:ncol(.)) %>%
+  filter(!is.na(bray))
 
@@
                taxlevel==6 &
-               str_detect(rankID, filter(community, str_detect(taxon, "^Cyanobacteria$"))$rankID)) %>%
+               str_detect(rankID, paste0(filter(community, str_detect(taxon, "^Cyanobacteria$"))$rankID, "\\."))) %>%
   filter_at(6:ncol(.), any_vars(. >= 1))
@@
                taxlevel==6 &
-               str_detect(rankID, filter(community, str_detect(taxon, "^Cyanobacteria$"))$rankID)) %>%
+               str_detect(rankID, paste0(filter(community, str_detect(taxon, "^Cyanobacteria$"))$rankID, "\\."))) %>%
   mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
@@
                taxlevel==6 &
-               str_detect(rankID, filter(community, str_detect(taxon, "^Alphaproteobacteria$"))$rankID)) %>%
+               str_detect(rankID, paste0(filter(community, str_detect(taxon, "^Alphaproteobacteria$"))$rankID, "\\."))) %>%
   filter_at(6:ncol(.), any_vars(. >= 2))
@@
                taxlevel==6 &
-               str_detect(rankID, filter(community, str_detect(taxon, "^Alphaproteobacteria$"))$rankID)) %>%
+               str_detect(rankID, paste0(filter(community, str_detect(taxon, "^Alphaproteobacteria$"))$rankID, "\\."))) %>%
   mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
@@
 All these methods result in different removal efficiencies but none was enabling a complete removal of attached microbial cells based on our experience.
-In the present study, we applied a removal protocol [@Korlevica, submitted] based on direct cellular lysis [@Burke2009].
+In the present study, we applied a removal protocol [@Korlevic, submitted] based on direct cellular lysis [@Burke2009].
 The application of a direct lysis procedure coupled with a high sampling frequency and Illumina amplicon sequencing has enabled us to described in detail the bacterial and archaeal communities associated with the surfaces of two marine macrophytes, *C.
@@
 Leaves and thalli were washed three times with sterile artificial seawater (ASW) to remove loosely attached microbial cells.
-Ambient seawater was collected in 10 \si{\l} containers by diving and transported to the laboratory where the whole container volume was filtered through a 20 \si{\um} net.
+Ambient seawater was collected in 10 \si{\l} containers by diving and transported to the laboratory where 10 -- 20 \si{\l} were filtered through a 20 \si{\um} net.
 The filtrate was further sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate membrane filters (Whatman, United Kingdom) using a peristaltic pump.
@@
 nodosa* and *C.
-cylindracea* was isolated using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevica, submitted; @Massana1997].
+cylindracea* was isolated using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevic, submitted; @Massana1997].
 Briefly, leaves and thalli are incubated in a lysis buffer and treated with lysozyme and proteinase K.
@@
 ### Illumina 16S rRNA sequencing
-Illumina MiSeq sequencing of the V4 region of the 16S rRNA gene was performed as described previously [@Korlevica, submitted].
+Illumina MiSeq sequencing of the V4 region of the 16S rRNA gene was performed as described previously [@Korlevic, submitted].
 The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
@@
 ## Acknowledgments
-This work was founded by the Croatian Science Foundation through the MICRO-SEAGRASS project (IP-2016-06-7118).
+This work was founded by the Croatian Science Foundation through the MICRO-SEAGRASS project (project number IP-2016-06-7118).
+ZZ and GJH were supported by the Austrian Science Fund (FWF) project ARTEMIS (project number P28781-B21).
 We would like to thank Margareta Buterer for technical support, Paolo Paliaga for help during sampling and the University Computing Center of the University of Zagreb for access to the computer cluster Isabella.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) bacterial and archaeal sequences on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 1 \\si{\\percent}) bacterial and archaeal sequences on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) cyanobacterial sequences on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 1 \\si{\\percent}) cyanobacterial sequences on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) sequences within the \\textit{Bacteroidota} on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 2 \\si{\\percent}) sequences within the \\textit{Bacteroidota} on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) alphaproteobacterial sequences on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 2 \\si{\\percent}) alphaproteobacterial sequences on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 2 \\si{\\percent}) gammaproteobacterial sequences on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 1 \\si{\\percent}) gammaproteobacterial sequences on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) sequences within the \\textit{Desulfobacterota} on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 1 \\si{\\percent}) sequences within the \\textit{Desulfobacterota} on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.

### 2020-11-16 submission/manuscript.tex
Coauthors review done.

@@
 
-\usepackage{lineno}
-\renewcommand{\linenumberfont}{\normalfont\tiny}
+%\usepackage{lineno}
+%\renewcommand{\linenumberfont}{\normalfont\tiny}
 
@@
 \newpage
-\linenumbers
+
 \sisetup{mode=text}
@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,071
-different OTUs with 382 to 2,031 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,082
+different OTUs with 379 to 2,017 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
 and diversity the observed number of OTUs, Chao1, ACE, Exponential
-Shannon (Jost, 2006) and Inverse Simpson were calculated after
+Shannon and Inverse Simpson (Jost, 2006) were calculated after
 normalization through rarefaction. Generally, richness estimators and
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,686.1 ±
-142.8 OTUs{]} and monospecific {[}Number of OTUs, 1,729.5 ± 159.6
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,064.1 ± 210.0
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,698.7 ±
+139.6 OTUs{]} and monospecific {[}Number of OTUs, 1,735.2 ± 160.1
+OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,052.8 ± 216.1
 OTUs) and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 525.4 ± 145.1 OTUs)
+ambient seawater (Number of OTUs, 528.8 ± 145.4 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (41.4 -- 52.4 \si{\percent}). To further
+proportion of total sequences (37.8 -- 53.6 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\leq\) 52.4
+accounting for a high proportion of sequences (\(\leq\) 53.6
 \si{\percent}) (\autoref{shared}). A very similar proportion of
@@
 microbial cells. Ambient seawater was collected in 10 \si{\l} containers
-by diving and transported to the laboratory where the whole container
-volume was filtered through a 20 \si{\um} net. The filtrate was further
+by diving and transported to the laboratory where 10 -- 20 \si{\l} were
+filtered through a 20 \si{\um} net. The filtrate was further
 sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate
@@
 This work was founded by the Croatian Science Foundation through the
-MICRO-SEAGRASS project (IP-2016-06-7118). We would like to thank
-Margareta Buterer for technical support, Paolo Paliaga for help during
-sampling and the University Computing Center of the University of Zagreb
-for access to the computer cluster Isabella.
+MICRO-SEAGRASS project (project number IP-2016-06-7118). ZZ and GJH were
+supported by the Austrian Science Fund (FWF) project ARTEMIS (project
+number P28781-B21). We would like to thank Margareta Buterer for
+technical support, Paolo Paliaga for help during sampling and the
+University Computing Center of the University of Zagreb for access to
+the computer cluster Isabella.
 
@@
 
-\leavevmode\hypertarget{ref-Korlevica}{}%
+\leavevmode\hypertarget{ref-Korlevic}{}%
 Korlević, M., Markovski, M., Zhao, Z., Herndl, G.J., and Najdek, M.
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) bacterial and archaeal sequences on the surfaces of the macrophytes \textit{C.
+\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 1 \si{\percent}) bacterial and archaeal sequences on the surfaces of the macrophytes \textit{C.
 nodosa} (mixed settlement) and \textit{C.
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) cyanobacterial sequences on the surfaces of the macrophytes \textit{C.
+\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 1 \si{\percent}) cyanobacterial sequences on the surfaces of the macrophytes \textit{C.
 nodosa} (mixed settlement) and \textit{C.
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) sequences within the \textit{Bacteroidota} on the surfaces of the macrophytes \textit{C.
+\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 2 \si{\percent}) sequences within the \textit{Bacteroidota} on the surfaces of the macrophytes \textit{C.
 nodosa} (mixed settlement) and \textit{C.
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) alphaproteobacterial sequences on the surfaces of the macrophytes \textit{C.
+\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 2 \si{\percent}) alphaproteobacterial sequences on the surfaces of the macrophytes \textit{C.
 nodosa} (mixed settlement) and \textit{C.
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant (> 2 \si{\percent}) gammaproteobacterial sequences on the surfaces of the macrophytes \textit{C.
+\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 1 \si{\percent}) gammaproteobacterial sequences on the surfaces of the macrophytes \textit{C.
 nodosa} (mixed settlement) and \textit{C.
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) sequences within the \textit{Desulfobacterota} on the surfaces of the macrophytes \textit{C.
+\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 1 \si{\percent}) sequences within the \textit{Desulfobacterota} on the surfaces of the macrophytes \textit{C.
 nodosa} (mixed settlement) and \textit{C.

### 2020-11-16 code/plot_calculators.R

@@ -58,7 +58,7 @@ estimators_indices_metadata <- inner_join(metadata, estimators_indices, by=c("ID
 theme <- theme(text=element_text(family="Times"), line=element_line(color="black"),
                panel.border=element_rect(fill=NA), panel.background=element_blank(),
                panel.grid=element_blank(), axis.line=element_blank(),
-               axis.text=element_text(size=12, color="black"), axis.text.x=element_text(angle=90, hjust=0.95, vjust=0.5),
+               axis.text=element_text(size=12, color="black"), axis.text.x=element_text(angle=90, hjust=0.95, vjust=1.75),
                axis.title=element_text(size=14, color="black"),
                plot.margin=unit(c(5.5, 5.5, 5.5, 5.5), "pt"), legend.position="none",
                plot.title=element_text(size=16, hjust=0.5))
@@ -89,6 +89,12 @@ p1 <- filter(estimators_indices_metadata, station=="F") %>%
   scale_shape_manual(values=shapes_p1) +
   scale_fill_manual(values=fills_p1) +
   scale_y_continuous(limits=c(300, 3000)) +
+  scale_x_date(breaks=seq(as.Date("2017-07-01"), as.Date("2018-11-01"), "months"),
+               labels=c("Jul 2017", "Aug 2017", "Sep 2017", "Oct 2017", "Nov 2017", "Dec 2017",
+                        "Jan 2018", "Feb 2018", "Mar 2018", "Apr 2018", "May 2018", "Jun 2018",
+                        "Jul 2018", "Aug 2018", "Sep 2018", "Oct 2018", ""),
+               limits=as.Date(c("2017-07-01", "2018-11-01")),
+               expand=c(0, 0)) +
   labs(x="", y="") +
   ggtitle(parse(text="bold('Seawater')")) +
   theme +
@@ -104,9 +110,15 @@ p2 <- filter(estimators_indices_metadata, station=="F") %>%
   scale_shape_manual(values=shapes_p2) +
   scale_fill_manual(values=fills_p2) +
   scale_y_continuous(limits=c(0, 150)) +
-  scale_x_date(date_break ="months" , date_labels="%b %Y") +
+  scale_x_date(breaks=seq(as.Date("2017-07-01"), as.Date("2018-11-01"), "months"),
[…]

### 2020-11-16 code/plot_community_bar_plot.R

@@ -87,7 +87,7 @@ theme <- theme(text=element_text(family="Times"), line=element_line(color="black
                axis.ticks.x=element_line(), axis.ticks.y.left=element_line(),
                axis.text.y=element_text(size=12, color="black"),
                axis.text.x=element_blank(), axis.title.y=element_text(size=14, color="black", vjust=-0.75),
-               panel.background=element_blank(), plot.margin=unit(c(5.5, 5.5, 5.5, -11), "pt"), legend.position="none",
+               panel.background=element_blank(), plot.margin=unit(c(44, 5.5, 5.5, -11), "pt"), legend.position="none",
                plot.title=element_text(size=16, hjust=0.13))
 
 # Plots generation
@@ -97,13 +97,17 @@ f <- filter(plot, station=="F") %>%
   scale_fill_manual(values=color, labels=names) + 
   labs(x=NULL, y="%") +
   ggtitle(parse(text="bold('Seawater')")) +
-  scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-07-01", "2018-10-15")),
+  scale_x_date(breaks=seq(as.Date("2017-07-01"), as.Date("2018-11-01"), "months"),
+               labels=c("Jul 2017", "Aug 2017", "Sep 2017", "Oct 2017", "Nov 2017", "Dec 2017",
+                        "Jan 2018", "Feb 2018", "Mar 2018", "Apr 2018", "May 2018", "Jun 2018",
+                        "Jul 2018", "Aug 2018", "Sep 2018", "Oct 2018", ""),
+               limits=as.Date(c("2017-07-01", "2018-11-01")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=c(0, 0), breaks=seq(0, 100, by=10)) +
   theme +
   theme(axis.text.x=element_text(size=12, color="black", angle=90,
-                                 hjust=1, vjust=0.5, margin=margin(t=5.5, unit="pt")),
-        axis.title.x=element_text(size=14, color="black"), plot.margin=unit(c(5.5, 5.5, 5.5, 5.5), "pt"),
+                                 hjust=1, vjust=2.5, margin=margin(t=5.5, unit="pt")),
[…]

### 2020-11-16 code/plot_community_bar_plot_alphaproteobacteria.R

@@ -76,10 +76,10 @@ names <- parse(text=case_when(str_detect(plot$taxon, "uncultured") ~ paste0("pla
                               plot$taxon=="Stappiaceae_ge" ~ "italic('Stappiaceae')",
                               plot$taxon=="HIMB11" ~ "plain('HIMB11')",
                               plot$taxon=="AEGEAN-169_marine_group_ge" ~ "plain('AEGEAN-169 Marine Group')",
-                              plot$taxon=="Clade_Ia" ~ "plain('Clade Ia')",
-                              plot$taxon=="Clade_Ib" ~ "plain('Clade Ib')",
-                              plot$taxon=="Clade_II_ge" ~ "plain('Clade II')",
-                              plot$taxon=="Clade_III_ge" ~ "plain('Clade III')",
+                              plot$taxon=="Clade_Ia" ~ "plain('Subclade Ia')",
+                              plot$taxon=="Clade_Ib" ~ "plain('Subclade Ib')",
+                              plot$taxon=="Clade_II_ge" ~ "plain('Subclade II')",
+                              plot$taxon=="Clade_III_ge" ~ "plain('Subclade III')",
                               plot$taxon=="Other_Alphaproteobacteria" ~ "plain('Other')~italic('Alphaproteobacteria')",
                               TRUE ~ paste0("italic('", plot$taxon, "')")))
 
@@ -207,6 +207,6 @@ p <- cowplot::ggdraw() +
   cowplot::draw_plot(fcam, x=0.317, y=0.317, width=0.683, height=0.228) +
   cowplot::draw_plot(fca, x=0.317, y=0, width=0.683, height=0.317) +
   cowplot::draw_plot(legend, x=0.120, y=0.280, width=0.100, height=0.200) +
-  cowplot::draw_line(x=c(0.135, 0.135), y=c(0.285, 0.344), size=0.5) +
-  cowplot::draw_label("SAR11 Clade", x=0.145, y=0.314, hjust=0,  fontfamily="Times", size=12)
+  cowplot::draw_line(x=c(0.160, 0.160), y=c(0.284, 0.344), size=0.5) +
+  cowplot::draw_label("SAR11 Clade", x=0.170, y=0.314, hjust=0,  fontfamily="Times", size=12)
 ggsave("results/figures/alphaproteobacteria_bar_plot.jpg", p, width=210, height=297, units="mm")

### 2020-11-16 code/plot_community_bar_plot_bacteroidota.R

@@ -122,13 +122,17 @@ f <- filter(plot, station=="F") %>%
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
   ggtitle(parse(text="bold('Seawater')")) +
-  scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-07-01", "2018-10-15")),
+  scale_x_date(breaks=seq(as.Date("2017-07-01"), as.Date("2018-11-01"), "months"),
+               labels=c("Jul 2017", "Aug 2017", "Sep 2017", "Oct 2017", "Nov 2017", "Dec 2017",
+                        "Jan 2018", "Feb 2018", "Mar 2018", "Apr 2018", "May 2018", "Jun 2018",
+                        "Jul 2018", "Aug 2018", "Sep 2018", "Oct 2018", ""),
+               limits=as.Date(c("2017-07-01", "2018-11-01")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=expand_scale(mult=c(0, 0.33)), breaks=seq(0, 100, by=10)) +
   annotate("segment", x=as.Date("2017-07-01"), y=0, xend=as.Date("2017-07-01"), yend=100.5, color="black", size=0.75) +
   theme +
   theme(axis.text.x=element_text(size=12, color="black", angle=90,
-                                 hjust=1, vjust=0.5, margin=margin(t=5.5, unit="pt")),
+                                 hjust=1, vjust=2.5, margin=margin(t=5.5, unit="pt")),
         axis.title.x=element_text(size=14, color="black"), plot.margin=unit(c(5.5, 5.5, 5.5, 5.5), "pt"),
         plot.title=element_text(size=16, hjust=0.5))
 
@@ -140,7 +144,11 @@ fcym <- filter(plot, station=="FCyM") %>%
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
   ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Mixed)')")) +
-  scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
+  scale_x_date(breaks=seq(as.Date("2017-11-01"), as.Date("2018-11-01"), "months"),
+               labels=c("Nov 2017", "Dec 2017",
[…]

### 2020-11-16 code/plot_community_bar_plot_chloroplast.R

@@ -72,7 +72,7 @@ theme <- theme(text=element_text(family="Times"), line=element_line(color="black
                axis.ticks.x=element_line(), axis.ticks.y.left=element_line(),
                axis.text.y=element_text(size=12, color="black"),
                axis.text.x=element_blank(), axis.title.y=element_text(size=14, color="black", vjust=-0.75),
-               panel.background=element_blank(), plot.margin=unit(c(5.5, 5.5, 5.5, -11), "pt"), legend.position="none",
+               panel.background=element_blank(), plot.margin=unit(c(44, 5.5, 5.5, -11), "pt"), legend.position="none",
                plot.title=element_text(size=16, hjust=0.13))
 
 # Plots generation
@@ -82,13 +82,17 @@ f <- filter(plot, station=="F") %>%
   scale_fill_manual(values=color, labels=names) + 
   labs(x=NULL, y="%") +
   ggtitle(parse(text="bold('Seawater')")) +
-  scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-07-01", "2018-10-15")),
+  scale_x_date(breaks=seq(as.Date("2017-07-01"), as.Date("2018-11-01"), "months"),
+               labels=c("Jul 2017", "Aug 2017", "Sep 2017", "Oct 2017", "Nov 2017", "Dec 2017",
+                        "Jan 2018", "Feb 2018", "Mar 2018", "Apr 2018", "May 2018", "Jun 2018",
+                        "Jul 2018", "Aug 2018", "Sep 2018", "Oct 2018", ""),
+               limits=as.Date(c("2017-07-01", "2018-11-01")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=c(0, 0), breaks=seq(0, 100, by=10)) +
   theme +
   theme(axis.text.x=element_text(size=12, color="black", angle=90,
-                                 hjust=1, vjust=0.5, margin=margin(t=5.5, unit="pt")),
-        axis.title.x=element_text(size=14, color="black"), plot.margin=unit(c(5.5, 5.5, 5.5, 5.5), "pt"),
+                                 hjust=1, vjust=2.5, margin=margin(t=5.5, unit="pt")),
[…]

### 2020-11-16 code/plot_community_bar_plot_cyanobacteria.R

@@ -119,13 +119,18 @@ f <- filter(plot, station=="F") %>%
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
   ggtitle(parse(text="bold('Seawater')")) +
-  scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-07-01", "2018-10-15")),
+  labs(x=NULL, y="%") +
+  scale_x_date(breaks=seq(as.Date("2017-07-01"), as.Date("2018-11-01"), "months"),
+               labels=c("Jul 2017", "Aug 2017", "Sep 2017", "Oct 2017", "Nov 2017", "Dec 2017",
+                        "Jan 2018", "Feb 2018", "Mar 2018", "Apr 2018", "May 2018", "Jun 2018",
+                        "Jul 2018", "Aug 2018", "Sep 2018", "Oct 2018", ""),
+               limits=as.Date(c("2017-07-01", "2018-11-01")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=expand_scale(mult=c(0, 0.33)), breaks=seq(0, 100, by=10)) +
   annotate("segment", x=as.Date("2017-07-01"), y=0, xend=as.Date("2017-07-01"), yend=100.5, color="black", size=0.75) +
   theme +
   theme(axis.text.x=element_text(size=12, color="black", angle=90,
-                                 hjust=1, vjust=0.5, margin=margin(t=5.5, unit="pt")),
+                                 hjust=1, vjust=2.5, margin=margin(t=5.5, unit="pt")),
         axis.title.x=element_text(size=14, color="black"), plot.margin=unit(c(5.5, 5.5, 5.5, 5.5), "pt"),
         plot.title=element_text(size=16, hjust=0.5))
 
@@ -137,7 +142,11 @@ fcym <- filter(plot, station=="FCyM") %>%
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
   ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Mixed)')")) +
-  scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
+  scale_x_date(breaks=seq(as.Date("2017-11-01"), as.Date("2018-11-01"), "months"),
[…]

### 2020-11-16 code/plot_community_bar_plot_desulfobacterota.R

@@ -120,13 +120,17 @@ f <- filter(plot, station=="F") %>%
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
   ggtitle(parse(text="bold('Seawater')")) +
-  scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-07-01", "2018-10-15")),
+  scale_x_date(breaks=seq(as.Date("2017-07-01"), as.Date("2018-11-01"), "months"),
+               labels=c("Jul 2017", "Aug 2017", "Sep 2017", "Oct 2017", "Nov 2017", "Dec 2017",
+                        "Jan 2018", "Feb 2018", "Mar 2018", "Apr 2018", "May 2018", "Jun 2018",
+                        "Jul 2018", "Aug 2018", "Sep 2018", "Oct 2018", ""),
+               limits=as.Date(c("2017-07-01", "2018-11-01")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=expand_scale(mult=c(0, 0.33)), breaks=seq(0, 100, by=10)) +
   annotate("segment", x=as.Date("2017-07-01"), y=0, xend=as.Date("2017-07-01"), yend=100.5, color="black", size=0.75) +
   theme +
   theme(axis.text.x=element_text(size=12, color="black", angle=90,
-                                 hjust=1, vjust=0.5, margin=margin(t=5.5, unit="pt")),
+                                 hjust=1, vjust=2.5, margin=margin(t=5.5, unit="pt")),
         axis.title.x=element_text(size=14, color="black"), plot.margin=unit(c(5.5, 5.5, 5.5, 5.5), "pt"),
         plot.title=element_text(size=16, hjust=0.5))
 
@@ -138,7 +142,11 @@ fcym <- filter(plot, station=="FCyM") %>%
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
   ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Mixed)')")) +
-  scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
+  scale_x_date(breaks=seq(as.Date("2017-11-01"), as.Date("2018-11-01"), "months"),
+               labels=c("Nov 2017", "Dec 2017",
[…]

### 2020-11-16 code/plot_community_bar_plot_gammaproteobacteria.R

@@ -122,13 +122,17 @@ f <- filter(plot, station=="F") %>%
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
   ggtitle(parse(text="bold('Seawater')")) +
-  scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-07-01", "2018-10-15")),
+  scale_x_date(breaks=seq(as.Date("2017-07-01"), as.Date("2018-11-01"), "months"),
+               labels=c("Jul 2017", "Aug 2017", "Sep 2017", "Oct 2017", "Nov 2017", "Dec 2017",
+                        "Jan 2018", "Feb 2018", "Mar 2018", "Apr 2018", "May 2018", "Jun 2018",
+                        "Jul 2018", "Aug 2018", "Sep 2018", "Oct 2018", ""),
+               limits=as.Date(c("2017-07-01", "2018-11-01")),
                expand=c(0, 0)) +
   scale_y_continuous(expand=expand_scale(mult=c(0, 0.33)), breaks=seq(0, 100, by=10)) +
   annotate("segment", x=as.Date("2017-07-01"), y=0, xend=as.Date("2017-07-01"), yend=100.5, color="black", size=0.75) +
   theme +
   theme(axis.text.x=element_text(size=12, color="black", angle=90,
-                                 hjust=1, vjust=0.5, margin=margin(t=5.5, unit="pt")),
+                                 hjust=1, vjust=2.5, margin=margin(t=5.5, unit="pt")),
         axis.title.x=element_text(size=14, color="black"), plot.margin=unit(c(5.5, 5.5, 5.5, 5.5), "pt"),
         plot.title=element_text(size=16, hjust=0.5))
 
@@ -140,7 +144,11 @@ fcym <- filter(plot, station=="FCyM") %>%
             family="Times", fontface="bold", angle=90, hjust=-1) +
   labs(x=NULL, y="%") +
   ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Mixed)')")) +
-  scale_x_date(date_break="months" , date_labels="%b %Y", limits=as.Date(c("2017-11-01", "2018-10-15")),
+  scale_x_date(breaks=seq(as.Date("2017-11-01"), as.Date("2018-11-01"), "months"),
+               labels=c("Nov 2017", "Dec 2017",
[…]

### 2020-11-16 code/plot_matrix.R

@@ -38,13 +38,19 @@ rarefied_metadata <- rarefied_metadata %>%
   select(-station)
 
 # Calculating matrices of similarity indices
-jaccard <- vegdist(rarefied_metadata, method="jaccard", binary=T)
-jaccard <- as_tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(jaccard)),
-                     .name_repair= ~c("V1", "V2", "jaccard"))
+jaccard <- as.matrix(vegdist(rarefied_metadata, method="jaccard", binary=T))
+jaccard[lower.tri(jaccard, diag=TRUE)] <- NA
+jaccard <- as_tibble(jaccard) %>%
+  add_column("V1"=rownames(jaccard), .before=TRUE) %>%
+  gather(key="V2", value="jaccard", 2:ncol(.)) %>%
+  filter(!is.na(jaccard))
 
-bray <- vegdist(rarefied_metadata, method="bray", binary=F)
-bray <- as_tibble(data.frame(t(combn(rownames(rarefied_metadata), 2)), as.numeric(bray)),
-                  .name_repair= ~c("V1", "V2", "bray"))
+bray <- as.matrix(vegdist(rarefied_metadata, method="bray", binary=F))
+bray[lower.tri(bray, diag=TRUE)] <- NA
+bray <- as_tibble(bray) %>%
+  add_column("V1"=rownames(bray), .before=TRUE) %>%
+  gather(key="V2", value="bray", 2:ncol(.)) %>%
+  filter(!is.na(bray))
 
 similarity <- inner_join(jaccard, bray, by=c("V1"="V1", "V2"="V2")) %>%
   mutate(jaccard=1-jaccard) %>%

### 2020-11-16 code/plot_seasonal_shared.R

@@ -55,7 +55,7 @@ distance_metadata <- inner_join(distance, metadata, by=c("V1"="ID")) %>%
 theme <- theme(text=element_text(family="Times"), line=element_line(color="black"),
                panel.border=element_rect(fill=NA), panel.background=element_blank(),
                panel.grid=element_blank(), axis.line=element_blank(),
-               axis.text=element_text(size=12, color="black"), axis.text.x=element_text(angle=90, hjust=0.95, vjust=0.5),
+               axis.text=element_text(size=12, color="black"), axis.text.x=element_text(angle=90, hjust=0.95, vjust=1.75),
                axis.title=element_text(size=14, color="black"),
                plot.margin=unit(c(5.5, 5.5, 5.5, 5.5), "pt"), legend.position="none",
                plot.title=element_text(size=16, hjust=0.5))
@@ -92,6 +92,12 @@ p1 <- filter(data, index=="bray") %>%
   scale_linetype_manual(values=lines) +
   scale_shape_manual(values=shapes) +
   scale_fill_manual(values=fills) +
+  scale_x_date(breaks=seq(as.Date("2017-07-01"), as.Date("2018-11-01"), "months"),
+               labels=c("Jul 2017", "Aug 2017", "Sep 2017", "Oct 2017", "Nov 2017", "Dec 2017",
+                        "Jan 2018", "Feb 2018", "Mar 2018", "Apr 2018", "May 2018", "Jun 2018",
+                        "Jul 2018", "Aug 2018", "Sep 2018", "Oct 2018", ""),
+               limits=as.Date(c("2017-07-01", "2018-11-01")),
+               expand=c(0, 0)) +
   scale_y_continuous(limits=c(0, 100)) +
   labs(x="", y="%") +
   ggtitle(parse(text="bold('Seawater')")) +
@@ -105,10 +111,16 @@ p2 <- filter(data, index=="jaccard") %>%
   scale_linetype_manual(values=lines) +
   scale_shape_manual(values=shapes) +
   scale_fill_manual(values=fills) +
-  scale_x_date(date_break ="months" , date_labels="%b %Y") +
+  scale_x_date(breaks=seq(as.Date("2017-07-01"), as.Date("2018-11-01"), "months"),
[…]

### 2020-11-17 submission/manuscript.Rmd
Summary shortened.

@@
 ## Summary
-Surfaces of marine macrophytes (seagrasses and macroalgae) are inhabited by diverse microbial communities.
+Surfaces of marine macrophytes are inhabited by diverse microbial communities.
 Most studies focusing on epiphytic communities of macrophytes did not take into account temporal changes or applied low sampling frequency approaches.
-Illumina sequencing of the V4 16S rRNA region was performed to determine the seasonal dynamics of epiphytic communities of the seagrass *Cymodocea nodosa* and invasive macroalga *Caulerpa cylindracea*.
-Leaves and thalli were sampled in a meadow of *Cymodocea nodosa* invaded by the invasive *Caulerpa cylindracea* and in a monospecific settlement of *Caulerpa cylindracea* located in the proximity of the meadow at monthly intervals.
-For comparison the ambient prokaryotic plankton community was also characterized.
+The seasonal dynamics of epiphytic microbial communities was determined in a meadow of *Cymodocea nodosa* invaded by *Caulerpa cylindracea* and in a monospecific settlement of *Caulerpa cylindracea* at monthly intervals.
+For comparison the ambient prokaryotic picoplankton community was also characterized.
 At the OTU level, the microbial community composition differed between the ambient water and the epiphytic communities exhibiting host-specificity.
@@
 Phylogenetic groups present throughout the year constituted most of the sequences, while less abundant taxa showed seasonal patterns connected to the macrophyte growth cycle.
-Taken together, epiphytic microbial communities of the seagrass *Cymodocea nodosa* and the macroalgae *Caulerpa cylindracea* appear to be host-specific and contain taxa that undergo successional changes.
+Taken together, epiphytic microbial communities of the seagrass *Cymodocea nodosa* and the macroalga *Caulerpa cylindracea* appear to be host-specific and contain taxa that undergo successional changes.
 

### 2020-11-17 submission/manuscript.tex
Summary shortened.

@@
 
-%\usepackage{lineno}
-%\renewcommand{\linenumberfont}{\normalfont\tiny}
+\usepackage{lineno}
+\renewcommand{\linenumberfont}{\normalfont\tiny}
 
@@
 
-Surfaces of marine macrophytes (seagrasses and macroalgae) are inhabited
-by diverse microbial communities. Most studies focusing on epiphytic
-communities of macrophytes did not take into account temporal changes or
-applied low sampling frequency approaches. Illumina sequencing of the V4
-16S rRNA region was performed to determine the seasonal dynamics of
-epiphytic communities of the seagrass \emph{Cymodocea nodosa} and
-invasive macroalga \emph{Caulerpa cylindracea}. Leaves and thalli were
-sampled in a meadow of \emph{Cymodocea nodosa} invaded by the invasive
-\emph{Caulerpa cylindracea} and in a monospecific settlement of
-\emph{Caulerpa cylindracea} located in the proximity of the meadow at
-monthly intervals. For comparison the ambient prokaryotic plankton
-community was also characterized. At the OTU level, the microbial
-community composition differed between the ambient water and the
-epiphytic communities exhibiting host-specificity. Also, successional
-changes were observed connected to the macrophyte growth cycle.
-Taxonomic analysis, however, showed similar high rank groups in the
-ambient water and the epiphytic communities, with the exception of
-\emph{Desulfobacterota}, which were only found on \emph{Caulerpa
-cylindracea}. \emph{Cyanobacteria} showed seasonal changes while other
-high rank taxa were present throughout the year. Phylogenetic groups
-present throughout the year constituted most of the sequences, while
-less abundant taxa showed seasonal patterns connected to the macrophyte
-growth cycle. Taken together, epiphytic microbial communities of the
-seagrass \emph{Cymodocea nodosa} and the macroalgae \emph{Caulerpa
-cylindracea} appear to be host-specific and contain taxa that undergo
-successional changes.
+Surfaces of marine macrophytes are inhabited by diverse microbial
+communities. Most studies focusing on epiphytic communities of
+macrophytes did not take into account temporal changes or applied low
+sampling frequency approaches. The seasonal dynamics of epiphytic
+microbial communities was determined in a meadow of \emph{Cymodocea
+nodosa} invaded by \emph{Caulerpa cylindracea} and in a monospecific
+settlement of \emph{Caulerpa cylindracea} at monthly intervals. For
+comparison the ambient prokaryotic picoplankton community was also
+characterized. At the OTU level, the microbial community composition
+differed between the ambient water and the epiphytic communities
+exhibiting host-specificity. Also, successional changes were observed
+connected to the macrophyte growth cycle. Taxonomic analysis, however,
+showed similar high rank groups in the ambient water and the epiphytic
+communities, with the exception of \emph{Desulfobacterota}, which were
+only found on \emph{Caulerpa cylindracea}. \emph{Cyanobacteria} showed
+seasonal changes while other high rank taxa were present throughout the
+year. Phylogenetic groups present throughout the year constituted most
+of the sequences, while less abundant taxa showed seasonal patterns
+connected to the macrophyte growth cycle. Taken together, epiphytic
+microbial communities of the seagrass \emph{Cymodocea nodosa} and the
+macroalga \emph{Caulerpa cylindracea} appear to be host-specific and
+contain taxa that undergo successional changes.
 
@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,082
-different OTUs with 379 to 2,017 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 16,986
+different OTUs with 361 to 2,026 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,698.7 ±
-139.6 OTUs{]} and monospecific {[}Number of OTUs, 1,735.2 ± 160.1
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,052.8 ± 216.1
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,678.1 ±
+143.6 OTUs{]} and monospecific {[}Number of OTUs, 1,734.2 ± 150.9
+OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,054.2 ± 213.0
 OTUs) and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 528.8 ± 145.4 OTUs)
+ambient seawater (Number of OTUs, 529.1 ± 136.4 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (37.8 -- 53.6 \si{\percent}). To further
+proportion of total sequences (36.9 -- 53.0 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\leq\) 53.6
+accounting for a high proportion of sequences (\(\leq\) 53.0
 \si{\percent}) (\autoref{shared}). A very similar proportion of

### 2020-11-20 submission/manuscript.Rmd
Formatting text.

@@
 \newpage
-[]: # (\linenumbers)
+\linenumbers
 \sisetup{mode=text}
@@
 ## Introduction
-Marine macrophytes (seagrasses and macroalgae) are important ecosystem engineers forming close associations with microorganism belonging to all three domains of life [@Tarquinio2019; @Egan2013].
+Marine macrophytes (seagrasses and macroalgae) are important ecosystem engineers forming close associations with microorganisms belonging to all three domains of life [@Tarquinio2019; @Egan2013].
 Microbes can live within macrophyte tissue as endophytes or can form epiphytic communities on surfaces of leaves and thalli [@Tarquinio2019; @Egan2013; @Hollants2013].
@@
 The majority of studies describing macrophyte epiphytic microbial communities did not include possible seasonal changes [@Ugarelli2019; @Lachnit2009; @Crump2008; @Burke2011a; @Roth-Schulze2016].
-If seasonal changes were taken into account, low temporal frequency, applied methodologies and/or limited number of analyzed host species did not allow a high taxonomic resolution [@Tujula2010; @Lachnit2011; @Miranda2013; @Michelou2013; @Mancuso2016].
+If seasonal changes were taken into account, low temporal frequency, applied methodologies and/or limited number of analysed host species did not allow a high taxonomic resolution [@Tujula2010; @Lachnit2011; @Miranda2013; @Michelou2013; @Mancuso2016].
 In the present study we describe the seasonal dynamics of bacterial and archaeal communities on the surfaces of the seagrass *Cymodocea nodosa* and siphonous macroalgae *Caulerpa cylindracea* determined on a mostly monthly scale.
@@
 nodosa* and the macroalga *C.
-cylindracea* were analyzed.
+cylindracea* were analysed.
 In addition, `r samples_seawater` samples (one of the samples was sequenced twice) originating from the ambient seawater were also processed for comparison.
@@
 In contrast, @Fahimipour2017 report in a global study of *Zostera marina*, similarities between the microbial community developed on leaves and in the ambient seawater.
-The discrepancy between our data and the study of @Fahimipour2017 could be explained by different seagrass species, methodological variations or biogeographic trends as @Fahimipour2017 analyzed samples from different locations throughout the northern hemisphere.
+The discrepancy between our data and the study of @Fahimipour2017 could be explained by different seagrass species, methodological variations or biogeographic trends as @Fahimipour2017 analysed samples from different locations throughout the northern hemisphere.
 It is possible that the microbial communities in ambient seawater and on leaves from the same location are differing but are still more similar to each other when compared to other sampling locations.
 Indeed, it was found that prokaryotic communities vary substantially between different sampling sites [@Bengtsson2017].
-When the taxonomic composition at high ranks was analyzed no such strong differentiation was noticed.
+When the taxonomic composition at high ranks was analysed no such strong differentiation was noticed.
 Phyla and classes such as *Actinobacteriota*, *Bacteroidota*, *Cyanobacteria*, *Alphaproteobacteria*, *Gammaproteobacteria* and *Verrucomicrobiota* were found in both ambient seawater as well as macrophyte associated, in agreement with previous studies [@Egan2013; @Michelou2013; @Burke2011a].
-In contrast, when low taxonomic ranks were analyzed (i.e., family and genus) a strong differentiation was observed (Figs.
+In contrast, when low taxonomic ranks were analysed (i.e., family and genus) a strong differentiation was observed (Figs.
 \ref{cyano} -- \ref{desulfo}).
@@
 
-Seasonal changes in richness in the epiphytic community were substantial as indicated by the proportion of OTUs ($\leq$ `r max(core$core_otus)` \si{\percent}) present at every sampling date.
-These persistent OTUs, however, were accounting for a high proportion of sequences ($\leq$ `r max(core$core_seq)` \si{\percent}) (\autoref{shared}).
+Seasonal changes in richness in the epiphytic community were substantial as indicated by the proportion of OTUs ($\leq$`r max(core$core_otus)` \si{\percent}) present at every sampling date.
+These persistent OTUs, however, were accounting for a high proportion of sequences ($\leq$`r max(core$core_seq)` \si{\percent}) (\autoref{shared}).
 A very similar proportion of persistent OTUs was reported in high-frequency sampling studies describing seasonal changes in picoplankton [@Gilbert2009; @Gilbert2012].
@@
 In the months of high cyanobacterial sequence abundances the majority of sequences from this group were classified as *Pleurocapsa*, a group known to colonized different living and non-living surfaces [@Longford2007; @Burns2004; @Mobberley2012; @Reisser2014].
-It is possible that during periods of low metabolic activity there is a reduced interaction and selection of the epiphytic community by the seagrass, causing leaves to become a suitable surface for nonspecific colonizers [@Zavodnik1998].
+It is possible that during periods of low metabolic activity there is a reduced interaction and selection of the epiphytic community by the seagrass, causing leaves to become a suitable surface for non-specific colonizers [@Zavodnik1998].
 *Pleurocapsa* was replaced in spring and summer by *Acrophormium*, *Phormidesmis* and sequences without known relatives within the *Nodosilineaceae*.
@@
 Other high rank taxa did not exhibit strong successional patterns.
-In every analyzed group, with the exception of *Desulfobacterota*, taxa present throughout the year in similar proportions and season specific taxa could be identified (Figs.
+In every analysed group, with the exception of *Desulfobacterota*, taxa present throughout the year in similar proportions and season specific taxa could be identified (Figs.
 \ref{bactero} and \ref{desulfo}).
@@
 cylindracea* is to be expected as part of the epiphytic community is in direct contact with the sediment.
-The *Desulfobacterota* community was dominated by *Desulfatitalea* and no relative *Desulfocapsaceae*, known sulfate sediment groups [@Higashioka2015; @Kuever2014].
+The *Desulfobacterota* community was dominated by *Desulfatitalea* and no relative *Desulfocapsaceae*, known sulphate sediment groups [@Higashioka2015; @Kuever2014].
 
 In temperate zones, marine macrophytes are exhibiting growth cycles, so it  is not surprising that the associated epiphytic microbial community is undergoing partial seasonal changes.
-In the present study, we could identify in every analyzed high rank taxa phylogenetic groups present throughout the year, comprising most of the sequences and a lower proportion of taxa showing seasonal patterns connected to the macrophyte growth cycle (Figs.
+In the present study, we could identify in every analysed high rank taxa phylogenetic groups present throughout the year, comprising most of the sequences and a lower proportion of taxa showing seasonal patterns connected to the macrophyte growth cycle (Figs.
 \ref{community} and \ref{desulfo}).
@@
 The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
-In the first PCR, the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used [@Caporaso2012; @Apprill2015; @Parada2016].
-These primers contained on their 5' end a tagged sequence.
+In the first PCR, the 515F ($5'$-GTGYCAGCMGCCGCGGTAA-$3'$) and 806R ($5'$-GGACTACNVGGGTWTCTAAT-$3'$) primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used [@Caporaso2012; @Apprill2015; @Parada2016].
+These primers contained on their $5'$ end a tagged sequence.
 Purified PCR products were sent for Illumina MiSeq sequencing at IMGM Laboratories, Martinsried, Germany.
@@
 
-Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations provided by the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
+Obtained sequences were analysed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations provided by the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
 For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 1 \\si{\\percent}) bacterial and archaeal sequences on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$1 \\si{\\percent}) bacterial and archaeal sequences on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 1 \\si{\\percent}) cyanobacterial sequences on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$1 \\si{\\percent}) cyanobacterial sequences on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 2 \\si{\\percent}) sequences within the \\textit{Bacteroidota} on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$2 \\si{\\percent}) sequences within the \\textit{Bacteroidota} on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 2 \\si{\\percent}) alphaproteobacterial sequences on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$2 \\si{\\percent}) alphaproteobacterial sequences on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 1 \\si{\\percent}) gammaproteobacterial sequences on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$1 \\si{\\percent}) gammaproteobacterial sequences on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 1 \\si{\\percent}) sequences within the \\textit{Desulfobacterota} on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$1 \\si{\\percent}) sequences within the \\textit{Desulfobacterota} on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.

### 2020-11-21 submission/manuscript.Rmd
Text formatting.

@@
 
-Seasonal changes in richness in the epiphytic community were substantial as indicated by the proportion of OTUs ($\leq$`r max(core$core_otus)` \si{\percent}) present at every sampling date.
-These persistent OTUs, however, were accounting for a high proportion of sequences ($\leq$`r max(core$core_seq)` \si{\percent}) (\autoref{shared}).
+Seasonal changes in richness in the epiphytic community were substantial as indicated by the proportion of OTUs ($\leq$ `r max(core$core_otus)` \si{\percent}) present at every sampling date.
+These persistent OTUs, however, were accounting for a high proportion of sequences ($\geq$ `r min(core$core_seq)` \si{\percent}) (\autoref{shared}).
 A very similar proportion of persistent OTUs was reported in high-frequency sampling studies describing seasonal changes in picoplankton [@Gilbert2009; @Gilbert2012].
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$1 \\si{\\percent}) bacterial and archaeal sequences on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 1 \\si{\\percent}) bacterial and archaeal sequences on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$1 \\si{\\percent}) cyanobacterial sequences on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 1 \\si{\\percent}) cyanobacterial sequences on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$2 \\si{\\percent}) sequences within the \\textit{Bacteroidota} on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 2 \\si{\\percent}) sequences within the \\textit{Bacteroidota} on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$2 \\si{\\percent}) alphaproteobacterial sequences on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 2 \\si{\\percent}) alphaproteobacterial sequences on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$1 \\si{\\percent}) gammaproteobacterial sequences on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 1 \\si{\\percent}) gammaproteobacterial sequences on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
@@
 
-```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$1 \\si{\\percent}) sequences within the \\textit{Desulfobacterota} on the surfaces of the macrophytes \\textit{C.
+```{r out.width="85%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 1 \\si{\\percent}) sequences within the \\textit{Desulfobacterota} on the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.

### 2020-11-21 submission/manuscript.tex
Text formatting.

@@
 \newpage
-
+\linenumbers
 \sisetup{mode=text}
@@
 Marine macrophytes (seagrasses and macroalgae) are important ecosystem
-engineers forming close associations with microorganism belonging to all
-three domains of life (Egan \emph{et al.}, 2013; Tarquinio \emph{et
+engineers forming close associations with microorganisms belonging to
+all three domains of life (Egan \emph{et al.}, 2013; Tarquinio \emph{et
 al.}, 2019). Microbes can live within macrophyte tissue as endophytes or
@@
 seasonal changes were taken into account, low temporal frequency,
-applied methodologies and/or limited number of analyzed host species did
+applied methodologies and/or limited number of analysed host species did
 not allow a high taxonomic resolution (Tujula \emph{et al.}, 2010;
@@
 associated with surfaces of the seagrass \emph{C. nodosa} and the
-macroalga \emph{C. cylindracea} were analyzed. In addition, 18 samples
+macroalga \emph{C. cylindracea} were analysed. In addition, 18 samples
 (one of the samples was sequenced twice) originating from the ambient
@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 16,986
-different OTUs with 361 to 2,026 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,079
+different OTUs with 384 to 2,062 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,678.1 ±
-143.6 OTUs{]} and monospecific {[}Number of OTUs, 1,734.2 ± 150.9
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,054.2 ± 213.0
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,684.1 ±
+133.8 OTUs{]} and monospecific {[}Number of OTUs, 1,731.9 ± 163.8
+OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,058.2 ± 214.5
 OTUs) and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 529.1 ± 136.4 OTUs)
+ambient seawater (Number of OTUs, 523.9 ± 138.7 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (36.9 -- 53.0 \si{\percent}). To further
+proportion of total sequences (39.5 -- 52.3 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 separation between spring, summer and autumn/winter samples (ANOSIM, R =
-0.53, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
+0.54, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
 communities a separation between summer and autumn/winter/spring samples
 was observed that was, however, not as strong as for \emph{C. nodosa}
-associated communities (ANOSIM, R = 0.31, \emph{p} \textless{} 0.01)
+associated communities (ANOSIM, R = 0.30, \emph{p} \textless{} 0.05)
 (\autoref{pcoa}).
@@
 explained by different seagrass species, methodological variations or
-biogeographic trends as Fahimipour \emph{et al.} (2017) analyzed samples
+biogeographic trends as Fahimipour \emph{et al.} (2017) analysed samples
 from different locations throughout the northern hemisphere. It is
@@
 sampling sites (Bengtsson \emph{et al.}, 2017). When the taxonomic
-composition at high ranks was analyzed no such strong differentiation
+composition at high ranks was analysed no such strong differentiation
 was noticed. Phyla and classes such as \emph{Actinobacteriota},
@@
 \emph{et al.}, 2013; Michelou \emph{et al.}, 2013). In contrast, when
-low taxonomic ranks were analyzed (i.e., family and genus) a strong
+low taxonomic ranks were analysed (i.e., family and genus) a strong
 differentiation was observed (Figs. \ref{cyano} -- \ref{desulfo}). A
@@
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\leq\) 53.0
+accounting for a high proportion of sequences (\(\geq\) 39.5
 \si{\percent}) (\autoref{shared}). A very similar proportion of
@@
 epiphytic community by the seagrass, causing leaves to become a suitable
-surface for nonspecific colonizers (Zavodnik \emph{et al.}, 1998).
+surface for non-specific colonizers (Zavodnik \emph{et al.}, 1998).
 \emph{Pleurocapsa} was replaced in spring and summer by
@@
 peak in autumn (Cardoso \emph{et al.}, 2019). Other high rank taxa did
-not exhibit strong successional patterns. In every analyzed group, with
+not exhibit strong successional patterns. In every analysed group, with
 the exception of \emph{Desulfobacterota}, taxa present throughout the
@@
 \emph{Desulfobacterota} community was dominated by \emph{Desulfatitalea}
-and no relative \emph{Desulfocapsaceae}, known sulfate sediment groups
+and no relative \emph{Desulfocapsaceae}, known sulphate sediment groups
 (Kuever, 2014; Higashioka \emph{et al.}, 2015).
@@
 is undergoing partial seasonal changes. In the present study, we could
-identify in every analyzed high rank taxa phylogenetic groups present
+identify in every analysed high rank taxa phylogenetic groups present
 throughout the year, comprising most of the sequences and a lower
@@
 The V4 region of the 16S rRNA gene was amplified using a two-step PCR
-procedure. In the first PCR, the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and
-806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome
+procedure. In the first PCR, the 515F
+(\(5'\)-GTGYCAGCMGCCGCGGTAA-\(3'\)) and 806R
+(\(5'\)-GGACTACNVGGGTWTCTAAT-\(3'\)) primers from the Earth Microbiome
 Project
@@
 were used (Caporaso \emph{et al.}, 2012; Apprill \emph{et al.}, 2015;
-Parada \emph{et al.}, 2016). These primers contained on their 5' end a
-tagged sequence. Purified PCR products were sent for Illumina MiSeq
+Parada \emph{et al.}, 2016). These primers contained on their \(5'\) end
+a tagged sequence. Purified PCR products were sent for Illumina MiSeq
 sequencing at IMGM Laboratories, Martinsried, Germany. Prior to
@@
 
-Obtained sequences were analyzed on the computer cluster Isabella
+Obtained sequences were analysed on the computer cluster Isabella
 (University Computing Center, University of Zagreb) using mothur

### 2020-11-21 submission/manuscript.tex

@@
 
+\usepackage{setspace}
+
 \begin{document}
@@
 \setlength\parindent{24pt}
+\doublespacing
 
@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,079
-different OTUs with 384 to 2,062 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,142
+different OTUs with 365 to 2,038 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,684.1 ±
-133.8 OTUs{]} and monospecific {[}Number of OTUs, 1,731.9 ± 163.8
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,058.2 ± 214.5
-OTUs) and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 523.9 ± 138.7 OTUs)
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,694.5 ±
+135.8 OTUs{]} and monospecific {[}Number of OTUs, 1,729 ± 159.6 OTUs{]})
+than for \emph{C. nodosa} (Number of OTUs, 1,061.9 ± 209.5 OTUs) and
+lowest values were obtained for the microbial community of the ambient
+seawater (Number of OTUs, 527.8 ± 146.1 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (39.5 -- 52.3 \si{\percent}). To further
+proportion of total sequences (41.8 -- 51.6 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 separation between spring, summer and autumn/winter samples (ANOSIM, R =
-0.54, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
+0.53, \emph{p} \textless{} 0.001). For \emph{C. cylindracea} associated
 communities a separation between summer and autumn/winter/spring samples
 was observed that was, however, not as strong as for \emph{C. nodosa}
-associated communities (ANOSIM, R = 0.30, \emph{p} \textless{} 0.05)
+associated communities (ANOSIM, R = 0.31, \emph{p} \textless{} 0.05)
 (\autoref{pcoa}).
@@
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\geq\) 39.5
+accounting for a high proportion of sequences (\(\geq\) 41.8
 \si{\percent}) (\autoref{shared}). A very similar proportion of

### 2021-02-23 submission/manuscript.Rmd
Text formatting.
Text format changed.

@@
 \setlength\parindent{24pt}
-
-## Summary
+\doublespacing
+
+## Abstract
 Surfaces of marine macrophytes are inhabited by diverse microbial communities.
@@
 For comparison, the microbial community of the ambient seawater was also characterized.
+
+\newpage
+## Materials and methods
+### Sampling
+Sampling was performed in the Bay of Funtana, northern Adriatic Sea (\ang{45;10;39} N, \ang{13;35;42} E).
+Leaves and thalli were sampled in a meadow of *Cymodocea nodosa* invaded by the invasive *Caulerpa cylindracea* (mixed settlement) and in a monospecific settlement of *Caulerpa cylindracea* located in the proximity of the meadow at approximately monthly intervals from December 2017 to October 2018 (\autoref{supp-nseq_notus}).
+Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with seawater collected at the sampling site.
+Upon arrival to the laboratory, *C.
+nodosa* leaves were cut into sections of 1 -- 2 \si{\cm}, while *C.
+cylindracea* thalli were cut into 5 -- 8 \si{\cm} long sections.
+Leaves and thalli were washed three times with sterile artificial seawater (ASW) to remove loosely attached microbial cells.
+Ambient seawater was collected in 10 \si{\l} containers by diving and transported to the laboratory where 10 -- 20 \si{\l} were filtered through a 20 \si{\um} net.
+The filtrate was further sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate membrane filters (Whatman, United Kingdom) using a peristaltic pump.
+Filters were briefly dried at room temperature and stored at \num{-80} \si{\degreeCelsius}.
+Seawater samples were also collected approximately monthly from July 2017 to October 2018.
+
+### DNA isolation
+DNA from surfaces of *C.
+nodosa* and *C.
+cylindracea* was isolated using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevic, submitted; @Massana1997].
+Briefly, leaves and thalli are incubated in a lysis buffer and treated with lysozyme and proteinase K.
+Following the incubations, the mixture containing lysed epiphytic cells was separated from the leaves and thalli and extracted using phenol-chloroform.
+Finally, the extracted DNA was precipitated using isopropanol.
+DNA from seawater picoplankton was extracted from 0.2 \si{\um} polycarbonate filters according to @Massana1997 with a slight modification.
+Following the phenol-chloroform extraction, 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added.
+DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 20,000 × *g* and 4 \si{\degreeCelsius} for 21 \si{\minute}.
+The pellet was washed twice with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged after each washing step at 20,000 × *g* and 4 \si{\degreeCelsius} for 5 \si{\minute}.
+Dried pellets were re-suspended in 50 -- 100 \si{\ul} of deionized water.
+
+### Illumina 16S rRNA sequencing
+Illumina MiSeq sequencing of the V4 region of the 16S rRNA gene was performed as described previously [@Korlevic, submitted].
+The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
+In the first PCR, the 515F ($5'$-GTGYCAGCMGCCGCGGTAA-$3'$) and 806R ($5'$-GGACTACNVGGGTWTCTAAT-$3'$) primers from the Earth Microbiome Project (https://earthmicrobiome.org/protocols-and-standards/16s/) were used [@Caporaso2012; @Apprill2015; @Parada2016].
+These primers contained on their $5'$ end a tagged sequence.
+Purified PCR products were sent for Illumina MiSeq sequencing at IMGM Laboratories, Martinsried, Germany.
+Prior to sequencing at IMGM, the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR.
+In addition, these primers contained adapter and sample-specific index sequences.
+Beside samples, a positive and negative control for each sequencing batch was sequenced.
+The negative control comprised PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
+Sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession number PRJEB37267.
+
+### Sequence analysis
+```{r error}
+# Calculation of the error rate based on the ATCC MSA-1002 mock community
+error <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.error.summary", na="") %>%
+  filter(numparents==1) %>%
+  mutate(mismatches_weight=weight*mismatches) %>%
+  mutate(total_weight=weight*total)
+error <- sum(error$mismatches_weight)/sum(error$total_weight)*100
+
+# Calculation of the number of sequences in the negative control
+nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
+  filter(taxon=="Root") %>%
+  select(NC_1)
+```
+
+Obtained sequences were analysed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations provided by the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
+For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
+Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019] combined with  packages  vegan (version 2.5-6) [@Oksanen2019], tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Wilke2018].
+The detailed analysis procedure including the R Markdown file are available in the GitHub repository (https://github.com/MicrobesRovinj/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021).
+Based on the ATCC MSA-1002 mock community included in the analysis an average sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
+In addition, the negative controls processed together with the samples yielded on average only `r nc` sequences after sequence quality curation.
 
@@
 
-\newpage
-## Experimental procedures
-### Sampling
-Sampling was performed in the Bay of Funtana, northern Adriatic Sea (\ang{45;10;39} N, \ang{13;35;42} E).
-Leaves and thalli were sampled in a meadow of *Cymodocea nodosa* invaded by the invasive *Caulerpa cylindracea* (mixed settlement) and in a monospecific settlement of *Caulerpa cylindracea* located in the proximity of the meadow at approximately monthly intervals from December 2017 to October 2018 (\autoref{supp-nseq_notus}).
-Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with seawater collected at the sampling site.
-Upon arrival to the laboratory, *C.
-nodosa* leaves were cut into sections of 1 -- 2 \si{\cm}, while *C.
-cylindracea* thalli were cut into 5 -- 8 \si{\cm} long sections.
-Leaves and thalli were washed three times with sterile artificial seawater (ASW) to remove loosely attached microbial cells.
-Ambient seawater was collected in 10 \si{\l} containers by diving and transported to the laboratory where 10 -- 20 \si{\l} were filtered through a 20 \si{\um} net.
-The filtrate was further sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate membrane filters (Whatman, United Kingdom) using a peristaltic pump.
-Filters were briefly dried at room temperature and stored at \num{-80} \si{\degreeCelsius}.
-Seawater samples were also collected approximately monthly from July 2017 to October 2018.
-
-### DNA isolation
-DNA from surfaces of *C.
-nodosa* and *C.
-cylindracea* was isolated using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevic, submitted; @Massana1997].
-Briefly, leaves and thalli are incubated in a lysis buffer and treated with lysozyme and proteinase K.
-Following the incubations, the mixture containing lysed epiphytic cells was separated from the leaves and thalli and extracted using phenol-chloroform.
-Finally, the extracted DNA was precipitated using isopropanol.
-DNA from seawater picoplankton was extracted from 0.2 \si{\um} polycarbonate filters according to @Massana1997 with a slight modification.
-Following the phenol-chloroform extraction, 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added.
-DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 20,000 × *g* and 4 \si{\degreeCelsius} for 21 \si{\minute}.
-The pellet was washed twice with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged after each washing step at 20,000 × *g* and 4 \si{\degreeCelsius} for 5 \si{\minute}.
-Dried pellets were re-suspended in 50 -- 100 \si{\ul} of deionized water.
-
-### Illumina 16S rRNA sequencing
-Illumina MiSeq sequencing of the V4 region of the 16S rRNA gene was performed as described previously [@Korlevic, submitted].
-The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
-In the first PCR, the 515F ($5'$-GTGYCAGCMGCCGCGGTAA-$3'$) and 806R ($5'$-GGACTACNVGGGTWTCTAAT-$3'$) primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used [@Caporaso2012; @Apprill2015; @Parada2016].
-These primers contained on their $5'$ end a tagged sequence.
-Purified PCR products were sent for Illumina MiSeq sequencing at IMGM Laboratories, Martinsried, Germany.
-Prior to sequencing at IMGM, the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR.
-In addition, these primers contained adapter and sample-specific index sequences.
-Beside samples, a positive and negative control for each sequencing batch was sequenced.
-The negative control comprised PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
-Sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession number PRJEB37267.
-
-### Sequence analysis
-```{r error}
-# Calculation of the error rate based on the ATCC MSA-1002 mock community
-error <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.error.summary", na="") %>%
-  filter(numparents==1) %>%
-  mutate(mismatches_weight=weight*mismatches) %>%
-  mutate(total_weight=weight*total)
-error <- sum(error$mismatches_weight)/sum(error$total_weight)*100
-
-# Calculation of the number of sequences in the negative control
-nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
-  filter(taxon=="Root") %>%
-  select(NC_1)
-```
-
-Obtained sequences were analysed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations provided by the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
-For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
-Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019] combined with  packages  vegan (version 2.5-6) [@Oksanen2019], tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Wilke2018].
-The detailed analysis procedure including the R Markdown file are available in the GitHub repository (https://github.com/MicrobesRovinj/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020).
-Based on the ATCC MSA-1002 mock community included in the analysis an average sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
-In addition, the negative controls processed together with the samples yielded on average only `r nc` sequences after sequence quality curation.
-
 ## Acknowledgments

### 2021-02-23 submission/manuscript.tex

--- 2020-11-21 submission/manuscript.tex
@@
 \setlength\parindent{24pt}
-
-\hypertarget{summary}{%
-\subsection{Summary}\label{summary}}
+\doublespacing
+
+\hypertarget{abstract}{%
+\subsection{Abstract}\label{abstract}}
 
@@
 engineers forming close associations with microorganisms belonging to
-all three domains of life (Egan \emph{et al.}, 2013; Tarquinio \emph{et
-al.}, 2019). Microbes can live within macrophyte tissue as endophytes or
-can form epiphytic communities on surfaces of leaves and thalli (Egan
-\emph{et al.}, 2013; Hollants \emph{et al.}, 2013; Tarquinio \emph{et
-al.}, 2019). Epiphytic and endophytic microbial communities exhibit a
-close functional relationship with the macrophyte host. It was proposed
-that this close relationship constitutes a holobiont, an integrated
-community where the macrophyte organism and its symbiotic partners
-support each other (Margulis, 1991; Egan \emph{et al.}, 2013; Tarquinio
-\emph{et al.}, 2019).
+all three domains of life (Egan et al., 2013; Tarquinio et al., 2019).
+Microbes can live within macrophyte tissue as endophytes or can form
+epiphytic communities on surfaces of leaves and thalli (Egan et al.,
+2013; Hollants et al., 2013; Tarquinio et al., 2019). Epiphytic and
+endophytic microbial communities exhibit a close functional relationship
+with the macrophyte host. It was proposed that this close relationship
+constitutes a holobiont, an integrated community where the macrophyte
+organism and its symbiotic partners support each other (Margulis, 1991;
+Egan et al., 2013; Tarquinio et al., 2019).
 
@@
 harbor cell abundances from 10\textsuperscript{2} to
-10\textsuperscript{7} \si{\cells\per\cm\squared} (Armstrong \emph{et
-al.}, 2000; Bengtsson \emph{et al.}, 2010; Burke and Thomas \emph{et
-al.}, 2011). In such an environment a number of positive and negative
-interactions between the macrophyte and the colonizing microorganisms
-have been described (Egan \emph{et al.}, 2013; Hollants \emph{et al.},
-2013; Tarquinio \emph{et al.}, 2019). Macrophytes can promote growth of
-associated microbes by nutrient exudation, while in return
-microorganisms may support macrophyte performance through improved
-nutrient availability, phytohormone production and protection from toxic
-compounds, oxidative stress, biofouling organisms and pathogens (Egan
-\emph{et al.}, 2013; Hollants \emph{et al.}, 2013; Tarquinio \emph{et
-al.}, 2019). Besides these positive interactions, macrophytes can
-negatively impact the associated microbes such as pathogenic bacteria by
-producing reactive oxygen species and secondary metabolites (Egan
-\emph{et al.}, 2013; Hollants \emph{et al.}, 2013; Tarquinio \emph{et
-al.}, 2019).
+10\textsuperscript{7} \si{\cells\per\cm\squared} (Armstrong et al.,
+2000; Bengtsson et al., 2010; Burke et al., 2011b). In such an
+environment a number of positive and negative interactions between the
+macrophyte and the colonizing microorganisms have been described (Egan
+et al., 2013; Hollants et al., 2013; Tarquinio et al., 2019).
+Macrophytes can promote growth of associated microbes by nutrient
+exudation, while in return microorganisms may support macrophyte
+performance through improved nutrient availability, phytohormone
+production and protection from toxic compounds, oxidative stress,
+biofouling organisms and pathogens (Egan et al., 2013; Hollants et al.,
+2013; Tarquinio et al., 2019). Besides these positive interactions,
+macrophytes can negatively impact the associated microbes such as
+pathogenic bacteria by producing reactive oxygen species and secondary
+metabolites (Egan et al., 2013; Hollants et al., 2013; Tarquinio et al.,
+2019).
 
@@
 \emph{Chloroflexi} and \emph{Verrucomicrobiota} (Crump and Koch, 2008;
-Tujula \emph{et al.}, 2010; Lachnit \emph{et al.}, 2011). In contrast,
-at lower taxonomic ranks host specific microbial communities were found
-(Lachnit \emph{et al.}, 2011; Roth-Schulze \emph{et al.}, 2016).
-Recently, it was shown that even different morphological niches within
-the same alga had a higher influence on the composition of the bacterial
-community than biogeography or environmental factors (Morrissey \emph{et
-al.}, 2019). While the microbial community composition varies between
-host species, metagenomic analyses revealed that the majority of the
-microbial functions are conserved (Burke and Peter Steinberg \emph{et
-al.}, 2011; Roth-Schulze \emph{et al.}, 2016). This discrepancy between
-the microbial taxonomic and functional composition might be explained by
-the lottery hypothesis. It postulates that an initial random
-colonization step takes places from a set of functionally equivalent
-taxonomic groups resulting in taxonomically different epiphytic
-communities sharing a core set of functional genes (Burke and Peter
-Steinberg \emph{et al.}, 2011; Roth-Schulze \emph{et al.}, 2016). In
-addition, some of the variation in the reported data could be attributed
-to different techniques used in these studies, such as different
-protocols for epiphytic cell detachment and/or DNA isolation, as no
-standard protocol has been yet established to study epiphytic
-communities (Ugarelli \emph{et al.}, 2019; Korlević \emph{et al.},
-submitted).
+Tujula et al., 2010; Lachnit et al., 2011). In contrast, at lower
+taxonomic ranks host specific microbial communities were found (Lachnit
+et al., 2011; Roth-Schulze et al., 2016). Recently, it was shown that
+even different morphological niches within the same alga had a higher
+influence on the composition of the bacterial community than
+biogeography or environmental factors (Morrissey et al., 2019). While
+the microbial community composition varies between host species,
+metagenomic analyses revealed that the majority of the microbial
+functions are conserved (Burke et al., 2011a; Roth-Schulze et al.,
+2016). This discrepancy between the microbial taxonomic and functional
+composition might be explained by the lottery hypothesis. It postulates
+that an initial random colonization step takes places from a set of
+functionally equivalent taxonomic groups resulting in taxonomically
+different epiphytic communities sharing a core set of functional genes
+(Burke et al., 2011a; Roth-Schulze et al., 2016). In addition, some of
+the variation in the reported data could be attributed to different
+techniques used in these studies, such as different protocols for
+epiphytic cell detachment and/or DNA isolation, as no standard protocol
+has been yet established to study epiphytic communities (Ugarelli et
+al., 2019; Korlević et al., submitted).
 
@@
 communities did not include possible seasonal changes (Crump and Koch,
-2008; Lachnit \emph{et al.}, 2009; Burke and Thomas \emph{et al.}, 2011;
-Roth-Schulze \emph{et al.}, 2016; Ugarelli \emph{et al.}, 2019). If
-seasonal changes were taken into account, low temporal frequency,
-applied methodologies and/or limited number of analysed host species did
-not allow a high taxonomic resolution (Tujula \emph{et al.}, 2010;
-Lachnit \emph{et al.}, 2011; Miranda \emph{et al.}, 2013; Michelou
-\emph{et al.}, 2013; Mancuso \emph{et al.}, 2016). In the present study
+2008; Lachnit et al., 2009; Burke et al., 2011b; Roth-Schulze et al.,
+2016; Ugarelli et al., 2019). If seasonal changes were taken into
+account, low temporal frequency, applied methodologies and/or limited
+number of analysed host species did not allow a high taxonomic
+resolution (Tujula et al., 2010; Lachnit et al., 2011; Michelou et al.,
+2013; Miranda et al., 2013; Mancuso et al., 2016). In the present study
 we describe the seasonal dynamics of bacterial and archaeal communities
@@
 seawater was also characterized.
+
+\newpage
+
+\hypertarget{materials-and-methods}{%
+\subsection{Materials and methods}\label{materials-and-methods}}
+
+\hypertarget{sampling}{%
+\subsubsection{Sampling}\label{sampling}}
+
+Sampling was performed in the Bay of Funtana, northern Adriatic Sea
+(\ang{45;10;39} N, \ang{13;35;42} E). Leaves and thalli were sampled in
+a meadow of \emph{Cymodocea nodosa} invaded by the invasive
+\emph{Caulerpa cylindracea} (mixed settlement) and in a monospecific
+settlement of \emph{Caulerpa cylindracea} located in the proximity of
+the meadow at approximately monthly intervals from December 2017 to
+October 2018 (\autoref{supp-nseq_notus}). Leaves and thalli were
+collected by diving and transported to the laboratory in containers
+placed on ice and filled with seawater collected at the sampling site.
+Upon arrival to the laboratory, \emph{C. nodosa} leaves were cut into
+sections of 1 -- 2 \si{\cm}, while \emph{C. cylindracea} thalli were cut
+into 5 -- 8 \si{\cm} long sections. Leaves and thalli were washed three
+times with sterile artificial seawater (ASW) to remove loosely attached
+microbial cells. Ambient seawater was collected in 10 \si{\l} containers
+by diving and transported to the laboratory where 10 -- 20 \si{\l} were
+filtered through a 20 \si{\um} net. The filtrate was further
+sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate
+membrane filters (Whatman, United Kingdom) using a peristaltic pump.
+Filters were briefly dried at room temperature and stored at \num{-80}
+\si{\degreeCelsius}. Seawater samples were also collected approximately
+monthly from July 2017 to October 2018.
+
+\hypertarget{dna-isolation}{%
+\subsubsection{DNA isolation}\label{dna-isolation}}
+
+DNA from surfaces of \emph{C. nodosa} and \emph{C. cylindracea} was
+isolated using a previously modified and adapted protocol that allows
+for a selective epiphytic DNA isolation (Massana et al., 1997; Korlević
+et al., submitted). Briefly, leaves and thalli are incubated in a lysis
+buffer and treated with lysozyme and proteinase K. Following the
+incubations, the mixture containing lysed epiphytic cells was separated
+from the leaves and thalli and extracted using phenol-chloroform.
+Finally, the extracted DNA was precipitated using isopropanol. DNA from
+seawater picoplankton was extracted from 0.2 \si{\um} polycarbonate
+filters according to Massana et al. (1997) with a slight modification.
+Following the phenol-chloroform extraction, 1/10 of chilled 3
+\si{\Molar} sodium acetate (pH 5.2) was added. DNA was precipitated by
+adding 1 volume of chilled isopropanol, incubating the mixtures
+overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 20,000 ×
+\emph{g} and 4 \si{\degreeCelsius} for 21 \si{\minute}. The pellet was
+washed twice with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and
+centrifuged after each washing step at 20,000 × \emph{g} and 4
+\si{\degreeCelsius} for 5 \si{\minute}. Dried pellets were re-suspended
+in 50 -- 100 \si{\ul} of deionized water.
+
+\hypertarget{illumina-16s-rrna-sequencing}{%
+\subsubsection{Illumina 16S rRNA
+sequencing}\label{illumina-16s-rrna-sequencing}}
+
+Illumina MiSeq sequencing of the V4 region of the 16S rRNA gene was
+performed as described previously (Korlević et al., submitted). The V4
+region of the 16S rRNA gene was amplified using a two-step PCR
+procedure. In the first PCR, the 515F
+(\(5'\)-GTGYCAGCMGCCGCGGTAA-\(3'\)) and 806R
+(\(5'\)-GGACTACNVGGGTWTCTAAT-\(3'\)) primers from the Earth Microbiome
+Project (\url{https://earthmicrobiome.org/protocols-and-standards/16s/})
+were used (Caporaso et al., 2012; Apprill et al., 2015; Parada et al.,
+2016). These primers contained on their \(5'\) end a tagged sequence.
+Purified PCR products were sent for Illumina MiSeq sequencing at IMGM
+Laboratories, Martinsried, Germany. Prior to sequencing at IMGM, the
+second PCR amplification of the two-step PCR procedure was performed
+using primers targeting the tagged region incorporated in the first PCR.
+In addition, these primers contained adapter and sample-specific index
+sequences. Beside samples, a positive and negative control for each
+sequencing batch was sequenced. The negative control comprised PCR
+reactions without DNA template, while for a positive control a mock
+community composed of evenly mixed DNA material originating from 20
+bacterial strains (ATCC MSA-1002, ATCC, USA) was used. Sequences
+obtained in this study have been deposited in the European Nucleotide
+Archive (ENA) at EMBL-EBI under accession number PRJEB37267.
+
+\hypertarget{sequence-analysis}{%
+\subsubsection{Sequence analysis}\label{sequence-analysis}}
+
+Obtained sequences were analysed on the computer cluster Isabella
+(University Computing Center, University of Zagreb) using mothur
+(version 1.43.0) (Schloss et al., 2009) according to the MiSeq Standard
+Operating Procedure (MiSeq SOP; \url{https://mothur.org/wiki/MiSeq_SOP})
+(Kozich et al., 2013) and recommendations provided by the Riffomonas
+project to enhance data reproducibility
+(\url{http://www.riffomonas.org/}). For alignment and classification of
+sequences the SILVA SSU Ref NR 99 database (release 138;
+\url{http://www.arb-silva.de}) was used (Quast et al., 2013; Yilmaz et
+al., 2014). Pipeline data processing and visualization was done using R
+(version 3.6.0) (R Core Team, 2019) combined with packages vegan
+(version 2.5-6) (Oksanen et al., 2019), tidyverse (version 1.3.0)
+(Wickham et al., 2019) and multiple other packages ({\textbf{???}};
+Neuwirth, 2014; Xie, 2014, 2015, 2019; Wilke, 2018; Xie et al., 2018;
+Allaire et al., 2019; Zhu, 2019). The detailed analysis procedure
+including the R Markdown file are available in the GitHub repository
+(\url{https://github.com/MicrobesRovinj/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021}).
+Based on the ATCC MSA-1002 mock community included in the analysis an
+average sequencing error rate of 0.01 \si{\percent} was determined,
+which is in line with previously reported values for next-generation
+sequencing data (Kozich et al., 2013; Schloss et al., 2016). In
+addition, the negative controls processed together with the samples
+yielded on average only 2 sequences after sequence quality curation.
 
@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,079
-different OTUs with 384 to 2,062 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,025
+different OTUs with 347 to 1,977 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,684.1 ±
-133.8 OTUs{]} and monospecific {[}Number of OTUs, 1,731.9 ± 163.8
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,058.2 ± 214.5
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,683.5 ±
+128.5 OTUs{]} and monospecific {[}Number of OTUs, 1,729.9 ± 145.8
+OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,052.2 ± 211.4
 OTUs) and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 523.9 ± 138.7 OTUs)
+ambient seawater (Number of OTUs, 529.2 ± 146.8 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (39.5 -- 52.3 \si{\percent}). To further
+proportion of total sequences (40.2 -- 52.1 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 Surfaces of marine macrophytes harbor biofilms consisting of diverse
-microbial taxa (Egan \emph{et al.}, 2013; Tarquinio \emph{et al.},
-2019). No standard protocol has been developed to study these
-macophyte-associated microbes (Ugarelli \emph{et al.}, 2019). Different
-procedures for removal of microbial cells from host surfaces are
-described, such as host tissue shaking (Nõges \emph{et al.}, 2010),
-scraping (Uku \emph{et al.}, 2007), swabbing (Mancuso \emph{et al.},
-2016) and ultrasonication (Cai \emph{et al.}, 2014). All these methods
+microbial taxa (Egan et al., 2013; Tarquinio et al., 2019). No standard
+protocol has been developed to study these macophyte-associated microbes
+(Ugarelli et al., 2019). Different procedures for removal of microbial
+cells from host surfaces are described, such as host tissue shaking
+(Nõges et al., 2010), scraping (Uku et al., 2007), swabbing (Mancuso et
+al., 2016) and ultrasonication (Cai et al., 2014). All these methods
 result in different removal efficiencies but none was enabling a
 complete removal of attached microbial cells based on our experience. In
-the present study, we applied a removal protocol (Korlević \emph{et
-al.}, submitted) based on direct cellular lysis (Burke \emph{et al.},
-2009). The application of a direct lysis procedure coupled with a high
-sampling frequency and Illumina amplicon sequencing has enabled us to
-described in detail the bacterial and archaeal communities associated
-with the surfaces of two marine macrophytes, \emph{C. nodosa} and
-\emph{C. cylindracea}.
+the present study, we applied a removal protocol (Korlević et al.,
+submitted) based on direct cellular lysis (Burke et al., 2009). The
+application of a direct lysis procedure coupled with a high sampling
+frequency and Illumina amplicon sequencing has enabled us to described
+in detail the bacterial and archaeal communities associated with the
+surfaces of two marine macrophytes, \emph{C. nodosa} and \emph{C.
+cylindracea}.
 
@@
 larger set of inhabitable microniches existing on macrophyte surfaces
-than in the ambient seawater (Ugarelli \emph{et al.}, 2019). The highest
+than in the ambient seawater (Ugarelli et al., 2019). The highest
 richness observed for \emph{C. cylindracea} might be partly due to its
@@
 two seasons than in autumn and winter (M. Najdek, personal
-communication; Zavodnik \emph{et al.}, 1998; Ruitton \emph{et al.},
-2005). During their main growth season in spring and summer macrophytes
-exhibit a more dynamic chemical interaction with the surface community
-probably causing an increase in the number of inhabitable microniches
-(Borges and Champenois, 2015; Rickert \emph{et al.}, 2016).
+communication; Zavodnik et al., 1998; Ruitton et al., 2005). During
+their main growth season in spring and summer macrophytes exhibit a more
+dynamic chemical interaction with the surface community probably causing
+an increase in the number of inhabitable microniches (Borges and
+Champenois, 2015; Rickert et al., 2016).
 
@@
 ambient seawater communities at the level of OTUs, in agreement with
-most published studies (Burke and Thomas \emph{et al.}, 2011; Michelou
-\emph{et al.}, 2013; Roth-Schulze \emph{et al.}, 2016; Mancuso \emph{et
-al.}, 2016; Crump \emph{et al.}, 2018; Ugarelli \emph{et al.}, 2019).
-This indicates that marine macrophytes are a selecting factor from the
-pool of microbial taxa present in the ambient seawater, modifying the
-microbial community once the macrophyte associated microbial biofilm
-develops (Salaün \emph{et al.}, 2012; Michelou \emph{et al.}, 2013). In
-contrast, Fahimipour \emph{et al.} (2017) report in a global study of
-\emph{Zostera marina}, similarities between the microbial community
-developed on leaves and in the ambient seawater. The discrepancy between
-our data and the study of Fahimipour \emph{et al.} (2017) could be
-explained by different seagrass species, methodological variations or
-biogeographic trends as Fahimipour \emph{et al.} (2017) analysed samples
-from different locations throughout the northern hemisphere. It is
-possible that the microbial communities in ambient seawater and on
+most published studies (Burke et al., 2011b; Michelou et al., 2013;
+Mancuso et al., 2016; Roth-Schulze et al., 2016; Crump et al., 2018;
+Ugarelli et al., 2019). This indicates that marine macrophytes are a
+selecting factor from the pool of microbial taxa present in the ambient
+seawater, modifying the microbial community once the macrophyte
+associated microbial biofilm develops (Salaün et al., 2012; Michelou et
+al., 2013). In contrast, Fahimipour et al. (2017) report in a global
+study of \emph{Zostera marina}, similarities between the microbial
+community developed on leaves and in the ambient seawater. The
+discrepancy between our data and the study of Fahimipour et al. (2017)
+could be explained by different seagrass species, methodological
+variations or biogeographic trends as Fahimipour et al. (2017) analysed
+samples from different locations throughout the northern hemisphere. It
+is possible that the microbial communities in ambient seawater and on
 leaves from the same location are differing but are still more similar
@@
 found that prokaryotic communities vary substantially between different
-sampling sites (Bengtsson \emph{et al.}, 2017). When the taxonomic
-composition at high ranks was analysed no such strong differentiation
-was noticed. Phyla and classes such as \emph{Actinobacteriota},
-\emph{Bacteroidota}, \emph{Cyanobacteria}, \emph{Alphaproteobacteria},
+sampling sites (Bengtsson et al., 2017). When the taxonomic composition
+at high ranks was analysed no such strong differentiation was noticed.
+Phyla and classes such as \emph{Actinobacteriota}, \emph{Bacteroidota},
+\emph{Cyanobacteria}, \emph{Alphaproteobacteria},
 \emph{Gammaproteobacteria} and \emph{Verrucomicrobiota} were found in
 both ambient seawater as well as macrophyte associated, in agreement
-with previous studies (Burke and Thomas \emph{et al.}, 2011; Egan
-\emph{et al.}, 2013; Michelou \emph{et al.}, 2013). In contrast, when
-low taxonomic ranks were analysed (i.e., family and genus) a strong
-differentiation was observed (Figs. \ref{cyano} -- \ref{desulfo}). A
-similar differentiation at lower microbial taxonomic ranks between
-ambient seawater and macrophytes was described for other macrophyte
-species as well (Egan \emph{et al.}, 2013; Michelou \emph{et al.}, 2013;
-Ugarelli \emph{et al.}, 2019).
+with previous studies (Burke et al., 2011b; Egan et al., 2013; Michelou
+et al., 2013). In contrast, when low taxonomic ranks were analysed
+(i.e., family and genus) a strong differentiation was observed (Figs.
+\ref{cyano} -- \ref{desulfo}). A similar differentiation at lower
+microbial taxonomic ranks between ambient seawater and macrophytes was
+described for other macrophyte species as well (Egan et al., 2013;
+Michelou et al., 2013; Ugarelli et al., 2019).
 
@@
 community is host-specific or whether there are generalist taxa
-characteristic to all or many macrophytes (Egan \emph{et al.}, 2013).
-Similar to previously described differences between microbial
-communities in the ambient seawater and on macrophytes, at high
-taxonomic ranks no major difference between the microbial communities
-associated with different hosts was observed. The only high rank phylum
-that was differing between \emph{C. nodosa} and \emph{C. cylindracea}
-was \emph{Desulfobacterota}, with more abundant sequences in the
-\emph{C. cylindracea} associated community. As already mentioned, the
-rhizoids and part of the stolon are in contact with the sediment. Thus
-\emph{Desulfobacterota} are probably a part of the epiphytic community
-that was in contact with the sediment. Similar high rank taxa found in
-this study were described to be specific for other species of
-macrophytes (Burke and Thomas \emph{et al.}, 2011; Lachnit \emph{et
-al.}, 2011; Mancuso \emph{et al.}, 2016; Bengtsson \emph{et al.}, 2017).
-In contrast to high taxonomic ranks, a substantial differentiation
-between host specific communities was found supporting the notion that
-macrophyte associated microbial communities might be host-specific.
-Host-specificity was also observed for other species of macroalgae and
-seagrasses (Lachnit \emph{et al.}, 2011; Roth-Schulze \emph{et al.},
-2016; Ugarelli \emph{et al.}, 2019; Morrissey \emph{et al.}, 2019).
-Taken together, at high taxonomic ranks a core set of taxa could be
-described that is characteristic for all or many macrophytes, while at
-low taxonomic ranks a community specific to host species was identified
-(Figs. \ref{pcoa} and \ref{community}) (Egan \emph{et al.}, 2013).
+characteristic to all or many macrophytes (Egan et al., 2013). Similar
+to previously described differences between microbial communities in the
+ambient seawater and on macrophytes, at high taxonomic ranks no major
+difference between the microbial communities associated with different
+hosts was observed. The only high rank phylum that was differing between
+\emph{C. nodosa} and \emph{C. cylindracea} was \emph{Desulfobacterota},
+with more abundant sequences in the \emph{C. cylindracea} associated
+community. As already mentioned, the rhizoids and part of the stolon are
+in contact with the sediment. Thus \emph{Desulfobacterota} are probably
+a part of the epiphytic community that was in contact with the sediment.
+Similar high rank taxa found in this study were described to be specific
+for other species of macrophytes (Burke et al., 2011b; Lachnit et al.,
+2011; Mancuso et al., 2016; Bengtsson et al., 2017). In contrast to high
+taxonomic ranks, a substantial differentiation between host specific
+communities was found supporting the notion that macrophyte associated
+microbial communities might be host-specific. Host-specificity was also
+observed for other species of macroalgae and seagrasses (Lachnit et al.,
+2011; Roth-Schulze et al., 2016; Morrissey et al., 2019; Ugarelli et
+al., 2019). Taken together, at high taxonomic ranks a core set of taxa
+could be described that is characteristic for all or many macrophytes,
+while at low taxonomic ranks a community specific to host species was
+identified (Figs. \ref{pcoa} and \ref{community}) (Egan et al., 2013).
 
@@
 present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\geq\) 39.5
+accounting for a high proportion of sequences (\(\geq\) 40.2
 \si{\percent}) (\autoref{shared}). A very similar proportion of
 persistent OTUs was reported in high-frequency sampling studies
-describing seasonal changes in picoplankton (Gilbert \emph{et al.},
-2009, 2012). In comparison to the seawater community, a lower degree of
+describing seasonal changes in picoplankton (Gilbert et al., 2009,
+2012). In comparison to the seawater community, a lower degree of
 seasonal shifts was observed for the macrophyte surface associated
@@
 Similar seasonal changes in the epiphytic community were also described
-for other macroalgae (Tujula \emph{et al.}, 2010; Lachnit \emph{et al.},
-2011). Higher seasonal stability of \emph{C. cylindracea} surface
-communities than in \emph{C. nodosa} was also observed in the higher
-proportion of shared communities between two consecutive sampling dates
-in \emph{C. cylindracea}.
+for other macroalgae (Tujula et al., 2010; Lachnit et al., 2011). Higher
+seasonal stability of \emph{C. cylindracea} surface communities than in
+\emph{C. nodosa} was also observed in the higher proportion of shared
+communities between two consecutive sampling dates in \emph{C.
+cylindracea}.
 
@@
 that is known to partially co-extract DNA from planktonic eukaryotes
-(Korlević \emph{et al.}, 2015). Strong seasonal fluctuations of high
-rank epiphytic taxa were not observed, with the exception of
+(Korlević et al., 2015). Strong seasonal fluctuations of high rank
+epiphytic taxa were not observed, with the exception of
 \emph{Cyanobacteria}. Cyanobacterial sequences were more pronounced in
@@
 group were classified as \emph{Pleurocapsa}, a group known to colonized
-different living and non-living surfaces (Burns \emph{et al.}, 2004;
-Longford \emph{et al.}, 2007; Mobberley \emph{et al.}, 2012; Reisser
-\emph{et al.}, 2014). It is possible that during periods of low
-metabolic activity there is a reduced interaction and selection of the
-epiphytic community by the seagrass, causing leaves to become a suitable
-surface for non-specific colonizers (Zavodnik \emph{et al.}, 1998).
-\emph{Pleurocapsa} was replaced in spring and summer by
-\emph{Acrophormium}, \emph{Phormidesmis} and sequences without known
-relatives within the \emph{Nodosilineaceae}. A study of coastal
-microbial mats found also a higher proportion of \emph{Nodosilineaceae}
-sequences in summer, while \emph{Phormidesmis} sequences were at their
-peak in autumn (Cardoso \emph{et al.}, 2019). Other high rank taxa did
-not exhibit strong successional patterns. In every analysed group, with
-the exception of \emph{Desulfobacterota}, taxa present throughout the
-year in similar proportions and season specific taxa could be identified
-(Figs. \ref{bactero} and \ref{desulfo}). Within \emph{Bacteroidota}
-different groups within the family \emph{Saprospiraceae} (e.g.
-\emph{Lewinella}, \emph{Portibacter} and \emph{Rubidimonas}) were
-detected across all seasons. Members of this family are often found in
-association with macrophytes and it is suggested that they are involved
-in the hydrolysis and utilization of complex carbon sources (Burke and
-Thomas \emph{et al.}, 2011; McIlroy and Nielsen, 2014; Crump \emph{et
-al.}, 2018). In contrast, the families \emph{Flavobacteriaceae} and
-\emph{Rhodothermaceae} showed seasonal patterns, with
+different living and non-living surfaces (Burns et al., 2004; Longford
+et al., 2007; Mobberley et al., 2012; Reisser et al., 2014). It is
+possible that during periods of low metabolic activity there is a
+reduced interaction and selection of the epiphytic community by the
+seagrass, causing leaves to become a suitable surface for non-specific
+colonizers (Zavodnik et al., 1998). \emph{Pleurocapsa} was replaced in
+spring and summer by \emph{Acrophormium}, \emph{Phormidesmis} and
+sequences without known relatives within the \emph{Nodosilineaceae}. A
+study of coastal microbial mats found also a higher proportion of
+\emph{Nodosilineaceae} sequences in summer, while \emph{Phormidesmis}
+sequences were at their peak in autumn (Cardoso et al., 2019). Other
+high rank taxa did not exhibit strong successional patterns. In every
+analysed group, with the exception of \emph{Desulfobacterota}, taxa
+present throughout the year in similar proportions and season specific
+taxa could be identified (Figs. \ref{bactero} and \ref{desulfo}). Within
+\emph{Bacteroidota} different groups within the family
+\emph{Saprospiraceae} (e.g. \emph{Lewinella}, \emph{Portibacter} and
+\emph{Rubidimonas}) were detected across all seasons. Members of this
+family are often found in association with macrophytes and it is
+suggested that they are involved in the hydrolysis and utilization of
+complex carbon sources (Burke et al., 2011b; McIlroy and Nielsen, 2014;
+Crump et al., 2018). In contrast, the families \emph{Flavobacteriaceae}
+and \emph{Rhodothermaceae} showed seasonal patterns, with
 \emph{Flavobacteriaceae} being more pronounced from November to June and
@@
 metabolically versatile family is often associated with macrophyte
-surfaces and usually is one of the most abundant groups (Burke and
-Thomas \emph{et al.}, 2011; Michelou \emph{et al.}, 2013; Pujalte
-\emph{et al.}, 2014; Mancuso \emph{et al.}, 2016). In addition,
-\emph{Hyphomonadaceae} were found in all samples. Interestingly, some of
-the species within this group contain stalks on their cells, which can
-be used to attach to the macrophyte surface (Weidner \emph{et al.},
-2000; Abraham and Rohde, 2014). Within the \emph{Gammaproteobacteria},
-sequences without known representatives were the most pronounced group
-present throughout the year (\autoref{gamma}). In addition,
-\emph{Granulosicoccus} was also found in almost all samples.
-\emph{Gammaproteobacteria} are often a major constituent of macrophyte
-epiphytic communities (Burke and Thomas \emph{et al.}, 2011; Michelou
-\emph{et al.}, 2013; Crump \emph{et al.}, 2018). Beside these two
-groups, other less abundant, taxa showed seasonal and host-specific
-patterns. For example, \emph{C. cylindracea} harbored \emph{Thioploca},
-a known sulfur sediment bacteria and \emph{Cellvibrionaceae}, a family
-with cultured members known as polysaccharide degraders (Jørgensen and
-Gallardo, 1999; Xie \emph{et al.}, 2017). \emph{Desulfobacterota} were
-found only associated with \emph{C. cylindracea} and no group within
-this phylum showed seasonal patterns (\autoref{desulfo}). The presence
-of this phylum only on \emph{C. cylindracea} is to be expected as part
-of the epiphytic community is in direct contact with the sediment. The
-\emph{Desulfobacterota} community was dominated by \emph{Desulfatitalea}
-and no relative \emph{Desulfocapsaceae}, known sulphate sediment groups
-(Kuever, 2014; Higashioka \emph{et al.}, 2015).
+surfaces and usually is one of the most abundant groups (Burke et al.,
+2011b; Michelou et al., 2013; Pujalte et al., 2014; Mancuso et al.,
+2016). In addition, \emph{Hyphomonadaceae} were found in all samples.
+Interestingly, some of the species within this group contain stalks on
+their cells, which can be used to attach to the macrophyte surface
+(Weidner et al., 2000; Abraham and Rohde, 2014). Within the
+\emph{Gammaproteobacteria}, sequences without known representatives were
+the most pronounced group present throughout the year (\autoref{gamma}).
+In addition, \emph{Granulosicoccus} was also found in almost all
+samples. \emph{Gammaproteobacteria} are often a major constituent of
+macrophyte epiphytic communities (Burke et al., 2011b; Michelou et al.,
+2013; Crump et al., 2018). Beside these two groups, other less abundant,
+taxa showed seasonal and host-specific patterns. For example, \emph{C.
+cylindracea} harbored \emph{Thioploca}, a known sulfur sediment bacteria
+and \emph{Cellvibrionaceae}, a family with cultured members known as
+polysaccharide degraders (Jørgensen and Gallardo, 1999; Xie et al.,
+2017). \emph{Desulfobacterota} were found only associated with \emph{C.
+cylindracea} and no group within this phylum showed seasonal patterns
+(\autoref{desulfo}). The presence of this phylum only on \emph{C.
+cylindracea} is to be expected as part of the epiphytic community is in
+direct contact with the sediment. The \emph{Desulfobacterota} community
+was dominated by \emph{Desulfatitalea} and no relative
+\emph{Desulfocapsaceae}, known sulphate sediment groups (Kuever, 2014;
+Higashioka et al., 2015).
 
@@
 hosts showed that the majority of functions could be found in every
-community, indicating functional redundancy (Roth-Schulze \emph{et al.},
-2016). This difference between phylogenetic varaibility and functional
+community, indicating functional redundancy (Roth-Schulze et al., 2016).
+This difference between phylogenetic varaibility and functional
 stability was explained by the lottery hypothesis assuming an initial
 random colonization step performed by a set of functionally equivalent
-taxonomic groups (Burke and Peter Steinberg \emph{et al.}, 2011;
-Roth-Schulze \emph{et al.}, 2016). It is possible that functional
-redundancy is a characteristic of high abundance taxa detected to be
-present throughout the year, while seasonal and/or host-specific
-functions are an attribute of taxa displaying successional patterns.
-Further studies connecting taxonomy with functional properties will be
-required to elucidate the degree of functional redundancy or specificity
-in epiphytic microbial communities.
-
-\newpage
-
-\hypertarget{experimental-procedures}{%
-\subsection{Experimental procedures}\label{experimental-procedures}}
-
-\hypertarget{sampling}{%
-\subsubsection{Sampling}\label{sampling}}
-
-Sampling was performed in the Bay of Funtana, northern Adriatic Sea
-(\ang{45;10;39} N, \ang{13;35;42} E). Leaves and thalli were sampled in
-a meadow of \emph{Cymodocea nodosa} invaded by the invasive
-\emph{Caulerpa cylindracea} (mixed settlement) and in a monospecific
-settlement of \emph{Caulerpa cylindracea} located in the proximity of
-the meadow at approximately monthly intervals from December 2017 to
-October 2018 (\autoref{supp-nseq_notus}). Leaves and thalli were
-collected by diving and transported to the laboratory in containers
-placed on ice and filled with seawater collected at the sampling site.
-Upon arrival to the laboratory, \emph{C. nodosa} leaves were cut into
-sections of 1 -- 2 \si{\cm}, while \emph{C. cylindracea} thalli were cut
-into 5 -- 8 \si{\cm} long sections. Leaves and thalli were washed three
-times with sterile artificial seawater (ASW) to remove loosely attached
-microbial cells. Ambient seawater was collected in 10 \si{\l} containers
-by diving and transported to the laboratory where 10 -- 20 \si{\l} were
-filtered through a 20 \si{\um} net. The filtrate was further
-sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate
-membrane filters (Whatman, United Kingdom) using a peristaltic pump.
-Filters were briefly dried at room temperature and stored at \num{-80}
-\si{\degreeCelsius}. Seawater samples were also collected approximately
-monthly from July 2017 to October 2018.
-
-\hypertarget{dna-isolation}{%
-\subsubsection{DNA isolation}\label{dna-isolation}}
-
-DNA from surfaces of \emph{C. nodosa} and \emph{C. cylindracea} was
-isolated using a previously modified and adapted protocol that allows
-for a selective epiphytic DNA isolation (Massana \emph{et al.}, 1997;
-Korlević \emph{et al.}, submitted). Briefly, leaves and thalli are
-incubated in a lysis buffer and treated with lysozyme and proteinase K.
-Following the incubations, the mixture containing lysed epiphytic cells
-was separated from the leaves and thalli and extracted using
-phenol-chloroform. Finally, the extracted DNA was precipitated using
-isopropanol. DNA from seawater picoplankton was extracted from 0.2
-\si{\um} polycarbonate filters according to Massana \emph{et al.} (1997)
-with a slight modification. Following the phenol-chloroform extraction,
-1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added. DNA was
-precipitated by adding 1 volume of chilled isopropanol, incubating the
-mixtures overnight at \num{-20} \si{\degreeCelsius} and centrifuging at
-20,000 × \emph{g} and 4 \si{\degreeCelsius} for 21 \si{\minute}. The
-pellet was washed twice with 500 \si{\ul} of chilled 70 \si{\percent}
-ethanol and centrifuged after each washing step at 20,000 × \emph{g} and
-4 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets were
-re-suspended in 50 -- 100 \si{\ul} of deionized water.
-
-\hypertarget{illumina-16s-rrna-sequencing}{%
-\subsubsection{Illumina 16S rRNA
-sequencing}\label{illumina-16s-rrna-sequencing}}
-
-Illumina MiSeq sequencing of the V4 region of the 16S rRNA gene was
-performed as described previously (Korlević \emph{et al.}, submitted).
-The V4 region of the 16S rRNA gene was amplified using a two-step PCR
-procedure. In the first PCR, the 515F
-(\(5'\)-GTGYCAGCMGCCGCGGTAA-\(3'\)) and 806R
-(\(5'\)-GGACTACNVGGGTWTCTAAT-\(3'\)) primers from the Earth Microbiome
-Project
-(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
-were used (Caporaso \emph{et al.}, 2012; Apprill \emph{et al.}, 2015;
-Parada \emph{et al.}, 2016). These primers contained on their \(5'\) end
-a tagged sequence. Purified PCR products were sent for Illumina MiSeq
-sequencing at IMGM Laboratories, Martinsried, Germany. Prior to
-sequencing at IMGM, the second PCR amplification of the two-step PCR
-procedure was performed using primers targeting the tagged region
-incorporated in the first PCR. In addition, these primers contained
-adapter and sample-specific index sequences. Beside samples, a positive
-and negative control for each sequencing batch was sequenced. The
-negative control comprised PCR reactions without DNA template, while for
-a positive control a mock community composed of evenly mixed DNA
-material originating from 20 bacterial strains (ATCC MSA-1002, ATCC,
-USA) was used. Sequences obtained in this study have been deposited in
-the European Nucleotide Archive (ENA) at EMBL-EBI under accession number
-PRJEB37267.
-
-\hypertarget{sequence-analysis}{%
-\subsubsection{Sequence analysis}\label{sequence-analysis}}
-
-Obtained sequences were analysed on the computer cluster Isabella
-(University Computing Center, University of Zagreb) using mothur
-(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
-Standard Operating Procedure (MiSeq SOP;
-\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
-and recommendations provided by the Riffomonas project to enhance data
-reproducibility (\url{http://www.riffomonas.org/}). For alignment and
-classification of sequences the SILVA SSU Ref NR 99 database (release
-138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
-Yilmaz \emph{et al.}, 2014). Pipeline data processing and visualization
-was done using R (version 3.6.0) (R Core Team, 2019) combined with
-packages vegan (version 2.5-6) (Oksanen \emph{et al.}, 2019), tidyverse
-(version 1.3.0) (Wickham \emph{et al.}, 2019) and multiple other
-packages (Xie, 2014, 2015, 2019, 2020; Neuwirth, 2014; Xie \emph{et
-al.}, 2018; Wilke, 2018; Allaire \emph{et al.}, 2019; Zhu, 2019). The
-detailed analysis procedure including the R Markdown file are available
-in the GitHub repository
-(\url{https://github.com/MicrobesRovinj/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020}).
-Based on the ATCC MSA-1002 mock community included in the analysis an
-average sequencing error rate of 0.01 \si{\percent} was determined,
-which is in line with previously reported values for next-generation
-sequencing data (Kozich \emph{et al.}, 2013; Schloss \emph{et al.},
-2016). In addition, the negative controls processed together with the
-samples yielded on average only 2 sequences after sequence quality
-curation.
+taxonomic groups (Burke et al., 2011a; Roth-Schulze et al., 2016). It is
+possible that functional redundancy is a characteristic of high
+abundance taxa detected to be present throughout the year, while
+seasonal and/or host-specific functions are an attribute of taxa
+displaying successional patterns. Further studies connecting taxonomy
+with functional properties will be required to elucidate the degree of
+functional redundancy or specificity in epiphytic microbial communities.
 
@@
 \leavevmode\hypertarget{ref-Abraham2014}{}%
-Abraham, W.R. and Rohde, M. (2014) The family \emph{Hyphomonadaceae}. In
-\emph{The Prokaryotes: Alphaproteobacteria and Betaproteobacteria}.
-Rosenberg, E., DeLong, E.F., Lory, S., Stackebrandt, E., and Thompson,
-F. (eds). Berlin, Heidelberg: Springer-Verlag, pp. 283--299.
+Abraham, W. R., and Rohde, M. (2014). ``The family
+\emph{Hyphomonadaceae},'' in \emph{The Prokaryotes: Alphaproteobacteria
+and Betaproteobacteria}, eds. E. Rosenberg, E. F. DeLong, S. Lory, E.
+Stackebrandt, and F. Thompson (Berlin, Heidelberg: Springer-Verlag),
+283--299.
+doi:\href{https://doi.org/10.1007/978-3-642-30197-1_260}{10.1007/978-3-642-30197-1\_260}.
 
 \leavevmode\hypertarget{ref-Allaire2019}{}%
-Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
-A., et al. (2019) rmarkdown: Dynamic documents for R.
+Allaire, J. J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
+A., et al. (2019). \emph{rmarkdown: Dynamic documents for R}.
 
 \leavevmode\hypertarget{ref-Apprill2015}{}%
-Apprill, A., McNally, S., Parsons, R., and Weber, L. (2015) Minor
+Apprill, A., McNally, S., Parsons, R., and Weber, L. (2015). Minor
 revision to V4 region SSU rRNA 806R gene primer greatly increases
-detection of SAR11 bacterioplankton. \emph{Aquat Microb Ecol}
-\textbf{75}: 129--137.
+detection of SAR11 bacterioplankton. \emph{Aquat. Microb. Ecol.} 75,
+129--137. doi:\href{https://doi.org/10.3354/ame01753}{10.3354/ame01753}.
 
 \leavevmode\hypertarget{ref-Armstrong2000}{}%
-Armstrong, E., Rogerson, A., and Leftley, J. (2000) The abundance of
-heterotrophic protists associated with intertidal seaweeds. \emph{Estuar
-Coast Shelf Sci} \textbf{50}: 415--424.
+Armstrong, E., Rogerson, A., and Leftley, J. (2000). The abundance of
+heterotrophic protists associated with intertidal seaweeds.
+\emph{Estuar. Coast. Shelf Sci.} 50, 415--424.
+doi:\href{https://doi.org/10.1006/ECSS.1999.0577}{10.1006/ECSS.1999.0577}.
 
 \leavevmode\hypertarget{ref-Bengtsson2017}{}%
-Bengtsson, M.M., Bühler, A., Brauer, A., Dahlke, S., Schubert, H., and
-Blindow, I. (2017) Eelgrass leaf surface microbiomes are locally
-variable and highly correlated with epibiotic eukaryotes. \emph{Front
-Microbiol} \textbf{8}: 1312.
+Bengtsson, M. M., Bühler, A., Brauer, A., Dahlke, S., Schubert, H., and
+Blindow, I. (2017). Eelgrass leaf surface microbiomes are locally
+variable and highly correlated with epibiotic eukaryotes. \emph{Front.
+Microbiol.} 8, 1312.
+doi:\href{https://doi.org/10.3389/fmicb.2017.01312}{10.3389/fmicb.2017.01312}.
 
 \leavevmode\hypertarget{ref-Bengtsson2010}{}%
-Bengtsson, M., Sjøtun, K., and Øvreås, L. (2010) Seasonal dynamics of
-bacterial biofilms on the kelp \emph{Laminaria hyperborea}. \emph{Aquat
-Microb Ecol} \textbf{60}: 71--83.
+Bengtsson, M., Sjøtun, K., and Øvreås, L. (2010). Seasonal dynamics of
+bacterial biofilms on the kelp \emph{Laminaria hyperborea}. \emph{Aquat.
+Microb. Ecol.} 60, 71--83.
+doi:\href{https://doi.org/10.3354/ame01409}{10.3354/ame01409}.
 
 \leavevmode\hypertarget{ref-Borges2015}{}%
-Borges, A.V. and Champenois, W. (2015) Seasonal and spatial variability
-of dimethylsulfoniopropionate (DMSP) in the Mediterranean seagrass
-\emph{Posidonia oceanica}. \emph{Aquat Bot} \textbf{125}: 72--79.
+Borges, A. V., and Champenois, W. (2015). Seasonal and spatial
+variability of dimethylsulfoniopropionate (DMSP) in the Mediterranean
+seagrass \emph{Posidonia oceanica}. \emph{Aquat. Bot.} 125, 72--79.
+doi:\href{https://doi.org/10.1016/j.aquabot.2015.05.008}{10.1016/j.aquabot.2015.05.008}.
 
 \leavevmode\hypertarget{ref-Burke2009}{}%
-Burke, C., Kjelleberg, S., and Thomas, T. (2009) Selective extraction of
-bacterial DNA from the surfaces of macroalgae. \emph{Appl Environ
-Microbiol} \textbf{75}: 252--256.
+Burke, C., Kjelleberg, S., and Thomas, T. (2009). Selective extraction
+of bacterial DNA from the surfaces of macroalgae. \emph{Appl. Environ.
+Microbiol.} 75, 252--256.
+doi:\href{https://doi.org/10.1128/AEM.01630-08}{10.1128/AEM.01630-08}.
 
@@
 Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
-(2011) Bacterial community assembly based on functional genes rather
-than species. \emph{Proc Natl Acad Sci U S A} \textbf{108}:
-14288--14293.
+(2011a). Bacterial community assembly based on functional genes rather
+than species. \emph{Proc. Natl. Acad. Sci. U. S. A.} 108, 14288--14293.
+doi:\href{https://doi.org/10.1073/pnas.1101591108}{10.1073/pnas.1101591108}.
 
@@
 Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
-(2011) Composition, uniqueness and variability of the epiphytic
+(2011b). Composition, uniqueness and variability of the epiphytic
 bacterial community of the green alga \emph{Ulva australis}. \emph{ISME
-J} \textbf{5}: 590--600.
+J.} 5, 590--600.
+doi:\href{https://doi.org/10.1038/ismej.2010.164}{10.1038/ismej.2010.164}.
 
 \leavevmode\hypertarget{ref-Burns2004}{}%
-Burns, B.P., Goh, F., Allen, M., and Neilan, B.A. (2004) Microbial
+Burns, B. P., Goh, F., Allen, M., and Neilan, B. A. (2004). Microbial
 diversity of extant stromatolites in the hypersaline marine environment
-of Shark Bay, Australia. \emph{Environ Microbiol} \textbf{6}:
-1096--1101.
+of Shark Bay, Australia. \emph{Environ. Microbiol.} 6, 1096--1101.
+doi:\href{https://doi.org/10.1111/j.1462-2920.2004.00651.x}{10.1111/j.1462-2920.2004.00651.x}.
 
 \leavevmode\hypertarget{ref-Cai2014}{}%
-Cai, X., Gao, G., Yang, J., Tang, X., Dai, J., Chen, D., and Song, Y.
-(2014) An ultrasonic method for separation of epiphytic microbes from
-freshwater submerged macrophytes. \emph{J Basic Microbiol} \textbf{54}:
+Cai, X., Gao, G., Yang, J., Tang, X., Dai, J., Chen, D., et al. (2014).
+An ultrasonic method for separation of epiphytic microbes from
+freshwater submerged macrophytes. \emph{J. Basic Microbiol.} 54,
 758--761.
+doi:\href{https://doi.org/10.1002/jobm.201300041}{10.1002/jobm.201300041}.
 
 \leavevmode\hypertarget{ref-Caporaso2012}{}%
-Caporaso, J.G., Lauber, C.L., Walters, W.A., Berg-Lyons, D., Huntley,
-J., Fierer, N., et al. (2012) Ultra-high-throughput microbial community
-analysis on the Illumina HiSeq and MiSeq platforms. \emph{ISME J}
-\textbf{6}: 1621--1624.
+Caporaso, J. G., Lauber, C. L., Walters, W. A., Berg-Lyons, D., Huntley,
+J., Fierer, N., et al. (2012). Ultra-high-throughput microbial community
+analysis on the Illumina HiSeq and MiSeq platforms. \emph{ISME J.} 6,
+1621--1624.
+doi:\href{https://doi.org/10.1038/ismej.2012.8}{10.1038/ismej.2012.8}.
 
 \leavevmode\hypertarget{ref-Cardoso2019}{}%
-Cardoso, D.C., Cretoiu, M.S., Stal, L.J., and Bolhuis, H. (2019)
-Seasonal development of a coastal microbial mat. \emph{Sci Rep}
-\textbf{9}: 9035.
+Cardoso, D. C., Cretoiu, M. S., Stal, L. J., and Bolhuis, H. (2019).
+Seasonal development of a coastal microbial mat. \emph{Sci. Rep.} 9,
+9035.
+doi:\href{https://doi.org/10.1038/s41598-019-45490-8}{10.1038/s41598-019-45490-8}.
 
 \leavevmode\hypertarget{ref-Crump2008}{}%
-Crump, B.C. and Koch, E.W. (2008) Attached bacterial populations shared
-by four species of aquatic angiosperms. \emph{Appl Environ Microbiol}
-\textbf{74}: 5948--5957.
+Crump, B. C., and Koch, E. W. (2008). Attached bacterial populations
+shared by four species of aquatic angiosperms. \emph{Appl. Environ.
+Microbiol.} 74, 5948--5957.
+doi:\href{https://doi.org/10.1128/AEM.00952-08}{10.1128/AEM.00952-08}.
 
 \leavevmode\hypertarget{ref-Crump2018}{}%
-Crump, B.C., Wojahn, J.M., Tomas, F., and Mueller, R.S. (2018)
+Crump, B. C., Wojahn, J. M., Tomas, F., and Mueller, R. S. (2018).
 Metatranscriptomics and amplicon sequencing reveal mutualisms in
-seagrass microbiomes. \emph{Front Microbiol} \textbf{9}: 388.
+seagrass microbiomes. \emph{Front. Microbiol.} 9, 388.
+doi:\href{https://doi.org/10.3389/fmicb.2018.00388}{10.3389/fmicb.2018.00388}.
 
@@
 Egan, S., Harder, T., Burke, C., Steinberg, P., Kjelleberg, S., and
-Thomas, T. (2013) The seaweed holobiont: Understanding seaweed-bacteria
-interactions. \emph{FEMS Microbiol Rev} \textbf{37}: 462--476.
+Thomas, T. (2013). The seaweed holobiont: Understanding seaweed-bacteria
+interactions. \emph{FEMS Microbiol. Rev.} 37, 462--476.
+doi:\href{https://doi.org/10.1111/1574-6976.12011}{10.1111/1574-6976.12011}.
 
 \leavevmode\hypertarget{ref-Fahimipour2017}{}%
-Fahimipour, A.K., Kardish, M.R., Lang, J.M., Green, J.L., Eisen, J.A.,
-and Stachowicz, J.J. (2017) Global-scale structure of the eelgrass
-microbiome. \emph{Appl Environ Microbiol} \textbf{83}: e03391--16.
+Fahimipour, A. K., Kardish, M. R., Lang, J. M., Green, J. L., Eisen, J.
+A., and Stachowicz, J. J. (2017). Global-scale structure of the eelgrass
+microbiome. \emph{Appl. Environ. Microbiol.} 83, e03391--16.
+doi:\href{https://doi.org/10.1128/AEM.03391-16}{10.1128/AEM.03391-16}.
 
 \leavevmode\hypertarget{ref-Gilbert2009}{}%
-Gilbert, J.A., Field, D., Swift, P., Newbold, L., Oliver, A., Smyth, T.,
-et al. (2009) The seasonal structure of microbial communities in the
-Western English Channel. \emph{Environ Microbiol} \textbf{11}:
-3132--3139.
+Gilbert, J. A., Field, D., Swift, P., Newbold, L., Oliver, A., Smyth,
+T., et al. (2009). The seasonal structure of microbial communities in
+the Western English Channel. \emph{Environ. Microbiol.} 11, 3132--3139.
+doi:\href{https://doi.org/10.1111/j.1462-2920.2009.02017.x}{10.1111/j.1462-2920.2009.02017.x}.
 
 \leavevmode\hypertarget{ref-Gilbert2012}{}%
-Gilbert, J.A., Steele, J.A., Caporaso, J.G., Steinbrück, L., Reeder, J.,
-Temperton, B., et al. (2012) Defining seasonal marine microbial
-community dynamics. \emph{ISME J} \textbf{6}: 298--308.
+Gilbert, J. A., Steele, J. A., Caporaso, J. G., Steinbrück, L., Reeder,
+J., Temperton, B., et al. (2012). Defining seasonal marine microbial
+community dynamics. \emph{ISME J.} 6, 298--308.
+doi:\href{https://doi.org/10.1038/ismej.2011.107}{10.1038/ismej.2011.107}.
 
 \leavevmode\hypertarget{ref-Higashioka2015}{}%
-Higashioka, Y., Kojima, H., Watanabe, T., and Fukui, M. (2015) Draft
+Higashioka, Y., Kojima, H., Watanabe, T., and Fukui, M. (2015). Draft
 genome sequence of \emph{Desulfatitalea tepidiphila}
-S28bF\textsuperscript{T}. \emph{Genome Announc} \textbf{3}: e01326--15.
+S28bF\textsuperscript{T}. \emph{Genome Announc.} 3, e01326--15.
+doi:\href{https://doi.org/10.1128/genomeA.01326-15}{10.1128/genomeA.01326-15}.
 
 \leavevmode\hypertarget{ref-Hollants2013}{}%
-Hollants, J., Leliaert, F., De Clerck, O., and Willems, A. (2013) What
+Hollants, J., Leliaert, F., De Clerck, O., and Willems, A. (2013). What
 we can learn from sushi: A review on seaweed-bacterial associations.
-\emph{FEMS Microbiol Ecol} \textbf{83}:
+\emph{FEMS Microbiol. Ecol.} 83.
+doi:\href{https://doi.org/10.1111/j.1574-6941.2012.01446.x}{10.1111/j.1574-6941.2012.01446.x}.
 
 \leavevmode\hypertarget{ref-Jost2006}{}%
-Jost, L. (2006) Entropy and diversity. \emph{Oikos} \textbf{113}:
-363--375.
+Jost, L. (2006). Entropy and diversity. \emph{Oikos} 113, 363--375.
+doi:\href{https://doi.org/10.1111/j.2006.0030-1299.14714.x}{10.1111/j.2006.0030-1299.14714.x}.
 
 \leavevmode\hypertarget{ref-Jorgensen1999}{}%
-Jørgensen, B.B. and Gallardo, V.A. (1999) \emph{Thioploca} spp.:
-filamentous sulfur bacteria with nitrate vacuoles. \emph{FEMS Microbiol
-Ecol} \textbf{28}: 301--313.
+Jørgensen, B. B., and Gallardo, V. A. (1999). \emph{Thioploca} spp.:
+filamentous sulfur bacteria with nitrate vacuoles. \emph{FEMS Microbiol.
+Ecol.} 28, 301--313.
+doi:\href{https://doi.org/10.1111/j.1574-6941.1999.tb00585.x}{10.1111/j.1574-6941.1999.tb00585.x}.
 
 \leavevmode\hypertarget{ref-Korlevic}{}%
-Korlević, M., Markovski, M., Zhao, Z., Herndl, G.J., and Najdek, M.
+Korlević, M., Markovski, M., Zhao, Z., Herndl, G. J., and Najdek, M.
 Selective DNA and protein isolation from marine macrophyte surfaces.
@@
 Korlević, M., Pop Ristova, P., Garić, R., Amann, R., and Orlić, S.
-(2015) Bacterial diversity in the South Adriatic Sea during a strong,
-deep winter convection year. \emph{Appl Environ Microbiol} \textbf{81}:
+(2015). Bacterial diversity in the South Adriatic Sea during a strong,
+deep winter convection year. \emph{Appl. Environ. Microbiol.} 81,
 1715--1726.
+doi:\href{https://doi.org/10.1128/AEM.03410-14}{10.1128/AEM.03410-14}.
 
 \leavevmode\hypertarget{ref-Kozich2013}{}%
-Kozich, J.J., Westcott, S.L., Baxter, N.T., Highlander, S.K., and
-Schloss, P.D. (2013) Development of a dual-index sequencing strategy and
-curation pipeline for analyzing amplicon sequence data on the MiSeq
-Illumina sequencing platform. \emph{Appl Environ Microbiol} \textbf{79}:
+Kozich, J. J., Westcott, S. L., Baxter, N. T., Highlander, S. K., and
+Schloss, P. D. (2013). Development of a dual-index sequencing strategy
+and curation pipeline for analyzing amplicon sequence data on the MiSeq
+Illumina sequencing platform. \emph{Appl. Environ. Microbiol.} 79,
 5112--5120.
+doi:\href{https://doi.org/10.1128/AEM.01043-13}{10.1128/AEM.01043-13}.
 
 \leavevmode\hypertarget{ref-Kuever2014}{}%
-Kuever, J. (2014) The family \emph{Desulfobulbaceae}. In \emph{The
-Prokaryotes: Deltaproteobacteria and Epsilonproteobacteria}. Rosenberg,
-E., DeLong, E.F., Lory, S., Stackebrandt, E., and Thompson, F. (eds).
-Berlin, Heidelberg: Springer-Verlag, pp. 75--86.
+Kuever, J. (2014). ``The family \emph{Desulfobulbaceae},'' in \emph{The
+Prokaryotes: Deltaproteobacteria and Epsilonproteobacteria}, eds. E.
+Rosenberg, E. F. DeLong, S. Lory, E. Stackebrandt, and F. Thompson
+(Berlin, Heidelberg: Springer-Verlag), 75--86.
+doi:\href{https://doi.org/10.1007/978-3-642-39044-9_267}{10.1007/978-3-642-39044-9\_267}.
 
 \leavevmode\hypertarget{ref-Lachnit2009}{}%
-Lachnit, T., Blümel, M., Imhoff, J.F., and Wahl, M. (2009) Specific
+Lachnit, T., Blümel, M., Imhoff, J. F., and Wahl, M. (2009). Specific
 epibacterial communities on macroalgae: Phylogeny matters more than
-habitat. \emph{Aquat Biol} \textbf{5}: 181--186.
+habitat. \emph{Aquat. Biol.} 5, 181--186.
+doi:\href{https://doi.org/10.3354/ab00149}{10.3354/ab00149}.
 
 \leavevmode\hypertarget{ref-Lachnit2011}{}%
-Lachnit, T., Meske, D., Wahl, M., Harder, T., and Schmitz, R. (2011)
+Lachnit, T., Meske, D., Wahl, M., Harder, T., and Schmitz, R. (2011).
 Epibacterial community patterns on marine macroalgae are host-specific
-but temporally variable. \emph{Environ Microbiol} \textbf{13}: 655--665.
+but temporally variable. \emph{Environ. Microbiol.} 13, 655--665.
+doi:\href{https://doi.org/10.1111/j.1462-2920.2010.02371.x}{10.1111/j.1462-2920.2010.02371.x}.
 
@@
 Longford, S., Tujula, N., Crocetti, G., Holmes, A., Holmström, C.,
-Kjelleberg, S., et al. (2007) Comparisons of diversity of bacterial
-communities associated with three sessile marine eukaryotes. \emph{Aquat
-Microb Ecol} \textbf{48}: 217--229.
+Kjelleberg, S., et al. (2007). Comparisons of diversity of bacterial
+communities associated with three sessile marine eukaryotes.
+\emph{Aquat. Microb. Ecol.} 48, 217--229.
+doi:\href{https://doi.org/10.3354/ame048217}{10.3354/ame048217}.
 
 \leavevmode\hypertarget{ref-Mancuso2016}{}%
-Mancuso, F.P., D'Hondt, S., Willems, A., Airoldi, L., and De Clerck, O.
-(2016) Diversity and temporal dynamics of the epiphytic bacterial
+Mancuso, F. P., D'Hondt, S., Willems, A., Airoldi, L., and De Clerck, O.
+(2016). Diversity and temporal dynamics of the epiphytic bacterial
 communities associated with the canopy-forming seaweed \emph{Cystoseira
-compressa} (Esper) Gerloff and Nizamuddin. \emph{Front Microbiol}
-\textbf{7}: 476.
+compressa} (Esper) Gerloff and Nizamuddin. \emph{Front Microbiol} 7,
+476.
+doi:\href{https://doi.org/10.3389/fmicb.2016.00476}{10.3389/fmicb.2016.00476}.
 
 \leavevmode\hypertarget{ref-Margulis1991}{}%
-Margulis, L. (1991) Symbiogenesis and symbionticism. In \emph{Symbiosis
-as a Source of Evolutionary Innovation: Speciation and Morphogenesis}.
-Margulis, L. and Fester, R. (eds). Cambridge, Massachusetts: The MIT
-Press, pp. 1--14.
+Margulis, L. (1991). ``Symbiogenesis and symbionticism,'' in
+\emph{Symbiosis as a Source of Evolutionary Innovation: Speciation and
+Morphogenesis}, eds. L. Margulis and R. Fester (Cambridge,
+Massachusetts: The MIT Press), 1--14.
 
 \leavevmode\hypertarget{ref-Massana1997}{}%
-Massana, R., Murray, A.E., Preston, C.M., and DeLong, E.F. (1997)
+Massana, R., Murray, A. E., Preston, C. M., and DeLong, E. F. (1997).
 Vertical distribution and phylogenetic characterization of marine
-planktonic \emph{Archaea} in the Santa Barbara Channel. \emph{Appl
-Environ Microbiol} \textbf{63}: 50--56.
+planktonic \emph{Archaea} in the Santa Barbara Channel. \emph{Appl.
+Environ. Microbiol.} 63, 50--56.
 
 \leavevmode\hypertarget{ref-McIlroy2014}{}%
-McIlroy, S.J. and Nielsen, P.H. (2014) The family \emph{Saprospiraceae}.
-In \emph{The Prokaryotes: Other Major Lineages of Bacteria and the
-Archaea}. Rosenberg, E., DeLong, E.F., Lory, S., Stackebrandt, E., and
-Thompson, F. (eds). Berlin, Heidelberg: Springer-Verlag, pp. 863--889.
+McIlroy, S. J., and Nielsen, P. H. (2014). ``The family
+\emph{Saprospiraceae},'' in \emph{The Prokaryotes: Other Major Lineages
+of Bacteria and the Archaea}, eds. E. Rosenberg, E. F. DeLong, S. Lory,
+E. Stackebrandt, and F. Thompson (Berlin, Heidelberg: Springer-Verlag),
+863--889.
+doi:\href{https://doi.org/10.1007/978-3-642-38954-2_138}{10.1007/978-3-642-38954-2\_138}.
 
 \leavevmode\hypertarget{ref-Michelou2013}{}%
-Michelou, V.K., Caporaso, J.G., Knight, R., and Palumbi, S.R. (2013) The
-ecology of microbial communities associated with \emph{Macrocystis
-pyrifera}. \emph{PloS One} \textbf{8}: e67480.
+Michelou, V. K., Caporaso, J. G., Knight, R., and Palumbi, S. R. (2013).
+The ecology of microbial communities associated with \emph{Macrocystis
+pyrifera}. \emph{PloS One} 8, e67480.
+doi:\href{https://doi.org/10.1371/journal.pone.0067480}{10.1371/journal.pone.0067480}.
 
 \leavevmode\hypertarget{ref-Miranda2013}{}%
-Miranda, L.N., Hutchison, K., Grossman, A.R., and Brawley, S.H. (2013)
-Diversity and abundance of the bacterial community of the red macroalga
-\emph{Porphyra umbilicalis}: Did bacterial farmers produce macroalgae?
-\emph{PloS One} \textbf{8}: e58269.
+Miranda, L. N., Hutchison, K., Grossman, A. R., and Brawley, S. H.
+(2013). Diversity and abundance of the bacterial community of the red
+macroalga \emph{Porphyra umbilicalis}: Did bacterial farmers produce
+macroalgae? \emph{PloS One} 8, e58269.
+doi:\href{https://doi.org/10.1371/journal.pone.0058269}{10.1371/journal.pone.0058269}.
 
 \leavevmode\hypertarget{ref-Mobberley2012}{}%
-Mobberley, J.M., Ortega, M.C., and Foster, J.S. (2012) Comparative
+Mobberley, J. M., Ortega, M. C., and Foster, J. S. (2012). Comparative
 microbial diversity analyses of modern marine thrombolitic mats by
-barcoded pyrosequencing. \emph{Environ Microbiol} \textbf{14}: 82--100.
+barcoded pyrosequencing. \emph{Environ. Microbiol.} 14, 82--100.
+doi:\href{https://doi.org/10.1111/j.1462-2920.2011.02509.x}{10.1111/j.1462-2920.2011.02509.x}.
 
 \leavevmode\hypertarget{ref-Morrissey2019}{}%
-Morrissey, K.L., Çavas, L., Willems, A., and De Clerck, O. (2019)
+Morrissey, K. L., Çavas, L., Willems, A., and De Clerck, O. (2019).
 Disentangling the influence of environment, host specificity and thallus
 differentiation on bacterial communities in siphonous green seaweeds.
-\emph{Front Microbiol} \textbf{10}: 717.
+\emph{Front. Microbiol.} 10, 717.
+doi:\href{https://doi.org/10.3389/fmicb.2019.00717}{10.3389/fmicb.2019.00717}.
 
 \leavevmode\hypertarget{ref-Neuwirth2014}{}%
-Neuwirth, E. (2014) RColorBrewer: ColorBrewer palettes.
+Neuwirth, E. (2014). \emph{RColorBrewer: ColorBrewer palettes}.
 
 \leavevmode\hypertarget{ref-Noges2010}{}%
-Nõges, T., Luup, H., and Feldmann, T. (2010) Primary production of
+Nõges, T., Luup, H., and Feldmann, T. (2010). Primary production of
 aquatic macrophytes and their epiphytes in two shallow lakes (Peipsi and
-Võrtsjärv) in Estonia. \emph{Aquat Ecol} \textbf{44}: 83--92.
+Võrtsjärv) in Estonia. \emph{Aquat. Ecol.} 44, 83--92.
+doi:\href{https://doi.org/10.1007/s10452-009-9249-4}{10.1007/s10452-009-9249-4}.
 
 \leavevmode\hypertarget{ref-Oksanen2019}{}%
-Oksanen, J., Blanchet, F.G., Friendly, M., Kindt, R., Legendre, P.,
-McGlinn, D., et al. (2019) vegan: Community ecology package.
+Oksanen, J., Blanchet, F. G., Friendly, M., Kindt, R., Legendre, P.,
+McGlinn, D., et al. (2019). \emph{vegan: Community ecology package}.
 
 \leavevmode\hypertarget{ref-Parada2016}{}%
-Parada, A.E., Needham, D.M., and Fuhrman, J.A. (2016) Every base
+Parada, A. E., Needham, D. M., and Fuhrman, J. A. (2016). Every base
 matters: Assessing small subunit rRNA primers for marine microbiomes
 with mock communities, time series and global field samples.
-\emph{Environ Microbiol} \textbf{18}: 1403--1414.
+\emph{Environ. Microbiol.} 18, 1403--1414.
+doi:\href{https://doi.org/10.1111/1462-2920.13023}{10.1111/1462-2920.13023}.
 
 \leavevmode\hypertarget{ref-Pujalte2014}{}%
-Pujalte, M.J., Lucena, T., Ruvira, M.A., Arahal, D.R., and Macián, M.C.
-(2014) The family \emph{Rhodobacteraceae}. In \emph{The Prokaryotes:
-Alphaproteobacteria and Betaproteobacteria}. Rosenberg, E., DeLong,
-E.F., Lory, S., Stackebrandt, E., and Thompson, F. (eds). Berlin,
-Heidelberg: Springer-Verlag, pp. 439--512.
+Pujalte, M. J., Lucena, T., Ruvira, M. A., Arahal, D. R., and Macián, M.
+C. (2014). ``The family \emph{Rhodobacteraceae},'' in \emph{The
+Prokaryotes: Alphaproteobacteria and Betaproteobacteria}, eds. E.
+Rosenberg, E. F. DeLong, S. Lory, E. Stackebrandt, and F. Thompson
+(Berlin, Heidelberg: Springer-Verlag), 439--512.
+doi:\href{https://doi.org/10.1007/978-3-642-30197-1_377}{10.1007/978-3-642-30197-1\_377}.
 
@@
 Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
-et al. (2013) The SILVA ribosomal RNA gene database project: Improved
-data processing and web-based tools. \emph{Nucleic Acids Res}
-\textbf{41}: D590--D596.
+et al. (2013). The SILVA ribosomal RNA gene database project: Improved
+data processing and web-based tools. \emph{Nucleic Acids Res.} 41,
+D590--D596.
+doi:\href{https://doi.org/10.1093/nar/gks1219}{10.1093/nar/gks1219}.
 
 \leavevmode\hypertarget{ref-RCoreTeam2019}{}%
-R Core Team (2019) R: A language and environment for statistical
-computing, Vienna, Austria: R Foundation for Statistical Computing.
+R Core Team (2019). \emph{R: A language and environment for statistical
+computing}. Vienna, Austria: R Foundation for Statistical Computing.
 
 \leavevmode\hypertarget{ref-Reisser2014}{}%
-Reisser, J., Shaw, J., Hallegraeff, G., Proietti, M., Barnes, D.K.A.,
-Thums, M., et al. (2014) Millimeter-sized marine plastics: A new pelagic
-habitat for microorganisms and invertebrates. \emph{PloS One}
-\textbf{9}: e100289.
+Reisser, J., Shaw, J., Hallegraeff, G., Proietti, M., Barnes, D. K. A.,
+Thums, M., et al. (2014). Millimeter-sized marine plastics: A new
+pelagic habitat for microorganisms and invertebrates. \emph{PloS One} 9,
+e100289.
+doi:\href{https://doi.org/10.1371/journal.pone.0100289}{10.1371/journal.pone.0100289}.
 
 \leavevmode\hypertarget{ref-Reyes2001}{}%
-Reyes, J. and Sansón, M. (2001) Biomass and production of the epiphytes
-on the leaves of \emph{Cymodocea nodosa} in the Canary Islands.
-\emph{Bot Mar} \textbf{44}: 307--313.
+Reyes, J., and Sansón, M. (2001). Biomass and production of the
+epiphytes on the leaves of \emph{Cymodocea nodosa} in the Canary
+Islands. \emph{Bot. Mar.} 44, 307--313.
+doi:\href{https://doi.org/10.1515/BOT.2001.039}{10.1515/BOT.2001.039}.
 
 \leavevmode\hypertarget{ref-Rickert2016}{}%
-Rickert, E., Wahl, M., Link, H., Richter, H., and Pohnert, G. (2016)
+Rickert, E., Wahl, M., Link, H., Richter, H., and Pohnert, G. (2016).
 Seasonal variations in surface metabolite composition of \emph{Fucus
 vesiculosus} and \emph{Fucus serratus} from the Baltic Sea. \emph{PloS
-One} \textbf{11}: e0168196.
+One} 11, e0168196.
+doi:\href{https://doi.org/10.1371/journal.pone.0168196}{10.1371/journal.pone.0168196}.
 
 \leavevmode\hypertarget{ref-Roth-Schulze2016}{}%
-Roth-Schulze, A.J., Zozaya-Valdés, E., Steinberg, P.D., and Thomas, T.
-(2016) Partitioning of functional and taxonomic diversity in
-surface-associated microbial communities. \emph{Environ Microbiol}
-\textbf{18}: 4391--4402.
+Roth-Schulze, A. J., Zozaya-Valdés, E., Steinberg, P. D., and Thomas, T.
+(2016). Partitioning of functional and taxonomic diversity in
+surface-associated microbial communities. \emph{Environ. Microbiol.} 18,
+4391--4402.
+doi:\href{https://doi.org/10.1111/1462-2920.13325}{10.1111/1462-2920.13325}.
 
 \leavevmode\hypertarget{ref-Ruitton2005}{}%
-Ruitton, S., Verlaque, M., and Boudouresque, C.F. (2005) Seasonal
+Ruitton, S., Verlaque, M., and Boudouresque, C. F. (2005). Seasonal
 changes of the introduced \emph{Caulerpa racemosa} var.
 \emph{cylindracea} (Caulerpales, Chlorophyta) at the northwest limit of
-its Mediterranean range. \emph{Aquat Bot} \textbf{82}: 55--70.
+its Mediterranean range. \emph{Aquat. Bot.} 82, 55--70.
+doi:\href{https://doi.org/10.1016/j.aquabot.2005.02.008}{10.1016/j.aquabot.2005.02.008}.
 
 \leavevmode\hypertarget{ref-Salaun2012}{}%
-Salaün, S., La Barre, S., Santos-Goncalvez, M.D., Potin, P., Haras, D.,
-and Bazire, A. (2012) Influence of exudates of the kelp \emph{Laminaria
+Salaün, S., La Barre, S., Santos-Goncalvez, M. D., Potin, P., Haras, D.,
+and Bazire, A. (2012). Influence of exudates of the kelp \emph{Laminaria
 digitata} on biofilm formation of associated and exogenous bacterial
-epiphytes. \emph{Microb Ecol} \textbf{64}: 359--369.
+epiphytes. \emph{Microb. Ecol.} 64, 359--369.
+doi:\href{https://doi.org/10.1007/s00248-012-0048-4}{10.1007/s00248-012-0048-4}.
 
 \leavevmode\hypertarget{ref-Schloss2016}{}%
-Schloss, P.D., Jenior, M.L., Koumpouras, C.C., Westcott, S.L., and
-Highlander, S.K. (2016) Sequencing 16S rRNA gene fragments using the
-PacBio SMRT DNA sequencing system. \emph{PeerJ} \textbf{4}: e1869.
+Schloss, P. D., Jenior, M. L., Koumpouras, C. C., Westcott, S. L., and
+Highlander, S. K. (2016). Sequencing 16S rRNA gene fragments using the
+PacBio SMRT DNA sequencing system. \emph{PeerJ} 4, e1869.
+doi:\href{https://doi.org/10.7717/peerj.1869}{10.7717/peerj.1869}.
 
 \leavevmode\hypertarget{ref-Schloss2009}{}%
-Schloss, P.D., Westcott, S.L., Ryabin, T., Hall, J.R., Hartmann, M.,
-Hollister, E.B., et al. (2009) Introducing mothur: Open-source,
+Schloss, P. D., Westcott, S. L., Ryabin, T., Hall, J. R., Hartmann, M.,
+Hollister, E. B., et al. (2009). Introducing mothur: Open-source,
 platform-independent, community-supported software for describing and
-comparing microbial communities. \emph{Appl Environ Microbiol}
-\textbf{75}: 7537--7541.
+comparing microbial communities. \emph{Appl. Environ. Microbiol.} 75,
+7537--7541.
+doi:\href{https://doi.org/10.1128/AEM.01541-09}{10.1128/AEM.01541-09}.
 
 \leavevmode\hypertarget{ref-Tarquinio2019}{}%
-Tarquinio, F., Hyndes, G.A., Laverock, B., Koenders, A., and Säwström,
-C. (2019) The seagrass holobiont: Understanding seagrass-bacteria
+Tarquinio, F., Hyndes, G. A., Laverock, B., Koenders, A., and Säwström,
+C. (2019). The seagrass holobiont: Understanding seagrass-bacteria
 interactions and their role in seagrass ecosystem functioning.
-\emph{FEMS Microbiol Lett} \textbf{366}: fnz057.
+\emph{FEMS Microbiol. Lett.} 366, fnz057.
+doi:\href{https://doi.org/10.1093/femsle/fnz057}{10.1093/femsle/fnz057}.
 
 \leavevmode\hypertarget{ref-Tujula2010}{}%
-Tujula, N.A., Crocetti, G.R., Burke, C., Thomas, T., Holmström, C., and
-Kjelleberg, S. (2010) Variability and abundance of the epiphytic
+Tujula, N. A., Crocetti, G. R., Burke, C., Thomas, T., Holmström, C.,
+and Kjelleberg, S. (2010). Variability and abundance of the epiphytic
 bacterial community associated with a green marine \emph{Ulvacean} alga.
-\emph{ISME J} \textbf{4}: 301--311.
+\emph{ISME J.} 4, 301--311.
+doi:\href{https://doi.org/10.1038/ismej.2009.107}{10.1038/ismej.2009.107}.
 
 \leavevmode\hypertarget{ref-Ugarelli2019}{}%
-Ugarelli, K., Laas, P., and Stingl, U. (2019) The microbial communities
+Ugarelli, K., Laas, P., and Stingl, U. (2019). The microbial communities
 of leaves and roots associated with turtle grass (\emph{Thalassia
@@
 distinct from seawater and sediment communities, but are similar between
-species and sampling sites. \emph{Microorganisms} \textbf{7}: 4.
+species and sampling sites. \emph{Microorganisms} 7, 4.
+doi:\href{https://doi.org/10.3390/microorganisms7010004}{10.3390/microorganisms7010004}.
 
 \leavevmode\hypertarget{ref-Uku2007}{}%
-Uku, J., Björk, M., Bergman, B., and Díez, B. (2007) Characterization
+Uku, J., Björk, M., Bergman, B., and Díez, B. (2007). Characterization
 and comparison of prokaryotic epiphytes associated with three East
-African seagrasses. \emph{J Phycol} \textbf{43}: 768--779.
+African seagrasses. \emph{J. Phycol.} 43, 768--779.
+doi:\href{https://doi.org/10.1111/j.1529-8817.2007.00371.x}{10.1111/j.1529-8817.2007.00371.x}.
 
 \leavevmode\hypertarget{ref-Weidner2000}{}%
-Weidner, S., Arnold, W., Stackebrandt, E., and Pühler, A. (2000)
+Weidner, S., Arnold, W., Stackebrandt, E., and Pühler, A. (2000).
 Phylogenetic analysis of bacterial communities associated with leaves of
 the seagrass \emph{Halophila stipulacea} by a culture-independent
-small-subunit rRNA gene approach. \emph{Microb Ecol} \textbf{39}:
-22--31.
+small-subunit rRNA gene approach. \emph{Microb. Ecol.} 39, 22--31.
 
 \leavevmode\hypertarget{ref-Wickham2019}{}%
-Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.D., François,
-R., et al. (2019) Welcome to the tidyverse. \emph{J Open Source Softw}
-\textbf{4}: 1686.
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
+François, R., et al. (2019). Welcome to the tidyverse. \emph{J. Open
+Source Softw.} 4, 1686.
+doi:\href{https://doi.org/10.21105/joss.01686}{10.21105/joss.01686}.
 
 \leavevmode\hypertarget{ref-Wilke2018}{}%
-Wilke, C.O. (2018) cowplot: Streamlined plot theme and plot annotations
-for 'ggplot2'.
+Wilke, C. O. (2018). \emph{cowplot: Streamlined plot theme and plot
+annotations for 'ggplot2'}.
+
+\leavevmode\hypertarget{ref-Xie2014}{}%
+Xie, Y. (2014). `` knitr: A comprehensive tool for reproducible research
+in R,'' in \emph{Implementing Reproducible Computational Research}, eds.
+V. Stodden, F. Leisch, and R. D. Peng (New York: Chapman and Hall/CRC),
+3--32.
 
 \leavevmode\hypertarget{ref-Xie2015}{}%
-Xie, Y. (2015) Dynamic Documents with R and knitr, 2nd ed. Boca Raton,
-Florida: Chapman and Hall/CRC.
-
-\leavevmode\hypertarget{ref-Xie2014}{}%
-Xie, Y. (2014) knitr: A comprehensive tool for reproducible research in
-R. In \emph{Implementing Reproducible Computational Research}. Stodden,
-V., Leisch, F., and Peng, R.D. (eds). New York: Chapman and Hall/CRC,
-pp. 3--32.
+Xie, Y. (2015). \emph{Dynamic Documents with R and knitr}. 2nd ed. Boca
+Raton, Florida: Chapman and Hall/CRC.
 
 \leavevmode\hypertarget{ref-Xie2019a}{}%
-Xie, Y. (2019) TinyTeX: A lightweight, cross-platform, and
+Xie, Y. (2019). TinyTeX: A lightweight, cross-platform, and
 easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
-\textbf{40}: 30--32.
-
-\leavevmode\hypertarget{ref-Xie2020}{}%
-Xie, Y. (2020) tinytex: Helper functions to install and maintain 'TeX
-Live', and compile 'LaTeX' documents.
+40, 30--32.
 
 \leavevmode\hypertarget{ref-Xie2018}{}%
-Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: The
-Definitive Guide, 1st ed. Boca Raton, Florida: Chapman and Hall/CRC.
+Xie, Y., Allaire, J. J., and Grolemund, G. (2018). \emph{R Markdown: The
+Definitive Guide}. 1st ed. Boca Raton, Florida: Chapman and Hall/CRC.
 
 \leavevmode\hypertarget{ref-Xie2017}{}%
-Xie, Z., Lin, W., and Luo, J. (2017) Comparative phenotype and genome
+Xie, Z., Lin, W., and Luo, J. (2017). Comparative phenotype and genome
 analysis of \emph{Cellvibrio} sp. PR1, a xylanolytic and agarolytic
-bacterium from the Pearl River. \emph{BioMed Res Int} \textbf{2017}:
-6304248.
+bacterium from the Pearl River. \emph{BioMed Res. Int.} 2017, 6304248.
+doi:\href{https://doi.org/10.1155/2017/6304248}{10.1155/2017/6304248}.
 
 \leavevmode\hypertarget{ref-Yilmaz2014}{}%
-Yilmaz, P., Parfrey, L.W., Yarza, P., Gerken, J., Pruesse, E., Quast,
-C., et al. (2014) The SILVA and "All-Species Living Tree Project (LTP)"
-taxonomic frameworks. \emph{Nucleic Acids Res} \textbf{42}: D643--D648.
+Yilmaz, P., Parfrey, L. W., Yarza, P., Gerken, J., Pruesse, E., Quast,
+C., et al. (2014). The SILVA and "All-Species Living Tree Project (LTP)"
+taxonomic frameworks. \emph{Nucleic Acids Res.} 42, D643--D648.
+doi:\href{https://doi.org/10.1093/nar/gkt1209}{10.1093/nar/gkt1209}.
 
 \leavevmode\hypertarget{ref-Zavodnik1998}{}%
-Zavodnik, N., Travizi, A., and de Rosa, S. (1998) Seasonal variations in
-the rate of photosynthetic activity and chemical composition of the
-seagrass \emph{Cymodocea nodosa} (Ucr.) Asch. \emph{Sci Mar}
-\textbf{62}: 301--309.
+Zavodnik, N., Travizi, A., and de Rosa, S. (1998). Seasonal variations
+in the rate of photosynthetic activity and chemical composition of the
+seagrass \emph{Cymodocea nodosa} (Ucr.) Asch. \emph{Sci. Mar.} 62,
+301--309.
+doi:\href{https://doi.org/10.3989/scimar.1998.62n4301}{10.3989/scimar.1998.62n4301}.
 
 \leavevmode\hypertarget{ref-Zhu2019}{}%
-Zhu, H. (2019) kableExtra: Construct complex table with 'kable' and pipe
-syntax.
+Zhu, H. (2019). \emph{kableExtra: Construct complex table with 'kable'
+and pipe syntax}.
 

### 2021-04-09 submission/manuscript.tex
Randomly rarefied data bug corrected.

@@
-\documentclass[12pt,]{article}
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+  12pt,
+]{article}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
   \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Seasonal Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=1.0in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
-            pdftitle={Seasonal Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
 \usepackage{graphicx,grffile}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{\textbf{Seasonal Dynamics of Epiphytic Microbial Communities on Marine
-Macrophyte Surfaces}}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-    \author{}
-    \preauthor{}\postauthor{}
-    \date{}
-    \predate{}\postdate{}
-
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
 \usepackage{times} % Times New Roman font
@@
 \usepackage{setspace}
+
+\title{\textbf{Seasonal Dynamics of Epiphytic Microbial Communities on Marine
+Macrophyte Surfaces}}
+\author{}
+\date{\vspace{-2.5em}}
 
@@
 (version 2.5-6) (Oksanen et al., 2019), tidyverse (version 1.3.0)
-(Wickham et al., 2019) and multiple other packages ({\textbf{???}};
-Neuwirth, 2014; Xie, 2014, 2015, 2019; Wilke, 2018; Xie et al., 2018;
-Allaire et al., 2019; Zhu, 2019). The detailed analysis procedure
-including the R Markdown file are available in the GitHub repository
+(Wickham et al., 2019) and multiple other packages (Neuwirth, 2014; Xie,
+2014, 2015, 2019a, 2019b; Wilke, 2018; Xie et al., 2018; Allaire et al.,
+2019; Zhu, 2019). The detailed analysis procedure including the R
+Markdown file are available in the GitHub repository
 (\url{https://github.com/MicrobesRovinj/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021}).
@@
 
-Sequencing of the 16S rRNA V4 region yielded a total of 1.8 million
+Sequencing of the 16S rRNA V4 region yielded a total of 1.7 million
 sequences after quality curation and exclusion of sequences without
@@
 were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,025
-different OTUs with 347 to 1,977 OTUs per sample
+\autoref{supp-nseq_notus}) through rarefaction resulting in 17,201
+different OTUs with 352 to 2,062 OTUs per sample
 (\autoref{supp-calculators}). To determine seasonal changes in richness
@@
 diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,683.5 ±
-128.5 OTUs{]} and monospecific {[}Number of OTUs, 1,729.9 ± 145.8
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,052.2 ± 211.4
+found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,688.4 ±
+136.6 OTUs{]} and monospecific {[}Number of OTUs, 1,750.4 ± 165.7
+OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,063.7 ± 210.6
 OTUs) and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 529.2 ± 146.8 OTUs)
+ambient seawater (Number of OTUs, 531.0 ± 143.9 OTUs)
 (\autoref{supp-calculators}). Seasonal changes did not reveal such large
@@
 present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (40.2 -- 52.1 \si{\percent}). To further
+proportion of total sequences (40.2 -- 53.2 \si{\percent}). To further
 disentangle the environmental and seasonal community dissimilarity a
@@
 separation between spring, summer and autumn/winter samples (ANOSIM, R =
-0.54, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
+0.56, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
 communities a separation between summer and autumn/winter/spring samples
 was observed that was, however, not as strong as for \emph{C. nodosa}
-associated communities (ANOSIM, R = 0.30, \emph{p} \textless{} 0.05)
+associated communities (ANOSIM, R = 0.30, \emph{p} \textless{} 0.01)
 (\autoref{pcoa}).
@@
 without known relatives within the \emph{Flavobacteriaceae} (no relative
-\emph{Flavobacteriaceae})and uncultured \emph{Rhodothermaceae} were
+\emph{Flavobacteriaceae}) and uncultured \emph{Rhodothermaceae} were
 found. Some groups showed minor seasonal changes such as no relative
@@
 \emph{Bacteroidota} different groups within the family
-\emph{Saprospiraceae} (e.g. \emph{Lewinella}, \emph{Portibacter} and
+\emph{Saprospiraceae} (e.g.~\emph{Lewinella}, \emph{Portibacter} and
 \emph{Rubidimonas}) were detected across all seasons. Members of this
@@
 
-This work was founded by the Croatian Science Foundation through the
+This work was funded by the Croatian Science Foundation through the
 MICRO-SEAGRASS project (project number IP-2016-06-7118). ZZ and GJH were
@@
 \leavevmode\hypertarget{ref-Armstrong2000}{}%
-Armstrong, E., Rogerson, A., and Leftley, J. (2000). The abundance of
+Armstrong, E., Rogerson, A., and Leftley, J. W. (2000). The abundance of
 heterotrophic protists associated with intertidal seaweeds.
@@
 \leavevmode\hypertarget{ref-Xie2019a}{}%
-Xie, Y. (2019). TinyTeX: A lightweight, cross-platform, and
+Xie, Y. (2019a). TinyTeX: A lightweight, cross-platform, and
 easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
 40, 30--32.
+
+\leavevmode\hypertarget{ref-Xie2019b}{}%
+Xie, Y. (2019b). \emph{tinytex: Helper functions to install and maintain
+'TeX Live', and compile 'LaTeX' documents}.
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{../results/figures/matrix}
+{\centering \includegraphics[width=0.7\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/matrix}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{../results/figures/seasonal_shared}
+{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/seasonal_shared}
 
@@
 
-{\centering \includegraphics[width=0.55\linewidth]{../results/figures/pcoa_figure}
+{\centering \includegraphics[width=0.55\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/pcoa_figure}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{../results/figures/community_bar_plot}
+{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/community_bar_plot}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{../results/figures/cyanobacteria_bar_plot}
+{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/cyanobacteria_bar_plot}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{../results/figures/bacteroidota_bar_plot}
+{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/bacteroidota_bar_plot}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{../results/figures/alphaproteobacteria_bar_plot}
+{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/alphaproteobacteria_bar_plot}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{../results/figures/gammaproteobacteria_bar_plot}
+{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/gammaproteobacteria_bar_plot}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{../results/figures/desulfobacterota_bar_plot}
+{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/desulfobacterota_bar_plot}
 

### 2021-04-09 submission/supplementary.Rmd
Coauthors review done.
Randomly rarefied data bug corrected.

@@
 
-\linenumbers
+[]: # (\linenumbers)
 \sisetup{mode=text}
@@
 cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\\label{rarefaction}", fig.pos="H"}
-include_graphics("../results/figures/rarefaction.jpg")
+include_graphics(normalizePath("results/figures/rarefaction.jpg"))
 ```
@@
 cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\\label{calculators}", fig.pos="H"}
-include_graphics("../results/figures/calculators.jpg")
+include_graphics(normalizePath("results/figures/calculators.jpg"))
 ```
@@
 cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\\label{chloroplast}", fig.pos="H"}
-include_graphics("../results/figures/chloroplast_bar_plot.jpg")
+include_graphics(normalizePath("results/figures/chloroplast_bar_plot.jpg"))
 ```

### 2021-04-09 submission/supplementary.tex
Coauthors review done.
Randomly rarefied data bug corrected.

@@
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
   \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Supplementary Information},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=1.0in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
-            pdftitle={Supplementary Information},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
 \usepackage{graphicx,grffile}
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{\textbf{Supplementary Information}}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-  \subtitle{\textbf{Seasonal Dynamics of Epiphytic Microbial Communities on Marine
-Macrophyte Surfaces}}
-  \author{}
-    \preauthor{}\postauthor{}
-    \date{}
-    \predate{}\postdate{}
-
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
 \usepackage{times} % Times New Roman font
@@
 
+\title{\textbf{Supplementary Information}}
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
+\subtitle{\textbf{Seasonal Dynamics of Epiphytic Microbial Communities on Marine
+Macrophyte Surfaces}}
+\author{}
+\date{\vspace{-2.5em}}
+
 \begin{document}
@@
 
-\linenumbers
 \sisetup{mode=text}
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{../results/figures/rarefaction}
+{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/rarefaction}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{../results/figures/calculators}
+{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/calculators}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{../results/figures/chloroplast_bar_plot}
+{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/chloroplast_bar_plot}
 
@@
 \endhead
-\
+
 \endfoot

### 2021-04-09 code/plot_matrix.R

@@ -2,22 +2,14 @@
 # code/plot_matrix.R
 # 
 # A script to plot a matrix of similarity indices.
-# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared
+# Dependencies: results/numerical/rarefied.Rdata
 #               data/raw/metadata.csv
 # Produces: results/figures/matrix.jpg
 #
 #################################################################################################################
 
-# Loading OTU/sample data
-shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")
-
-# Generation of random rarefied community data
-rarefied <- shared %>%
-  select(-label, -Group, -numOtus) %>%
-  rrarefy(., min(rowSums(.))) %>%
-  as_tibble(.name_repair="unique") %>%
-  add_column("Group"=shared$Group, .before=TRUE) %>%
-  select_if(list(~ !is.numeric(.) || sum(.)!=0))
+# Loading rarefied community data created during the generation of the richness and diversity calculators plot
+load(file="results/numerical/rarefied.Rdata")
 
 # Loading metadata 
 metadata <- read_tsv("data/raw/metadata.csv")
@@ -26,16 +18,13 @@ metadata <- read_tsv("data/raw/metadata.csv")
 rarefied_metadata <- inner_join(rarefied, metadata, by=c("Group"="ID")) %>%
   mutate(date=as.Date(date, "%d.%m.%Y")) %>%
   filter(date >= "2017-11-01") %>%
-  select_if(funs(!is.numeric(.) || sum(.)!=0)) %>%
+  select(!where(is.numeric) | where(~is.numeric(.x) && sum(.x)!=0)) %>%
   group_by(station) %>%
-  summarise_at(vars(starts_with("Otu")), sum)
+  summarise(across(starts_with("Otu"), sum), .groups="drop")
 
 # Copying the sample labels to the rows (input for library vegan)
-row.names(rarefied_metadata) <- rarefied_metadata$station
-
-# Removing column containing sample labels
 rarefied_metadata <- rarefied_metadata %>%
-  select(-station)
+  column_to_rownames("station")
 
 # Calculating matrices of similarity indices
 jaccard <- as.matrix(vegdist(rarefied_metadata, method="jaccard", binary=T))

### 2021-04-09 code/plot_pcoa.R

@@ -2,22 +2,14 @@
 # plot_pcoa.R
 # 
 # A script to generate the PCoA figure.
-# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared
+# Dependencies: results/numerical/rarefied.Rdata
 #               data/raw/metadata.csv
 # Produces: results/figures/pcoa_figure.jpg
 #
 #################################################################################################################
 
-# Loading OTU/sample data
-shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")
-
-# Generation of random rarefied community data
-rarefied <- shared %>%
-  select(-label, -Group, -numOtus) %>%
-  rrarefy(., min(rowSums(.))) %>%
-  as_tibble(.name_repair="unique") %>%
-  add_column("Group"=shared$Group, .before=TRUE) %>%
-  select_if(list(~!is.numeric(.) || sum(.)!=0))
+# Loading rarefied community data created during the generation of the richness and diversity calculators plot
+load(file="results/numerical/rarefied.Rdata")
 
 # Loading metadata 
 metadata <- read_tsv("data/raw/metadata.csv")
@@ -56,7 +48,7 @@ coordinates <- inner_join(metadata, coordinates, by=c("label"="Group"))
 p1 <- ggplot() +
   geom_point(data=coordinates, aes(x=A1, y=A2, fill=station), shape=21, size=3, stroke=0.5) +
   scale_fill_manual(name=NULL,
-                      values=c("#66A61E", "#E6AB02", "#A6761D", "#666666"),
+                      values=c("#66A61E", "#666666", "#A6761D", "#E6AB02"),
                       breaks=c("F", "FCyM", "FCaM", "FCa"),
                       labels=c("Seawater",
                                parse(text="italic('Cymodocea nodosa')~'(Mixed)'"),
@@ -71,7 +63,7 @@ p1 <- ggplot() +
 
 # Cymodocea nodosa samples
 rarefied_metadata_select <- filter(rarefied_metadata, station=="FCyM") %>%
-  select_if(list(~!is.numeric(.) || sum(.)!=0))
+  select(!where(is.numeric) | where(~ is.numeric(.x) && sum(.x)!=0))
 
[…]

### 2021-04-09 code/plot_rarefaction.R

@@ -22,7 +22,8 @@ metadata <- read_tsv("data/raw/metadata.csv") %>%
   arrange(date) %>%
   mutate(date=format(date, "%d %B %Y")) %>%
   mutate(date=str_replace(date, "^0", "")) %>%
-  mutate(date=if_else(label=="24/4-18 F-2", paste0(date, "a"), date))
+  mutate(date=if_else(label=="24/4-18 F-1", paste0(date, "a"), date)) %>%
+  mutate(date=if_else(label=="24/4-18 F-2", paste0(date, "b"), date))
 
 # Joining metadata and input data
 metadata_rarefaction <- inner_join(metadata, rarefaction, by=c("ID"="sample"))
@@ -42,12 +43,13 @@ color_type <- tribble(~date, ~color, ~line_type,
                       "27 March 2018", "#A6CEE3", "dotted",
                       "24 April 2018", "#1F78B4", "dotted",
                       "24 April 2018a", "#B2DF8A", "dotted",
-                      "22 May 2018", "#33A02C", "dotted",
-                      "19 June 2018", "#FB9A99", "dotted",
-                      "10 July 2018", "#E31A1C", "dotted",
-                      "9 August 2018", "#FDBF6F", "dotted",
-                      "4 September 2018", "#FF7F00", "dotted",
-                      "5 October 2018", "#CAB2D6", "dotted")
+                      "24 April 2018b", "#33A02C", "dotted",
+                      "22 May 2018", "#FB9A99", "dotted",
+                      "19 June 2018", "#E31A1C", "dotted",
+                      "10 July 2018", "#FDBF6F", "dotted",
+                      "9 August 2018", "#FF7F00", "dotted",
+                      "4 September 2018", "#CAB2D6", "dotted",
+                      "5 October 2018", "#6A3D9A", "dotted")
 
 # Generating a common theme for plots
 theme <- theme(text=element_text(family="Times"), line=element_line(color="black"),
@@ -102,7 +104,7 @@ p4 <- filter(metadata_rarefaction, str_detect(label, "FCa$")) %>%
 
 # Generating a plot with all samples to extract a common legend
 p <- ggplot(metadata_rarefaction, aes(x=numsampled, y=sobs, group=ID,
-            color=factor(date, levels=unique(date)), linetype=factor(date, levels=unique(date)))) +
+            color=factor(date, levels=color_type$date), linetype=factor(date, levels=color_type$date))) +
   geom_line(size=1.5) +
[…]

### 2021-04-09 submission/manuscript.Rmd

@@
 		if(abs(x - round(x)) < .Machine$double.eps^0.5){
-			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
+			paste(format(x, big.mark=',', digits=0, scientific=FALSE))
 		} else {
-			paste(format(x,big.mark=',', digits=1, nsmall=1, scientific=FALSE))
+			paste(format(x, big.mark=',', digits=1, nsmall=1, scientific=FALSE))
 		}
@@
 For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
-Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019] combined with  packages  vegan (version 2.5-6) [@Oksanen2019], tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Wilke2018].
+Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019] combined with  packages  vegan (version 2.5-6) [@Oksanen2019], tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019b; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Wilke2018].
 The detailed analysis procedure including the R Markdown file are available in the GitHub repository (https://github.com/MicrobesRovinj/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021).
@@
 million_sequences <- filter(community, taxon=="Archaea" | taxon=="Bacteria") %>%
-  select(total) %>%
+  select(6:ncol(.)) %>%
   sum()/1e6
@@
   mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
-  gather(key="Group", value="abundance", 6:(ncol(.))) %>%
+  gather(key="Group", value="abundance", 6:ncol(.)) %>%
   group_by(taxon) %>%
@@
 
-# Generation of random rarefied community data
-rarefied <- shared %>%
-  select(-label, -Group, -numOtus) %>%
-  rrarefy(., min(rowSums(.))) %>%
-  as_tibble(.name_repair="unique") %>%
-  add_column("Group"=shared$Group, .before=TRUE) %>%
-  select_if(list(~ !is.numeric(.) || sum(.)!=0))
+# Loading rarefied community data created during the generation of the richness and diversity calculators plot
+load(file="results/numerical/rarefied.Rdata")
 
@@
 ```{r calculators}
+# Loading rarefied community data created during the generation of the richness and diversity calculators plot
+load(file="results/numerical/rarefied.Rdata")
+
 # Copying the sample labels to the rows (input for library vegan)
-row.names(rarefied) <- rarefied$Group
-
-# Removing column containing sample labels
 rarefied <- rarefied %>%
-  select(-Group)
+  column_to_rownames("Group")
 
@@
 cylindracea* (mixed [Number of OTUs, `r S.obs_fcam$mean` ± `r S.obs_fcam$sd` OTUs] and monospecific [Number of OTUs, `r S.obs_fca$mean` ± `r S.obs_fca$sd` OTUs]) than for *C.
-nodosa* (Number of OTUs, `r S.obs_fcym$mean` ± `r S.obs_fcym$sd` OTUs) and lowest values were obtained for the microbial community of the ambient seawater (Number of OTUs, `r S.obs_f$mean` ± `r S.obs_f$sd` OTUs) (\autoref{supp-calculators}).
+nodosa* (Number of OTUs, `r S.obs_fcym$mean` ± `r S.obs_fcym$sd` OTUs) and lowest values were obtained for the microbial community of the ambient seawater (Number of OTUs, `r format(S.obs_f$mean, digits=1, nsmall=1)` ± `r S.obs_f$sd` OTUs) (\autoref{supp-calculators}).
 Seasonal changes did not reveal such large dissimilarities.
@@
 ```{r seasonal_shared}
-# Generation of random rarefied community data
-rarefied <- shared %>%
-  select(-label, -Group, -numOtus) %>%
-  rrarefy(., min(rowSums(.))) %>%
-  as_tibble(.name_repair="unique") %>%
-  add_column("Group"=shared$Group, .before=TRUE) %>%
-  select_if(list(~ !is.numeric(.) || sum(.)!=0))
+# Loading rarefied community data created during the generation of the richness and diversity calculators plot
+load(file="results/numerical/rarefied.Rdata")
 
@@
   filter(date >= "2017-11-01") %>%
-  select_if(list(~ !is.numeric(.) || sum(.)!=0)) %>%
+  select(!where(is.numeric) | where(~is.numeric(.x) && sum(.x)!=0)) %>%
   group_by(station) %>%
-  summarise_at(vars(starts_with("Otu")), sum)
+  summarise(across(starts_with("Otu"), sum), .groups="drop")
 
 # Copying the sample labels to the rows (input for library vegan)
-row.names(rarefied_metadata) <- rarefied_metadata$station
-
-# Removing column containing sample labels
 rarefied_metadata <- rarefied_metadata %>%
-  select(-station)
+  column_to_rownames("station")
 
@@
 ```{r core_otus}
+# Loading rarefied community data created during the generation of the richness and diversity calculators plot
+load(file="results/numerical/rarefied.Rdata")
+
 # Joining metadata with OTU/sample data
-metadata_rarefied <- inner_join(metadata, rarefied, by=c("ID"="Group")) %>%
+metadata_rarefied <- read_tsv("data/raw/metadata.csv") %>%
+  inner_join(., rarefied, by=c("ID"="Group")) %>%
   mutate(date=as.Date(date, "%d.%m.%Y"))
@@
 ```{r anosim}
-# Generation of random rarefied community data and preserving sample names as row names
-rarefied <- shared %>%
-  select(-label, -Group, -numOtus) %>%
-  rrarefy(., min(rowSums(.))) %>%
-  as_tibble(.name_repair="unique") %>%
-  select_if(list(~!is.numeric(.) || sum(.)!=0))
-rownames(rarefied) <- shared$Group
-rarefied <- rarefied %>%
-  add_column("Group"=shared$Group, .before=TRUE)
+# Loading rarefied community data created during the generation of the richness and diversity calculators plot
+load(file="results/numerical/rarefied.Rdata")
 
@@
 The ambient seawater community was characterized by the NS4 and NS5 marine groups, uncultured *Cryomorphaceae*, uncultured *Flavobacteriaceae*, NS11-12 marine group, *Balneola*, uncultured *Balneolaceae* and *Formosa*.
-In contrast, in surface associated communities *Lewinella*, *Portibacter*, *Rubidimonas*, sequences without known relatives within the *Saprospiraceae* (no relative *Saprospiraceae*), uncultured *Saprospiraceae*, (sequences without known relatives within the *Flavobacteriaceae* (no relative *Flavobacteriaceae*)and uncultured *Rhodothermaceae* were found.
+In contrast, in surface associated communities *Lewinella*, *Portibacter*, *Rubidimonas*, sequences without known relatives within the *Saprospiraceae* (no relative *Saprospiraceae*), uncultured *Saprospiraceae*, (sequences without known relatives within the *Flavobacteriaceae* (no relative *Flavobacteriaceae*) and uncultured *Rhodothermaceae* were found.
 Some groups showed minor seasonal changes such as no relative *Flavobacteriaceae* whose sequences were more abundant from November 2017 until June 2018.
@@
 ## Acknowledgments
-This work was founded by the Croatian Science Foundation through the MICRO-SEAGRASS project (project number IP-2016-06-7118).
+This work was funded by the Croatian Science Foundation through the MICRO-SEAGRASS project (project number IP-2016-06-7118).
 ZZ and GJH were supported by the Austrian Science Fund (FWF) project ARTEMIS (project number P28781-B21).
@@
 cylindracea} (mixed and monospecific settlement) and communities in the ambient seawater.\\label{matrix}", fig.pos="H"}
-include_graphics("../results/figures/matrix.jpg")
+include_graphics(normalizePath("results/figures/matrix.jpg"))
 ```
@@
 cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\\label{shared}", fig.pos="H"}
-include_graphics("../results/figures/seasonal_shared.jpg")
+include_graphics(normalizePath("results/figures/seasonal_shared.jpg"))
 ```
@@
 The proportion of explained variation by each axis is shown on the corresponding axis in parentheses.\\label{pcoa}", fig.pos="H"}
-include_graphics("../results/figures/pcoa_figure.jpg")
+include_graphics(normalizePath("results/figures/pcoa_figure.jpg"))
 ```
@@
 NR -- No Relative (sequences without known relatives within the corresponding group)\\label{community}", fig.pos="H"}
-include_graphics("../results/figures/community_bar_plot.jpg")
+include_graphics(normalizePath("results/figures/community_bar_plot.jpg"))
 ```
@@
 NR -- No Relative (sequences without known relatives within the corresponding group)\\label{cyano}", fig.pos="H"}
-include_graphics("../results/figures/cyanobacteria_bar_plot.jpg")
+include_graphics(normalizePath("results/figures/cyanobacteria_bar_plot.jpg"))
 ```
@@
 NR -- No Relative (sequences without known relatives within the corresponding group)\\label{bactero}", fig.pos="H"}
-include_graphics("../results/figures/bacteroidota_bar_plot.jpg")
+include_graphics(normalizePath("results/figures/bacteroidota_bar_plot.jpg"))
 ```
@@
 NR -- No Relative (sequences without known relatives within the corresponding group)\\label{alpha}", fig.pos="H"}
-include_graphics("../results/figures/alphaproteobacteria_bar_plot.jpg")
+include_graphics(normalizePath("results/figures/alphaproteobacteria_bar_plot.jpg"))
 ```
@@
 NR -- No Relative (sequences without known relatives within the corresponding group)\\label{gamma}", fig.pos="H"}
-include_graphics("../results/figures/gammaproteobacteria_bar_plot.jpg")
+include_graphics(normalizePath("results/figures/gammaproteobacteria_bar_plot.jpg"))
 
@@
 NR -- No Relative (sequences without known relatives within the corresponding group)\\label{desulfo}", fig.pos="H"}
-include_graphics("../results/figures/desulfobacterota_bar_plot.jpg")
-```
+include_graphics(normalizePath("results/figures/desulfobacterota_bar_plot.jpg"))
+```

### 2021-07-09 results/numerical/calculator_statistics.tex
Review done.

\begingroup\fontsize{9}{11}\selectfont

\begin{longtable}[t]{>{\centering\arraybackslash}m{6em}ccccc}
\caption{\label{tab:calculator_statistics}Statistic of richness estimator and diversity index parameters for different community types. Parameters were tested by applying the Kruskal-Wallis \textit{H} test followed by a pairwise comparison using the Mann-Whitney \textit{U} test. Bonferroni correction was used to address the problem of multiple comparisons.\label{calculator_statistics}}\\
\toprule
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{3}{c}{\textbf{Kruskal-Wallis \textit{H} test}} & \multicolumn{2}{c}{\textbf{Mann-Whitney \textit{U} test}} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-6}
\textbf{Parameter} & \textbf{\textit{H}} & \textbf{\textit{df}} & \textbf{\textit{p}} & \textbf{Comparisons Between Community Types} & \textbf{\textit{p}}\\
\midrule
\endfirsthead
\caption[]{Statistic of richness estimator and diversity index parameters for different community types. Parameters were tested by applying the Kruskal-Wallis \textit{H} test followed by a pairwise comparison using the Mann-Whitney \textit{U} test. Bonferroni correction was used to address the problem of multiple comparisons.\label{calculator_statistics} \textit{(continued)}}\\
\toprule
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{3}{c}{\textbf{Kruskal-Wallis \textit{H} test}} & \multicolumn{2}{c}{\textbf{Mann-Whitney \textit{U} test}} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-6}
\textbf{Parameter} & \textbf{\textit{H}} & \textbf{\textit{df}} & \textbf{\textit{p}} & \textbf{Comparisons Between Community Types} & \textbf{\textit{p}}\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
 &  &  &  & Seawater – \textit{Cymodocea nodosa} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Monospecific) & < 0.0001\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Mixed) & < 0.001\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & < 0.001\\
\nopagebreak
\multirow{-6}{6em}{\centering\arraybackslash Observed No. of OTUs} & \multirow{-6}{*}{\centering\arraybackslash 44.8} & \multirow{-6}{*}{\centering\arraybackslash 3} & \multirow{-6}{*}{\centering\arraybackslash < 0.0001} & \textit{Caulerpa cylindracea} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & 1.00\\
\cmidrule{1-6}\pagebreak[0]
 &  &  &  & Seawater – \textit{Cymodocea nodosa} (Mixed) & < 0.01\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Monospecific) & < 0.0001\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & < 0.001\\
\nopagebreak
\multirow{-6}{6em}{\centering\arraybackslash Chao1} & \multirow{-6}{*}{\centering\arraybackslash 38.3} & \multirow{-6}{*}{\centering\arraybackslash 3} & \multirow{-6}{*}{\centering\arraybackslash < 0.0001} & \textit{Caulerpa cylindracea} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & 1.00\\
\cmidrule{1-6}\pagebreak[0]
 &  &  &  & Seawater – \textit{Cymodocea nodosa} (Mixed) & < 0.05\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Monospecific) & < 0.0001\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & < 0.001\\
\nopagebreak
\multirow{-6}{6em}{\centering\arraybackslash ACE} & \multirow{-6}{*}{\centering\arraybackslash 37.1} & \multirow{-6}{*}{\centering\arraybackslash 3} & \multirow{-6}{*}{\centering\arraybackslash < 0.0001} & \textit{Caulerpa cylindracea} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & 1.00\\
\cmidrule{1-6}\pagebreak[0]
 &  &  &  & Seawater – \textit{Cymodocea nodosa} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Monospecific) & < 0.0001\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Mixed) & < 0.01\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & < 0.0001\\
\nopagebreak
\multirow{-6}{6em}{\centering\arraybackslash Exponential Shannon} & \multirow{-6}{*}{\centering\arraybackslash 43.3} & \multirow{-6}{*}{\centering\arraybackslash 3} & \multirow{-6}{*}{\centering\arraybackslash < 0.0001} & \textit{Caulerpa cylindracea} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & 0.26\\
\cmidrule{1-6}\pagebreak[0]
 &  &  &  & Seawater – \textit{Cymodocea nodosa} (Mixed) & < 0.01\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Monospecific) & < 0.0001\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Mixed) & 1.00\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & < 0.01\\
\nopagebreak
\multirow{-6}{6em}{\centering\arraybackslash Inverse Simpson} & \multirow{-6}{*}{\centering\arraybackslash 34.8} & \multirow{-6}{*}{\centering\arraybackslash 3} & \multirow{-6}{*}{\centering\arraybackslash < 0.0001} & \textit{Caulerpa cylindracea} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & 0.26\\*
\end{longtable}
\endgroup{}

### 2021-07-09 submission/manuscript.Rmd
Review done.

@@
   pdf_document: 
-    keep_tex: true
+    keep_tex: yes
     includes:
@@
 
-3\. NIOZ, Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research, Utrecht University, The Netherlands
+3\. Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research (NIOZ), Utrecht University, The Netherlands
 
@@
 Also, successional changes were observed connected to the macrophyte growth cycle.
-Taxonomic analysis, however, showed similar high rank groups in the ambient water and the epiphytic communities, with the exception of *Desulfobacterota*, which were only found on *Caulerpa cylindracea*.
+Taxonomic analysis, however, showed similar high rank taxa (phyla and classes) in the ambient water and the epiphytic communities, with the exception of *Desulfobacterota*, which were only found on *Caulerpa cylindracea*.
 *Cyanobacteria* showed seasonal changes while other high rank taxa were present throughout the year.
-Phylogenetic groups present throughout the year constituted most of the sequences, while less abundant taxa showed seasonal patterns connected to the macrophyte growth cycle.
+In months of high *Cyanobacteria* presence the majority of cyanobacterial sequences were classified as *Pleurocapsa*.
+Phylogenetic groups present throughout the year (e.g.
+*Saprospiraceae*, *Rhodobacteraceae*, members without known relatives within *Gammaproteobacteria*, *Desulfatitalea* and members without known relatives within *Desulfocapsaceae*) constituted most of the sequences, while less abundant taxa showed seasonal patterns connected to the macrophyte growth cycle.
 Taken together, epiphytic microbial communities of the seagrass *Cymodocea nodosa* and the macroalga *Caulerpa cylindracea* appear to be host-specific and contain taxa that undergo successional changes.
@@
 Marine macrophytes (seagrasses and macroalgae) are important ecosystem engineers forming close associations with microorganisms belonging to all three domains of life [@Tarquinio2019; @Egan2013].
-Microbes can live within macrophyte tissue as endophytes or can form epiphytic communities on surfaces of leaves and thalli [@Tarquinio2019; @Egan2013; @Hollants2013].
+Microbes can live within macrophyte tissue as endophytes or form epiphytic communities on surfaces of leaves, roots, rhizomes and thalli [@Tarquinio2019; @Egan2013; @Hollants2013; @Aires2015].
 Epiphytic and endophytic microbial communities exhibit a close functional relationship with the macrophyte host.
-It was proposed that this close relationship constitutes a holobiont, an integrated community where the macrophyte organism and its symbiotic partners support each other [@Margulis1991; @Tarquinio2019; @Egan2013].
-
-Biofilms of microbial epiphytes can contain diverse taxonomic groups and harbor cell abundances from 10^2^ to 10^7^ \si{\cells\per\cm\squared} [@Armstrong2000; @Bengtsson2010; @Burke2011a].
+It has been proposed that this close relationship constitutes a holobiont, an integrated community where the macrophyte organism and its symbiotic partners support each other [@Margulis1991].
+In addition, as suggested by the hologenome theory endophytic microbes play a critical role in the adaptation and evolution of the host species [@Aires2015].
+
+Biofilms of microbial epiphytes can contain diverse taxonomic groups and harbour cell abundances from 10^2^ to 10^7^ \si{\cells\per\cm\squared} [@Armstrong2000; @Bengtsson2010; @Burke2011a].
 In such an environment a number of positive and negative interactions between the macrophyte and the colonizing microorganisms have been described [@Tarquinio2019; @Egan2013; @Hollants2013].
-Macrophytes can promote growth of associated microbes by nutrient exudation, while in return microorganisms may support macrophyte performance through improved nutrient availability, phytohormone production and protection from toxic compounds, oxidative stress, biofouling organisms and pathogens [@Tarquinio2019; @Egan2013; @Hollants2013].
-Besides these positive interactions, macrophytes can  negatively impact the associated microbes such as pathogenic bacteria by producing reactive oxygen species and secondary metabolites [@Tarquinio2019; @Egan2013; @Hollants2013].
+Macrophytes can promote growth of associated microbes by nutrient exudation [@Wood1981], while in return microorganisms may support macrophyte performance through improved nutrient availability [@Nielsen2001; @deOliveira2012], phytohormone production [@Celdran2012; @Matsuo2003] and protection from toxic compounds [@Kusel2006], oxidative stress [@Sanchez-Amat2010], biofouling organisms [@Dobretsov2002] and pathogens [@Penesyan2009].
+Besides these positive interactions, macrophytes can  negatively impact the associated microbes by producing reactive oxygen species [@Weinberger2007] and secondary metabolites [@Saha2011].
 
 All these ecological roles are carried out by a taxonomically diverse community of microorganisms.
-At higher taxonomic ranks a core set of macrophyte epiphytic taxa was described consisting mainly of *Alphaproteobacteria*, *Gammaproteobacteria*, *Desulfobacterota*, *Bacteroidota*, *Cyanobacteria*, *Actinobacteriota*, *Firmicutes*, *Planctomycetota*, *Chloroflexi* and *Verrucomicrobiota* [@Lachnit2011; @Crump2008; @Tujula2010].
-In contrast, at lower taxonomic ranks host specific microbial communities were found [@Roth-Schulze2016; @Lachnit2011].
-Recently, it was shown that even different morphological niches within the same alga had a higher influence on the composition of the bacterial community than biogeography or environmental factors [@Morrissey2019].
-While the microbial community composition varies between host species, metagenomic analyses revealed that the majority of the microbial functions are conserved [@Burke2011; @Roth-Schulze2016].
-This discrepancy between the microbial taxonomic and functional composition might be explained by the lottery hypothesis.
-It postulates that an initial random colonization step takes places from a set of functionally equivalent taxonomic groups resulting in taxonomically different epiphytic communities sharing a core set of functional genes [@Burke2011; @Roth-Schulze2016].
-In addition, some of the variation in the reported data could be attributed to different techniques used in these studies, such as different protocols for epiphytic cell detachment and/or DNA isolation, as no standard protocol has been yet established to study epiphytic communities [@Korlevic, submitted; @Ugarelli2019].
-
-The majority of studies describing macrophyte epiphytic microbial communities did not include possible seasonal changes [@Ugarelli2019; @Lachnit2009; @Crump2008; @Burke2011a; @Roth-Schulze2016].
-If seasonal changes were taken into account, low temporal frequency, applied methodologies and/or limited number of analysed host species did not allow a high taxonomic resolution [@Tujula2010; @Lachnit2011; @Miranda2013; @Michelou2013; @Mancuso2016].
-In the present study we describe the seasonal dynamics of bacterial and archaeal communities on the surfaces of the seagrass *Cymodocea nodosa* and siphonous macroalgae *Caulerpa cylindracea* determined on a mostly monthly scale.
+At higher taxonomic ranks (phyla and classes) microbial taxa, such as *Alphaproteobacteria*, *Gammaproteobacteria*, *Bacteroidota* and *Cyanobacteria*, have been associated with surfaces of seagrass leaves and macroalgal thalli [@Lachnit2011; @Crump2008; @Tujula2010; @Egan2013; @Tarquinio2019; @Ugarelli2019].
+While similar high rank taxa have been found on surfaces of different macrophyte species, in order to describe new ecological patterns it is also necessary to focus on lower taxonomic ranks (genus and OTUs) which tend to be host-specific [@Roth-Schulze2016; @Lachnit2011; @Hollants2013].
+While the microbial community composition can vary between host species, metagenomic analyses revealed that the majority of microbial functions are conserved, showing that different epiphytic microbial species could be functionally similar [@Burke2011; @Roth-Schulze2016; @Cucio2018].
+This discrepancy between the microbial taxonomic and functional composition might be explained by the lottery hypothesis [@Sale1976].
+It postulates that an initial random colonization step takes place from a set of functionally equivalent taxonomic groups resulting in taxonomically different epiphytic communities sharing a core set of functional genes [@Burke2011; @Roth-Schulze2016; @Schmidt2015a; @Stratil2013].
+
+Seagrasses are know to form close relationships with microbial communities associated with the surfaces of leaves, roots and rhizomes [@Cucio2016; @Ugarelli2019; @Wang2020; @Ettinger2020; @Crump2018].
+For different seagrass species a distinct microbial community from ambient seawater or bulk sediment has been reported, however no species specific communities have been found [@Cucio2016; @Ugarelli2019; @Crump2018].
+It seems that seagrasses are selecting the associated microbial community but these microbes have not coevolved with their seagrass plant host.
+Similar to seagrasses, siphonous macroalgae of the genus *Caulerpa* are also closely associated with their microbial communities [@Aires2015; @Aires2013; @Morrissey2019; @Rizzo2016a; @Stabili2017].
+While some studies have found similar culturable bacterial groups associated with the surface of a *Caulerpa* species from different geographic locations [@Stabili2017], other have reported large compositional differences that were mainly attributed to different host species of this genus, biogeography and nutrient levels [@Morrissey2019] again raising the question to which extent are associated communities host-specific.
+
+Since marine macrophytes in temperate zones are exhibiting seasonal changes in growth and physiology [@Agostini2003; @Najdek2020a] it is important to verify if and how surface associated microbial communities are affected by these changes.
+The majority of studies describing macrophyte epiphytic microbial communities have not included possible seasonal changes [@Ugarelli2019; @Lachnit2009; @Crump2008; @Burke2011a; @Roth-Schulze2016].
+If seasonal changes have been taken into account, low temporal frequency, applied methodologies and/or limited number of analysed host species did not allow for a detailed taxonomic analysis [@Tujula2010; @Lachnit2011; @Miranda2013; @Michelou2013; @Mancuso2016; @Bengtsson2012].
+In the present study we performed a descriptive analysis of seasonal bacterial and archaeal community dynamics on the surfaces of the seagrass *Cymodocea nodosa*, an abundant seagrass species in the Mediterranean [@Short2001], and siphonous macroalga *Caulerpa cylindracea*, one of the most invasive macroalgal species [@Klein2008; @Boudouresque2009].
 Bacterial and archaeal epiphytes were sampled in a meadow of *C.
@@
 For comparison, the microbial community of the ambient seawater was also characterized.
+The presence of both macrophytes in the same area enabled (i) the assessment of differences in the bacterial and archaeal communities between host species and settlements of *C.
+cylindracea* and (ii) the evaluation of differences between surface associated and free living (ambient seawater) communities.
+In addition, these differences were evaluated on a monthly scale providing insight into seasonal changes (iii).
 
@@
 Sampling was performed in the Bay of Funtana, northern Adriatic Sea (\ang{45;10;39} N, \ang{13;35;42} E).
-Leaves and thalli were sampled in a meadow of *Cymodocea nodosa* invaded by the invasive *Caulerpa cylindracea* (mixed settlement) and in a monospecific settlement of *Caulerpa cylindracea* located in the proximity of the meadow at approximately monthly intervals from December 2017 to October 2018 (\autoref{supp-nseq_notus}).
+The sea floor in the bay is partly covered by the invasive macroalga *C.
+cylindracea* that can be found in a monospecific settlement or mixed with the seagrass *C.
+nodosa* (\autoref{map}).
+*C.
+nodosa* leaves were retrieved from a meadow of *C.
+nodosa* invaded by the invasive *C.
+cylindracea* (mixed settlement; depth, 2 -- 2.5 \si{\m}), while *C.
+cylindracea* thalli were sampled in the same invaded meadow (mixed settlement; depth, 2 -- 2.5 \si{\m}) and in a monospecific settlement (depth, 1 -- 1.5 \si{\m}) of *C.
+cylindracea* located in the proximity (20 -- 50 \si{\m}) of the invaded meadow at approximately monthly intervals from November 2017 to October 2018 (\autoref{supp-nseq_notus}).
 Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with seawater collected at the sampling site.
@@
 nodosa* and *C.
-cylindracea* was isolated using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevic, submitted; @Massana1997].
-Briefly, leaves and thalli are incubated in a lysis buffer and treated with lysozyme and proteinase K.
+cylindracea* was isolated from a pool of leaves or thalli on the sampling day using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevic2021; @Massana1997].
+Briefly, leaves and thalli were incubated in a lysis buffer and treated with lysozyme and proteinase K.
 Following the incubations, the mixture containing lysed epiphytic cells was separated from the leaves and thalli and extracted using phenol-chloroform.
@@
 DNA from seawater picoplankton was extracted from 0.2 \si{\um} polycarbonate filters according to @Massana1997 with a slight modification.
-Following the phenol-chloroform extraction, 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added.
-DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 20,000 × *g* and 4 \si{\degreeCelsius} for 21 \si{\minute}.
+Following the phenol-chloroform extraction, 1/10 of 3 \si{\Molar} sodium acetate (pH 5.2) was added.
+DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 20,000 × *g* and 4 \si{\degreeCelsius} for 20 \si{\minute}.
 The pellet was washed twice with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged after each washing step at 20,000 × *g* and 4 \si{\degreeCelsius} for 5 \si{\minute}.
@@
 ### Illumina 16S rRNA sequencing
-Illumina MiSeq sequencing of the V4 region of the 16S rRNA gene was performed as described previously [@Korlevic, submitted].
+Illumina MiSeq sequencing of the V4 region of the 16S rRNA gene was performed as described previously [@Korlevic2021].
 The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
@@
 The negative control comprised PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
-Sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession number PRJEB37267.
-
-### Sequence analysis
-```{r error}
-# Calculation of the error rate based on the ATCC MSA-1002 mock community
-error <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.error.summary", na="") %>%
-  filter(numparents==1) %>%
-  mutate(mismatches_weight=weight*mismatches) %>%
-  mutate(total_weight=weight*total)
-error <- sum(error$mismatches_weight)/sum(error$total_weight)*100
-
-# Calculation of the number of sequences in the negative control
-nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
-  filter(taxon=="Root") %>%
-  select(NC_1)
-```
-
-Obtained sequences were analysed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations provided by the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
-For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
-Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019] combined with  packages  vegan (version 2.5-6) [@Oksanen2019], tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019b; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Wilke2018].
-The detailed analysis procedure including the R Markdown file are available in the GitHub repository (https://github.com/MicrobesRovinj/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021).
-Based on the ATCC MSA-1002 mock community included in the analysis an average sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
-In addition, the negative controls processed together with the samples yielded on average only `r nc` sequences after sequence quality curation.
-
-\newpage
-## Results
-
+Sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under the accession number PRJEB37267 (https://www.ebi.ac.uk/ena/browser/view/PRJEB37267).
+
+### Sequence and data analysis
 ```{r community}
@@
 ```
+
+```{r error}
+# Calculation of the error rate based on the ATCC MSA-1002 mock community
+error <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.error.summary", na="") %>%
+  filter(numparents==1) %>%
+  mutate(mismatches_weight=weight*mismatches) %>%
+  mutate(total_weight=weight*total)
+error <- sum(error$mismatches_weight)/sum(error$total_weight)*100
+
+# Calculation of the number of sequences in the negative control
+nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
+  filter(taxon=="Root") %>%
+  select(NC_1)
+```
+
+Obtained sequences were analysed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations provided by the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
+For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
+Sequences were clustered into operational taxonomic units (OTUs) at a similarity level of 97 \si{\percent}.
+
+Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019] combined with packages vegan (version 2.5-6) [@Oksanen2019], tidyverse (version 1.3.0) [@Wickham2017; @Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019; @Xie2019a; @Xie2019b; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Wilke2019].
+Observed number of OTUs, Chao1, ACE, exponential of the Shannon diversity index and Inverse Simpson index were calculated after normalization to the minimum number of reads per sample using vegan's function `rrarefy` to account for different sequencing depths [@Oksanen2019].
+Chao1 and ACE estimators were calculated using vegan's function `estimateR`, while Shannon and Inverse Simpson diversity indices were retrieved using vegan's function `diversity` [@Oksanen2019].
+To express both diversity indices in terms of effective number of OTUs the exponential of the Shannon diversity index was calculated [@Jost2006].
+The proportions of shared OTUs and communities between samples and community types (seawater, *C.
+nodosa* [mixed], *C.
+cylindracea* [mixed] and *C.
+cylindracea* [monospecific]) were expressed as the Jaccard's (on presence/absence data) and Bray-Curtis similarity coefficient, respectively.
+The coefficients were calculated on the OTU data table using vegan's function `vegdist` and converted from dissimilarities to similarities [@Oksanen2019; @Borcard2011; @Legendre2012].
+The Principal Coordinates Analysis (PCoA) was performed on Bray-Curtis dissimilarities based on OTU abundances using the function `cmdscale` [@Borcard2011; @Legendre2012].
+Differences between communities were tested by performing the Analysis of Similarities (ANOSIM) using the vegan's function `anosim` and 1000 permutations [@Oksanen2019], while differences in relative contributions or proportions of shared OTUs and communities were tested by applying the Mann–Whitney *U* test using the function `wilcox.test`.
+In addition, differences between community type estimators or indices were tested by performing the Kruskal-Wallis *H* test (function `kruskal.test`) followed by a pairwise comparison using the Mann-Whitney *U* test (function `pairwise.wilcox`).
+Bonferroni correction was used to address the problem of multiple comparisons.
+
+A total of `r million_sequences` million sequences after quality curation and exclusion of sequences without known relatives (no relative sequences) and eukaryotic, chloroplast and mitochondrial sequences were obtained (\autoref{supp-nseq_notus}).
+The number of reads per sample ranged between `r min_max$min` and `r min_max$max` sequences (\autoref{supp-nseq_notus}).
+Even when the highest sequencing effort was applied the rarefaction curves did not level off as commonly observed in high-throughput 16S rRNA amplicon sequencing approaches (\autoref{supp-rarefaction}).
+Following quality curation and exclusion of sequences as mentioned above reads were clustered into `r no_otus` different OTUs.
+Read numbers were normalized to the minimum number of sequences (`r min_max$min`, \autoref{supp-nseq_notus}) using previously mentioned vegan's function `rrarefy` resulting in `r no_otus_rarefied` different OTUs ranging from `r min(otu_min_max)` to `r max(otu_min_max)` OTUs per sample (\autoref{supp-calculators}).
+Based on the ATCC MSA-1002 mock community included in the analysis an average sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
+In addition, the negative controls processed together with the samples yielded only `r nc` sequences after sequence quality curation.
+The detailed analysis procedure including the R Markdown file is available as a GitHub repository (https://github.com/MicrobesRovinj/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021).
+
+\newpage
+## Results
 
@@
             mean_invsimpson=mean(invsimpson), sd_invsimpson=sd(invsimpson))
-```
-
-Sequencing of the 16S rRNA V4 region yielded a total of `r million_sequences` million sequences after quality curation and exclusion of sequences without known relatives (no relative sequences) and eukaryotic, chloroplast and mitochondrial sequences (\autoref{supp-nseq_notus}).
+
+# Performing the Kruskal-Wallis H test
+kruskal_S.obs <- kruskal.test(S.obs ~ station, data=estimators_indices_metadata)
+kruskal_S.obs <- case_when(kruskal_S.obs$p.value < 0.05 & kruskal_S.obs$p.value >= 0.01 ~ 0.05,
+                           kruskal_S.obs$p.value < 0.01 & kruskal_S.obs$p.value >= 0.001 ~ 0.01,
+                           kruskal_S.obs$p.value < 0.001 & kruskal_S.obs$p.value >= 0.0001 ~ 0.001,
+                           kruskal_S.obs$p.value < 0.0001 ~ 0.0001,
+                           TRUE ~ kruskal_S.obs$p.value)
+```
+
 A total of `r samples_epiphytes` samples originating from epiphytic archaeal and bacterial communities associated with surfaces of the seagrass *C.
@@
 In addition, `r samples_seawater` samples (one of the samples was sequenced twice) originating from the ambient seawater were also processed for comparison.
-The number of reads per sample ranged between `r min_max$min` and `r min_max$max` sequences (\autoref{supp-nseq_notus}).
-Even when the highest sequencing effort was applied the rarefaction curves did not level off as commonly observed in high-throughput 16S rRNA amplicon sequencing approaches (\autoref{supp-rarefaction}).
-Following quality curation and exclusion of sequences as mentioned above reads were clustered into `r no_otus` different OTUs at a similarity level of 97 \si{\percent}.
-Read numbers were normalized to the minimum number of sequences (`r min_max$min`, \autoref{supp-nseq_notus}) through rarefaction resulting in `r no_otus_rarefied` different OTUs with `r min(otu_min_max)` to `r max(otu_min_max)` OTUs per sample (\autoref{supp-calculators}).
-To determine seasonal changes in richness and diversity the observed number of OTUs, Chao1, ACE, Exponential Shannon and Inverse Simpson [@Jost2006] were calculated after normalization through rarefaction.
 Generally, richness estimators and diversity indices showed similar trends.
@@
 cylindracea* (mixed [Number of OTUs, `r S.obs_fcam$mean` ± `r S.obs_fcam$sd` OTUs] and monospecific [Number of OTUs, `r S.obs_fca$mean` ± `r S.obs_fca$sd` OTUs]) than for *C.
-nodosa* (Number of OTUs, `r S.obs_fcym$mean` ± `r S.obs_fcym$sd` OTUs) and lowest values were obtained for the microbial community of the ambient seawater (Number of OTUs, `r format(S.obs_f$mean, digits=1, nsmall=1)` ± `r S.obs_f$sd` OTUs) (\autoref{supp-calculators}).
-Seasonal changes did not reveal such large dissimilarities.
+nodosa* (Number of OTUs, `r S.obs_fcym$mean` ± `r S.obs_fcym$sd` OTUs) and lowest values were obtained for the microbial community of the ambient seawater (Number of OTUs, `r format(S.obs_f$mean, nsmall=1)` ± `r S.obs_f$sd` OTUs) (Kruskal-Wallis, *p* < `r kruskal_S.obs`) (\autoref{supp-calculators} and Tables \ref{supp-calculators_community_type} and \ref{supp-calculator_statistics}).
+Temporal changes did not reveal such large dissimilarities.
 *C.
-nodosa* communities showed a slow increase towards the end of the study, while *C.
+nodosa* communities showed a slow increase in all calculated richness estimators towards the end of the study, while *C.
 cylindracea* (mixed and monospecific) communities were characterized by slightly higher values in spring and summer than in autumn and winter (\autoref{supp-calculators}).
@@
 
+```{r seasonal_shared_temporal}
+# Loading rarefied community data created during the generation of the richness and diversity calculators plot
+load(file="results/numerical/rarefied.Rdata")
+
+# Copying the sample labels to the rows (input for library vegan)
+rarefied <- rarefied %>%
+  column_to_rownames("Group")
+
+# Calculating dissimilarity indices
+jaccard <- as.matrix(vegdist(rarefied, method="jaccard", binary=T))
+jaccard <- as_tibble(jaccard) %>%
+  add_column("V1"=rownames(jaccard), .before=TRUE) %>%
+  gather(key="V2", value="jaccard", 2:ncol(.))
+
+bray <- as.matrix(vegdist(rarefied, method="bray", binary=F))
+bray <- as_tibble(bray) %>%
+  add_column("V1"=rownames(bray), .before=TRUE) %>%
+  gather(key="V2", value="bray", 2:ncol(.))
+
+distance <- inner_join(jaccard, bray, by=c("V1"="V1", "V2"="V2")) %>%
+  mutate_at(c("V1", "V2"), list(~ as.character(.)))
+
+# Loading metadata 
+metadata <- read_tsv("data/raw/metadata.csv")
+
+# Joining metadata with dissimilarity index data
+distance_metadata <- inner_join(metadata, distance, by=c("ID"="V1")) %>%
+  inner_join(., metadata, by=c("V2"="ID")) %>%
+  mutate(date.x=as.Date(date.x, "%d.%m.%Y")) %>%
+  mutate(date.y=as.Date(date.y, "%d.%m.%Y")) %>%
+  mutate(jaccard=(1-jaccard)*100) %>%
+  mutate(bray=(1-bray)*100) %>%
+  gather(key="index", value="value", jaccard, bray)
+
+# Filtering out consecutive sampling points
+distance_metadata <- distance_metadata %>%
+  filter((date.x=="2017-07-13" & date.y=="2017-07-27") |
+         (date.x=="2017-07-27" & date.y=="2017-08-10") |
+         (date.x=="2017-08-10" & date.y=="2017-08-24") |
+         (date.x=="2017-08-24" & date.y=="2017-09-19") |
+         (date.x=="2017-09-19" & date.y=="2017-10-12") |
+         (date.x=="2017-10-12" & date.y=="2017-11-23") |
+         (date.x=="2017-11-23" & date.y=="2017-12-04") |
+         (date.x=="2017-12-04" & date.y=="2017-12-14") |
+         (date.x=="2017-11-23" & date.y=="2017-12-14" & (station.x=="F" & station.y=="F")) |
+         (date.x=="2017-12-14" & date.y=="2018-02-13") |
+         (date.x=="2018-02-13" & date.y=="2018-03-27") |
+         (date.x=="2017-12-14" & date.y=="2018-03-27" & (station.x=="FCaM" & station.y=="FCaM")) |
+         (date.x=="2018-03-27" & date.y=="2018-04-24") |
+         (date.x=="2018-04-24" & date.y=="2018-05-22") |
+         (date.x=="2018-05-22" & date.y=="2018-06-19") |
+         (date.x=="2018-06-19" & date.y=="2018-07-10") |
+         (date.x=="2018-07-10" & date.y=="2018-08-09") |
+         (date.x=="2018-08-09" & date.y=="2018-09-04") |
+         (date.x=="2018-09-04" & date.y=="2018-10-05")) %>%
+  filter((station.x=="F" & station.y=="F") |
+         (station.x=="FCyM" & station.y=="FCyM") |
+         (station.x=="FCaM" & station.y=="FCaM") |
+         (station.x=="FCa" & station.y=="FCa"))
+
+# Summarising the data
+seasonal_shared_temporal <- distance_metadata %>%
+  group_by(index, station.x) %>%
+  summarise(mean=mean(value), sd=sd(value))
+
+# Performing the Mann–Whitney U test
+jaccard_test <- wilcox.test(filter(distance_metadata, index=="jaccard" & station.x=="F")$value,
+                            filter(distance_metadata, index=="jaccard" & station.x!="F")$value)
+
+jaccard_test <- case_when(jaccard_test$p.value < 0.05 & jaccard_test$p.value >= 0.01 ~ 0.05,
+                          jaccard_test$p.value < 0.01 & jaccard_test$p.value >= 0.001 ~ 0.01,
+                          jaccard_test$p.value < 0.001 & jaccard_test$p.value >= 0.0001 ~ 0.001,
+                          jaccard_test$p.value < 0.0001 ~ 0.0001,
+                          TRUE ~ jaccard_test$p.value)
+
+bray_test <- wilcox.test(filter(distance_metadata, index=="bray" & station.x=="F")$value,
+                            filter(distance_metadata, index=="bray" & station.x!="F")$value)
+
+bray_test <- case_when(bray_test$p.value < 0.05 & bray_test$p.value >= 0.01 ~ 0.05,
+                       bray_test$p.value < 0.01 & bray_test$p.value >= 0.001 ~ 0.01,
+                       bray_test$p.value < 0.001 & bray_test$p.value >= 0.0001 ~ 0.001,
+                       bray_test$p.value < 0.0001 ~ 0.0001,
+                       TRUE ~ bray_test$p.value)
+```
+
 ```{r core_otus}
@@
 
-# Finding the core OTUs for each environment
+# Finding the core OTUs of each environment
+# Cymodocea nodosa (mixed)
 core <- metadata_rarefied %>%
@@
 
+# Caulerpa cylindracea (mixed)
 core <- metadata_rarefied %>%
@@
 
+# Caulerpa cylindracea (monospecific)
 core <- metadata_rarefied %>%
@@
 
-To determine the proportion of shared archaeal and bacterial OTUs and communities sampled in different environments the Jaccard's Similarity Coefficient on presence-absence data and Bray-Curtis Similarity Coefficient, respectively, were calculated.
-Coefficients were determined after normalization through rarefaction and binning of samples from the particular environment.
-The highest proportion of shared OTUs and community was found between mixed and monospecific  *C.
-cylindracea* (Jaccard, `r format(filter(similarity, V1=="FCa", V2=="FCaM")$jaccard, digits=2, nsmall=2)`; Bray-Curtis, `r format(filter(similarity, V1=="FCa", V2=="FCaM")$bray, digits=2, nsmall=2)`), while lower shared values were calculated between ambient seawater and epiphytic communities (\autoref{matrix}).
-Shared proportion between *C.
-nodosa* and *C.
-cylindracea* were approximately in-between the values of mixed and monospecific *C.
-cylindracea*.
-To assess seasonal changes in the proportion of shared OTUs and communities the Jaccard's and Bray-Curtis Similarity Coefficients were calculated between consecutive sampling points (\autoref{shared}).
-Both coefficients showed similar trends.
-Temporal proportional changes were more pronounced for ambient seawater than for *C.
-nodosa* and especially *C.
-cylindracea* associated communities (\autoref{shared}).
-In addition, only `r min(core$core_otus)` -- `r max(core$core_otus)` \si{\percent} of OTUs from each surface associated community were present at all seasons.
-These persistent OTUs constituted a high proportion of total sequences (`r min(core$core_seq)` -- `r max(core$core_seq)` \si{\percent}).
-To further disentangle the environmental and seasonal community dissimilarity a Principal Coordinates Analysis (PCoA) based on Bray-Curtis distances and OTU abundances was applied.
 A clear separation between ambient seawater and surface associated communities was found (\autoref{pcoa}).
 In addition, a separation of epiphytic bacterial and archaeal communities based on host species was detected.
-This separation was further supported by ANOSIM (R = `r format(anosim_all$statistic, digits=2, nsmall=2)`, *p* < `r anosim_all_signif`).
+This separation was further supported by ANOSIM (R = `r format(round(anosim_all$statistic, 2), nsmall=2)`, *p* < `r anosim_all_signif`).
+The highest proportion of shared OTUs and community was found between mixed and monospecific  *C.
+cylindracea* (Jaccard, `r format(round(filter(similarity, V1=="FCa", V2=="FCaM")$jaccard, 2), nsmall=2)`; Bray-Curtis, `r format(round(filter(similarity, V1=="FCa", V2=="FCaM")$bray, 2), nsmall=2)`), while lower shared values were calculated between ambient seawater and epiphytic communities (Jaccard, `r format(round(min(filter(similarity, V1=="F")$jaccard), 2), nsmall=2)` -- `r format(round(max(filter(similarity, V1=="F")$jaccard), 2), nsmall=2)`; Bray-Curtis, `r format(round(min(filter(similarity, V1=="F")$bray), 2), nsmall=2)` -- `r format(round(max(filter(similarity, V1=="F")$bray), 2), nsmall=2)`).
+Shared proportions of OTUs and communities between *C.
+nodosa* and *C.
+cylindracea* (either mixed or monospecific) were approximately in-between the values obtained for the comparison of ambient seawater with all other communities and for the comparison of the mixed and monospecific *C.
+cylindracea* associated community.
 Seasonal changes of *C.
-nodosa* associated communities indicated a separation between spring, summer and autumn/winter samples (ANOSIM, R = `r format(anosim_fcym$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fcym_signif`).
+nodosa* associated communities indicated a separation between spring, summer and autumn/winter samples (ANOSIM, R = `r format(round(anosim_fcym$statistic, 2), nsmall=2)`, *p* < `r anosim_fcym_signif`).
 For *C.
 cylindracea* associated communities a separation between summer and autumn/winter/spring samples was observed that was, however, not as strong as for *C.
-nodosa* associated communities (ANOSIM, R = `r format(anosim_fca$statistic, digits=2, nsmall=2)`, *p* < `r anosim_fca_signif`) (\autoref{pcoa}).
+nodosa* associated communities (ANOSIM, R = `r format(round(anosim_fca$statistic, 2), nsmall=2)`, *p* < `r anosim_fca_signif`) (\autoref{pcoa}).
+Shared proportions of OTUs between consecutive sampling points were lower for ambient seawater (`r filter(seasonal_shared_temporal, index=="jaccard", station.x=="F")$mean` ± `r filter(seasonal_shared_temporal, index=="jaccard", station.x=="F")$sd` \si{\percent}) than for *C.
+nodosa* (`r filter(seasonal_shared_temporal, index=="jaccard", station.x=="FCyM")$mean` ± `r filter(seasonal_shared_temporal, index=="jaccard", station.x=="FCyM")$sd` \si{\percent}) and *C.
+cylindracea* (mixed [`r filter(seasonal_shared_temporal, index=="jaccard", station.x=="FCaM")$mean` ± `r filter(seasonal_shared_temporal, index=="jaccard", station.x=="FCaM")$sd` \si{\percent}] and monospecific [`r filter(seasonal_shared_temporal, index=="jaccard", station.x=="FCa")$mean` ± `r filter(seasonal_shared_temporal, index=="jaccard", station.x=="FCa")$sd` \si{\percent}]) associated communities (*p* < `r jaccard_test`), while mean proportions of shared communities between consecutive sampling points did not show such a difference (seawater, `r filter(seasonal_shared_temporal, index=="bray", station.x=="F")$mean` ± `r filter(seasonal_shared_temporal, index=="bray", station.x=="F")$sd` \si{\percent}; *C.
+nodosa*, `r filter(seasonal_shared_temporal, index=="bray", station.x=="FCyM")$mean` ± `r filter(seasonal_shared_temporal, index=="bray", station.x=="FCyM")$sd` \si{\percent}; *C.
+cylindracea* [mixed], `r filter(seasonal_shared_temporal, index=="bray", station.x=="FCaM")$mean` ± `r filter(seasonal_shared_temporal, index=="bray", station.x=="FCaM")$sd` \si{\percent}; *C.
+cylindracea* [monospecific], `r filter(seasonal_shared_temporal, index=="bray", station.x=="FCa")$mean` ± `r filter(seasonal_shared_temporal, index=="bray", station.x=="FCa")$sd` \si{\percent}) (*p* = `r bray_test`), although in ambient seawater higher fluctuations could be observed (\autoref{shared}).
+In addition, only `r min(core$core_otus)` -- `r max(core$core_otus)` \si{\percent} of OTUs from each surface associated community were present at all seasons.
+These persistent OTUs constituted a high proportion of total sequences (`r min(core$core_seq)` -- `r max(core$core_seq)` \si{\percent}) and were mainly contributing to abundant phylogenetic groups present throughout the year, e.g.
+the no relative *Rhodobacteraceae* in the case of *C.
+nodosa* or taxa within *Desulfobacterota* in the case of *C.
+cylindracea* (see below) (\autoref{supp-core_otus_taxonomy_table}).
 
@@
   summarise(mean=mean(abundance), sd=sd(abundance))
-```
-
-The taxonomic composition of both, macrophyte associated and ambient seawater communities was dominated by bacterial (`r filter(archaea_bacteria, taxon=="Bacteria")$mean` ± `r filter(archaea_bacteria, taxon=="Bacteria")$sd` \si{\percent}) over archaeal sequences (`r filter(archaea_bacteria, taxon=="Archaea")$mean` ± `r filter(archaea_bacteria, taxon=="Archaea")$sd` \si{\percent}) (\autoref{community}).
-Higher relative abundances of chloroplast related sequences were only observed in surface associated communities, with higher values in autumn/winter (`r filter(chloroplast, season=="Winter")$mean` ± `r filter(chloroplast, season=="Winter")$sd` \si{\percent}) than in spring/summer (`r filter(chloroplast, season=="Summer")$mean` ± `r filter(chloroplast, season=="Summer")$sd` \si{\percent}) (\autoref{supp-chloroplast}).
+
+# Performing the Mann–Whitney U test
+chloroplast_test <- plot %>%
+  filter(station!="F" & taxon!="Other") %>%
+  mutate(season=str_replace(season, "Autumn", "Winter")) %>%
+  mutate(season=str_replace(season, "Spring", "Summer")) %>%
+  select(taxon, season, abundance)
+
+chloroplast_test <- wilcox.test(filter(chloroplast_test, season=="Winter")$abundance,
+                                filter(chloroplast_test, season=="Summer")$abundance)
+
+chloroplast_test <- case_when(chloroplast_test$p.value < 0.05 & chloroplast_test$p.value >= 0.01 ~ 0.05,
+                              chloroplast_test$p.value < 0.01 & chloroplast_test$p.value >= 0.001 ~ 0.01,
+                              chloroplast_test$p.value < 0.001 & chloroplast_test$p.value >= 0.0001 ~ 0.001,
+                              chloroplast_test$p.value < 0.0001 ~ 0.0001,
+                              TRUE ~ chloroplast_test$p.value)
+```
+
+The taxonomic composition of both, macrophyte associated and ambient seawater community, was dominated by bacterial (`r filter(archaea_bacteria, taxon=="Bacteria")$mean` ± `r filter(archaea_bacteria, taxon=="Bacteria")$sd` \si{\percent}) over archaeal sequences (`r filter(archaea_bacteria, taxon=="Archaea")$mean` ± `r filter(archaea_bacteria, taxon=="Archaea")$sd` \si{\percent}) (\autoref{community}).
+Higher relative abundances of chloroplast related sequences were only observed in surface associated communities, with higher values in autumn/winter (`r filter(chloroplast, season=="Winter")$mean` ± `r filter(chloroplast, season=="Winter")$sd` \si{\percent}) than in spring/summer (`r filter(chloroplast, season=="Summer")$mean` ± `r filter(chloroplast, season=="Summer")$sd` \si{\percent}) (*p* < `r chloroplast_test`) (\autoref{supp-chloroplast}).
 Generally, at higher taxonomic ranks (phylum-class), epiphytic and ambient seawater microbial communities were composed of similar bacterial taxa.
@@
 nodosa* communities with the addition of *Desulfobacterota* (\autoref{community}).
-Larger differences between environments and host species were observed at lower taxonomic ranks (Figs.
-\ref{cyano} -- \ref{desulfo}).
+Larger differences between environments and host species were observed at lower taxonomic ranks (Figures \ref{cyano} -- \ref{desulfo}).
 
@@
   summarise(mean=mean(abundance), sd=sd(abundance))
+
+# Performing the Mann–Whitney U test
+cyano_station_season_whole_test <- inner_join(metadata, whole, by=c("ID"="Group")) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y"))
+
+cyano_station_season_whole_test <- wilcox.test(filter(cyano_station_season_whole_test, season=="Autumn" & station!="F")$abundance,
+                                               filter(cyano_station_season_whole_test, season!="Autumn" & station!="F")$abundance)
+
+cyano_station_season_whole_test <- case_when(cyano_station_season_whole_test$p.value < 0.05 & cyano_station_season_whole_test$p.value >= 0.01 ~ 0.05,
+                                             cyano_station_season_whole_test$p.value < 0.01 & cyano_station_season_whole_test$p.value >= 0.001 ~ 0.01,
+                                             cyano_station_season_whole_test$p.value < 0.001 & cyano_station_season_whole_test$p.value >= 0.0001 ~ 0.001,
+                                             cyano_station_season_whole_test$p.value < 0.0001 ~ 0.0001,
+                                             TRUE ~ cyano_station_season_whole_test$p.value)
 ```
@@
 nodosa* (`r filter(cyano_station_season_whole, station=="FCyM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCyM" & season=="Autumn")$sd` \si{\percent}) and *C.
-cylindracea* mixed (`r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$sd` \si{\percent}) and monospecific (`r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$sd` \si{\percent}) associated communities in autumn and for ambient seawater communities in winter (`r filter(cyano_station_season_whole, station=="F" & season=="Winter")$mean` ± `r filter(cyano_station_season_whole, station=="F" & season=="Winter")$sd` \si{\percent}).
+cylindracea* mixed (`r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$sd` \si{\percent}) and monospecific (`r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$sd` \si{\percent}) associated communities in autumn (*p* < `r cyano_station_season_whole_test`) and for ambient seawater communities in winter (`r filter(cyano_station_season_whole, station=="F" & season=="Winter")$mean` ± `r filter(cyano_station_season_whole, station=="F" & season=="Winter")$sd` \si{\percent}).
 Large taxonomic differences between surface associated and ambient seawater cyanobacterial communities were observed.
 Ambient seawater communities were mainly comprised of *Cyanobium* and *Synechococcus*, while surface associated communities were comprised of *Pleurocapsa* and sequences within the class *Cyanobacteriia* that could not be further classified (no relative *Cyanobacteriia*) (\autoref{cyano}).
-In addition, seasonal changes in surface associated communities were observed in *Pleurocapsa* and no relative *Cyanobacteriia* comprising larger proportions in autumn and winter and *Acrophormium*, *Phormidesmis* and sequences without known relatives within the *Nodosilineaceae* (no relative *Nodosilineaceae*) in spring and summer (\autoref{cyano}).
+In addition, seasonal changes in surface associated cyanobacterial communities were observed with *Pleurocapsa* and no relative *Cyanobacteriia* comprising larger proportions of *Cyanobacteria* in autumn and winter and *Acrophormium*, *Phormidesmis* and sequences without known relatives within the *Nodosilineaceae* (no relative *Nodosilineaceae*) in spring and summer (\autoref{cyano}).
 
@@
 Sequences classified as *Bacteroidota* comprised, on average, `r bactero_all_whole$mean` ± `r bactero_all_whole$sd` \si{\percent} of all sequences (\autoref{bactero}).
-Similarly to *Cyanobacteria*, large differences in the taxonomic composition between ambient seawater and surface associated communities were found (\autoref{bactero}).
+Similar to *Cyanobacteria*, large differences in the taxonomic composition between ambient seawater and surface associated communities were found (\autoref{bactero}).
 The ambient seawater community was characterized by the NS4 and NS5 marine groups, uncultured *Cryomorphaceae*, uncultured *Flavobacteriaceae*, NS11-12 marine group, *Balneola*, uncultured *Balneolaceae* and *Formosa*.
-In contrast, in surface associated communities *Lewinella*, *Portibacter*, *Rubidimonas*, sequences without known relatives within the *Saprospiraceae* (no relative *Saprospiraceae*), uncultured *Saprospiraceae*, (sequences without known relatives within the *Flavobacteriaceae* (no relative *Flavobacteriaceae*) and uncultured *Rhodothermaceae* were found.
+In contrast, in surface associated communities *Lewinella*, *Portibacter*, *Rubidimonas*, sequences without known relatives within the *Saprospiraceae* (no relative *Saprospiraceae*), uncultured *Saprospiraceae*, sequences without known relatives within the *Flavobacteriaceae* (no relative *Flavobacteriaceae*) and uncultured *Rhodothermaceae* were found.
 Some groups showed minor seasonal changes such as no relative *Flavobacteriaceae* whose sequences were more abundant from November 2017 until June 2018.
@@
 
-# Calculating the mean abundance and standard deviation of Rhodobacteraceae_unclassified within Alpharoteobacteria
+# Calculating the mean abundance and standard deviation of Rhodobacteraceae_unclassified within Alphaproteobacteria
 rhodo <- plot %>%
-  filter(season!="F" & taxon=="Rhodobacteraceae_unclassified") %>%
+  filter(station!="F" & taxon=="Rhodobacteraceae_unclassified") %>%
   summarise(mean=mean(abundance), sd=sd(abundance))
@@
 ## Discussion
-Surfaces of marine macrophytes harbor biofilms consisting of diverse microbial taxa [@Egan2013; @Tarquinio2019].
-No standard protocol has been developed to study these macophyte-associated microbes [@Ugarelli2019].
-Different procedures for removal of microbial cells from host surfaces are described, such as host tissue shaking [@Noges2010], scraping [@Uku2007], swabbing [@Mancuso2016] and ultrasonication [@Cai2014].
-All these methods result in different removal efficiencies but none was enabling a complete removal of attached microbial cells based on our experience.
-In the present study, we applied a removal protocol [@Korlevic, submitted] based on direct cellular lysis [@Burke2009].
-The application of a direct lysis procedure coupled with a high sampling frequency and Illumina amplicon sequencing has enabled us to described in detail the bacterial and archaeal communities associated with the surfaces of two marine macrophytes, *C.
+In the present study, we applied a selective epiphytic DNA isolation procedure based on direct cellular lysis [@Korlevic2021] coupled with a monthly sampling and Illumina amplicon sequencing to describe in detail the bacterial and archaeal communities associated with the surfaces of two marine macrophytes, *C.
 nodosa* and *C.
 cylindracea*.
-
-In the present study, highest richness was observed for *C.
+Highest richness was observed for *C.
 cylindracea* (mixed and monospecific) followed by *C.
 nodosa* and lowest richness was found in ambient seawater microbial communities.
-Higher richness of microbial communities associated with seagrasses than in ambient seawater was described earlier and could be attributed to a larger set of inhabitable microniches existing on macrophyte surfaces than in the ambient seawater [@Ugarelli2019].
+Higher richness of microbial communities associated with macrophytes than in ambient seawater has been described earlier [@Ugarelli2019; @Mancuso2016] and could be attributed to a larger set of inhabitable microniches existing on macrophyte surfaces than in the ambient seawater.
 The highest richness observed for *C.
@@
 The stolon of *C.
-cylindracea* is attached to the sediment surface with rhizoids and thus, the stolon and rhizoids are in a direct contact with the sediment.
-In addition, seasonal differences in richness observed for surface attached communities indicated a slightly higher richness in spring and summer.
-This pattern could be explained by a higher macrophyte growth in these two seasons than in autumn and winter [M.
-Najdek, personal communication; @Zavodnik1998; @Ruitton2005].
+cylindracea* is attached to the sediment surface with rhizoids and thus, the stolon and rhizoids are in direct contact with the sediment.
+Also, studies have shown that the presence of *C.
+cylindracea* can alter the content and biochemical composition of sedimentary organic matter [@Rizzo2017; @Pusceddu2016; @Rizzo2020] possibly further expanding the number of inhabitable microniches and thus causing the observed increase in richness.
+Seasonal differences in richness observed for surface attached communities indicated a slightly higher richness in spring and summer.
+This pattern could be explained by a more intense macrophyte growth in these two seasons than in autumn and winter [@Najdek2020a; @Zavodnik1998; @Ruitton2005].
 During their main growth season in spring and summer macrophytes exhibit a more dynamic chemical interaction with the surface community probably causing an increase in the number of inhabitable microniches [@Rickert2016; @Borges2015].
-
-We observed a strong differentiation between the surface attached and ambient seawater communities at the level of OTUs, in agreement with most published studies [@Ugarelli2019; @Roth-Schulze2016; @Burke2011a; @Michelou2013; @Crump2018; @Mancuso2016].
-This indicates that marine macrophytes are a selecting factor from the pool of microbial taxa present in the ambient seawater, modifying the microbial community once the macrophyte associated microbial biofilm develops [@Salaun2012; @Michelou2013].
-In contrast, @Fahimipour2017 report in a global study of *Zostera marina*, similarities between the microbial community developed on leaves and in the ambient seawater.
-The discrepancy between our data and the study of @Fahimipour2017 could be explained by different seagrass species, methodological variations or biogeographic trends as @Fahimipour2017 analysed samples from different locations throughout the northern hemisphere.
-It is possible that the microbial communities in ambient seawater and on leaves from the same location are differing but are still more similar to each other when compared to other sampling locations.
-Indeed, it was found that prokaryotic communities vary substantially between different sampling sites [@Bengtsson2017].
-When the taxonomic composition at high ranks was analysed no such strong differentiation was noticed.
-Phyla and classes such as *Actinobacteriota*, *Bacteroidota*, *Cyanobacteria*, *Alphaproteobacteria*, *Gammaproteobacteria* and *Verrucomicrobiota* were found in both ambient seawater as well as macrophyte associated, in agreement with previous studies [@Egan2013; @Michelou2013; @Burke2011a].
-In contrast, when low taxonomic ranks were analysed (i.e., family and genus) a strong differentiation was observed (Figs.
-\ref{cyano} -- \ref{desulfo}).
-A similar differentiation at lower microbial taxonomic ranks between ambient seawater and macrophytes was described for other macrophyte species as well [@Egan2013; @Michelou2013; @Ugarelli2019].
-
-Beside differences between ambient seawater and surface associated microbial communities, it is unclear whether the prokaryotic epiphytic community is host-specific or whether there are generalist taxa characteristic to all or many macrophytes [@Egan2013].
-Similar to previously described differences between microbial communities in the ambient seawater and on macrophytes, at high taxonomic ranks no major difference between the microbial communities associated with different hosts was observed.
-The only high rank phylum that was differing between *C.
-nodosa* and *C.
-cylindracea* was *Desulfobacterota*, with more abundant sequences in the *C.
-cylindracea* associated community.
-As already mentioned, the rhizoids and part of the stolon are in contact with the sediment.
-Thus *Desulfobacterota* are probably a part of the epiphytic community that was in contact with the sediment.
-Similar high rank taxa found in this study were described to be specific for other species of macrophytes [@Burke2011a; @Lachnit2011; @Bengtsson2017; @Mancuso2016].
-In contrast to high taxonomic ranks, a substantial differentiation between host specific communities was found supporting the notion that macrophyte associated microbial communities might be host-specific.
-Host-specificity was also observed for other species of macroalgae and seagrasses [@Lachnit2011; @Roth-Schulze2016; @Ugarelli2019; @Morrissey2019].
-Taken together, at high taxonomic ranks a core set of taxa could be described that is characteristic for all or many macrophytes, while at low taxonomic ranks a community specific to host species was identified (Figs.
-\ref{pcoa} and \ref{community}) [@Egan2013].
-
-Seasonal changes in richness in the epiphytic community were substantial as indicated by the proportion of OTUs ($\leq$ `r max(core$core_otus)` \si{\percent}) present at every sampling date.
-These persistent OTUs, however, were accounting for a high proportion of sequences ($\geq$ `r min(core$core_seq)` \si{\percent}) (\autoref{shared}).
-A very similar proportion of persistent OTUs was reported in high-frequency sampling studies describing seasonal changes in picoplankton [@Gilbert2009; @Gilbert2012].
-In comparison to the seawater community, a lower degree of seasonal shifts was observed for the macrophyte surface associated communities.
-It appears that microniches at the surfaces of macrophytes are providing more stable conditions than the ambient seawater.
+Proportions of shared epiphytic OTUs between consecutive sampling points were low also indicated by the proportion of OTUs ($\leq$ `r max(core$core_otus)` \si{\percent}) present at every sampling date (\autoref{shared}).
+These persistent OTUs, however, accounted for a high proportion of sequences ($\geq$ `r min(core$core_seq)` \si{\percent}), as is often the case with similar high-frequency sampling studies [@Gilbert2009; @Gilbert2012].
+In comparison to the seawater community, higher values of shared OTUs between consecutive sampling points were observed for the macrophyte surface associated communities.
+It appears that macrophyte surfaces are providing more stable conditions than the ambient seawater.
+
+We observed a strong differentiation between the surface attached and ambient seawater communities at the level of OTUs which is in agreement with most published studies [@Ugarelli2019; @Roth-Schulze2016; @Burke2011a; @Michelou2013; @Crump2018; @Mancuso2016; @Sanders-Smith2020].
+This indicates that marine macrophytes are selecting microorganisms from the pool of microbial taxa present in the ambient seawater, modifying the microbial community once the macrophyte associated microbial biofilm develops [@Salaun2012; @Michelou2013].
+In addition, similar to the study of @Roth-Schulze2016 seagrass and macroalgae specific microbial communities were identified, while no difference between *C.
+cylindracea* settlements was observed indicating that seagrass and macroalgae specific metabolism is involved in the selection and development of the associated biofilm.
 At the level of OTUs seasonal changes of *C.
 nodosa* and *C.
-cylindracea* associated communities were identified that could be linked to the growth cycle of the seagrass and macroalgae (M.
-Najdek, personal communication).
+cylindracea* associated communities were identified that could be linked to the growth cycle of the seagrass and macroalgae [@Najdek2020a; @Agostini2003].
 *C.
-nodosa* was characterized by a spring community during maximum seagrass proliferation, a summer community during the highest standing stock of *C.nodosa* and an autumn/winter community during the decay of seagrass biomass.
+nodosa* was characterized by a spring community during maximum seagrass proliferation, a summer community during the highest standing stock of *C.
+nodosa* and an autumn/winter community during the decay of seagrass biomass.
 In contrast, *C.
 cylindracea* started to proliferate in late spring and was characterized only by a summer community during high growth rates and by an autumn/winter/spring community when the biomass was at the peak and decaying thereafter.
-Similar seasonal changes in the epiphytic community were also described for other macroalgae [@Lachnit2011; @Tujula2010].
-Higher seasonal stability of *C.
-cylindracea* surface communities than in *C.
-nodosa* was also observed in the higher proportion of shared communities between two consecutive sampling dates in *C.
+Similar seasonal changes in the epiphytic community have also been described for other macroalgae [@Lachnit2011; @Tujula2010].
+
+The taxonomic analysis showed higher chloroplast sequence abundances in autumn/winter than in spring/summer.
+This pattern is not surprising as seagrasses harbour more algal epiphytes during autumn/winter than in spring/summer [@Reyes2001].
+Furthermore, we used an adapted DNA isolation protocol that is known to partially co-extract DNA from planktonic eukaryotes [@Korlevic2015].
+In general, the taxonomic analysis identified epiphytic phylogenetic groups present throughout the year comprising most of the reads, and taxa present in lower proportions showing seasonal patterns.
+The first group was comprised of members of the *Bacteroidota* family *Saprospiraceae*, the alphaproteobacterial *Rhodobacteraceae* and *Hyphomonadaceae*, the gammaproteobacterial genus *Granulosicoccus*, sequences without known relatives within *Gammaproteobacteria* and various taxa within *Desulfobacterota* (Figures \ref{bactero} -- \ref{desulfo}).
+All these groups were found on all host species, with the exception of *Desulfobacterota* that was characteristic for *C.
 cylindracea*.
-
-Chloroplast sequence abundances were higher in autumn/winter than in spring/summer.
-This pattern is not surprising as seagrasses harbor more algal epiphytes during autumn/winter than in spring/summer [@Reyes2001].
-Furthermore, we used an adapted DNA isolation protocol that is known to partially co-extract DNA from planktonic eukaryotes [@Korlevic2015].
-Strong seasonal fluctuations of high rank epiphytic taxa were not observed, with the exception of *Cyanobacteria*.
+In addition, the persistence of *Rhodobacteraceae* in the case of *C.
+nodosa* and *Desulfobacterota* in the case of *C.
+cylindracea* could be observed in the taxonomic classification of OTUs present at every sampling date.
+Within the *Bacteroidota* different groups within *Saprospiraceae* (e.g.
+*Lewinella*, *Portibacter* and *Rubidimonas*) were identified to be persistent.
+It has been suggested that members of this family are important in the hydrolysis and utilization of complex organic sources [@McIlroy2014].
+Surface attached life style would be beneficial to these microbes as they could thrive on products of host cellular breakdown or by-products of host metabolism, so it not surprising that they are often found associated with macrophyte surfaces [@Burke2011a; @Crump2018; @McIlroy2014].
+*Rhodobacteraceae* are often detected on macrophyte surfaces and usually are one of the most abundant groups [@Burke2011a; @Michelou2013; @Mancuso2016].
+The functional association between macrophytes and members of this groups is difficult to assess based on 16S rRNA analysis as this family is phenotypically, metabolically, and ecologically very diverse [@Pujalte2014].
+However, some interesting metabolic capacities linked to this group were described.
+Genomic analysis of *Rhodobacteraceae* strains and metatranscriptomic sequencing of seagrass microbiomes revealed the potential for biosynthesis of indole-3-acetic acid (IAA), a plant hormone [@Simon2017], indicating a possible intake by seagrasses.
+However, another study found no effect of IAA on *C.
+nodosa* growth showing the complexity of macrophyte--microbes interactions [@Munoz1995].
+Another persistent alphaproteobacterial family was the *Hyphomonadaceae*, a group that contain species with stalks used to attach cells to different surfaces [@Abraham2014].
+This group has been previously associated with seagrass surfaces [@Weidner2000] and it is believed that possessing stalks could be an advantage to keep the cells in the proximity of exudate excreted by the host [@Abraham2014; @Weidner2000].
+
+Within the *Gammaproteobacteria*, sequences without known representatives were the most pronounced group present throughout the year.
+*Gammaproteobacteria* are often a major constituent of macrophyte epiphytic communities [@Michelou2013; @Burke2011a; @Crump2018].
+A study has attributed the expression of enzymes for the degradation of galactose-based algal polymers to this class indicating their possible involvement into epibiotic algal biofilm control [@Crump2018].
+In addition, *Granulosicoccus* was also found in almost all samples.
+A species of this genus has been isolated from the leaf surface of the seagrass *Zostera marina* [@Kurilenko2010], while sequences related to this genus have been found on the surfaces of macroalgae [@Bengtsson2012; @Lachnit2011], including *C.
+cylindracea* [@Rizzo2016], indicating this group preference for macrophyte surfaces.
+It is possible that bacteria of this genus can thrive on exudates of different macrophytes as it is known from cultivated members that they can utilise various sugars and amino acids [@Ivanova2014].
+The presence of *Desulfobacterota* only on *C.
+cylindracea* is to be expected as part of the epiphytic community is in direct contact with the sediment.
+The *Desulfobacterota* community was comprised of known sulphate sediment groups such as the *Desulfatitalea* and no relative *Desulfocapsaceae* [@Higashioka2015; @Kuever2014].
+Sequences related to sulphur cycling bacterial groups have been found in *Caulerpa* endophytic and epiphytic communities [@Aires2013].
+It is possible that these groups are involved into enhanced sulphate reduction rates observed in sediments underlying *Caulerpa* settlements causing unsuitable conditions to sulphide-sensitive seagrasses [@Holmer2009].
+
+The only high rank taxonomic group showing strong seasonal fluctuations was *Cyanobacteria*.
 Cyanobacterial sequences were more pronounced in November and December than in spring and summer.
-In the months of high cyanobacterial sequence abundances the majority of sequences from this group were classified as *Pleurocapsa*, a group known to colonized different living and non-living surfaces [@Longford2007; @Burns2004; @Mobberley2012; @Reisser2014].
-It is possible that during periods of low metabolic activity there is a reduced interaction and selection of the epiphytic community by the seagrass, causing leaves to become a suitable surface for non-specific colonizers [@Zavodnik1998].
-*Pleurocapsa* was replaced in spring and summer by *Acrophormium*, *Phormidesmis* and sequences without known relatives within the *Nodosilineaceae*.
-A study of coastal microbial mats found also a higher proportion of *Nodosilineaceae* sequences in summer, while *Phormidesmis* sequences were at their peak in autumn [@Cardoso2019].
-Other high rank taxa did not exhibit strong successional patterns.
-In every analysed group, with the exception of *Desulfobacterota*, taxa present throughout the year in similar proportions and season specific taxa could be identified (Figs.
-\ref{bactero} and \ref{desulfo}).
-Within *Bacteroidota* different groups within the family *Saprospiraceae* (e.g.
-*Lewinella*, *Portibacter* and *Rubidimonas*) were detected across all seasons.
-Members of this family are often found in association with macrophytes and it is suggested that they are involved in the hydrolysis and utilization of complex carbon sources [@Burke2011a; @Crump2018; @McIlroy2014].
-In contrast, the families *Flavobacteriaceae* and *Rhodothermaceae* showed seasonal patterns, with *Flavobacteriaceae* being more pronounced from November to June and *Rhodothermaceae* from June to October (\autoref{bactero}).
-Within *Alphaproteobacteria* the family *Rhodobacteraceae* comprised the majority of sequences throughout the year (\autoref{alpha}).
-This metabolically versatile family is often associated with macrophyte surfaces and usually is one of the most abundant groups [@Burke2011a; @Michelou2013; @Pujalte2014; @Mancuso2016].
-In addition, *Hyphomonadaceae* were found in all samples.
-Interestingly, some of the species within this group contain stalks on their cells, which can be used to attach to the macrophyte surface [@Weidner2000; @Abraham2014].
-Within the *Gammaproteobacteria*, sequences without known representatives were the most pronounced group present throughout the year (\autoref{gamma}).
-In addition, *Granulosicoccus* was also found in almost all samples.
-*Gammaproteobacteria* are often a major constituent of macrophyte epiphytic communities [@Michelou2013; @Burke2011a; @Crump2018].
-Beside these two groups, other less abundant, taxa showed seasonal and host-specific patterns.
-For example, *C.
-cylindracea* harbored *Thioploca*, a known sulfur sediment bacteria and *Cellvibrionaceae*, a family with cultured members known as polysaccharide degraders [@Jorgensen1999; @Xie2017].
-*Desulfobacterota* were found only associated with *C.
-cylindracea* and no group within this phylum showed seasonal patterns (\autoref{desulfo}).
-The presence of this phylum only on *C.
-cylindracea* is to be expected as part of the epiphytic community is in direct contact with the sediment.
-The *Desulfobacterota* community was dominated by *Desulfatitalea* and no relative *Desulfocapsaceae*, known sulphate sediment groups [@Higashioka2015; @Kuever2014].
-
-In temperate zones, marine macrophytes are exhibiting growth cycles, so it  is not surprising that the associated epiphytic microbial community is undergoing partial seasonal changes.
-In the present study, we could identify in every analysed high rank taxa phylogenetic groups present throughout the year, comprising most of the sequences and a lower proportion of taxa showing seasonal patterns connected to the macrophyte growth cycle (Figs.
-\ref{community} and \ref{desulfo}).
+In the months of high cyanobacterial sequence abundances the majority of sequences from this group were classified as *Pleurocapsa*, a group known to colonize different living and non-living surfaces [@Longford2007; @Burns2004; @Mobberley2012; @Reisser2014; @Kolda2020].
+While we observed a strong temporal pattern for this group, a study of surface sediment cyanobacterial communities did not find any seasonal dynamics for *Pleurocapsa* [@Kolda2020], indicating a possibility that there is a reduced selection of the epiphytic community by the seagrass during periods of low photosynthetic activity [@Zavodnik1998], causing leaves to become a suitable surface for non-specific colonizers.
+Beside all these groups comprising most of the sequences, a set of taxa present in lower proportions and showing seasonal patterns was identified.
+This group was comprised of e.g.
+*Bacteroidota* sequences without known relatives within *Flavobacteriaceae* and *Rhodothermaceae*, the alphaproteobacterial *Amylibacter* and the gammaproteobacterial *Methylotenera*, *Reinekea* and sequences without known relatives within *Cellvibrionaceae* (Figures \ref{bactero} and \ref{gamma}).
+
+It is possible that *Flavobacteriaceae* and *Rhodothermaceae* are occupying similar niches with *Rhodothermaceae* being more adapted to higher temperatures as it is known that culturable members of this family exhibit mesophilic and thermophilic characteristics [@Park2014].
+This would explain why we observed a higher presence of *Rhodothermaceae* in the warmer period of the year.
+A strain belonging to the *Rhodobacteraceae* genus *Amylibacter* has been isolated from the surface of a green macroalga indicating that members of this group can exhibit surface attached life style [@Nedashkovskaya2016].
+In addition, since this is a relatively novel genus it is possible that novel taxa within *Rhodobacteraceae* will be described in the future elucidating the taxonomy of the currently high proportion of *Rhodobacteraceae* sequences without known relatives.
+The genus *Methylotenera* belongs to the methylotrophic family *Methylophilaceae*, a group capable of oxidising non-methane single-carbon compounds such as methanol and methylamine [@Chistoserdova2018].
+Interestingly, angiosperms produce methanol during cell-wall synthesis [@Nemecek-Marshall1995; @Dorokhov2018], so it is not surprising that we found members of this genus only on *C.
+nodosa* and in spring, during a period of maximum seagrass proliferation.
+Other studies have also found *Methylotenera*-specific sequences associated with seagrass roots and leaves indicating that this group members are important constituents of the seagrass microbiome [@Crump2018; @Sanders-Smith2020].
+Genomic and physiological analyses of cultivated *Cellvibrionaceae* and *Reinekea* members showed the capabilities to use important algal polysaccharides [@Xie2017; @Avci2017] indicating their possible involvement into the degradation of *C.
+cylindracea* polysaccharides and/or the control of its algal epiphytes.
+
+The epiphytic microbial community associated with marine macrophytes is undergoing seasonal changes that can be attributed to the fluctuations of environmental conditions, the growth cycle of macrophytes inhabiting temperate zones or the combined effect of both.
+In the present study, we could identify in analysed high rank taxa phylogenetic groups present throughout the year, comprising most of the sequences and a lower proportion of taxa showing seasonal patterns connected to the macrophyte growth cycle (Figures \ref{community} and \ref{desulfo}).
 Studies focusing on functional comparisons between communities associated with different hosts showed that the majority of functions could be found in every community, indicating functional redundancy [@Roth-Schulze2016].
-This difference between phylogenetic varaibility and functional stability was explained by the lottery hypothesis assuming an initial random colonization step performed by a set of functionally equivalent taxonomic groups [@Burke2011; @Roth-Schulze2016].
+This difference between phylogenetic variability and functional stability has been explained by the lottery hypothesis assuming an initial random colonization step performed by a set of functionally equivalent taxonomic groups [@Burke2011; @Roth-Schulze2016].
 It is possible that functional redundancy is a characteristic of high abundance taxa detected to be present throughout the year, while seasonal and/or host-specific functions are an attribute of taxa displaying successional patterns.
@@
 \interlinepenalty=10000
-\setlength{\emergencystretch}{3.5em}
+\setlength{\emergencystretch}{3.7em}
 <div id="refs"></div>
@@
 ## Figure legends
-**\autoref{matrix}.** \nameref{matrix}
+**\autoref{map}.** \nameref{map}
+
+**\autoref{pcoa}.** \nameref{pcoa}
 
@@
 
-**\autoref{pcoa}.** \nameref{pcoa}
-
 **\autoref{community}.** \nameref{community}
@@
 
+\newpage
 ## Figures
-```{r out.width="70%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal OTUs (Jaccard's similarity coefficient) and shared bacterial and archaeal communities (Bray-Curtis similarity coefficient) between assemblages associated with the surfaces of the macrophytes \\textit{C.
+```{r out.width="100%", fig.align="center", fig.cap="Location of the mixed (\\textit{C.
+nodosa} and \\textit{C.
+cylindracea}) and monospecific (\\textit{C.
+cylindracea}) settlement in the Bay of Funtana, northern Adriatic Sea (© OpenStreetMap contributors, www.openstreetmap.org/copyright).\\label{map}", fig.pos="H"}
+include_graphics(normalizePath("results/figures/map.jpg"))
+```
+
+```{r out.width="55%", fig.align="center", fig.cap="Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
-cylindracea} (mixed and monospecific settlement) and communities in the ambient seawater.\\label{matrix}", fig.pos="H"}
-include_graphics(normalizePath("results/figures/matrix.jpg"))
-```
-
-```{r out.width="85%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal communities (Bray-Curtis similarity coefficient) and shared bacterial and archaeal OTUs (Jaccard's similarity coefficient) between consecutive sampling dates and from the surfaces of the macrophytes \\textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
+Samples from different environments or seasons are labeled in different color and shape.
+The proportion of explained variation by each axis is shown on the corresponding axis in parentheses.\\label{pcoa}", fig.pos="H"}
+include_graphics(normalizePath("results/figures/pcoa_figure.jpg"))
+```
+
+```{r out.width="100%", fig.align="center", fig.cap="Proportion of shared bacterial and archaeal communities (Bray-Curtis similarity coefficient) and shared bacterial and archaeal OTUs (Jaccard's similarity coefficient) between consecutive sampling dates and from the surfaces of the macrophytes \\textit{C.
 nodosa} (mixed settlement) and \\textit{C.
@@
 include_graphics(normalizePath("results/figures/seasonal_shared.jpg"))
-```
-
-```{r out.width="55%", fig.align="center", fig.cap="Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of the macrophytes \\textit{C.
-nodosa} (mixed settlement) and \\textit{C.
-cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
-Samples from the same environment or same season are labeled in different colors.
-The proportion of explained variation by each axis is shown on the corresponding axis in parentheses.\\label{pcoa}", fig.pos="H"}
-include_graphics(normalizePath("results/figures/pcoa_figure.jpg"))
 ```

### 2021-07-09 submission/manuscript.tex
Review done.

@@
 ]{article}
+\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{amssymb,amsmath}
 \usepackage{ifxetex,ifluatex}
@@
 \usepackage[margin=1.0in]{geometry}
-\usepackage{graphicx,grffile}
+\usepackage{graphicx}
 \makeatletter
@@
 \renewcommand{\thefigure}{\arabic{figure}}
-\renewcommand{\figurename}{Fig.}
+\renewcommand{\figurename}{Figure}
 \captionsetup{justification=raggedright,singlelinecheck=false}
@@
 
-\renewcommand{\figureautorefname}{Fig.}
+\renewcommand{\figureautorefname}{Figure}
 
@@
 
-\title{\textbf{Seasonal Dynamics of Epiphytic Microbial Communities on Marine
-Macrophyte Surfaces}}
+\DeclareSIUnit{\litre}{l}
+\ifluatex
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+
+\title{\textbf{Seasonal Dynamics of Epiphytic Microbial Communities on
+Marine Macrophyte Surfaces}}
 \author{}
@@
 
-3. NIOZ, Department of Marine Microbiology and Biogeochemistry, Royal
-Netherlands Institute for Sea Research, Utrecht University, The
+3. Department of Marine Microbiology and Biogeochemistry, Royal
+Netherlands Institute for Sea Research (NIOZ), Utrecht University, The
 Netherlands
@@
 connected to the macrophyte growth cycle. Taxonomic analysis, however,
-showed similar high rank groups in the ambient water and the epiphytic
-communities, with the exception of \emph{Desulfobacterota}, which were
-only found on \emph{Caulerpa cylindracea}. \emph{Cyanobacteria} showed
-seasonal changes while other high rank taxa were present throughout the
-year. Phylogenetic groups present throughout the year constituted most
-of the sequences, while less abundant taxa showed seasonal patterns
-connected to the macrophyte growth cycle. Taken together, epiphytic
-microbial communities of the seagrass \emph{Cymodocea nodosa} and the
-macroalga \emph{Caulerpa cylindracea} appear to be host-specific and
-contain taxa that undergo successional changes.
+showed similar high rank taxa (phyla and classes) in the ambient water
+and the epiphytic communities, with the exception of
+\emph{Desulfobacterota}, which were only found on \emph{Caulerpa
+cylindracea}. \emph{Cyanobacteria} showed seasonal changes while other
+high rank taxa were present throughout the year. In months of high
+\emph{Cyanobacteria} presence the majority of cyanobacterial sequences
+were classified as \emph{Pleurocapsa}. Phylogenetic groups present
+throughout the year (e.g.~\emph{Saprospiraceae},
+\emph{Rhodobacteraceae}, members without known relatives within
+\emph{Gammaproteobacteria}, \emph{Desulfatitalea} and members without
+known relatives within \emph{Desulfocapsaceae}) constituted most of the
+sequences, while less abundant taxa showed seasonal patterns connected
+to the macrophyte growth cycle. Taken together, epiphytic microbial
+communities of the seagrass \emph{Cymodocea nodosa} and the macroalga
+\emph{Caulerpa cylindracea} appear to be host-specific and contain taxa
+that undergo successional changes.
 
@@
 all three domains of life (Egan et al., 2013; Tarquinio et al., 2019).
-Microbes can live within macrophyte tissue as endophytes or can form
-epiphytic communities on surfaces of leaves and thalli (Egan et al.,
-2013; Hollants et al., 2013; Tarquinio et al., 2019). Epiphytic and
-endophytic microbial communities exhibit a close functional relationship
-with the macrophyte host. It was proposed that this close relationship
-constitutes a holobiont, an integrated community where the macrophyte
-organism and its symbiotic partners support each other (Margulis, 1991;
-Egan et al., 2013; Tarquinio et al., 2019).
+Microbes can live within macrophyte tissue as endophytes or form
+epiphytic communities on surfaces of leaves, roots, rhizomes and thalli
+(Egan et al., 2013; Hollants et al., 2013; Aires et al., 2015; Tarquinio
+et al., 2019). Epiphytic and endophytic microbial communities exhibit a
+close functional relationship with the macrophyte host. It has been
+proposed that this close relationship constitutes a holobiont, an
+integrated community where the macrophyte organism and its symbiotic
+partners support each other (Margulis, 1991). In addition, as suggested
+by the hologenome theory endophytic microbes play a critical role in the
+adaptation and evolution of the host species (Aires et al., 2015).
 
 Biofilms of microbial epiphytes can contain diverse taxonomic groups and
-harbor cell abundances from 10\textsuperscript{2} to
+harbour cell abundances from 10\textsuperscript{2} to
 10\textsuperscript{7} \si{\cells\per\cm\squared} (Armstrong et al.,
@@
 Macrophytes can promote growth of associated microbes by nutrient
-exudation, while in return microorganisms may support macrophyte
-performance through improved nutrient availability, phytohormone
-production and protection from toxic compounds, oxidative stress,
-biofouling organisms and pathogens (Egan et al., 2013; Hollants et al.,
-2013; Tarquinio et al., 2019). Besides these positive interactions,
-macrophytes can negatively impact the associated microbes such as
-pathogenic bacteria by producing reactive oxygen species and secondary
-metabolites (Egan et al., 2013; Hollants et al., 2013; Tarquinio et al.,
-2019).
+exudation (Wood and Hayasaka, 1981), while in return microorganisms may
+support macrophyte performance through improved nutrient availability
+(Nielsen et al., 2001; de Oliveira et al., 2012), phytohormone
+production (Matsuo et al., 2003; Celdrán et al., 2012) and protection
+from toxic compounds (Küsel et al., 2006), oxidative stress
+(Sanchez-Amat et al., 2010), biofouling organisms (Dobretsov and Qian,
+2002) and pathogens (Penesyan et al., 2009). Besides these positive
+interactions, macrophytes can negatively impact the associated microbes
+by producing reactive oxygen species (Weinberger, 2007) and secondary
+metabolites (Saha et al., 2011).
 
 All these ecological roles are carried out by a taxonomically diverse
-community of microorganisms. At higher taxonomic ranks a core set of
-macrophyte epiphytic taxa was described consisting mainly of
-\emph{Alphaproteobacteria}, \emph{Gammaproteobacteria},
-\emph{Desulfobacterota}, \emph{Bacteroidota}, \emph{Cyanobacteria},
-\emph{Actinobacteriota}, \emph{Firmicutes}, \emph{Planctomycetota},
-\emph{Chloroflexi} and \emph{Verrucomicrobiota} (Crump and Koch, 2008;
-Tujula et al., 2010; Lachnit et al., 2011). In contrast, at lower
-taxonomic ranks host specific microbial communities were found (Lachnit
-et al., 2011; Roth-Schulze et al., 2016). Recently, it was shown that
-even different morphological niches within the same alga had a higher
-influence on the composition of the bacterial community than
-biogeography or environmental factors (Morrissey et al., 2019). While
-the microbial community composition varies between host species,
-metagenomic analyses revealed that the majority of the microbial
-functions are conserved (Burke et al., 2011a; Roth-Schulze et al.,
-2016). This discrepancy between the microbial taxonomic and functional
-composition might be explained by the lottery hypothesis. It postulates
-that an initial random colonization step takes places from a set of
-functionally equivalent taxonomic groups resulting in taxonomically
-different epiphytic communities sharing a core set of functional genes
-(Burke et al., 2011a; Roth-Schulze et al., 2016). In addition, some of
-the variation in the reported data could be attributed to different
-techniques used in these studies, such as different protocols for
-epiphytic cell detachment and/or DNA isolation, as no standard protocol
-has been yet established to study epiphytic communities (Ugarelli et
-al., 2019; Korlević et al., submitted).
-
-The majority of studies describing macrophyte epiphytic microbial
-communities did not include possible seasonal changes (Crump and Koch,
-2008; Lachnit et al., 2009; Burke et al., 2011b; Roth-Schulze et al.,
-2016; Ugarelli et al., 2019). If seasonal changes were taken into
-account, low temporal frequency, applied methodologies and/or limited
-number of analysed host species did not allow a high taxonomic
-resolution (Tujula et al., 2010; Lachnit et al., 2011; Michelou et al.,
+community of microorganisms. At higher taxonomic ranks (phyla and
+classes) microbial taxa, such as \emph{Alphaproteobacteria},
+\emph{Gammaproteobacteria}, \emph{Bacteroidota} and
+\emph{Cyanobacteria}, have been associated with surfaces of seagrass
+leaves and macroalgal thalli (Crump and Koch, 2008; Tujula et al., 2010;
+Lachnit et al., 2011; Egan et al., 2013; Tarquinio et al., 2019;
+Ugarelli et al., 2019). While similar high rank taxa have been found on
+surfaces of different macrophyte species, in order to describe new
+ecological patterns it is also necessary to focus on lower taxonomic
+ranks (genus and OTUs) which tend to be host-specific (Lachnit et al.,
+2011; Hollants et al., 2013; Roth-Schulze et al., 2016). While the
+microbial community composition can vary between host species,
+metagenomic analyses revealed that the majority of microbial functions
+are conserved, showing that different epiphytic microbial species could
+be functionally similar (Burke et al., 2011a; Roth-Schulze et al., 2016;
+Cúcio et al., 2018). This discrepancy between the microbial taxonomic
+and functional composition might be explained by the lottery hypothesis
+(Sale, 1976). It postulates that an initial random colonization step
+takes place from a set of functionally equivalent taxonomic groups
+resulting in taxonomically different epiphytic communities sharing a
+core set of functional genes (Burke et al., 2011a; Stratil et al., 2013;
+Schmidt et al., 2015; Roth-Schulze et al., 2016).
+
+Seagrasses are know to form close relationships with microbial
+communities associated with the surfaces of leaves, roots and rhizomes
+(Cúcio et al., 2016; Crump et al., 2018; Ugarelli et al., 2019; Ettinger
+and Eisen, 2020; Wang et al., 2020). For different seagrass species a
+distinct microbial community from ambient seawater or bulk sediment has
+been reported, however no species specific communities have been found
+(Cúcio et al., 2016; Crump et al., 2018; Ugarelli et al., 2019). It
+seems that seagrasses are selecting the associated microbial community
+but these microbes have not coevolved with their seagrass plant host.
+Similar to seagrasses, siphonous macroalgae of the genus \emph{Caulerpa}
+are also closely associated with their microbial communities (Aires et
+al., 2013, 2015; Rizzo et al., 2016b; Stabili et al., 2017; Morrissey et
+al., 2019). While some studies have found similar culturable bacterial
+groups associated with the surface of a \emph{Caulerpa} species from
+different geographic locations (Stabili et al., 2017), other have
+reported large compositional differences that were mainly attributed to
+different host species of this genus, biogeography and nutrient levels
+(Morrissey et al., 2019) again raising the question to which extent are
+associated communities host-specific.
+
+Since marine macrophytes in temperate zones are exhibiting seasonal
+changes in growth and physiology (Agostini et al., 2003; Najdek et al.,
+2020) it is important to verify if and how surface associated microbial
+communities are affected by these changes. The majority of studies
+describing macrophyte epiphytic microbial communities have not included
+possible seasonal changes (Crump and Koch, 2008; Lachnit et al., 2009;
+Burke et al., 2011b; Roth-Schulze et al., 2016; Ugarelli et al., 2019).
+If seasonal changes have been taken into account, low temporal
+frequency, applied methodologies and/or limited number of analysed host
+species did not allow for a detailed taxonomic analysis (Tujula et al.,
+2010; Lachnit et al., 2011; Bengtsson et al., 2012; Michelou et al.,
 2013; Miranda et al., 2013; Mancuso et al., 2016). In the present study
-we describe the seasonal dynamics of bacterial and archaeal communities
-on the surfaces of the seagrass \emph{Cymodocea nodosa} and siphonous
-macroalgae \emph{Caulerpa cylindracea} determined on a mostly monthly
-scale. Bacterial and archaeal epiphytes were sampled in a meadow of
-\emph{C. nodosa} invaded by the invasive \emph{C. cylindracea} and in a
-locality of only \emph{C. cylindracea} located in the proximity of the
-seagrass meadow. For comparison, the microbial community of the ambient
-seawater was also characterized.
+we performed a descriptive analysis of seasonal bacterial and archaeal
+community dynamics on the surfaces of the seagrass \emph{Cymodocea
+nodosa}, an abundant seagrass species in the Mediterranean (Short et
+al., 2001), and siphonous macroalga \emph{Caulerpa cylindracea}, one of
+the most invasive macroalgal species (Klein and Verlaque, 2008;
+Boudouresque et al., 2009). Bacterial and archaeal epiphytes were
+sampled in a meadow of \emph{C. nodosa} invaded by the invasive \emph{C.
+cylindracea} and in a locality of only \emph{C. cylindracea} located in
+the proximity of the seagrass meadow. For comparison, the microbial
+community of the ambient seawater was also characterized. The presence
+of both macrophytes in the same area enabled (i) the assessment of
+differences in the bacterial and archaeal communities between host
+species and settlements of \emph{C. cylindracea} and (ii) the evaluation
+of differences between surface associated and free living (ambient
+seawater) communities. In addition, these differences were evaluated on
+a monthly scale providing insight into seasonal changes (iii).
 
@@
 Sampling was performed in the Bay of Funtana, northern Adriatic Sea
-(\ang{45;10;39} N, \ang{13;35;42} E). Leaves and thalli were sampled in
-a meadow of \emph{Cymodocea nodosa} invaded by the invasive
-\emph{Caulerpa cylindracea} (mixed settlement) and in a monospecific
-settlement of \emph{Caulerpa cylindracea} located in the proximity of
-the meadow at approximately monthly intervals from December 2017 to
-October 2018 (\autoref{supp-nseq_notus}). Leaves and thalli were
-collected by diving and transported to the laboratory in containers
-placed on ice and filled with seawater collected at the sampling site.
-Upon arrival to the laboratory, \emph{C. nodosa} leaves were cut into
-sections of 1 -- 2 \si{\cm}, while \emph{C. cylindracea} thalli were cut
-into 5 -- 8 \si{\cm} long sections. Leaves and thalli were washed three
-times with sterile artificial seawater (ASW) to remove loosely attached
-microbial cells. Ambient seawater was collected in 10 \si{\l} containers
-by diving and transported to the laboratory where 10 -- 20 \si{\l} were
-filtered through a 20 \si{\um} net. The filtrate was further
-sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate
-membrane filters (Whatman, United Kingdom) using a peristaltic pump.
-Filters were briefly dried at room temperature and stored at \num{-80}
-\si{\degreeCelsius}. Seawater samples were also collected approximately
-monthly from July 2017 to October 2018.
+(\ang{45;10;39} N, \ang{13;35;42} E). The sea floor in the bay is partly
+covered by the invasive macroalga \emph{C. cylindracea} that can be
+found in a monospecific settlement or mixed with the seagrass \emph{C.
+nodosa} (\autoref{map}). \emph{C. nodosa} leaves were retrieved from a
+meadow of \emph{C. nodosa} invaded by the invasive \emph{C. cylindracea}
+(mixed settlement; depth, 2 -- 2.5 \si{\m}), while \emph{C. cylindracea}
+thalli were sampled in the same invaded meadow (mixed settlement; depth,
+2 -- 2.5 \si{\m}) and in a monospecific settlement (depth, 1 -- 1.5
+\si{\m}) of \emph{C. cylindracea} located in the proximity (20 -- 50
+\si{\m}) of the invaded meadow at approximately monthly intervals from
+November 2017 to October 2018 (\autoref{supp-nseq_notus}). Leaves and
+thalli were collected by diving and transported to the laboratory in
+containers placed on ice and filled with seawater collected at the
+sampling site. Upon arrival to the laboratory, \emph{C. nodosa} leaves
+were cut into sections of 1 -- 2 \si{\cm}, while \emph{C. cylindracea}
+thalli were cut into 5 -- 8 \si{\cm} long sections. Leaves and thalli
+were washed three times with sterile artificial seawater (ASW) to remove
+loosely attached microbial cells. Ambient seawater was collected in 10
+\si{\l} containers by diving and transported to the laboratory where 10
+-- 20 \si{\l} were filtered through a 20 \si{\um} net. The filtrate was
+further sequentially filtered through 3 \si{\um} and 0.2 \si{\um}
+polycarbonate membrane filters (Whatman, United Kingdom) using a
+peristaltic pump. Filters were briefly dried at room temperature and
+stored at \num{-80} \si{\degreeCelsius}. Seawater samples were also
+collected approximately monthly from July 2017 to October 2018.
 
@@
 DNA from surfaces of \emph{C. nodosa} and \emph{C. cylindracea} was
-isolated using a previously modified and adapted protocol that allows
-for a selective epiphytic DNA isolation (Massana et al., 1997; Korlević
-et al., submitted). Briefly, leaves and thalli are incubated in a lysis
-buffer and treated with lysozyme and proteinase K. Following the
-incubations, the mixture containing lysed epiphytic cells was separated
-from the leaves and thalli and extracted using phenol-chloroform.
-Finally, the extracted DNA was precipitated using isopropanol. DNA from
-seawater picoplankton was extracted from 0.2 \si{\um} polycarbonate
-filters according to Massana et al. (1997) with a slight modification.
-Following the phenol-chloroform extraction, 1/10 of chilled 3
-\si{\Molar} sodium acetate (pH 5.2) was added. DNA was precipitated by
-adding 1 volume of chilled isopropanol, incubating the mixtures
-overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 20,000 ×
-\emph{g} and 4 \si{\degreeCelsius} for 21 \si{\minute}. The pellet was
-washed twice with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and
-centrifuged after each washing step at 20,000 × \emph{g} and 4
-\si{\degreeCelsius} for 5 \si{\minute}. Dried pellets were re-suspended
-in 50 -- 100 \si{\ul} of deionized water.
+isolated from a pool of leaves or thalli on the sampling day using a
+previously modified and adapted protocol that allows for a selective
+epiphytic DNA isolation (Massana et al., 1997; Korlević et al., 2021).
+Briefly, leaves and thalli were incubated in a lysis buffer and treated
+with lysozyme and proteinase K. Following the incubations, the mixture
+containing lysed epiphytic cells was separated from the leaves and
+thalli and extracted using phenol-chloroform. Finally, the extracted DNA
+was precipitated using isopropanol. DNA from seawater picoplankton was
+extracted from 0.2 \si{\um} polycarbonate filters according to Massana
+et al. (1997) with a slight modification. Following the
+phenol-chloroform extraction, 1/10 of 3 \si{\Molar} sodium acetate (pH
+5.2) was added. DNA was precipitated by adding 1 volume of chilled
+isopropanol, incubating the mixtures overnight at \num{-20}
+\si{\degreeCelsius} and centrifuging at 20,000 × \emph{g} and 4
+\si{\degreeCelsius} for 20 \si{\minute}. The pellet was washed twice
+with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged
+after each washing step at 20,000 × \emph{g} and 4 \si{\degreeCelsius}
+for 5 \si{\minute}. Dried pellets were re-suspended in 50 -- 100
+\si{\ul} of deionized water.
 
@@
 Illumina MiSeq sequencing of the V4 region of the 16S rRNA gene was
-performed as described previously (Korlević et al., submitted). The V4
-region of the 16S rRNA gene was amplified using a two-step PCR
-procedure. In the first PCR, the 515F
-(\(5'\)-GTGYCAGCMGCCGCGGTAA-\(3'\)) and 806R
+performed as described previously (Korlević et al., 2021). The V4 region
+of the 16S rRNA gene was amplified using a two-step PCR procedure. In
+the first PCR, the 515F (\(5'\)-GTGYCAGCMGCCGCGGTAA-\(3'\)) and 806R
 (\(5'\)-GGACTACNVGGGTWTCTAAT-\(3'\)) primers from the Earth Microbiome
@@
 obtained in this study have been deposited in the European Nucleotide
-Archive (ENA) at EMBL-EBI under accession number PRJEB37267.
-
-\hypertarget{sequence-analysis}{%
-\subsubsection{Sequence analysis}\label{sequence-analysis}}
+Archive (ENA) at EMBL-EBI under the accession number PRJEB37267
+(\url{https://www.ebi.ac.uk/ena/browser/view/PRJEB37267}).
+
+\hypertarget{sequence-and-data-analysis}{%
+\subsubsection{Sequence and data
+analysis}\label{sequence-and-data-analysis}}
 
@@
 \url{http://www.arb-silva.de}) was used (Quast et al., 2013; Yilmaz et
-al., 2014). Pipeline data processing and visualization was done using R
-(version 3.6.0) (R Core Team, 2019) combined with packages vegan
-(version 2.5-6) (Oksanen et al., 2019), tidyverse (version 1.3.0)
-(Wickham et al., 2019) and multiple other packages (Neuwirth, 2014; Xie,
-2014, 2015, 2019a, 2019b; Wilke, 2018; Xie et al., 2018; Allaire et al.,
-2019; Zhu, 2019). The detailed analysis procedure including the R
-Markdown file are available in the GitHub repository
+al., 2014). Sequences were clustered into operational taxonomic units
+(OTUs) at a similarity level of 97 \si{\percent}.
+
+Pipeline data processing and visualization was done using R (version
+3.6.0) (R Core Team, 2019) combined with packages vegan (version 2.5-6)
+(Oksanen et al., 2019), tidyverse (version 1.3.0) (Wickham, 2017;
+Wickham et al., 2019) and multiple other packages (Neuwirth, 2014; Xie,
+2014, 2015, 2019a, 2019b, 2019c; Xie et al., 2018; Allaire et al., 2019;
+Wilke, 2019; Zhu, 2019). Observed number of OTUs, Chao1, ACE,
+exponential of the Shannon diversity index and Inverse Simpson index
+were calculated after normalization to the minimum number of reads per
+sample using vegan's function \texttt{rrarefy} to account for different
+sequencing depths (Oksanen et al., 2019). Chao1 and ACE estimators were
+calculated using vegan's function \texttt{estimateR}, while Shannon and
+Inverse Simpson diversity indices were retrieved using vegan's function
+\texttt{diversity} (Oksanen et al., 2019). To express both diversity
+indices in terms of effective number of OTUs the exponential of the
+Shannon diversity index was calculated (Jost, 2006). The proportions of
+shared OTUs and communities between samples and community types
+(seawater, \emph{C. nodosa} {[}mixed{]}, \emph{C. cylindracea}
+{[}mixed{]} and \emph{C. cylindracea} {[}monospecific{]}) were expressed
+as the Jaccard's (on presence/absence data) and Bray-Curtis similarity
+coefficient, respectively. The coefficients were calculated on the OTU
+data table using vegan's function \texttt{vegdist} and converted from
+dissimilarities to similarities (Borcard et al., 2011; Legendre and
+Legendre, 2012; Oksanen et al., 2019). The Principal Coordinates
+Analysis (PCoA) was performed on Bray-Curtis dissimilarities based on
+OTU abundances using the function \texttt{cmdscale} (Borcard et al.,
+2011; Legendre and Legendre, 2012). Differences between communities were
+tested by performing the Analysis of Similarities (ANOSIM) using the
+vegan's function \texttt{anosim} and 1000 permutations (Oksanen et al.,
+2019), while differences in relative contributions or proportions of
+shared OTUs and communities were tested by applying the Mann--Whitney
+\emph{U} test using the function \texttt{wilcox.test}. In addition,
+differences between community type estimators or indices were tested by
+performing the Kruskal-Wallis \emph{H} test (function
+\texttt{kruskal.test}) followed by a pairwise comparison using the
+Mann-Whitney \emph{U} test (function \texttt{pairwise.wilcox}).
+Bonferroni correction was used to address the problem of multiple
+comparisons.
+
+A total of 1.7 million sequences after quality curation and exclusion of
+sequences without known relatives (no relative sequences) and
+eukaryotic, chloroplast and mitochondrial sequences were obtained
+(\autoref{supp-nseq_notus}). The number of reads per sample ranged
+between 8,408 and 77,463 sequences (\autoref{supp-nseq_notus}). Even
+when the highest sequencing effort was applied the rarefaction curves
+did not level off as commonly observed in high-throughput 16S rRNA
+amplicon sequencing approaches (\autoref{supp-rarefaction}). Following
+quality curation and exclusion of sequences as mentioned above reads
+were clustered into 28,750 different OTUs. Read numbers were normalized
+to the minimum number of sequences (8,408, \autoref{supp-nseq_notus})
+using previously mentioned vegan's function \texttt{rrarefy} resulting
+in 17,201 different OTUs ranging from 352 to 2,062 OTUs per sample
+(\autoref{supp-calculators}). Based on the ATCC MSA-1002 mock community
+included in the analysis an average sequencing error rate of 0.01
+\si{\percent} was determined, which is in line with previously reported
+values for next-generation sequencing data (Kozich et al., 2013; Schloss
+et al., 2016). In addition, the negative controls processed together
+with the samples yielded only 2 sequences after sequence quality
+curation. The detailed analysis procedure including the R Markdown file
+is available as a GitHub repository
 (\url{https://github.com/MicrobesRovinj/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021}).
-Based on the ATCC MSA-1002 mock community included in the analysis an
-average sequencing error rate of 0.01 \si{\percent} was determined,
-which is in line with previously reported values for next-generation
-sequencing data (Kozich et al., 2013; Schloss et al., 2016). In
-addition, the negative controls processed together with the samples
-yielded on average only 2 sequences after sequence quality curation.
 
@@
 
-Sequencing of the 16S rRNA V4 region yielded a total of 1.7 million
-sequences after quality curation and exclusion of sequences without
-known relatives (no relative sequences) and eukaryotic, chloroplast and
-mitochondrial sequences (\autoref{supp-nseq_notus}). A total of 35
-samples originating from epiphytic archaeal and bacterial communities
-associated with surfaces of the seagrass \emph{C. nodosa} and the
-macroalga \emph{C. cylindracea} were analysed. In addition, 18 samples
-(one of the samples was sequenced twice) originating from the ambient
-seawater were also processed for comparison. The number of reads per
-sample ranged between 8,408 and 77,463 sequences
-(\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
-applied the rarefaction curves did not level off as commonly observed in
-high-throughput 16S rRNA amplicon sequencing approaches
-(\autoref{supp-rarefaction}). Following quality curation and exclusion
-of sequences as mentioned above reads were clustered into 28,750
-different OTUs at a similarity level of 97 \si{\percent}. Read numbers
-were normalized to the minimum number of sequences (8,408,
-\autoref{supp-nseq_notus}) through rarefaction resulting in 17,201
-different OTUs with 352 to 2,062 OTUs per sample
-(\autoref{supp-calculators}). To determine seasonal changes in richness
-and diversity the observed number of OTUs, Chao1, ACE, Exponential
-Shannon and Inverse Simpson (Jost, 2006) were calculated after
-normalization through rarefaction. Generally, richness estimators and
-diversity indices showed similar trends. On average, higher values were
-found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,688.4 ±
-136.6 OTUs{]} and monospecific {[}Number of OTUs, 1,750.4 ± 165.7
-OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,063.7 ± 210.6
-OTUs) and lowest values were obtained for the microbial community of the
-ambient seawater (Number of OTUs, 531.0 ± 143.9 OTUs)
-(\autoref{supp-calculators}). Seasonal changes did not reveal such large
-dissimilarities. \emph{C. nodosa} communities showed a slow increase
-towards the end of the study, while \emph{C. cylindracea} (mixed and
-monospecific) communities were characterized by slightly higher values
-in spring and summer than in autumn and winter
-(\autoref{supp-calculators}).
-
-To determine the proportion of shared archaeal and bacterial OTUs and
-communities sampled in different environments the Jaccard's Similarity
-Coefficient on presence-absence data and Bray-Curtis Similarity
-Coefficient, respectively, were calculated. Coefficients were determined
-after normalization through rarefaction and binning of samples from the
-particular environment. The highest proportion of shared OTUs and
+A total of 35 samples originating from epiphytic archaeal and bacterial
+communities associated with surfaces of the seagrass \emph{C. nodosa}
+and the macroalga \emph{C. cylindracea} were analysed. In addition, 18
+samples (one of the samples was sequenced twice) originating from the
+ambient seawater were also processed for comparison. Generally, richness
+estimators and diversity indices showed similar trends. On average,
+higher values were found for \emph{C. cylindracea} (mixed {[}Number of
+OTUs, 1,688.4 ± 136.6 OTUs{]} and monospecific {[}Number of OTUs,
+1,750.4 ± 165.7 OTUs{]}) than for \emph{C. nodosa} (Number of OTUs,
+1,063.7 ± 210.6 OTUs) and lowest values were obtained for the microbial
+community of the ambient seawater (Number of OTUs, 531.0 ± 143.9 OTUs)
+(Kruskal-Wallis, \emph{p} \textless{} 0.0001)
+(\autoref{supp-calculators} and Tables
+\ref{supp-calculators_community_type} and
+\ref{supp-calculator_statistics}). Temporal changes did not reveal such
+large dissimilarities. \emph{C. nodosa} communities showed a slow
+increase in all calculated richness estimators towards the end of the
+study, while \emph{C. cylindracea} (mixed and monospecific) communities
+were characterized by slightly higher values in spring and summer than
+in autumn and winter (\autoref{supp-calculators}).
+
+A clear separation between ambient seawater and surface associated
+communities was found (\autoref{pcoa}). In addition, a separation of
+epiphytic bacterial and archaeal communities based on host species was
+detected. This separation was further supported by ANOSIM (R = 0.96,
+\emph{p} \textless{} 0.001). The highest proportion of shared OTUs and
 community was found between mixed and monospecific \emph{C. cylindracea}
 (Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared values were
-calculated between ambient seawater and epiphytic communities
-(\autoref{matrix}). Shared proportion between \emph{C. nodosa} and
-\emph{C. cylindracea} were approximately in-between the values of mixed
-and monospecific \emph{C. cylindracea}. To assess seasonal changes in
-the proportion of shared OTUs and communities the Jaccard's and
-Bray-Curtis Similarity Coefficients were calculated between consecutive
-sampling points (\autoref{shared}). Both coefficients showed similar
-trends. Temporal proportional changes were more pronounced for ambient
-seawater than for \emph{C. nodosa} and especially \emph{C. cylindracea}
-associated communities (\autoref{shared}). In addition, only 0.4 -- 1.0
-\si{\percent} of OTUs from each surface associated community were
-present at all seasons. These persistent OTUs constituted a high
-proportion of total sequences (40.2 -- 53.2 \si{\percent}). To further
-disentangle the environmental and seasonal community dissimilarity a
-Principal Coordinates Analysis (PCoA) based on Bray-Curtis distances and
-OTU abundances was applied. A clear separation between ambient seawater
-and surface associated communities was found (\autoref{pcoa}). In
-addition, a separation of epiphytic bacterial and archaeal communities
-based on host species was detected. This separation was further
-supported by ANOSIM (R = 0.96, \emph{p} \textless{} 0.001). Seasonal
-changes of \emph{C. nodosa} associated communities indicated a
-separation between spring, summer and autumn/winter samples (ANOSIM, R =
-0.56, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
-communities a separation between summer and autumn/winter/spring samples
-was observed that was, however, not as strong as for \emph{C. nodosa}
-associated communities (ANOSIM, R = 0.30, \emph{p} \textless{} 0.01)
-(\autoref{pcoa}).
+calculated between ambient seawater and epiphytic communities (Jaccard,
+0.10 -- 0.11; Bray-Curtis, 0.05 -- 0.06). Shared proportions of OTUs and
+communities between \emph{C. nodosa} and \emph{C. cylindracea} (either
+mixed or monospecific) were approximately in-between the values obtained
+for the comparison of ambient seawater with all other communities and
+for the comparison of the mixed and monospecific \emph{C. cylindracea}
+associated community. Seasonal changes of \emph{C. nodosa} associated
+communities indicated a separation between spring, summer and
+autumn/winter samples (ANOSIM, R = 0.56, \emph{p} \textless{} 0.01). For
+\emph{C. cylindracea} associated communities a separation between summer
+and autumn/winter/spring samples was observed that was, however, not as
+strong as for \emph{C. nodosa} associated communities (ANOSIM, R = 0.30,
+\emph{p} \textless{} 0.05) (\autoref{pcoa}). Shared proportions of OTUs
+between consecutive sampling points were lower for ambient seawater
+(19.6 ± 2.5 \si{\percent}) than for \emph{C. nodosa} (28.3 ± 5.2
+\si{\percent}) and \emph{C. cylindracea} (mixed {[}26.3 ± 2.1
+\si{\percent}{]} and monospecific {[}27.2 ± 2.0 \si{\percent}{]})
+associated communities (\emph{p} \textless{} 0.0001), while mean
+proportions of shared communities between consecutive sampling points
+did not show such a difference (seawater, 57.4 ± 14.7 \si{\percent};
+\emph{C. nodosa}, 53.4 ± 9.3 \si{\percent}; \emph{C. cylindracea}
+{[}mixed{]}, 55.0 ± 7.0 \si{\percent}; \emph{C. cylindracea}
+{[}monospecific{]}, 55.1 ± 5.2 \si{\percent}) (\emph{p} = 0.1), although
+in ambient seawater higher fluctuations could be observed
+(\autoref{shared}). In addition, only 0.4 -- 1.0 \si{\percent} of OTUs
+from each surface associated community were present at all seasons.
+These persistent OTUs constituted a high proportion of total sequences
+(40.2 -- 53.2 \si{\percent}) and were mainly contributing to abundant
+phylogenetic groups present throughout the year, e.g.~the no relative
+\emph{Rhodobacteraceae} in the case of \emph{C. nodosa} or taxa within
+\emph{Desulfobacterota} in the case of \emph{C. cylindracea} (see below)
+(\autoref{supp-core_otus_taxonomy_table}).
 
 The taxonomic composition of both, macrophyte associated and ambient
-seawater communities was dominated by bacterial (99.1 ± 2.1
+seawater community, was dominated by bacterial (99.1 ± 2.1
 \si{\percent}) over archaeal sequences (0.9 ± 2.1 \si{\percent})
@@
 higher values in autumn/winter (37.2 ± 11.2 \si{\percent}) than in
-spring/summer (20.9 ± 9.7 \si{\percent}) (\autoref{supp-chloroplast}).
-Generally, at higher taxonomic ranks (phylum-class), epiphytic and
-ambient seawater microbial communities were composed of similar
-bacterial taxa. Ambient seawater communities were mainly comprised of
-\emph{Actinobacteriota}, \emph{Bacteroidota}, \emph{Cyanobacteria},
-\emph{Alphaproteobacteria}, \emph{Gammaproteobacteria} and
-\emph{Verrucomicrobiota}. Communities associated with \emph{C. nodosa}
-consisted additionally of \emph{Planctomycetota} contributing more in
-summer 2018 than in other seasons. In addition, communities from mixed
-and monospecific \emph{C. cylindracea} were similar and characterized by
-the same groups as ambient seawater and \emph{C. nodosa} communities
-with the addition of \emph{Desulfobacterota} (\autoref{community}).
-Larger differences between environments and host species were observed
-at lower taxonomic ranks (Figs. \ref{cyano} -- \ref{desulfo}).
+spring/summer (20.9 ± 9.7 \si{\percent}) (\emph{p} \textless{} 0.0001)
+(\autoref{supp-chloroplast}). Generally, at higher taxonomic ranks
+(phylum-class), epiphytic and ambient seawater microbial communities
+were composed of similar bacterial taxa. Ambient seawater communities
+were mainly comprised of \emph{Actinobacteriota}, \emph{Bacteroidota},
+\emph{Cyanobacteria}, \emph{Alphaproteobacteria},
+\emph{Gammaproteobacteria} and \emph{Verrucomicrobiota}. Communities
+associated with \emph{C. nodosa} consisted additionally of
+\emph{Planctomycetota} contributing more in summer 2018 than in other
+seasons. In addition, communities from mixed and monospecific \emph{C.
+cylindracea} were similar and characterized by the same groups as
+ambient seawater and \emph{C. nodosa} communities with the addition of
+\emph{Desulfobacterota} (\autoref{community}). Larger differences
+between environments and host species were observed at lower taxonomic
+ranks (Figures \ref{cyano} -- \ref{desulfo}).
 
@@
 cylindracea} mixed (7.7 ± 3.9 \si{\percent}) and monospecific (7.8 ± 2.4
-\si{\percent}) associated communities in autumn and for ambient seawater
-communities in winter (8.8 ± 7.5 \si{\percent}). Large taxonomic
-differences between surface associated and ambient seawater
-cyanobacterial communities were observed. Ambient seawater communities
-were mainly comprised of \emph{Cyanobium} and \emph{Synechococcus},
-while surface associated communities were comprised of
-\emph{Pleurocapsa} and sequences within the class \emph{Cyanobacteriia}
-that could not be further classified (no relative \emph{Cyanobacteriia})
-(\autoref{cyano}). In addition, seasonal changes in surface associated
-communities were observed in \emph{Pleurocapsa} and no relative
-\emph{Cyanobacteriia} comprising larger proportions in autumn and winter
-and \emph{Acrophormium}, \emph{Phormidesmis} and sequences without known
+\si{\percent}) associated communities in autumn (\emph{p} \textless{}
+0.0001) and for ambient seawater communities in winter (8.8 ± 7.5
+\si{\percent}). Large taxonomic differences between surface associated
+and ambient seawater cyanobacterial communities were observed. Ambient
+seawater communities were mainly comprised of \emph{Cyanobium} and
+\emph{Synechococcus}, while surface associated communities were
+comprised of \emph{Pleurocapsa} and sequences within the class
+\emph{Cyanobacteriia} that could not be further classified (no relative
+\emph{Cyanobacteriia}) (\autoref{cyano}). In addition, seasonal changes
+in surface associated cyanobacterial communities were observed with
+\emph{Pleurocapsa} and no relative \emph{Cyanobacteriia} comprising
+larger proportions of \emph{Cyanobacteria} in autumn and winter and
+\emph{Acrophormium}, \emph{Phormidesmis} and sequences without known
 relatives within the \emph{Nodosilineaceae} (no relative
@@
 Sequences classified as \emph{Bacteroidota} comprised, on average, 19.2
-± 5.5 \si{\percent} of all sequences (\autoref{bactero}). Similarly to
+± 5.5 \si{\percent} of all sequences (\autoref{bactero}). Similar to
 \emph{Cyanobacteria}, large differences in the taxonomic composition
@@
 relatives within the \emph{Saprospiraceae} (no relative
-\emph{Saprospiraceae}), uncultured \emph{Saprospiraceae}, (sequences
+\emph{Saprospiraceae}), uncultured \emph{Saprospiraceae}, sequences
 without known relatives within the \emph{Flavobacteriaceae} (no relative
@@
 \emph{Hyphomonadaceae}). Representatives of no relative
-\emph{Rhodobacteraceae} comprised on average 40.6 ± 23.2 \si{\percent}
+\emph{Rhodobacteraceae} comprised on average 54.7 ± 11.5 \si{\percent}
 of all alphaproteobacterial sequences in the epiphytic community
@@
 
-Surfaces of marine macrophytes harbor biofilms consisting of diverse
-microbial taxa (Egan et al., 2013; Tarquinio et al., 2019). No standard
-protocol has been developed to study these macophyte-associated microbes
-(Ugarelli et al., 2019). Different procedures for removal of microbial
-cells from host surfaces are described, such as host tissue shaking
-(Nõges et al., 2010), scraping (Uku et al., 2007), swabbing (Mancuso et
-al., 2016) and ultrasonication (Cai et al., 2014). All these methods
-result in different removal efficiencies but none was enabling a
-complete removal of attached microbial cells based on our experience. In
-the present study, we applied a removal protocol (Korlević et al.,
-submitted) based on direct cellular lysis (Burke et al., 2009). The
-application of a direct lysis procedure coupled with a high sampling
-frequency and Illumina amplicon sequencing has enabled us to described
-in detail the bacterial and archaeal communities associated with the
+In the present study, we applied a selective epiphytic DNA isolation
+procedure based on direct cellular lysis (Korlević et al., 2021) coupled
+with a monthly sampling and Illumina amplicon sequencing to describe in
+detail the bacterial and archaeal communities associated with the
 surfaces of two marine macrophytes, \emph{C. nodosa} and \emph{C.
-cylindracea}.
-
-In the present study, highest richness was observed for \emph{C.
-cylindracea} (mixed and monospecific) followed by \emph{C. nodosa} and
-lowest richness was found in ambient seawater microbial communities.
-Higher richness of microbial communities associated with seagrasses than
-in ambient seawater was described earlier and could be attributed to a
-larger set of inhabitable microniches existing on macrophyte surfaces
-than in the ambient seawater (Ugarelli et al., 2019). The highest
-richness observed for \emph{C. cylindracea} might be partly due to its
-contact with the sediment. The stolon of \emph{C. cylindracea} is
-attached to the sediment surface with rhizoids and thus, the stolon and
-rhizoids are in a direct contact with the sediment. In addition,
-seasonal differences in richness observed for surface attached
-communities indicated a slightly higher richness in spring and summer.
-This pattern could be explained by a higher macrophyte growth in these
-two seasons than in autumn and winter (M. Najdek, personal
-communication; Zavodnik et al., 1998; Ruitton et al., 2005). During
-their main growth season in spring and summer macrophytes exhibit a more
-dynamic chemical interaction with the surface community probably causing
-an increase in the number of inhabitable microniches (Borges and
-Champenois, 2015; Rickert et al., 2016).
+cylindracea}. Highest richness was observed for \emph{C. cylindracea}
+(mixed and monospecific) followed by \emph{C. nodosa} and lowest
+richness was found in ambient seawater microbial communities. Higher
+richness of microbial communities associated with macrophytes than in
+ambient seawater has been described earlier (Mancuso et al., 2016;
+Ugarelli et al., 2019) and could be attributed to a larger set of
+inhabitable microniches existing on macrophyte surfaces than in the
+ambient seawater. The highest richness observed for \emph{C.
+cylindracea} might be partly due to its contact with the sediment. The
+stolon of \emph{C. cylindracea} is attached to the sediment surface with
+rhizoids and thus, the stolon and rhizoids are in direct contact with
+the sediment. Also, studies have shown that the presence of \emph{C.
+cylindracea} can alter the content and biochemical composition of
+sedimentary organic matter (Pusceddu et al., 2016; Rizzo et al., 2017,
+2020) possibly further expanding the number of inhabitable microniches
+and thus causing the observed increase in richness. Seasonal differences
+in richness observed for surface attached communities indicated a
+slightly higher richness in spring and summer. This pattern could be
+explained by a more intense macrophyte growth in these two seasons than
+in autumn and winter (Zavodnik et al., 1998; Ruitton et al., 2005;
+Najdek et al., 2020). During their main growth season in spring and
+summer macrophytes exhibit a more dynamic chemical interaction with the
+surface community probably causing an increase in the number of
+inhabitable microniches (Borges and Champenois, 2015; Rickert et al.,
+2016). Proportions of shared epiphytic OTUs between consecutive sampling
+points were low also indicated by the proportion of OTUs (\(\leq\) 1.0
+\si{\percent}) present at every sampling date (\autoref{shared}). These
+persistent OTUs, however, accounted for a high proportion of sequences
+(\(\geq\) 40.2 \si{\percent}), as is often the case with similar
+high-frequency sampling studies (Gilbert et al., 2009; Gilbert et al.,
+2012). In comparison to the seawater community, higher values of shared
+OTUs between consecutive sampling points were observed for the
+macrophyte surface associated communities. It appears that macrophyte
+surfaces are providing more stable conditions than the ambient seawater.
 
 We observed a strong differentiation between the surface attached and
-ambient seawater communities at the level of OTUs, in agreement with
-most published studies (Burke et al., 2011b; Michelou et al., 2013;
+ambient seawater communities at the level of OTUs which is in agreement
+with most published studies (Burke et al., 2011b; Michelou et al., 2013;
 Mancuso et al., 2016; Roth-Schulze et al., 2016; Crump et al., 2018;
-Ugarelli et al., 2019). This indicates that marine macrophytes are a
-selecting factor from the pool of microbial taxa present in the ambient
-seawater, modifying the microbial community once the macrophyte
-associated microbial biofilm develops (Salaün et al., 2012; Michelou et
-al., 2013). In contrast, Fahimipour et al. (2017) report in a global
-study of \emph{Zostera marina}, similarities between the microbial
-community developed on leaves and in the ambient seawater. The
-discrepancy between our data and the study of Fahimipour et al. (2017)
-could be explained by different seagrass species, methodological
-variations or biogeographic trends as Fahimipour et al. (2017) analysed
-samples from different locations throughout the northern hemisphere. It
-is possible that the microbial communities in ambient seawater and on
-leaves from the same location are differing but are still more similar
-to each other when compared to other sampling locations. Indeed, it was
-found that prokaryotic communities vary substantially between different
-sampling sites (Bengtsson et al., 2017). When the taxonomic composition
-at high ranks was analysed no such strong differentiation was noticed.
-Phyla and classes such as \emph{Actinobacteriota}, \emph{Bacteroidota},
-\emph{Cyanobacteria}, \emph{Alphaproteobacteria},
-\emph{Gammaproteobacteria} and \emph{Verrucomicrobiota} were found in
-both ambient seawater as well as macrophyte associated, in agreement
-with previous studies (Burke et al., 2011b; Egan et al., 2013; Michelou
-et al., 2013). In contrast, when low taxonomic ranks were analysed
-(i.e., family and genus) a strong differentiation was observed (Figs.
-\ref{cyano} -- \ref{desulfo}). A similar differentiation at lower
-microbial taxonomic ranks between ambient seawater and macrophytes was
-described for other macrophyte species as well (Egan et al., 2013;
-Michelou et al., 2013; Ugarelli et al., 2019).
-
-Beside differences between ambient seawater and surface associated
-microbial communities, it is unclear whether the prokaryotic epiphytic
-community is host-specific or whether there are generalist taxa
-characteristic to all or many macrophytes (Egan et al., 2013). Similar
-to previously described differences between microbial communities in the
-ambient seawater and on macrophytes, at high taxonomic ranks no major
-difference between the microbial communities associated with different
-hosts was observed. The only high rank phylum that was differing between
-\emph{C. nodosa} and \emph{C. cylindracea} was \emph{Desulfobacterota},
-with more abundant sequences in the \emph{C. cylindracea} associated
-community. As already mentioned, the rhizoids and part of the stolon are
-in contact with the sediment. Thus \emph{Desulfobacterota} are probably
-a part of the epiphytic community that was in contact with the sediment.
-Similar high rank taxa found in this study were described to be specific
-for other species of macrophytes (Burke et al., 2011b; Lachnit et al.,
-2011; Mancuso et al., 2016; Bengtsson et al., 2017). In contrast to high
-taxonomic ranks, a substantial differentiation between host specific
-communities was found supporting the notion that macrophyte associated
-microbial communities might be host-specific. Host-specificity was also
-observed for other species of macroalgae and seagrasses (Lachnit et al.,
-2011; Roth-Schulze et al., 2016; Morrissey et al., 2019; Ugarelli et
-al., 2019). Taken together, at high taxonomic ranks a core set of taxa
-could be described that is characteristic for all or many macrophytes,
-while at low taxonomic ranks a community specific to host species was
-identified (Figs. \ref{pcoa} and \ref{community}) (Egan et al., 2013).
-
-Seasonal changes in richness in the epiphytic community were substantial
-as indicated by the proportion of OTUs (\(\leq\) 1.0 \si{\percent})
-present at every sampling date. These persistent OTUs, however, were
-accounting for a high proportion of sequences (\(\geq\) 40.2
-\si{\percent}) (\autoref{shared}). A very similar proportion of
-persistent OTUs was reported in high-frequency sampling studies
-describing seasonal changes in picoplankton (Gilbert et al., 2009,
-2012). In comparison to the seawater community, a lower degree of
-seasonal shifts was observed for the macrophyte surface associated
-communities. It appears that microniches at the surfaces of macrophytes
-are providing more stable conditions than the ambient seawater. At the
-level of OTUs seasonal changes of \emph{C. nodosa} and \emph{C.
-cylindracea} associated communities were identified that could be linked
-to the growth cycle of the seagrass and macroalgae (M. Najdek, personal
-communication). \emph{C. nodosa} was characterized by a spring community
-during maximum seagrass proliferation, a summer community during the
-highest standing stock of \emph{C.nodosa} and an autumn/winter community
-during the decay of seagrass biomass. In contrast, \emph{C. cylindracea}
+Ugarelli et al., 2019; Sanders-Smith et al., 2020). This indicates that
+marine macrophytes are selecting microorganisms from the pool of
+microbial taxa present in the ambient seawater, modifying the microbial
+community once the macrophyte associated microbial biofilm develops
+(Salaün et al., 2012; Michelou et al., 2013). In addition, similar to
+the study of Roth-Schulze et al. (2016) seagrass and macroalgae specific
+microbial communities were identified, while no difference between
+\emph{C. cylindracea} settlements was observed indicating that seagrass
+and macroalgae specific metabolism is involved in the selection and
+development of the associated biofilm. At the level of OTUs seasonal
+changes of \emph{C. nodosa} and \emph{C. cylindracea} associated
+communities were identified that could be linked to the growth cycle of
+the seagrass and macroalgae (Agostini et al., 2003; Najdek et al.,
+2020). \emph{C. nodosa} was characterized by a spring community during
+maximum seagrass proliferation, a summer community during the highest
+standing stock of \emph{C. nodosa} and an autumn/winter community during
+the decay of seagrass biomass. In contrast, \emph{C. cylindracea}
 started to proliferate in late spring and was characterized only by a
@@
 community when the biomass was at the peak and decaying thereafter.
-Similar seasonal changes in the epiphytic community were also described
-for other macroalgae (Tujula et al., 2010; Lachnit et al., 2011). Higher
-seasonal stability of \emph{C. cylindracea} surface communities than in
-\emph{C. nodosa} was also observed in the higher proportion of shared
-communities between two consecutive sampling dates in \emph{C.
-cylindracea}.
-
-Chloroplast sequence abundances were higher in autumn/winter than in
-spring/summer. This pattern is not surprising as seagrasses harbor more
-algal epiphytes during autumn/winter than in spring/summer (Reyes and
-Sansón, 2001). Furthermore, we used an adapted DNA isolation protocol
-that is known to partially co-extract DNA from planktonic eukaryotes
-(Korlević et al., 2015). Strong seasonal fluctuations of high rank
-epiphytic taxa were not observed, with the exception of
-\emph{Cyanobacteria}. Cyanobacterial sequences were more pronounced in
-November and December than in spring and summer. In the months of high
-cyanobacterial sequence abundances the majority of sequences from this
-group were classified as \emph{Pleurocapsa}, a group known to colonized
-different living and non-living surfaces (Burns et al., 2004; Longford
-et al., 2007; Mobberley et al., 2012; Reisser et al., 2014). It is
-possible that during periods of low metabolic activity there is a
-reduced interaction and selection of the epiphytic community by the
-seagrass, causing leaves to become a suitable surface for non-specific
-colonizers (Zavodnik et al., 1998). \emph{Pleurocapsa} was replaced in
-spring and summer by \emph{Acrophormium}, \emph{Phormidesmis} and
-sequences without known relatives within the \emph{Nodosilineaceae}. A
-study of coastal microbial mats found also a higher proportion of
-\emph{Nodosilineaceae} sequences in summer, while \emph{Phormidesmis}
-sequences were at their peak in autumn (Cardoso et al., 2019). Other
-high rank taxa did not exhibit strong successional patterns. In every
-analysed group, with the exception of \emph{Desulfobacterota}, taxa
-present throughout the year in similar proportions and season specific
-taxa could be identified (Figs. \ref{bactero} and \ref{desulfo}). Within
-\emph{Bacteroidota} different groups within the family
-\emph{Saprospiraceae} (e.g.~\emph{Lewinella}, \emph{Portibacter} and
-\emph{Rubidimonas}) were detected across all seasons. Members of this
-family are often found in association with macrophytes and it is
-suggested that they are involved in the hydrolysis and utilization of
-complex carbon sources (Burke et al., 2011b; McIlroy and Nielsen, 2014;
-Crump et al., 2018). In contrast, the families \emph{Flavobacteriaceae}
-and \emph{Rhodothermaceae} showed seasonal patterns, with
-\emph{Flavobacteriaceae} being more pronounced from November to June and
-\emph{Rhodothermaceae} from June to October (\autoref{bactero}). Within
-\emph{Alphaproteobacteria} the family \emph{Rhodobacteraceae} comprised
-the majority of sequences throughout the year (\autoref{alpha}). This
-metabolically versatile family is often associated with macrophyte
-surfaces and usually is one of the most abundant groups (Burke et al.,
-2011b; Michelou et al., 2013; Pujalte et al., 2014; Mancuso et al.,
-2016). In addition, \emph{Hyphomonadaceae} were found in all samples.
-Interestingly, some of the species within this group contain stalks on
-their cells, which can be used to attach to the macrophyte surface
-(Weidner et al., 2000; Abraham and Rohde, 2014). Within the
-\emph{Gammaproteobacteria}, sequences without known representatives were
-the most pronounced group present throughout the year (\autoref{gamma}).
-In addition, \emph{Granulosicoccus} was also found in almost all
-samples. \emph{Gammaproteobacteria} are often a major constituent of
+Similar seasonal changes in the epiphytic community have also been
+described for other macroalgae (Tujula et al., 2010; Lachnit et al.,
+2011).
+
+The taxonomic analysis showed higher chloroplast sequence abundances in
+autumn/winter than in spring/summer. This pattern is not surprising as
+seagrasses harbour more algal epiphytes during autumn/winter than in
+spring/summer (Reyes and Sansón, 2001). Furthermore, we used an adapted
+DNA isolation protocol that is known to partially co-extract DNA from
+planktonic eukaryotes (Korlević et al., 2015). In general, the taxonomic
+analysis identified epiphytic phylogenetic groups present throughout the
+year comprising most of the reads, and taxa present in lower proportions
+showing seasonal patterns. The first group was comprised of members of
+the \emph{Bacteroidota} family \emph{Saprospiraceae}, the
+alphaproteobacterial \emph{Rhodobacteraceae} and \emph{Hyphomonadaceae},
+the gammaproteobacterial genus \emph{Granulosicoccus}, sequences without
+known relatives within \emph{Gammaproteobacteria} and various taxa
+within \emph{Desulfobacterota} (Figures \ref{bactero} -- \ref{desulfo}).
+All these groups were found on all host species, with the exception of
+\emph{Desulfobacterota} that was characteristic for \emph{C.
+cylindracea}. In addition, the persistence of \emph{Rhodobacteraceae} in
+the case of \emph{C. nodosa} and \emph{Desulfobacterota} in the case of
+\emph{C. cylindracea} could be observed in the taxonomic classification
+of OTUs present at every sampling date. Within the \emph{Bacteroidota}
+different groups within \emph{Saprospiraceae} (e.g.~\emph{Lewinella},
+\emph{Portibacter} and \emph{Rubidimonas}) were identified to be
+persistent. It has been suggested that members of this family are
+important in the hydrolysis and utilization of complex organic sources
+(McIlroy and Nielsen, 2014). Surface attached life style would be
+beneficial to these microbes as they could thrive on products of host
+cellular breakdown or by-products of host metabolism, so it not
+surprising that they are often found associated with macrophyte surfaces
+(Burke et al., 2011b; McIlroy and Nielsen, 2014; Crump et al., 2018).
+\emph{Rhodobacteraceae} are often detected on macrophyte surfaces and
+usually are one of the most abundant groups (Burke et al., 2011b;
+Michelou et al., 2013; Mancuso et al., 2016). The functional association
+between macrophytes and members of this groups is difficult to assess
+based on 16S rRNA analysis as this family is phenotypically,
+metabolically, and ecologically very diverse (Pujalte et al., 2014).
+However, some interesting metabolic capacities linked to this group were
+described. Genomic analysis of \emph{Rhodobacteraceae} strains and
+metatranscriptomic sequencing of seagrass microbiomes revealed the
+potential for biosynthesis of indole-3-acetic acid (IAA), a plant
+hormone (Simon et al., 2017), indicating a possible intake by
+seagrasses. However, another study found no effect of IAA on \emph{C.
+nodosa} growth showing the complexity of macrophyte--microbes
+interactions (Muñoz, 1995). Another persistent alphaproteobacterial
+family was the \emph{Hyphomonadaceae}, a group that contain species with
+stalks used to attach cells to different surfaces (Abraham and Rohde,
+2014). This group has been previously associated with seagrass surfaces
+(Weidner et al., 2000) and it is believed that possessing stalks could
+be an advantage to keep the cells in the proximity of exudate excreted
+by the host (Weidner et al., 2000; Abraham and Rohde, 2014).
+
+Within the \emph{Gammaproteobacteria}, sequences without known
+representatives were the most pronounced group present throughout the
+year. \emph{Gammaproteobacteria} are often a major constituent of
 macrophyte epiphytic communities (Burke et al., 2011b; Michelou et al.,
-2013; Crump et al., 2018). Beside these two groups, other less abundant,
-taxa showed seasonal and host-specific patterns. For example, \emph{C.
-cylindracea} harbored \emph{Thioploca}, a known sulfur sediment bacteria
-and \emph{Cellvibrionaceae}, a family with cultured members known as
-polysaccharide degraders (Jørgensen and Gallardo, 1999; Xie et al.,
-2017). \emph{Desulfobacterota} were found only associated with \emph{C.
-cylindracea} and no group within this phylum showed seasonal patterns
-(\autoref{desulfo}). The presence of this phylum only on \emph{C.
-cylindracea} is to be expected as part of the epiphytic community is in
-direct contact with the sediment. The \emph{Desulfobacterota} community
-was dominated by \emph{Desulfatitalea} and no relative
-\emph{Desulfocapsaceae}, known sulphate sediment groups (Kuever, 2014;
-Higashioka et al., 2015).
-
-In temperate zones, marine macrophytes are exhibiting growth cycles, so
-it is not surprising that the associated epiphytic microbial community
-is undergoing partial seasonal changes. In the present study, we could
-identify in every analysed high rank taxa phylogenetic groups present
+2013; Crump et al., 2018). A study has attributed the expression of
+enzymes for the degradation of galactose-based algal polymers to this
+class indicating their possible involvement into epibiotic algal biofilm
+control (Crump et al., 2018). In addition, \emph{Granulosicoccus} was
+also found in almost all samples. A species of this genus has been
+isolated from the leaf surface of the seagrass \emph{Zostera marina}
+(Kurilenko et al., 2010), while sequences related to this genus have
+been found on the surfaces of macroalgae (Lachnit et al., 2011;
+Bengtsson et al., 2012), including \emph{C. cylindracea} (Rizzo et al.,
+2016a), indicating this group preference for macrophyte surfaces. It is
+possible that bacteria of this genus can thrive on exudates of different
+macrophytes as it is known from cultivated members that they can utilise
+various sugars and amino acids (Ivanova and Webb, 2014). The presence of
+\emph{Desulfobacterota} only on \emph{C. cylindracea} is to be expected
+as part of the epiphytic community is in direct contact with the
+sediment. The \emph{Desulfobacterota} community was comprised of known
+sulphate sediment groups such as the \emph{Desulfatitalea} and no
+relative \emph{Desulfocapsaceae} (Kuever, 2014; Higashioka et al.,
+2015). Sequences related to sulphur cycling bacterial groups have been
+found in \emph{Caulerpa} endophytic and epiphytic communities (Aires et
+al., 2013). It is possible that these groups are involved into enhanced
+sulphate reduction rates observed in sediments underlying
+\emph{Caulerpa} settlements causing unsuitable conditions to
+sulphide-sensitive seagrasses (Holmer et al., 2009).
+
+The only high rank taxonomic group showing strong seasonal fluctuations
+was \emph{Cyanobacteria}. Cyanobacterial sequences were more pronounced
+in November and December than in spring and summer. In the months of
+high cyanobacterial sequence abundances the majority of sequences from
+this group were classified as \emph{Pleurocapsa}, a group known to
+colonize different living and non-living surfaces (Burns et al., 2004;
+Longford et al., 2007; Mobberley et al., 2012; Reisser et al., 2014;
+Kolda et al., 2020). While we observed a strong temporal pattern for
+this group, a study of surface sediment cyanobacterial communities did
+not find any seasonal dynamics for \emph{Pleurocapsa} (Kolda et al.,
+2020), indicating a possibility that there is a reduced selection of the
+epiphytic community by the seagrass during periods of low photosynthetic
+activity (Zavodnik et al., 1998), causing leaves to become a suitable
+surface for non-specific colonizers. Beside all these groups comprising
+most of the sequences, a set of taxa present in lower proportions and
+showing seasonal patterns was identified. This group was comprised of
+e.g.~\emph{Bacteroidota} sequences without known relatives within
+\emph{Flavobacteriaceae} and \emph{Rhodothermaceae}, the
+alphaproteobacterial \emph{Amylibacter} and the gammaproteobacterial
+\emph{Methylotenera}, \emph{Reinekea} and sequences without known
+relatives within \emph{Cellvibrionaceae} (Figures \ref{bactero} and
+\ref{gamma}).
+
+It is possible that \emph{Flavobacteriaceae} and \emph{Rhodothermaceae}
+are occupying similar niches with \emph{Rhodothermaceae} being more
+adapted to higher temperatures as it is known that culturable members of
+this family exhibit mesophilic and thermophilic characteristics (Park et
+al., 2014). This would explain why we observed a higher presence of
+\emph{Rhodothermaceae} in the warmer period of the year. A strain
+belonging to the \emph{Rhodobacteraceae} genus \emph{Amylibacter} has
+been isolated from the surface of a green macroalga indicating that
+members of this group can exhibit surface attached life style
+(Nedashkovskaya et al., 2016). In addition, since this is a relatively
+novel genus it is possible that novel taxa within
+\emph{Rhodobacteraceae} will be described in the future elucidating the
+taxonomy of the currently high proportion of \emph{Rhodobacteraceae}
+sequences without known relatives. The genus \emph{Methylotenera}
+belongs to the methylotrophic family \emph{Methylophilaceae}, a group
+capable of oxidising non-methane single-carbon compounds such as
+methanol and methylamine (Chistoserdova and Kalyuzhnaya, 2018).
+Interestingly, angiosperms produce methanol during cell-wall synthesis
+(Nemecek-Marshall et al., 1995; Dorokhov et al., 2018), so it is not
+surprising that we found members of this genus only on \emph{C. nodosa}
+and in spring, during a period of maximum seagrass proliferation. Other
+studies have also found \emph{Methylotenera}-specific sequences
+associated with seagrass roots and leaves indicating that this group
+members are important constituents of the seagrass microbiome (Crump et
+al., 2018; Sanders-Smith et al., 2020). Genomic and physiological
+analyses of cultivated \emph{Cellvibrionaceae} and \emph{Reinekea}
+members showed the capabilities to use important algal polysaccharides
+(Avcı et al., 2017; Xie et al., 2017) indicating their possible
+involvement into the degradation of \emph{C. cylindracea}
+polysaccharides and/or the control of its algal epiphytes.
+
+The epiphytic microbial community associated with marine macrophytes is
+undergoing seasonal changes that can be attributed to the fluctuations
+of environmental conditions, the growth cycle of macrophytes inhabiting
+temperate zones or the combined effect of both. In the present study, we
+could identify in analysed high rank taxa phylogenetic groups present
 throughout the year, comprising most of the sequences and a lower
 proportion of taxa showing seasonal patterns connected to the macrophyte
-growth cycle (Figs. \ref{community} and \ref{desulfo}). Studies focusing
-on functional comparisons between communities associated with different
-hosts showed that the majority of functions could be found in every
-community, indicating functional redundancy (Roth-Schulze et al., 2016).
-This difference between phylogenetic varaibility and functional
-stability was explained by the lottery hypothesis assuming an initial
-random colonization step performed by a set of functionally equivalent
-taxonomic groups (Burke et al., 2011a; Roth-Schulze et al., 2016). It is
-possible that functional redundancy is a characteristic of high
-abundance taxa detected to be present throughout the year, while
+growth cycle (Figures \ref{community} and \ref{desulfo}). Studies
+focusing on functional comparisons between communities associated with
+different hosts showed that the majority of functions could be found in
+every community, indicating functional redundancy (Roth-Schulze et al.,
+2016). This difference between phylogenetic variability and functional
+stability has been explained by the lottery hypothesis assuming an
+initial random colonization step performed by a set of functionally
+equivalent taxonomic groups (Burke et al., 2011a; Roth-Schulze et al.,
+2016). It is possible that functional redundancy is a characteristic of
+high abundance taxa detected to be present throughout the year, while
 seasonal and/or host-specific functions are an attribute of taxa
@@
 
-\interlinepenalty=10000 \setlength{\emergencystretch}{3.5em}
+\interlinepenalty=10000
+\setlength{\emergencystretch}{3.7em}
 
 \hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
 \leavevmode\hypertarget{ref-Abraham2014}{}%
-Abraham, W. R., and Rohde, M. (2014). ``The family
-\emph{Hyphomonadaceae},'' in \emph{The Prokaryotes: Alphaproteobacteria
-and Betaproteobacteria}, eds. E. Rosenberg, E. F. DeLong, S. Lory, E.
-Stackebrandt, and F. Thompson (Berlin, Heidelberg: Springer-Verlag),
-283--299.
+Abraham, W. R., and Rohde, M. (2014). {``The family
+{\emph{Hyphomonadaceae}},''} in \emph{The {Prokaryotes}:
+{Alphaproteobacteria} and {Betaproteobacteria}}, eds. E. Rosenberg, E.
+F. DeLong, S. Lory, E. Stackebrandt, and F. Thompson ({Berlin,
+Heidelberg}: {Springer-Verlag}), 283--299.
 doi:\href{https://doi.org/10.1007/978-3-642-30197-1_260}{10.1007/978-3-642-30197-1\_260}.
+
+\leavevmode\hypertarget{ref-Agostini2003}{}%
+Agostini, S., Pergent, G., and Marchand, B. (2003). Growth and primary
+production of {\emph{Cymodocea nodosa}} in a coastal lagoon.
+\emph{Aquat. Bot.} 76, 185--193.
+doi:\href{https://doi.org/10.1016/S0304-3770(03)00049-4}{10.1016/S0304-3770(03)00049-4}.
+
+\leavevmode\hypertarget{ref-Aires2015}{}%
+Aires, T., Moalic, Y., Serrao, E. A., and Arnaud-Haond, S. (2015).
+Hologenome theory supported by cooccurrence networks of species-specific
+bacterial communities in siphonous algae ({\emph{Caulerpa}}). \emph{FEMS
+Microbiol. Ecol.} 91, fiv067.
+doi:\href{https://doi.org/10.1093/femsec/fiv067}{10.1093/femsec/fiv067}.
+
+\leavevmode\hypertarget{ref-Aires2013}{}%
+Aires, T., Serrão, E. A., Kendrick, G., Duarte, C. M., and Arnaud-Haond,
+S. (2013). Invasion is a community affair: {Clandestine} followers in
+the bacterial community associated to green algae, {\emph{Caulerpa
+racemosa}}, track the invasion source. \emph{PLoS One} 8, e68429.
+doi:\href{https://doi.org/10.1371/journal.pone.0068429}{10.1371/journal.pone.0068429}.
 
@@
 Allaire, J. J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
-A., et al. (2019). \emph{rmarkdown: Dynamic documents for R}.
+A., et al. (2019). \emph{{rmarkdown}: Dynamic documents for {R}}.
 
@@
 Apprill, A., McNally, S., Parsons, R., and Weber, L. (2015). Minor
-revision to V4 region SSU rRNA 806R gene primer greatly increases
-detection of SAR11 bacterioplankton. \emph{Aquat. Microb. Ecol.} 75,
+revision to {V4} region {SSU rRNA 806R} gene primer greatly increases
+detection of {SAR11} bacterioplankton. \emph{Aquat. Microb. Ecol.} 75,
 129--137. doi:\href{https://doi.org/10.3354/ame01753}{10.3354/ame01753}.
@@
 
-\leavevmode\hypertarget{ref-Bengtsson2017}{}%
-Bengtsson, M. M., Bühler, A., Brauer, A., Dahlke, S., Schubert, H., and
-Blindow, I. (2017). Eelgrass leaf surface microbiomes are locally
-variable and highly correlated with epibiotic eukaryotes. \emph{Front.
-Microbiol.} 8, 1312.
-doi:\href{https://doi.org/10.3389/fmicb.2017.01312}{10.3389/fmicb.2017.01312}.
+\leavevmode\hypertarget{ref-Avci2017}{}%
+Avcı, B., Hahnke, R. L., Chafee, M., Fischer, T., Gruber-Vodicka, H.,
+Tegetmeyer, H. E., et al. (2017). Genomic and physiological analyses of
+`{{\emph{Reinekea forsetii}}' reveal a versatile opportunistic lifestyle
+during spring algae blooms}. \emph{Environ. Microbiol.} 19, 1209--1221.
+doi:\href{https://doi.org/10.1111/1462-2920.13646}{10.1111/1462-2920.13646}.
+
+\leavevmode\hypertarget{ref-Bengtsson2012}{}%
+Bengtsson, M. M., Sjøtun, K., Lanzén, A., and Øvreås, L. (2012).
+Bacterial diversity in relation to secondary production and succession
+on surfaces of the kelp {\emph{Laminaria hyperborea}}. \emph{ISME J.} 6,
+2188--2198.
+doi:\href{https://doi.org/10.1038/ismej.2012.67}{10.1038/ismej.2012.67}.
 
@@
 Bengtsson, M., Sjøtun, K., and Øvreås, L. (2010). Seasonal dynamics of
-bacterial biofilms on the kelp \emph{Laminaria hyperborea}. \emph{Aquat.
-Microb. Ecol.} 60, 71--83.
+bacterial biofilms on the kelp {\emph{Laminaria hyperborea}}.
+\emph{Aquat. Microb. Ecol.} 60, 71--83.
 doi:\href{https://doi.org/10.3354/ame01409}{10.3354/ame01409}.
+
+\leavevmode\hypertarget{ref-Borcard2011}{}%
+Borcard, D., Gillet, F., and Legendre, P. (2011). \emph{Numerical
+ecology with {R}}. 1st ed. {New York}: {Springer-Verlag}
+doi:\href{https://doi.org/10.1007/978-1-4419-7976-6}{10.1007/978-1-4419-7976-6}.
 
@@
 Borges, A. V., and Champenois, W. (2015). Seasonal and spatial
-variability of dimethylsulfoniopropionate (DMSP) in the Mediterranean
-seagrass \emph{Posidonia oceanica}. \emph{Aquat. Bot.} 125, 72--79.
+variability of dimethylsulfoniopropionate ({DMSP}) in the
+{Mediterranean} seagrass {\emph{Posidonia oceanica}}. \emph{Aquat. Bot.}
+125, 72--79.
 doi:\href{https://doi.org/10.1016/j.aquabot.2015.05.008}{10.1016/j.aquabot.2015.05.008}.
 
-\leavevmode\hypertarget{ref-Burke2009}{}%
-Burke, C., Kjelleberg, S., and Thomas, T. (2009). Selective extraction
-of bacterial DNA from the surfaces of macroalgae. \emph{Appl. Environ.
-Microbiol.} 75, 252--256.
-doi:\href{https://doi.org/10.1128/AEM.01630-08}{10.1128/AEM.01630-08}.
+\leavevmode\hypertarget{ref-Boudouresque2009}{}%
+Boudouresque, C. F., Bernard, G., Pergent, G., Shili, A., and Verlaque,
+M. (2009). Regression of {Mediterranean} seagrasses caused by natural
+processes and anthropogenic disturbances and stress: A critical review.
+\emph{Botanica Marina} 52, 395--418.
+doi:\href{https://doi.org/10.1515/BOT.2009.057}{10.1515/BOT.2009.057}.
 
@@
 (2011b). Composition, uniqueness and variability of the epiphytic
-bacterial community of the green alga \emph{Ulva australis}. \emph{ISME
-J.} 5, 590--600.
+bacterial community of the green alga {\emph{Ulva australis}}.
+\emph{ISME J.} 5, 590--600.
 doi:\href{https://doi.org/10.1038/ismej.2010.164}{10.1038/ismej.2010.164}.
@@
 diversity of extant stromatolites in the hypersaline marine environment
-of Shark Bay, Australia. \emph{Environ. Microbiol.} 6, 1096--1101.
+of {Shark Bay}, {Australia}. \emph{Environ. Microbiol.} 6, 1096--1101.
 doi:\href{https://doi.org/10.1111/j.1462-2920.2004.00651.x}{10.1111/j.1462-2920.2004.00651.x}.
-
-\leavevmode\hypertarget{ref-Cai2014}{}%
-Cai, X., Gao, G., Yang, J., Tang, X., Dai, J., Chen, D., et al. (2014).
-An ultrasonic method for separation of epiphytic microbes from
-freshwater submerged macrophytes. \emph{J. Basic Microbiol.} 54,
-758--761.
-doi:\href{https://doi.org/10.1002/jobm.201300041}{10.1002/jobm.201300041}.
 
@@
 J., Fierer, N., et al. (2012). Ultra-high-throughput microbial community
-analysis on the Illumina HiSeq and MiSeq platforms. \emph{ISME J.} 6,
-1621--1624.
+analysis on the {Illumina HiSeq} and {MiSeq} platforms. \emph{ISME J.}
+6, 1621--1624.
 doi:\href{https://doi.org/10.1038/ismej.2012.8}{10.1038/ismej.2012.8}.
 
-\leavevmode\hypertarget{ref-Cardoso2019}{}%
-Cardoso, D. C., Cretoiu, M. S., Stal, L. J., and Bolhuis, H. (2019).
-Seasonal development of a coastal microbial mat. \emph{Sci. Rep.} 9,
-9035.
-doi:\href{https://doi.org/10.1038/s41598-019-45490-8}{10.1038/s41598-019-45490-8}.
+\leavevmode\hypertarget{ref-Celdran2012}{}%
+Celdrán, D., Espinosa, E., Sánchez-Amat, A., and Marín, A. (2012).
+Effects of epibiotic bacteria on leaf growth and epiphytes of the
+seagrass {\emph{Posidonia oceanica}}. \emph{Mar. Ecol. Prog. Ser.} 456,
+21--27. doi:\href{https://doi.org/10.3354/meps09672}{10.3354/meps09672}.
+
+\leavevmode\hypertarget{ref-Chistoserdova2018}{}%
+Chistoserdova, L., and Kalyuzhnaya, M. G. (2018). Current trends in
+methylotrophy. \emph{Trends Microbiol.} 26, 703--714.
+doi:\href{https://doi.org/10.1016/j.tim.2018.01.011}{10.1016/j.tim.2018.01.011}.
 
@@
 
+\leavevmode\hypertarget{ref-Cucio2016}{}%
+Cúcio, C., Engelen, A. H., Costa, R., and Muyzer, G. (2016). Rhizosphere
+microbiomes of european seagrasses are selected by the plant, but are
+not species specific. \emph{Front. Microbiol.} 7, 440.
+doi:\href{https://doi.org/10.3389/fmicb.2016.00440}{10.3389/fmicb.2016.00440}.
+
+\leavevmode\hypertarget{ref-Cucio2018}{}%
+Cúcio, C., Overmars, L., Engelen, A. H., and Muyzer, G. (2018).
+Metagenomic analysis shows the presence of bacteria related to
+free-living forms of sulfur-oxidizing chemolithoautotrophic symbionts in
+the rhizosphere of the seagrass {\emph{Zostera marina}}. \emph{Front.
+Mar. Sci.} 5, 171.
+doi:\href{https://doi.org/10.3389/fmars.2018.00171}{10.3389/fmars.2018.00171}.
+
+\leavevmode\hypertarget{ref-deOliveira2012}{}%
+de Oliveira, L. S., Gregoracci, G. B., Silva, G. G. Z., Salgado, L. T.,
+Filho, G. A., Alves-Ferreira, M., et al. (2012). Transcriptomic analysis
+of the red seaweed {\emph{Laurencia dendroidea}} ({Florideophyceae},
+{Rhodophyta}) and its microbiome. \emph{BMC Genomics} 13, 487.
+doi:\href{https://doi.org/10.1186/1471-2164-13-487}{10.1186/1471-2164-13-487}.
+
+\leavevmode\hypertarget{ref-Dobretsov2002}{}%
+Dobretsov, S. V., and Qian, P.-Y. (2002). Effect of bacteria associated
+with the green alga {\emph{Ulva reticulata}} on marine micro- and
+macrofouling. \emph{Biofouling} 18, 217--228.
+doi:\href{https://doi.org/10.1080/08927010290013026}{10.1080/08927010290013026}.
+
+\leavevmode\hypertarget{ref-Dorokhov2018}{}%
+Dorokhov, Y. L., Sheshukova, E. V., and Komarova, T. V. (2018). Methanol
+in plant life. \emph{Front. Plant Sci.} 9, 1623.
+doi:\href{https://doi.org/10.3389/fpls.2018.01623}{10.3389/fpls.2018.01623}.
+
 \leavevmode\hypertarget{ref-Egan2013}{}%
@@
 
-\leavevmode\hypertarget{ref-Fahimipour2017}{}%
-Fahimipour, A. K., Kardish, M. R., Lang, J. M., Green, J. L., Eisen, J.
-A., and Stachowicz, J. J. (2017). Global-scale structure of the eelgrass
-microbiome. \emph{Appl. Environ. Microbiol.} 83, e03391--16.
-doi:\href{https://doi.org/10.1128/AEM.03391-16}{10.1128/AEM.03391-16}.
+\leavevmode\hypertarget{ref-Ettinger2020}{}%
+Ettinger, C. L., and Eisen, J. A. (2020). Fungi, bacteria and oomycota
+opportunistically isolated from the seagrass, {\emph{Zostera marina}}.
+\emph{PLoS One} 15, e0236135.
+doi:\href{https://doi.org/10.1371/journal.pone.0236135}{10.1371/journal.pone.0236135}.
 
@@
 T., et al. (2009). The seasonal structure of microbial communities in
-the Western English Channel. \emph{Environ. Microbiol.} 11, 3132--3139.
+the {Western English Channel}. \emph{Environ. Microbiol.} 11,
+3132--3139.
 doi:\href{https://doi.org/10.1111/j.1462-2920.2009.02017.x}{10.1111/j.1462-2920.2009.02017.x}.
@@
 Higashioka, Y., Kojima, H., Watanabe, T., and Fukui, M. (2015). Draft
-genome sequence of \emph{Desulfatitalea tepidiphila}
-S28bF\textsuperscript{T}. \emph{Genome Announc.} 3, e01326--15.
+genome sequence of {\emph{Desulfatitalea tepidiphila}}
+{S28bF}{\textsuperscript{T}}. \emph{Genome Announc.} 3, e01326--15.
 doi:\href{https://doi.org/10.1128/genomeA.01326-15}{10.1128/genomeA.01326-15}.
@@
 we can learn from sushi: A review on seaweed-bacterial associations.
-\emph{FEMS Microbiol. Ecol.} 83.
+\emph{FEMS Microbiol. Ecol.} 83, 1--16.
 doi:\href{https://doi.org/10.1111/j.1574-6941.2012.01446.x}{10.1111/j.1574-6941.2012.01446.x}.
+
+\leavevmode\hypertarget{ref-Holmer2009}{}%
+Holmer, M., Marbà, N., Lamote, M., and Duarte, C. M. (2009).
+Deterioration of sediment quality in seagrass meadows ({\emph{Posidonia
+oceanica}}) invaded by macroalgae ({\emph{Caulerpa}} sp.).
+\emph{Estuaries Coast} 32, 456--466.
+doi:\href{https://doi.org/10.1007/s12237-009-9133-4}{10.1007/s12237-009-9133-4}.
+
+\leavevmode\hypertarget{ref-Ivanova2014}{}%
+Ivanova, E. P., and Webb, H. K. (2014). {``The family
+{\emph{Granulosicoccaceae}},''} in \emph{The {Prokaryotes}:
+{Gammaproteobacteria}}, eds. E. Rosenberg, E. F. DeLong, S. Lory, E.
+Stackebrandt, and F. Thompson ({Berlin, Heidelberg}: {Springer-Verlag}),
+315--317.
+doi:\href{https://doi.org/10.1007/978-3-642-38922-1_247}{10.1007/978-3-642-38922-1\_247}.
 
@@
 
-\leavevmode\hypertarget{ref-Jorgensen1999}{}%
-Jørgensen, B. B., and Gallardo, V. A. (1999). \emph{Thioploca} spp.:
-filamentous sulfur bacteria with nitrate vacuoles. \emph{FEMS Microbiol.
-Ecol.} 28, 301--313.
-doi:\href{https://doi.org/10.1111/j.1574-6941.1999.tb00585.x}{10.1111/j.1574-6941.1999.tb00585.x}.
-
-\leavevmode\hypertarget{ref-Korlevic}{}%
+\leavevmode\hypertarget{ref-Klein2008}{}%
+Klein, J., and Verlaque, M. (2008). The {\emph{Caulerpa racemosa}}
+invasion: A critical review. \emph{Mar. Pollut. Bull.} 56, 205--225.
+doi:\href{https://doi.org/10.1016/j.marpolbul.2007.09.043}{10.1016/j.marpolbul.2007.09.043}.
+
+\leavevmode\hypertarget{ref-Kolda2020}{}%
+Kolda, A., Ljubešić, Z., Gavrilović, A., Jug-Dujaković, J., Pikelj, K.,
+and Kapetanović, D. (2020). Metabarcoding {\emph{Cyanobacteria}} in
+coastal waters and sediment in central and southern {Adriatic Sea}.
+\emph{Acta Bot. Croat.} 79, 157--169.
+doi:\href{https://doi.org/10.37427/botcro-2020-021}{10.37427/botcro-2020-021}.
+
+\leavevmode\hypertarget{ref-Korlevic2021}{}%
 Korlević, M., Markovski, M., Zhao, Z., Herndl, G. J., and Najdek, M.
-Selective DNA and protein isolation from marine macrophyte surfaces.
+(2021). Selective {DNA} and protein isolation from marine macrophyte
+surfaces. \emph{Front. Microbiol.} 12, 665999.
+doi:\href{https://doi.org/10.3389/fmicb.2021.665999}{10.3389/fmicb.2021.665999}.
 
@@
 Korlević, M., Pop Ristova, P., Garić, R., Amann, R., and Orlić, S.
-(2015). Bacterial diversity in the South Adriatic Sea during a strong,
+(2015). Bacterial diversity in the {South Adriatic Sea} during a strong,
 deep winter convection year. \emph{Appl. Environ. Microbiol.} 81,
@@
 Schloss, P. D. (2013). Development of a dual-index sequencing strategy
-and curation pipeline for analyzing amplicon sequence data on the MiSeq
-Illumina sequencing platform. \emph{Appl. Environ. Microbiol.} 79,
+and curation pipeline for analyzing amplicon sequence data on the {MiSeq
+Illumina} sequencing platform. \emph{Appl. Environ. Microbiol.} 79,
 5112--5120.
@@
 \leavevmode\hypertarget{ref-Kuever2014}{}%
-Kuever, J. (2014). ``The family \emph{Desulfobulbaceae},'' in \emph{The
-Prokaryotes: Deltaproteobacteria and Epsilonproteobacteria}, eds. E.
-Rosenberg, E. F. DeLong, S. Lory, E. Stackebrandt, and F. Thompson
-(Berlin, Heidelberg: Springer-Verlag), 75--86.
+Kuever, J. (2014). {``The family {\emph{Desulfobulbaceae}},''} in
+\emph{The {Prokaryotes}: {Deltaproteobacteria} and
+{Epsilonproteobacteria}}, eds. E. Rosenberg, E. F. DeLong, S. Lory, E.
+Stackebrandt, and F. Thompson ({Berlin, Heidelberg}: {Springer-Verlag}),
+75--86.
 doi:\href{https://doi.org/10.1007/978-3-642-39044-9_267}{10.1007/978-3-642-39044-9\_267}.
+
+\leavevmode\hypertarget{ref-Kurilenko2010}{}%
+Kurilenko, V. V., Christen, R., Zhukova, N. V., Kalinovskaya, N. I.,
+Mikhailov, V. V., Crawford, R. J., et al. (2010).
+{{\emph{Granulosicoccus coccoides}} sp. nov., isolated from leaves of
+seagrass ({\emph{Zostera marina}})}. \emph{Int. J. Syst. Evol.
+Microbiol.} 60, 972--976.
+doi:\href{https://doi.org/10.1099/ijs.0.013516-0}{10.1099/ijs.0.013516-0}.
+
+\leavevmode\hypertarget{ref-Kusel2006}{}%
+Küsel, K., Trinkwalter, T., Drake, H. L., and Devereux, R. (2006).
+Comparative evaluation of anaerobic bacterial communities associated
+with roots of submerged macrophytes growing in marine or brackish water
+sediments. \emph{J. Exp. Mar. Biol. Ecol.} 337, 49--58.
+doi:\href{https://doi.org/10.1016/j.jembe.2006.06.004}{10.1016/j.jembe.2006.06.004}.
 
@@
 doi:\href{https://doi.org/10.1111/j.1462-2920.2010.02371.x}{10.1111/j.1462-2920.2010.02371.x}.
+
+\leavevmode\hypertarget{ref-Legendre2012}{}%
+Legendre, P., and Legendre, L. (2012). \emph{Numerical ecology}. 3rd ed.
+{Amsterdam}: {Elsevier}.
 
@@
 (2016). Diversity and temporal dynamics of the epiphytic bacterial
-communities associated with the canopy-forming seaweed \emph{Cystoseira
-compressa} (Esper) Gerloff and Nizamuddin. \emph{Front Microbiol} 7,
-476.
+communities associated with the canopy-forming seaweed {\emph{Cystoseira
+compressa}} ({Esper}) {Gerloff} and {Nizamuddin}. \emph{Front.
+Microbiol.} 7, 476.
 doi:\href{https://doi.org/10.3389/fmicb.2016.00476}{10.3389/fmicb.2016.00476}.
@@
 \leavevmode\hypertarget{ref-Margulis1991}{}%
-Margulis, L. (1991). ``Symbiogenesis and symbionticism,'' in
-\emph{Symbiosis as a Source of Evolutionary Innovation: Speciation and
-Morphogenesis}, eds. L. Margulis and R. Fester (Cambridge,
-Massachusetts: The MIT Press), 1--14.
+Margulis, L. (1991). {``Symbiogenesis and symbionticism,''} in
+\emph{Symbiosis as a {Source} of {Evolutionary Innovation}: {Speciation}
+and {Morphogenesis}}, eds. L. Margulis and R. Fester ({Cambridge,
+Massachusetts}: {The MIT Press}), 1--14.
 
@@
 Vertical distribution and phylogenetic characterization of marine
-planktonic \emph{Archaea} in the Santa Barbara Channel. \emph{Appl.
+planktonic {\emph{Archaea}} in the {Santa Barbara Channel}. \emph{Appl.
 Environ. Microbiol.} 63, 50--56.
 
+\leavevmode\hypertarget{ref-Matsuo2003}{}%
+Matsuo, Y., Suzuki, M., Kasai, H., Shizuri, Y., and Harayama, S. (2003).
+Isolation and phylogenetic characterization of bacteria capable of
+inducing differentiation in the green alga {\emph{Monostroma
+oxyspermum}}. \emph{Environ. Microbiol.} 5, 25--35.
+doi:\href{https://doi.org/10.1046/j.1462-2920.2003.00382.x}{10.1046/j.1462-2920.2003.00382.x}.
+
 \leavevmode\hypertarget{ref-McIlroy2014}{}%
-McIlroy, S. J., and Nielsen, P. H. (2014). ``The family
-\emph{Saprospiraceae},'' in \emph{The Prokaryotes: Other Major Lineages
-of Bacteria and the Archaea}, eds. E. Rosenberg, E. F. DeLong, S. Lory,
-E. Stackebrandt, and F. Thompson (Berlin, Heidelberg: Springer-Verlag),
-863--889.
+McIlroy, S. J., and Nielsen, P. H. (2014). {``The family
+{\emph{Saprospiraceae}},''} in \emph{The {Prokaryotes}: {Other Major
+Lineages} of {Bacteria} and the {Archaea}}, eds. E. Rosenberg, E. F.
+DeLong, S. Lory, E. Stackebrandt, and F. Thompson ({Berlin, Heidelberg}:
+{Springer-Verlag}), 863--889.
 doi:\href{https://doi.org/10.1007/978-3-642-38954-2_138}{10.1007/978-3-642-38954-2\_138}.
@@
 Michelou, V. K., Caporaso, J. G., Knight, R., and Palumbi, S. R. (2013).
-The ecology of microbial communities associated with \emph{Macrocystis
-pyrifera}. \emph{PloS One} 8, e67480.
+The ecology of microbial communities associated with {\emph{Macrocystis
+pyrifera}}. \emph{PLoS One} 8, e67480.
 doi:\href{https://doi.org/10.1371/journal.pone.0067480}{10.1371/journal.pone.0067480}.
@@
 (2013). Diversity and abundance of the bacterial community of the red
-macroalga \emph{Porphyra umbilicalis}: Did bacterial farmers produce
-macroalgae? \emph{PloS One} 8, e58269.
+macroalga {\emph{Porphyra umbilicalis}}: Did bacterial farmers produce
+macroalgae? \emph{PLoS One} 8, e58269.
 doi:\href{https://doi.org/10.1371/journal.pone.0058269}{10.1371/journal.pone.0058269}.
@@
 
+\leavevmode\hypertarget{ref-Munoz1995}{}%
+Muñoz, J. T. (1995). Effects of some plant growth regulators on the
+growth of the seagrass {\emph{Cymodocea nodosa}} ({Ucria}) {Ascherson}.
+\emph{Aquat. Bot.} 51, 311--318.
+doi:\href{https://doi.org/10.1016/0304-3770(95)00481-E}{10.1016/0304-3770(95)00481-E}.
+
+\leavevmode\hypertarget{ref-Najdek2020a}{}%
+Najdek, M., Korlević, M., Paliaga, P., Markovski, M., Ivančić, I.,
+Iveša, L., et al. (2020). Effects of the invasion of {\emph{Caulerpa
+cylindracea}} in a {\emph{Cymodocea nodosa}} meadow in the {Northern
+Adriatic Sea}. \emph{Front. Mar. Sci.} 7, 602055.
+doi:\href{https://doi.org/10.3389/fmars.2020.602055}{10.3389/fmars.2020.602055}.
+
+\leavevmode\hypertarget{ref-Nedashkovskaya2016}{}%
+Nedashkovskaya, O. I., Kukhlevskiy, A. D., Zhukova, N. V., and Kim, S.
+B. (2016). {{\emph{Amylibacter ulvae}} sp. nov., a new
+alphaproteobacterium isolated from the Pacific green alga {\emph{Ulva
+fenestrata}}}. \emph{Arch. Microbiol.} 198, 251--256.
+doi:\href{https://doi.org/10.1007/s00203-015-1185-1}{10.1007/s00203-015-1185-1}.
+
+\leavevmode\hypertarget{ref-Nemecek-Marshall1995}{}%
+Nemecek-Marshall, M., MacDonald, R. C., Franzen, J. J., Wojciechowski,
+C. L., and Fall, R. (1995). Methanol emission from leaves (enzymatic
+detection of gas-phase methanol and relation of methanol fluxes to
+stomatal conductance and leaf development). \emph{Plant Physiol.} 108,
+1359--1368.
+doi:\href{https://doi.org/10.1104/pp.108.4.1359}{10.1104/pp.108.4.1359}.
+
 \leavevmode\hypertarget{ref-Neuwirth2014}{}%
-Neuwirth, E. (2014). \emph{RColorBrewer: ColorBrewer palettes}.
-
-\leavevmode\hypertarget{ref-Noges2010}{}%
-Nõges, T., Luup, H., and Feldmann, T. (2010). Primary production of
-aquatic macrophytes and their epiphytes in two shallow lakes (Peipsi and
-Võrtsjärv) in Estonia. \emph{Aquat. Ecol.} 44, 83--92.
-doi:\href{https://doi.org/10.1007/s10452-009-9249-4}{10.1007/s10452-009-9249-4}.
+Neuwirth, E. (2014). \emph{{RColorBrewer}: {ColorBrewer} palettes}.
+
+\leavevmode\hypertarget{ref-Nielsen2001}{}%
+Nielsen, L. B., Finster, K., Welsh, D. T., Donelly, A., Herbert, R. A.,
+de Wit, R., et al. (2001). Sulphate reduction and nitrogen fixation
+rates associated with roots, rhizomes and sediments from {\emph{Zostera
+noltii}} and {\emph{Spartina maritima}} meadows. \emph{Environ.
+Microbiol.} 3, 63--71.
+doi:\href{https://doi.org/doi.org/10.1046/j.1462-2920.2001.00160.x}{doi.org/10.1046/j.1462-2920.2001.00160.x}.
 
@@
 Oksanen, J., Blanchet, F. G., Friendly, M., Kindt, R., Legendre, P.,
-McGlinn, D., et al. (2019). \emph{vegan: Community ecology package}.
+McGlinn, D., et al. (2019). \emph{{vegan}: Community ecology package}.
 
@@
 Parada, A. E., Needham, D. M., and Fuhrman, J. A. (2016). Every base
-matters: Assessing small subunit rRNA primers for marine microbiomes
+matters: Assessing small subunit {rRNA} primers for marine microbiomes
 with mock communities, time series and global field samples.
@@
 
+\leavevmode\hypertarget{ref-Park2014}{}%
+Park, S., Akira, Y., and Kogure, K. (2014). {``The family
+{\emph{Rhodothermaceae}},''} in \emph{The {Prokaryotes}: {Other Major
+Lineages} of {Bacteria} and the {Archaea}}, eds. E. Rosenberg, E. F.
+DeLong, S. Lory, E. Stackebrandt, and F. Thompson ({Berlin, Heidelberg}:
+{Springer-Verlag}), 849--856.
+doi:\href{https://doi.org/10.1007/978-3-642-38954-2_141}{10.1007/978-3-642-38954-2\_141}.
+
+\leavevmode\hypertarget{ref-Penesyan2009}{}%
+Penesyan, A., Marshall-Jones, Z., Holmstrom, C., Kjelleberg, S., and
+Egan, S. (2009). Antimicrobial activity observed among cultured marine
+epiphytic bacteria reflects their potential as a source of new drugs.
+\emph{FEMS Microbiol. Ecol.} 69, 113--124.
+doi:\href{https://doi.org/10.1111/j.1574-6941.2009.00688.x}{10.1111/j.1574-6941.2009.00688.x}.
+
 \leavevmode\hypertarget{ref-Pujalte2014}{}%
 Pujalte, M. J., Lucena, T., Ruvira, M. A., Arahal, D. R., and Macián, M.
-C. (2014). ``The family \emph{Rhodobacteraceae},'' in \emph{The
-Prokaryotes: Alphaproteobacteria and Betaproteobacteria}, eds. E.
+C. (2014). {``The family {\emph{Rhodobacteraceae}},''} in \emph{The
+{Prokaryotes}: {Alphaproteobacteria} and {Betaproteobacteria}}, eds. E.
 Rosenberg, E. F. DeLong, S. Lory, E. Stackebrandt, and F. Thompson
-(Berlin, Heidelberg: Springer-Verlag), 439--512.
+({Berlin, Heidelberg}: {Springer-Verlag}), 439--512.
 doi:\href{https://doi.org/10.1007/978-3-642-30197-1_377}{10.1007/978-3-642-30197-1\_377}.
+
+\leavevmode\hypertarget{ref-Pusceddu2016}{}%
+Pusceddu, A., Fraschetti, S., Scopa, M., Rizzo, L., and Danovaro, R.
+(2016). Meiofauna communities, nematode diversity and {C} degradation
+rates in seagrass ({{\emph{Posidonia oceanica}} L.) and unvegetated
+sediments invaded by the algae {\emph{Caulerpa cylindracea}} (Sonder)}.
+\emph{Mar. Environ. Res.} 119, 88--99.
+doi:\href{https://doi.org/10.1016/j.marenvres.2016.05.015}{10.1016/j.marenvres.2016.05.015}.
 
@@
 Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
-et al. (2013). The SILVA ribosomal RNA gene database project: Improved
-data processing and web-based tools. \emph{Nucleic Acids Res.} 41,
-D590--D596.
+et al. (2013). The {SILVA} ribosomal {RNA} gene database project:
+Improved data processing and web-based tools. \emph{Nucleic Acids Res.}
+41, D590--D596.
 doi:\href{https://doi.org/10.1093/nar/gks1219}{10.1093/nar/gks1219}.
@@
 \leavevmode\hypertarget{ref-RCoreTeam2019}{}%
-R Core Team (2019). \emph{R: A language and environment for statistical
-computing}. Vienna, Austria: R Foundation for Statistical Computing.
+R Core Team (2019). \emph{R: {A} language and environment for
+statistical computing}. {Vienna, Austria}: {R Foundation for Statistical
+Computing}.
 
@@
 Thums, M., et al. (2014). Millimeter-sized marine plastics: A new
-pelagic habitat for microorganisms and invertebrates. \emph{PloS One} 9,
+pelagic habitat for microorganisms and invertebrates. \emph{PLoS One} 9,
 e100289.
@@
 Reyes, J., and Sansón, M. (2001). Biomass and production of the
-epiphytes on the leaves of \emph{Cymodocea nodosa} in the Canary
-Islands. \emph{Bot. Mar.} 44, 307--313.
+epiphytes on the leaves of {\emph{Cymodocea nodosa}} in the {Canary
+Islands}. \emph{Botanica Marina} 44, 307--313.
 doi:\href{https://doi.org/10.1515/BOT.2001.039}{10.1515/BOT.2001.039}.
@@
 Rickert, E., Wahl, M., Link, H., Richter, H., and Pohnert, G. (2016).
-Seasonal variations in surface metabolite composition of \emph{Fucus
-vesiculosus} and \emph{Fucus serratus} from the Baltic Sea. \emph{PloS
-One} 11, e0168196.
+Seasonal variations in surface metabolite composition of {\emph{Fucus
+vesiculosus}} and {\emph{Fucus serratus}} from the {Baltic Sea}.
+\emph{PLoS One} 11, e0168196.
 doi:\href{https://doi.org/10.1371/journal.pone.0168196}{10.1371/journal.pone.0168196}.
+
+\leavevmode\hypertarget{ref-Rizzo2016}{}%
+Rizzo, L., Fraschetti, S., Alifano, P., Pizzolante, G., and Stabili, L.
+(2016a). The alien species {\emph{Caulerpa cylindracea}} and its
+associated bacteria in the {Mediterranean Sea}. \emph{Mar. Biol.} 163,
+4.
+doi:\href{https://doi.org/10.1007/s00227-015-2775-9}{10.1007/s00227-015-2775-9}.
+
+\leavevmode\hypertarget{ref-Rizzo2016a}{}%
+Rizzo, L., Fraschetti, S., Alifano, P., Tredici, M. S., and Stabili, L.
+(2016b). Association of {\emph{Vibrio}} community with the {Atlantic
+Mediterranean} invasive alga {\emph{Caulerpa cylindracea{\emph{}}}}.
+\emph{J. Exp. Mar. Biol. Ecol.} 475, 129--136.
+doi:\href{https://doi.org/10.1016/j.jembe.2015.11.013}{10.1016/j.jembe.2015.11.013}.
+
+\leavevmode\hypertarget{ref-Rizzo2020}{}%
+Rizzo, L., Pusceddu, A., Bianchelli, S., and Fraschetti, S. (2020).
+Potentially combined effect of the invasive seaweed {\emph{Caulerpa
+cylindracea}} ({Sonder}) and sediment deposition rates on organic matter
+and meiofaunal assemblages. \emph{Mar. Environ. Res.} 159, 104966.
+doi:\href{https://doi.org/10.1016/j.marenvres.2020.104966}{10.1016/j.marenvres.2020.104966}.
+
+\leavevmode\hypertarget{ref-Rizzo2017}{}%
+Rizzo, L., Pusceddu, A., Stabili, L., Alifano, P., and Fraschetti, S.
+(2017). Potential effects of an invasive seaweed ({\emph{Caulerpa
+cylindracea}}, {Sonder}) on sedimentary organic matter and microbial
+metabolic activities. \emph{Sci. Rep.} 7, 12113.
+doi:\href{https://doi.org/10.1038/s41598-017-12556-4}{10.1038/s41598-017-12556-4}.
 
@@
 Ruitton, S., Verlaque, M., and Boudouresque, C. F. (2005). Seasonal
-changes of the introduced \emph{Caulerpa racemosa} var.
-\emph{cylindracea} (Caulerpales, Chlorophyta) at the northwest limit of
-its Mediterranean range. \emph{Aquat. Bot.} 82, 55--70.
+changes of the introduced {\emph{Caulerpa racemosa}} var.
+{\emph{cylindracea}} ({Caulerpales}, {Chlorophyta}) at the northwest
+limit of its {Mediterranean} range. \emph{Aquat. Bot.} 82, 55--70.
 doi:\href{https://doi.org/10.1016/j.aquabot.2005.02.008}{10.1016/j.aquabot.2005.02.008}.
+
+\leavevmode\hypertarget{ref-Saha2011}{}%
+Saha, M., Rempt, M., Grosser, K., Pohnert, G., and Weinberger, F.
+(2011). Surface-associated fucoxanthin mediates settlement of bacterial
+epiphytes on the rockweed {\emph{Fucus vesiculosus}}. \emph{Biofouling}
+27, 423--433.
+doi:\href{https://doi.org/10.1080/08927014.2011.580841}{10.1080/08927014.2011.580841}.
 
@@
 Salaün, S., La Barre, S., Santos-Goncalvez, M. D., Potin, P., Haras, D.,
-and Bazire, A. (2012). Influence of exudates of the kelp \emph{Laminaria
-digitata} on biofilm formation of associated and exogenous bacterial
-epiphytes. \emph{Microb. Ecol.} 64, 359--369.
+and Bazire, A. (2012). Influence of exudates of the kelp
+{\emph{Laminaria digitata}} on biofilm formation of associated and
+exogenous bacterial epiphytes. \emph{Microb. Ecol.} 64, 359--369.
 doi:\href{https://doi.org/10.1007/s00248-012-0048-4}{10.1007/s00248-012-0048-4}.
+
+\leavevmode\hypertarget{ref-Sale1976}{}%
+Sale, P. F. (1976). Reef fish lottery. \emph{Nat. Hist.} 85, 60--65.
+
+\leavevmode\hypertarget{ref-Sanchez-Amat2010}{}%
+Sanchez-Amat, A., Solano, F., and Lucas-Elío, P. (2010). Finding new
+enzymes from bacterial physiology: A successful approach illustrated by
+the detection of novel oxidases in {\emph{Marinomonas mediterranea}}.
+\emph{Mar. Drugs} 8, 519--541.
+doi:\href{https://doi.org/10.3390/md8030519}{10.3390/md8030519}.
+
+\leavevmode\hypertarget{ref-Sanders-Smith2020}{}%
+Sanders-Smith, R., Segovia, B. T., Forbes, C., Hessing-Lewis, M.,
+Morien, E., Lemay, M. A., et al. (2020). Host-specificity and core taxa
+of seagrass leaf microbiome identified across tissue age and
+geographical regions. \emph{Front. Ecol. Evol.} 8.
+doi:\href{https://doi.org/10.3389/fevo.2020.605304}{10.3389/fevo.2020.605304}.
 
@@
 Schloss, P. D., Jenior, M. L., Koumpouras, C. C., Westcott, S. L., and
-Highlander, S. K. (2016). Sequencing 16S rRNA gene fragments using the
-PacBio SMRT DNA sequencing system. \emph{PeerJ} 4, e1869.
+Highlander, S. K. (2016). Sequencing {16S rRNA} gene fragments using the
+{PacBio SMRT DNA} sequencing system. \emph{PeerJ} 4, e1869.
 doi:\href{https://doi.org/10.7717/peerj.1869}{10.7717/peerj.1869}.
@@
 
+\leavevmode\hypertarget{ref-Schmidt2015a}{}%
+Schmidt, V. T., Smith, K. F., Melvin, D. W., and Amaral-Zettler, L. A.
+(2015). Community assembly of a euryhaline fish microbiome during
+salinity acclimation. \emph{Mol. Ecol.} 24, 2537--2550.
+doi:\href{https://doi.org/10.1111/mec.13177}{10.1111/mec.13177}.
+
+\leavevmode\hypertarget{ref-Short2001}{}%
+Short, F. T., Coles, R. G., and Pergent-Martini, C. (2001). {``Global
+seagrass distribution,''} in \emph{Global {Seagrass Research Methods}},
+eds. F. T. Short and R. G. Coles ({Amsterdam}: {Elsevier Science B.V.}),
+5--30.
+
+\leavevmode\hypertarget{ref-Simon2017}{}%
+Simon, M., Scheuner, C., Meier-Kolthoff, J. P., Brinkhoff, T.,
+Wagner-Döbler, I., Ulbrich, M., et al. (2017). Phylogenomics of
+{\emph{Rhodobacteraceae}} reveals evolutionary adaptation to marine and
+non-marine habitats. \emph{ISME J.} 11, 1483--1499.
+doi:\href{https://doi.org/10.1038/ismej.2016.198}{10.1038/ismej.2016.198}.
+
+\leavevmode\hypertarget{ref-Stabili2017}{}%
+Stabili, L., Rizzo, L., Pizzolante, G., Alifano, P., and Fraschetti, S.
+(2017). Spatial distribution of the culturable bacterial community
+associated with the invasive alga {\emph{Caulerpa cylindracea}} in the
+{Mediterranean Sea}. \emph{Mar. Environ. Res.} 125, 90--98.
+doi:\href{https://doi.org/10.1016/j.marenvres.2017.02.001}{10.1016/j.marenvres.2017.02.001}.
+
+\leavevmode\hypertarget{ref-Stratil2013}{}%
+Stratil, S. B., Neulinger, S. C., Knecht, H., Friedrichs, A. K., and
+Wahl, M. (2013). Temperature-driven shifts in the epibiotic bacterial
+community composition of the brown macroalga {\emph{Fucus vesiculosus}}.
+\emph{Microbiologyopen} 2, 338--349.
+doi:\href{https://doi.org/10.1002/mbo3.79}{10.1002/mbo3.79}.
+
 \leavevmode\hypertarget{ref-Tarquinio2019}{}%
@@
 and Kjelleberg, S. (2010). Variability and abundance of the epiphytic
-bacterial community associated with a green marine \emph{Ulvacean} alga.
-\emph{ISME J.} 4, 301--311.
+bacterial community associated with a green marine {\emph{Ulvacean}}
+alga. \emph{ISME J.} 4, 301--311.
 doi:\href{https://doi.org/10.1038/ismej.2009.107}{10.1038/ismej.2009.107}.
@@
 Ugarelli, K., Laas, P., and Stingl, U. (2019). The microbial communities
-of leaves and roots associated with turtle grass (\emph{Thalassia
-testudinum}) and manatee grass (\emph{Syringodium filliforme}) are
+of leaves and roots associated with turtle grass ({\emph{Thalassia
+testudinum}}) and manatee grass ({\emph{Syringodium filliforme}}) are
 distinct from seawater and sediment communities, but are similar between
@@
 
-\leavevmode\hypertarget{ref-Uku2007}{}%
-Uku, J., Björk, M., Bergman, B., and Díez, B. (2007). Characterization
-and comparison of prokaryotic epiphytes associated with three East
-African seagrasses. \emph{J. Phycol.} 43, 768--779.
-doi:\href{https://doi.org/10.1111/j.1529-8817.2007.00371.x}{10.1111/j.1529-8817.2007.00371.x}.
+\leavevmode\hypertarget{ref-Wang2020}{}%
+Wang, L., Tomas, F., and Mueller, R. S. (2020). Nutrient enrichment
+increases size of {\emph{Zostera marina}} shoots and enriches for sulfur
+and nitrogen cycling bacteria in root-associated microbiomes. \emph{FEMS
+Microbiol. Ecol.} 96, fiaa129.
+doi:\href{https://doi.org/10.1093/femsec/fiaa129}{10.1093/femsec/fiaa129}.
 
@@
 Phylogenetic analysis of bacterial communities associated with leaves of
-the seagrass \emph{Halophila stipulacea} by a culture-independent
-small-subunit rRNA gene approach. \emph{Microb. Ecol.} 39, 22--31.
+the seagrass {\emph{Halophila stipulacea}} by a culture-independent
+small-subunit {rRNA} gene approach. \emph{Microb. Ecol.} 39, 22--31.
+doi:\href{https://doi.org/doi.org/10.1007/s002489900194}{doi.org/10.1007/s002489900194}.
+
+\leavevmode\hypertarget{ref-Weinberger2007}{}%
+Weinberger, F. (2007). Pathogen-induced defense and innate immunity in
+macroalgae. \emph{Biol. Bull.} 213, 290--302.
+doi:\href{https://doi.org/10.2307/25066646}{10.2307/25066646}.
+
+\leavevmode\hypertarget{ref-Wickham2017}{}%
+Wickham, H. (2017). \emph{{tidyverse}: Easily install and load the
+'tidyverse'}.
 
@@
 
-\leavevmode\hypertarget{ref-Wilke2018}{}%
-Wilke, C. O. (2018). \emph{cowplot: Streamlined plot theme and plot
-annotations for 'ggplot2'}.
+\leavevmode\hypertarget{ref-Wilke2019}{}%
+Wilke, C. O. (2019). \emph{{cowplot}: Streamlined plot theme and plot
+annotations for {'ggplot2'}}.
+
+\leavevmode\hypertarget{ref-Wood1981}{}%
+Wood, D. C., and Hayasaka, S. S. (1981). Chemotaxis of rhizoplane
+bacteria to amino acids comprising eelgrass ({{\emph{Zostera marina}}
+L.) root exudate}. \emph{J. Exp. Mar. Biol. Ecol.} 50, 153--161.
+doi:\href{https://doi.org/10.1016/0022-0981(81)90047-2}{10.1016/0022-0981(81)90047-2}.
 
 \leavevmode\hypertarget{ref-Xie2014}{}%
-Xie, Y. (2014). `` knitr: A comprehensive tool for reproducible research
-in R,'' in \emph{Implementing Reproducible Computational Research}, eds.
-V. Stodden, F. Leisch, and R. D. Peng (New York: Chapman and Hall/CRC),
-3--32.
+Xie, Y. (2014). {``{knitr}: A comprehensive tool for reproducible
+research in {R},''} in \emph{Implementing {Reproducible Computational
+Research}}, eds. V. Stodden, F. Leisch, and R. D. Peng ({New York}:
+{Chapman and Hall/CRC}), 3--32.
 
 \leavevmode\hypertarget{ref-Xie2015}{}%
-Xie, Y. (2015). \emph{Dynamic Documents with R and knitr}. 2nd ed. Boca
-Raton, Florida: Chapman and Hall/CRC.
+Xie, Y. (2015). \emph{Dynamic {Documents} with {R} and knitr}. 2nd ed.
+{Boca Raton, Florida}: {Chapman and Hall/CRC}.
+
+\leavevmode\hypertarget{ref-Xie2019}{}%
+Xie, Y. (2019a). \emph{{knitr}: A general-purpose package for dynamic
+report generation in {R}}.
 
 \leavevmode\hypertarget{ref-Xie2019a}{}%
-Xie, Y. (2019a). TinyTeX: A lightweight, cross-platform, and
-easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
-40, 30--32.
+Xie, Y. (2019b). {TinyTeX}: A lightweight, cross-platform, and
+easy-to-maintain {LaTeX} distribution based on {TeX Live}.
+\emph{TUGboat} 40, 30--32.
 
 \leavevmode\hypertarget{ref-Xie2019b}{}%
-Xie, Y. (2019b). \emph{tinytex: Helper functions to install and maintain
-'TeX Live', and compile 'LaTeX' documents}.
+Xie, Y. (2019c). \emph{{tinytex}: Helper functions to install and
+maintain '{TeX Live}', and compile '{LaTeX}' documents}.
 
 \leavevmode\hypertarget{ref-Xie2018}{}%
-Xie, Y., Allaire, J. J., and Grolemund, G. (2018). \emph{R Markdown: The
-Definitive Guide}. 1st ed. Boca Raton, Florida: Chapman and Hall/CRC.
+Xie, Y., Allaire, J. J., and Grolemund, G. (2018). \emph{R {Markdown}:
+{The Definitive Guide}}. 1st ed. {Boca Raton, Florida}: {Chapman and
+Hall/CRC}.
 
@@
 Xie, Z., Lin, W., and Luo, J. (2017). Comparative phenotype and genome
-analysis of \emph{Cellvibrio} sp. PR1, a xylanolytic and agarolytic
-bacterium from the Pearl River. \emph{BioMed Res. Int.} 2017, 6304248.
+analysis of {\emph{Cellvibrio}} sp. {PR1}, a xylanolytic and agarolytic
+bacterium from the {Pearl River}. \emph{BioMed Res. Int.} 2017, 6304248.
 doi:\href{https://doi.org/10.1155/2017/6304248}{10.1155/2017/6304248}.
@@
 Yilmaz, P., Parfrey, L. W., Yarza, P., Gerken, J., Pruesse, E., Quast,
-C., et al. (2014). The SILVA and "All-Species Living Tree Project (LTP)"
-taxonomic frameworks. \emph{Nucleic Acids Res.} 42, D643--D648.
+C., et al. (2014). The {SILVA} and "{All}-{Species Living Tree Project}
+({LTP})" taxonomic frameworks. \emph{Nucleic Acids Res.} 42, D643--D648.
 doi:\href{https://doi.org/10.1093/nar/gkt1209}{10.1093/nar/gkt1209}.
@@
 in the rate of photosynthetic activity and chemical composition of the
-seagrass \emph{Cymodocea nodosa} (Ucr.) Asch. \emph{Sci. Mar.} 62,
+seagrass {\emph{Cymodocea nodosa}} ({Ucr}.) {Asch}. \emph{Sci. Mar.} 62,
 301--309.
@@
 \leavevmode\hypertarget{ref-Zhu2019}{}%
-Zhu, H. (2019). \emph{kableExtra: Construct complex table with 'kable'
+Zhu, H. (2019). \emph{{kableExtra}: Construct complex table with 'kable'
 and pipe syntax}.
+
+\end{CSLReferences}
 
@@
 
-\textbf{\autoref{matrix}.} \nameref{matrix}
+\textbf{\autoref{map}.} \nameref{map}
+
+\textbf{\autoref{pcoa}.} \nameref{pcoa}
 
@@
 
-\textbf{\autoref{pcoa}.} \nameref{pcoa}
-
 \textbf{\autoref{community}.} \nameref{community}
@@
 \textbf{\autoref{desulfo}.} \nameref{desulfo}
+
+\newpage
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/matrix}
+{\centering \includegraphics[width=1\linewidth]{/home/marino/Work/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/map}
 
@@
 
-\caption{Proportion of shared bacterial and archaeal OTUs (Jaccard's similarity coefficient) and shared bacterial and archaeal communities (Bray-Curtis similarity coefficient) between assemblages associated with the surfaces of the macrophytes \textit{C.
-nodosa} (mixed settlement) and \textit{C.
-cylindracea} (mixed and monospecific settlement) and communities in the ambient seawater.\label{matrix}}\label{fig:unnamed-chunk-1}
+\caption{Location of the mixed (\textit{C.
+nodosa} and \textit{C.
+cylindracea}) and monospecific (\textit{C.
+cylindracea}) settlement in the Bay of Funtana, northern Adriatic Sea (© OpenStreetMap contributors, www.openstreetmap.org/copyright).\label{map}}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/seasonal_shared}
-
-}
-
-\caption{Proportion of shared bacterial and archaeal communities (Bray-Curtis similarity coefficient) and shared bacterial and archaeal OTUs (Jaccard's similarity coefficient) between consecutive sampling dates and from the surfaces of the macrophytes \textit{C.
-nodosa} (mixed settlement) and \textit{C.
-cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\label{shared}}\label{fig:unnamed-chunk-2}
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.55\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/pcoa_figure}
+{\centering \includegraphics[width=0.55\linewidth]{/home/marino/Work/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/pcoa_figure}
 
@@
 cylindracea} (mixed and monospecific settlement) and in the ambient seawater.
-Samples from the same environment or same season are labeled in different colors.
-The proportion of explained variation by each axis is shown on the corresponding axis in parentheses.\label{pcoa}}\label{fig:unnamed-chunk-3}
+Samples from different environments or seasons are labeled in different color and shape.
+The proportion of explained variation by each axis is shown on the corresponding axis in parentheses.\label{pcoa}}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/community_bar_plot}
+{\centering \includegraphics[width=1\linewidth]{/home/marino/Work/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/seasonal_shared}
+
+}
+
+\caption{Proportion of shared bacterial and archaeal communities (Bray-Curtis similarity coefficient) and shared bacterial and archaeal OTUs (Jaccard's similarity coefficient) between consecutive sampling dates and from the surfaces of the macrophytes \textit{C.
+nodosa} (mixed settlement) and \textit{C.
+cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\label{shared}}\label{fig:unnamed-chunk-3}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.85\linewidth]{/home/marino/Work/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/community_bar_plot}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/cyanobacteria_bar_plot}
+{\centering \includegraphics[width=0.85\linewidth]{/home/marino/Work/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/cyanobacteria_bar_plot}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/bacteroidota_bar_plot}
+{\centering \includegraphics[width=0.85\linewidth]{/home/marino/Work/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/bacteroidota_bar_plot}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/alphaproteobacteria_bar_plot}
+{\centering \includegraphics[width=0.85\linewidth]{/home/marino/Work/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/alphaproteobacteria_bar_plot}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/gammaproteobacteria_bar_plot}
+{\centering \includegraphics[width=0.85\linewidth]{/home/marino/Work/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/gammaproteobacteria_bar_plot}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/desulfobacterota_bar_plot}
+{\centering \includegraphics[width=0.85\linewidth]{/home/marino/Work/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/desulfobacterota_bar_plot}
 

### 2021-07-09 submission/supplementary.Rmd
Review done.

@@
 ---
-title: "**Supplementary Information**"
+title: "**Supplementary Material**"
 subtitle: "**Seasonal Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces**"
@@
 
-\vspace{20mm}
+\vspace{10mm}
 Marino Korlević^1$*$^, Marsej Markovski^1^, Zihao Zhao^2^, Gerhard J. Herndl^2,3^, Mirjana Najdek^1^
@@
 ```
+
+```{r calculators}
+# Loading rarefied community data created during the generation of the richness and diversity calculators plot
+load(file="results/numerical/rarefied.Rdata")
+
+# Copying the sample labels to the rows (input for library vegan)
+rarefied <- rarefied %>%
+  column_to_rownames("Group")
+
+# Calculating Chao1 and ACE species estimators
+estimators <- estimateR(rarefied)
+estimators <- as_tibble(t(estimators), .name_repair="unique") %>%
+  add_column("Group"=colnames(estimators), .before=TRUE)
+
+# Calculating diversity indices
+shannon <- diversity(rarefied, index="shannon")
+shannon <- tibble(Group=names(shannon), shannon)
+invsimpson <- diversity(rarefied, index="invsimpson")
+invsimpson <- tibble(Group=names(invsimpson), invsimpson)
+
+# Transforming the Shannon entropy to the effective number of species
+# (http://www.loujost.com/Statistics%20and%20Physics/Diversity%20and%20Similarity/EffectiveNumberOfSpecies.htm)
+shannon <- mutate(shannon, shannon=exp(shannon)) %>%
+  rename(eshannon=shannon)
+
+# Joining together estimators and indices
+estimators_indices <- inner_join(estimators, shannon, by="Group") %>%
+  inner_join(., invsimpson, by="Group")
+
+# Loading metadata 
+metadata <- read_tsv("data/raw/metadata.csv")
+
+# Joining metadata with estimators and indices
+estimators_indices_metadata <- inner_join(metadata, estimators_indices, by=c("ID"="Group")) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y"))
+
+# Calculating the mean and standard deviation for each environment
+calculators_environemnt <- estimators_indices_metadata %>%
+  group_by(station) %>%
+  summarise(S.obs_mean=mean(S.obs), S.obs_sd=sd(S.obs),
+            S.chao1_mean=mean(S.chao1), S.chao1_sd=sd(S.chao1),
+            S.ACE_mean=mean(S.ACE), S.ACE_sd=sd(S.ACE),
+            eshannon_mean=mean(eshannon), eshannon_sd=sd(eshannon),
+            invsimpson_mean=mean(invsimpson), invsimpson_sd=sd(invsimpson)) %>%
+  arrange(match(station, c("F", "FCyM", "FCaM", "FCa"))) %>%
+  mutate(station=str_replace(station, "^F$", "Seawater")) %>%
+  mutate(station=str_replace(station, "^FCyM$", "\\\\textit{Cymodocea nodosa} (Mixed)")) %>%
+  mutate(station=str_replace(station, "^FCaM$", "\\\\textit{Caulerpa cylindracea} (Mixed)")) %>%
+  mutate(station=str_replace(station, "^FCa$", "\\\\textit{Caulerpa cylindracea} (Monospecific)")) %>%
+  mutate_at(vars(2:ncol(.)), ~ format(., big.mark=",", digits=1, nsmall=1, scientific=FALSE)) %>%
+  mutate(S.obs=paste(S.obs_mean, " ± ", S.obs_sd),
+         S.chao1=paste(S.chao1_mean, " ± ", S.chao1_sd),
+         S.ACE=paste(S.ACE_mean, " ± ", S.ACE_sd),
+         eshannon=paste(eshannon_mean, " ± ", eshannon_sd),
+         invsimpson=paste(invsimpson_mean, " ± ", invsimpson_sd)) %>%
+  select(!contains("_"))
+
+# Generating the table
+kable(calculators_environemnt, format="latex", booktabs=TRUE, linesep="", longtable=TRUE, escape=FALSE, col.names=c("Community Type", "Observed No.
+of OTUs", "Chao1", "ACE", "Exponential Shannon", "Inverse Simpson"), caption="Richness estimators and diversity indices of each community type.
+Calculations were done after normalization to the minimum number of reads per sample.\\label{calculators_community_type}", align="c") %>%
+  kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9) %>%
+  row_spec(0, bold=TRUE) %>%
+  column_spec(1, width="10em", latex_valign="m") %>%
+  column_spec(2, width="10em", latex_valign="m") %>%
+  column_spec(5:6, width="6em", latex_valign="m")
+```
+
+\newpage
+```{r calculator_statistics}
+# Performing the Kruskal-Wallis H test
+# Observed No. of OTUs
+kruskal <- kruskal.test(formula=S.obs ~ station, data=estimators_indices_metadata)
+pairwise.wilcox <- pairwise.wilcox.test(estimators_indices_metadata$S.obs, estimators_indices_metadata$station,
+                                              p.adjust.method="bonferroni", exact=FALSE)
+pairwise <- pairwise.wilcox$p.value %>%
+  as_tibble() %>%
+  add_column("V"=rownames(pairwise.wilcox$p.value), .before=TRUE)
+calculators_environemnt_test <-  tribble(~parameter, ~KW_H, ~KW_df, ~KW_p, ~pairwise, ~MW_p,
+                                         "S.obs", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCyM", filter(pairwise, V=="FCyM")$F,
+                                         "S.obs", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCaM", filter(pairwise, V=="FCaM")$F,
+                                         "S.obs", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCa", filter(pairwise, V=="FCa")$F,
+                                         "S.obs", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCaM", filter(pairwise, V=="FCyM")$FCaM,
+                                         "S.obs", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCa", filter(pairwise, V=="FCyM")$FCa,
+                                         "S.obs", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCaM – FCa", filter(pairwise, V=="FCaM")$FCa)
+
+# Chao1
+kruskal <- kruskal.test(formula=S.chao1 ~ station, data=estimators_indices_metadata)
+pairwise.wilcox <- pairwise.wilcox.test(estimators_indices_metadata$S.chao1, estimators_indices_metadata$station,
+                                              p.adjust.method="bonferroni")
+pairwise <- pairwise.wilcox$p.value %>%
+  as_tibble() %>%
+  add_column("V"=rownames(pairwise.wilcox$p.value), .before=TRUE)
+test_data <- tribble(~parameter, ~KW_H, ~KW_df, ~KW_p, ~pairwise, ~MW_p,
+                     "S.chao1", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCyM", filter(pairwise, V=="FCyM")$F,
+                     "S.chao1", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCaM", filter(pairwise, V=="FCaM")$F,
+                     "S.chao1", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCa", filter(pairwise, V=="FCa")$F,
+                     "S.chao1", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCaM", filter(pairwise, V=="FCyM")$FCaM,
+                     "S.chao1", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCa", filter(pairwise, V=="FCyM")$FCa,
+                     "S.chao1", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCaM – FCa", filter(pairwise, V=="FCaM")$FCa)
+calculators_environemnt_test <- rbind(calculators_environemnt_test, test_data)
+
+# ACE
+kruskal <- kruskal.test(formula=S.ACE ~ station, data=estimators_indices_metadata)
+pairwise.wilcox <- pairwise.wilcox.test(estimators_indices_metadata$S.ACE, estimators_indices_metadata$station,
+                                              p.adjust.method="bonferroni")
+pairwise <- pairwise.wilcox$p.value %>%
+  as_tibble() %>%
+  add_column("V"=rownames(pairwise.wilcox$p.value), .before=TRUE)
+test_data <- tribble(~parameter, ~KW_H, ~KW_df, ~KW_p, ~pairwise, ~MW_p,
+                     "S.ACE", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCyM", filter(pairwise, V=="FCyM")$F,
+                     "S.ACE", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCaM", filter(pairwise, V=="FCaM")$F,
+                     "S.ACE", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCa", filter(pairwise, V=="FCa")$F,
+                     "S.ACE", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCaM", filter(pairwise, V=="FCyM")$FCaM,
+                     "S.ACE", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCa", filter(pairwise, V=="FCyM")$FCa,
+                     "S.ACE", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCaM – FCa", filter(pairwise, V=="FCaM")$FCa)
+calculators_environemnt_test <- rbind(calculators_environemnt_test, test_data)
+
+# Exponential Shannon
+kruskal <- kruskal.test(formula=eshannon ~ station, data=estimators_indices_metadata)
+pairwise.wilcox <- pairwise.wilcox.test(estimators_indices_metadata$eshannon, estimators_indices_metadata$station,
+                                              p.adjust.method="bonferroni")
+pairwise <- pairwise.wilcox$p.value %>%
+  as_tibble() %>%
+  add_column("V"=rownames(pairwise.wilcox$p.value), .before=TRUE)
+test_data <- tribble(~parameter, ~KW_H, ~KW_df, ~KW_p, ~pairwise, ~MW_p,
+                     "eshannon", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCyM", filter(pairwise, V=="FCyM")$F,
+                     "eshannon", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCaM", filter(pairwise, V=="FCaM")$F,
+                     "eshannon", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCa", filter(pairwise, V=="FCa")$F,
+                     "eshannon", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCaM", filter(pairwise, V=="FCyM")$FCaM,
+                     "eshannon", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCa", filter(pairwise, V=="FCyM")$FCa,
+                     "eshannon", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCaM – FCa", filter(pairwise, V=="FCaM")$FCa)
+calculators_environemnt_test <- rbind(calculators_environemnt_test, test_data)
+
+# Inverse Simpson
+kruskal <- kruskal.test(formula=invsimpson ~ station, data=estimators_indices_metadata)
+pairwise.wilcox <- pairwise.wilcox.test(estimators_indices_metadata$invsimpson, estimators_indices_metadata$station,
+                                              p.adjust.method="bonferroni")
+pairwise <- pairwise.wilcox$p.value %>%
+  as_tibble() %>%
+  add_column("V"=rownames(pairwise.wilcox$p.value), .before=TRUE)
+test_data <- tribble(~parameter, ~KW_H, ~KW_df, ~KW_p, ~pairwise, ~MW_p,
+                     "invsimpson", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCyM", filter(pairwise, V=="FCyM")$F,
+                     "invsimpson", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCaM", filter(pairwise, V=="FCaM")$F,
+                     "invsimpson", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCa", filter(pairwise, V=="FCa")$F,
+                     "invsimpson", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCaM", filter(pairwise, V=="FCyM")$FCaM,
+                     "invsimpson", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCa", filter(pairwise, V=="FCyM")$FCa,
+                     "invsimpson", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCaM – FCa", filter(pairwise, V=="FCaM")$FCa)
+calculators_environemnt_test <- rbind(calculators_environemnt_test, test_data)
+
+# Formatting the statistics table
+calculators_environemnt_test <- calculators_environemnt_test %>%
+  mutate(pairwise=str_replace(pairwise, "F – FCyM", "Seawater – \\\\textit{Cymodocea nodosa} (Mixed)")) %>%
+  mutate(pairwise=str_replace(pairwise, "F – FCaM", "Seawater – \\\\textit{Caulerpa cylindracea} (Mixed)")) %>%
+  mutate(pairwise=str_replace(pairwise, "F – FCa", "Seawater – \\\\textit{Caulerpa cylindracea} (Monospecific)")) %>%
+  mutate(pairwise=str_replace(pairwise, "FCyM – FCaM", "\\\\textit{Cymodocea nodosa} (Mixed) – \\\\textit{Caulerpa cylindracea} (Mixed)")) %>%
+  mutate(pairwise=str_replace(pairwise, "FCyM – FCa", "\\\\textit{Cymodocea nodosa} (Mixed) – \\\\textit{Caulerpa cylindracea} (Monospecific)")) %>%
+  mutate(pairwise=str_replace(pairwise, "FCaM – FCa", "\\\\textit{Caulerpa cylindracea} (Mixed) – \\\\textit{Caulerpa cylindracea} (Monospecific)")) %>%
+  mutate(KW_p=case_when(KW_p < 0.05 & KW_p >= 0.01 ~ "< 0.05",
+                        KW_p < 0.01 & KW_p >= 0.001 ~ "< 0.01",
+                        KW_p < 0.001 & KW_p >= 0.0001 ~ "< 0.001",
+                        KW_p < 0.0001 ~ "< 0.0001",
+                        TRUE ~ as.character(format(round(KW_p, 2), nsmall=2, scientific=FALSE)))) %>%
+  mutate(MW_p=case_when(MW_p < 0.05 & MW_p >= 0.01 ~ "< 0.05",
+                        MW_p < 0.01 & MW_p >= 0.001 ~ "< 0.01",
+                        MW_p < 0.001 & MW_p >= 0.0001 ~ "< 0.001",
+                        MW_p < 0.0001 ~ "< 0.0001",
+                        TRUE ~ as.character(format(round(MW_p, 2), nsmall=2, scientific=FALSE)))) %>%
+  mutate(parameter=str_replace(parameter, "S.obs", "Observed No. of OTUs")) %>%
+  mutate(parameter=str_replace(parameter, "S.chao1", "Chao1")) %>%
+  mutate(parameter=str_replace(parameter, "S.ACE", "ACE")) %>%
+  mutate(parameter=str_replace(parameter, "eshannon", "Exponential Shannon")) %>%
+  mutate(parameter=str_replace(parameter, "invsimpson", "Inverse Simpson")) %>%
+  mutate(KW_H=format(KW_H, digits=1, nsmall=1))
+
+# Generating the table
+calculator_statistics <- kable(calculators_environemnt_test, format="latex", booktabs=TRUE, linesep="", longtable=TRUE, escape=FALSE, col.names=c("Parameter", "\\textit{H}", "\\textit{df}", "\\textit{p}", "Comparisons Between Community Types", "\\textit{p}"), caption="Statistic of richness estimator and diversity index parameters for different community types.
+Parameters were tested by applying the Kruskal-Wallis \\textit{H} test followed by a pairwise comparison using the Mann-Whitney \\textit{U} test.
+Bonferroni correction was used to address the problem of multiple comparisons.\\label{calculator_statistics}", align="c") %>%
+  add_header_above(c(" ", "Kruskal-Wallis \\\\textit{H} test"=3, "Mann-Whitney \\\\textit{U} test"=2), bold=TRUE, escape=FALSE) %>%
+  kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9) %>%
+  row_spec(0, bold=TRUE) %>%
+  column_spec(1, width="6em", latex_valign="m") %>%
+  collapse_rows(columns=1:2, valign="middle", latex_hline="major")
+
+# Saving the table
+cat(calculator_statistics, file="results/numerical/calculator_statistics.tex")
+
+# Merging rows in column H, df and p by parameter
+system("sed -i 's/ &  & 3 & < 0.0001 / \\&  \\&  \\&  /' results/numerical/calculator_statistics.tex")
+system("sed -i 's/& 3 & < 0.0001 &/\\& \\\\multirow{-6}{*}{\\\\centering\\\\arraybackslash 3} \\& \\\\multirow{-6}{*}{\\\\centering\\\\arraybackslash < 0.0001} \\&/' results/numerical/calculator_statistics.tex")
+```
+\input{../results/numerical/calculator_statistics.tex}
+
+```{r core_otus_taxonomy}
+# Loading rarefied community data created during the generation of the richness and diversity calculators plot
+load(file="results/numerical/rarefied.Rdata")
+
+# Joining metadata with OTU/sample data
+metadata_rarefied <- read_tsv("data/raw/metadata.csv") %>%
+  inner_join(., rarefied, by=c("ID"="Group")) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y"))
+
+# Finding the core OTUs of each environment and their taxonomy
+# Cymodocea nodosa (mixed)
+core <- metadata_rarefied %>%
+  filter(station=="FCyM") %>%
+  select_if(~ all(. != 0)) %>%
+  select(starts_with("Otu"))
+
+total <- metadata_rarefied %>%
+  filter(station=="FCyM") %>%
+  select(starts_with("Otu"))
+
+core_otus_fcym <- length(core)/length(total) * 100
+core_seq_fcym <- sum(core)/sum(total) * 100
+
+core_otu_sum <- as.list(colSums(core)) %>%
+  as_tibble() %>%
+  gather(key="OTU", value="size_normalized", 1:ncol(.))
+
+otu_taxonomy_fcym <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.0.03.cons.taxonomy") %>%
+  inner_join(core_otu_sum, ., by=c("OTU"="OTU")) %>%
+  select(-Size)
+
+# Caulerpa cylindracea (mixed)
+core <- metadata_rarefied %>%
+  filter(station=="FCaM") %>%
+  select_if(~ all(. != 0)) %>%
+  select(starts_with("Otu"))
+
+total <- metadata_rarefied %>%
+  filter(station=="FCaM") %>%
+  select(starts_with("Otu"))
+
+core_otus_fcam <- length(core)/length(total) * 100
+core_seq_fcam <- sum(core)/sum(total) * 100
+
+core_otu_sum <- as.list(colSums(core)) %>%
+  as_tibble() %>%
+  gather(key="OTU", value="size_normalized", 1:ncol(.))
+
+otu_taxonomy_fcam <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.0.03.cons.taxonomy") %>%
+  inner_join(core_otu_sum, ., by=c("OTU"="OTU")) %>%
+  select(-Size)
+
+# Caulerpa cylindracea (monospecific)
+core <- metadata_rarefied %>%
+  filter(station=="FCa") %>%
+  select_if(~ all(. != 0)) %>%
+  select(starts_with("Otu"))
+
+total <- metadata_rarefied %>%
+  filter(station=="FCa") %>%
+  select(starts_with("Otu"))
+
+core_otus_fca <- length(core)/length(total) * 100
+core_seq_fca <- sum(core)/sum(total) * 100
+
+core_otu_sum <- as.list(colSums(core)) %>%
+  as_tibble() %>%
+  gather(key="OTU", value="size_normalized", 1:ncol(.))
+
+otu_taxonomy_fca <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.0.03.cons.taxonomy") %>%
+  inner_join(core_otu_sum, ., by=c("OTU"="OTU")) %>%
+  select(-Size)
+```
+
+\newpage
+\blandscape
+```{r core_otus_taxonomy_table}
+otu_taxonomy_fcym <- otu_taxonomy_fcym %>%
+  mutate(environment="\\textit{Cymodocea nodosa} (Mixed)", .before=OTU) %>%
+  arrange(desc(size_normalized)) %>%
+  slice(1:10)
+
+otu_taxonomy_fcam <- otu_taxonomy_fcam %>%
+  mutate(environment="\\textit{Caulerpa cylindracea} (Mixed)", .before=OTU) %>%
+  arrange(desc(size_normalized)) %>%
+  slice(1:10)
+
+otu_taxonomy_fca <- otu_taxonomy_fca %>%
+  mutate(environment="\\textit{Caulerpa cylindracea} (Monospecific)", .before=OTU) %>%
+  arrange(desc(size_normalized)) %>%
+  slice(1:10)
+
+otu_taxonomy <- rbind(otu_taxonomy_fcym,
+                      otu_taxonomy_fcam,
+                      otu_taxonomy_fca) %>%
+  mutate(Taxonomy=str_replace_all(Taxonomy, ";", "; ")) %>%
+  mutate(Taxonomy=str_replace_all(Taxonomy, "\\([^\\)]+\\)", "")) %>%
+  mutate(Taxonomy=str_replace_all(Taxonomy, "_", " ")) %>%
+  mutate(Taxonomy=str_replace(Taxonomy, "; $", "}")) %>%
+  mutate(Taxonomy=str_replace_all(Taxonomy, "^", "\\\\textit{")) %>%
+  mutate(Taxonomy=str_replace_all(Taxonomy, "; ", "}; \\\\textit{")) %>%
+  mutate(Taxonomy=str_replace(Taxonomy, "\\\\textit\\{Rhodobacteraceae unclassified\\}", "\\\\textit{Rhodobacteraceae} (NR)")) %>%
+  mutate(Taxonomy=str_replace(Taxonomy, "\\\\textit\\{Pleurocapsa PCC-7319\\}", "\\\\textit{Pleurocapsa}")) %>%
+  mutate(Taxonomy=str_replace(Taxonomy, "\\\\textit\\{SEEP-SRB4\\}", "SEEP-SRB4")) %>%
+  mutate(Taxonomy=str_replace(Taxonomy, "\\\\textit\\{Desulfocapsaceae unclassified\\}", "\\\\textit{Desulfocapsaceae} (NR)")) %>%
+  mutate(Taxonomy=str_replace(Taxonomy, "\\\\textit\\{Desulfobacteraceae unclassified\\}", "\\\\textit{Desulfobacteraceae} (NR)"))
+
+# Generating the table
+kable(otu_taxonomy, format="latex", booktabs=TRUE, linesep="", longtable=TRUE, escape=FALSE, col.names=c("Community Type", "OTU Number", "No.
+of Sequences", "OTU Taxonomy"), caption="Taxonomic classification of OTUs present at every sampling date in the each community type.
+Only ten OTUs with the highest number of sequences after normalization to the minimum number of reads per sample are shown.
+NR -- No Relative (sequences without known relatives within the corresponding group)\\label{core_otus_taxonomy_table}", align=c("c", "c", "c", "l")) %>%
+  kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9) %>%
+  row_spec(0, bold=TRUE) %>%
+  column_spec(1, width="8em", latex_valign="m") %>%
+  collapse_rows(columns=1, valign="middle", latex_hline="major")
+```
+\elandscape

### 2021-07-09 code/plot_calculators.R

@@ -138,7 +138,7 @@ p1 <- filter(estimators_indices_metadata, station=="FCyM") %>%
   scale_linetype_manual(values=lines_p1) +
   scale_shape_manual(values=shapes_p1) +
   scale_fill_manual(values=fills_p1) +
-  scale_y_continuous(limits=c(300, 2500)) +
+  scale_y_continuous(limits=c(200, 2600), breaks=c(500, 1000, 1500, 2000, 2500)) +
   scale_x_date(breaks=seq(as.Date("2017-11-01"), as.Date("2018-11-01"), "months"),
                labels=c("Nov 2017", "Dec 2017",
                         "Jan 2018", "Feb 2018", "Mar 2018", "Apr 2018", "May 2018", "Jun 2018",

### 2021-07-09 code/plot_pcoa.R

@@ -48,7 +48,7 @@ coordinates <- inner_join(metadata, coordinates, by=c("label"="Group"))
 p1 <- ggplot() +
   geom_point(data=coordinates, aes(x=A1, y=A2, fill=station), shape=21, size=3, stroke=0.5) +
   scale_fill_manual(name=NULL,
-                      values=c("#66A61E", "#666666", "#A6761D", "#E6AB02"),
+                      values=c("F"="#1F78B4", "FCyM"="#666666", "FCaM"="#FDBF6F", "FCa"="#FF7F00"),
                       breaks=c("F", "FCyM", "FCaM", "FCa"),
                       labels=c("Seawater",
                                parse(text="italic('Cymodocea nodosa')~'(Mixed)'"),
@@ -78,8 +78,7 @@ p2 <- ggplot() +
   geom_point(data=coordinates, aes(x=A1, y=A2, fill=season), shape=21, size=3, stroke=0.5) +
   scale_fill_manual(name=NULL,
                     breaks=c("Spring", "Summer", "Autumn", "Winter"),
-                    values=c("Spring"="#377EB8", "Summer"="#4DAF4A",
-                             "Autumn"="#E41A1C", "Winter"="#984EA3")) +
+                    values=c("Spring"="#1F78B4", "Summer"="#666666", "Autumn"="#FDBF6F", "Winter"="#FF7F00")) +
   labs(x=paste0("PCoA I (",format(round(spe.b.pcoa$eig[1]/sum(spe.b.pcoa$eig)*100, digits=2), nsmall=2)," %)"), 
        y=paste0("PCoA II (",format(round(spe.b.pcoa$eig[2]/sum(spe.b.pcoa$eig)*100, digits=2), nsmall=2)," %)")) +
   ggtitle(parse(text="bolditalic('Cymodocea nodosa')~bold('(Mixed)')")) +
@@ -104,8 +103,7 @@ p3 <- ggplot() +
   geom_point(data=coordinates, aes(x=A1, y=A2, fill=season, shape=station), size=3, stroke=0.5) +
   scale_fill_manual(name=NULL,
                     breaks=c("Spring", "Summer", "Autumn", "Winter"),
-                    values=c("Spring"="#377EB8", "Summer"="#4DAF4A",
-                             "Autumn"="#E41A1C", "Winter"="#984EA3")) +
[…]

### 2021-07-09 code/plot_seasonal_shared.R

@@ -46,253 +46,90 @@ distance_metadata <- inner_join(metadata, distance, by=c("ID"="V1")) %>%
 theme <- theme(text=element_text(family="Times"), line=element_line(color="black"),
                panel.border=element_rect(fill=NA), panel.background=element_blank(),
                panel.grid=element_blank(), axis.line=element_blank(),
-               axis.text=element_text(size=12, color="black"), axis.text.x=element_text(angle=90, hjust=0.95, vjust=1.75),
-               axis.title=element_text(size=14, color="black"),
+               axis.text=element_text(size=16, color="black"), axis.text.x=element_text(angle=90, hjust=0.95, vjust=1.75),
+               axis.title=element_text(size=18, color="black"),
                plot.margin=unit(c(5.5, 5.5, 5.5, 5.5), "pt"), legend.position="none",
-               plot.title=element_text(size=16, hjust=0.5))
+               plot.title=element_text(face="bold", size=20, hjust=0.5))
 
-# Defining line types, dot shapes and fill dot colors
-lines <- c("jaccard"="dotted", "bray"="solid")
-shapes <- c("jaccard"=21, "bray"=23)
-fills <- c("jaccard"="white", "bray"="black")
+# Formatting group names and defining line types, dot shapes and fill dot colours
+group_name <- c("F"=expression("Seawater"),
+                "FCyM"=expression(paste(italic("Cymodocea nodosa"), plain(" (Mixed)"))),
+                "FCaM"=expression(paste(italic("Caulerpa cylindracea"), plain(" (Mixed)"))),
+                "FCa"=expression(paste(italic("Caulerpa cylindracea"), plain(" (Monospecific)"))))
+lines <- c("F"="dotted", "FCyM"="solid", "FCaM"="solid", "FCa"="solid")
+shapes <- c("F"=21, "FCyM"=21, "FCaM"=23, "FCa"=25)
+fills <- c("F"="white", "FCyM"="black", "FCaM"="black", "FCa"="black")
 
 # Plots generation
-# Seawater samples
-data <- filter(distance_metadata, station.x=="F" & station.y=="F") %>%
+data <- distance_metadata %>%
[…]

### 2021-07-09 submission/supplementary.tex

@@
 \hypersetup{
-  pdftitle={Supplementary Information},
+  pdftitle={Supplementary Material},
   hidelinks,
@@
 \renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\figurename}{Fig.}
+\renewcommand{\figurename}{Figure}
 \renewcommand{\thetable}{S\arabic{table}}
@@
 \usepackage{xcolor}
+\ifluatex
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
 
-\title{\textbf{Supplementary Information}}
+\title{\textbf{Supplementary Material}}
 \usepackage{etoolbox}
@@
 \makeatother
-\subtitle{\textbf{Seasonal Dynamics of Epiphytic Microbial Communities on Marine
-Macrophyte Surfaces}}
+\subtitle{\textbf{Seasonal Dynamics of Epiphytic Microbial Communities
+on Marine Macrophyte Surfaces}}
 \author{}
@@
 
-\vspace{20mm}
+\vspace{10mm}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/rarefaction}
+{\centering \includegraphics[width=0.85\linewidth]{/home/marino/Desktop/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/rarefaction}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/calculators}
+{\centering \includegraphics[width=0.85\linewidth]{/home/marino/Desktop/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/calculators}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/mydisk/POSAO2/isabella/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/chloroplast_bar_plot}
+{\centering \includegraphics[width=0.85\linewidth]{/home/marino/Desktop/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/chloroplast_bar_plot}
 
@@
 
+\begingroup\fontsize{9}{11}\selectfont
+
+\begin{longtable}[t]{>{\centering\arraybackslash}m{10em}>{\centering\arraybackslash}m{10em}cc>{\centering\arraybackslash}m{6em}>{\centering\arraybackslash}m{6em}}
+\caption{\label{tab:calculators}Richness estimators and diversity indices of each community type.
+Calculations were done after normalization to the minimum number of reads per sample.\label{calculators_community_type}}\\
+\toprule
+\textbf{Community Type} & \textbf{Observed No. of OTUs} & \textbf{Chao1} & \textbf{ACE} & \textbf{Exponential Shannon} & \textbf{Inverse Simpson}\\
+\midrule
+\endfirsthead
+\caption[]{Richness estimators and diversity indices of each community type.
+Calculations were done after normalization to the minimum number of reads per sample.\label{calculators_community_type} \textit{(continued)}}\\
+\toprule
+\textbf{Community Type} & \textbf{Observed No. of OTUs} & \textbf{Chao1} & \textbf{ACE} & \textbf{Exponential Shannon} & \textbf{Inverse Simpson}\\
+\midrule
+\endhead
+
+\endfoot
+\bottomrule
+\endlastfoot
+Seawater & 531.0  ±  143.9 & 1,244.5  ±  392.7 & 1,441.1  ±  458.7 & 62.2  ±   20.5 & 26.7  ±   8.0\\
+\textit{Cymodocea nodosa} (Mixed) & 1,063.7  ±  210.6 & 1,703.2  ±  409.9 & 1,838.8  ±  437.3 & 237.8  ±  100.4 & 76.9  ±  46.6\\
+\textit{Caulerpa cylindracea} (Mixed) & 1,688.4  ±  136.6 & 2,936.9  ±  460.2 & 3,220.7  ±  497.2 & 410.4  ±   92.5 & 102.2  ±  47.2\\
+\textit{Caulerpa cylindracea} (Monospecific) & 1,750.4  ±  165.7 & 2,911.0  ±  575.0 & 3,171.7  ±  622.1 & 496.4  ±   87.1 & 144.4  ±  42.6\\*
+\end{longtable}
+\endgroup{}
+
+\newpage
+
+\input{../results/numerical/calculator_statistics.tex}
+
+\newpage
+\blandscape
+\begingroup\fontsize{9}{11}\selectfont
+
+\begin{longtable}[t]{>{\centering\arraybackslash}m{8em}ccl}
+\caption{\label{tab:core_otus_taxonomy_table}Taxonomic classification of OTUs present at every sampling date in the each community type.
+Only ten OTUs with the highest number of sequences after normalization to the minimum number of reads per sample are shown.
+NR -- No Relative (sequences without known relatives within the corresponding group)\label{core_otus_taxonomy_table}}\\
+\toprule
+\textbf{Community Type} & \textbf{OTU Number} & \textbf{No. of Sequences} & \textbf{OTU Taxonomy}\\
+\midrule
+\endfirsthead
+\caption[]{Taxonomic classification of OTUs present at every sampling date in the each community type.
+Only ten OTUs with the highest number of sequences after normalization to the minimum number of reads per sample are shown.
+NR -- No Relative (sequences without known relatives within the corresponding group)\label{core_otus_taxonomy_table} \textit{(continued)}}\\
+\toprule
+\textbf{Community Type} & \textbf{OTU Number} & \textbf{No. of Sequences} & \textbf{OTU Taxonomy}\\
+\midrule
+\endhead
+
+\endfoot
+\bottomrule
+\endlastfoot
+ & Otu00014 & 5092 & \textit{Bacteria}; \textit{Proteobacteria}; \textit{Alphaproteobacteria}; \textit{Rhodobacterales}; \textit{Rhodobacteraceae}; \textit{Rhodobacteraceae} (NR)\\
+\nopagebreak
+ & Otu00020 & 3188 & \textit{Bacteria}; \textit{Proteobacteria}; \textit{Alphaproteobacteria}; \textit{Rhodobacterales}; \textit{Rhodobacteraceae}; \textit{Rhodobacteraceae} (NR)\\
+\nopagebreak
+ & Otu00019 & 3026 & \textit{Bacteria}; \textit{Cyanobacteria}; \textit{Cyanobacteriia}; \textit{Cyanobacteriales}; \textit{Xenococcaceae}; \textit{Pleurocapsa}\\
+\nopagebreak
+ & Otu00030 & 2014 & \textit{Bacteria}; \textit{Proteobacteria}; \textit{Alphaproteobacteria}; \textit{Rhodobacterales}; \textit{Rhodobacteraceae}; \textit{Rhodobacteraceae} (NR)\\
+\nopagebreak
+ & Otu00013 & 1884 & \textit{Bacteria}; \textit{Proteobacteria}; \textit{Alphaproteobacteria}; \textit{Rhodobacterales}; \textit{Rhodobacteraceae}; \textit{Rhodobacteraceae} (NR)\\
+\nopagebreak
+ & Otu00032 & 1586 & \textit{Bacteria}; \textit{Verrucomicrobiota}; \textit{Verrucomicrobiae}; \textit{Opitutales}; \textit{Puniceicoccaceae}; \textit{Lentimonas}\\
+\nopagebreak
+ & Otu00059 & 1469 & \textit{Bacteria}; \textit{Proteobacteria}; \textit{Gammaproteobacteria}; \textit{Burkholderiales}; \textit{Methylophilaceae}; \textit{Methylotenera}\\
+\nopagebreak
+ & Otu00005 & 1337 & \textit{Bacteria}; \textit{Proteobacteria}; \textit{Alphaproteobacteria}; \textit{Rhodobacterales}; \textit{Rhodobacteraceae}; \textit{Rhodobacteraceae} (NR)\\
+\nopagebreak
+ & Otu00050 & 1264 & \textit{Bacteria}; \textit{Proteobacteria}; \textit{Alphaproteobacteria}; \textit{Rhodobacterales}; \textit{Rhodobacteraceae}; \textit{Rhodobacteraceae} (NR)\\
+\nopagebreak
+\multirow{-10}{8em}{\centering\arraybackslash \textit{Cymodocea nodosa} (Mixed)} & Otu00029 & 1162 & \textit{Bacteria}; \textit{Proteobacteria}; \textit{Alphaproteobacteria}; \textit{Rhodobacterales}; \textit{Rhodobacteraceae}; \textit{Rhodobacteraceae} (NR)\\
+\cmidrule{1-4}\pagebreak[0]
+ & Otu00011 & 4051 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobacteria}; \textit{Desulfobacterales}; \textit{Desulfosarcinaceae}; \textit{Desulfatitalea}\\
+\nopagebreak
+ & Otu00010 & 3436 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobulbia}; \textit{Desulfobulbales}; \textit{Desulfocapsaceae}; \textit{Desulfocapsaceae} (NR)\\
+\nopagebreak
+ & Otu00013 & 2531 & \textit{Bacteria}; \textit{Proteobacteria}; \textit{Alphaproteobacteria}; \textit{Rhodobacterales}; \textit{Rhodobacteraceae}; \textit{Rhodobacteraceae} (NR)\\
+\nopagebreak
+ & Otu00021 & 1859 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobulbia}; \textit{Desulfobulbales}; \textit{Desulfocapsaceae}; \textit{Desulfocapsaceae} (NR)\\
+\nopagebreak
+ & Otu00025 & 1481 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobulbia}; \textit{Desulfobulbales}; \textit{Desulfocapsaceae}; SEEP-SRB4\\
+\nopagebreak
+ & Otu00037 & 1209 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobulbia}; \textit{Desulfobulbales}; \textit{Desulfobulbaceae}; \textit{Desulfobulbus}\\
+\nopagebreak
+ & Otu00033 & 1189 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobulbia}; \textit{Desulfobulbales}; \textit{Desulfocapsaceae}; \textit{Desulfocapsaceae} (NR)\\
+\nopagebreak
+ & Otu00044 & 1042 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobulbia}; \textit{Desulfobulbales}; \textit{Desulfocapsaceae}; \textit{Desulfocapsaceae} (NR)\\
+\nopagebreak
+ & Otu00019 & 1012 & \textit{Bacteria}; \textit{Cyanobacteria}; \textit{Cyanobacteriia}; \textit{Cyanobacteriales}; \textit{Xenococcaceae}; \textit{Pleurocapsa}\\
+\nopagebreak
+\multirow{-10}{8em}{\centering\arraybackslash \textit{Caulerpa cylindracea} (Mixed)} & Otu00046 & 982 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobulbia}; \textit{Desulfobulbales}; \textit{Desulfocapsaceae}; \textit{Desulfotalea}\\
+\cmidrule{1-4}\pagebreak[0]
+ & Otu00010 & 3572 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobulbia}; \textit{Desulfobulbales}; \textit{Desulfocapsaceae}; \textit{Desulfocapsaceae} (NR)\\
+\nopagebreak
+ & Otu00011 & 2977 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobacteria}; \textit{Desulfobacterales}; \textit{Desulfosarcinaceae}; \textit{Desulfatitalea}\\
+\nopagebreak
+ & Otu00013 & 2109 & \textit{Bacteria}; \textit{Proteobacteria}; \textit{Alphaproteobacteria}; \textit{Rhodobacterales}; \textit{Rhodobacteraceae}; \textit{Rhodobacteraceae} (NR)\\
+\nopagebreak
+ & Otu00021 & 1931 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobulbia}; \textit{Desulfobulbales}; \textit{Desulfocapsaceae}; \textit{Desulfocapsaceae} (NR)\\
+\nopagebreak
+ & Otu00025 & 1912 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobulbia}; \textit{Desulfobulbales}; \textit{Desulfocapsaceae}; SEEP-SRB4\\
+\nopagebreak
+ & Otu00033 & 1613 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobulbia}; \textit{Desulfobulbales}; \textit{Desulfocapsaceae}; \textit{Desulfocapsaceae} (NR)\\
+\nopagebreak
+ & Otu00040 & 1289 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobacteria}; \textit{Desulfobacterales}; \textit{Desulfobacteraceae}; \textit{Desulfobacteraceae} (NR)\\
+\nopagebreak
+ & Otu00046 & 1089 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobulbia}; \textit{Desulfobulbales}; \textit{Desulfocapsaceae}; \textit{Desulfotalea}\\
+\nopagebreak
+ & Otu00037 & 1058 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobulbia}; \textit{Desulfobulbales}; \textit{Desulfobulbaceae}; \textit{Desulfobulbus}\\
+\nopagebreak
+\multirow{-10}{8em}{\centering\arraybackslash \textit{Caulerpa cylindracea} (Monospecific)} & Otu00058 & 950 & \textit{Bacteria}; \textit{Desulfobacterota}; \textit{Desulfobacteria}; \textit{Desulfobacterales}; \textit{Desulfosarcinaceae}; \textit{Desulfatitalea}\\*
+\end{longtable}
+\endgroup{}
+\elandscape
+
 \end{document}

### 2021-08-10 submission/manuscript.Rmd
Editor review done.

@@
 nodosa* and *C.
-cylindracea* was isolated from a pool of leaves or thalli on the sampling day using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevic2021; @Massana1997].
+cylindracea* was isolated from a pool of leaves (1 g wet weight) or thalli (2 g wet weight) on the sampling day using a previously modified and adapted protocol that allows for a selective epiphytic DNA isolation [@Korlevic2021; @Massana1997].
 Briefly, leaves and thalli were incubated in a lysis buffer and treated with lysozyme and proteinase K.
@@
 Dried pellets were re-suspended in 50 -- 100 \si{\ul} of deionized water.
+One DNA sample originating from seawater picoplankton was obtained per each sampling point.
 
@@
 
-\newpage
 ## Results
@@
 
+\newpage
 ## Acknowledgments

### 2021-08-10 submission/manuscript.tex

@@
 DNA from surfaces of \emph{C. nodosa} and \emph{C. cylindracea} was
-isolated from a pool of leaves or thalli on the sampling day using a
-previously modified and adapted protocol that allows for a selective
-epiphytic DNA isolation (Massana et al., 1997; Korlević et al., 2021).
-Briefly, leaves and thalli were incubated in a lysis buffer and treated
-with lysozyme and proteinase K. Following the incubations, the mixture
-containing lysed epiphytic cells was separated from the leaves and
-thalli and extracted using phenol-chloroform. Finally, the extracted DNA
-was precipitated using isopropanol. DNA from seawater picoplankton was
-extracted from 0.2 \si{\um} polycarbonate filters according to Massana
-et al. (1997) with a slight modification. Following the
-phenol-chloroform extraction, 1/10 of 3 \si{\Molar} sodium acetate (pH
-5.2) was added. DNA was precipitated by adding 1 volume of chilled
-isopropanol, incubating the mixtures overnight at \num{-20}
-\si{\degreeCelsius} and centrifuging at 20,000 × \emph{g} and 4
-\si{\degreeCelsius} for 20 \si{\minute}. The pellet was washed twice
-with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and centrifuged
-after each washing step at 20,000 × \emph{g} and 4 \si{\degreeCelsius}
-for 5 \si{\minute}. Dried pellets were re-suspended in 50 -- 100
-\si{\ul} of deionized water.
+isolated from a pool of leaves (1 g wet weight) or thalli (2 g wet
+weight) on the sampling day using a previously modified and adapted
+protocol that allows for a selective epiphytic DNA isolation (Massana et
+al., 1997; Korlević et al., 2021). Briefly, leaves and thalli were
+incubated in a lysis buffer and treated with lysozyme and proteinase K.
+Following the incubations, the mixture containing lysed epiphytic cells
+was separated from the leaves and thalli and extracted using
+phenol-chloroform. Finally, the extracted DNA was precipitated using
+isopropanol. DNA from seawater picoplankton was extracted from 0.2
+\si{\um} polycarbonate filters according to Massana et al. (1997) with a
+slight modification. Following the phenol-chloroform extraction, 1/10 of
+3 \si{\Molar} sodium acetate (pH 5.2) was added. DNA was precipitated by
+adding 1 volume of chilled isopropanol, incubating the mixtures
+overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 20,000 ×
+\emph{g} and 4 \si{\degreeCelsius} for 20 \si{\minute}. The pellet was
+washed twice with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and
+centrifuged after each washing step at 20,000 × \emph{g} and 4
+\si{\degreeCelsius} for 5 \si{\minute}. Dried pellets were re-suspended
+in 50 -- 100 \si{\ul} of deionized water. One DNA sample originating
+from seawater picoplankton was obtained per each sampling point.
 
@@
 (\url{https://github.com/MicrobesRovinj/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021}).
-
-\newpage
 
@@
 
+\newpage
+
 \hypertarget{acknowledgments}{%

### 2021-08-17 submission/manuscript.Rmd
Production done.

@@
 
-Seagrasses are know to form close relationships with microbial communities associated with the surfaces of leaves, roots and rhizomes [@Cucio2016; @Ugarelli2019; @Wang2020; @Ettinger2020; @Crump2018].
+Seagrasses are known to form close relationships with microbial communities associated with the surfaces of leaves, roots and rhizomes [@Cucio2016; @Ugarelli2019; @Wang2020; @Ettinger2020; @Crump2018].
 For different seagrass species a distinct microbial community from ambient seawater or bulk sediment has been reported, however no species specific communities have been found [@Cucio2016; @Ugarelli2019; @Crump2018].

### 2021-08-17 submission/manuscript.tex
Production done.

@@
 
-Seagrasses are know to form close relationships with microbial
+Seagrasses are known to form close relationships with microbial
 communities associated with the surfaces of leaves, roots and rhizomes

### 2021-08-17 submission/supplementary.tex
Production done.

@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/home/marino/Desktop/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/rarefaction}
+{\centering \includegraphics[width=0.85\linewidth]{/home/marino/Work/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/rarefaction}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/home/marino/Desktop/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/calculators}
+{\centering \includegraphics[width=0.85\linewidth]{/home/marino/Work/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/calculators}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/home/marino/Desktop/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/chloroplast_bar_plot}
+{\centering \includegraphics[width=0.85\linewidth]{/home/marino/Work/repositories/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021/results/figures/chloroplast_bar_plot}
 
@@
 \begin{longtable}[t]{>{\centering\arraybackslash}m{8em}ccl}
-\caption{\label{tab:core_otus_taxonomy_table}Taxonomic classification of OTUs present at every sampling date in the each community type.
+\caption{\label{tab:core_otus_taxonomy_table}Taxonomic classification of OTUs present at every sampling date in each community type.
 Only ten OTUs with the highest number of sequences after normalization to the minimum number of reads per sample are shown.
@@
 \endfirsthead
-\caption[]{Taxonomic classification of OTUs present at every sampling date in the each community type.
+\caption[]{Taxonomic classification of OTUs present at every sampling date in each community type.
 Only ten OTUs with the highest number of sequences after normalization to the minimum number of reads per sample are shown.
